### 2022-06-10 simulate_cutoff.Rmd
all the updates

---
title: "Simulate Cut Off"
author: "Erin M. Buchanan"
date: "Last Knitted: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
set.seed(895893)
library(dplyr)
library(ggplot2)
library(tidyr)
library(broom)
library(papaja)
```

# Introduction

This document outlines using the suggested procedure to simulate the number of participants necessary to accurately measure items. There are two key issues these ideas should address that we know about power:

1) we should see differences in projected sample sizes based on the variability in the variance for those items (i.e., heterogeneity should increase projected sample size)
2) we should see projected sample sizes that "level off" when pilot data increases. As with regular power estimates, studies can be "overpowered" to detect an effect, and this same idea should be present. For example, if one has a 500 person pilot study, our simulations should suggest a point at which items are likely measured well, which may have happened well before 500.

# Method

## Data Simulation

*Population*. The data was simulated using the `rnorm` function assuming a normal distribution for 30 rating scale type items rounded to two decimal places. The item means were simulated as $\mu$ = 4 points with a $\sigma$  of .25 points. The item $\sigma$s were simulated as 2 points with a low variability ($\sigma$ = .2), medium variability (.4), and high variability (.8). Each population was simulated with 1000 data points.

```{r}
## add room to vary as part of the equation

# mu <- rnorm(30, 4, .25)
# sigma1 <- rnorm(30, 2, .2)
# sigma2 <- rnorm(30, 2, .4)
# sigma3 <- rnorm(30, 2, .8)

mu <- rnorm(30, 847.48, 150)
sigma1 <- rnorm(30, 350, 100)
sigma2 <- rnorm(30, 350, 150)
sigma3 <- rnorm(30, 350, 200)

population1 <- data.frame(
  item = rep(1:30, 1000), 
  score = rnorm(1000*30, mean = mu, sd = sigma1)
  )

population2 <- data.frame(
  item = rep(1:30, 1000), 
  score = rnorm(1000*30, mean = mu, sd = sigma2)
  )

population3 <- data.frame(
  item = rep(1:30, 1000), 
  score = rnorm(1000*30, mean = mu, sd = sigma3)
  )

#evidence that they are simulated correctly
tapply(population1$score, population1$item, mean) # around 4
tapply(population1$score, population1$item, sd) # around 2
sd(tapply(population1$score, population1$item, sd)) # around .2

tapply(population2$score, population2$item, mean) # around 4
tapply(population2$score, population2$item, sd) # around 2
sd(tapply(population2$score, population2$item, sd)) # around .4

tapply(population3$score, population3$item, mean) # around 4
tapply(population3$score, population3$item, sd) # around 2
sd(tapply(population3$score, population3$item, sd)) # around .8
```

*Samples*. Each population was then sampled as if a researcher was conducting a pilot study. The sample sizes started at 20 participants per item increasing in units of 5 up to 100 participants.

```{r}
# create pilot samples from 20 to 100
samples1 <- samples2 <- samples3 <- list() 
sizes <- c(20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75, 80, 85, 90, 95, 100)
for (i in 1:length(sizes)){
  samples1[[i]] <- population1 %>% 
    group_by(item) %>% 
    slice_sample(n = sizes[i])

  samples2[[i]] <- population2 %>% 
    group_by(item) %>% 
    slice_sample(n = sizes[i])

  samples3[[i]] <- population3 %>% 
    group_by(item) %>% 
    slice_sample(n = sizes[i])
}
```

*AIPE Criterions*. The standard errors of each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as standard error functions as the main component in the formula for normal distribution confidence intervals. Standard errors were calculated at each decile of the items up to 90% (0% smallest SE, 10% ..., 90% largest SE). The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet AIPE goals, while the higher deciles would represent less strict criterions for AIPE goals.

```{r}
# calculate the SEs and the cutoff scores 
SES1 <- SES2 <- SES3 <- list()
cutoffs1 <- cutoffs2 <- cutoffs3 <- list()
for (i in 1:length(samples1)){
  SES1[[i]] <- tapply(samples1[[i]]$score,
                     samples1[[i]]$item,
                     function (x){ sd(x)/sqrt(length(x))})
  SES2[[i]] <- tapply(samples2[[i]]$score,
                   samples2[[i]]$item,
                   function (x){ sd(x)/sqrt(length(x))})
  SES3[[i]] <- tapply(samples3[[i]]$score,
                 samples3[[i]]$item,
                 function (x){ sd(x)/sqrt(length(x))})

  cutoffs1[[i]] <- quantile(SES1[[i]], probs = seq(0, .9, by = .1))
  cutoffs2[[i]] <- quantile(SES2[[i]], probs = seq(0, .9, by = .1))
  cutoffs3[[i]] <- quantile(SES3[[i]], probs = seq(0, .9, by = .1))
}
```

## AIPE Simulation

In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items. Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 500 to determine what the new sample size suggestion would be. We assume that samples over 500 may be considered too large for many researchers who do not work in teams. The standard error of each item was calculated for each suggested sample size by pilot sample size by population type.

```{r}
# sequence of sample sizes to try
samplesize_values <- seq(20, 1000, 5)

# place to store everything
sampled_values1 <- sampled_values2 <- sampled_values3 <- list()

# loop over the samples
for (i in 1:length(samples1)){

  # create a blank table for us to save the values in 
  sim_table1 <- matrix(NA, 
                      nrow = length(samplesize_values), 
                      ncol = 30)

  # make it a data frame
  sim_table1 <- sim_table2 <- sim_table3 <- as.data.frame(sim_table1)

  # add a place for sample size values 
  sim_table1$sample_size <- sim_table2$sample_size <- sim_table3$sample_size <- NA

  # loop over pilot sample sizes
  for (q in 1:length(samplesize_values)){

    # temp dataframe that samples and summarizes
    temp <- samples1[[i]] %>% 
      group_by(item) %>% 
      slice_sample(n = samplesize_values[q], replace = T) %>% 
      summarize(se = sd(score)/sqrt(length(score)))

    sim_table1[q, 1:30] <- temp$se
    sim_table1[q, 31] <- samplesize_values[q]

    temp <- samples2[[i]] %>% 
      group_by(item) %>% 
      slice_sample(n = samplesize_values[q], replace = T) %>% 
      summarize(se = sd(score)/sqrt(length(score)))

    sim_table2[q, 1:30] <- temp$se
    sim_table2[q, 31] <- samplesize_values[q]

    temp <- samples3[[i]] %>% 
      group_by(item) %>% 
      slice_sample(n = samplesize_values[q], replace = T) %>% 
      summarize(se = sd(score)/sqrt(length(score)))

    sim_table3[q, 1:30] <- temp$se
    sim_table3[q, 31] <- samplesize_values[q]

    } # end pilot sample loop

  sampled_values1[[i]] <- sim_table1
  sampled_values2[[i]] <- sim_table2
  sampled_values3[[i]] <- sim_table3

} # end all sample loop 
```

Next, the percent of items that fall below the cutoff scores, and thus, would be considered "well-measured" were calculated for each decile by sample.

```{r}
summary_list1 <- summary_list2 <- summary_list3 <- list()
for (i in 1:length(sampled_values1)){

  summary_list1[[i]] <- sampled_values1[[i]] %>% 
    pivot_longer(cols = -c(sample_size)) %>% 
    rename(item = name, se = value) %>% 
    group_by(sample_size) %>% 
    summarize(Percent_Below0 = sum(se <= cutoffs1[[i]][1])/30,
             Percent_Below10 = sum(se <= cutoffs1[[i]][2])/30,
             Percent_Below20 = sum(se <= cutoffs1[[i]][3])/30,
             Percent_Below30 = sum(se <= cutoffs1[[i]][4])/30,
             Percent_Below40 = sum(se <= cutoffs1[[i]][5])/30,
             Percent_Below50 = sum(se <= cutoffs1[[i]][6])/30, 
             Percent_Below60 = sum(se <= cutoffs1[[i]][7])/30, 
             Percent_Below70 = sum(se <= cutoffs1[[i]][8])/30, 
             Percent_Below80 = sum(se <= cutoffs1[[i]][9])/30, 
             Percent_Below90 = sum(se <= cutoffs1[[i]][10])/30) %>% 
    mutate(original_n = sizes[i])

  summary_list2[[i]] <- sampled_values2[[i]] %>% 
    pivot_longer(cols = -c(sample_size)) %>% 
    rename(item = name, se = value) %>% 
    group_by(sample_size) %>% 
    summarize(Percent_Below0 = sum(se <= cutoffs2[[i]][1])/30,
             Percent_Below10 = sum(se <= cutoffs2[[i]][2])/30,
             Percent_Below20 = sum(se <= cutoffs2[[i]][3])/30,
             Percent_Below30 = sum(se <= cutoffs2[[i]][4])/30,
             Percent_Below40 = sum(se <= cutoffs2[[i]][5])/30,
             Percent_Below50 = sum(se <= cutoffs2[[i]][6])/30, 
             Percent_Below60 = sum(se <= cutoffs2[[i]][7])/30, 
             Percent_Below70 = sum(se <= cutoffs2[[i]][8])/30, 
             Percent_Below80 = sum(se <= cutoffs2[[i]][9])/30, 
             Percent_Below90 = sum(se <= cutoffs2[[i]][10])/30) %>% 
    mutate(original_n = sizes[i])

  summary_list3[[i]] <- sampled_values3[[i]] %>% 
    pivot_longer(cols = -c(sample_size)) %>% 
    rename(item = name, se = value) %>% 
    group_by(sample_size) %>% 
    summarize(Percent_Below0 = sum(se <= cutoffs3[[i]][1])/30,
             Percent_Below10 = sum(se <= cutoffs3[[i]][2])/30,
             Percent_Below20 = sum(se <= cutoffs3[[i]][3])/30,
             Percent_Below30 = sum(se <= cutoffs3[[i]][4])/30,
             Percent_Below40 = sum(se <= cutoffs3[[i]][5])/30,
             Percent_Below50 = sum(se <= cutoffs3[[i]][6])/30, 
             Percent_Below60 = sum(se <= cutoffs3[[i]][7])/30, 
             Percent_Below70 = sum(se <= cutoffs3[[i]][8])/30, 
             Percent_Below80 = sum(se <= cutoffs3[[i]][9])/30, 
             Percent_Below90 = sum(se <= cutoffs3[[i]][10])/30) %>% 
    mutate(original_n = sizes[i])
}

summary_DF <- bind_rows(summary_list1, 
                        summary_list2, 
                        summary_list3)
summary_DF$source <- rep(c("low", "med", "high"), each = length(sampled_values1)*length(samplesize_values))
```

From this data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90%, and 95% of the items fall below the cutoff criterion. These values were chosen as popular measures of "power" in which one could determine the minimum suggested sample size (potentially 80% of items) and the maximum suggested sample size (potentially 90%).

```{r}
summary_long_80 <- summary_DF %>% 
  pivot_longer(cols = -c(sample_size, original_n, source)) %>% 
  filter(value >= .80) %>% 
  arrange(sample_size, original_n, source, name) %>% 
  group_by(original_n, name, source) %>% 
  slice_head(n = 1) %>% 
  mutate(power = 80)

summary_long_85 <- summary_DF %>% 
  pivot_longer(cols = -c(sample_size, original_n, source)) %>% 
  filter(value >= .85) %>% 
  arrange(sample_size, original_n, source, name) %>% 
  group_by(original_n, name, source) %>% 
  slice_head(n = 1) %>% 
  mutate(power = 85)

summary_long_90 <- summary_DF %>% 
  pivot_longer(cols = -c(sample_size, original_n, source)) %>% 
  filter(value >= .90) %>% 
  arrange(sample_size, original_n, source, name) %>% 
  group_by(original_n, name, source) %>% 
  slice_head(n = 1) %>% 
  mutate(power = 90)

summary_long_95 <- summary_DF %>% 
  pivot_longer(cols = -c(sample_size, original_n, source)) %>% 
  filter(value >= .95) %>% 
  arrange(sample_size, original_n, source, name) %>% 
  group_by(original_n, name, source) %>% 
  slice_head(n = 1) %>% 
  mutate(power = 95)

summary_long <- rbind(summary_long_80, 
                      summary_long_85,
                      summary_long_90,
                      summary_long_95)

summary_long$source <- factor(summary_long$source, 
                              levels = c("low", "med", "high"),
                              labels = c("Low", "Medium", "High"))
summary_long$sd_original <- .2
summary_long$sd_original[summary_long$source == "Medium"] <- .4
summary_long$sd_original[summary_long$source == "High"] <- .8
summary_long$name <- gsub("Percent_Below", "Decile ", summary_long$name)
```

# Results

## Differences in Item Variance

We examined if this procedure is sensitive to differences in item heterogeneity, as we should expect to collect larger samples if we wish to have a large number of items reach a threshold of acceptable variance; potentially, assuring we *could* average them if a researcher did not wish to use a more complex analysis such as multilevel modeling.

The figure below illustrates the potential minimum sample size for 80% of items to achieve a desired cutoff score. The black dots denote the original sample size against the suggested sample size. By comparing the facets, we can determine that our suggested procedure does capture the differences in heterogeneity. As heterogeneity increases in item variances, the proposed sample size also increases, especially at stricter cutoffs. Missing cutoff points where sample sizes proposed would be higher than 500.

```{r}
ggplot(summary_long %>% filter(power == 80), aes(original_n, sample_size, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ source) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

## Sample Size Sensitivity to Pilot Data Size

In our second question, we examined if the suggested procedure was sensitive to the amount of information present in the pilot data. Larger pilot data is more informative, and therefore, we should expect a lower projected sample size. As shown in the figure below for only the low variability data, we do not find this effect. These simplistic simulations from the pilot data would nearly always suggest a larger sample size - mostly in a linear trend increasing with sample sizes. This result comes from the nature of the procedure - if we base our estimates on some SE cutoff, we will almost always need a bit more people for items to meet those goals. This result does not achieve our second goal.

```{r}
ggplot(summary_long %>% filter(source == "Low"), aes(original_n, sample_size, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

Therefore, we suggest using a correction factor on the simulation procedure to account for the known asymptotic nature of power (i.e., at larger sample sizes power increases level off). For this function, we combined a correction factor for upward biasing of effect sizes (Hedges' correction) with the formula for exponential decay calculations. The decay factor is calculated as follows:

$$ 1 - \sqrt{\frac{N_{pilot} - min(N_{sim})}{N_{pilot}}}^{log_2(N_{pilot})}$$
$N_{pilot}$ indicates the sample size of the pilot data minus the minimum sample size for simulation to ensure that the smallest sample sizes do not decay (e.g., the formula zeroes out). This value is raised to the power of $log_2$ of the sample size of the pilot data, which decreases the impact of the decay to smaller increments for increasing sample sizes. This value is then multiplied by the proposed sample size. As show in the figure below, this correction factor produces the dsired quality of maintaining that small pilot studies should *increase* sample size, and that sample size suggestions level off as pilot study data sample size increases.

```{r}
decay <- 1-sqrt((summary_long$original_n-20)/summary_long$original_n)^log2(summary_long$original_n)

summary_long$new_sample <- summary_long$sample_size*decay

ggplot(summary_long %>% filter(source == "Low"), aes(original_n, new_sample, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

## Corrections for Individual Researchers

We have portrayed that this procedure, with a correction factor, can perform as desired. However, within real scenarios, researchers will only have one pilot sample, not many as show here. What should the researcher do to correct their sample size on their own pilot data?

To explore if we could recover the new suggested sample size from data a researcher would have, we used linear models to create a formula for calculation. First, the corrected sample size was predicted by the original suggested sample size. Next, the standard deviation of the item standard deviations was added to the equation. Last, we included the pilot sample size.

```{r}
user_model <- lm(new_sample ~ sample_size, data = summary_long)
user_print <- apa_print(user_model)

user_model2 <- lm(new_sample ~ sample_size + sd_original, data = summary_long)
user_print2 <- apa_print(user_model2)

user_model3 <- lm(new_sample ~ sample_size + sd_original + original_n, data = summary_long)
user_print3 <- apa_print(user_model3)

change_table <- tidy(anova(user_model, user_model2, user_model3))
```

The first model using original sample size to predict new sample size was significant, `r user_print$full_result$modelfit$r2`, capturing nearly 90% of the variance. The second model with item standard deviation was better then the first model *F*(`r change_table$df[2]`, `r change_table$res.df[2]`) = `r change_table$statistic[2]`, *p* < .001, `r user_print2$estimate$modelfit$r2`. The addition of the original pilot sample size was also significant, *F*(`r change_table$df[3]`, `r change_table$res.df[3]`) = `r change_table$statistic[3]`, *p* < .001, `r user_print3$estimate$modelfit$r2`.

As shown in the table below, the new suggested sample size is proportional to the original suggested sample size (i.e., *b* < 1), which reduces the sample size suggestion. As variability increases, the suggested sample size also increases to capture differences in heterogeneity shown above. Last, in order to correct for large pilot data, the original pilot sample size decreases the new suggested sample size. This formula approximation captures 96% of the variance in sample size scores and should allow a researcher to estimate based on their own data.

```{r results = 'asis'}
apa_table(tidy(user_model3))
```

## Choosing an Appropriate Cutoff

```{r}
by_cutoff <- list()
R2 <- list()

for (cutoff in unique(summary_long$name)){
  by_cutoff[[cutoff]] <- lm(new_sample ~ sample_size + sd_original + original_n, data = summary_long  %>% filter(name == cutoff))
  R2[cutoff] <- summary(by_cutoff[[cutoff]])$r.squared
}

R2
```

Last, we examine the question of an appropriate SE decile. All graphs for power, variability, and correction are presented below. If we examine the $R^2$ values for each decile of our regression equation separately, we find that the 10% (`r round(R2[["Decile 10"]], digits = 3)`) and 30% deciles (`r round(R2[["Decile 30"]], digits = 3)`) represent the best match to our corrected sample size suggestions. The 10% is likely unfeasible for many researchers, as the sample sizes are often quite large. The 30% decile, in the corrected format, appears to meet all goals: 1) increases with heterogeneity, 2) higher suggested values for lower samples and a leveling effect at larger pilot data, and ...

### Low Variability

```{r}
ggplot(summary_long %>% filter(source == "Low"), aes(original_n, sample_size, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

```{r}
ggplot(summary_long %>% filter(source == "Low"), aes(original_n, new_sample, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

### Medium Varability

```{r}
ggplot(summary_long %>% filter(source == "Medium"), aes(original_n, sample_size, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

```{r}
summary_long$new_sample <- summary_long$sample_size*decay

ggplot(summary_long %>% filter(source == "Medium"), aes(original_n, new_sample, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

### High Variability

```{r}
ggplot(summary_long %>% filter(source == "High"), aes(original_n, sample_size, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

```{r}
ggplot(summary_long %>% filter(source == "High"), aes(original_n, new_sample, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

### 2022-06-14 simulate_cutoff.Rmd
added new variable

@@
 
+More real introduction on power, current trends to AIPE, and why items are key.
+
 This document outlines using the suggested procedure to simulate the number of participants necessary to accurately measure items.
@@
 
+# Explain the Procedure
+
+Notes forthcoming.
+
 # Method
@@
 *Population*.
-The data was simulated using the `rnorm` function assuming a normal distribution for 30 rating scale type items rounded to two decimal places.
-The item means were simulated as $\mu$ = 4 points with a $\sigma$  of .25 points.
-The item $\sigma$s were simulated as 2 points with a low variability ($\sigma$ = .2), medium variability (.4), and high variability (.8).
+The data was simulated using the `rnorm` function assuming a normal distribution for 30 scale type items.
 Each population was simulated with 1000 data points.
+No items were rounded for this simulation.
 
-```{r}
-## add room to vary as part of the equation
+First, the scale of the data was manipulated by creating three sets of scales.
+The first scale was mimicked after small rating scales (i.e., 1-7 type style) using a $\mu$ = 4 with a $\sigma$ = .25 around the mean to create item mean variability.
+The second scale included a larger potential distribution of scores with a $\mu$ = 50 ($\sigma$ = 10) imitating a 0-100 scale.
+Last, the final scale included a $\mu$ = 1000 ($\sigma$ = 150) simulating a study that may include response latency data in the milliseconds.
+While there are many potential scales, these three represent a large number of potential variables in the social sciences.
+As we are suggesting variances as a key factor for estimating sample sizes, the scale of the data is influential on the amount of *potential* variance.
+Smaller ranges of data (1-7) cannot necessarily have the same variance as larger ranges (0-100).
 
-# mu <- rnorm(30, 4, .25)
-# sigma1 <- rnorm(30, 2, .2)
-# sigma2 <- rnorm(30, 2, .4)
-# sigma3 <- rnorm(30, 2, .8)
+Next, item variance heterogeneity was included by manipulating the potential $\sigma$ for each individual item.
+For small scales, the $\sigma$ = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items.
+For the medium scale of data, $\sigma$ = 20 with a variance of 4, 8, and 12.
+Last, for the large scale of data, $\sigma$ = 350 with a variance of 100, 150, and 200 for heterogeneity.
 
-mu <- rnorm(30, 847.48, 150)
-sigma1 <- rnorm(30, 350, 100)
-sigma2 <- rnorm(30, 350, 150)
-sigma3 <- rnorm(30, 350, 200)
+```{r echo = F}
+# small potential variability overall, sort of 1-7ish scale
+mu1 <- rnorm(30, 4, .25)
+sigma1.1 <- rnorm(30, 2, .2)
+sigma2.1 <- rnorm(30, 2, .4)
+sigma3.1 <- rnorm(30, 2, .8)
+
+# medium potential variability 0 to 100 scale
+mu2 <- rnorm(30, 50, 10)
+sigma1.2 <- rnorm(30, 20, 4)
+sigma2.2 <- rnorm(30, 20, 8)
+sigma3.2 <- rnorm(30, 20, 12)
+
+# large potential variability in the 1000s scale
+
+mu3 <- rnorm(30, 1000, 150)
+sigma1.3 <- rnorm(30, 350, 100)
+sigma2.3 <- rnorm(30, 350, 150)
+sigma3.3 <- rnorm(30, 350, 200)
+
+while(sum(sigma3.3 < 0) > 0){
+  sigma3.3 <- rnorm(30, 350, 200)
+}
 
 population1 <- data.frame(
-  item = rep(1:30, 1000), 
-  score = rnorm(1000*30, mean = mu, sd = sigma1)
+  item = rep(1:30, 1000*3),
+  scale = rep(1:3, each = 1000*30),
+  score = c(rnorm(1000*30, mean = mu1, sd = sigma1.1),
+            rnorm(1000*30, mean = mu2, sd = sigma1.2),
+            rnorm(1000*30, mean = mu3, sd = sigma1.3))
   )
@@
 population2 <- data.frame(
-  item = rep(1:30, 1000), 
-  score = rnorm(1000*30, mean = mu, sd = sigma2)
+  item = rep(1:30, 1000*3),
+  scale = rep(1:3, each = 1000*30),
+  score = c(rnorm(1000*30, mean = mu1, sd = sigma2.1),
+            rnorm(1000*30, mean = mu2, sd = sigma2.2),
+            rnorm(1000*30, mean = mu3, sd = sigma2.3))
   )
@@
 population3 <- data.frame(
-  item = rep(1:30, 1000), 
-  score = rnorm(1000*30, mean = mu, sd = sigma3)
+  item = rep(1:30, 1000*3),
+  scale = rep(1:3, each = 1000*30),
+  score = c(rnorm(1000*30, mean = mu1, sd = sigma3.1),
+            rnorm(1000*30, mean = mu2, sd = sigma3.2),
+            rnorm(1000*30, mean = mu3, sd = sigma3.3))
   )
@@
 #evidence that they are simulated correctly
-tapply(population1$score, population1$item, mean) # around 4
-tapply(population1$score, population1$item, sd) # around 2
-sd(tapply(population1$score, population1$item, sd)) # around .2
+tapply(population1$score, list(population1$item, 
+                               population1$scale), mean)
+tapply(population1$score, list(population1$item,
+                               population1$scale), sd)
 
-tapply(population2$score, population2$item, mean) # around 4
-tapply(population2$score, population2$item, sd) # around 2
-sd(tapply(population2$score, population2$item, sd)) # around .4
+tapply(population2$score, list(population2$item, 
+                               population2$scale), mean)
+tapply(population2$score, list(population2$item,
+                               population2$scale), sd)
 
-tapply(population3$score, population3$item, mean) # around 4
-tapply(population3$score, population3$item, sd) # around 2
-sd(tapply(population3$score, population3$item, sd)) # around .8
+tapply(population3$score, list(population3$item, 
+                               population3$scale), mean)
+tapply(population3$score, list(population3$item,
+                               population3$scale), sd) 
 ```
@@
   samples1[[i]] <- population1 %>% 
-    group_by(item) %>% 
+    group_by(item, scale) %>% 
     slice_sample(n = sizes[i])
@@
   samples2[[i]] <- population2 %>% 
-    group_by(item) %>% 
+    group_by(item, scale) %>% 
     slice_sample(n = sizes[i])
@@
   samples3[[i]] <- population3 %>% 
-    group_by(item) %>% 
+    group_by(item, scale) %>% 
     slice_sample(n = sizes[i])
@@
   SES1[[i]] <- tapply(samples1[[i]]$score,
-                     samples1[[i]]$item,
+                     list(samples1[[i]]$item,
+                          samples1[[i]]$scale),
                      function (x){ sd(x)/sqrt(length(x))})
   SES2[[i]] <- tapply(samples2[[i]]$score,
-                   samples2[[i]]$item,
+                   list(samples2[[i]]$item,
+                        samples2[[i]]$scale),
                    function (x){ sd(x)/sqrt(length(x))})
   SES3[[i]] <- tapply(samples3[[i]]$score,
-                 samples3[[i]]$item,
+                 list(samples2[[i]]$item,
+                      samples2[[i]]$scale),
                  function (x){ sd(x)/sqrt(length(x))})
 
-  cutoffs1[[i]] <- quantile(SES1[[i]], probs = seq(0, .9, by = .1))
-  cutoffs2[[i]] <- quantile(SES2[[i]], probs = seq(0, .9, by = .1))
-  cutoffs3[[i]] <- quantile(SES3[[i]], probs = seq(0, .9, by = .1))
+  cutoffs1[[i]] <- apply(as.data.frame(SES1[[i]]), 2, 
+                         quantile, 
+                         probs = seq(0, .9, by = .1))
+  cutoffs2[[i]] <- apply(as.data.frame(SES2[[i]]), 2, 
+                         quantile, 
+                         probs = seq(0, .9, by = .1))
+  cutoffs3[[i]] <- apply(as.data.frame(SES3[[i]]), 2, 
+                         quantile, 
+                         probs = seq(0, .9, by = .1))
 }
@@
 Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 500 to determine what the new sample size suggestion would be.
-We assume that samples over 500 may be considered too large for many researchers who do not work in teams.
+We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds.
 The standard error of each item was calculated for each suggested sample size by pilot sample size by population type.
@@
 # sequence of sample sizes to try
-samplesize_values <- seq(20, 1000, 5)
+samplesize_values <- seq(20, 500, 5)
 
@@
                       nrow = length(samplesize_values), 
-                      ncol = 30)
+                      ncol = 30*3)
 
@@
     temp <- samples1[[i]] %>% 
-      group_by(item) %>% 
+      group_by(item, scale) %>% 
       slice_sample(n = samplesize_values[q], replace = T) %>% 
-      summarize(se = sd(score)/sqrt(length(score)))
+      summarize(se = sd(score)/sqrt(length(score)),
+                .groups = "keep")
 
-    sim_table1[q, 1:30] <- temp$se
-    sim_table1[q, 31] <- samplesize_values[q]
+    sim_table1[q, 1:90] <- temp$se
+    sim_table1[q, 91] <- samplesize_values[q]
 
     temp <- samples2[[i]] %>% 
-      group_by(item) %>% 
+      group_by(item, scale) %>% 
       slice_sample(n = samplesize_values[q], replace = T) %>% 
-      summarize(se = sd(score)/sqrt(length(score)))
+      summarize(se = sd(score)/sqrt(length(score)),
+                .groups = "keep")
 
-    sim_table2[q, 1:30] <- temp$se
-    sim_table2[q, 31] <- samplesize_values[q]
+    sim_table2[q, 1:90] <- temp$se
+    sim_table2[q, 91] <- samplesize_values[q]
 
     temp <- samples3[[i]] %>% 
-      group_by(item) %>% 
+      group_by(item, scale) %>% 
       slice_sample(n = samplesize_values[q], replace = T) %>% 
-      summarize(se = sd(score)/sqrt(length(score)))
+      summarize(se = sd(score)/sqrt(length(score)),
+                .groups = "keep")
 
-    sim_table3[q, 1:30] <- temp$se
-    sim_table3[q, 31] <- samplesize_values[q]
+    sim_table3[q, 1:90] <- temp$se
+    sim_table3[q, 91] <- samplesize_values[q]
 
@@
 
+  # summary list 1 ----
   summary_list1[[i]] <- sampled_values1[[i]] %>% 
@@
     rename(item = name, se = value) %>% 
-    group_by(sample_size) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs1[[i]][1])/30,
-             Percent_Below10 = sum(se <= cutoffs1[[i]][2])/30,
-             Percent_Below20 = sum(se <= cutoffs1[[i]][3])/30,
-             Percent_Below30 = sum(se <= cutoffs1[[i]][4])/30,
-             Percent_Below40 = sum(se <= cutoffs1[[i]][5])/30,
-             Percent_Below50 = sum(se <= cutoffs1[[i]][6])/30, 
-             Percent_Below60 = sum(se <= cutoffs1[[i]][7])/30, 
-             Percent_Below70 = sum(se <= cutoffs1[[i]][8])/30, 
-             Percent_Below80 = sum(se <= cutoffs1[[i]][9])/30, 
-             Percent_Below90 = sum(se <= cutoffs1[[i]][10])/30) %>% 
-    mutate(original_n = sizes[i])
+    mutate(scale = rep(1:3, 30*length(samplesize_values))) %>% 
+    mutate(item = rep(rep(1:30, each = 3), length(samplesize_values)))
 
+  # cut offs for 1
+  temp1.1 <- summary_list1[[i]] %>% 
+    filter(scale == "1") %>% 
+    group_by(sample_size, scale) %>% 
+    summarize(Percent_Below0 = sum(se <= cutoffs1[[i]][1, 1])/30,
+           Percent_Below10 = sum(se <= cutoffs1[[i]][2, 1])/30,
+           Percent_Below20 = sum(se <= cutoffs1[[i]][3, 1])/30,
+           Percent_Below30 = sum(se <= cutoffs1[[i]][4, 1])/30,
+           Percent_Below40 = sum(se <= cutoffs1[[i]][5, 1])/30,
+           Percent_Below50 = sum(se <= cutoffs1[[i]][6, 1])/30, 
+           Percent_Below60 = sum(se <= cutoffs1[[i]][7, 1])/30, 
+           Percent_Below70 = sum(se <= cutoffs1[[i]][8, 1])/30, 
+           Percent_Below80 = sum(se <= cutoffs1[[i]][9, 1])/30, 
+           Percent_Below90 = sum(se <= cutoffs1[[i]][10, 1])/30, 
+           .groups = "keep") %>% 
+    mutate(original_n = sizes[i], 
+           source = "low")
+
+  temp1.2 <- summary_list1[[i]] %>% 
+    filter(scale == "2") %>% 
+    group_by(sample_size, scale) %>% 
+    summarize(Percent_Below0 = sum(se <= cutoffs1[[i]][1, 2])/30,
+           Percent_Below10 = sum(se <= cutoffs1[[i]][2, 2])/30,
+           Percent_Below20 = sum(se <= cutoffs1[[i]][3, 2])/30,
+           Percent_Below30 = sum(se <= cutoffs1[[i]][4, 2])/30,
+           Percent_Below40 = sum(se <= cutoffs1[[i]][5, 2])/30,
+           Percent_Below50 = sum(se <= cutoffs1[[i]][6, 2])/30, 
+           Percent_Below60 = sum(se <= cutoffs1[[i]][7, 2])/30, 
+           Percent_Below70 = sum(se <= cutoffs1[[i]][8, 2])/30, 
+           Percent_Below80 = sum(se <= cutoffs1[[i]][9, 2])/30, 
+           Percent_Below90 = sum(se <= cutoffs1[[i]][10, 2])/30, 
+           .groups = "keep") %>% 
+    mutate(original_n = sizes[i],
+           source = "low")
+
+  temp1.3 <- summary_list1[[i]] %>% 
+    filter(scale == "3") %>% 
+    group_by(sample_size, scale) %>% 
+    summarize(Percent_Below0 = sum(se <= cutoffs1[[i]][1, 3])/30,
+           Percent_Below10 = sum(se <= cutoffs1[[i]][2, 3])/30,
+           Percent_Below20 = sum(se <= cutoffs1[[i]][3, 3])/30,
+           Percent_Below30 = sum(se <= cutoffs1[[i]][4, 3])/30,
+           Percent_Below40 = sum(se <= cutoffs1[[i]][5, 3])/30,
+           Percent_Below50 = sum(se <= cutoffs1[[i]][6, 3])/30, 
+           Percent_Below60 = sum(se <= cutoffs1[[i]][7, 3])/30, 
+           Percent_Below70 = sum(se <= cutoffs1[[i]][8, 3])/30, 
+           Percent_Below80 = sum(se <= cutoffs1[[i]][9, 3])/30, 
+           Percent_Below90 = sum(se <= cutoffs1[[i]][10, 3])/30, 
+           .groups = "keep") %>% 
+    mutate(original_n = sizes[i], 
+           source = "low")
+
+  #rejoin 
+  summary_list1[[i]] <- bind_rows(temp1.1, temp1.2, temp1.3)
+
+  # summary list 2 ----
   summary_list2[[i]] <- sampled_values2[[i]] %>% 
@@
     rename(item = name, se = value) %>% 
-    group_by(sample_size) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs2[[i]][1])/30,
-             Percent_Below10 = sum(se <= cutoffs2[[i]][2])/30,
-             Percent_Below20 = sum(se <= cutoffs2[[i]][3])/30,
-             Percent_Below30 = sum(se <= cutoffs2[[i]][4])/30,
-             Percent_Below40 = sum(se <= cutoffs2[[i]][5])/30,
-             Percent_Below50 = sum(se <= cutoffs2[[i]][6])/30, 
-             Percent_Below60 = sum(se <= cutoffs2[[i]][7])/30, 
-             Percent_Below70 = sum(se <= cutoffs2[[i]][8])/30, 
-             Percent_Below80 = sum(se <= cutoffs2[[i]][9])/30, 
-             Percent_Below90 = sum(se <= cutoffs2[[i]][10])/30) %>% 
-    mutate(original_n = sizes[i])
+    mutate(scale = rep(1:3, 30*length(samplesize_values))) %>% 
+    mutate(item = rep(rep(1:30, each = 3), length(samplesize_values)))
 
+  # cut offs for 2
+  temp2.1 <- summary_list2[[i]] %>% 
+    filter(scale == "1") %>% 
+    group_by(sample_size, scale) %>% 
+    summarize(Percent_Below0 = sum(se <= cutoffs2[[i]][1, 1])/30,
+           Percent_Below10 = sum(se <= cutoffs2[[i]][2, 1])/30,
+           Percent_Below20 = sum(se <= cutoffs2[[i]][3, 1])/30,
+           Percent_Below30 = sum(se <= cutoffs2[[i]][4, 1])/30,
+           Percent_Below40 = sum(se <= cutoffs2[[i]][5, 1])/30,
+           Percent_Below50 = sum(se <= cutoffs2[[i]][6, 1])/30, 
+           Percent_Below60 = sum(se <= cutoffs2[[i]][7, 1])/30, 
+           Percent_Below70 = sum(se <= cutoffs2[[i]][8, 1])/30, 
+           Percent_Below80 = sum(se <= cutoffs2[[i]][9, 1])/30, 
+           Percent_Below90 = sum(se <= cutoffs2[[i]][10, 1])/30, 
+           .groups = "keep") %>% 
+    mutate(original_n = sizes[i],
+           source = "med")
+
+  temp2.2 <- summary_list2[[i]] %>% 
+    filter(scale == "2") %>% 
+    group_by(sample_size, scale) %>% 
+    summarize(Percent_Below0 = sum(se <= cutoffs2[[i]][1, 2])/30,
+           Percent_Below10 = sum(se <= cutoffs2[[i]][2, 2])/30,
+           Percent_Below20 = sum(se <= cutoffs2[[i]][3, 2])/30,
+           Percent_Below30 = sum(se <= cutoffs2[[i]][4, 2])/30,
+           Percent_Below40 = sum(se <= cutoffs2[[i]][5, 2])/30,
+           Percent_Below50 = sum(se <= cutoffs2[[i]][6, 2])/30, 
+           Percent_Below60 = sum(se <= cutoffs2[[i]][7, 2])/30, 
+           Percent_Below70 = sum(se <= cutoffs2[[i]][8, 2])/30, 
+           Percent_Below80 = sum(se <= cutoffs2[[i]][9, 2])/30, 
+           Percent_Below90 = sum(se <= cutoffs2[[i]][10, 2])/30, 
+           .groups = "keep") %>% 
+    mutate(original_n = sizes[i], 
+           source = "med")
+
+  temp2.3 <- summary_list2[[i]] %>% 
+    filter(scale == "3") %>% 
+    group_by(sample_size, scale) %>% 
+    summarize(Percent_Below0 = sum(se <= cutoffs2[[i]][1, 3])/30,
+           Percent_Below10 = sum(se <= cutoffs2[[i]][2, 3])/30,
+           Percent_Below20 = sum(se <= cutoffs2[[i]][3, 3])/30,
+           Percent_Below30 = sum(se <= cutoffs2[[i]][4, 3])/30,
+           Percent_Below40 = sum(se <= cutoffs2[[i]][5, 3])/30,
+           Percent_Below50 = sum(se <= cutoffs2[[i]][6, 3])/30, 
+           Percent_Below60 = sum(se <= cutoffs2[[i]][7, 3])/30, 
+           Percent_Below70 = sum(se <= cutoffs2[[i]][8, 3])/30, 
+           Percent_Below80 = sum(se <= cutoffs2[[i]][9, 3])/30, 
+           Percent_Below90 = sum(se <= cutoffs2[[i]][10, 3])/30, 
+           .groups = "keep") %>% 
+    mutate(original_n = sizes[i],
+           source = "med")
+
+  #rejoin 
+  summary_list2[[i]] <- bind_rows(temp2.1, temp2.2, temp2.3)
+
+  # summary list 3 ----
   summary_list3[[i]] <- sampled_values3[[i]] %>% 
@@
     rename(item = name, se = value) %>% 
-    group_by(sample_size) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs3[[i]][1])/30,
-             Percent_Below10 = sum(se <= cutoffs3[[i]][2])/30,
-             Percent_Below20 = sum(se <= cutoffs3[[i]][3])/30,
-             Percent_Below30 = sum(se <= cutoffs3[[i]][4])/30,
-             Percent_Below40 = sum(se <= cutoffs3[[i]][5])/30,
-             Percent_Below50 = sum(se <= cutoffs3[[i]][6])/30, 
-             Percent_Below60 = sum(se <= cutoffs3[[i]][7])/30, 
-             Percent_Below70 = sum(se <= cutoffs3[[i]][8])/30, 
-             Percent_Below80 = sum(se <= cutoffs3[[i]][9])/30, 
-             Percent_Below90 = sum(se <= cutoffs3[[i]][10])/30) %>% 
-    mutate(original_n = sizes[i])
+    mutate(scale = rep(1:3, 30*length(samplesize_values))) %>% 
+    mutate(item = rep(rep(1:30, each = 3), length(samplesize_values)))
+
+  # cut offs for 3 
+  temp3.1 <- summary_list3[[i]] %>% 
+    filter(scale == "1") %>% 
+    group_by(sample_size, scale) %>% 
+    summarize(Percent_Below0 = sum(se <= cutoffs3[[i]][1, 1])/30,
+           Percent_Below10 = sum(se <= cutoffs3[[i]][2, 1])/30,
+           Percent_Below20 = sum(se <= cutoffs3[[i]][3, 1])/30,
+           Percent_Below30 = sum(se <= cutoffs3[[i]][4, 1])/30,
+           Percent_Below40 = sum(se <= cutoffs3[[i]][5, 1])/30,
+           Percent_Below50 = sum(se <= cutoffs3[[i]][6, 1])/30, 
+           Percent_Below60 = sum(se <= cutoffs3[[i]][7, 1])/30, 
+           Percent_Below70 = sum(se <= cutoffs3[[i]][8, 1])/30, 
+           Percent_Below80 = sum(se <= cutoffs3[[i]][9, 1])/30, 
+           Percent_Below90 = sum(se <= cutoffs3[[i]][10, 1])/30, 
+           .groups = "keep") %>% 
+    mutate(original_n = sizes[i],
+           source = "high")
+
+  temp3.2 <- summary_list3[[i]] %>% 
+    filter(scale == "2") %>% 
+    group_by(sample_size, scale) %>% 
+    summarize(Percent_Below0 = sum(se <= cutoffs3[[i]][1, 2])/30,
+           Percent_Below10 = sum(se <= cutoffs3[[i]][2, 2])/30,
+           Percent_Below20 = sum(se <= cutoffs3[[i]][3, 2])/30,
+           Percent_Below30 = sum(se <= cutoffs3[[i]][4, 2])/30,
+           Percent_Below40 = sum(se <= cutoffs3[[i]][5, 2])/30,
+           Percent_Below50 = sum(se <= cutoffs3[[i]][6, 2])/30, 
+           Percent_Below60 = sum(se <= cutoffs3[[i]][7, 2])/30, 
+           Percent_Below70 = sum(se <= cutoffs3[[i]][8, 2])/30, 
+           Percent_Below80 = sum(se <= cutoffs3[[i]][9, 2])/30, 
+           Percent_Below90 = sum(se <= cutoffs3[[i]][10, 2])/30, 
+           .groups = "keep") %>% 
+    mutate(original_n = sizes[i],
+           source = "high")
+
+  temp3.3 <- summary_list3[[i]] %>% 
+    filter(scale == "3") %>% 
+    group_by(sample_size, scale) %>% 
+    summarize(Percent_Below0 = sum(se <= cutoffs3[[i]][1, 3])/30,
+           Percent_Below10 = sum(se <= cutoffs3[[i]][2, 3])/30,
+           Percent_Below20 = sum(se <= cutoffs3[[i]][3, 3])/30,
+           Percent_Below30 = sum(se <= cutoffs3[[i]][4, 3])/30,
+           Percent_Below40 = sum(se <= cutoffs3[[i]][5, 3])/30,
+           Percent_Below50 = sum(se <= cutoffs3[[i]][6, 3])/30, 
+           Percent_Below60 = sum(se <= cutoffs3[[i]][7, 3])/30, 
+           Percent_Below70 = sum(se <= cutoffs3[[i]][8, 3])/30, 
+           Percent_Below80 = sum(se <= cutoffs3[[i]][9, 3])/30, 
+           Percent_Below90 = sum(se <= cutoffs3[[i]][10, 3])/30, 
+           .groups = "keep") %>% 
+    mutate(original_n = sizes[i],
+           source = "high")
+
+  #rejoin 
+  summary_list3[[i]] <- bind_rows(temp3.1, temp3.2, temp3.3)
+
   }
@@
                         summary_list3)
-summary_DF$source <- rep(c("low", "med", "high"), each = length(sampled_values1)*length(samplesize_values))
 ```
@@
 summary_long_80 <- summary_DF %>% 
-  pivot_longer(cols = -c(sample_size, original_n, source)) %>% 
+  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
   filter(value >= .80) %>% 
-  arrange(sample_size, original_n, source, name) %>% 
-  group_by(original_n, name, source) %>% 
+  arrange(sample_size, original_n, source, scale, name) %>% 
+  group_by(original_n, name, source, scale) %>% 
   slice_head(n = 1) %>% 
@@
 summary_long_85 <- summary_DF %>% 
-  pivot_longer(cols = -c(sample_size, original_n, source)) %>% 
+  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
   filter(value >= .85) %>% 
-  arrange(sample_size, original_n, source, name) %>% 
-  group_by(original_n, name, source) %>% 
+  arrange(sample_size, original_n, source, scale, name) %>% 
+  group_by(original_n, name, source, scale) %>% 
   slice_head(n = 1) %>% 
@@
 summary_long_90 <- summary_DF %>% 
-  pivot_longer(cols = -c(sample_size, original_n, source)) %>% 
+  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
   filter(value >= .90) %>% 
-  arrange(sample_size, original_n, source, name) %>% 
-  group_by(original_n, name, source) %>% 
+  arrange(sample_size, original_n, source, scale, name) %>% 
+  group_by(original_n, name, source, scale) %>% 
   slice_head(n = 1) %>% 
@@
 summary_long_95 <- summary_DF %>% 
-  pivot_longer(cols = -c(sample_size, original_n, source)) %>% 
+  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
   filter(value >= .95) %>% 
-  arrange(sample_size, original_n, source, name) %>% 
-  group_by(original_n, name, source) %>% 
+  arrange(sample_size, original_n, source, scale, name) %>% 
+  group_by(original_n, name, source, scale) %>% 
   slice_head(n = 1) %>% 
@@
                               labels = c("Low", "Medium", "High"))
+
 summary_long$sd_original <- .2
-summary_long$sd_original[summary_long$source == "Medium"] <- .4
-summary_long$sd_original[summary_long$source == "High"] <- .8
+summary_long$sd_original[summary_long$source == "Medium" &
+                           summary_long$scale == "1"] <- .4
+summary_long$sd_original[summary_long$source == "High" &
+                           summary_long$source == "1"] <- .8
+
+summary_long$sd_original[summary_long$source == "Low" &
+                           summary_long$scale == "2"] <- 4
+summary_long$sd_original[summary_long$source == "Medium" &
+                           summary_long$scale == "2"] <- 8
+summary_long$sd_original[summary_long$source == "High" &
+                           summary_long$source == "2"] <- 12
+
+summary_long$sd_original[summary_long$source == "Low" &
+                           summary_long$scale == "3"] <- 100
+summary_long$sd_original[summary_long$source == "Medium" &
+                           summary_long$scale == "3"] <- 150
+summary_long$sd_original[summary_long$source == "High" &
+                           summary_long$source == "3"] <- 200
+
 summary_long$name <- gsub("Percent_Below", "Decile ", summary_long$name)
@@
 user_print3 <- apa_print(user_model3)
+
+user_model4 <- lm(new_sample ~ sample_size + sd_original + original_n + scale, data = summary_long)
+user_print4 <- apa_print(user_model4)
 

### 2022-06-17 simulate_cutoff.Rmd
finalizing stimuli power

@@
 
-More real introduction on power, current trends to AIPE, and why items are key.
-
 This document outlines using the suggested procedure to simulate the number of participants necessary to accurately measure items.
@@
 
-Notes forthcoming.
+1) Take your original pilot data.
+2) Find a critical score for an accurately measured item using SE.
+3) Sample from your pilot data using sample sizes starting at 20 participants and increase in units of 5 up to a value you feel is the maximum sample size.
+4) Calculate the number of items that meet your critical score at each sample size.
+5) Find the smallest sample size for 80%, 85%, 90%, and 95% power. 
+6) Designate your minimum sample size from these values, the stopping rule used from step 2, and the maximum sample size from these values.
 
@@
 
-```{r echo = F}
+```{r}
 # small potential variability overall, sort of 1-7ish scale
@@
 cutoffs1 <- cutoffs2 <- cutoffs3 <- list()
+sd_items1 <- sd_items2 <- sd_items3 <- list()
 for (i in 1:length(samples1)){
+  sd_items1[[i]] <- samples1[[i]] %>% group_by(item, scale) %>% 
+    summarize(sd = sd(score), .groups = "keep") %>% 
+    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
+  sd_items2[[i]] <- samples2[[i]] %>% group_by(item, scale) %>% 
+    summarize(sd = sd(score), .groups = "keep") %>% 
+    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
+  sd_items3[[i]] <- samples3[[i]] %>% group_by(item, scale) %>% 
+    summarize(sd = sd(score), .groups = "keep") %>% 
+    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
+
   SES1[[i]] <- tapply(samples1[[i]]$score,
@@
                               levels = c("low", "med", "high"),
-                              labels = c("Low", "Medium", "High"))
+                              labels = c("Low Variance", "Medium Variance", "High Variance"))
 
-summary_long$sd_original <- .2
-summary_long$sd_original[summary_long$source == "Medium" &
-                           summary_long$scale == "1"] <- .4
-summary_long$sd_original[summary_long$source == "High" &
-                           summary_long$source == "1"] <- .8
+summary_long$scale2 <- factor(summary_long$scale, 
+                              levels = c(1:3),
+                              labels = c("Small Scale", "Medium Scale", "Large Scale"))
 
-summary_long$sd_original[summary_long$source == "Low" &
-                           summary_long$scale == "2"] <- 4
-summary_long$sd_original[summary_long$source == "Medium" &
-                           summary_long$scale == "2"] <- 8
-summary_long$sd_original[summary_long$source == "High" &
-                           summary_long$source == "2"] <- 12
-
-summary_long$sd_original[summary_long$source == "Low" &
-                           summary_long$scale == "3"] <- 100
-summary_long$sd_original[summary_long$source == "Medium" &
-                           summary_long$scale == "3"] <- 150
-summary_long$sd_original[summary_long$source == "High" &
-                           summary_long$source == "3"] <- 200
+sd_items <- bind_rows(sd_items1, sd_items2, sd_items3)
+sd_items$original_n <- rep(rep(sizes, each = 3), 3)
+sd_items$source <- rep(c("Low Variance", "Medium Variance", "High Variance"), each = 3*length(sizes))
+summary_long <- summary_long %>% 
+  full_join(sd_items, 
+            by = c("original_n" = "original_n", "scale" = "scale", 
+                   "source" = "source"))
 
@@
 ```{r}
+summary_long$source <- factor(summary_long$source, 
+                              levels = c("Low Variance", "Medium Variance", "High Variance"))
+
 ggplot(summary_long %>% filter(power == 80), aes(original_n, sample_size, color = name)) + 
@@
   theme_classic() + 
-  facet_wrap(~ source) + 
+  facet_wrap(~ source*scale2) + 
   xlab("Original Sample Size") + 
@@
 Larger pilot data is more informative, and therefore, we should expect a lower projected sample size.
-As shown in the figure below for only the low variability data, we do not find this effect.
+As shown in the figure below for only the low variability and small scale data, we do not find this effect.
 These simplistic simulations from the pilot data would nearly always suggest a larger sample size - mostly in a linear trend increasing with sample sizes.
@@
 ```{r}
-ggplot(summary_long %>% filter(source == "Low"), aes(original_n, sample_size, color = name)) + 
+ggplot(summary_long %>% filter(source == "Low Variance") %>% filter(scale == 1), aes(original_n, sample_size, color = name)) + 
   geom_point() +  
@@
 This value is then multiplied by the proposed sample size.
-As show in the figure below, this correction factor produces the dsired quality of maintaining that small pilot studies should *increase* sample size, and that sample size suggestions level off as pilot study data sample size increases.
+As show in the figure below, this correction factor produces the desired quality of maintaining that small pilot studies should *increase* sample size, and that sample size suggestions level off as pilot study data sample size increases.
 
@@
 
-ggplot(summary_long %>% filter(source == "Low"), aes(original_n, new_sample, color = name)) + 
+ggplot(summary_long %>% filter(source == "Low Variance") %>% filter(scale == 1), aes(original_n, new_sample, color = name)) + 
   geom_point() +  
@@
 We have portrayed that this procedure, with a correction factor, can perform as desired.
-However, within real scenarios, researchers will only have one pilot sample, not many as show here.
+However, within real scenarios, researchers will only have one pilot sample, not the various simulated samples shown above.
 What should the researcher do to correct their sample size on their own pilot data?
@@
 First, the corrected sample size was predicted by the original suggested sample size.
-Next, the standard deviation of the item standard deviations was added to the equation.
+Next, the standard deviation of the item standard deviations was added to the equation to recreate heterogeneity estimates.
 Last, we included the pilot sample size.
+The scale of the data is embedded into the standard deviation of the items (*r* = `r round(cor(summary_long$scale, summary_long$sd_item), digits = 2)`), and therefore, this variable was not included separately.
+The standard deviation of the item's standard deviation embeds both heterogeneity and potential scale size.
 
@@
 
-user_model2 <- lm(new_sample ~ sample_size + sd_original, data = summary_long)
+user_model2 <- lm(new_sample ~ sample_size + sd_item, data = summary_long)
 user_print2 <- apa_print(user_model2)
 
-user_model3 <- lm(new_sample ~ sample_size + sd_original + original_n, data = summary_long)
+user_model3 <- lm(new_sample ~ sample_size + sd_item + original_n, data = summary_long)
 user_print3 <- apa_print(user_model3)
-
-user_model4 <- lm(new_sample ~ sample_size + sd_original + original_n + scale, data = summary_long)
-user_print4 <- apa_print(user_model4)
 
@@
 
-The first model using original sample size to predict new sample size was significant, `r user_print$full_result$modelfit$r2`, capturing nearly 90% of the variance.
+The first model using original sample size to predict new sample size was significant, `r user_print$full_result$modelfit$r2`, capturing nearly 90% of the variance, `r user_print$estimate$sample_size`.
 The second model with item standard deviation was better then the first model *F*(`r change_table$df[2]`, `r change_table$res.df[2]`) = `r change_table$statistic[2]`, *p* < .001, `r user_print2$estimate$modelfit$r2`.
+The item standard deviation predictor was significant, `r user_print2$full_result$sd_item`.
 The addition of the original pilot sample size was also significant, *F*(`r change_table$df[3]`, `r change_table$res.df[3]`) = `r change_table$statistic[3]`, *p* < .001, `r user_print3$estimate$modelfit$r2`.
 
-As shown in the table below, the new suggested sample size is proportional to the original suggested sample size (i.e., *b* < 1), which reduces the sample size suggestion.
-As variability increases, the suggested sample size also increases to capture differences in heterogeneity shown above.
+As shown in the table below for our final model, the new suggested sample size is proportional to the original suggested sample size (i.e., *b* < 1), which reduces the sample size suggestion.
+As variability increases, the suggested sample size also increases to capture differences in heterogeneity shown above; however, this predictor is not significant in the final model, and only contributes a small portion of overall variance.
 Last, in order to correct for large pilot data, the original pilot sample size decreases the new suggested sample size.
@@
 for (cutoff in unique(summary_long$name)){
-  by_cutoff[[cutoff]] <- lm(new_sample ~ sample_size + sd_original + original_n, data = summary_long  %>% filter(name == cutoff))
+  by_cutoff[[cutoff]] <- lm(new_sample ~ sample_size + sd_item + original_n, data = summary_long  %>% filter(name == cutoff))
   R2[cutoff] <- summary(by_cutoff[[cutoff]])$r.squared
@@
 Last, we examine the question of an appropriate SE decile.
-All graphs for power, variability, and correction are presented below.
-If we examine the $R^2$ values for each decile of our regression equation separately, we find that the 10% (`r round(R2[["Decile 10"]], digits = 3)`) and 30% deciles (`r round(R2[["Decile 30"]], digits = 3)`) represent the best match to our corrected sample size suggestions.
-The 10% is likely unfeasible for many researchers, as the sample sizes are often quite large.
-The 30% decile, in the corrected format, appears to meet all goals: 1) increases with heterogeneity, 2) higher suggested values for lower samples and a leveling effect at larger pilot data, and ...
+All graphs for power, heterogeneity, scale, and correction are presented below.
+First, the 0%, 10%, and 20% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale, and heterogeneity.If we examine the $R^2$ values for each decile of our regression equation separately, we find that the 50% (`r round(R2[["Decile 50"]], digits = 3)`) represents the best match to our corrected sample size suggestions.
+The 50% decile, in the corrected format, appears to meet all goals: 1) increases with heterogeneity and scale of data, and 2) higher suggested values for small original samples and a leveling effect at larger pilot data.
+
+The formula for finding the corrected sample size using a 50% decile is: $Final Sample = 0.637 \times X_{Suggested Sample Size} + 0.002 \times X_{SD Items} + -0.546 \times X_{Pilot Sample Size}$.
+
+## Conclusions
+
+1) Take your original pilot data
+2) Find a critical score for an accurately measured item using SE _*at the 50% decile of the item SE values*_.
+3) Sample from your pilot data using sample sizes starting at 20 participants and increase in units of 5 up to a value you feel is the maximum sample size. 
+4) Calculate the number of items that meet your critical score at each sample size.
+5) Find the smallest sample size for 80%, 85%, 90%, and 95% power. 
+6) _*Apply the correction to the smallest sample sizes*_.
+7) Designate your minimum sample size from these values, the stopping rule used from step 2, and the maximum sample size from these values.
 
@@
 ```{r}
-ggplot(summary_long %>% filter(source == "Low"), aes(original_n, sample_size, color = name)) + 
+ggplot(summary_long %>% filter(source == "Low Variance"), aes(original_n, sample_size, color = name)) + 
   geom_point() +  
@@
   theme_classic() + 
-  facet_wrap(~ power) + 
+  facet_wrap(~ scale2*power) + 
   xlab("Original Sample Size") + 
@@
 ```{r}
-ggplot(summary_long %>% filter(source == "Low"), aes(original_n, new_sample, color = name)) + 
+ggplot(summary_long %>% filter(source == "Low Variance"), aes(original_n, new_sample, color = name)) + 
   geom_point() +  
@@
 ```{r}
-ggplot(summary_long %>% filter(source == "Medium"), aes(original_n, sample_size, color = name)) + 
+ggplot(summary_long %>% filter(source == "Medium Variance"), aes(original_n, sample_size, color = name)) + 
   geom_point() +  
@@
 
-ggplot(summary_long %>% filter(source == "Medium"), aes(original_n, new_sample, color = name)) + 
+ggplot(summary_long %>% filter(source == "Medium Variance"), aes(original_n, new_sample, color = name)) + 
   geom_point() +  
@@
 ```{r}
-ggplot(summary_long %>% filter(source == "High"), aes(original_n, sample_size, color = name)) + 
+ggplot(summary_long %>% filter(source == "High Variance"), aes(original_n, sample_size, color = name)) + 
   geom_point() +  
@@
 ```{r}
-ggplot(summary_long %>% filter(source == "High"), aes(original_n, new_sample, color = name)) + 
+ggplot(summary_long %>% filter(source == "High Variance"), aes(original_n, new_sample, color = name)) + 
   geom_point() +  

### 2022-06-22 manuscript_draft.Rmd
updating manuscript

--- 2022-06-17 simulate_cutoff.Rmd
@@
 ---
-title: "Simulate Cut Off"
-author: "Erin M. Buchanan"
-date: "Last Knitted: `r Sys.Date()`"
-output: html_document
+title             : "Power to the Stimuli: Not the Effect"
+shorttitle        : "Title"
+
+author: 
+  - name          : "Erin M. Buchanan"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       : "326 Market St., Harrisburg, PA, 17101"
+    email         : "ebuchanan@harrisburgu.edu"
+    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+      - Conceptualization
+      - Writing - Original Draft Preparation
+      - Writing - Review & Editing
+      - Resources
+      - Validation 
+      - Visualization 
+      - Project Administration 
+      - Formal Analysis
+  - name          : "Other Folks as Per Order on Doc"
+    affiliation   : "2"
+    role:
+      - Writing - Review & Editing
+      - Data Curation
+      - Resources
+
+affiliation:
+  - id            : "1"
+    institution   : "Harrisburg University of Science and Technology"
+  - id            : "2"
+    institution   : "Other Instituions"
+
+authornote: |
+  The authors would like to thank K.D. Valentine for her assistance in formulation of correction scores.
+
+abstract: |
+  We will add an abstract that explains that we are going to 1) talk about this cool procedure, 2) show some simulations that the procedure works, and 3) give two examples of the procedure in action.
+You should try it out!
+
+keywords          : "power, sampling, accuracy in parameter estimation"
+
+# bibliography      : ["r-references.bib"]
+
+floatsintext      : no
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+linenumbers       : yes
+mask              : no
+draft             : no
+
+documentclass     : "apa6"
+classoption       : "man"
+output            : papaja::apa6_docx
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-## Libraries
-
-```{r}
+```{r setup, include = FALSE}
 set.seed(895893)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
 library(dplyr)
@@
 library(papaja)
+library(rio)
 ```
 
-# Introduction
+This part we will put in a real introduction. Things we need to discuss:
 
-This document outlines using the suggested procedure to simulate the number of participants necessary to accurately measure items.
-There are two key issues these ideas should address that we know about power:
+  - Repeated measures data with items
+  - Accuracy in Parameter Estimation 
+  - Adaptive sampling
+  - MLM designs account for item heterogeneity but a lot of people anova
 
-1) we should see differences in projected sample sizes based on the variability in the variance for those items (i.e., heterogeneity should increase projected sample size)
-2) we should see projected sample sizes that "level off" when pilot data increases.
+## Simulating Sample Size
+
+Using ideas from accuracy in parameter estimation, we suggest the following procedure to determine a sample size for each item:
+
+1) Use pilot data that closely resembles your intended data collection, on the same or similar items that will be used in the study.
+In this procedure, we will assume that the pilot data is representative of a larger population of sampled items that you intend to assess.
+2) Calculate the standard error of each item from the pilot data to create a cutoff score for when an item is "accurately measured".
+The simulations below will explore what criterion to use when determining the cutoff score from the pilot data.
+3) Sample, with replacement, from your pilot data using sample sizes starting at 20 participants and increase in small units (e.g., 20, 25, 30) up to a value that you consider the maximum sample size.
+We will demonstrate example maximum sample sizes based on the data simulation below; however, a practical maximum sample size may be determined by time (e.g, one semester data collection) or researcher resources (e.g., 200 participants worth of funding).
+While 20 participants would likely represent an underpowered study, we simply suggest this starting minimum for simulation purposes.
+4) For each simulated sample, calculate the standard error for each item, and use these values to ascertain the percentage of items that meet the cutoff score determined in step 2. 
+5) Find the minimum sample size that meets 80%, 85%, 90%, and 95% of the items.
+We recommend these scores to ensure that most items are accurately measured, in a similar vein to common power criteria suggestions.
+Each researcher can determine which of these is their minimum or maximum sample size (e.g., individual can choose to use 80% as a minimum and 90% as a maximum).
+6) Report these values, and designate a minimum sample size, the cutoff criterion, and the maximum sample size.
+Each researcher should also report if they plan to use an adaptive design, which would stop data collection after meeting the cutoff criterion for each item.
+
+## Key Issues
+
+Given the long history of research on power, there are a few key issues that this procedure should address:
+
+1) We should see differences in projected sample sizes based on the variability in the variance for those items (i.e., heterogeneity should increase projected sample size).
+2) We should see projected sample sizes that "level off" when pilot data increases.
 As with regular power estimates, studies can be "overpowered" to detect an effect, and this same idea should be present.
 For example, if one has a 500 person pilot study, our simulations should suggest a point at which items are likely measured well, which may have happened well before 500.
-
-# Explain the Procedure
-
-1) Take your original pilot data.
-2) Find a critical score for an accurately measured item using SE.
-3) Sample from your pilot data using sample sizes starting at 20 participants and increase in units of 5 up to a value you feel is the maximum sample size.
-4) Calculate the number of items that meet your critical score at each sample size.
-5) Find the smallest sample size for 80%, 85%, 90%, and 95% power. 
-6) Designate your minimum sample size from these values, the stopping rule used from step 2, and the maximum sample size from these values.
 
@@
 While there are many potential scales, these three represent a large number of potential variables in the social sciences.
-As we are suggesting variances as a key factor for estimating sample sizes, the scale of the data is influential on the amount of *potential* variance.
+As we are suggesting item variances as a key factor for estimating sample sizes, the scale of the data is influential on the amount of *potential* variance.
 Smaller ranges of data (1-7) cannot necessarily have the same variance as larger ranges (0-100).
@@
 For small scales, the $\sigma$ = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items.
-For the medium scale of data, $\sigma$ = 20 with a variance of 4, 8, and 12.
-Last, for the large scale of data, $\sigma$ = 350 with a variance of 100, 150, and 200 for heterogeneity.
+For the medium scale of data, $\sigma$ = 25 with a variance of 4, 8, and 16.
+Last, for the large scale of data, $\sigma$ = 400 with a variance of 50, 100, and 200 for heterogeneity.
 
-```{r}
+```{r sim_pop, include = F, echo = F}
 # small potential variability overall, sort of 1-7ish scale
@@
 mu2 <- rnorm(30, 50, 10)
-sigma1.2 <- rnorm(30, 20, 4)
-sigma2.2 <- rnorm(30, 20, 8)
-sigma3.2 <- rnorm(30, 20, 12)
+sigma1.2 <- rnorm(30, 25, 4)
+sigma2.2 <- rnorm(30, 25, 8)
+sigma3.2 <- rnorm(30, 25, 16)
+
+while(sum(sigma3.2 < 0) > 0){
+  sigma3.2 <- rnorm(30, 25, 16)
+}
 
 # large potential variability in the 1000s scale
-
 mu3 <- rnorm(30, 1000, 150)
-sigma1.3 <- rnorm(30, 350, 100)
-sigma2.3 <- rnorm(30, 350, 150)
-sigma3.3 <- rnorm(30, 350, 200)
+sigma1.3 <- rnorm(30, 400, 50)
+sigma2.3 <- rnorm(30, 400, 100)
+sigma3.3 <- rnorm(30, 400, 200)
 
 while(sum(sigma3.3 < 0) > 0){
-  sigma3.3 <- rnorm(30, 350, 200)
+  sigma3.3 <- rnorm(30, 400, 200)
 }
@@
   )
-
-#evidence that they are simulated correctly
-tapply(population1$score, list(population1$item, 
-                               population1$scale), mean)
-tapply(population1$score, list(population1$item,
-                               population1$scale), sd)
-
-tapply(population2$score, list(population2$item, 
-                               population2$scale), mean)
-tapply(population2$score, list(population2$item,
-                               population2$scale), sd)
-
-tapply(population3$score, list(population3$item, 
-                               population3$scale), mean)
-tapply(population3$score, list(population3$item,
-                               population3$scale), sd) 
 ```
@@
 
-```{r}
+```{r sim_sample, include = F, echo = F}
 # create pilot samples from 20 to 100
@@
 
-*AIPE Criterions*.
+*Cutoff Score Criterions*.
 The standard errors of each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as standard error functions as the main component in the formula for normal distribution confidence intervals.
-Standard errors were calculated at each decile of the items up to 90% (0% smallest SE, 10% ..., 90% largest SE).
-The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet AIPE goals, while the higher deciles would represent less strict criterions for AIPE goals.
+Standard errors were calculated at each decile of the items up to 90% (i.e., 0% smallest SE, 10% ..., 90% largest SE).
+The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criterions for cutoff scores.
 
-```{r}
+```{r calc_SE, include = F, echo = F}
 # calculate the SEs and the cutoff scores 
@@
 
-## AIPE Simulation
+## Researcher Sample Simulation
 
@@
 
-```{r}
+```{r sim_sim, include = F, echo = F}
 # sequence of sample sizes to try
@@
 Next, the percent of items that fall below the cutoff scores, and thus, would be considered "well-measured" were calculated for each decile by sample.
+From this data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90%, and 95% of the items fall below the cutoff criterion.
+These values were chosen as popular measures of "power" in which one could determine the minimum suggested sample size (potentially 80% of items) and the maximum suggested sample size (potentially 90%).
 
-```{r}
+```{r calc_percent, include = F, echo = F}
 summary_list1 <- summary_list2 <- summary_list3 <- list()
@@
 
-From this data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90%, and 95% of the items fall below the cutoff criterion.
-These values were chosen as popular measures of "power" in which one could determine the minimum suggested sample size (potentially 80% of items) and the maximum suggested sample size (potentially 90%).
-
-```{r}
+```{r calc_percent_combine, include = F, echo = F}
 summary_long_80 <- summary_DF %>% 
@@
 
-The figure below illustrates the potential minimum sample size for 80% of items to achieve a desired cutoff score.
+Figure \@ref(fig:item-figure) illustrates the potential minimum sample size for 80% of items to achieve a desired cutoff score.
 The black dots denote the original sample size against the suggested sample size.
@@
 
-```{r}
+```{r item-figure, echo = F, fig.cap="Add a good caption here."}
 summary_long$source <- factor(summary_long$source, 
@@
   facet_wrap(~ source*scale2) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
+  xlab("Pilot Sample Size") + 
+  ylab("Projected Sample Size") + 
   scale_color_discrete(name = "Cutoff Score")
@@
 
-## Sample Size Sensitivity to Pilot Data Size
+## Projected Sample Size Sensitivity to Pilot Sample Size
 
@@
 Larger pilot data is more informative, and therefore, we should expect a lower projected sample size.
-As shown in the figure below for only the low variability and small scale data, we do not find this effect.
-These simplistic simulations from the pilot data would nearly always suggest a larger sample size - mostly in a linear trend increasing with sample sizes.
-This result comes from the nature of the procedure - if we base our estimates on some SE cutoff, we will almost always need a bit more people for items to meet those goals.
+As shown in Figure \@ref(fig:sensitive-figure) for only the low variability and small scale data, we do not find this effect.
+These simulations from the pilot data would nearly always suggest a larger sample size - mostly in a linear trend increasing with sample sizes.
+This result comes from the nature of the procedure - if we base our estimates on a SE cutoff, we will almost always need a bit more people for items to meet those goals.
 This result does not achieve our second goal.
 
-```{r}
+```{r sensitive-figure, echo = F, fig.cap="Add good description here."}
 ggplot(summary_long %>% filter(source == "Low Variance") %>% filter(scale == 1), aes(original_n, sample_size, color = name)) + 
@@
   facet_wrap(~ power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
+  xlab("Pilot Sample Size") + 
+  ylab("Projected Sample Size") + 
   scale_color_discrete(name = "Cutoff Score")
@@
 Therefore, we suggest using a correction factor on the simulation procedure to account for the known asymptotic nature of power (i.e., at larger sample sizes power increases level off).
-For this function, we combined a correction factor for upward biasing of effect sizes (Hedges' correction) with the formula for exponential decay calculations.
-The decay factor is calculated as follows:
+For this function in our simulation study, we combined a correction factor for upward biasing of effect sizes (Hedges' correction) with the formula for exponential decay calculations.
+The decay factor was calculated as follows:
 
-$$ 1 - \sqrt{\frac{N_{pilot} - min(N_{sim})}{N_{pilot}}}^{log_2(N_{pilot})}$$
-$N_{pilot}$ indicates the sample size of the pilot data minus the minimum sample size for simulation to ensure that the smallest sample sizes do not decay (e.g., the formula zeroes out).
+$$ 1 - \sqrt{\frac{N_{Pilot} - min(N_{Simulation})}{N_{Pilot}}}^{log_2(N_{Pilot})}$$
+
+$N_{Pilot}$ indicates the sample size of the pilot data minus the minimum simulated sample size to ensure that the smallest sample sizes do not decay (i.e., the formula zeroes out).
 This value is raised to the power of $log_2$ of the sample size of the pilot data, which decreases the impact of the decay to smaller increments for increasing sample sizes.
-This value is then multiplied by the proposed sample size.
-As show in the figure below, this correction factor produces the desired quality of maintaining that small pilot studies should *increase* sample size, and that sample size suggestions level off as pilot study data sample size increases.
+This value is then multiplied by the projected sample size.
+As shown in Figure \@ref(fig:corrected-figure), this correction factor produces the desired quality of maintaining that small pilot studies should *increase* sample size, and that sample size suggestions level off as pilot study data sample size increases.
 
-```{r}
+```{r corrected-figure, echo = F, fig.cap="A corrected figure update this caption."}
 decay <- 1-sqrt((summary_long$original_n-20)/summary_long$original_n)^log2(summary_long$original_n)
@@
   facet_wrap(~ power) + 
-  xlab("Original Sample Size") + 
+  xlab("Pilot Sample Size") + 
   ylab("Suggested Sample Size") + 
@@
 However, within real scenarios, researchers will only have one pilot sample, not the various simulated samples shown above.
-What should the researcher do to correct their sample size on their own pilot data?
+What should the researcher do to correct their projected sample size from their own pilot data simulations?
 
-To explore if we could recover the new suggested sample size from data a researcher would have, we used linear models to create a formula for calculation.
-First, the corrected sample size was predicted by the original suggested sample size.
+To explore if we could recover the new projected sample size from data a researcher would have, we used linear models to create a formula for researcher correction.
+First, the corrected projected sample size was predicted by the original projected sample size.
 Next, the standard deviation of the item standard deviations was added to the equation to recreate heterogeneity estimates.
+The scale of the data is embedded into the standard deviation of the items (*r* = `r round(cor(summary_long$scale, summary_long$sd_item), digits = 2)`), and therefore, this variable was not included separately.
 Last, we included the pilot sample size.
-The scale of the data is embedded into the standard deviation of the items (*r* = `r round(cor(summary_long$scale, summary_long$sd_item), digits = 2)`), and therefore, this variable was not included separately.
-The standard deviation of the item's standard deviation embeds both heterogeneity and potential scale size.
-
-```{r}
+```{r regressions, include = F, echo = F}
 user_model <- lm(new_sample ~ sample_size, data = summary_long)
 user_print <- apa_print(user_model)
+user_sum <- summary(user_model)
 
@@
 
-The first model using original sample size to predict new sample size was significant, `r user_print$full_result$modelfit$r2`, capturing nearly 90% of the variance, `r user_print$estimate$sample_size`.
-The second model with item standard deviation was better then the first model *F*(`r change_table$df[2]`, `r change_table$res.df[2]`) = `r change_table$statistic[2]`, *p* < .001, `r user_print2$estimate$modelfit$r2`.
+The first model using pilot sample size to predict new sample size was significant, *F*(`r user_sum$fstatistic[2]`, `r user_sum$fstatistic[3]`) = `r round(user_sum$fstatistic[1], digits = 2)`, *p* < .001, `r user_print$estimate$modelfit$r2`, capturing nearly 90% of the variance, `r user_print$estimate$sample_size`.
+The second model with item standard deviation was better than the first model *F*(`r change_table$df[2]`, `r format(change_table$res.df[2], big.mark = "")`) = `r round(change_table$statistic[2], digits = 2)`, *p* < .001, `r user_print2$estimate$modelfit$r2`.
 The item standard deviation predictor was significant, `r user_print2$full_result$sd_item`.
-The addition of the original pilot sample size was also significant, *F*(`r change_table$df[3]`, `r change_table$res.df[3]`) = `r change_table$statistic[3]`, *p* < .001, `r user_print3$estimate$modelfit$r2`.
+The addition of the original pilot sample size was also significant, *F*(`r change_table$df[3]`, `r format(change_table$res.df[3], big.mark = "")`) = `r round(change_table$statistic[3], digits = 2)`, *p* < .001, `r user_print3$estimate$modelfit$r2`.
 
-As shown in the table below for our final model, the new suggested sample size is proportional to the original suggested sample size (i.e., *b* < 1), which reduces the sample size suggestion.
+As shown in the final model Table \@ref(tab:table-correction), the new suggested sample size is proportional to the original suggested sample size (i.e., *b* < 1), which reduces the sample size suggestion.
 As variability increases, the suggested sample size also increases to capture differences in heterogeneity shown above; however, this predictor is not significant in the final model, and only contributes a small portion of overall variance.
@@
 
-```{r results = 'asis'}
-apa_table(tidy(user_model3))
+```{r table-correction, results = 'asis'}
+table_correct <- tidy(user_model3)
+table_correct$term <- c("Intercept", "Projected Sample Size", 
+                        "Item SD", "Pilot Sample Size")
+table_correct$estimate <- printnum(table_correct$estimate, digits = 3)
+table_correct$std.error <- printnum(table_correct$std.error, digits = 3)
+table_correct$statistic <- printnum(table_correct$statistic, digits = 3)
+table_correct$p.value <- printnum(table_correct$p.value, digits = 3,
+                                  zero = FALSE, gt1 = F)
+apa_table(table_correct, 
+          col.names = c("Term", "Estimate", "$SD$", "$t$", "$p$"),
+          caption = "Parameters for All Decile Cutoff Scores")
 ```
@@
 
-```{r}
+```{r R2_cutoff, include = F, echo = F}
 by_cutoff <- list()
@@
 }
-
-R2
 ```
@@
 Last, we examine the question of an appropriate SE decile.
-All graphs for power, heterogeneity, scale, and correction are presented below.
-First, the 0%, 10%, and 20% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale, and heterogeneity.If we examine the $R^2$ values for each decile of our regression equation separately, we find that the 50% (`r round(R2[["Decile 50"]], digits = 3)`) represents the best match to our corrected sample size suggestions.
+All graphs for power, heterogeneity, scale, and correction are presented online.
+First, the 0%, 10%, and 20% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale, and heterogeneity.
+If we examine the $R^2$ values for each decile of our regression equation separately, we find that the 50% (`r round(R2[["Decile 50"]], digits = 3)`) represents the best match to our corrected sample size suggestions.
 The 50% decile, in the corrected format, appears to meet all goals: 1) increases with heterogeneity and scale of data, and 2) higher suggested values for small original samples and a leveling effect at larger pilot data.
+Figure \@ref(fig:decile-figure) illustrates the corrected scores for simulations at the 50% decile recommended cutoff for item standard errors.
 
-The formula for finding the corrected sample size using a 50% decile is: $Final Sample = 0.637 \times X_{Suggested Sample Size} + 0.002 \times X_{SD Items} + -0.546 \times X_{Pilot Sample Size}$.
-
-## Conclusions
-
-1) Take your original pilot data
-2) Find a critical score for an accurately measured item using SE _*at the 50% decile of the item SE values*_.
-3) Sample from your pilot data using sample sizes starting at 20 participants and increase in units of 5 up to a value you feel is the maximum sample size. 
-4) Calculate the number of items that meet your critical score at each sample size.
-5) Find the smallest sample size for 80%, 85%, 90%, and 95% power. 
-6) _*Apply the correction to the smallest sample sizes*_.
-7) Designate your minimum sample size from these values, the stopping rule used from step 2, and the maximum sample size from these values.
-
-### Low Variability
-
-```{r}
-ggplot(summary_long %>% filter(source == "Low Variance"), aes(original_n, sample_size, color = name)) + 
+```{r decile-figure, fig.cap = "A picture of the 50% cutoff."}
+ggplot((summary_long %>% filter(name == "Decile 50")), 
+       aes(original_n, new_sample, color = source)) + 
   geom_point() + 
@@
   facet_wrap(~ scale2*power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
+  xlab("Pilot Sample Size") + 
+  ylab("Projected Sample Size") + 
+  scale_color_discrete(name = "Heterogeneity") +
+  theme(legend.position="bottom")
+
 ```
 
-```{r}
-ggplot(summary_long %>% filter(source == "Low Variance"), aes(original_n, new_sample, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ scale2*power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
+The formula for finding the corrected sample size using a 50% decile is: $N_{Corrected Projected} = 39.269 + 0.700 \times X_{N_{Projected}} + 0.003 \times X_{SD Items} - 0.694 \times X_{N_{Pilot}}$.
+The suggested sample size will be estimated from the 80%, 85%, 90%, or 95% selection at the 50% decile as shown above.
+The item SD can be calculated directly from the data, and the pilot sample size is the sample size of the data from which a researcher is simulating their samples.
+Therefore, we will recommend the 50% decile of the item standard errors for step 2 of our suggested simulation procedure, and to correct the projected sample sizes found in step 5 using the correction equation above.
+While the estimated coefficients could change given variations on our simulation parameters, the general size and pattern of coefficients was consistent, and therefore, we believe this correction equation should work for a variety of use cases.
+
+# Examples
+
+```{r import_data, include = F, echo = F}
+conDF <- import("./data/concreteness_trial.rda")
+elpDF <- import("./data/ELPDecisionData.zip")
 ```
 
-### Medium Varability
+In this section, we provide two examples of the suggested procedure.
+The first example includes concreteness ratings from Brysbaert et al.
+[CITE, 2014].
+Instructions given to participants denoted the difference between concrete (i.e., "refers to something that exists in reality") and abstract (i.e., "something you cannot experience directly through your senses or actions") terms.
+Participants were then asked to rate concreteness of terms using a 1 (abstract) to 5 (concrete) scale.
+This data represents a small scale dataset that could be used as pilot data for a study using concrete word ratings.
+The data is available at OSF LINK [CITE].
+The second dataset includes a large scale dataset with response latencies, the English Lexicon Project [CITE, Balota et al.].
+The English Lexicon Project consists of lexical decision response latencies for English words.
+In a lexical decision task, participants simply select "word" for real words (e.g., *dog*) and "nonword" for pseudowords (e.g., *wug*).
+The trial level data is available here [https://elexicon.wustl.edu/, CITE].
+Critically, in each of these examples, the individual trial level data for each item is available to simulate and calculate standard errors on.
+Data that has been summarized could potentially be used, as long as the original standard deviations for each item were present.
+From the mean and standard deviation for each item, a simulated pilot dataset could be generated for estimating new sample sizes.
+All code to estimate sample sizes is provided on our OSF page.
 
-```{r}
-ggplot(summary_long %>% filter(source == "Medium Variance"), aes(original_n, sample_size, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ scale2*power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
+## Concreteness Ratings
+
+```{r con_estimate, echo = F, include = F}
+# sample the data
+con_words <- sample(unique(conDF$Word), size = 100, replace = F)
+con_small <- conDF %>% filter(Word %in% con_words)
+con_small$Word <- droplevels(con_small$Word)
+pilot_size <- round(mean(tapply(con_small$Rating, con_small$Word, length)))
+
+# figure out data loss
+con_small$Rating[con_small$Rating == "n" | con_small$Rating == "N"] <- NA
+con_small$Rating <- as.numeric(con_small$Rating)
+data_loss <- con_small %>% 
+  group_by(Word) %>% 
+  summarize(percent_correct = sum(!is.na(Rating))/n(), .groups = "keep")
+
+con_use <- con_small %>% filter(!is.na(Rating))
+
+# cutoff score 
+SE <- tapply(con_use$Rating, con_use$Word, function (x) { sd(x)/sqrt(length(x))})
+cutoff <- quantile(SE, probs = .50)
 ```
 
-```{r}
-summary_long$new_sample <- summary_long$sample_size*decay
+The concreteness ratings data includes `r length(unique(conDF$Word))` concepts that were rated for their concreteness.
+In our fictional study for this example, we selected 100 random words to show participants.
+In the original study, not every participant rated every word, which created uneven sample sizes for each word.
+In our random sample of 100 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size` as our pilot sample size for this example.
+All "do not know" ratings were set as missing data.
+The 50% decile for items standard error was `r round(cutoff, digits = 3)` for our cutoff criterion.
 
-ggplot(summary_long %>% filter(source == "Medium Variance"), aes(original_n, new_sample, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ scale2*power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
+```{r con_sample, include = F, echo = F}
+# sequence of sample sizes to try
+samplesize_values <- seq(20, 300, 5)
+
+# create a blank table for us to save the values in 
+sim_table <- matrix(NA, 
+                    nrow = length(samplesize_values), 
+                    ncol = length(unique(con_use$Word)))
+
+# make it a data frame
+sim_table <- as.data.frame(sim_table)
+
+# add a place for sample size values 
+sim_table$sample_size <- NA
+
+# loop over sample sizes
+for (i in 1:length(samplesize_values)){
+
+  # temp dataframe that samples and summarizes
+  temp <- con_use %>% 
+    group_by(Word) %>% 
+    sample_n(samplesize_values[i], replace = T) %>% 
+    summarize(se = sd(Rating)/sqrt(length(Rating)))
+
+  colnames(sim_table)[1:length(unique(con_use$Word))] <- temp$Word
+  sim_table[i, 1:length(unique(con_use$Word))] <- temp$se
+  sim_table[i, "sample_size"] <- samplesize_values[i]
+  }
+
+# figure out potential samples
+final_sample <- 
+  sim_table %>% 
+  pivot_longer(cols = -c(sample_size)) %>% 
+  rename(item = name, se = value) %>% 
+  group_by(sample_size) %>% 
+  summarize(Percent_Below = sum(se <= cutoff)/length(unique(con_use$Word))) %>% 
+  filter(Percent_Below >= .80) %>% 
+  arrange(Percent_Below)
+
+# apply correction
+# calculate the SD of the item's SD 
+sd_items <- sd(tapply(con_use$Rating, con_use$Word, sd))
+
+final_sample$new_sample <- round(39.369 + 0.700*final_sample$sample_size + 0.003*sd_items - 0.694*pilot_size)
+
+final_sample$data_loss_sample <- round(final_sample$new_sample * 1/mean(data_loss$percent_correct))
 ```
 
-### High Variability
+The pilot data was then simulated, with replacement, with samples from 20 to 300 increasing in units of 5.
+On each sample, the percent of items below the cutoff score were calculated.
+After applying our correction equation, we find that a sample size of `r final_sample$new_sample[1]` would allow for at least 80% of items to meet the cutoff criterion.
+The sample sizes for 85% (`r final_sample$new_sample[2]`), 90% (`r final_sample$new_sample[2]`), and 95% (`r final_sample$new_sample[3]`) are also options for sample size suggestions.
+Finally, we calculated the potential amount of data retention given that participants could indicate they did not know a word ($M_{correct}$ = `r round(mean(data_loss$percent_correct), digits = 2)`, *SD* = `r round(sd(data_loss$percent_correct), digits = 2)`).
+In order to account for this facet, the potential sample sizes were multiplied by 1/`r round(mean(data_loss$percent_correct), digits = 2)`, which results in a suggested sample of `r final_sample$data_loss_sample[1]`, `r final_sample$data_loss_sample[2]`, and `r final_sample$data_loss_sample[3]`.
+Therefore, we could designate our minimum sample per item as `r final_sample$data_loss_sample[1]`, stopping rule of `r cutoff`, and maximum sample size of `r final_sample$data_loss_sample[3]`.
 
-```{r}
-ggplot(summary_long %>% filter(source == "High Variance"), aes(original_n, sample_size, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ scale2*power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
+## Response Latencies
+
+```{r elp_estimate, echo = F, include = F}
+# sample the data
+elp_fake <- sample(unique(elpDF$Stimulus[elpDF$Type == 0]), size = 500, replace = F)
+elp_real <- sample(unique(elpDF$Stimulus[elpDF$Type == 1]), size = 500, replace = F)
+elp_small <- elpDF %>% filter(Stimulus %in% c(elp_fake, elp_real))
+pilot_size <- round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)))
+
+# figure out data loss
+elp_small$RT[elp_small$Accuracy == "0"] <- NA
+data_loss <- elp_small %>% 
+  group_by(Stimulus, Type) %>% 
+  summarize(percent_correct = sum(!is.na(RT))/n(), .groups = "keep")
+
+elp_use <- elp_small %>% filter(!is.na(RT))
+
+# cutoff score 
+SE <- tapply(elp_use$RT, elp_use$Stimulus, function (x) { sd(x)/sqrt(length(x))})
+cutoff <- quantile(SE, probs = .50, na.rm = T)
+
+SE_word <- tapply(elp_use$RT[elp_use$Type == 1], elp_use$Stimulus[elp_use$Type == 1], function (x) { sd(x)/sqrt(length(x))})
+SE_nw <- tapply(elp_use$RT[elp_use$Type == 0], elp_use$Stimulus[elp_use$Type == 0], function (x) { sd(x)/sqrt(length(x))})
+cutoff_word <- quantile(SE_word, probs = .50, na.rm = T)
+cutoff_nw <- quantile(SE_nw, probs = .50, na.rm = T)
 ```
 
-```{r}
-ggplot(summary_long %>% filter(source == "High Variance"), aes(original_n, new_sample, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ scale2*power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
+The ELP response latency data includes `r length(unique(elpDF$Stimulus))` word-forms, `r length(unique(elpDF$Stimulus[elpDF$Type == 0]))` that are listed as non-words, and `r length(unique(elpDF$Stimulus[elpDF$Type == 1]))` real words.
+For our example study, we will randomly select 500 real words and 500 non-words to show participants.
+The average pilot sample size for this random sample was `r format(round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2), nsmall = 2)` (*SD* = `r round(sd(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2)`), and *n* = `r pilot_size` will be our pilot size for this example.
+Again, participants are expected to make mistakes, and we calculated percent correct as `r round(mean(data_loss$percent_correct), digits = 2)`, which was roughly even in the two stimulus categories: $M_{word}$ = `r round(mean(data_loss$percent_correct[data_loss$Type == 1]), digits = 2)` and $M_{non-word}$ = `r round(mean(data_loss$percent_correct[data_loss$Type == 0]), digits = 2)`.
+The 50% decile for items standard error was `r round(cutoff, digits = 3)` for our cutoff criterion.
+We additionally checked to ensure that the two stimulus types did not have very different cutoff criterions: 50% decile $SE_{words}$ = `r round(cutoff_word, digits = 3)`, 50% decile $SE_{nonwords}$ = `r round(cutoff_nw, digits =3)`.
+In this scenario, we could choose to go with the lower SE to be more conservative (i.e., higher projected sample size).
+Given the values were close for large scale data, we used the 50% decile of all stimuli taken together.
+
+```{r elp_sample, include = F, echo = F}
+# sequence of sample sizes to try
+samplesize_values <- seq(20, 300, 5)
+
+# create a blank table for us to save the values in 
+sim_table <- matrix(NA, 
+                    nrow = length(samplesize_values), 
+                    ncol = length(unique(elp_use$Stimulus)))
+
+# make it a data frame
+sim_table <- as.data.frame(sim_table)
+
+# add a place for sample size values 
+sim_table$sample_size <- NA
+
+# loop over sample sizes
+for (i in 1:length(samplesize_values)){
+
+  # temp dataframe that samples and summarizes
+  temp <- elp_use %>% 
+    group_by(Stimulus) %>% 
+    sample_n(samplesize_values[i], replace = T) %>% 
+    summarize(se = sd(RT)/sqrt(length(RT)))
+
+  colnames(sim_table)[1:length(unique(elp_use$Stimulus))] <- temp$Stimulus
+  sim_table[i, 1:length(unique(elp_use$Stimulus))] <- temp$se
+  sim_table[i, "sample_size"] <- samplesize_values[i]
+  }
+
+# figure out potential samples
+final_sample <- 
+  sim_table %>% 
+  pivot_longer(cols = -c(sample_size)) %>% 
+  rename(item = name, se = value) %>% 
+  group_by(sample_size) %>% 
+  summarize(Percent_Below = sum(se <= cutoff)/length(unique(elp_use$Word))) %>% 
+  filter(Percent_Below >= .80) %>% 
+  arrange(Percent_Below)
+
+# apply correction
+# calculate the SD of the item's SD 
+sd_items <- sd(tapply(elp_use$RT, elp_use$Stimulus, sd, na.rm = T), na.rm = T)
+
+final_sample$new_sample <- round(39.369 + 0.700*final_sample$sample_size + 0.003*sd_items - 0.694*pilot_size)
+
+final_sample$data_loss_sample <- round(final_sample$new_sample * 1/mean(data_loss$percent_correct))
 ```
+
+The pilot response latency data was then simulated in the same way as described above.
+After calculating the percent below our cutoff score, we applied the correction to the projected sample sizes.
+A sample size of `r final_sample$new_sample[1]` would equate to 80% of the items reaching our cutoff, along with 85% (`r final_sample$new_sample[2]`), 90% (`r final_sample$new_sample[2]`), and 95% (`r final_sample$new_sample[3]`).
+Again, we adjusted for data loss given that participants are expected to incorrectly answer items, resulting in a suggested sample of `r final_sample$data_loss_sample[1]`, `r final_sample$data_loss_sample[2]`, and `r final_sample$data_loss_sample[3]`.
+One other possible consideration for this study is potential fatigue in showing participants 1000 target items.
+Therefore, we could designate in our research design that each participant will only receive 500 of the target items.
+We would need to double our sample sizes to account for splitting of the items across multiple sets of participants.
+Our minimum sample size for the entire study could be `r final_sample$data_loss_sample[1]*2`, stopping rule of `r round(cutoff, digits = 3)`, and maximum sample size of `r final_sample$data_loss_sample[3]*2`.
+This study would benefit from an adaptive design, where smaller sets items are randomly sampled for participants until they reach the minimum sample size or the cutoff criteria.
+At this point, items are probabilistically sampled (e.g., higher selection probability for items that have not reached a minimum or stopping rule) until all items have reached criteria.
+
+## Additional Materials
+
+While the examples in this manuscript are traditionally cognitive linguistics focused, any research using repeated items can benefit from newer sampling techniques.
+Therefore, we provide XX example vignettes on our OSF page/GitHub site for this manuscript across a range of examples of data types provided by the authors of this manuscript.
+Examples include ...
+ADD HERE AFTER DONE WITH VIGNETTES.
+
+# Discussion
+
+To be added once people say this idea doesn't suck.
+
+\newpage
+
+# References
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+
+<div id="refs" custom-style="Bibliography"></div>
+\endgroup

### 2022-06-22 manuscript_draft.tex
updating manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Power to the Stimuli: Not the Effect},
  pdfauthor={Erin M. Buchanan1 \& Other Folks as Per Order on Doc2},
  pdflang={en-EN},
  pdfkeywords={power, sampling, accuracy in parameter estimation},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{power, sampling, accuracy in parameter estimation}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Power to the Stimuli: Not the Effect}
\author{Erin M. Buchanan\textsuperscript{1} \& Other Folks as Per Order on Doc\textsuperscript{2}}
\date{}

\shorttitle{Title}

\authornote{

The authors would like to thank K.D. Valentine for her assistance in formulation of correction scores.

The authors made the following contributions. Erin M. Buchanan: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing, Resources, Validation, Visualization, Project Administration, Formal Analysis; Other Folks as Per Order on Doc: Writing - Review \& Editing, Data Curation, Resources.

Correspondence concerning this article should be addressed to Erin M. Buchanan, 326 Market St., Harrisburg, PA, 17101. E-mail: \href{mailto:ebuchanan@harrisburgu.edu}{\nolinkurl{ebuchanan@harrisburgu.edu}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Harrisburg University of Science and Technology\\\textsuperscript{2} Other Instituions}

\abstract{%
We will add an abstract that explains that we are going to 1) talk about this cool procedure, 2) show some simulations that the procedure works, and 3) give two examples of the procedure in action. You should try it out!
}

\begin{document}
\maketitle

This part we will put in a real introduction. Things we need to discuss:

\begin{itemize}
\tightlist
\item
  Repeated measures data with items
\item
  Accuracy in Parameter Estimation
\item
  Adaptive sampling
\item
  MLM designs account for item heterogeneity but a lot of people anova
\end{itemize}

\hypertarget{simulating-sample-size}{%
\subsection{Simulating Sample Size}\label{simulating-sample-size}}

Using ideas from accuracy in parameter estimation, we suggest the following procedure to determine a sample size for each item:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Use pilot data that closely resembles your intended data collection, on the same or similar items that will be used in the study. In this procedure, we will assume that the pilot data is representative of a larger population of sampled items that you intend to assess.
\item
  Calculate the standard error of each item from the pilot data to create a cutoff score for when an item is ``accurately measured''. The simulations below will explore what criterion to use when determining the cutoff score from the pilot data.
\item
  Sample, with replacement, from your pilot data using sample sizes starting at 20 participants and increase in small units (e.g., 20, 25, 30) up to a value that you consider the maximum sample size. We will demonstrate example maximum sample sizes based on the data simulation below; however, a practical maximum sample size may be determined by time (e.g, one semester data collection) or researcher resources (e.g., 200 participants worth of funding). While 20 participants would likely represent an underpowered study, we simply suggest this starting minimum for simulation purposes.
\item
  For each simulated sample, calculate the standard error for each item, and use these values to ascertain the percentage of items that meet the cutoff score determined in step 2.
\item
  Find the minimum sample size that meets 80\%, 85\%, 90\%, and 95\% of the items. We recommend these scores to ensure that most items are accurately measured, in a similar vein to common power criteria suggestions. Each researcher can determine which of these is their minimum or maximum sample size (e.g., individual can choose to use 80\% as a minimum and 90\% as a maximum).
\item
  Report these values, and designate a minimum sample size, the cutoff criterion, and the maximum sample size. Each researcher should also report if they plan to use an adaptive design, which would stop data collection after meeting the cutoff criterion for each item.
\end{enumerate}

\hypertarget{key-issues}{%
\subsection{Key Issues}\label{key-issues}}

Given the long history of research on power, there are a few key issues that this procedure should address:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  We should see differences in projected sample sizes based on the variability in the variance for those items (i.e., heterogeneity should increase projected sample size).
\item
  We should see projected sample sizes that ``level off'' when pilot data increases. As with regular power estimates, studies can be ``overpowered'' to detect an effect, and this same idea should be present. For example, if one has a 500 person pilot study, our simulations should suggest a point at which items are likely measured well, which may have happened well before 500.
\end{enumerate}

\hypertarget{method}{%
\section{Method}\label{method}}

\hypertarget{data-simulation}{%
\subsection{Data Simulation}\label{data-simulation}}

\emph{Population}. The data was simulated using the \texttt{rnorm} function assuming a normal distribution for 30 scale type items. Each population was simulated with 1000 data points. No items were rounded for this simulation.

First, the scale of the data was manipulated by creating three sets of scales. The first scale was mimicked after small rating scales (i.e., 1-7 type style) using a \(\mu\) = 4 with a \(\sigma\) = .25 around the mean to create item mean variability. The second scale included a larger potential distribution of scores with a \(\mu\) = 50 (\(\sigma\) = 10) imitating a 0-100 scale. Last, the final scale included a \(\mu\) = 1000 (\(\sigma\) = 150) simulating a study that may include response latency data in the milliseconds. While there are many potential scales, these three represent a large number of potential variables in the social sciences. As we are suggesting item variances as a key factor for estimating sample sizes, the scale of the data is influential on the amount of \emph{potential} variance. Smaller ranges of data (1-7) cannot necessarily have the same variance as larger ranges (0-100).

Next, item variance heterogeneity was included by manipulating the potential \(\sigma\) for each individual item. For small scales, the \(\sigma\) = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items. For the medium scale of data, \(\sigma\) = 25 with a variance of 4, 8, and 16. Last, for the large scale of data, \(\sigma\) = 400 with a variance of 50, 100, and 200 for heterogeneity.

\emph{Samples}. Each population was then sampled as if a researcher was conducting a pilot study. The sample sizes started at 20 participants per item increasing in units of 5 up to 100 participants.

\emph{Cutoff Score Criterions}. The standard errors of each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as standard error functions as the main component in the formula for normal distribution confidence intervals. Standard errors were calculated at each decile of the items up to 90\% (i.e., 0\% smallest SE, 10\% \ldots, 90\% largest SE). The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criterions for cutoff scores.

\hypertarget{researcher-sample-simulation}{%
\subsection{Researcher Sample Simulation}\label{researcher-sample-simulation}}

In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items. Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 500 to determine what the new sample size suggestion would be. We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds. The standard error of each item was calculated for each suggested sample size by pilot sample size by population type.

Next, the percent of items that fall below the cutoff scores, and thus, would be considered ``well-measured'' were calculated for each decile by sample. From this data, we pinpoint the smallest suggested sample size at which 80\%, 85\%, 90\%, and 95\% of the items fall below the cutoff criterion. These values were chosen as popular measures of ``power'' in which one could determine the minimum suggested sample size (potentially 80\% of items) and the maximum suggested sample size (potentially 90\%).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{differences-in-item-variance}{%
\subsection{Differences in Item Variance}\label{differences-in-item-variance}}

We examined if this procedure is sensitive to differences in item heterogeneity, as we should expect to collect larger samples if we wish to have a large number of items reach a threshold of acceptable variance; potentially, assuring we \emph{could} average them if a researcher did not wish to use a more complex analysis such as multilevel modeling.

Figure \ref{fig:item-figure} illustrates the potential minimum sample size for 80\% of items to achieve a desired cutoff score. The black dots denote the original sample size against the suggested sample size. By comparing the facets, we can determine that our suggested procedure does capture the differences in heterogeneity. As heterogeneity increases in item variances, the proposed sample size also increases, especially at stricter cutoffs. Missing cutoff points where sample sizes proposed would be higher than 500.

\begin{figure}
\centering
\includegraphics{manuscript_draft_files/figure-latex/item-figure-1.pdf}
\caption{\label{fig:item-figure}Add a good caption here.}
\end{figure}

\hypertarget{projected-sample-size-sensitivity-to-pilot-sample-size}{%
\subsection{Projected Sample Size Sensitivity to Pilot Sample Size}\label{projected-sample-size-sensitivity-to-pilot-sample-size}}

In our second question, we examined if the suggested procedure was sensitive to the amount of information present in the pilot data. Larger pilot data is more informative, and therefore, we should expect a lower projected sample size. As shown in Figure \ref{fig:sensitive-figure} for only the low variability and small scale data, we do not find this effect. These simulations from the pilot data would nearly always suggest a larger sample size - mostly in a linear trend increasing with sample sizes. This result comes from the nature of the procedure - if we base our estimates on a SE cutoff, we will almost always need a bit more people for items to meet those goals. This result does not achieve our second goal.

\begin{figure}
\centering
\includegraphics{manuscript_draft_files/figure-latex/sensitive-figure-1.pdf}
\caption{\label{fig:sensitive-figure}Add good description here.}
\end{figure}

Therefore, we suggest using a correction factor on the simulation procedure to account for the known asymptotic nature of power (i.e., at larger sample sizes power increases level off). For this function in our simulation study, we combined a correction factor for upward biasing of effect sizes (Hedges' correction) with the formula for exponential decay calculations. The decay factor was calculated as follows:

\[ 1 - \sqrt{\frac{N_{Pilot} - min(N_{Simulation})}{N_{Pilot}}}^{log_2(N_{Pilot})}\]

\(N_{Pilot}\) indicates the sample size of the pilot data minus the minimum simulated sample size to ensure that the smallest sample sizes do not decay (i.e., the formula zeroes out). This value is raised to the power of \(log_2\) of the sample size of the pilot data, which decreases the impact of the decay to smaller increments for increasing sample sizes. This value is then multiplied by the projected sample size. As show in Figure \ref{fig:corrected-figure}, this correction factor produces the desired quality of maintaining that small pilot studies should \emph{increase} sample size, and that sample size suggestions level off as pilot study data sample size increases.

\begin{figure}
\centering
\includegraphics{manuscript_draft_files/figure-latex/corrected-figure-1.pdf}
\caption{\label{fig:corrected-figure}A corrected figure update this caption.}
\end{figure}

\hypertarget{corrections-for-individual-researchers}{%
\subsection{Corrections for Individual Researchers}\label{corrections-for-individual-researchers}}

We have portrayed that this procedure, with a correction factor, can perform as desired. However, within real scenarios, researchers will only have one pilot sample, not the various simulated samples shown above. What should the researcher do to correct their projected sample size from their own pilot data simulations?

To explore if we could recover the new projected sample size from data a researcher would have, we used linear models to create a formula for researcher correction. First, the corrected projected sample size was predicted by the original projected sample size. Next, the standard deviation of the item standard deviations was added to the equation to recreate heterogeneity estimates. The scale of the data is embedded into the standard deviation of the items (\emph{r} = 0.81), and therefore, this variable was not included separately. Last, we included the pilot sample size.

The first model using pilot sample size to predict new sample size was significant, \emph{F}(1, 5666) = 48,042.64, \emph{p} \textless{} .001, \(R^2 = .89\), 90\% CI \([0.89, 0.90]\), capturing nearly 90\% of the variance, \(b = 0.62\), 95\% CI \([0.62, 0.63]\). The second model with item standard deviation was better then the first model \emph{F}(1, 5665) = 55.21, \emph{p} \textless{} .001, \(R^2 = .89\), 90\% CI \([0.89, 0.90]\). The item standard deviation predictor was significant, \(b = 0.02\), 95\% CI \([0.01, 0.03]\), \(t(5665) = 4.54\), \(p < .001\). The addition of the original pilot sample size was also significant, \emph{F}(1, 5664) = 9,529.83, \emph{p} \textless{} .001, \(R^2 = .96\), 90\% CI \([0.96, 0.96]\).

As shown in the final model Table \ref{tab:table-correction}, the new suggested sample size is proportional to the original suggested sample size (i.e., \emph{b} \textless{} 1), which reduces the sample size suggestion. As variability increases, the suggested sample size also increases to capture differences in heterogeneity shown above; however, this predictor is not significant in the final model, and only contributes a small portion of overall variance. Last, in order to correct for large pilot data, the original pilot sample size decreases the new suggested sample size. This formula approximation captures 96\% of the variance in sample size scores and should allow a researcher to estimate based on their own data.

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:table-correction}Parameters for All Decile Cutoff Scores}

\begin{tabular}{lllll}
\toprule
Term & Estimate & $SD$ & $t$ & $p$\\
\midrule
Intercept & 39.269 & 0.437 & 89.843 & < .001\\
Projected Sample Size & 0.700 & 0.002 & 366.669 & < .001\\
Item SD & 0.003 & 0.003 & 0.952 & .341\\
Pilot Sample Size & -0.695 & 0.007 & -97.621 & < .001\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{choosing-an-appropriate-cutoff}{%
\subsection{Choosing an Appropriate Cutoff}\label{choosing-an-appropriate-cutoff}}

Last, we examine the question of an appropriate SE decile. All graphs for power, heterogeneity, scale, and correction are presented online. First, the 0\%, 10\%, and 20\% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale, and heterogeneity. If we examine the \(R^2\) values for each decile of our regression equation separately, we find that the 50\% (0.96) represents the best match to our corrected sample size suggestions. The 50\% decile, in the corrected format, appears to meet all goals: 1) increases with heterogeneity and scale of data, and 2) higher suggested values for small original samples and a leveling effect at larger pilot data. Figure \ref{fig:decile-figure} illustrates the corrected scores for simulations at the 50\% decile recommended cutoff for item standard errors.

\begin{figure}
\centering
\includegraphics{manuscript_draft_files/figure-latex/decile-figure-1.pdf}
\caption{\label{fig:decile-figure}A picture of the 50\% cutoff.}
\end{figure}

The formula for finding the corrected sample size using a 50\% decile is: \(N_{Corrected Projected} = 39.269 + 0.700 \times X_{N_{Projected}} + 0.003 \times X_{SD Items} - 0.694 \times X_{N_{Pilot}}\). The suggested sample size will be estimated from the 80\%, 85\%, 90\%, or 95\% selection at the 50\% decile as shown above. The item SD can be calculated directly from the data, and the pilot sample size is the sample size of the data from which a researcher is simulating their samples. Therefore, we will recommend the 50\% decile of the item standard errors for step 2 of our suggested simulation procedure, and to correct the projected sample sizes found in step 5 using the correction equation above. While the estimated coefficients could change given variations on our simulation parameters, the general size and pattern of coefficients was consistent, and therefore, we believe this correction equation should work for a variety of use cases.

\hypertarget{examples}{%
\section{Examples}\label{examples}}

In this section, we provide two examples of the suggested procedure. The first example includes concreteness ratings from Brysbaert et al.~{[}CITE, 2014{]}. Instructions given to participants denoted the difference between concrete (i.e., ``refers to something that exists in reality'') and abstract (i.e., ``something you cannot experience directly through your senses or actions'') terms. Participants were then asked to rate concreteness of terms using a 1 (abstract) to 5 (concrete) scale. This data represents a small scale dataset that could be used as pilot data for a study using concrete word ratings. The data is avaliable at OSF LINK {[}CITE{]}. The second dataset includes a large scale dataset with response latencies, the English Lexicon Project {[}CITE, Balota et al.{]}. The English Lexicon Project consists of lexical decision response latencies for English words. In a lexical decision task, participants simply select ``word'' for real words (e.g., \emph{dog}) and ``nonword'' for pseudowords (e.g., \emph{wug}). The trial level data is avaliable here {[}\url{https://elexicon.wustl.edu/}, CITE{]}. Critically, in each of these examples, the individual trial level data for each item is available to simulate and calculate standard errors on. Data that has been summarized could potentially be used, as long as the original standard deviations for each item were present. From the mean and standard deviation for each item, a simulated pilot dataset could be generated for estimating new sample sizes. All code to estimate sample sizes is provided on our OSF page.

\hypertarget{concreteness-ratings}{%
\subsection{Concreteness Ratings}\label{concreteness-ratings}}

The concreteness ratings data includes 63039 concepts that were rated for their concreteness. In our fictional study for this example, we selected 100 random words to show participants. In the original study, not every particpant rated every word, which created uneven sample sizes for each word. In our random sample of 100 words, the average pilot sample size was 27.96 (\emph{SD} = 1.47), and we will use 28 as our pilot sample size for this example. All ``do not know'' ratings were set as missing data. The 50\% decile for items standard error was 0.25 for our cutoff criterion.

The pilot data was then simulated, with replacement, with samples from 20 to 300 increasing in units of 5. On each sample, the percent of items below the cutoff score were calculated. After applying our correction equation, we find that a sample size of 44 would allow for at least 80\% of items to meet the cutoff criterion. The sample sizes for 85\% (48), 90\% (48), and 95\% (51) are also options for sample size suggestions. Finally, we calculated the potential amount of data retention given that participants could indicate they did not know a word (\(M_{correct}\) = 0.82, \emph{SD} = 0.24). In order to account for this facet, the potential sample sizes were multiplied by 1/0.82, which results in a suggested sample of 54, 59, and 63. Therefore, we could designate our minimum sample per item as 54, stopping rule of 0.25, and maximum sample size of 63.

\hypertarget{response-latencies}{%
\subsection{Response Latencies}\label{response-latencies}}

The ELP response latency data includes 80962 word-forms, 40481 that are listed as non-words, and 40481 real words. For our example study, we will randomly select 500 real words and 500 non-words to show participants. The average pilot sample size for this random sample was 32.68 (\emph{SD} = 0.64), and \emph{n} = 33 will be our pilot size for this example. Again, participants are expected to make mistakes, and we calculated percent correct as 0.86, which was roughly even in the two stimulus categories: \(M_{word}\) = 0.84 and \(M_{non-word}\) = 0.87. The 50\% decile for items standard error was 61.32 for our cutoff criterion. We additionally checked to ensure that the two stimulus types did not have very different cutoff criterions: 50\% decile \(SE_{words}\) = 56.94, 50\% decile \(SE_{nonwords}\) = 65.61. In this scenario, we could chose to go with the lower SE to be more conservative (i.e., higher projected sample size). Given the values were close for large scale data, we used the 50\% decile of all stimuli taken together.

The pilot response latency data was then simulated in the same way as described above. After calculating the percent below our cutoff score, we applied the correction to the projected sample sizes. A sample size of 31 would equate to 80\% of the items reaching our cutoff, along with 85\% (34), 90\% (34), and 95\% (38). Again, we adjusted for data loss given that participants are expected to incorrectly answer items, resulting in a suggested sample of 36, 40, and 44. One other possible consideration for this study is potential fatigue in showing participants 1000 target items. Therefore, we could designate in our research design that each participant will only receive 500 of the target items. We would need to double our sample sizes to account for splitting of the items across multiple sets of participants. Our minimum sample size for the entire study could be 72, stopping rule of 61.32, and maximum sample size of 88. This study would benefit from an adaptive design, where smaller sets items are randomly sampled for participants until they reach the minimum sample size or the cutoff criteria. At this point, items are probablistically sampled (e.g., higher selection probability for items that have not reached a minimum or stopping rule) until all items have reached criteria.

\hypertarget{additional-materials}{%
\subsection{Additional Materials}\label{additional-materials}}

While the examples in this manuscript are traditionally cognitive linguistics focused, any research using repeated items can benefit from newer sampling techniques. Therefore, we provide XX example vignettes on our OSF page/GitHub site for this manuscript across a range of examples of data types provided by the authors of this manuscript. Examples include \ldots{} ADD HERE AFTER DONE WITH VIGNETTES.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

To be added once people say this idea doesn't suck.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\end{CSLReferences}

\endgroup

\end{document}

### 2022-06-22 simulation_study/simulate_cutoff.Rmd
updating manuscript

--- 2022-06-17 simulate_cutoff.Rmd
@@
 For small scales, the $\sigma$ = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items.
-For the medium scale of data, $\sigma$ = 20 with a variance of 4, 8, and 12.
-Last, for the large scale of data, $\sigma$ = 350 with a variance of 100, 150, and 200 for heterogeneity.
+For the medium scale of data, $\sigma$ = 25 with a variance of 4, 8, and 16.
+Last, for the large scale of data, $\sigma$ = 400 with a variance of 50, 100, and 200 for heterogeneity.
 
@@
 The first model using original sample size to predict new sample size was significant, `r user_print$full_result$modelfit$r2`, capturing nearly 90% of the variance, `r user_print$estimate$sample_size`.
-The second model with item standard deviation was better then the first model *F*(`r change_table$df[2]`, `r change_table$res.df[2]`) = `r change_table$statistic[2]`, *p* < .001, `r user_print2$estimate$modelfit$r2`.
+The second model with item standard deviation was better then the first model *F*(`r change_table$df[2]`, `r change_table$res.df[2]`) = `r round(change_table$statistic[2], digits = 2)`, *p* < .001, `r user_print2$estimate$modelfit$r2`.
 The item standard deviation predictor was significant, `r user_print2$full_result$sd_item`.
-The addition of the original pilot sample size was also significant, *F*(`r change_table$df[3]`, `r change_table$res.df[3]`) = `r change_table$statistic[3]`, *p* < .001, `r user_print3$estimate$modelfit$r2`.
+The addition of the original pilot sample size was also significant, *F*(`r change_table$df[3]`, `r change_table$res.df[3]`) = `r round(change_table$statistic[3], digits = 2)`, *p* < .001, `r user_print3$estimate$modelfit$r2`.
 
@@
 
-The formula for finding the corrected sample size using a 50% decile is: $Final Sample = 0.637 \times X_{Suggested Sample Size} + 0.002 \times X_{SD Items} + -0.546 \times X_{Pilot Sample Size}$.
+The formula for finding the corrected sample size using a 50% decile is: $Final Sample = 39.269 + 0.700 \times X_{Suggested Sample Size} + 0.003 \times X_{SD Items} + -0.695 \times X_{Pilot Sample Size}$.
+The suggested sample size will be estimated from the 80%, 85%, 90%, or 95% selection at the 50% decile as shown above.
+The item SD can be calculated directly from the data, and the pilot sample size is the sample size of the data from which a researcher is simulating their samples.
 

### 2022-07-21 manuscript_draft.Rmd
adding information about residuals

@@
 library(rio)
+library(nlme)
+library(MuMIn)
 ```
@@
 
+```{r test-assume, include = F, echo = F}
+resid_model <- scale(residuals(user_model3))
+fit_model <- scale(fitted.values(user_model3))
+hist(resid_model)
+qqnorm(resid_model); abline(0,1)
+plot(fit_model, resid_model); abline(v = 0); abline(h = 0)
+
+# user_model_lme <- lme(new_sample ~ sample_size + sd_item + original_n, 
+#                       data = summary_long, 
+#                       random = list(~1|scale2, ~scale2|power, ~scale2|name))
+# r.squaredGLMM(user_model_lme)
+# resid_model <- scale(residuals(user_model_lme))
+# fit_model <- scale(fitted.values(user_model_lme))
+# hist(resid_model)
+# qqnorm(resid_model); abline(0,1)
+# plot(fit_model, resid_model); abline(v = 0); abline(h = 0)
+```
+
 The first model using pilot sample size to predict new sample size was significant, *F*(`r user_sum$fstatistic[2]`, `r user_sum$fstatistic[3]`) = `r round(user_sum$fstatistic[1], digits = 2)`, *p* < .001, `r user_print$estimate$modelfit$r2`, capturing nearly 90% of the variance, `r user_print$estimate$sample_size`.
@@
 }
+```
+
+```{r test-decile, include = F, echo = F}
+resid_model <- scale(residuals(by_cutoff[["Decile 50"]]))
+fit_model <- scale(fitted.values(by_cutoff[["Decile 50"]]))
+hist(resid_model)
+qqnorm(resid_model); abline(0,1)
+plot(fit_model, resid_model); abline(v = 0); abline(h = 0)
+
+# by scale, power, source
+separate_lms <- list()
+R2_lms <- list()
+
+for (scale_loop in unique(summary_long$scale)){
+  for (power_loop in unique(summary_long$power)){
+    for (source_loop in unique(summary_long$source)){
+      temp <- summary_long %>% 
+        filter(scale == scale_loop) %>% 
+        filter(power == power_loop) %>% 
+        filter(source == source_loop) %>% 
+        filter(name == "Decile 50")
+
+      separate_lms[[paste(scale_loop, power_loop, source_loop, sep = "_")]] <- lm(new_sample ~ sample_size + sd_item + original_n, data = temp)
+      R2_lms[[paste(scale_loop, power_loop, source_loop, sep = "_")]] <- summary(separate_lms[[paste(scale_loop, power_loop, source_loop, sep = "_")]])$r.squared
+    }
+  }
+}
+
+# resid_model <- scale(residuals(separate_lms[[10]]))
+# fit_model <- scale(fitted.values(separate_lms[[10]]))
+# hist(resid_model)
+# qqnorm(resid_model); abline(0,1)
+# plot(fit_model, resid_model); abline(v = 0); abline(h = 0)
 ```

### 2022-08-12 manuscript_draft.Rmd
updating with simulations outside the manuscript

@@
 
-```{r sim_pop, include = F, echo = F}
-# small potential variability overall, sort of 1-7ish scale
-mu1 <- rnorm(30, 4, .25)
-sigma1.1 <- rnorm(30, 2, .2)
-sigma2.1 <- rnorm(30, 2, .4)
-sigma3.1 <- rnorm(30, 2, .8)
-
-# medium potential variability 0 to 100 scale
-mu2 <- rnorm(30, 50, 10)
-sigma1.2 <- rnorm(30, 25, 4)
-sigma2.2 <- rnorm(30, 25, 8)
-sigma3.2 <- rnorm(30, 25, 16)
-
-while(sum(sigma3.2 < 0) > 0){
-  sigma3.2 <- rnorm(30, 25, 16)
-}
-
-# large potential variability in the 1000s scale
-mu3 <- rnorm(30, 1000, 150)
-sigma1.3 <- rnorm(30, 400, 50)
-sigma2.3 <- rnorm(30, 400, 100)
-sigma3.3 <- rnorm(30, 400, 200)
-
-while(sum(sigma3.3 < 0) > 0){
-  sigma3.3 <- rnorm(30, 400, 200)
-}
-
-population1 <- data.frame(
-  item = rep(1:30, 1000*3),
-  scale = rep(1:3, each = 1000*30),
-  score = c(rnorm(1000*30, mean = mu1, sd = sigma1.1),
-            rnorm(1000*30, mean = mu2, sd = sigma1.2),
-            rnorm(1000*30, mean = mu3, sd = sigma1.3))
-  )
-
-population2 <- data.frame(
-  item = rep(1:30, 1000*3),
-  scale = rep(1:3, each = 1000*30),
-  score = c(rnorm(1000*30, mean = mu1, sd = sigma2.1),
-            rnorm(1000*30, mean = mu2, sd = sigma2.2),
-            rnorm(1000*30, mean = mu3, sd = sigma2.3))
-  )
-
-population3 <- data.frame(
-  item = rep(1:30, 1000*3),
-  scale = rep(1:3, each = 1000*30),
-  score = c(rnorm(1000*30, mean = mu1, sd = sigma3.1),
-            rnorm(1000*30, mean = mu2, sd = sigma3.2),
-            rnorm(1000*30, mean = mu3, sd = sigma3.3))
-  )
-```
-
 *Samples*. Each population was then sampled as if a researcher was conducting a pilot study. The sample sizes started at 20 participants per item increasing in units of 5 up to 100 participants.
-
-```{r sim_sample, include = F, echo = F}
-# create pilot samples from 20 to 100
-samples1 <- samples2 <- samples3 <- list() 
-sizes <- c(20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75, 80, 85, 90, 95, 100)
-for (i in 1:length(sizes)){
-  samples1[[i]] <- population1 %>% 
-    group_by(item, scale) %>% 
-    slice_sample(n = sizes[i])
-
-  samples2[[i]] <- population2 %>% 
-    group_by(item, scale) %>% 
-    slice_sample(n = sizes[i])
-
-  samples3[[i]] <- population3 %>% 
-    group_by(item, scale) %>% 
-    slice_sample(n = sizes[i])
-}
-```
 
@@
 
-```{r calc_SE, include = F, echo = F}
-# calculate the SEs and the cutoff scores 
-SES1 <- SES2 <- SES3 <- list()
-cutoffs1 <- cutoffs2 <- cutoffs3 <- list()
-sd_items1 <- sd_items2 <- sd_items3 <- list()
-for (i in 1:length(samples1)){
-  sd_items1[[i]] <- samples1[[i]] %>% group_by(item, scale) %>% 
-    summarize(sd = sd(score), .groups = "keep") %>% 
-    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
-  sd_items2[[i]] <- samples2[[i]] %>% group_by(item, scale) %>% 
-    summarize(sd = sd(score), .groups = "keep") %>% 
-    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
-  sd_items3[[i]] <- samples3[[i]] %>% group_by(item, scale) %>% 
-    summarize(sd = sd(score), .groups = "keep") %>% 
-    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
-
-  SES1[[i]] <- tapply(samples1[[i]]$score,
-                     list(samples1[[i]]$item,
-                          samples1[[i]]$scale),
-                     function (x){ sd(x)/sqrt(length(x))})
-  SES2[[i]] <- tapply(samples2[[i]]$score,
-                   list(samples2[[i]]$item,
-                        samples2[[i]]$scale),
-                   function (x){ sd(x)/sqrt(length(x))})
-  SES3[[i]] <- tapply(samples3[[i]]$score,
-                 list(samples2[[i]]$item,
-                      samples2[[i]]$scale),
-                 function (x){ sd(x)/sqrt(length(x))})
-
-  cutoffs1[[i]] <- apply(as.data.frame(SES1[[i]]), 2, 
-                         quantile, 
-                         probs = seq(0, .9, by = .1))
-  cutoffs2[[i]] <- apply(as.data.frame(SES2[[i]]), 2, 
-                         quantile, 
-                         probs = seq(0, .9, by = .1))
-  cutoffs3[[i]] <- apply(as.data.frame(SES3[[i]]), 2, 
-                         quantile, 
-                         probs = seq(0, .9, by = .1))
-}
-```
-
 ## Researcher Sample Simulation
@@
 In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items.
-Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 500 to determine what the new sample size suggestion would be.
-We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds.
+Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 2000 to determine what the new sample size suggestion would be.
+We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds; however, the sample size simulations were estimated over this amount to determine the pattern of suggested sample sizes (i.e., the function between original sample size and proposed sample size).
 The standard error of each item was calculated for each suggested sample size by pilot sample size by population type.
-
-```{r sim_sim, include = F, echo = F}
-# sequence of sample sizes to try
-samplesize_values <- seq(20, 500, 5)
-
-# place to store everything
-sampled_values1 <- sampled_values2 <- sampled_values3 <- list()
-
-# loop over the samples
-for (i in 1:length(samples1)){
-
-  # create a blank table for us to save the values in 
-  sim_table1 <- matrix(NA, 
-                      nrow = length(samplesize_values), 
-                      ncol = 30*3)
-
-  # make it a data frame
-  sim_table1 <- sim_table2 <- sim_table3 <- as.data.frame(sim_table1)
-
-  # add a place for sample size values 
-  sim_table1$sample_size <- sim_table2$sample_size <- sim_table3$sample_size <- NA
-
-  # loop over pilot sample sizes
-  for (q in 1:length(samplesize_values)){
-
-    # temp dataframe that samples and summarizes
-    temp <- samples1[[i]] %>% 
-      group_by(item, scale) %>% 
-      slice_sample(n = samplesize_values[q], replace = T) %>% 
-      summarize(se = sd(score)/sqrt(length(score)),
-                .groups = "keep")
-
-    sim_table1[q, 1:90] <- temp$se
-    sim_table1[q, 91] <- samplesize_values[q]
-
-    temp <- samples2[[i]] %>% 
-      group_by(item, scale) %>% 
-      slice_sample(n = samplesize_values[q], replace = T) %>% 
-      summarize(se = sd(score)/sqrt(length(score)),
-                .groups = "keep")
-
-    sim_table2[q, 1:90] <- temp$se
-    sim_table2[q, 91] <- samplesize_values[q]
-
-    temp <- samples3[[i]] %>% 
-      group_by(item, scale) %>% 
-      slice_sample(n = samplesize_values[q], replace = T) %>% 
-      summarize(se = sd(score)/sqrt(length(score)),
-                .groups = "keep")
-
-    sim_table3[q, 1:90] <- temp$se
-    sim_table3[q, 91] <- samplesize_values[q]
-
-    } # end pilot sample loop
-
-  sampled_values1[[i]] <- sim_table1
-  sampled_values2[[i]] <- sim_table2
-  sampled_values3[[i]] <- sim_table3
-
-} # end all sample loop 
-```
 
@@
 
-```{r calc_percent, include = F, echo = F}
-summary_list1 <- summary_list2 <- summary_list3 <- list()
-for (i in 1:length(sampled_values1)){
-
-  # summary list 1 ----
-  summary_list1[[i]] <- sampled_values1[[i]] %>% 
-    pivot_longer(cols = -c(sample_size)) %>% 
-    rename(item = name, se = value) %>% 
-    mutate(scale = rep(1:3, 30*length(samplesize_values))) %>% 
-    mutate(item = rep(rep(1:30, each = 3), length(samplesize_values)))
-
-  # cut offs for 1
-  temp1.1 <- summary_list1[[i]] %>% 
-    filter(scale == "1") %>% 
-    group_by(sample_size, scale) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs1[[i]][1, 1])/30,
-           Percent_Below10 = sum(se <= cutoffs1[[i]][2, 1])/30,
-           Percent_Below20 = sum(se <= cutoffs1[[i]][3, 1])/30,
-           Percent_Below30 = sum(se <= cutoffs1[[i]][4, 1])/30,
-           Percent_Below40 = sum(se <= cutoffs1[[i]][5, 1])/30,
-           Percent_Below50 = sum(se <= cutoffs1[[i]][6, 1])/30, 
-           Percent_Below60 = sum(se <= cutoffs1[[i]][7, 1])/30, 
-           Percent_Below70 = sum(se <= cutoffs1[[i]][8, 1])/30, 
-           Percent_Below80 = sum(se <= cutoffs1[[i]][9, 1])/30, 
-           Percent_Below90 = sum(se <= cutoffs1[[i]][10, 1])/30, 
-           .groups = "keep") %>% 
-    mutate(original_n = sizes[i], 
-           source = "low")
-
-  temp1.2 <- summary_list1[[i]] %>% 
-    filter(scale == "2") %>% 
-    group_by(sample_size, scale) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs1[[i]][1, 2])/30,
-           Percent_Below10 = sum(se <= cutoffs1[[i]][2, 2])/30,
-           Percent_Below20 = sum(se <= cutoffs1[[i]][3, 2])/30,
-           Percent_Below30 = sum(se <= cutoffs1[[i]][4, 2])/30,
-           Percent_Below40 = sum(se <= cutoffs1[[i]][5, 2])/30,
-           Percent_Below50 = sum(se <= cutoffs1[[i]][6, 2])/30, 
-           Percent_Below60 = sum(se <= cutoffs1[[i]][7, 2])/30, 
-           Percent_Below70 = sum(se <= cutoffs1[[i]][8, 2])/30, 
-           Percent_Below80 = sum(se <= cutoffs1[[i]][9, 2])/30, 
-           Percent_Below90 = sum(se <= cutoffs1[[i]][10, 2])/30, 
-           .groups = "keep") %>% 
-    mutate(original_n = sizes[i],
-           source = "low")
-
-  temp1.3 <- summary_list1[[i]] %>% 
-    filter(scale == "3") %>% 
-    group_by(sample_size, scale) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs1[[i]][1, 3])/30,
-           Percent_Below10 = sum(se <= cutoffs1[[i]][2, 3])/30,
-           Percent_Below20 = sum(se <= cutoffs1[[i]][3, 3])/30,
-           Percent_Below30 = sum(se <= cutoffs1[[i]][4, 3])/30,
-           Percent_Below40 = sum(se <= cutoffs1[[i]][5, 3])/30,
-           Percent_Below50 = sum(se <= cutoffs1[[i]][6, 3])/30, 
-           Percent_Below60 = sum(se <= cutoffs1[[i]][7, 3])/30, 
-           Percent_Below70 = sum(se <= cutoffs1[[i]][8, 3])/30, 
-           Percent_Below80 = sum(se <= cutoffs1[[i]][9, 3])/30, 
-           Percent_Below90 = sum(se <= cutoffs1[[i]][10, 3])/30, 
-           .groups = "keep") %>% 
-    mutate(original_n = sizes[i], 
-           source = "low")
-
-  #rejoin 
-  summary_list1[[i]] <- bind_rows(temp1.1, temp1.2, temp1.3)
-
-  # summary list 2 ----
-  summary_list2[[i]] <- sampled_values2[[i]] %>% 
-    pivot_longer(cols = -c(sample_size)) %>% 
-    rename(item = name, se = value) %>% 
-    mutate(scale = rep(1:3, 30*length(samplesize_values))) %>% 
-    mutate(item = rep(rep(1:30, each = 3), length(samplesize_values)))
-
-  # cut offs for 2
-  temp2.1 <- summary_list2[[i]] %>% 
-    filter(scale == "1") %>% 
-    group_by(sample_size, scale) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs2[[i]][1, 1])/30,
-           Percent_Below10 = sum(se <= cutoffs2[[i]][2, 1])/30,
-           Percent_Below20 = sum(se <= cutoffs2[[i]][3, 1])/30,
-           Percent_Below30 = sum(se <= cutoffs2[[i]][4, 1])/30,
-           Percent_Below40 = sum(se <= cutoffs2[[i]][5, 1])/30,
-           Percent_Below50 = sum(se <= cutoffs2[[i]][6, 1])/30, 
-           Percent_Below60 = sum(se <= cutoffs2[[i]][7, 1])/30, 
-           Percent_Below70 = sum(se <= cutoffs2[[i]][8, 1])/30, 
-           Percent_Below80 = sum(se <= cutoffs2[[i]][9, 1])/30, 
-           Percent_Below90 = sum(se <= cutoffs2[[i]][10, 1])/30, 
-           .groups = "keep") %>% 
-    mutate(original_n = sizes[i],
-           source = "med")
-
-  temp2.2 <- summary_list2[[i]] %>% 
-    filter(scale == "2") %>% 
-    group_by(sample_size, scale) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs2[[i]][1, 2])/30,
-           Percent_Below10 = sum(se <= cutoffs2[[i]][2, 2])/30,
-           Percent_Below20 = sum(se <= cutoffs2[[i]][3, 2])/30,
-           Percent_Below30 = sum(se <= cutoffs2[[i]][4, 2])/30,
-           Percent_Below40 = sum(se <= cutoffs2[[i]][5, 2])/30,
-           Percent_Below50 = sum(se <= cutoffs2[[i]][6, 2])/30, 
-           Percent_Below60 = sum(se <= cutoffs2[[i]][7, 2])/30, 
-           Percent_Below70 = sum(se <= cutoffs2[[i]][8, 2])/30, 
-           Percent_Below80 = sum(se <= cutoffs2[[i]][9, 2])/30, 
-           Percent_Below90 = sum(se <= cutoffs2[[i]][10, 2])/30, 
-           .groups = "keep") %>% 
-    mutate(original_n = sizes[i], 
-           source = "med")
-
-  temp2.3 <- summary_list2[[i]] %>% 
-    filter(scale == "3") %>% 
-    group_by(sample_size, scale) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs2[[i]][1, 3])/30,
-           Percent_Below10 = sum(se <= cutoffs2[[i]][2, 3])/30,
-           Percent_Below20 = sum(se <= cutoffs2[[i]][3, 3])/30,
-           Percent_Below30 = sum(se <= cutoffs2[[i]][4, 3])/30,
-           Percent_Below40 = sum(se <= cutoffs2[[i]][5, 3])/30,
-           Percent_Below50 = sum(se <= cutoffs2[[i]][6, 3])/30, 
-           Percent_Below60 = sum(se <= cutoffs2[[i]][7, 3])/30, 
-           Percent_Below70 = sum(se <= cutoffs2[[i]][8, 3])/30, 
-           Percent_Below80 = sum(se <= cutoffs2[[i]][9, 3])/30, 
-           Percent_Below90 = sum(se <= cutoffs2[[i]][10, 3])/30, 
-           .groups = "keep") %>% 
-    mutate(original_n = sizes[i],
-           source = "med")
-
-  #rejoin 
-  summary_list2[[i]] <- bind_rows(temp2.1, temp2.2, temp2.3)
-
-  # summary list 3 ----
-  summary_list3[[i]] <- sampled_values3[[i]] %>% 
-    pivot_longer(cols = -c(sample_size)) %>% 
-    rename(item = name, se = value) %>% 
-    mutate(scale = rep(1:3, 30*length(samplesize_values))) %>% 
-    mutate(item = rep(rep(1:30, each = 3), length(samplesize_values)))
-
-  # cut offs for 3 
-  temp3.1 <- summary_list3[[i]] %>% 
-    filter(scale == "1") %>% 
-    group_by(sample_size, scale) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs3[[i]][1, 1])/30,
-           Percent_Below10 = sum(se <= cutoffs3[[i]][2, 1])/30,
-           Percent_Below20 = sum(se <= cutoffs3[[i]][3, 1])/30,
-           Percent_Below30 = sum(se <= cutoffs3[[i]][4, 1])/30,
-           Percent_Below40 = sum(se <= cutoffs3[[i]][5, 1])/30,
-           Percent_Below50 = sum(se <= cutoffs3[[i]][6, 1])/30, 
-           Percent_Below60 = sum(se <= cutoffs3[[i]][7, 1])/30, 
-           Percent_Below70 = sum(se <= cutoffs3[[i]][8, 1])/30, 
-           Percent_Below80 = sum(se <= cutoffs3[[i]][9, 1])/30, 
-           Percent_Below90 = sum(se <= cutoffs3[[i]][10, 1])/30, 
-           .groups = "keep") %>% 
-    mutate(original_n = sizes[i],
-           source = "high")
-
-  temp3.2 <- summary_list3[[i]] %>% 
-    filter(scale == "2") %>% 
-    group_by(sample_size, scale) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs3[[i]][1, 2])/30,
-           Percent_Below10 = sum(se <= cutoffs3[[i]][2, 2])/30,
-           Percent_Below20 = sum(se <= cutoffs3[[i]][3, 2])/30,
-           Percent_Below30 = sum(se <= cutoffs3[[i]][4, 2])/30,
-           Percent_Below40 = sum(se <= cutoffs3[[i]][5, 2])/30,
-           Percent_Below50 = sum(se <= cutoffs3[[i]][6, 2])/30, 
-           Percent_Below60 = sum(se <= cutoffs3[[i]][7, 2])/30, 
-           Percent_Below70 = sum(se <= cutoffs3[[i]][8, 2])/30, 
-           Percent_Below80 = sum(se <= cutoffs3[[i]][9, 2])/30, 
-           Percent_Below90 = sum(se <= cutoffs3[[i]][10, 2])/30, 
-           .groups = "keep") %>% 
-    mutate(original_n = sizes[i],
-           source = "high")
-
-  temp3.3 <- summary_list3[[i]] %>% 
-    filter(scale == "3") %>% 
-    group_by(sample_size, scale) %>% 
-    summarize(Percent_Below0 = sum(se <= cutoffs3[[i]][1, 3])/30,
-           Percent_Below10 = sum(se <= cutoffs3[[i]][2, 3])/30,
-           Percent_Below20 = sum(se <= cutoffs3[[i]][3, 3])/30,
-           Percent_Below30 = sum(se <= cutoffs3[[i]][4, 3])/30,
-           Percent_Below40 = sum(se <= cutoffs3[[i]][5, 3])/30,
-           Percent_Below50 = sum(se <= cutoffs3[[i]][6, 3])/30, 
-           Percent_Below60 = sum(se <= cutoffs3[[i]][7, 3])/30, 
-           Percent_Below70 = sum(se <= cutoffs3[[i]][8, 3])/30, 
-           Percent_Below80 = sum(se <= cutoffs3[[i]][9, 3])/30, 
-           Percent_Below90 = sum(se <= cutoffs3[[i]][10, 3])/30, 
-           .groups = "keep") %>% 
-    mutate(original_n = sizes[i],
-           source = "high")
-
-  #rejoin 
-  summary_list3[[i]] <- bind_rows(temp3.1, temp3.2, temp3.3)
-
-  }
-
-summary_DF <- bind_rows(summary_list1, 
-                        summary_list2, 
-                        summary_list3)
-```
-
-```{r calc_percent_combine, include = F, echo = F}
-summary_long_80 <- summary_DF %>% 
-  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
-  filter(value >= .80) %>% 
-  arrange(sample_size, original_n, source, scale, name) %>% 
-  group_by(original_n, name, source, scale) %>% 
-  slice_head(n = 1) %>% 
-  mutate(power = 80)
-
-summary_long_85 <- summary_DF %>% 
-  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
-  filter(value >= .85) %>% 
-  arrange(sample_size, original_n, source, scale, name) %>% 
-  group_by(original_n, name, source, scale) %>% 
-  slice_head(n = 1) %>% 
-  mutate(power = 85)
-
-summary_long_90 <- summary_DF %>% 
-  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
-  filter(value >= .90) %>% 
-  arrange(sample_size, original_n, source, scale, name) %>% 
-  group_by(original_n, name, source, scale) %>% 
-  slice_head(n = 1) %>% 
-  mutate(power = 90)
-
-summary_long_95 <- summary_DF %>% 
-  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
-  filter(value >= .95) %>% 
-  arrange(sample_size, original_n, source, scale, name) %>% 
-  group_by(original_n, name, source, scale) %>% 
-  slice_head(n = 1) %>% 
-  mutate(power = 95)
-
-summary_long <- rbind(summary_long_80, 
-                      summary_long_85,
-                      summary_long_90,
-                      summary_long_95)
-
-summary_long$source <- factor(summary_long$source, 
-                              levels = c("low", "med", "high"),
-                              labels = c("Low Variance", "Medium Variance", "High Variance"))
-
-summary_long$scale2 <- factor(summary_long$scale, 
-                              levels = c(1:3),
-                              labels = c("Small Scale", "Medium Scale", "Large Scale"))
-
-sd_items <- bind_rows(sd_items1, sd_items2, sd_items3)
-sd_items$original_n <- rep(rep(sizes, each = 3), 3)
-sd_items$source <- rep(c("Low Variance", "Medium Variance", "High Variance"), each = 3*length(sizes))
-summary_long <- summary_long %>% 
-  full_join(sd_items, 
-            by = c("original_n" = "original_n", "scale" = "scale", 
-                   "source" = "source"))
-
-summary_long$name <- gsub("Percent_Below", "Decile ", summary_long$name)
-```
+In this procedure, we identified three points of simulation: the population, the samples from those populations, and the researcher sample simulation.
+In order to minimize any potentially random quirks, we ran each of these samples 100 times for a total of 1000000 simulations.
+The average of these simulations is presented in the results.
 
 # Results
+
+```{r import-data}
+
+```
 
@@
 
-```{r R2_cutoff, include = F, echo = F}
+```{r R2-cutoff, include = F, echo = F}
 by_cutoff <- list()
@@
 
-```{r import_data, include = F, echo = F}
+```{r import-data-example, include = F, echo = F}
 conDF <- import("./data/concreteness_trial.rda")
@@
 
-```{r con_estimate, echo = F, include = F}
+```{r con-estimate, echo = F, include = F}
 # sample the data
@@
 
-```{r con_sample, include = F, echo = F}
+```{r con-sample, include = F, echo = F}
 # sequence of sample sizes to try
@@
 
-```{r elp_estimate, echo = F, include = F}
+```{r elp-estimate, echo = F, include = F}
 # sample the data
@@
 
-```{r elp_sample, include = F, echo = F}
+```{r elp-sample, include = F, echo = F}
 # sequence of sample sizes to try

### 2022-08-12 simulation_study/simulate_cutoff.Rmd
updating with simulations outside the manuscript

--- 2022-06-14 simulate_cutoff.Rmd
@@
 ---
-title: "Simulate Cut Off"
+title: "Methods Power Study"
 author: "Erin M. Buchanan"
@@
 library(papaja)
+library(rio)
 ```
 
-# Introduction
-
-More real introduction on power, current trends to AIPE, and why items are key.
-
-This document outlines using the suggested procedure to simulate the number of participants necessary to accurately measure items.
-There are two key issues these ideas should address that we know about power:
-
-1) we should see differences in projected sample sizes based on the variability in the variance for those items (i.e., heterogeneity should increase projected sample size)
-2) we should see projected sample sizes that "level off" when pilot data increases.
-As with regular power estimates, studies can be "overpowered" to detect an effect, and this same idea should be present.
-For example, if one has a 500 person pilot study, our simulations should suggest a point at which items are likely measured well, which may have happened well before 500.
-
-# Explain the Procedure
-
-Notes forthcoming.
+This document includes the informal write up for the method for the manuscript.
+We exported the simulations to this document in order to run the simulations separately from the manuscript for speed reasons (i.e., it's really slow to run the simulations along with the manuscript markdown).
+At the end of this document, a `simulated_summary_data.csv` file is created that is imported into the manuscript for plotting and other analyses.
+Please see the manuscript for formal write up.
 
@@
 *Population*.
-The data was simulated using the `rnorm` function assuming a normal distribution for 30 scale type items.
-Each population was simulated with 1000 data points.
-No items were rounded for this simulation.
 
-First, the scale of the data was manipulated by creating three sets of scales.
-The first scale was mimicked after small rating scales (i.e., 1-7 type style) using a $\mu$ = 4 with a $\sigma$ = .25 around the mean to create item mean variability.
-The second scale included a larger potential distribution of scores with a $\mu$ = 50 ($\sigma$ = 10) imitating a 0-100 scale.
-Last, the final scale included a $\mu$ = 1000 ($\sigma$ = 150) simulating a study that may include response latency data in the milliseconds.
-While there are many potential scales, these three represent a large number of potential variables in the social sciences.
-As we are suggesting variances as a key factor for estimating sample sizes, the scale of the data is influential on the amount of *potential* variance.
-Smaller ranges of data (1-7) cannot necessarily have the same variance as larger ranges (0-100).
+- Simulate the data with `rnorm` assuming a normal distribution for 30 scale items
+- Population is created with 1000 data points 
+- No rounding 
+- We will run this part 100 times, so create a function that does this for you.
 
-Next, item variance heterogeneity was included by manipulating the potential $\sigma$ for each individual item.
-For small scales, the $\sigma$ = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items.
-For the medium scale of data, $\sigma$ = 20 with a variance of 4, 8, and 12.
-Last, for the large scale of data, $\sigma$ = 350 with a variance of 100, 150, and 200 for heterogeneity.
+```{r sim_pop, include = T, echo = T}
+simulate_population <- function (x = 1){
 
-```{r echo = F}
   # small potential variability overall, sort of 1-7ish scale
@@
   mu2 <- rnorm(30, 50, 10)
-sigma1.2 <- rnorm(30, 20, 4)
-sigma2.2 <- rnorm(30, 20, 8)
-sigma3.2 <- rnorm(30, 20, 12)
+  sigma1.2 <- rnorm(30, 25, 4)
+  sigma2.2 <- rnorm(30, 25, 8)
+  sigma3.2 <- rnorm(30, 25, 16)
+
+  while(sum(sigma3.2 < 0) > 0){
+    sigma3.2 <- rnorm(30, 25, 16)
+  }
 
   # large potential variability in the 1000s scale
-
   mu3 <- rnorm(30, 1000, 150)
-sigma1.3 <- rnorm(30, 350, 100)
-sigma2.3 <- rnorm(30, 350, 150)
-sigma3.3 <- rnorm(30, 350, 200)
+  sigma1.3 <- rnorm(30, 400, 50)
+  sigma2.3 <- rnorm(30, 400, 100)
+  sigma3.3 <- rnorm(30, 400, 200)
 
   while(sum(sigma3.3 < 0) > 0){
-  sigma3.3 <- rnorm(30, 350, 200)
+    sigma3.3 <- rnorm(30, 400, 200)
   }
@@
 
-#evidence that they are simulated correctly
-tapply(population1$score, list(population1$item, 
-                               population1$scale), mean)
-tapply(population1$score, list(population1$item,
-                               population1$scale), sd)
-
-tapply(population2$score, list(population2$item, 
-                               population2$scale), mean)
-tapply(population2$score, list(population2$item,
-                               population2$scale), sd)
-
-tapply(population3$score, list(population3$item, 
-                               population3$scale), mean)
-tapply(population3$score, list(population3$item,
-                               population3$scale), sd) 
+  # return populations 
+  return(list(population1 = population1, 
+              population2 = population2, 
+              population3 = population3))
+}
 ```
 
-*Samples*. Each population was then sampled as if a researcher was conducting a pilot study. The sample sizes started at 20 participants per item increasing in units of 5 up to 100 participants.
+- Within that function, create the scale size variable: 
+  - Small scale data 1-7, mean 4, sd of mean .25
+  - Medium scale data 0-100, mean 50, sd of mean 10
+  - Large scale data no range, mean 1000, sd of mean 150
+  - These choices are meant to mimic popular scale choices 
+- Additionally, create the scale heterogeneity variable:
+  - Small scale heterogeneity, sd 2, sd of sd .2 .4 .8
+  - Medium scale heterogeneity, sd 25, sd of sd 4, 8, 16
+  - Large scale heterogeneity, sd 400, sd of sd 50, 100, 200
 
-```{r}
-# create pilot samples from 20 to 100
+*Samples*.
+
+- Create a function that does the samples that a researcher might do
+- Includes a start and stop of the number of participants you might consider 
+- Includes an increasing value between samples
+- We will also run this part 100 times
+
+```{r sim_sample, include = T, echo = T}
+simulate_samples <- function(start = 20, stop = 100, increase = 5,
+                             population1, population2, population3){
+
+  # save those samples for small medium large 
   samples1 <- samples2 <- samples3 <- list()
-sizes <- c(20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75, 80, 85, 90, 95, 100)
+
+  # create the list of sizes 
+  sizes <- seq(from = start, to = stop, by = increase)
+
+  # loop over sizes and create those samples 
   for (i in 1:length(sizes)){
@@
   }
+
+  # return your values
+  return(list(samples1 = samples1, 
+              samples2 = samples2, 
+              samples3 = samples3,
+              sizes = sizes))
+
+}
 ```
 
-*AIPE Criterions*.
-The standard errors of each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as standard error functions as the main component in the formula for normal distribution confidence intervals.
-Standard errors were calculated at each decile of the items up to 90% (0% smallest SE, 10% ..., 90% largest SE).
-The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet AIPE goals, while the higher deciles would represent less strict criterions for AIPE goals.
+*Cutoff Score Criterions*.
 
-```{r}
+- Calculate the SEs for the items at each decile 
+- Save as a function for later
+
+```{r calc_SE, include = T, echo = T}
+calculate_deciles <- function(samples1, samples2, samples3){
   # calculate the SEs and the cutoff scores 
@@
   cutoffs1 <- cutoffs2 <- cutoffs3 <- list()
+  sd_items1 <- sd_items2 <- sd_items3 <- list()
+
+  # loop and calculate for each sample size 
   for (i in 1:length(samples1)){
+  sd_items1[[i]] <- samples1[[i]] %>% group_by(item, scale) %>% 
+    summarize(sd = sd(score), .groups = "keep") %>% 
+    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
+  sd_items2[[i]] <- samples2[[i]] %>% group_by(item, scale) %>% 
+    summarize(sd = sd(score), .groups = "keep") %>% 
+    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
+  sd_items3[[i]] <- samples3[[i]] %>% group_by(item, scale) %>% 
+    summarize(sd = sd(score), .groups = "keep") %>% 
+    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
+
   SES1[[i]] <- tapply(samples1[[i]]$score,
@@
                          quantile, 
-                         probs = seq(0, .9, by = .1))
+                         probs = seq(0, .9, by = .1),
+                         na.rm = T)
   cutoffs2[[i]] <- apply(as.data.frame(SES2[[i]]), 2, 
                          quantile, 
-                         probs = seq(0, .9, by = .1))
+                         probs = seq(0, .9, by = .1),
+                         na.rm = T)
   cutoffs3[[i]] <- apply(as.data.frame(SES3[[i]]), 2, 
                          quantile, 
-                         probs = seq(0, .9, by = .1))
+                         probs = seq(0, .9, by = .1),
+                         na.rm = T)
+
+  }
+
+  # return information
+  return(list(SES1 = SES1, SES2 = SES2, SES3 = SES3, 
+              cutoffs1 = cutoffs1, cutoffs2 = cutoffs2, cutoffs3 = cutoffs3, 
+              sd_items1 = sd_items1, sd_items2 = sd_items2, sd_items3 = sd_items3))
+
 }
@@
 
-## AIPE Simulation
+## Researcher Sample Simulation
 
-In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items.
-Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 500 to determine what the new sample size suggestion would be.
-We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds.
-The standard error of each item was calculated for each suggested sample size by pilot sample size by population type.
+- Here we are simulating the researcher side of the equation assuming they have a sample pulled from population (which we simulated above)
+- We will simulate samples from 20 to 2000 ... mostly only to see the leveling out effect of the simulation procedure, as we likely believe that over 2000 is not reasonable for most researchers. 
+- Calculate the SEs here, to see if it matches our cutoff 
+- Save this as a function to run within our pipeline (100 populations and 100 samples)
 
-```{r}
+```{r sim_sim, include = T, echo = T}
+simulate_researcher <- function(samples1, samples2, samples3, 
+                                start = 20, stop = 2000, 
+                                increase = 5){
+
   # sequence of sample sizes to try
-samplesize_values <- seq(20, 500, 5)
+  samplesize_values <- seq(from = start, to = stop, by = increase)
 
@@
   } # end all sample loop
+
+  # return sampled values 
+  return(list(sampled_values1 = sampled_values1, 
+              sampled_values2 = sampled_values2, 
+              sampled_values3 = sampled_values3,
+              samplesize_values = samplesize_values))
+
+}
 ```
 
-Next, the percent of items that fall below the cutoff scores, and thus, would be considered "well-measured" were calculated for each decile by sample.
+- Next, calculate the percent of items falling below the decile scores 
+- Pick up 80, 85, 90, and 95% of items to mimic power 
+- Save a function for our pipeline
 
-```{r}
+```{r calc_percent, include = T, echo = T}
+calculate_percent <- function(sampled_values1, sampled_values2, sampled_values3,
+                              cutoffs1, cutoffs2, cutoffs3, 
+                              samplesize_values, sizes){
+
+  # create temporary storage
   summary_list1 <- summary_list2 <- summary_list3 <- list()
+
+  # loop and calculate
   for (i in 1:length(sampled_values1)){
@@
 
-  }
+    } # end loop and calculate
 
+  # create end summary 
   summary_DF <- bind_rows(summary_list1, 
@@
                         summary_list3)
+
+  # return values 
+  return(summary_DF)
+
+}
 ```
 
-From this data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90%, and 95% of the items fall below the cutoff criterion.
-These values were chosen as popular measures of "power" in which one could determine the minimum suggested sample size (potentially 80% of items) and the maximum suggested sample size (potentially 90%).
+- Last combine everything together grabbing the proposed sample at each decile given the 80, 85, 90, and 95 scores 
+- Create a function for our pipeline
 
-```{r}
+```{r calc_percent_combine, include = T, echo = T}
+calculate_proposed <- function(summary_DF){
+
+  # 80% summary 
   summary_long_80 <- summary_DF %>% 
@@
 
+  # 85% summary
   summary_long_85 <- summary_DF %>% 
@@
 
+  # 90% summary  
   summary_long_90 <- summary_DF %>% 
@@
 
+  # 95% summary
   summary_long_95 <- summary_DF %>% 
@@
 
+  # combine the summary together
   summary_long <- rbind(summary_long_80, 
@@
 
+  # return values
+  return(summary_long)
+
+}
+```
+
+## Simulation Pipeline
+
+- Put together all the simulation pieces 
+  - Simulate the population 
+  - Simulate samples within that population 
+  - Calculate the cutoffs and deciles 
+  - Simulate like the researcher would 
+  - Calculate the percent under the decile
+  - Create final summary
+
+- We will simulate multiple times at each simulation point:
+  - Populations X 100
+  - Samples X 100
+  - Simulate Researcher X 100
+
+```{r simulation_pipeline, include = T, echo = T}
+# a list to save this in 
+full_simulation <- list()
+
+for (i in 1:10){
+
+  # simulate population 
+  populations <- simulate_population(x)
+
+  for (p in 1:10){
+
+    # simulate samples from population
+    samples <- simulate_samples(start = 20, stop = 100, increase = 5, 
+                            population1 = populations$population1,
+                            population2 = populations$population2, 
+                            population3 = populations$population3)
+
+    # calculate deciles 
+    deciles <- calculate_deciles(samples1 = samples$samples1, 
+                             samples2 = samples$samples2,
+                             samples3 = samples$samples3)
+
+    # reset this with each new sample 
+    researcher_summary <- list()
+
+    for (q in 1:10){
+
+      # simulate researcher
+      researcher_data <- simulate_researcher(samples1 = samples$samples1, 
+                                       samples2 = samples$samples2,
+                                       samples3 = samples$samples3, 
+                                       start = 20, stop = 30, 
+                                       increase = 5)
+
+      # calculate percent under each decile
+      percent_items <- calculate_percent(
+        sampled_values1 = researcher_data$sampled_values1, 
+        sampled_values2 = researcher_data$sampled_values2, 
+        sampled_values3 = researcher_data$sampled_values3, 
+        cutoffs1 = deciles$cutoffs1, 
+        cutoffs2 = deciles$cutoffs2, 
+        cutoffs3 = deciles$cutoffs3, 
+        samplesize_values = researcher_data$samplesize_values,
+        sizes = samples$sizes)
+
+      # calculate final summary data 
+      researcher_summary[[q]] <- calculate_proposed(summary_DF = percent_items)
+
+    } # tell the researcher to do it 1000 times for normal simulation numbers
+
+    # after the researcher summary, summarize and stick in dataframe 
+    # so we have a dataframe of each population by sample 
+    full_simulation[[paste(i, "_", q, sep = "")]] <- 
+      bind_rows(researcher_summary) %>% 
+      group_by(scale, original_n, source, name, power) %>% 
+      summarize(sample_size = mean(sample_size),
+                value = mean(value), .groups = "keep")
+
+  } # simulate the sample 100 times for each population
+
+} # simulate the population 100 times
+
+```
+
+```{r}
+# clean up the stuff from the summary 
   summary_long$source <- factor(summary_long$source, 
                                 levels = c("low", "med", "high"),
-                              labels = c("Low", "Medium", "High"))
+                                labels = c("Low Variance", "Medium Variance", "High Variance"))
 
-summary_long$sd_original <- .2
-summary_long$sd_original[summary_long$source == "Medium" &
-                           summary_long$scale == "1"] <- .4
-summary_long$sd_original[summary_long$source == "High" &
-                           summary_long$source == "1"] <- .8
+  summary_long$scale2 <- factor(summary_long$scale, 
+                                levels = c(1:3),
+                                labels = c("Small Scale", "Medium Scale", "Large Scale"))
 
-summary_long$sd_original[summary_long$source == "Low" &
-                           summary_long$scale == "2"] <- 4
-summary_long$sd_original[summary_long$source == "Medium" &
-                           summary_long$scale == "2"] <- 8
-summary_long$sd_original[summary_long$source == "High" &
-                           summary_long$source == "2"] <- 12
-
-summary_long$sd_original[summary_long$source == "Low" &
-                           summary_long$scale == "3"] <- 100
-summary_long$sd_original[summary_long$source == "Medium" &
-                           summary_long$scale == "3"] <- 150
-summary_long$sd_original[summary_long$source == "High" &
-                           summary_long$source == "3"] <- 200
+  sd_items <- bind_rows(sd_items1, sd_items2, sd_items3)
+  sd_items$original_n <- rep(rep(sizes, each = 3), 3)
+  sd_items$source <- rep(c("Low Variance", "Medium Variance", "High Variance"), each = 3*length(sizes))
+  summary_long <- summary_long %>% 
+    full_join(sd_items, 
+              by = c("original_n" = "original_n", "scale" = "scale", 
+                     "source" = "source"))
 
@@
 ```
-
-# Results
-
-## Differences in Item Variance
-
-We examined if this procedure is sensitive to differences in item heterogeneity, as we should expect to collect larger samples if we wish to have a large number of items reach a threshold of acceptable variance; potentially, assuring we *could* average them if a researcher did not wish to use a more complex analysis such as multilevel modeling.
-
-The figure below illustrates the potential minimum sample size for 80% of items to achieve a desired cutoff score.
-The black dots denote the original sample size against the suggested sample size.
-By comparing the facets, we can determine that our suggested procedure does capture the differences in heterogeneity.
-As heterogeneity increases in item variances, the proposed sample size also increases, especially at stricter cutoffs.
-Missing cutoff points where sample sizes proposed would be higher than 500.
-
-```{r}
-ggplot(summary_long %>% filter(power == 80), aes(original_n, sample_size, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ source) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```
-
-## Sample Size Sensitivity to Pilot Data Size
-
-In our second question, we examined if the suggested procedure was sensitive to the amount of information present in the pilot data.
-Larger pilot data is more informative, and therefore, we should expect a lower projected sample size.
-As shown in the figure below for only the low variability data, we do not find this effect.
-These simplistic simulations from the pilot data would nearly always suggest a larger sample size - mostly in a linear trend increasing with sample sizes.
-This result comes from the nature of the procedure - if we base our estimates on some SE cutoff, we will almost always need a bit more people for items to meet those goals.
-This result does not achieve our second goal.
-
-```{r}
-ggplot(summary_long %>% filter(source == "Low"), aes(original_n, sample_size, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```
-
-Therefore, we suggest using a correction factor on the simulation procedure to account for the known asymptotic nature of power (i.e., at larger sample sizes power increases level off).
-For this function, we combined a correction factor for upward biasing of effect sizes (Hedges' correction) with the formula for exponential decay calculations.
-The decay factor is calculated as follows:
-
-$$ 1 - \sqrt{\frac{N_{pilot} - min(N_{sim})}{N_{pilot}}}^{log_2(N_{pilot})}$$
-$N_{pilot}$ indicates the sample size of the pilot data minus the minimum sample size for simulation to ensure that the smallest sample sizes do not decay (e.g., the formula zeroes out).
-This value is raised to the power of $log_2$ of the sample size of the pilot data, which decreases the impact of the decay to smaller increments for increasing sample sizes.
-This value is then multiplied by the proposed sample size.
-As show in the figure below, this correction factor produces the dsired quality of maintaining that small pilot studies should *increase* sample size, and that sample size suggestions level off as pilot study data sample size increases.
-
-```{r}
-decay <- 1-sqrt((summary_long$original_n-20)/summary_long$original_n)^log2(summary_long$original_n)
-
-summary_long$new_sample <- summary_long$sample_size*decay
-
-ggplot(summary_long %>% filter(source == "Low"), aes(original_n, new_sample, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```
-
-## Corrections for Individual Researchers
-
-We have portrayed that this procedure, with a correction factor, can perform as desired.
-However, within real scenarios, researchers will only have one pilot sample, not many as show here.
-What should the researcher do to correct their sample size on their own pilot data?
-
-To explore if we could recover the new suggested sample size from data a researcher would have, we used linear models to create a formula for calculation.
-First, the corrected sample size was predicted by the original suggested sample size.
-Next, the standard deviation of the item standard deviations was added to the equation.
-Last, we included the pilot sample size.
-
-```{r}
-user_model <- lm(new_sample ~ sample_size, data = summary_long)
-user_print <- apa_print(user_model)
-
-user_model2 <- lm(new_sample ~ sample_size + sd_original, data = summary_long)
-user_print2 <- apa_print(user_model2)
-
-user_model3 <- lm(new_sample ~ sample_size + sd_original + original_n, data = summary_long)
-user_print3 <- apa_print(user_model3)
-
-user_model4 <- lm(new_sample ~ sample_size + sd_original + original_n + scale, data = summary_long)
-user_print4 <- apa_print(user_model4)
-
-change_table <- tidy(anova(user_model, user_model2, user_model3))
-```
-
-The first model using original sample size to predict new sample size was significant, `r user_print$full_result$modelfit$r2`, capturing nearly 90% of the variance.
-The second model with item standard deviation was better then the first model *F*(`r change_table$df[2]`, `r change_table$res.df[2]`) = `r change_table$statistic[2]`, *p* < .001, `r user_print2$estimate$modelfit$r2`.
-The addition of the original pilot sample size was also significant, *F*(`r change_table$df[3]`, `r change_table$res.df[3]`) = `r change_table$statistic[3]`, *p* < .001, `r user_print3$estimate$modelfit$r2`.
-
-As shown in the table below, the new suggested sample size is proportional to the original suggested sample size (i.e., *b* < 1), which reduces the sample size suggestion.
-As variability increases, the suggested sample size also increases to capture differences in heterogeneity shown above.
-Last, in order to correct for large pilot data, the original pilot sample size decreases the new suggested sample size.
-This formula approximation captures 96% of the variance in sample size scores and should allow a researcher to estimate based on their own data.
-
-```{r results = 'asis'}
-apa_table(tidy(user_model3))
-```
-
-## Choosing an Appropriate Cutoff
-
-```{r}
-by_cutoff <- list()
-R2 <- list()
-
-for (cutoff in unique(summary_long$name)){
-  by_cutoff[[cutoff]] <- lm(new_sample ~ sample_size + sd_original + original_n, data = summary_long  %>% filter(name == cutoff))
-  R2[cutoff] <- summary(by_cutoff[[cutoff]])$r.squared
-}
-
-R2
-```
-
-Last, we examine the question of an appropriate SE decile.
-All graphs for power, variability, and correction are presented below.
-If we examine the $R^2$ values for each decile of our regression equation separately, we find that the 10% (`r round(R2[["Decile 10"]], digits = 3)`) and 30% deciles (`r round(R2[["Decile 30"]], digits = 3)`) represent the best match to our corrected sample size suggestions.
-The 10% is likely unfeasible for many researchers, as the sample sizes are often quite large.
-The 30% decile, in the corrected format, appears to meet all goals: 1) increases with heterogeneity, 2) higher suggested values for lower samples and a leveling effect at larger pilot data, and ...
-
-### Low Variability
-
-```{r}
-ggplot(summary_long %>% filter(source == "Low"), aes(original_n, sample_size, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```
-
-```{r}
-ggplot(summary_long %>% filter(source == "Low"), aes(original_n, new_sample, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```
-
-### Medium Varability
-
-```{r}
-ggplot(summary_long %>% filter(source == "Medium"), aes(original_n, sample_size, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```
-
-```{r}
-summary_long$new_sample <- summary_long$sample_size*decay
-
-ggplot(summary_long %>% filter(source == "Medium"), aes(original_n, new_sample, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```
-
-### High Variability
-
-```{r}
-ggplot(summary_long %>% filter(source == "High"), aes(original_n, sample_size, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```
-
-```{r}
-ggplot(summary_long %>% filter(source == "High"), aes(original_n, new_sample, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ power) + 
-  xlab("Original Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```

### 2022-08-12 manuscript_draft.Rmd
updating with new simulation code

@@
 
-*Samples*. Each population was then sampled as if a researcher was conducting a pilot study. The sample sizes started at 20 participants per item increasing in units of 5 up to 100 participants.
+*Samples*. Each population was then sampled as if a researcher was conducting a pilot study. The sample sizes started at 20 participants per item increasing in units of 10 up to 100 participants.
 
@@
 In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items.
-Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 2000 to determine what the new sample size suggestion would be.
+Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 2000 increasing in units of 20 to determine what the new sample size suggestion would be.
 We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds; however, the sample size simulations were estimated over this amount to determine the pattern of suggested sample sizes (i.e., the function between original sample size and proposed sample size).
@@
 
-In this procedure, we identified three points of simulation: the population, the samples from those populations, and the researcher sample simulation.
-In order to minimize any potentially random quirks, we ran each of these samples 100 times for a total of 1000000 simulations.
+In order to minimize any potentially random quirks, we simulated the sample selection from the population 100 times and the researcher simulation 100 times for each of those selections, resulting in 10000 simulations of all combinations of variables (i.e., scale of the data, heterogeneity, pilot study size, researcher simulation size).
 The average of these simulations is presented in the results.
@@
 ```{r import-data}
-
+summary_long <- import("simulation_study/simulated_summary_data.csv")
 ```

### 2022-08-12 simulation_study/simulate_summary.Rmd
updating with new simulation code

--- 2022-08-12 simulation_study/simulate_cutoff.Rmd
@@
 full_simulation <- list()
-
-for (i in 1:10){
-
-  # simulate population 
 populations <- simulate_population(x)
 
-  for (p in 1:10){
+for (p in 1:100){
 
   # simulate samples from population
-    samples <- simulate_samples(start = 20, stop = 100, increase = 5, 
+  samples <- simulate_samples(start = 20, stop = 100, increase = 10, 
                           population1 = populations$population1,
@@
 
+  cat(paste(p, Sys.time(), "\n"))
+
   # reset this with each new sample 
@@
 
-    for (q in 1:10){
+  for (q in 1:100){
 
@@
                                      samples3 = samples$samples3, 
-                                       start = 20, stop = 30, 
-                                       increase = 5)
+                                     start = 20, stop = 2000, 
+                                     increase = 20)
 
@@
 
-    } # tell the researcher to do it 1000 times for normal simulation numbers
+  } # researcher 100 simulations
 
@@
   # so we have a dataframe of each population by sample 
-    full_simulation[[paste(i, "_", q, sep = "")]] <- 
+  full_simulation[[paste(p, Sys.time(), sep = "_")]] <-
    bind_rows(researcher_summary) %>%
@@
 
-  } # simulate the sample 100 times for each population
+  # writing out takes time but don't want to lose 100 researcher sims
+  # ran this across computers so make clear when run 
+  saveRDS(full_simulation, file = paste("simulation_", Sys.time(), ".Rdata", sep = ""))
 
-} # simulate the population 100 times
-
+} # simulate sampling 100 times from the population
 ```

### 2022-08-15 simulation_study/simulate_summary.Rmd
fix stupid mistake

@@
 ```{r}
-set.seed(895893)
+set.seed(895893) # 233232
 library(dplyr)
@@
 - We will simulate multiple times at each simulation point:
-  - Populations X 100
   - Samples X 100
@@
   # so we have a dataframe of each population by sample 
-  full_simulation[[paste(p, Sys.time(), sep = "_")]] <-
+  full_simulation[[p]] <-
    bind_rows(researcher_summary) %>%
@@
              value = mean(value), .groups = "keep")
+
+  # clean up the values from the summary while you have the data 
+  full_simulation[[p]]$source <- factor(full_simulation[[p]]$source, 
+                                levels = c("low", "med", "high"),
+                                labels = c("Low Variance", 
+                                           "Medium Variance", 
+                                           "High Variance"))
+
+  full_simulation[[p]]$scale2 <- factor(full_simulation[[p]]$scale, 
+                                levels = c(1:3),
+                                labels = c("Small Scale", 
+                                           "Medium Scale", 
+                                           "Large Scale"))
+
+  # grab the sd information from the pilot sample
+  # note this is not the researcher simulated sample 
+  sd_items <- bind_rows(deciles$sd_items1, deciles$sd_items2, deciles$sd_items3)
+  sd_items$source <- rep(c("Low Variance", 
+                           "Medium Variance", 
+                           "High Variance"), 
+                         each = 3*length(samples$sizes))
+  sd_items$original_n <- rep(rep(samples$sizes, each = 3), 3)
+
+  # add together that information 
+  full_simulation[[p]] <- full_simulation[[p]] %>% 
+    full_join(sd_items, 
+              by = c("original_n" = "original_n", 
+                     "scale" = "scale", 
+                     "source" = "source"))
+
+  full_simulation[[p]]$name <- gsub("Percent_Below", "Decile ", full_simulation[[p]]$name)
 
@@
 
+## Put Together Data
+
 ```{r}
-# clean up the stuff from the summary 
-  summary_long$source <- factor(summary_long$source, 
-                                levels = c("low", "med", "high"),
-                                labels = c("Low Variance", "Medium Variance", "High Variance"))
+sim_mac <- readRDS("simulation_2022-08-14 12:54:44.Rdata")
+sim_win
+sim_linux
 
-  summary_long$scale2 <- factor(summary_long$scale, 
-                                levels = c(1:3),
-                                labels = c("Small Scale", "Medium Scale", "Large Scale"))
+all_sim <- bind_rows(sim_mac, .id = "id")
 
-  sd_items <- bind_rows(sd_items1, sd_items2, sd_items3)
-  sd_items$original_n <- rep(rep(sizes, each = 3), 3)
-  sd_items$source <- rep(c("Low Variance", "Medium Variance", "High Variance"), each = 3*length(sizes))
-  summary_long <- summary_long %>% 
-    full_join(sd_items, 
-              by = c("original_n" = "original_n", "scale" = "scale", 
-                     "source" = "source"))
+summary_long <- all_sim %>% 
+  group_by(scale, original_n, source, name, power) %>% 
+  summarize(sample_size = mean(sample_size),
+             value = mean(value), .groups = "keep")
+```
 
-  summary_long$name <- gsub("Percent_Below", "Decile ", summary_long$name)
+## Export Final Data
+
+```{r}
+
 ```

### 2022-08-17 manuscript_draft.Rmd
fix stupid mistake
some mini updates with simulations

@@
 ```{r import-data}
-summary_long <- import("simulation_study/simulated_summary_data.csv")
+summary_long <- import("simulation_study/simulated_summary_data_update.csv")
+
+# reorder for graphs
+summary_long$source <- factor(summary_long$source, 
+                              levels = c("Low Variance", "Medium Variance", "High Variance"))
+summary_long$scale2 <- factor(summary_long$scale2, 
+                              levels = c("Small Scale", "Medium Scale", "Large Scale"))
+
+# a choice to eliminate the low deciles
+summary_long <- summary_long %>% filter(name != "Decile 0" &  name != "Decile 10" & name != "Decile 20")
 ```
@@
 ```{r item-figure, echo = F, fig.cap="Add a good caption here."}
-summary_long$source <- factor(summary_long$source, 
-                              levels = c("Low Variance", "Medium Variance", "High Variance"))
-
 ggplot(summary_long %>% filter(power == 80), aes(original_n, sample_size, color = name)) + 

### 2022-08-17 simulation_study/simulate_summary.Rmd
some mini updates with simulations

@@
 
-```{r}
-sim_mac <- readRDS("simulation_2022-08-14 12:54:44.Rdata")
-sim_win
-sim_linux
+```{r eval = F}
+sim_mac <- readRDS("simulation_2022-08-15 06:40:46.Rdata")
+sim_win <- readRDS("simulation_windows3.Rdata")
+sim_linux <- readRDS("simulation_linux4.Rdata")
 
-all_sim <- bind_rows(sim_mac, .id = "id")
+all_sim <- bind_rows(sim_mac, .id = "id") %>% 
+  bind_rows(sim_win, .id = "id") %>% 
+  bind_rows(sim_linux, .id = "id")
 
 summary_long <- all_sim %>% 
-  group_by(scale, original_n, source, name, power) %>% 
+  group_by(scale, original_n, source, name, power, scale2) %>% 
   summarize(sample_size = mean(sample_size),
-             value = mean(value), .groups = "keep")
+            value = mean(value), 
+            sd_item = mean(sd_item),
+            .groups = "keep",)
 ```
@@
 ```{r}
-
+export(summary_long, "simulated_summary_data_update.csv", row.names = F)
 ```

### 2022-09-12 simulation_study/simulate_summary.Rmd
data updates

--- 2022-08-15 simulation_study/simulate_summary.Rmd
@@
 
-```{r simulation_pipeline, include = T, echo = T}
+```{r simulation_pipeline, include = T, echo = T, eval = F}
 # a list to save this in 
@@
 
-```{r}
-sim_mac <- readRDS("simulation_2022-08-14 12:54:44.Rdata")
-sim_win
-sim_linux
-
-all_sim <- bind_rows(sim_mac, .id = "id")
+```{r eval = F}
+all_sim <- bind_rows(
+  bind_rows(readRDS("simulation_linux4_0815.Rdata")), 
+  bind_rows(readRDS("simulation_linux8_0815.Rdata")), 
+  bind_rows(readRDS("simulation_linux8_0816.Rdata")),
+  bind_rows(readRDS("simulation_linux36_0819.Rdata")),
+  bind_rows(readRDS("simulation_windows3_0815.Rdata")), 
+  bind_rows(readRDS("simulation_windows39_0819.Rdata")),
+  bind_rows(readRDS("simulation_mac_0815.Rdata")), 
+  bind_rows(readRDS("simulation_mac_0817.Rdata"))
+) %>% unique()
 
 summary_long <- all_sim %>% 
-  group_by(scale, original_n, source, name, power) %>% 
-  summarize(sample_size = mean(sample_size),
-             value = mean(value), .groups = "keep")
+  group_by(scale, original_n, source, name, power, scale2) %>% 
+  summarize(
+    n = n(),
+    sample_size = mean(sample_size),
+    value = mean(value), 
+    sd_item = mean(sd_item),
+    .groups = "keep")
+
 ```

### 2023-03-23 manuscript_draft.Rmd
bam bam bam

@@
 abstract: |
-  We will add an abstract that explains that we are going to 1) talk about this cool procedure, 2) show some simulations that the procedure works, and 3) give two examples of the procedure in action.
-You should try it out!
+
+  Sample size planning for research studies often focuses on obtaining a significant result given a specified level of power, alpha, and proposed effect size.
+This planning generally requires prior knowledge of study design and a statistical analysis to calculate the proposed sample size.
+However, there may not be just one specific testable analysis from which to derive power [@silberzahn2018many] or even a hypothesis to test for the project (e.g., stimuli database creation).
+Newer power and sample size planning suggestions include Accuracy in Parameter Estimation [AIPE, @kelley2007, @maxwell2008] and simulation of proposed analyses [@chalmers2020].
+These toolkits provide flexibility in traditional power analyses that focus on the if-this-then-that approach, yet, both AIPE and simulation require either a specific parameter (e.g., mean, effect size, etc.) or statistical test for planning sample size.
+In this tutorial, we explore how these latter two approaches can be combined to accommodate studies that may not have a specific hypothesis test or wish to account for the potential of a multiverse of analyses.
+Specifically, the examples focus on studies that implement multiple items and suggest that sample sizes can be planned to measure those items adequately and accurately, regardless of statistical test.
+Results show that pilot data can be used to determine a sample size that represents well-measured data, and multiple code vignettes will be provided for researchers to adapt and apply to their own measures.
 
@@
 
-# bibliography      : ["r-references.bib"]
+bibliography      : ["references.bib"]
 
@@
 output            : papaja::apa6_docx
+bibliography: references.bib
 ---
@@
 
-This part we will put in a real introduction. Things we need to discuss:
+Statistical power and power analyses are arguably one of the most important components to planning a research study [@cohen1990].
+Yet, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts [@hardwicke2020, @doi:10.1177/1745691620979806].
+The replication "crisis" and credibility revolution have shown that many published studies in psychology are underpowered [@opensciencecollaboration2015, @vazire2018].
+Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020, @d.chambers2014, @nosek2014] and then should summarily include a power analysis to determine the sample size necessary to detect the expected effect.
+Given the combined issues of publish-or-perish and that most non-significant results do not turn in published manuscripts, one may expect that power analysis would be especially critical for early career researchers [@rosenthal1979, @simmons2011].
+Potentially, it is uninformative to publish studies that are underpowered [@halpern2002], but it can be difficult to know if a bad power analysis is better than no power analysis.
+A recent review of power analyses found in psychology journal articles indicates that researchers did not provide enough information to understand their power analyses and often chose effect sizes that were inappropriately justified [@beribisky2019].
 
-  - Repeated measures data with items
-  - Accuracy in Parameter Estimation 
-  - Adaptive sampling
-  - MLM designs account for item heterogeneity but a lot of people anova
+One potential solution to the power analysis problem is the plethora of tools made available for researchers to simply power.
+G\*Power is one of the most popular free power software options [@faul2007, @erdfelder1996] that provides a simple point and click graphical user interface for power.
+Web-based tools have also sprung up for overall and statistical test specific sample size planning including powerandsamplesize.com and <https://designingexperiments.com> [@anderson2017].
+Coding based packages, such as *pwr* [@champely2017], *faux* [@debruine2021], and *SimDesign* [@chalmers2020] can be used to examine power and sample size planning using *R*, usually with simulation.
+Researchers have to be careful using any toolkit, as errors can occur with the over-reliance on software [@nuijten2016].
+When computing sample size estimates it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result [@williamson2000, @batterham2005].
+
+Changes in publication practices and research design have also shown a new wrinkle to providing a sample size plan for a research study.
+While statistics courses often suggest that a specific research design leads to a specific statistical test, multiple Many Analysts papers have shown that - given the same data and hypothesis - researchers can come up with multiple ways to analyze the data [@silberzahn2018many, @coretta].
+Research projects often have multiple testable hypothesis, however, it is unclear which hypothesis or test the sample size planning should be performed on.
+Further, an entire set of common research publications may not even have a hypothesis to examine within their project, as they are simply providing a large, quality dataset for future reuse [i.e., stimuli database creation [@buchanan2019].
+The increased ability to compute complex statistical analyses, such as multilevel modeling, has pushed researchers with repeated measures designs to abandon creating person-level averages just to be able to use a traditional ANOVA [@brysbaert2018].
+These analyses have also made it clear that we should be careful to assume that all items in a research study have the "effect", as there is often variability in their impact on the outcomes of the study.
+
+In this manuscript, we show a proposed method to account for variability in item effects, a potential lack of hypothesis test (or simply not a good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analyses choices.
+These methods are inspired by newer sample size planning methods including Accuracy in Parameter Estimation [AIPE, \@kelley2007, \@maxwell2008] and the ability to simulate proposed data for item estimates [@rousselet].
+AIPE shifts the focus away from finding a *significant* *p*-value to finding a parameter that is "accurately measured".
+For example, a researcher may wish to detect a specific sized correlation in a study, *r* = .35.
+They could then use AIPE to estimate the sample size needed to find a "sufficiently narrow" confidence interval around that correlation.
+Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and confidence intervals.
+Therefore, they could decide that their 95% confidence interval should be approximately [.20, .50], and sufficiently narrow was defined as a width of .30 or .15 on each side.
+While confidence intervals are related to Null Hypothesis Significance Testing (i.e., 95% confidence intervals that do not include zero would indicate a significant difference from zero at $\alpha$ \< .05), AIPE procedures instead suggest a sample size that should obtain that width of a confidence interval, regardless if it includes zero.
+
+```{r echo = F, eval = F}
+library(MBESS)
+ss.aipe.R2(Population.R2 = 0.35^2, # our correlation squared
+           conf.level = 0.95, # confidence interval 
+           width = 0.30^2, # the width squared
+           p = 1) # one predictor X to Y
+```
+
+In this approach, a researcher could use sequential testing to estimate their parameter of interest after each participant to determine if they have achieved their expected width of the confidence interval around that parameter.
+In this procedure, one should consider setting a minimum sample size (i.e., based on known data collection ability), use the confidence interval width as a stopping rule (i.e., stop data collect when the CI is narrow), and use the estimated sample size from the AIPE calculations as a potential maximum sample size.
+By defining each of these components, a research could ensure a feasible minimum sample size, a way to cease data collection when goals have been met, and a stopping rule to ensure an actual end to data collection.
+Given pilot or previously collected data, we should be able to leverage the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for repeated measures studies with expected variabillity in parameter estimates for items.
 
@@
 
-Using ideas from accuracy in parameter estimation, we suggest the following procedure to determine a sample size for each item:
+Using ideas from newer power methods, we suggest the following procedure to determine a sample size for each item:
 
@@
 Last, we included the pilot sample size.
+
 ```{r regressions, include = F, echo = F}
@@
 
-The first model using pilot sample size to predict new sample size was significant, *F*(`r user_sum$fstatistic[2]`, `r user_sum$fstatistic[3]`) = `r round(user_sum$fstatistic[1], digits = 2)`, *p* < .001, `r user_print$estimate$modelfit$r2`, capturing nearly 90% of the variance, `r user_print$estimate$sample_size`.
-The second model with item standard deviation was better than the first model *F*(`r change_table$df[2]`, `r format(change_table$res.df[2], big.mark = "")`) = `r round(change_table$statistic[2], digits = 2)`, *p* < .001, `r user_print2$estimate$modelfit$r2`.
+The first model using pilot sample size to predict new sample size was significant, *F*(`r user_sum$fstatistic[2]`, `r user_sum$fstatistic[3]`) = `r round(user_sum$fstatistic[1], digits = 2)`, *p* \< .001, `r user_print$estimate$modelfit$r2`, capturing nearly 90% of the variance, `r user_print$estimate$sample_size`.
+The second model with item standard deviation was better than the first model *F*(`r change_table$df[2]`, `r format(change_table$res.df[2], big.mark = "")`) = `r round(change_table$statistic[2], digits = 2)`, *p* \< .001, `r user_print2$estimate$modelfit$r2`.
 The item standard deviation predictor was significant, `r user_print2$full_result$sd_item`.
-The addition of the original pilot sample size was also significant, *F*(`r change_table$df[3]`, `r format(change_table$res.df[3], big.mark = "")`) = `r round(change_table$statistic[3], digits = 2)`, *p* < .001, `r user_print3$estimate$modelfit$r2`.
+The addition of the original pilot sample size was also significant, *F*(`r change_table$df[3]`, `r format(change_table$res.df[3], big.mark = "")`) = `r round(change_table$statistic[3], digits = 2)`, *p* \< .001, `r user_print3$estimate$modelfit$r2`.
 
-As shown in the final model Table \@ref(tab:table-correction), the new suggested sample size is proportional to the original suggested sample size (i.e., *b* < 1), which reduces the sample size suggestion.
+As shown in the final model Table \@ref(tab:table-correction), the new suggested sample size is proportional to the original suggested sample size (i.e., *b* \< 1), which reduces the sample size suggestion.
 As variability increases, the suggested sample size also increases to capture differences in heterogeneity shown above; however, this predictor is not significant in the final model, and only contributes a small portion of overall variance.
@@
 In a lexical decision task, participants simply select "word" for real words (e.g., *dog*) and "nonword" for pseudowords (e.g., *wug*).
-The trial level data is available here [https://elexicon.wustl.edu/, CITE].
+The trial level data is available here [<https://elexicon.wustl.edu/>, CITE].
 Critically, in each of these examples, the individual trial level data for each item is available to simulate and calculate standard errors on.
@@
 
+```{=tex}
 \begingroup
@@
 \setlength{\leftskip}{0.5in}
+```
+::: {#refs custom-style="Bibliography"}
+:::
 
-<div id="refs" custom-style="Bibliography"></div>
+```{=tex}
 \endgroup
+```

### 2023-03-29 manuscript_draft.Rmd
so much all the updates

@@
 
-bibliography      : ["references.bib"]
-
 floatsintext      : no
@@
 The increased ability to compute complex statistical analyses, such as multilevel modeling, has pushed researchers with repeated measures designs to abandon creating person-level averages just to be able to use a traditional ANOVA [@brysbaert2018].
-These analyses have also made it clear that we should be careful to assume that all items in a research study have the "effect", as there is often variability in their impact on the outcomes of the study.
+These analyses have also made it clear that we should be careful to assume that all items in a research study have the same "effect", as there is often variability in their impact on the outcomes of the study.
 
@@
 In this approach, a researcher could use sequential testing to estimate their parameter of interest after each participant to determine if they have achieved their expected width of the confidence interval around that parameter.
-In this procedure, one should consider setting a minimum sample size (i.e., based on known data collection ability), use the confidence interval width as a stopping rule (i.e., stop data collect when the CI is narrow), and use the estimated sample size from the AIPE calculations as a potential maximum sample size.
+One would set a minimum sample size (i.e., based on known data collection ability), use the confidence interval width as a stopping rule (i.e., stop data collect when the CI is narrow), and use the estimated sample size from the AIPE calculations as a potential maximum sample size.
 By defining each of these components, a research could ensure a feasible minimum sample size, a way to cease data collection when goals have been met, and a stopping rule to ensure an actual end to data collection.
@@
 
-Using ideas from newer power methods, we suggest the following procedure to determine a sample size for each item:
+Using these ideas, we suggest the following procedure to determine a sample size for each item:
 
@@
 In this section, we provide two examples of the suggested procedure.
-The first example includes concreteness ratings from Brysbaert et al.
-[CITE, 2014].
+The first example includes concreteness ratings from @brysbaert2014.
 Instructions given to participants denoted the difference between concrete (i.e., "refers to something that exists in reality") and abstract (i.e., "something you cannot experience directly through your senses or actions") terms.
@@
 This data represents a small scale dataset that could be used as pilot data for a study using concrete word ratings.
-The data is available at OSF LINK [CITE].
-The second dataset includes a large scale dataset with response latencies, the English Lexicon Project [CITE, Balota et al.].
+The data is available at <https://osf.io/qpmf4/>.
+The second dataset includes a large scale dataset with response latencies, the English Lexicon Project [@balota2007].
 The English Lexicon Project consists of lexical decision response latencies for English words.
 In a lexical decision task, participants simply select "word" for real words (e.g., *dog*) and "nonword" for pseudowords (e.g., *wug*).
-The trial level data is available here [<https://elexicon.wustl.edu/>, CITE].
+The trial level data is available here [<https://elexicon.wustl.edu/>].
 Critically, in each of these examples, the individual trial level data for each item is available to simulate and calculate standard errors on.
@@
 While the examples in this manuscript are traditionally cognitive linguistics focused, any research using repeated items can benefit from newer sampling techniques.
-Therefore, we provide XX example vignettes on our OSF page/GitHub site for this manuscript across a range of examples of data types provided by the authors of this manuscript.
-Examples include ...
-ADD HERE AFTER DONE WITH VIGNETTES.
+Therefore, we provide XX example vignettes and code examples on our OSF page/GitHub site for this manuscript across a range of examples of data types provided by the authors of this manuscript.
+Examples include psycholinguistics, social psychology, COVID related data, and cognitive psychology.
 
@@
 
-To be added once people say this idea doesn't suck.
+In this manuscript, we demonstrated a method using AIPE and simulation/bootstrapping to estimate a minimum and maximum sample size along with a rule for stopping data collection based on narrow confidence intervals on the parameter of interest.
+We believe this procedure is specifically useful for studies with multiple items that intend on using item level focused analyses; however, the utility of measuring each item well can extend to many analysis choices.
+By focusing on gathering quality data, we can suggest that the data is useful, regardless of outcome of any hypothesis test.
+
+One limitation of these methods would be using datasets with very large numbers of items to simulate what might happen within one study.
+For example, the English Lexicon Project includes thousands of items, and by the time we would simulate for all of those, it would likely suggest needing thousands of participants for *most items* to reach criterion.
+Alternatively, as the number of items increases, you also could potentially see very small estimates for sample size due to the correction factor (as with large numbers of items, you could find many items with standard errors below the 50% decile).
+Therefore, it would be beneficial to consider only simulating with what a participant would reasonably complete in a study.
+On the other side, small numbers of repeated items usually result in higher sample sizes proposed from the original pilot data.
+This result occurs because the smaller number of items means more samples for nearly all to reach the cutoff criteria.
+These results are not too different than what we might expect for a power analysis using a multilevel model - larger number of items tends to decrease necessary sample size, while smaller numbers of items tend to increase sample size.
+
+Second, these methods do not ensure the normal interpretation of power, wherein you know would find a specific effect for a specific test, $\alpha$, and so on.
+As discussed in the introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis, and many of the guesses within a traditional power analysis are just that - best guesses for various parameters.
+These methods could be used together to strengthen our understanding of sample size necessary for both a hypothesis test and well tuned estimation.
 

### 2023-03-29 manuscript_draft.tex
so much all the updates

@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Power to the Stimuli: Not the Effect},
-  pdfauthor={Erin M. Buchanan1 \& Other Folks as Per Order on Doc2},
-  pdflang={en-EN},
-  pdfkeywords={power, sampling, accuracy in parameter estimation},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx}
@@
 
-% \usepackage{etoolbox}
 \makeatletter
-\patchcmd{\HyOrg@maketitle}
+\usepackage{etoolbox}
+\patchcmd{\maketitle}
   {\section{\normalfont\normalsize\abstractname}}
@@
   {}{\typeout{Failed to patch abstract.}}
-\patchcmd{\HyOrg@maketitle}
+\patchcmd{\maketitle}
   {\section{\protect\normalfont{\@title}}}
@@
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Power to the Stimuli: Not the Effect},
+  pdfauthor={Erin M. Buchanan1 \& Other Folks as Per Order on Doc2},
+  pdflang={en-EN},
+  pdfkeywords={power, sampling, accuracy in parameter estimation},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 \abstract{%
-We will add an abstract that explains that we are going to 1) talk about this cool procedure, 2) show some simulations that the procedure works, and 3) give two examples of the procedure in action.
-You should try it out!
+Sample size planning for research studies often focuses on obtaining a significant result given a specified level of power, alpha, and proposed effect size.
+This planning generally requires prior knowledge of study design and a statistical analysis to calculate the proposed sample size.
+However, there may not be just one specific testable analysis from which to derive power (Silberzahn et al., 2018) or even a hypothesis to test for the project (e.g., stimuli database creation).
+Newer power and sample size planning suggestions include Accuracy in Parameter Estimation Maxwell, Kelley, \& Rausch (2008) and simulation of proposed analyses (Chalmers \& Adkins, 2020).
+These toolkits provide flexibility in traditional power analyses that focus on the if-this-then-that approach, yet, both AIPE and simulation require either a specific parameter (e.g., mean, effect size, etc.) or statistical test for planning sample size.
+In this tutorial, we explore how these latter two approaches can be combined to accommodate studies that may not have a specific hypothesis test or wish to account for the potential of a multiverse of analyses.
+Specifically, the examples focus on studies that implement multiple items and suggest that sample sizes can be planned to measure those items adequately and accurately, regardless of statistical test.
+Results show that pilot data can be used to determine a sample size that represents well-measured data, and multiple code vignettes will be provided for researchers to adapt and apply to their own measures.
 }
@@
 
-This part we will put in a real introduction. Things we need to discuss:
+Statistical power and power analyses are arguably one of the most important components to planning a research study (Cohen, 1990).
+Yet, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts Hardwicke et al.
+(2022).
+The replication ``crisis'' and credibility revolution have shown that many published studies in psychology are underpowered Vazire (2018).
+Pre-registration of a study involves outlining the study and hypotheses before data collection begins Nosek \& Lakens (2014) and then should summarily include a power analysis to determine the sample size necessary to detect the expected effect.
+Given the combined issues of publish-or-perish and that most non-significant results do not turn in published manuscripts, one may expect that power analysis would be especially critical for early career researchers Simmons, Nelson, \& Simonsohn (2011).
+Potentially, it is uninformative to publish studies that are underpowered (Halpern, 2002), but it can be difficult to know if a bad power analysis is better than no power analysis.
+A recent review of power analyses found in psychology journal articles indicates that researchers did not provide enough information to understand their power analyses and often chose effect sizes that were inappropriately justified (Beribisky, 2019).
 
-\begin{itemize}
-\tightlist
-\item
-  Repeated measures data with items
-\item
-  Accuracy in Parameter Estimation
-\item
-  Adaptive sampling
-\item
-  MLM designs account for item heterogeneity but a lot of people anova
-\end{itemize}
+One potential solution to the power analysis problem is the plethora of tools made available for researchers to simply power.
+G*Power is one of the most popular free power software options Erdfelder, Faul, \& Buchner (1996) that provides a simple point and click graphical user interface for power.
+Web-based tools have also sprung up for overall and statistical test specific sample size planning including powerandsamplesize.com and \url{https://designingexperiments.com} (Anderson, Kelley, \& Maxwell, 2017).
+Coding based packages, such as \emph{pwr} (Champely et al., 2017), \emph{faux} (DeBruine, 2021), and \emph{SimDesign} (Chalmers \& Adkins, 2020) can be used to examine power and sample size planning using \emph{R}, usually with simulation.
+Researchers have to be careful using any toolkit, as errors can occur with the over-reliance on software (Nuijten, Hartgerink, Assen, Epskamp, \& Wicherts, 2016).
+When computing sample size estimates it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result Batterham \& Atkinson (2005).
+
+Changes in publication practices and research design have also shown a new wrinkle to providing a sample size plan for a research study.
+While statistics courses often suggest that a specific research design leads to a specific statistical test, multiple Many Analysts papers have shown that - given the same data and hypothesis - researchers can come up with multiple ways to analyze the data Coretta, Casillas, {[}participating authors{]}, \& Roettger (n.d.).
+Research projects often have multiple testable hypothesis, however, it is unclear which hypothesis or test the sample size planning should be performed on.
+Further, an entire set of common research publications may not even have a hypothesis to examine within their project, as they are simply providing a large, quality dataset for future reuse {[}i.e., stimuli database creation (Buchanan, Valentine, \& Maxwell, 2019).
+The increased ability to compute complex statistical analyses, such as multilevel modeling, has pushed researchers with repeated measures designs to abandon creating person-level averages just to be able to use a traditional ANOVA (Brysbaert \& Stevens, 2018).
+These analyses have also made it clear that we should be careful to assume that all items in a research study have the same ``effect'', as there is often variability in their impact on the outcomes of the study.
+
+In this manuscript, we show a proposed method to account for variability in item effects, a potential lack of hypothesis test (or simply not a good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analyses choices.
+These methods are inspired by newer sample size planning methods including Accuracy in Parameter Estimation {[}AIPE, @kelley2007, @maxwell2008{]} and the ability to simulate proposed data for item estimates (Rousselet, Pernet, \& Wilcox, n.d.).
+AIPE shifts the focus away from finding a \emph{significant} \emph{p}-value to finding a parameter that is ``accurately measured''.
+For example, a researcher may wish to detect a specific sized correlation in a study, \emph{r} = .35.
+They could then use AIPE to estimate the sample size needed to find a ``sufficiently narrow'' confidence interval around that correlation.
+Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and confidence intervals.
+Therefore, they could decide that their 95\% confidence interval should be approximately {[}.20, .50{]}, and sufficiently narrow was defined as a width of .30 or .15 on each side.
+While confidence intervals are related to Null Hypothesis Significance Testing (i.e., 95\% confidence intervals that do not include zero would indicate a significant difference from zero at \(\alpha\) \textless{} .05), AIPE procedures instead suggest a sample size that should obtain that width of a confidence interval, regardless if it includes zero.
+
+In this approach, a researcher could use sequential testing to estimate their parameter of interest after each participant to determine if they have achieved their expected width of the confidence interval around that parameter.
+One would set a minimum sample size (i.e., based on known data collection ability), use the confidence interval width as a stopping rule (i.e., stop data collect when the CI is narrow), and use the estimated sample size from the AIPE calculations as a potential maximum sample size.
+By defining each of these components, a research could ensure a feasible minimum sample size, a way to cease data collection when goals have been met, and a stopping rule to ensure an actual end to data collection.
+Given pilot or previously collected data, we should be able to leverage the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for repeated measures studies with expected variabillity in parameter estimates for items.
 
@@
 
-Using ideas from accuracy in parameter estimation, we suggest the following procedure to determine a sample size for each item:
+Using these ideas, we suggest the following procedure to determine a sample size for each item:
 
@@
 
-\emph{Samples}. Each population was then sampled as if a researcher was conducting a pilot study. The sample sizes started at 20 participants per item increasing in units of 5 up to 100 participants.
+\emph{Samples}. Each population was then sampled as if a researcher was conducting a pilot study. The sample sizes started at 20 participants per item increasing in units of 10 up to 100 participants.
 
@@
 In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items.
-Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 500 to determine what the new sample size suggestion would be.
-We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds.
+Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 2000 increasing in units of 20 to determine what the new sample size suggestion would be.
+We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds; however, the sample size simulations were estimated over this amount to determine the pattern of suggested sample sizes (i.e., the function between original sample size and proposed sample size).
 The standard error of each item was calculated for each suggested sample size by pilot sample size by population type.
@@
 These values were chosen as popular measures of ``power'' in which one could determine the minimum suggested sample size (potentially 80\% of items) and the maximum suggested sample size (potentially 90\%).
+
+In order to minimize any potentially random quirks, we simulated the sample selection from the population 100 times and the researcher simulation 100 times for each of those selections, resulting in 10000 simulations of all combinations of variables (i.e., scale of the data, heterogeneity, pilot study size, researcher simulation size).
+The average of these simulations is presented in the results.
 
@@
 This value is then multiplied by the projected sample size.
-As show in Figure \ref{fig:corrected-figure}, this correction factor produces the desired quality of maintaining that small pilot studies should \emph{increase} sample size, and that sample size suggestions level off as pilot study data sample size increases.
+As shown in Figure \ref{fig:corrected-figure}, this correction factor produces the desired quality of maintaining that small pilot studies should \emph{increase} sample size, and that sample size suggestions level off as pilot study data sample size increases.
 
@@
 Next, the standard deviation of the item standard deviations was added to the equation to recreate heterogeneity estimates.
-The scale of the data is embedded into the standard deviation of the items (\emph{r} = 0.81), and therefore, this variable was not included separately.
+The scale of the data is embedded into the standard deviation of the items (\emph{r} = 0.80), and therefore, this variable was not included separately.
 Last, we included the pilot sample size.
 
-The first model using pilot sample size to predict new sample size was significant, \emph{F}(1, 5666) = 48,042.64, \emph{p} \textless{} .001, \(R^2 = .89\), 90\% CI \([0.89, 0.90]\), capturing nearly 90\% of the variance, \(b = 0.62\), 95\% CI \([0.62, 0.63]\).
-The second model with item standard deviation was better then the first model \emph{F}(1, 5665) = 55.21, \emph{p} \textless{} .001, \(R^2 = .89\), 90\% CI \([0.89, 0.90]\).
-The item standard deviation predictor was significant, \(b = 0.02\), 95\% CI \([0.01, 0.03]\), \(t(5665) = 4.54\), \(p < .001\).
-The addition of the original pilot sample size was also significant, \emph{F}(1, 5664) = 9,529.83, \emph{p} \textless{} .001, \(R^2 = .96\), 90\% CI \([0.96, 0.96]\).
+The first model using pilot sample size to predict new sample size was significant, \emph{F}(1, 2266) = 23,280.26, \emph{p} \textless{} .001, \(R^2 = .91\), 90\% CI \([0.91, 0.92]\), capturing nearly 90\% of the variance, \(b = 0.53\), 95\% CI \([0.52, 0.54]\).
+The second model with item standard deviation was better than the first model \emph{F}(1, NULL) = 13.59, \emph{p} \textless{} .001, \(R^2 = .91\), 90\% CI \([0.91, 0.92]\).
+The item standard deviation predictor was significant, \(b = 0.01\), 95\% CI \([0.00, 0.02]\), \(t(2265) = 2.20\), \(p = .028\).
+The addition of the original pilot sample size was also significant, \emph{F}(1, NULL) = 4,101.10, \emph{p} \textless{} .001, \(R^2 = .97\), 90\% CI \([0.97, 0.97]\).
 
@@
 \midrule
-Intercept & 39.269 & 0.437 & 89.843 & < .001\\
-Projected Sample Size & 0.700 & 0.002 & 366.669 & < .001\\
-Item SD & 0.003 & 0.003 & 0.952 & .341\\
-Pilot Sample Size & -0.695 & 0.007 & -97.621 & < .001\\
+Intercept & 34.549 & 0.425 & 81.264 & < .001\\
+Projected Sample Size & 0.621 & 0.003 & 247.039 & < .001\\
+Item SD & 0.000 & 0.003 & 0.014 & .989\\
+Pilot Sample Size & -0.483 & 0.008 & -64.040 & < .001\\
 \bottomrule
@@
 First, the 0\%, 10\%, and 20\% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale, and heterogeneity.
-If we examine the \(R^2\) values for each decile of our regression equation separately, we find that the 50\% (0.96) represents the best match to our corrected sample size suggestions.
+If we examine the \(R^2\) values for each decile of our regression equation separately, we find that the 50\% (0.97) represents the best match to our corrected sample size suggestions.
 The 50\% decile, in the corrected format, appears to meet all goals: 1) increases with heterogeneity and scale of data, and 2) higher suggested values for small original samples and a leveling effect at larger pilot data.
@@
 In this section, we provide two examples of the suggested procedure.
-The first example includes concreteness ratings from Brysbaert et al.~{[}CITE, 2014{]}.
+The first example includes concreteness ratings from Brysbaert, Warriner, and Kuperman (2014).
 Instructions given to participants denoted the difference between concrete (i.e., ``refers to something that exists in reality'') and abstract (i.e., ``something you cannot experience directly through your senses or actions'') terms.
@@
 This data represents a small scale dataset that could be used as pilot data for a study using concrete word ratings.
-The data is avaliable at OSF LINK {[}CITE{]}.
-The second dataset includes a large scale dataset with response latencies, the English Lexicon Project {[}CITE, Balota et al.{]}.
+The data is available at \url{https://osf.io/qpmf4/}.
+The second dataset includes a large scale dataset with response latencies, the English Lexicon Project (Balota et al., 2007).
 The English Lexicon Project consists of lexical decision response latencies for English words.
 In a lexical decision task, participants simply select ``word'' for real words (e.g., \emph{dog}) and ``nonword'' for pseudowords (e.g., \emph{wug}).
-The trial level data is avaliable here {[}\url{https://elexicon.wustl.edu/}, CITE{]}.
+The trial level data is available here {[}\url{https://elexicon.wustl.edu/}{]}.
 Critically, in each of these examples, the individual trial level data for each item is available to simulate and calculate standard errors on.
@@
 In our fictional study for this example, we selected 100 random words to show participants.
-In the original study, not every particpant rated every word, which created uneven sample sizes for each word.
-In our random sample of 100 words, the average pilot sample size was 27.96 (\emph{SD} = 1.47), and we will use 28 as our pilot sample size for this example.
+In the original study, not every participant rated every word, which created uneven sample sizes for each word.
+In our random sample of 100 words, the average pilot sample size was 28.15 (\emph{SD} = 1.59), and we will use 28 as our pilot sample size for this example.
 All ``do not know'' ratings were set as missing data.
@@
 The sample sizes for 85\% (48), 90\% (48), and 95\% (51) are also options for sample size suggestions.
-Finally, we calculated the potential amount of data retention given that participants could indicate they did not know a word (\(M_{correct}\) = 0.82, \emph{SD} = 0.24).
-In order to account for this facet, the potential sample sizes were multiplied by 1/0.82, which results in a suggested sample of 54, 59, and 63.
-Therefore, we could designate our minimum sample per item as 54, stopping rule of 0.25, and maximum sample size of 63.
+Finally, we calculated the potential amount of data retention given that participants could indicate they did not know a word (\(M_{correct}\) = 0.79, \emph{SD} = 0.24).
+In order to account for this facet, the potential sample sizes were multiplied by 1/0.79, which results in a suggested sample of 56, 61, and 65.
+Therefore, we could designate our minimum sample per item as 56, stopping rule of 0.25, and maximum sample size of 65.
 
@@
 For our example study, we will randomly select 500 real words and 500 non-words to show participants.
-The average pilot sample size for this random sample was 32.68 (\emph{SD} = 0.64), and \emph{n} = 33 will be our pilot size for this example.
-Again, participants are expected to make mistakes, and we calculated percent correct as 0.86, which was roughly even in the two stimulus categories: \(M_{word}\) = 0.84 and \(M_{non-word}\) = 0.87.
-The 50\% decile for items standard error was 61.32 for our cutoff criterion.
-We additionally checked to ensure that the two stimulus types did not have very different cutoff criterions: 50\% decile \(SE_{words}\) = 56.94, 50\% decile \(SE_{nonwords}\) = 65.61.
-In this scenario, we could chose to go with the lower SE to be more conservative (i.e., higher projected sample size).
+The average pilot sample size for this random sample was 32.72 (\emph{SD} = 0.64), and \emph{n} = 33 will be our pilot size for this example.
+Again, participants are expected to make mistakes, and we calculated percent correct as 0.85, which was roughly even in the two stimulus categories: \(M_{word}\) = 0.82 and \(M_{non-word}\) = 0.88.
+The 50\% decile for items standard error was 61.23 for our cutoff criterion.
+We additionally checked to ensure that the two stimulus types did not have very different cutoff criterions: 50\% decile \(SE_{words}\) = 58.98, 50\% decile \(SE_{nonwords}\) = 62.61.
+In this scenario, we could choose to go with the lower SE to be more conservative (i.e., higher projected sample size).
 Given the values were close for large scale data, we used the 50\% decile of all stimuli taken together.
@@
 A sample size of 31 would equate to 80\% of the items reaching our cutoff, along with 85\% (34), 90\% (34), and 95\% (38).
-Again, we adjusted for data loss given that participants are expected to incorrectly answer items, resulting in a suggested sample of 36, 40, and 44.
+Again, we adjusted for data loss given that participants are expected to incorrectly answer items, resulting in a suggested sample of 36, 40, and 45.
 One other possible consideration for this study is potential fatigue in showing participants 1000 target items.
@@
 We would need to double our sample sizes to account for splitting of the items across multiple sets of participants.
-Our minimum sample size for the entire study could be 72, stopping rule of 61.32, and maximum sample size of 88.
+Our minimum sample size for the entire study could be 72, stopping rule of 61.23, and maximum sample size of 90.
 This study would benefit from an adaptive design, where smaller sets items are randomly sampled for participants until they reach the minimum sample size or the cutoff criteria.
-At this point, items are probablistically sampled (e.g., higher selection probability for items that have not reached a minimum or stopping rule) until all items have reached criteria.
+At this point, items are probabilistically sampled (e.g., higher selection probability for items that have not reached a minimum or stopping rule) until all items have reached criteria.
 
@@
 While the examples in this manuscript are traditionally cognitive linguistics focused, any research using repeated items can benefit from newer sampling techniques.
-Therefore, we provide XX example vignettes on our OSF page/GitHub site for this manuscript across a range of examples of data types provided by the authors of this manuscript.
-Examples include \ldots{} ADD HERE AFTER DONE WITH VIGNETTES.
+Therefore, we provide XX example vignettes and code examples on our OSF page/GitHub site for this manuscript across a range of examples of data types provided by the authors of this manuscript.
+Examples include psycholinguistics, social psychology, COVID related data, and cognitive psychology.
 
@@
 
-To be added once people say this idea doesn't suck.
+In this manuscript, we demonstrated a method using AIPE and simulation/bootstrapping to estimate a minimum and maximum sample size along with a rule for stopping data collection based on narrow confidence intervals on the parameter of interest.
+We believe this procedure is specifically useful for studies with multiple items that intend on using item level focused analyses; however, the utility of measuring each item well can extend to many analysis choices.
+By focusing on gathering quality data, we can suggest that the data is useful, regardless of outcome of any hypothesis test.
+
+One limitation of these methods would be using datasets with very large numbers of items to simulate what might happen within one study.
+For example, the English Lexicon Project includes thousands of items, and by the time we would simulate for all of those, it would likely suggest needing thousands of participants for \emph{most items} to reach criterion.
+Alternatively, as the number of items increases, you also could potentially see very small estimates for sample size due to the correction factor (as with large numbers of items, you could find many items with standard errors below the 50\% decile).
+Therefore, it would be beneficial to consider only simulating with what a participant would reasonably complete in a study.
+On the other side, small numbers of repeated items usually result in higher sample sizes proposed from the original pilot data.
+This result occurs because the smaller number of items means more samples for nearly all to reach the cutoff criteria.
+These results are not too different than what we might expect for a power analysis using a multilevel model - larger number of items tends to decrease necessary sample size, while smaller numbers of items tend to increase sample size.
+
+Second, these methods do not ensure the normal interpretation of power, wherein you know would find a specific effect for a specific test, \(\alpha\), and so on.
+As discussed in the introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis, and many of the guesses within a traditional power analysis are just that - best guesses for various parameters.
+These methods could be used together to strengthen our understanding of sample size necessary for both a hypothesis test and well tuned estimation.
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-anderson2017}{}}%
+Anderson, S.
+F., Kelley, K., \& Maxwell, S.
+E.
+(2017).
+Sample-Size Planning for More Accurate Statistical Power: A Method Adjusting Sample Effect Sizes for Publication Bias and Uncertainty.
+\emph{Psychological Science}, \emph{28}(11), 1547--1562.
+\url{https://doi.org/10.1177/0956797617723724}
+
+\leavevmode\vadjust pre{\hypertarget{ref-balota2007}{}}%
+Balota, D.
+A., Yap, M.
+J., Hutchison, K.
+A., Cortese, M.
+J., Kessler, B., Loftis, B., \ldots{} Treiman, R.
+(2007).
+The English Lexicon Project.
+\emph{Behavior Research Methods}, \emph{39}(3), 445--459.
+\url{https://doi.org/10.3758/BF03193014}
+
+\leavevmode\vadjust pre{\hypertarget{ref-batterham2005}{}}%
+Batterham, A.
+M., \& Atkinson, G.
+(2005).
+How big does my sample need to be?
+A primer on the murky world of sample size estimation.
+\emph{Physical Therapy in Sport}, \emph{6}(3), 153--163.
+\url{https://doi.org/10.1016/j.ptsp.2005.05.004}
+
+\leavevmode\vadjust pre{\hypertarget{ref-beribisky2019}{}}%
+Beribisky, N.
+(2019).
+\emph{A Multi-Faceted Mess: A Review of Statistical Power Analysis in Psychology Journal Articles}.
+Retrieved from \url{https://yorkspace.library.yorku.ca/xmlui/handle/10315/36719}
+
+\leavevmode\vadjust pre{\hypertarget{ref-brysbaert2018}{}}%
+Brysbaert, M., \& Stevens, M. (2018). Power Analysis and Effect Size in Mixed Effects Models: A Tutorial. \emph{Journal of Cognition}, \emph{1}(1), 9. \url{https://doi.org/10.5334/joc.10}
+
+\leavevmode\vadjust pre{\hypertarget{ref-brysbaert2014}{}}%
+Brysbaert, M., Warriner, A.
+B., \& Kuperman, V.
+(2014).
+Concreteness ratings for 40 thousand generally known English word lemmas.
+\emph{Behavior Research Methods}, \emph{46}(3), 904--911.
+\url{https://doi.org/10.3758/s13428-013-0403-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-buchanan2019}{}}%
+Buchanan, E.
+M., Valentine, K.
+D., \& Maxwell, N.
+P.
+(2019).
+LAB: Linguistic Annotated Bibliography {\textendash} a searchable portal for normed database information.
+\emph{Behavior Research Methods}, \emph{51}(4), 1878--1888.
+\url{https://doi.org/10.3758/s13428-018-1130-8}
+
+\leavevmode\vadjust pre{\hypertarget{ref-chalmers2020}{}}%
+Chalmers, R.
+P., \& Adkins, M.
+C.
+(2020).
+Writing effective and reliable monte carlo simulations with the SimDesign package.
+\emph{The Quantitative Methods for Psychology}, \emph{16}(4), 248--280.
+\url{https://doi.org/10.20982/tqmp.16.4.p248}
+
+\leavevmode\vadjust pre{\hypertarget{ref-champely2017}{}}%
+Champely, S., Ekstrom, C., Dalgaard, P., Gill, J., Weibelzahl, S., Anandkumar, A., \ldots{} De Rosario, H. (2017). \emph{Pwr: Basic functions for power analysis}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cohen1990}{}}%
+Cohen, J. (1990). Things I have learned (so far). \emph{American Psychologist}, \emph{45}(12), 1304--1312. \url{https://doi.org/10.1037/0003-066X.45.12.1304}
+
+\leavevmode\vadjust pre{\hypertarget{ref-coretta}{}}%
+Coretta, S., Casillas, J., {[}participating authors{]}, \& Roettger, T.
+B.
+(n.d.).
+Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses.
+\emph{Advances in Methods and Practices in Psychological Science}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-d.chambers2014}{}}%
+D.
+Chambers, C., Feredoes, E., D.
+Muthukumaraswamy, S., J.
+Etchells, P., \& 1 Cardiff University Brain Research Imaging Centre, School of Psychology, Cardiff University; (2014).
+Instead of {``}playing the game{''} it is time to change the rules: Registered Reports at AIMS Neuroscience and beyond.
+\emph{AIMS Neuroscience}, \emph{1}(1), 4--17.
+\url{https://doi.org/10.3934/Neuroscience.2014.1.4}
+
+\leavevmode\vadjust pre{\hypertarget{ref-debruine2021}{}}%
+DeBruine, L. (2021). \emph{Faux: Simulation for factorial designs}. Zenodo. \url{https://doi.org/10.5281/ZENODO.2669586}
+
+\leavevmode\vadjust pre{\hypertarget{ref-erdfelder1996}{}}%
+Erdfelder, E., Faul, F., \& Buchner, A.
+(1996).
+GPOWER: A general power analysis program.
+\emph{Behavior Research Methods, Instruments, \& Computers}, \emph{28}(1), 1--11.
+\url{https://doi.org/10.3758/BF03203630}
+
+\leavevmode\vadjust pre{\hypertarget{ref-faul2007}{}}%
+Faul, F., Erdfelder, E., Lang, A.-G., \& Buchner, A.
+(2007).
+G*Power 3: A flexible statistical power analysis program for the social, behavioral, and biomedical sciences.
+\emph{Behavior Research Methods}, \emph{39}(2), 175--191.
+\url{https://doi.org/10.3758/BF03193146}
+
+\leavevmode\vadjust pre{\hypertarget{ref-halpern2002}{}}%
+Halpern, S. D. (2002). The Continuing Unethical Conduct of Underpowered Clinical Trials. \emph{JAMA}, \emph{288}(3), 358. \url{https://doi.org/10.1001/jama.288.3.358}
+
+\leavevmode\vadjust pre{\hypertarget{ref-doi:10.1177ux2f1745691620979806}{}}%
+Hardwicke, T.
+E., Thibault, R.
+T., Kosie, J.
+E., Wallach, J.
+D., Kidwell, M.
+C., \& Ioannidis, J.
+P.
+A.
+(2022).
+Estimating the prevalence of transparency and reproducibility-related research practices in psychology (2014{\textendash}2017).
+\emph{Perspectives on Psychological Science}, \emph{17}(1), 239--251.
+\url{https://doi.org/10.1177/1745691620979806}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hardwicke2020}{}}%
+Hardwicke, T.
+E., Wallach, J.
+D., Kidwell, M.
+C., Bendixen, T., Crüwell, S., \& Ioannidis, J.
+P.
+A.
+(2020).
+An empirical assessment of transparency and reproducibility-related research practices in the social sciences (2014{\textendash}2017).
+\emph{Royal Society Open Science}, \emph{7}(2), 190806.
+\url{https://doi.org/10.1098/rsos.190806}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kelley2007}{}}%
+Kelley, K.
+(2007).
+Sample size planning for the coefficient of variation from the accuracy in parameter estimation approach.
+\emph{Behavior Research Methods}, \emph{39}(4), 755--766.
+\url{https://doi.org/10.3758/BF03192966}
+
+\leavevmode\vadjust pre{\hypertarget{ref-maxwell2008}{}}%
+Maxwell, S.
+E., Kelley, K., \& Rausch, J.
+R.
+(2008).
+Sample size planning for statistical power and accuracy in parameter estimation.
+\emph{Annual Review of Psychology}, \emph{59}, 537--563.
+\url{https://doi.org/10.1146/annurev.psych.59.103006.093735}
+
+\leavevmode\vadjust pre{\hypertarget{ref-nosek2014}{}}%
+Nosek, B.
+A., \& Lakens, D.
+(2014).
+Registered Reports: A Method to Increase the Credibility of Published Results.
+\emph{Social Psychology}, \emph{45}(3), 137--141.
+\url{https://doi.org/10.1027/1864-9335/a000192}
+
+\leavevmode\vadjust pre{\hypertarget{ref-nuijten2016}{}}%
+Nuijten, M.
+B., Hartgerink, C.
+H.
+J., Assen, M.
+A.
+L.
+M.
+van, Epskamp, S., \& Wicherts, J.
+M.
+(2016).
+The prevalence of statistical reporting errors in psychology (1985{\textendash}2013).
+\emph{Behavior Research Methods}, \emph{48}(4), 1205--1226.
+\url{https://doi.org/10.3758/s13428-015-0664-2}
+
+\leavevmode\vadjust pre{\hypertarget{ref-opensciencecollaboration2015}{}}%
+Open Science Collaboration. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716--aac4716. \url{https://doi.org/10.1126/science.aac4716}
+
+\leavevmode\vadjust pre{\hypertarget{ref-rosenthal1979}{}}%
+Rosenthal, R. (1979). The file drawer problem and tolerance for null results. \emph{Psychological Bulletin}, \emph{86}(3), 638--641. \url{https://doi.org/10.1037/0033-2909.86.3.638}
+
+\leavevmode\vadjust pre{\hypertarget{ref-rousselet}{}}%
+Rousselet, G., Pernet, D.
+C., \& Wilcox, R.
+R.
+(n.d.).
+\emph{An introduction to the bootstrap: A versatile method to make inferences by using data-driven simulations}.
+\url{https://doi.org/10.31234/osf.io/h8ft7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-silberzahn2018many}{}}%
+Silberzahn, R., Uhlmann, E.
+L., Martin, D.
+P., Anselmi, P., Aust, F., Awtrey, E., \ldots{} others.
+(2018).
+Many analysts, one data set: Making transparent how variations in analytic choices affect results.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(3), 337356.
+
+\leavevmode\vadjust pre{\hypertarget{ref-simmons2011}{}}%
+Simmons, J.
+P., Nelson, L.
+D., \& Simonsohn, U.
+(2011).
+False-positive psychology: Undisclosed flexibility in data collection and analysis allows presenting anything as significant.
+\emph{Psychological Science}, \emph{22}(11), 1359--1366.
+\url{https://doi.org/10.1177/0956797611417632}
+
+\leavevmode\vadjust pre{\hypertarget{ref-stewart2020}{}}%
+Stewart, S., Rinke, E.
+M., McGarrigle, R., Lynott, D., Lunny, C., Lautarescu, A., \ldots{} Crook, Z.
+(2020).
+\emph{Pre-registration and registered reports: A primer from UKRN}.
+\url{https://doi.org/10.31219/osf.io/8v2n7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vazire2018}{}}%
+Vazire, S.
+(2018).
+Implications of the Credibility Revolution for Productivity, Creativity, and Progress.
+\emph{Perspectives on Psychological Science}, \emph{13}(4), 411--417.
+\url{https://doi.org/10.1177/1745691617751884}
+
+\leavevmode\vadjust pre{\hypertarget{ref-williamson2000}{}}%
+Williamson, P., Hutton, J.
+L., Bliss, J., Blunt, J., Campbell, M.
+J., \& Nicholson, R.
+(2000).
+Statistical review by research ethics committees.
+\emph{Journal of the Royal Statistical Society: Series A (Statistics in Society)}, \emph{163}(1), 5--13.
+\url{https://doi.org/10.1111/1467-985X.00152}
+
 \end{CSLReferences}

### 2023-05-17 manuscript_draft.Rmd
updating draft

@@
 ---
-title             : "Power to the Stimuli: Not the Effect"
+title             : "Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli"
 shorttitle        : "Title"
@@
 
-keywords          : "power, sampling, accuracy in parameter estimation"
+  The planning of the sample size for research studies often focuses on obtaining a significant result given a specified level of power, significance, and an anticipated effect size.
+This planning requires prior knowledge of the study design and a statistical analysis to calculate the proposed sample size.
+However, there may not be one specific testable analysis from which to derive power (Silberzahn et al., 2018) or a hypothesis to test for the project (e.g., creation of a stimuli database).
+Modern power and sample size planning suggestions include accuracy in parameter estimation (AIPE; Maxwell et al., 2008) and simulation of proposed analyses (Chalmers & Adkins, 2020).
+These toolkits provide flexibility in traditional power analyses that focus on the if-this, then-that approach, yet, both AIPE and simulation require either a specific parameter (e.g., mean, effect size, etc.) or statistical test for planning sample size.
+In this tutorial, we explore how AIPE and simulation approaches can be combined to accommodate studies that may not have a specific hypothesis test or wish to account for the potential of a multiverse of analyses.
+Specifically, the examples focus on studies that adopt multiple items and suggest that sample sizes can be planned to measure those items adequately and precisely, regardless of statistical test.
+We demonstrate that pilot data can be used to determine a sample size that represents well-measured data.
+This tutorial also provides multiple code vignettes that researchers can adapt and apply to their own measures.
+
+keywords          : "accuracy in parameter estimation, power, sampling, simulation, hypothesis testing"
 

### 2023-08-03 manuscript_draft.Rmd
update the power

@@
 title             : "Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli"
-shorttitle        : "Title"
+shorttitle        : "AIPE SIMULATION"
 
@@
 
-  Sample size planning for research studies often focuses on obtaining a significant result given a specified level of power, alpha, and proposed effect size.
-This planning generally requires prior knowledge of study design and a statistical analysis to calculate the proposed sample size.
-However, there may not be just one specific testable analysis from which to derive power [@silberzahn2018many] or even a hypothesis to test for the project (e.g., stimuli database creation).
-Newer power and sample size planning suggestions include Accuracy in Parameter Estimation [AIPE, @kelley2007, @maxwell2008] and simulation of proposed analyses [@chalmers2020].
-These toolkits provide flexibility in traditional power analyses that focus on the if-this-then-that approach, yet, both AIPE and simulation require either a specific parameter (e.g., mean, effect size, etc.) or statistical test for planning sample size.
-In this tutorial, we explore how these latter two approaches can be combined to accommodate studies that may not have a specific hypothesis test or wish to account for the potential of a multiverse of analyses.
-Specifically, the examples focus on studies that implement multiple items and suggest that sample sizes can be planned to measure those items adequately and accurately, regardless of statistical test.
-Results show that pilot data can be used to determine a sample size that represents well-measured data, and multiple code vignettes will be provided for researchers to adapt and apply to their own measures.
-
   The planning of the sample size for research studies often focuses on obtaining a significant result given a specified level of power, significance, and an anticipated effect size.
 This planning requires prior knowledge of the study design and a statistical analysis to calculate the proposed sample size.
-However, there may not be one specific testable analysis from which to derive power (Silberzahn et al., 2018) or a hypothesis to test for the project (e.g., creation of a stimuli database).
-Modern power and sample size planning suggestions include accuracy in parameter estimation (AIPE; Maxwell et al., 2008) and simulation of proposed analyses (Chalmers & Adkins, 2020).
+However, there may not be one specific testable analysis from which to derive power [@silberzahn2018many] or a hypothesis to test for the project (e.g., creation of a stimuli database).
+Modern power and sample size planning suggestions include accuracy in parameter estimation [AIPE, @kelley2007; @maxwell2008] and simulation of proposed analyses [@chalmers2020].
 These toolkits provide flexibility in traditional power analyses that focus on the if-this, then-that approach, yet, both AIPE and simulation require either a specific parameter (e.g., mean, effect size, etc.) or statistical test for planning sample size.
@@
 
-Statistical power and power analyses are arguably one of the most important components to planning a research study [@cohen1990].
-Yet, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts [@hardwicke2020, @doi:10.1177/1745691620979806].
-The replication "crisis" and credibility revolution have shown that many published studies in psychology are underpowered [@opensciencecollaboration2015, @vazire2018].
-Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020, @d.chambers2014, @nosek2014] and then should summarily include a power analysis to determine the sample size necessary to detect the expected effect.
-Given the combined issues of publish-or-perish and that most non-significant results do not turn in published manuscripts, one may expect that power analysis would be especially critical for early career researchers [@rosenthal1979, @simmons2011].
-Potentially, it is uninformative to publish studies that are underpowered [@halpern2002], but it can be difficult to know if a bad power analysis is better than no power analysis.
-A recent review of power analyses found in psychology journal articles indicates that researchers did not provide enough information to understand their power analyses and often chose effect sizes that were inappropriately justified [@beribisky2019].
+Statistical power and power analyses are arguably one of the most important components in planning a research study [@cohen1990].
+However, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts [@hardwicke2020; @doi:10.1177/1745691620979806].
+The replication \"crisis\" and credibility revolution have shown that published studies in psychology are underpowered [@opensciencecollaboration2015; @vazire2018].
+Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020; @d.chambers2014; @nosek2014], and details of a power analyses or other limitations on resources are often used to provide justification for the pre-registered sample quota.
+Given the combined issues of publish-or-perish and that most non-significant results do not result in published manuscripts, one may expect that power analysis would be especially critical for early career researchers [@rosenthal1979; @simmons2011].
+At best, an underpowered study provides limited insight [@halpern2002], and it can be difficult to know if a poorly implemented power analysis is better than no power analysis.
 
-One potential solution to the power analysis problem is the plethora of tools made available for researchers to simply power.
-G\*Power is one of the most popular free power software options [@faul2007, @erdfelder1996] that provides a simple point and click graphical user interface for power.
-Web-based tools have also sprung up for overall and statistical test specific sample size planning including powerandsamplesize.com and <https://designingexperiments.com> [@anderson2017].
-Coding based packages, such as *pwr* [@champely2017], *faux* [@debruine2021], and *SimDesign* [@chalmers2020] can be used to examine power and sample size planning using *R*, usually with simulation.
-Researchers have to be careful using any toolkit, as errors can occur with the over-reliance on software [@nuijten2016].
-When computing sample size estimates it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result [@williamson2000, @batterham2005].
+A recent review of power analyses found - across psychology journal articles - researchers did not provide enough information to understand their power analyses and often chose effect sizes that were unjustified [@beribisky2019].
+One solution to this power analysis problem is the plethora of tools made available for researchers to make power computations accessible to non-statisticians with the caveat that a solid education in power is necessary to use these tools.
+G\*Power is one of the most popular free power software options [@faul2007; @erdfelder1996] that provides a simple point and click graphical user interface for power calculations [however, see @brysbaert].
+Web-based tools have also sprung up for overall and statistical test specific sample size planning including <https://powerandsamplesize.com> and <https://designingexperiments.com> [@anderson2017].
+*R*-coding based packages, such as *pwr* [@champely2017], *faux* [@debruine2021], and *SimDesign* [@chalmers2020], can be used to examine power and sample size planning usually with simulation.
+Researchers must be careful using any toolkit, as errors can occur with the over-reliance on software [e.g., it should not be a substitute for critical thinking, @nuijten2016].
+When computing sample size estimates, it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result [@batterham2005].
+For example, if any parameter estimated by the researcher was not found in the study (i.e., a smaller effect size than used for the power analysis), then the sample size estimate tied to that specific level of power may be incorrect.
+
+STOPPED HERE
 
@@
 Examples include psycholinguistics, social psychology, COVID related data, and cognitive psychology.
+The code for this manuscript can be found in *papaja* [@aust2022] formated Rmarkdown document.
+To
 

### 2023-08-21 manuscript_draft.Rmd
figured out trackdown!

@@
 
+```{r trackdown, eval = F}
+library(trackdown)
+# initial set up 
+# upload_file(file = "manuscript_draft.Rmd",
+#             gfile = "power2stim_V2_trackdown",
+#             gpath = "Studies/PSA007 SPAM-L/Manuscripts/Power",
+#             shared_drive = "Psychological Science Accelerator")
+
+# from google 
+download_file(file = "manuscript_draft.Rmd", 
+            gfile = "power2stim_V2_trackdown",
+            gpath = "Studies/PSA007 SPAM-L/Manuscripts/Power",
+            shared_drive = "Psychological Science Accelerator")
+
+# to google
+update_file(file = "manuscript_draft.Rmd", 
+            gfile = "power2stim_V2_trackdown",
+            gpath = "Studies/PSA007 SPAM-L/Manuscripts/Power",
+            shared_drive = "Psychological Science Accelerator",
+            rich_text = TRUE,
+            hide_code = TRUE)
+```
+
 Statistical power and power analyses are arguably one of the most important components in planning a research study [@cohen1990].
 However, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts [@hardwicke2020; @doi:10.1177/1745691620979806].
-The replication \"crisis\" and credibility revolution have shown that published studies in psychology are underpowered [@opensciencecollaboration2015; @vazire2018].
+The replication "crisis" and credibility revolution have shown that published studies in psychology are underpowered [@opensciencecollaboration2015; @vazire2018].
 Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020; @d.chambers2014; @nosek2014], and details of a power analyses or other limitations on resources are often used to provide justification for the pre-registered sample quota.

### 2023-08-30 manuscript_draft.Rmd
simulations done!

--- 2023-08-03 manuscript_draft.Rmd
@@
 shorttitle    : "AIPE SIMULATION"
-
 author: 
@@
    - Resources
-
 affiliation:
@@
   institution  : "Other Instituions"
-
 authornote: |
  The authors would like to thank K.D. Valentine for her assistance in formulation of correction scores. 
-
 abstract: |
-
  The planning of the sample size for research studies often focuses on obtaining a significant result given a specified level of power, significance, and an anticipated effect size.
@@
 keywords     : "accuracy in parameter estimation, power, sampling, simulation, hypothesis testing"
-
 floatsintext   : no
@@
 draft       : no
-
 documentclass   : "apa6"
@@
 library(MuMIn)
+library(pwr)
+library(latex2exp)
+```
+
+```{r trackdown, eval = F}
+library(trackdown)
+# initial set up 
+# upload_file(file = "manuscript_draft.Rmd",
+#       gfile = "power2stim_V2_trackdown",
+#       gpath = "Studies/PSA007 SPAM-L/Manuscripts/Power",
+#       shared_drive = "Psychological Science Accelerator")
+# from google 
+download_file(file = "manuscript_draft.Rmd", 
+      gfile = "power2stim_V2_trackdown",
+      gpath = "Studies/PSA007 SPAM-L/Manuscripts/Power",
+      shared_drive = "Psychological Science Accelerator")
+# to google
+update_file(file = "manuscript_draft.Rmd", 
+      gfile = "power2stim_V2_trackdown",
+      gpath = "Studies/PSA007 SPAM-L/Manuscripts/Power",
+      shared_drive = "Psychological Science Accelerator",
+      rich_text = TRUE,
+      hide_code = TRUE)
 ```
@@
 However, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts [@hardwicke2020; @doi:10.1177/1745691620979806].
-The replication \"crisis\" and credibility revolution have shown that published studies in psychology are underpowered [@opensciencecollaboration2015; @vazire2018].
+The replication "crisis" and credibility revolution have shown that published studies in psychology are underpowered [@opensciencecollaboration2015; @vazire2018].
 Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020; @d.chambers2014; @nosek2014], and details of a power analyses or other limitations on resources are often used to provide justification for the pre-registered sample quota.
@@
 
-STOPPED HERE
+Changes in publication practices and research design have also created a new wrinkle in providing a sample size plan for a research study.
+While statistics courses often suggest that a specific research design leads to a specific statistical test, meta-science work has shown that given the same data and hypothesis, researchers can come up with multiple ways to analyze the data [@silberzahn2018many, @coretta].
+Therefore, a single power analysis only supports the specific analysis in which the researcher expects to test.
+Analyses may evolve during the research project or be subject to secondary analysis; thus, power and sample size estimation based on one analysis is potentially less useful than previously imagined.
+Further, research projects often have multiple testable hypotheses, but it is unclear which hypothesis or test should be used to estimate sample size with a power analysis.
+Last, research publications may not even have a specific, testable hypothesis, as some publications are intended to curate a large dataset for future reuse [i.e., stimuli database creation, @buchanan2019].
 
-Changes in publication practices and research design have also shown a new wrinkle to providing a sample size plan for a research study.
-While statistics courses often suggest that a specific research design leads to a specific statistical test, multiple Many Analysts papers have shown that - given the same data and hypothesis - researchers can come up with multiple ways to analyze the data [@silberzahn2018many, @coretta].
-Research projects often have multiple testable hypothesis, however, it is unclear which hypothesis or test the sample size planning should be performed on.
-Further, an entire set of common research publications may not even have a hypothesis to examine within their project, as they are simply providing a large, quality dataset for future reuse [i.e., stimuli database creation [@buchanan2019].
-The increased ability to compute complex statistical analyses, such as multilevel modeling, has pushed researchers with repeated measures designs to abandon creating person-level averages just to be able to use a traditional ANOVA [@brysbaert2018].
-These analyses have also made it clear that we should be careful to assume that all items in a research study have the same "effect", as there is often variability in their impact on the outcomes of the study.
+In light of these analytical (or lack thereof) concerns, we propose a new method that combines accuracy in parameter estimation [AIPE, @kelley2007, @maxwell2008] and bootstrapped simulation on pilot data [@rousselet].
+This method accounts for a potential lack of hypothesis test (or simply not a good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices.
+Additionally, we consider the nature of cognitive research designs that use multiple items to measure the phenomena of interest.
+For example, semantic priming is measured with multiple paired stimuli [@meyer1971], which traditionally was analyzed by creating person or item-level averages for an ANOVA [@brysbaert2018].
+However, the use of multilevel models with random effects for the stimuli used in a study have shown that we should be careful to assume that all items of a research study have the same "effect", as there is often variability in their impact on the outcomes of the study.
 
-In this manuscript, we show a proposed method to account for variability in item effects, a potential lack of hypothesis test (or simply not a good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analyses choices.
-These methods are inspired by newer sample size planning methods including Accuracy in Parameter Estimation [AIPE, \@kelley2007, \@maxwell2008] and the ability to simulate proposed data for item estimates [@rousselet].
-AIPE shifts the focus away from finding a *significant* *p*-value to finding a parameter that is "accurately measured".
-For example, a researcher may wish to detect a specific sized correlation in a study, *r* = .35.
+## Accuracy in Parameter Estimation
+
+AIPE shifts the focus away from finding a significant p-value to finding a parameter that is "accurately measured".
+For example, researchers may wish to detect a specific correlation in a study, *r* = .35.
 They could then use AIPE to estimate the sample size needed to find a "sufficiently narrow" confidence interval around that correlation.
 Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and confidence intervals.
-Therefore, they could decide that their 95% confidence interval should be approximately [.20, .50], and sufficiently narrow was defined as a width of .30 or .15 on each side.
-While confidence intervals are related to Null Hypothesis Significance Testing (i.e., 95% confidence intervals that do not include zero would indicate a significant difference from zero at $\alpha$ \< .05), AIPE procedures instead suggest a sample size that should obtain that width of a confidence interval, regardless if it includes zero.
+Therefore, they could decide that their 95% confidence interval should be approximately between .20 and .50, and sufficiently narrow could be defined as a width of .30 or .15 on each side.
+While confidence intervals are related to null hypothesis significance testing (i.e., 95% confidence intervals that do not include zero would indicate a significant difference from zero at $\alpha$ \< .05), AIPE procedures suggest how we can define a sample size with a given width of confidence interval, regardless of whether it includes zero.
+
+## Bootstrapping and Simulation
+
+Bootstrapping involves using data obtained to simulate similar datasets by drawing from the original data with replacement [@efron2000; @rousselet].
+Bootstrapping is a form of data simulation that allows one to calculate parameter estimates, confidence intervals, and more to simulate the potential population distribution, shape, and bias.
+Simulation is often paired with making up data for testing analyses and hypotheses based on proposed effect size or suggested population means.
+Generally, we would suggest starting with pilot data of a smaller sample size to understand the variability in potential items used to represent your phenomenon, especially if they are to be used in a larger study.
+However, given some background knowledge about the potential items, one could simulate example pilot data to use in.
+similar manner in our suggested procedure.
+Pilot or simulated data would be used to estimate the variability within items and select a "sufficiently narrow" window for item's confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
+Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein many items meet the required confidence interval.
 
@@
 
-In this approach, a researcher could use sequential testing to estimate their parameter of interest after each participant to determine if they have achieved their expected width of the confidence interval around that parameter.
-One would set a minimum sample size (i.e., based on known data collection ability), use the confidence interval width as a stopping rule (i.e., stop data collect when the CI is narrow), and use the estimated sample size from the AIPE calculations as a potential maximum sample size.
-By defining each of these components, a research could ensure a feasible minimum sample size, a way to cease data collection when goals have been met, and a stopping rule to ensure an actual end to data collection.
-Given pilot or previously collected data, we should be able to leverage the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for repeated measures studies with expected variabillity in parameter estimates for items.
+## Sequential Testing
 
-## Simulating Sample Size
+Researchers could then use sequential testing to estimate their parameter of interest after each participant's data was collected to determine whether they have achieved their expected width of the confidence interval around that parameter.
+One would set a minimum sample size (i.e., based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the CI is narrow, as defined above).
+Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (i.e., we expect 90% of items to meet our stopping rule with 100 participants based on simulation).
+By defining each of these components, researchers could ensure a feasible minimum sample size, a way to stop data collection when goals have been met, and a maximum sample size rule to ensure an actual end to data collection.
+Therefore, we proposed that we should be able to leverage the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for repeated measures studies with expected variability in parameter estimates for items.
+
+## Proposed Method to Calculate Sample Size
 
@@
 
-1)  Use pilot data that closely resembles your intended data collection, on the same or similar items that will be used in the study.
-In this procedure, we will assume that the pilot data is representative of a larger population of sampled items that you intend to assess.
-2)  Calculate the standard error of each item from the pilot data to create a cutoff score for when an item is "accurately measured".
-The simulations below will explore what criterion to use when determining the cutoff score from the pilot data.
-3)  Sample, with replacement, from your pilot data using sample sizes starting at 20 participants and increase in small units (e.g., 20, 25, 30) up to a value that you consider the maximum sample size.
-We will demonstrate example maximum sample sizes based on the data simulation below; however, a practical maximum sample size may be determined by time (e.g, one semester data collection) or researcher resources (e.g., 200 participants worth of funding).
-While 20 participants would likely represent an underpowered study, we simply suggest this starting minimum for simulation purposes.
-4)  For each simulated sample, calculate the standard error for each item, and use these values to ascertain the percentage of items that meet the cutoff score determined in step 2.
-5)  Find the minimum sample size that meets 80%, 85%, 90%, and 95% of the items.
-We recommend these scores to ensure that most items are accurately measured, in a similar vein to common power criteria suggestions.
-Each researcher can determine which of these is their minimum or maximum sample size (e.g., individual can choose to use 80% as a minimum and 90% as a maximum).
-6)  Report these values, and designate a minimum sample size, the cutoff criterion, and the maximum sample size.
-Each researcher should also report if they plan to use an adaptive design, which would stop data collection after meeting the cutoff criterion for each item.
+1)  Use pilot data that closely resembles data you intend to collect.
+This dataset should contain items that are identical or similar to those that will be collected in the study.
+In this procedure, it is important to ensure that the data is representative of a larger population of sampled items that you intend to assess.
+Generally, pilot data sample sizes will be small, as the goal would be to determine how many items would be necessary to reach a "stable" standard error.
 
-## Key Issues
+2)  Calculate the standard error (SE) in the subset of pilot data associated with each item to create a cut-off score that defines when items are considered "accurately measured".
+Next, the simulations below will explore what criterion should be used to determine the cut-off score from the pilot data.
 
-Given the long history of research on power, there are a few key issues that this procedure should address:
+3)  Sample, with replacement, from your pilot data using sample sizes starting at 20 participants and increase in small units up to a value that you consider the maximum sample size.
+We will demonstrate example maximum sample sizes based on the data simulation below; however, a practical maximum sample size may be determined by time (e.g., one semester data collection) or resources (e.g., 200 participants worth of funding).
+Although 20 participants would likely yield imprecise estimates, we suggest this starting minimum for simulation purposes.
 
-1)  We should see differences in projected sample sizes based on the variability in the variance for those items (i.e., heterogeneity should increase projected sample size).
-2)  We should see projected sample sizes that "level off" when pilot data increases.
-As with regular power estimates, studies can be "overpowered" to detect an effect, and this same idea should be present.
-For example, if one has a 500 person pilot study, our simulations should suggest a point at which items are likely measured well, which may have happened well before 500.
+4)  For each simulation, calculate the standard error for each item, and use these values to determine the percentage of items that meet the cutoff score determined in Step 2.
 
-# Method
+5)  Find the minimum sample size so that 80%, 85%, 90%, and 95% of the items meet the cut-off score and can be considered accurately measured.
+We recommend these scores to ensure that most items are accurately measured, in a similar vein to the common power-criterion suggestions.
+Each researcher can determine which of these is their minimum or maximum sample size (e.g., individuals can choose to use 80% as a minimum and 90% as a maximum or use values from Step 3 based on resources).
 
-## Data Simulation
+6)  Report these values, and designate a minimum sample size, the cut-off criterion, and the maximum sample size.
+Each researcher should also report if they plan to use an adaptive design, which would stop data collection after meeting the cut-off criterion for each item.
 
-*Population*.
-The data was simulated using the `rnorm` function assuming a normal distribution for 30 scale type items.
-Each population was simulated with 1000 data points.
-No items were rounded for this simulation.
+These steps are summarized in Table \@ref(tab:table-summary) 1.
+We will first demonstrate the ideas behind the steps using open data [@brysbaert2014; @balota2007].
+This example will reveal a few areas of needed exploration for the steps.
+Next, we portray simulations for the proposed procedure and find solutions to streamline and improve the sample size estimation procedure.
+Finally, we include additional resources for researchers to use to implement the estimation procedure.
 
-First, the scale of the data was manipulated by creating three sets of scales.
-The first scale was mimicked after small rating scales (i.e., 1-7 type style) using a $\mu$ = 4 with a $\sigma$ = .25 around the mean to create item mean variability.
-The second scale included a larger potential distribution of scores with a $\mu$ = 50 ($\sigma$ = 10) imitating a 0-100 scale.
-Last, the final scale included a $\mu$ = 1000 ($\sigma$ = 150) simulating a study that may include response latency data in the milliseconds.
-While there are many potential scales, these three represent a large number of potential variables in the social sciences.
-As we are suggesting item variances as a key factor for estimating sample sizes, the scale of the data is influential on the amount of *potential* variance.
-Smaller ranges of data (1-7) cannot necessarily have the same variance as larger ranges (0-100).
-
-Next, item variance heterogeneity was included by manipulating the potential $\sigma$ for each individual item.
-For small scales, the $\sigma$ = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items.
-For the medium scale of data, $\sigma$ = 25 with a variance of 4, 8, and 16.
-Last, for the large scale of data, $\sigma$ = 400 with a variance of 50, 100, and 200 for heterogeneity.
-
-*Samples*. Each population was then sampled as if a researcher was conducting a pilot study. The sample sizes started at 20 participants per item increasing in units of 10 up to 100 participants.
-
-*Cutoff Score Criterions*.
-The standard errors of each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as standard error functions as the main component in the formula for normal distribution confidence intervals.
-Standard errors were calculated at each decile of the items up to 90% (i.e., 0% smallest SE, 10% ..., 90% largest SE).
-The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criterions for cutoff scores.
-
-## Researcher Sample Simulation
-
-In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items.
-Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 2000 increasing in units of 20 to determine what the new sample size suggestion would be.
-We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds; however, the sample size simulations were estimated over this amount to determine the pattern of suggested sample sizes (i.e., the function between original sample size and proposed sample size).
-The standard error of each item was calculated for each suggested sample size by pilot sample size by population type.
-
-Next, the percent of items that fall below the cutoff scores, and thus, would be considered "well-measured" were calculated for each decile by sample.
-From this data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90%, and 95% of the items fall below the cutoff criterion.
-These values were chosen as popular measures of "power" in which one could determine the minimum suggested sample size (potentially 80% of items) and the maximum suggested sample size (potentially 90%).
-
-In order to minimize any potentially random quirks, we simulated the sample selection from the population 100 times and the researcher simulation 100 times for each of those selections, resulting in 10000 simulations of all combinations of variables (i.e., scale of the data, heterogeneity, pilot study size, researcher simulation size).
-The average of these simulations is presented in the results.
-
-# Results
-
-```{r import-data}
-summary_long <- import("simulation_study/simulated_summary_data_update.csv")
-
-# reorder for graphs
-summary_long$source <- factor(summary_long$source, 
-                              levels = c("Low Variance", "Medium Variance", "High Variance"))
-summary_long$scale2 <- factor(summary_long$scale2, 
-                              levels = c("Small Scale", "Medium Scale", "Large Scale"))
-
-# a choice to eliminate the low deciles
-summary_long <- summary_long %>% filter(name != "Decile 0" &  name != "Decile 10" & name != "Decile 20")
+```{r table-summary, results = 'asis'}
+apa_table(data.frame(
+  "Step Number" = c("1", "2" , "3", "4", "5", "6"),
+  "Proposed Steps" = c(
+    "Use representative pilot data.",
+    "Calculate standard error of each of the items in the pilot data. Determine the appropriate SE for the stopping rule.",
+    "Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.", 
+    "Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cut-off score from Step 2.",
+    "Determine the sample size at which 80%, 85%, 90%, 95% of items are below the cut-off score.",
+    "Report all values. Designate one as the minimum sample size, the cut-off score as the stopping rule for adaptive designs, and the maximum sample size.",
+    "Updated Steps" = c("Use representative pilot data.",
+    "Calculate standard error of each of the items in the pilot data. Using the 40%, determine the cut-off and stopping rule for the standard error of the items.",
+    "Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.", 
+    "Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cut-off score from Step 2.",
+    "Determine the sample size at which 80%, 85%, 90%, 95% of items are below the cut-off score.
+Use the correction formula to adjust your proposed sample size based on pilot data size, power, and percent variability.",
+    "Report all values. Designate one as the minimum sample size, the cut-off score as the stopping rule for adaptive designs, and the maximum sample size.")
+  ), 
+  caption = "Proposed Procedure for Powering Studies with Multiple Items"
+))
 ```
 
-## Differences in Item Variance
-
-We examined if this procedure is sensitive to differences in item heterogeneity, as we should expect to collect larger samples if we wish to have a large number of items reach a threshold of acceptable variance; potentially, assuring we *could* average them if a researcher did not wish to use a more complex analysis such as multilevel modeling.
-
-Figure \@ref(fig:item-figure) illustrates the potential minimum sample size for 80% of items to achieve a desired cutoff score.
-The black dots denote the original sample size against the suggested sample size.
-By comparing the facets, we can determine that our suggested procedure does capture the differences in heterogeneity.
-As heterogeneity increases in item variances, the proposed sample size also increases, especially at stricter cutoffs.
-Missing cutoff points where sample sizes proposed would be higher than 500.
-
-```{r item-figure, echo = F, fig.cap="Add a good caption here."}
-ggplot(summary_long %>% filter(power == 80), aes(original_n, sample_size, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ source*scale2) + 
-  xlab("Pilot Sample Size") + 
-  ylab("Projected Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```
-
-## Projected Sample Size Sensitivity to Pilot Sample Size
-
-In our second question, we examined if the suggested procedure was sensitive to the amount of information present in the pilot data.
-Larger pilot data is more informative, and therefore, we should expect a lower projected sample size.
-As shown in Figure \@ref(fig:sensitive-figure) for only the low variability and small scale data, we do not find this effect.
-These simulations from the pilot data would nearly always suggest a larger sample size - mostly in a linear trend increasing with sample sizes.
-This result comes from the nature of the procedure - if we base our estimates on a SE cutoff, we will almost always need a bit more people for items to meet those goals.
-This result does not achieve our second goal.
-
-```{r sensitive-figure, echo = F, fig.cap="Add good description here."}
-ggplot(summary_long %>% filter(source == "Low Variance") %>% filter(scale == 1), aes(original_n, sample_size, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ power) + 
-  xlab("Pilot Sample Size") + 
-  ylab("Projected Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```
-
-Therefore, we suggest using a correction factor on the simulation procedure to account for the known asymptotic nature of power (i.e., at larger sample sizes power increases level off).
-For this function in our simulation study, we combined a correction factor for upward biasing of effect sizes (Hedges' correction) with the formula for exponential decay calculations.
-The decay factor was calculated as follows:
-
-$$ 1 - \sqrt{\frac{N_{Pilot} - min(N_{Simulation})}{N_{Pilot}}}^{log_2(N_{Pilot})}$$
-
-$N_{Pilot}$ indicates the sample size of the pilot data minus the minimum simulated sample size to ensure that the smallest sample sizes do not decay (i.e., the formula zeroes out).
-This value is raised to the power of $log_2$ of the sample size of the pilot data, which decreases the impact of the decay to smaller increments for increasing sample sizes.
-This value is then multiplied by the projected sample size.
-As shown in Figure \@ref(fig:corrected-figure), this correction factor produces the desired quality of maintaining that small pilot studies should *increase* sample size, and that sample size suggestions level off as pilot study data sample size increases.
-
-```{r corrected-figure, echo = F, fig.cap="A corrected figure update this caption."}
-decay <- 1-sqrt((summary_long$original_n-20)/summary_long$original_n)^log2(summary_long$original_n)
-
-summary_long$new_sample <- summary_long$sample_size*decay
-
-ggplot(summary_long %>% filter(source == "Low Variance") %>% filter(scale == 1), aes(original_n, new_sample, color = name)) + 
-  geom_point() +  
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ power) + 
-  xlab("Pilot Sample Size") + 
-  ylab("Suggested Sample Size") + 
-  scale_color_discrete(name = "Cutoff Score")
-```
-
-## Corrections for Individual Researchers
-
-We have portrayed that this procedure, with a correction factor, can perform as desired.
-However, within real scenarios, researchers will only have one pilot sample, not the various simulated samples shown above.
-What should the researcher do to correct their projected sample size from their own pilot data simulations?
-
-To explore if we could recover the new projected sample size from data a researcher would have, we used linear models to create a formula for researcher correction.
-First, the corrected projected sample size was predicted by the original projected sample size.
-Next, the standard deviation of the item standard deviations was added to the equation to recreate heterogeneity estimates.
-The scale of the data is embedded into the standard deviation of the items (*r* = `r round(cor(summary_long$scale, summary_long$sd_item), digits = 2)`), and therefore, this variable was not included separately.
-Last, we included the pilot sample size.
-
-```{r regressions, include = F, echo = F}
-user_model <- lm(new_sample ~ sample_size, data = summary_long)
-user_print <- apa_print(user_model)
-user_sum <- summary(user_model)
-
-user_model2 <- lm(new_sample ~ sample_size + sd_item, data = summary_long)
-user_print2 <- apa_print(user_model2)
-
-user_model3 <- lm(new_sample ~ sample_size + sd_item + original_n, data = summary_long)
-user_print3 <- apa_print(user_model3)
-
-change_table <- tidy(anova(user_model, user_model2, user_model3))
-```
-
-```{r test-assume, include = F, echo = F}
-resid_model <- scale(residuals(user_model3))
-fit_model <- scale(fitted.values(user_model3))
-hist(resid_model)
-qqnorm(resid_model); abline(0,1)
-plot(fit_model, resid_model); abline(v = 0); abline(h = 0)
-
-# user_model_lme <- lme(new_sample ~ sample_size + sd_item + original_n, 
-#                       data = summary_long, 
-#                       random = list(~1|scale2, ~scale2|power, ~scale2|name))
-# r.squaredGLMM(user_model_lme)
-# resid_model <- scale(residuals(user_model_lme))
-# fit_model <- scale(fitted.values(user_model_lme))
-# hist(resid_model)
-# qqnorm(resid_model); abline(0,1)
-# plot(fit_model, resid_model); abline(v = 0); abline(h = 0)
-```
-
-The first model using pilot sample size to predict new sample size was significant, *F*(`r user_sum$fstatistic[2]`, `r user_sum$fstatistic[3]`) = `r round(user_sum$fstatistic[1], digits = 2)`, *p* \< .001, `r user_print$estimate$modelfit$r2`, capturing nearly 90% of the variance, `r user_print$estimate$sample_size`.
-The second model with item standard deviation was better than the first model *F*(`r change_table$df[2]`, `r format(change_table$res.df[2], big.mark = "")`) = `r round(change_table$statistic[2], digits = 2)`, *p* \< .001, `r user_print2$estimate$modelfit$r2`.
-The item standard deviation predictor was significant, `r user_print2$full_result$sd_item`.
-The addition of the original pilot sample size was also significant, *F*(`r change_table$df[3]`, `r format(change_table$res.df[3], big.mark = "")`) = `r round(change_table$statistic[3], digits = 2)`, *p* \< .001, `r user_print3$estimate$modelfit$r2`.
-
-As shown in the final model Table \@ref(tab:table-correction), the new suggested sample size is proportional to the original suggested sample size (i.e., *b* \< 1), which reduces the sample size suggestion.
-As variability increases, the suggested sample size also increases to capture differences in heterogeneity shown above; however, this predictor is not significant in the final model, and only contributes a small portion of overall variance.
-Last, in order to correct for large pilot data, the original pilot sample size decreases the new suggested sample size.
-This formula approximation captures 96% of the variance in sample size scores and should allow a researcher to estimate based on their own data.
-
-```{r table-correction, results = 'asis'}
-table_correct <- tidy(user_model3)
-table_correct$term <- c("Intercept", "Projected Sample Size", 
-                        "Item SD", "Pilot Sample Size")
-table_correct$estimate <- printnum(table_correct$estimate, digits = 3)
-table_correct$std.error <- printnum(table_correct$std.error, digits = 3)
-table_correct$statistic <- printnum(table_correct$statistic, digits = 3)
-table_correct$p.value <- printnum(table_correct$p.value, digits = 3,
-                                  zero = FALSE, gt1 = F)
-apa_table(table_correct, 
-          col.names = c("Term", "Estimate", "$SD$", "$t$", "$p$"),
-          caption = "Parameters for All Decile Cutoff Scores")
-```
-
-## Choosing an Appropriate Cutoff
-
-```{r R2-cutoff, include = F, echo = F}
-by_cutoff <- list()
-R2 <- list()
-
-for (cutoff in unique(summary_long$name)){
-  by_cutoff[[cutoff]] <- lm(new_sample ~ sample_size + sd_item + original_n, data = summary_long  %>% filter(name == cutoff))
-  R2[cutoff] <- summary(by_cutoff[[cutoff]])$r.squared
-}
-```
-
-```{r test-decile, include = F, echo = F}
-resid_model <- scale(residuals(by_cutoff[["Decile 50"]]))
-fit_model <- scale(fitted.values(by_cutoff[["Decile 50"]]))
-hist(resid_model)
-qqnorm(resid_model); abline(0,1)
-plot(fit_model, resid_model); abline(v = 0); abline(h = 0)
-
-# by scale, power, source
-separate_lms <- list()
-R2_lms <- list()
-
-for (scale_loop in unique(summary_long$scale)){
-  for (power_loop in unique(summary_long$power)){
-    for (source_loop in unique(summary_long$source)){
-      temp <- summary_long %>% 
-        filter(scale == scale_loop) %>% 
-        filter(power == power_loop) %>% 
-        filter(source == source_loop) %>% 
-        filter(name == "Decile 50")
-
-      separate_lms[[paste(scale_loop, power_loop, source_loop, sep = "_")]] <- lm(new_sample ~ sample_size + sd_item + original_n, data = temp)
-      R2_lms[[paste(scale_loop, power_loop, source_loop, sep = "_")]] <- summary(separate_lms[[paste(scale_loop, power_loop, source_loop, sep = "_")]])$r.squared
-    }
-  }
-}
-
-# resid_model <- scale(residuals(separate_lms[[10]]))
-# fit_model <- scale(fitted.values(separate_lms[[10]]))
-# hist(resid_model)
-# qqnorm(resid_model); abline(0,1)
-# plot(fit_model, resid_model); abline(v = 0); abline(h = 0)
-```
-
-Last, we examine the question of an appropriate SE decile.
-All graphs for power, heterogeneity, scale, and correction are presented online.
-First, the 0%, 10%, and 20% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale, and heterogeneity.
-If we examine the $R^2$ values for each decile of our regression equation separately, we find that the 50% (`r round(R2[["Decile 50"]], digits = 3)`) represents the best match to our corrected sample size suggestions.
-The 50% decile, in the corrected format, appears to meet all goals: 1) increases with heterogeneity and scale of data, and 2) higher suggested values for small original samples and a leveling effect at larger pilot data.
-Figure \@ref(fig:decile-figure) illustrates the corrected scores for simulations at the 50% decile recommended cutoff for item standard errors.
-
-```{r decile-figure, fig.cap = "A picture of the 50% cutoff."}
-ggplot((summary_long %>% filter(name == "Decile 50")), 
-       aes(original_n, new_sample, color = source)) + 
-  geom_point() + 
-  geom_point(aes(original_n, original_n), color = "black") + 
-  geom_line() + 
-  theme_classic() + 
-  facet_wrap(~ scale2*power) + 
-  xlab("Pilot Sample Size") + 
-  ylab("Projected Sample Size") + 
-  scale_color_discrete(name = "Heterogeneity") +
-  theme(legend.position="bottom")
-
-```
-
-The formula for finding the corrected sample size using a 50% decile is: $N_{Corrected Projected} = 39.269 + 0.700 \times X_{N_{Projected}} + 0.003 \times X_{SD Items} - 0.694 \times X_{N_{Pilot}}$.
-The suggested sample size will be estimated from the 80%, 85%, 90%, or 95% selection at the 50% decile as shown above.
-The item SD can be calculated directly from the data, and the pilot sample size is the sample size of the data from which a researcher is simulating their samples.
-Therefore, we will recommend the 50% decile of the item standard errors for step 2 of our suggested simulation procedure, and to correct the projected sample sizes found in step 5 using the correction equation above.
-While the estimated coefficients could change given variations on our simulation parameters, the general size and pattern of coefficients was consistent, and therefore, we believe this correction equation should work for a variety of use cases.
-
-# Examples
+# Example
 
@@
 elpDF <- import("./data/ELPDecisionData.zip")
+
+# grab only the overlap for this fake study
+elpDF <- elpDF %>% 
+  filter(Stimulus %in% conDF$Word)
+
+conDF <- conDF %>% 
+  filter(Word %in% elpDF$Stimulus) 
 ```
@@
 Instructions given to participants denoted the difference between concrete (i.e., "refers to something that exists in reality") and abstract (i.e., "something you cannot experience directly through your senses or actions") terms.
-Participants were then asked to rate concreteness of terms using a 1 (abstract) to 5 (concrete) scale.
+Participants were then asked to rate concreteness of terms using a 1 (*abstract*) to 5 (*concrete*) scale.
 This data represents a small scale dataset that could be used as pilot data for a study using concrete word ratings.
 The data is available at <https://osf.io/qpmf4/>.
+
 The second dataset includes a large scale dataset with response latencies, the English Lexicon Project [@balota2007].
@@
 
-## Concreteness Ratings
+For this example, our researcher wants to determine the differences in response latencies for abstract and concrete words.
+They will select *n* = 40 words from the rating data that are split evenly into abstract and concrete ends of the rating scale.
+In the experiment, each participant will rate the words for their concreteness, and then complete a lexical decision task with these words as the object of interest.
+Using both datasets, we can determine the sample size necessary to ensure adequately measured ratings and response latencies.
 
 ```{r con-estimate, echo = F, include = F}
-# sample the data
-con_words <- sample(unique(conDF$Word), size = 100, replace = F)
+# sample the data of with 20 words in each category
+average_item_score <- conDF %>% 
+  filter(Rating != "N" & Rating != "n") %>% 
+  mutate(Rating = as.numeric(Rating)) %>% 
+  group_by(Word) %>% 
+  summarize(avgrating = mean(Rating))
+
+con_words <- c(
+  average_item_score %>% 
+    filter(avgrating <= 2) %>% 
+    sample_n(size = 20) %>% 
+    pull(Word), 
+  average_item_score %>% 
+    filter(avgrating >= 4) %>% 
+    sample_n(size = 20) %>% 
+    pull(Word)
+)
+
+# filter the data 
 con_small <- conDF %>% filter(Word %in% con_words)
+
+# drop levels so R isn't dumb
 con_small$Word <- droplevels(con_small$Word)
-pilot_size <- round(mean(tapply(con_small$Rating, con_small$Word, length)))
+
+# figure out what size the pilot data actually is for each rating 
+pilot_size_c <- round(mean(tapply(con_small$Rating, con_small$Word, length)))
 
@@
 con_small$Rating <- as.numeric(con_small$Rating)
-data_loss <- con_small %>% 
+data_loss_c <- con_small %>% 
  group_by(Word) %>% 
  summarize(percent_correct = sum(!is.na(Rating))/n(), .groups = "keep")
-
 con_use <- con_small %>% filter(!is.na(Rating))
 
-# cutoff score 
-SE <- tapply(con_use$Rating, con_use$Word, function (x) { sd(x)/sqrt(length(x))})
-cutoff <- quantile(SE, probs = .50)
+# cutoff score options
+SE_c <- tapply(con_use$Rating, con_use$Word, function (x) { sd(x)/sqrt(length(x))})
+cutoff_c <- quantile(SE_c, probs = seq(from = .1, to = .9, by = .1))
 ```
 
+*Step 1*.
 The concreteness ratings data includes `r length(unique(conDF$Word))` concepts that were rated for their concreteness.
-In our fictional study for this example, we selected 100 random words to show participants.
+We randomly selected *n* = 20 abstract words ($M_{Rating}$ \<= 2) and *n* = 20 concrete words ($M_{Rating}$ \>= 4).
 In the original study, not every participant rated every word, which created uneven sample sizes for each word.
-In our random sample of 100 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size` as our pilot sample size for this example.
-All "do not know" ratings were set as missing data.
-The 50% decile for items standard error was `r round(cutoff, digits = 3)` for our cutoff criterion.
+Further, participants were allowed to indicate they did not know a word, and this data was set to missing data.
+In our sample of 40 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size` as our pilot sample size for this example.
+
+```{r elp-estimate, echo = F, include = F}
+elp_small <- elpDF %>% filter(Stimulus %in% con_words)
+# elp_small$Stimulus <- droplevels(elp_small$Stimulus) already a character
+
+# figure out pilot size
+pilot_size_e <- round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)))
+
+# figure out data loss
+elp_small$RT[elp_small$Accuracy == "0"] <- NA
+data_loss_e <- elp_small %>% 
+ group_by(Stimulus, Type) %>% 
+ summarize(percent_correct = sum(!is.na(RT))/n(), .groups = "keep")
+elp_use <- elp_small %>% filter(!is.na(RT))
+
+# cutoff score 
+SE_e <- tapply(elp_use$RT, elp_use$Stimulus, function (x) { sd(x)/sqrt(length(x))})
+cutoff_e <- quantile(SE_e, probs = seq(from = .1, to = .9, by = .1))
+```
+
+The ELP response latency data includes `r length(unique(elpDF$Stimulus))` word-forms, `r length(unique(elpDF$Stimulus[elpDF$Type == 0]))` that are listed as non-words, and `r length(unique(elpDF$Stimulus[elpDF$Type == 1]))` real words.
+We selected the same words as the concreteness subset selected above.
+The average pilot sample size for this random sample was `r format(round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2), nsmall = 2)` (*SD* = `r round(sd(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2)`), and *n* = `r pilot_size_e` will be our pilot size for this example.
+
+*Step 2*.
+Table \@ref(tab:table-example) demonstrates the cut-off scores for deciles of the standard errors for the items for the concreteness ratings and lexical decision response latencies.
+A researcher could potentially pick any of these cut-offs or other percentage options not shown here.
+We will use simulation to determine the suggestion that best captures the balance of adequately powering our sample and feasibility.
+
+*Step 3-5*.
+The pilot data was then simulated, with replacement, with samples from 20 to 300 increasing in units of 5, separately for concreteness and lexical decision times (Step 3).
+The standard error of each item was calculated for the bootstrapped samples (Step 4), and the percentage of items below each potential cut-off was gathered (Step 5).
+The smallest sample size with at least 80%, 85%, 90%, and 95% of items below the cut-off are reported in Table \@ref(tab:table-example) (Step 5).
 
@@
 samplesize_values <- seq(20, 300, 5)
-
 # create a blank table for us to save the values in 
@@
           ncol = length(unique(con_use$Word)))
-
 # make it a data frame
 sim_table <- as.data.frame(sim_table)
-
 # add a place for sample size values 
 sim_table$sample_size <- NA
-
 # loop over sample sizes
@@
 # figure out potential samples
-final_sample <- 
+final_sample_c <- 
  sim_table %>% 
@@
  group_by(sample_size) %>% 
-  summarize(Percent_Below = sum(se <= cutoff)/length(unique(con_use$Word))) %>% 
-  filter(Percent_Below >= .80) %>% 
-  arrange(Percent_Below)
+ summarize(Percent_Below10 = sum(se <= cutoff_c["10%"])/length(unique(con_use$Word)),
+          Percent_Below20 = sum(se <= cutoff_c["20%"])/length(unique(con_use$Word)),
+          Percent_Below30 = sum(se <= cutoff_c["30%"])/length(unique(con_use$Word)),
+          Percent_Below40 = sum(se <= cutoff_c["40%"])/length(unique(con_use$Word)),
+          Percent_Below50 = sum(se <= cutoff_c["50%"])/length(unique(con_use$Word)),
+          Percent_Below60 = sum(se <= cutoff_c["60%"])/length(unique(con_use$Word)),
+          Percent_Below70 = sum(se <= cutoff_c["70%"])/length(unique(con_use$Word)),
+          Percent_Below80 = sum(se <= cutoff_c["80%"])/length(unique(con_use$Word)),
+          Percent_Below90 = sum(se <= cutoff_c["90%"])/length(unique(con_use$Word))) %>% 
+  pivot_longer(-sample_size) %>% 
+  filter(value >= .80) %>% 
+  group_by(name) %>% 
+  arrange(sample_size, value)
 
-# apply correction
+values_80_c <- final_sample_c %>% 
+  group_by(name) %>% 
+  arrange(sample_size, value) %>% 
+  slice_head()
+
+values_85_c <- final_sample_c %>% 
+  filter(value >= .85) %>% 
+  group_by(name) %>% 
+  arrange(sample_size, value) %>% 
+  slice_head()
+
+values_90_c <- final_sample_c %>% 
+  filter(value >= .90) %>% 
+  group_by(name) %>% 
+  arrange(sample_size, value) %>% 
+  slice_head()
+
+values_95_c <- final_sample_c %>% 
+  filter(value >= .95) %>% 
+  group_by(name) %>% 
+  arrange(sample_size, value) %>% 
+  slice_head()
+
 # calculate the SD of the item's SD 
-sd_items <- sd(tapply(con_use$Rating, con_use$Word, sd))
-
-final_sample$new_sample <- round(39.369 + 0.700*final_sample$sample_size + 0.003*sd_items - 0.694*pilot_size)
-
-final_sample$data_loss_sample <- round(final_sample$new_sample * 1/mean(data_loss$percent_correct))
+sd_items_c <- sd(tapply(con_use$Rating, con_use$Word, sd))
 ```
-
-The pilot data was then simulated, with replacement, with samples from 20 to 300 increasing in units of 5.
-On each sample, the percent of items below the cutoff score were calculated.
-After applying our correction equation, we find that a sample size of `r final_sample$new_sample[1]` would allow for at least 80% of items to meet the cutoff criterion.
-The sample sizes for 85% (`r final_sample$new_sample[2]`), 90% (`r final_sample$new_sample[2]`), and 95% (`r final_sample$new_sample[3]`) are also options for sample size suggestions.
-Finally, we calculated the potential amount of data retention given that participants could indicate they did not know a word ($M_{correct}$ = `r round(mean(data_loss$percent_correct), digits = 2)`, *SD* = `r round(sd(data_loss$percent_correct), digits = 2)`).
-In order to account for this facet, the potential sample sizes were multiplied by 1/`r round(mean(data_loss$percent_correct), digits = 2)`, which results in a suggested sample of `r final_sample$data_loss_sample[1]`, `r final_sample$data_loss_sample[2]`, and `r final_sample$data_loss_sample[3]`.
-Therefore, we could designate our minimum sample per item as `r final_sample$data_loss_sample[1]`, stopping rule of `r cutoff`, and maximum sample size of `r final_sample$data_loss_sample[3]`.
-
-## Response Latencies
-
-```{r elp-estimate, echo = F, include = F}
-# sample the data
-elp_fake <- sample(unique(elpDF$Stimulus[elpDF$Type == 0]), size = 500, replace = F)
-elp_real <- sample(unique(elpDF$Stimulus[elpDF$Type == 1]), size = 500, replace = F)
-elp_small <- elpDF %>% filter(Stimulus %in% c(elp_fake, elp_real))
-pilot_size <- round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)))
-
-# figure out data loss
-elp_small$RT[elp_small$Accuracy == "0"] <- NA
-data_loss <- elp_small %>% 
-  group_by(Stimulus, Type) %>% 
-  summarize(percent_correct = sum(!is.na(RT))/n(), .groups = "keep")
-
-elp_use <- elp_small %>% filter(!is.na(RT))
-
-# cutoff score 
-SE <- tapply(elp_use$RT, elp_use$Stimulus, function (x) { sd(x)/sqrt(length(x))})
-cutoff <- quantile(SE, probs = .50, na.rm = T)
-
-SE_word <- tapply(elp_use$RT[elp_use$Type == 1], elp_use$Stimulus[elp_use$Type == 1], function (x) { sd(x)/sqrt(length(x))})
-SE_nw <- tapply(elp_use$RT[elp_use$Type == 0], elp_use$Stimulus[elp_use$Type == 0], function (x) { sd(x)/sqrt(length(x))})
-cutoff_word <- quantile(SE_word, probs = .50, na.rm = T)
-cutoff_nw <- quantile(SE_nw, probs = .50, na.rm = T)
-```
-
-The ELP response latency data includes `r length(unique(elpDF$Stimulus))` word-forms, `r length(unique(elpDF$Stimulus[elpDF$Type == 0]))` that are listed as non-words, and `r length(unique(elpDF$Stimulus[elpDF$Type == 1]))` real words.
-For our example study, we will randomly select 500 real words and 500 non-words to show participants.
-The average pilot sample size for this random sample was `r format(round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2), nsmall = 2)` (*SD* = `r round(sd(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2)`), and *n* = `r pilot_size` will be our pilot size for this example.
-Again, participants are expected to make mistakes, and we calculated percent correct as `r round(mean(data_loss$percent_correct), digits = 2)`, which was roughly even in the two stimulus categories: $M_{word}$ = `r round(mean(data_loss$percent_correct[data_loss$Type == 1]), digits = 2)` and $M_{non-word}$ = `r round(mean(data_loss$percent_correct[data_loss$Type == 0]), digits = 2)`.
-The 50% decile for items standard error was `r round(cutoff, digits = 3)` for our cutoff criterion.
-We additionally checked to ensure that the two stimulus types did not have very different cutoff criterions: 50% decile $SE_{words}$ = `r round(cutoff_word, digits = 3)`, 50% decile $SE_{nonwords}$ = `r round(cutoff_nw, digits =3)`.
-In this scenario, we could choose to go with the lower SE to be more conservative (i.e., higher projected sample size).
-Given the values were close for large scale data, we used the 50% decile of all stimuli taken together.
 
@@
           ncol = length(unique(elp_use$Stimulus)))
-
 # make it a data frame
 sim_table <- as.data.frame(sim_table)
-
 # add a place for sample size values 
 sim_table$sample_size <- NA
-
 # loop over sample sizes
@@
  }
-
 # figure out potential samples
-final_sample <- 
+final_sample_e <- 
  sim_table %>% 
@@
  group_by(sample_size) %>% 
-  summarize(Percent_Below = sum(se <= cutoff)/length(unique(elp_use$Word))) %>% 
-  filter(Percent_Below >= .80) %>% 
-  arrange(Percent_Below)
+ summarize(Percent_Below10 = sum(se <= cutoff_e["10%"])/length(unique(elp_use$Stimulus)),
+          Percent_Below20 = sum(se <= cutoff_e["20%"])/length(unique(elp_use$Stimulus)),
+          Percent_Below30 = sum(se <= cutoff_e["30%"])/length(unique(elp_use$Stimulus)),
+          Percent_Below40 = sum(se <= cutoff_e["40%"])/length(unique(elp_use$Stimulus)),
+          Percent_Below50 = sum(se <= cutoff_e["50%"])/length(unique(elp_use$Stimulus)),
+          Percent_Below60 = sum(se <= cutoff_e["60%"])/length(unique(elp_use$Stimulus)),
+          Percent_Below70 = sum(se <= cutoff_e["70%"])/length(unique(elp_use$Stimulus)),
+          Percent_Below80 = sum(se <= cutoff_e["80%"])/length(unique(elp_use$Stimulus)),
+          Percent_Below90 = sum(se <= cutoff_e["90%"])/length(unique(elp_use$Stimulus))) %>% 
+  pivot_longer(-sample_size) %>% 
+  filter(value >= .80) %>% 
+  group_by(name) %>% 
+  arrange(sample_size, value)
 
-# apply correction
+values_80_e <- final_sample_e %>% 
+  group_by(name) %>% 
+  arrange(sample_size, value) %>% 
+  slice_head()
+
+values_85_e <- final_sample_e %>% 
+  filter(value >= .85) %>% 
+  group_by(name) %>% 
+  arrange(sample_size, value) %>% 
+  slice_head()
+
+values_90_e <- final_sample_e %>% 
+  filter(value >= .90) %>% 
+  group_by(name) %>% 
+  arrange(sample_size, value) %>% 
+  slice_head()
+
+values_95_e <- final_sample_e %>% 
+  filter(value >= .95) %>% 
+  group_by(name) %>% 
+  arrange(value, sample_size) %>% 
+  slice_head()
+
 # calculate the SD of the item's SD 
-sd_items <- sd(tapply(elp_use$RT, elp_use$Stimulus, sd, na.rm = T), na.rm = T)
-
-final_sample$new_sample <- round(39.369 + 0.700*final_sample$sample_size + 0.003*sd_items - 0.694*pilot_size)
-
-final_sample$data_loss_sample <- round(final_sample$new_sample * 1/mean(data_loss$percent_correct))
+sd_items_e <- sd(tapply(elp_use$RT, elp_use$Stimulus, sd, na.rm = T), na.rm = T)
 ```
 
-The pilot response latency data was then simulated in the same way as described above.
-After calculating the percent below our cutoff score, we applied the correction to the projected sample sizes.
-A sample size of `r final_sample$new_sample[1]` would equate to 80% of the items reaching our cutoff, along with 85% (`r final_sample$new_sample[2]`), 90% (`r final_sample$new_sample[2]`), and 95% (`r final_sample$new_sample[3]`).
-Again, we adjusted for data loss given that participants are expected to incorrectly answer items, resulting in a suggested sample of `r final_sample$data_loss_sample[1]`, `r final_sample$data_loss_sample[2]`, and `r final_sample$data_loss_sample[3]`.
-One other possible consideration for this study is potential fatigue in showing participants 1000 target items.
-Therefore, we could designate in our research design that each participant will only receive 500 of the target items.
-We would need to double our sample sizes to account for splitting of the items across multiple sets of participants.
-Our minimum sample size for the entire study could be `r final_sample$data_loss_sample[1]*2`, stopping rule of `r round(cutoff, digits = 3)`, and maximum sample size of `r final_sample$data_loss_sample[3]*2`.
-This study would benefit from an adaptive design, where smaller sets items are randomly sampled for participants until they reach the minimum sample size or the cutoff criteria.
-At this point, items are probabilistically sampled (e.g., higher selection probability for items that have not reached a minimum or stopping rule) until all items have reached criteria.
+```{r table-example, results = 'asis'}
+# concreteness before - elp before 
+  # sample size
+  # decile cut-offs? 
+  # number of participants for each decile? 
+  # 80, 85, 90, 95
 
-## Additional Materials
+concrete_cuts <- values_80_c %>% 
+  left_join(values_85_c, by = "name") %>% 
+  left_join(values_90_c, by = "name") %>% 
+  left_join(values_95_c, by = "name") %>%
+  ungroup() %>% 
+  rename("c_80" = sample_size.x, 
+         "c_85" = sample_size.y, 
+         "c_90" = sample_size.x.x, 
+         "c_95" = sample_size.y.y) %>% 
+  select(c_80, c_85, c_90, c_95)
+
+lexical_cuts <- values_80_e %>% 
+  left_join(values_85_e, by = "name") %>% 
+  left_join(values_90_e, by = "name") %>% 
+  left_join(values_95_e, by = "name") %>%
+  ungroup() %>% 
+  rename("e_80" = sample_size.x, 
+         "e_85" = sample_size.y, 
+         "e_90" = sample_size.x.x, 
+         "e_95" = sample_size.y.y) %>% 
+  select(e_80, e_85, e_90, e_95)
+
+results_table <- matrix(NA, nrow = 9, ncol = 11)
+results_table[ , 1] <- c("Decile 10", "Decile 20", "Decile 30", 
+                         "Decile 40", "Decile 50", "Decile 60", 
+                         "Decile 70", "Decile 80", "Decile 90")
+results_table[1 , 2:11] <- unlist(c(cutoff_c[1], concrete_cuts[1 , ], 
+                                    cutoff_e[1], lexical_cuts[1, ]))
+results_table[2 , 2:11] <- unlist(c(cutoff_c[2], concrete_cuts[2 , ], 
+                                    cutoff_e[2], lexical_cuts[2, ]))
+results_table[3 , 2:11] <- unlist(c(cutoff_c[3], concrete_cuts[3 , ], 
+                                    cutoff_e[3], lexical_cuts[3, ]))
+results_table[4 , 2:11] <- unlist(c(cutoff_c[4], concrete_cuts[4 , ], 
+                                    cutoff_e[4], lexical_cuts[4, ]))
+results_table[5 , 2:11] <- unlist(c(cutoff_c[5], concrete_cuts[5 , ], 
+                                    cutoff_e[5], lexical_cuts[5, ]))
+results_table[6 , 2:11] <- unlist(c(cutoff_c[6], concrete_cuts[6 , ], 
+                                    cutoff_e[6], lexical_cuts[6, ]))
+results_table[7 , 2:11] <- unlist(c(cutoff_c[7], concrete_cuts[7 , ], 
+                                    cutoff_e[7], lexical_cuts[7, ]))
+results_table[8 , 2:11] <- unlist(c(cutoff_c[8], concrete_cuts[8 , ],
+                                    cutoff_e[8], lexical_cuts[8, ]))
+results_table[9 , 2:11] <- unlist(c(cutoff_c[9], concrete_cuts[9 , ], 
+                                 cutoff_e[9], lexical_cuts[9, ]))
+
+results_table_example <- as.data.frame(results_table)
+colnames(results_table_example) <- c("Deciles", "C Decile SE", "C 80", "C 85", "C 90", "C 95",
+             "L Decile SE", "L 80", "L 85", "L 90", "L 95")
+
+apa_table(results_table_example, 
+          caption = "Sample Size Estimates by Decile for Concreteness Example",
+          note = "Esimates are based on meeting at least the minimum percent of items (e.g., 80%) but may be estimated over that amount (e.g., 82.5%).")
+```
+
+*Step 6*.
+In the last step, the researcher would indicate their smallest sample size, the cut-off standard error criterion if they wanted to adaptively test (e.g., examine the SE after each participant and stop data collection if all items reached criteria), and their maximum sample size.
+As mentioned earlier, the decile for a balanced standard error cut-off is unclear and without guidance, a potential set of researcher degrees of freedom [CITE].
+Even though both scales appear to converge on similar sample size suggestions for each decile and percent level, the impact of scale size (i.e., 1-5 versus 0-`r max(elp_small$RT, na.rm = T)`) and heterogeneity of item standard errors (concrete $SD_{SD}$ = `r apa_num(sd_items_c)` and lexical $SD_{SD}$ = `r apa_num(sd_items_e)`) is not obvious.
+Last, by selecting the ends of the distribution for our concreteness words, skew of the distribution may additionally impact our estimates.
+Each of these will be explored in our simulation.
+
+# Simulation Method
+
+In order to evaluate our approach, we used data simulation to create representative pilot datasets of several popular cognitive scales (1-7 measurements, 0-100 percentage measurements, and 0-3000 response latency scale data).
+For each of these scales, we also manipulated item heterogeneity by simulating small differences in item variances to large differences in item variances based on original scale size.
+On each of the simulated datasets, we applied the above proposed method to determine how the procedure would perform and evaluate what criteria should be used for cut-off selection (Step 2).
+This procedure was performed on distributions in the middle of the scale (i.e., normal) and at the ceiling of the scale (i.e., skewed).
+With this simulation, we will answer several questions:
+
+1)  How do pilot data influence sample size suggestions?
+A.
+How does scale size impact sample size estimations?
+In theory, the size of the scale used should not impact the power estimates; however, larger scales have a potential for more variability in their item standard deviations (see point C).
+B.
+How does distribution skew impact sample size estimations?
+Skew can potentially decrease heterogeneity (i.e., all items are at ceiling, and therefore, variance between item standard errors is low) or could increase heterogeneity (i.e., some items are skewed, while others are not).
+Therefore, we expect skew to impact the estimates in the same way as point C.
+C.
+How does heterogeneity impact sample size estimations?
+Heterogeneity should decrease power [@alexander1994; @rheinheimer2001], and thus, increased projected sample sizes should be proposed as heterogeneity of item variances increases.
+2)  Do the results match what one might expect for traditional power curves?
+Power curves are asymptotic, that is, they "level off" as sample size increases.
+Therefore, we expect that our procedure should also demonstrate a leveling off effect as pilot data sample size increases.
+For example, if one has a 500-person pilot study, our simulations should suggest a point at which items are likely measured well, which may have happened well before 500.
+3)  What should the suggested cut-off standard error decile be?
+
+## Data Simulation
+
+Table \@ref(tab:table-sim) presents the variables and information about the simulations as a summary.
+
+```{r table-sim, results = 'asis'}
+apa_table(data.frame(
+  Information = c("Minimum", "Maximum", "$\\mu$", "$Skewed \\mu$", "$\\sigma_{\\mu}$", 
+                  "$\\sigma$", "Small $\\sigma_{\\sigma}$", 
+                  "Medium $\\sigma_{\\sigma}$", "Large $\\sigma_{\\sigma}$"),
+  Likert = c(1, 7, 4, 6, .25, 2, .2, .4, .8),
+  Percent = c(0, 100, 50, 85, 10, 25, 4, 8, 16),
+  Milliseconds = c(0, 3000, 1000, 2500, 150, 400, 50, 100, 200)
+), caption = "Parameter Values for Data Simulation"
+)
+
+```
+
+*Population*.
+We simulated data for 30 items using the rnorm function assuming a normal distribution.
+Each item's population data was simulated with 1000 data points.
+Items were rounded to the nearest whole number to mimic scales generally collected by researchers.
+Items were also rounded to their appropriate scale end points (i.e., all items below 0 on a 1-7 scale were replaced with 1, etc.).
+
+*Data Scale*.
+First, the scale of the data was manipulated by creating three sets of scales.
+The first scale was mimicked after small rating scales (i.e., 1-7 type style) using a $\mu$ = 4 with a $\sigma$ = .25 around the mean to create item mean variability.
+The second scale included a larger potential distribution of scores with a $\mu$ = 50 ($\sigma$ = 10) imitating a 0-100 scale.
+Last, the final scale included a $\mu$ = 1000 ($\sigma$ = 150) simulating a study that may include response latency data in the milliseconds.
+For the skewed distributions, the item means were set to $\mu$ = 6, 85, and 2500 respectively with the same $\sigma$ values around the item means.
+Although there are many potential scales, these three represent a large number of potential variables commonly used in the social sciences.
+As we are suggesting item variances is a key factor for estimating sample sizes, the scale of the data is influential on the amount of potential variance.
+Smaller data ranges (1-7) cannot necessarily have the same variance as larger ranges (0-100).
+
+*Item Heterogeneity*.
+Next, item heterogeneity was included by manipulating the potential for each individual item.
+For small scales, the $\sigma$ = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items.
+For the medium scale of the data, $\sigma$ = 25 with a variance of 4, 8, and 16.
+Finally, for the large scale of the data, $\sigma$ = 400 with a variance of 50, 100, and 200 for heterogeneity.
+
+*Pilot Data Samples*.
+Each of the populations shown in Table \@ref(tab:table-sim) was then sampled as if a researcher was conducting a pilot study.
+The sample sizes started at 20 participants per item, increasing in units of 10 up to 100 participants.
+Each of these samples would correspond to Step 1 of the proposed method where a researcher would use pilot data to start their estimation.
+Therefore, the simulations included 3 scales X 3 heterogeneity X 2 normal/skewed X 9 pilot sample sizes representing a potential Step 1 of our procedure.
+
+## Researcher Sample Simulation
+
+In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items.
+Assuming that each pilot sample represents a dataset that a researcher has collected (Step 1), the standard errors for each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as SE functions as the main component of the formula for normal distribution confidence intervals.
+Standard errors were calculated at each decile of the items up to 90% (i.e., 0% smallest SE, 10% ..., 90% largest SE).
+The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cut-off scores, while the higher deciles would represent less strict criteria for cut-off scores.
+
+We then simulated samples of 20 to 2000 increasing in units of 20 to determine what the new sample size suggestion would be (Step 3).
+We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds.
+However, the sample size simulations were estimated over this amount to determine the pattern of suggested sample sizes (i.e., the function between original sample size and projected sample size).
+
+Next, the percentage of items that fall below the cut-off scores and therefore would be considered "well-measured" were calculated for each decile by sample (Step 4).
+From these data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90% and 95% of the items fall below the cut-off criterion (Step 5).
+These values were chosen as popular measures of "power" in which one could determine the minimum suggested sample size (potentially 80% of the items) and the maximum suggested sample size (selected from a higher percentage, such as 90% or 95%).
+
+In order to minimize the potential for random quirks to arise, we simulated the sample selection from the population 100 times and the researcher simulation 100 times for each of those selections.
+This resulted in 1,620,000 simulations of all combinations of variables (i.e., scale of the data, heterogeneity, data skew, pilot study size, researcher simulation size).
+The average of these simulations is presented in the results.
+
+# Simulation Results
+
+```{r import-data}
+summary_long <- bind_rows(
+  import("simulation_study/simulation_middle/simulated_summary_data_middle.csv") %>% 
+    mutate(data_type = "Normal"), 
+    import("simulation_study/simulation_ceiling/simulated_summary_data_ceiling.csv") %>% 
+    mutate(data_type = "Ceiling"))
+
+summary_long$scale_size <- factor(summary_long$scale_size,
+                                  levels = c("Likert", "Percent", "Milliseconds"))
+summary_long$variability <- factor(summary_long$variability, 
+                                   levels = c("Small Heterogeneity", 
+                                              "Medium Heterogeneity", 
+                                              "Large Heterogeneity"))
+
+### make a decision here ### 
+summary_long <- summary_long %>% filter(name != "Decile 0" & name != "Decile 10" & name != "Decile 20")
+```
+
+## Pilot Data Influence on Sample Size
+
+For each variable, the plot of the pilot sample size, projected sample size (i.e., what the simulation suggested), and power levels are presented below.
+The large number of variables means we cannot plot them all simultaneously, and therefore, we averaged the results across other variables for each plot.
+The entire datasets can be examined on our OSF page.
+
+### Scale Size
+
+```{r scale-size-figure, echo = F, fig.cap="Simulated pilot sample size and final projected sample size to achieve 80%, 85%, 90%, and 95% of items below threshold.
+These values are averaged over all other variables.
+Black dots represent original sample size for reference.
+"}
+likert_var <- sqrt(((7-1)^2)/4)
+percent_var <- sqrt(((100-0)^2)/4)
+millisecond_var <- sqrt(((3000-0)^2)/4)
+
+summary_long$percent_var <- ifelse(
+  summary_long$scale_size == "Likert", summary_long$sd_item/likert_var, ifelse(
+    summary_long$scale_size == "Percent", summary_long$sd_item/percent_var, 
+      summary_long$sd_item/millisecond_var
+  )
+)
+
+mean_var <- tapply(summary_long$percent_var, summary_long$scale_size, mean)
+
+ggplot(summary_long %>% 
+         group_by(original_n, power, scale_size) %>% 
+         summarize(sample_size_avg = mean(sample_size)),
+       aes(original_n, sample_size_avg, color = scale_size)) + 
+  geom_point() + 
+  geom_point(aes(original_n, original_n), color = "black") + 
+  geom_line() + 
+  theme_linedraw(base_size = 15) + 
+  xlab("Pilot Sample Size") + 
+  ylab("Projected Sample Size") + 
+  scale_color_discrete(name = "Scale Size") +
+  facet_wrap(~power)
+
+```
+
+Figure \@ref(fig:scale-size-figure) demonstrates the influence of scale size on the results separate by potential cut-off level.
+The black dots denote the original sample size for reference.
+Larger scales have more potential variability, and therefore, we see that percent and millisecond scales project a larger required sample size.
+This relationship does not appear to be linear with scale size, as percent scales often represent the highest projected sample size.
+Potentially, this finding is due to the larger proportion of possible variance -- the variance of the item standard deviations / total possible variance -- was largest for percent scales in this set of simulations ($p_{Percent}$ = `r apa_num(mean_var["Percent"], gt1 = FALSE)`.
+This finding may be an interaction with heterogeneity, as the Likert scale had the next highest percent variability in item standard errors ($p_{Likert}$ = `r apa_num(mean_var["Likert"], gt1 = FALSE)`, followed by milliseconds ($p_{Milliseconds}$ = `r apa_num(mean_var["Milliseconds"], gt1 = FALSE)`.
+
+### Skew
+
+```{r scale-skew-figure, echo = F, fig.cap="Simulated pilot sample size compared to projected sample size for ceiling versus normal distributions on each scale. " }
+ggplot(summary_long %>% 
+         group_by(original_n, power, data_type) %>% 
+         summarize(sample_size_avg = mean(sample_size)),
+       aes(original_n, sample_size_avg, color = data_type)) + 
+  geom_point() + 
+  geom_point(aes(original_n, original_n), color = "black") + 
+  geom_line() + 
+  theme_linedraw(base_size = 15) + 
+  xlab("Pilot Sample Size") + 
+  ylab("Projected Sample Size") + 
+  scale_color_discrete(name = "Scale Skew") +
+  facet_wrap(~power)
+```
+
+Figure \@ref(fig:scale-skew-figure) displays that ceiling distributions, averaged over all other variables, show slightly higher estimates than normal distributions.
+This result is consistent across scale type and heterogeneity, as results indicated that they are often the same or slightly higher for ceiling distributions.
+
+### Item Heterogeneity
+
+```{r scale-hetero-figure, echo = F, fig.cap="Add a good caption here."}
+ggplot(summary_long %>% 
+         group_by(original_n, power, variability) %>% 
+         summarize(sample_size_avg = mean(sample_size)),
+       aes(original_n, sample_size_avg, color = variability)) + 
+  geom_point() + 
+  geom_point(aes(original_n, original_n), color = "black") + 
+  geom_line() + 
+  theme_linedraw() + 
+  xlab("Pilot Sample Size") + 
+  ylab("Projected Sample Size") + 
+  scale_color_discrete(name = "Item Heterogeneity") +
+  facet_wrap(~power)
+```
+
+```{r strongest-predict}
+regression_DF <- summary_long %>% 
+  group_by(original_n, scale_size, variability, data_type) %>% 
+  summarize(sample_size = mean(sample_size),
+            sd_item = mean(sd_item),
+            percent_var = mean(percent_var))
+model.samplesize <- lm(sample_size ~ original_n + scale_size + percent_var + data_type,
+                       data = regression_DF)
+summary(model.samplesize, correlation = TRUE)
+```
+
+Figure \@ref(fig:scale-hetero-figure) displays the results for item heterogeneity for different levels of potential power.
+In this figure, we found that our suggested procedure does capture the differences in heterogeneity.
+As heterogeneity increases in item variances, the proposed sample size also increases.
+
+```{r table-predict-largest, results='asis'}
+temp_table <- broom::tidy(model.samplesize)
+temp_table$term <- c("Intercept", "Pilot Sample Size", "Scale: Likert v Percent", "Scale: Likert v Milllisecond", "Proportion Variability", "Data: Ceiling v Normal")
+temp_table$p.value <- apa_p(temp_table$p.value)
+temp_table$pr <- apa_p((temp_table$statistic / sqrt(temp_table$statistic^2 + model.samplesize$df.residual))^2)
+
+apa_table(
+  temp_table, 
+  col.names = c("Term", "Estimate", "$SE$", "$t$", "$p$", "$pr^2$"),
+  escape = FALSE
+)
+```
+
+Using a regression model, we predicted proposed sample size using pilot sample size, scale size, proportion variability (i.e., heterogeneity), and data type (normal, ceiling).
+As shown in Table \@ref(tab:table-predict-largest), the largest influence on proposed sample size is the original pilot sample size, followed by proportion of variance/heterogeneity, and then data and scale sizes.
+
+## Projected Sample Size Sensitivity to Pilot Sample Size
+
+In our second question, we examined if the suggested procedure was sensitive to the amount of information present in the pilot data.
+Larger pilot data is more informative, and therefore, we should expect a lower projected sample size.
+As shown in each figure presented already, we do not find this effect.
+These simulations from the pilot data would nearly always suggest a larger sample size - mostly in a linear trend increasing with sample sizes.
+This result comes from the nature of the procedure - if we base our estimates on a SE cut-off, we will almost always need a bit more people for items to meet those goals.
+This result does not achieve our second goal.
+
+Therefore, we suggest using a correction factor on the simulation procedure to account for the known asymptotic nature of power (i.e., at larger sample sizes power increases level off).
+For this function in our simulation study, we combined a correction factor for upward biasing of effect sizes (Hedges' correction) with the formula for exponential decay calculations.
+The decay factor was calculated as follows:
+
+$$ 1 - \sqrt{\frac{N_{Pilot} - min(N_{Simulation})}{N_{Pilot}}}^{log_2(N_{Pilot})}$$
+
+$N_{Pilot}$ indicates the sample size of the pilot data minus the minimum simulated sample size to ensure that the smallest sample sizes do not decay (i.e., the formula zeroes out).
+This value is raised to the power of $log_2$ of the sample size of the pilot data, which decreases the impact of the decay to smaller increments for increasing sample sizes.
+This value is then multiplied by the projected sample size.
+As shown in Figure \@ref(fig:corrected-figure), this correction factor produces the desired quality of maintaining that small pilot studies should *increase* sample size, and that sample size suggestions level off as pilot study data sample size increases.
+
+```{r corrected-figure, echo = F, fig.cap="A corrected figure update this caption."}
+decay <- 1-sqrt((summary_long$original_n-20)/summary_long$original_n)^log2(summary_long$original_n)
+summary_long$new_sample <- summary_long$sample_size*decay
+ggplot(summary_long %>% 
+         group_by(original_n, power, variability) %>% 
+         summarize(sample_size_avg = mean(sample_size),
+                   new_sample_size_avg = mean(new_sample)),
+       aes(original_n, new_sample_size_avg, color = variability)) + 
+  geom_point() + 
+  geom_point(aes(original_n, original_n), color = "black") + 
+  geom_line() + 
+  theme_linedraw() + 
+  xlab("Pilot Sample Size") + 
+  ylab("Projected Sample Size") + 
+  scale_color_discrete(name = "Item Heterogeneity") +
+  facet_wrap(~power)
+```
+
+## Corrections for Individual Researchers
+
+We have portrayed that this procedure, with a correction factor, can perform as desired.
+However, within real scenarios, researchers will only have one pilot sample, not the various simulated samples shown above.
+What should the researcher do to correct their projected sample size from their own pilot data simulations?
+
+To explore if we could recover the new projected sample size from data a researcher would have, we used regression models to create a formula for researcher correction.
+The researcher employing our procedure would have the possible following variables from their simulations on their (one) pilot dataset: 1) proposed sample size, 2) pilot sample size, 3) estimate of heterogeneity for the items, 4) and the estimate percent of items below the threshold.
+Given the non-linear nature of the correction, we added each variable and it's non-linear `log2` transform to the regression equation, as this function was used to create the correction.
+The intercept only model was used as a starting point (i.e., `corrected sample ~ 1`), and then all eight variables (each variable and their `log2` transform) were entered into a forward stepwise regression to capture the corrected scores with the most predictive values.
+Each variable was entered one at a time using the `step` function from the *stats* library in *R* [@R-base].
+
+```{r regressions, include = F, echo = F}
+intercept_only <- lm(new_sample ~ 1, data = summary_long)
+user_model3 <- lm(new_sample ~ sample_size + percent_var + original_n + power + log2(sample_size) + log2(percent_var) + log2(original_n) + log2(power), data = summary_long)
+
+forward <- step(intercept_only, direction='forward', scope=formula(user_model3), trace=0)
+
+model.summary <- tidy(forward)
+model.summary$AIC <- forward$anova$AIC
+```
+
+```{r test-assume, include = F, echo = F}
+resid_model <- scale(residuals(forward))
+fit_model <- scale(fitted.values(forward))
+hist(resid_model)
+qqnorm(resid_model); abline(0,1)
+plot(fit_model, resid_model); abline(v = 0); abline(h = 0)
+# definitely residuals are a mess, but this also includes all deciles
+```
+
+As shown in Table \@ref(tab:table-correction), all variables were included in the final equation, each contributing a significant change to the previous model, as defined by $\Delta$AIC \> 2 points change between each step of the model.
+Proposed sample size and original sample size were the largest predictors -- unsurprising given the correction formula employed -- followed by the percent "power" level and proportion of variance.
+This formula approximation captures $R^2$ = `r apa_print(forward)$estimate$modelfit$r2` of the variance in sample size scores and should allow a researcher to estimate based on their own data, `r apa_print(forward)$statistic$modelfit$r2`.
+We provide convenience functions in our additional materials to assist researchers in estimating the final adjusted sample size.
+
+```{r table-correction, results = 'asis'}
+table_correct <- model.summary
+table_correct$term <- c("Intercept", "Projected Sample Size", "Pilot Sample Size", 
+                        "Log2 Projected Sample Size", "Log2 Pilot Sample Size", 
+                        "Log2 Power", "Proportion Variability", 
+                        "Log2 Proportion Variability", "Power")
+table_correct$estimate <- printnum(table_correct$estimate, digits = 3)
+table_correct$std.error <- printnum(table_correct$std.error, digits = 3)
+table_correct$statistic <- printnum(table_correct$statistic, digits = 3)
+table_correct$p.value <- printnum(table_correct$p.value, digits = 3,
+                 zero = FALSE, gt1 = F)
+table_correct$AIC <- printnum(table_correct$AIC, digits = 2)
+apa_table(table_correct, 
+     col.names = c("Term", "Estimate", "$SE$", "$t$", "$p$", "AIC"),
+     caption = "Parameters for All Decile Cut-off Scores")
+```
+
+## Choosing an Appropriate Cut-off
+
+```{r R2-cutoff, include = F, echo = F}
+by_cutoff <- list()
+R2 <- list()
+for (cutoff in unique(summary_long$name)){
+ by_cutoff[[cutoff]] <- lm(new_sample ~ sample_size + percent_var + original_n + power + log2(sample_size) + log2(percent_var) + log2(original_n) + log2(power), data = summary_long %>% filter(name == cutoff))
+ R2[cutoff] <- summary(by_cutoff[[cutoff]])$r.squared
+}
+```
+
+```{r test-decile, include = F, echo = F}
+resid_model <- scale(residuals(by_cutoff[["Decile 40"]]))
+fit_model <- scale(fitted.values(by_cutoff[["Decile 40"]]))
+# hist(resid_model)
+# qqnorm(resid_model); abline(0,1)
+# plot(fit_model, resid_model); abline(v = 0); abline(h = 0)
+# by scale, power, source
+separate_lms <- list()
+R2_lms <- list()
+for (scale_loop in unique(summary_long$scale_size)){
+ for (power_loop in unique(summary_long$power)){
+  for (source_loop in unique(summary_long$variability)){
+   temp <- summary_long %>% 
+    filter(scale_size == scale_loop) %>% 
+    filter(power == power_loop) %>% 
+    filter(variability == source_loop) %>% 
+    filter(name == "Decile 50")
+
+   separate_lms[[paste(scale_loop, power_loop, source_loop, sep = "_")]] <- lm(new_sample ~ sample_size + sd_item + original_n, data = temp)
+   R2_lms[[paste(scale_loop, power_loop, source_loop, sep = "_")]] <- summary(separate_lms[[paste(scale_loop, power_loop, source_loop, sep = "_")]])$r.squared
+  }
+ }
+}
+# resid_model <- scale(residuals(separate_lms[[10]]))
+# fit_model <- scale(fitted.values(separate_lms[[10]]))
+# hist(resid_model)
+# qqnorm(resid_model); abline(0,1)
+# plot(fit_model, resid_model); abline(v = 0); abline(h = 0)
+```
+
+Last, we examine the question of an appropriate SE decile.
+First, the 0%, 10%, and 20% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale size, and heterogeneity.
+If we examine the $R^2$ values for each decile of our regression equation separately, we find that the values are all $R^2$ > .99 with very little differences between them.
+Figure \@ref(fig:decile-figure) illustrates the corrected scores for simulations at the the 40% and 50% decile recommended cut-off for item standard errors.
+For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 40% decile range, especially for scales with larger potential variance.
+Therefore, we would suggest the 40% decile to overpower each item to determine the minimum and maximum sample size.
+
+```{r decile-figure, fig.cap = "Comparison of the cutoffs for 40% and 50% deciles across heterogeneity (rows), powering of items (columns), and scale size (shape)."}
+summary_long %>% 
+  filter(name == "Decile 40" | name == "Decile 50") %>% 
+  group_by(variability, original_n, scale_size, name, power) %>% 
+  summarize(new_sample = mean(new_sample)) %>% 
+  ggplot(., aes(original_n, new_sample, color = name)) + 
+  geom_point(aes(shape = scale_size)) + 
+  geom_point(aes(original_n, original_n), color = "black") + 
+  theme_classic(base_size = 15) + 
+  facet_wrap(~ variability*power) + 
+  xlab("Pilot Sample Size") + 
+  ylab("Corrected Sample Size") + 
+  scale_shape_discrete(name = "Scale Size") +
+  scale_color_discrete(name = "Cut-off Decile") + 
+  theme(legend.position="bottom")
+```
+
+The final formula for 40% decile correction is provided in Table \@ref(tab:table-decile). Proportion of variance can be calculated with the following:
+
+$$\frac{SD_{Item SD}}{\sqrt{\frac{(Maximum - Minimum)^2}{4}}}$$
+where maximum and minimum are the max and min values found in the scale (or the data, if the scale is unbounded).
+This formula would be applied in Step 5 of the proposed procedure.
+While the estimated coefficients could change given variations on our simulation parameters, the general size and pattern of coefficients was consistent, and therefore, we believe this correction equation should work for a variety of use cases.
+We will now demonstrate the final procedure on the examples provided earlier.
+
+```{r table-decile, results = 'asis'}
+table_correct <- tidy(summary(by_cutoff$`Decile 40`))
+table_correct <- table_correct[c(1, 2, 4, 6, 8, 9, 3, 7, 5) , ]
+table_correct$term <- c("Intercept", "Projected Sample Size", "Pilot Sample Size", 
+                        "Log2 Projected Sample Size", "Log2 Pilot Sample Size", 
+                        "Log2 Power", "Proportion Variability", 
+                        "Log2 Proportion Variability", "Power")
+table_correct$estimate <- printnum(table_correct$estimate, digits = 3)
+table_correct$std.error <- printnum(table_correct$std.error, digits = 3)
+table_correct$statistic <- printnum(table_correct$statistic, digits = 3)
+table_correct$p.value <- printnum(table_correct$p.value, digits = 3,
+                 zero = FALSE, gt1 = F)
+apa_table(table_correct, 
+     col.names = c("Term", "Estimate", "$SE$", "$t$", "$p$"),
+     caption = "Parameters for 40% Decile Cut-off Scores")
+```
+
+# Updated Example
+
+```{r table-scores-updated, results = 'asis'}
+c_sample <- c(results_table_example$`C 80`[4], 
+              results_table_example$`C 85`[4],
+              results_table_example$`C 90`[4],
+              results_table_example$`C 95`[4])
+
+e_sample <- c(results_table_example$`L 80`[4], 
+              results_table_example$`L 85`[4],
+              results_table_example$`L 90`[4],
+              results_table_example$`L 95`[4])
+
+con_prop <- sd_items_c / sqrt(((max(con_use$Rating) - min(con_use$Rating))^2)/4)
+ldt_prop <- sd_items_e / sqrt(((max(elp_use$RT) - min(elp_use$RT))^2)/4)
+
+table_correction_applied <- matrix(NA, ncol = 10, nrow = 9)
+table_correction_applied[1 , ] <- c("Formula", as.numeric(table_correct$estimate))
+table_correction_applied[2 , ] <- c("Concrete 80", 1, c_sample[1], 
+                                    pilot_size_c, log2(as.numeric(c_sample[1])), 
+                                    log2(pilot_size_c), log2(80), con_prop, 
+                                    log2(con_prop), 80)
+
+table_correction_applied[3 , ] <- c("Concrete 85", 1, c_sample[2], 
+                                    pilot_size_c, log2(as.numeric(c_sample[2])), 
+                                    log2(pilot_size_c), log2(85), con_prop, 
+                                    log2(con_prop), 85)
+
+table_correction_applied[4 , ] <- c("Concrete 90", 1, c_sample[3], 
+                                    pilot_size_c, log2(as.numeric(c_sample[3])), 
+                                    log2(pilot_size_c), log2(90), con_prop, 
+                                    log2(con_prop), 90)
+
+table_correction_applied[5 , ] <- c("Concrete 95", 1, c_sample[4], 
+                                    pilot_size_c, log2(as.numeric(c_sample[4])), 
+                                    log2(pilot_size_c), log2(95), con_prop, 
+                                    log2(con_prop), 95)
+
+table_correction_applied[6 , ] <- c("LDT 80", 1, e_sample[1], 
+                                    pilot_size_e, log2(as.numeric(e_sample[1])), 
+                                    log2(pilot_size_e), log2(80), ldt_prop, 
+                                    log2(ldt_prop), 80)
+
+table_correction_applied[7 , ] <- c("LDT 85", 1, e_sample[2], 
+                                    pilot_size_e, log2(as.numeric(e_sample[2])), 
+                                    log2(pilot_size_e), log2(85), ldt_prop, 
+                                    log2(ldt_prop), 85)
+
+table_correction_applied[8 , ] <- c("LDT 90", 1, e_sample[3], 
+                                    pilot_size_e, log2(as.numeric(e_sample[3])), 
+                                    log2(pilot_size_e), log2(90), ldt_prop, 
+                                    log2(ldt_prop), 90)
+
+table_correction_applied[9 , ] <- c("LDT 95", 1, e_sample[4], 
+                                    pilot_size_e, log2(as.numeric(e_sample[4])), 
+                                    log2(pilot_size_e), log2(95), ldt_prop, 
+                                    log2(ldt_prop), 95)
+
+table_correction_applied <- as.data.frame(table_correction_applied) 
+colnames(table_correction_applied) <- c("Formula Estimate", "Intercept", 
+                                        "Projected Sample Size", "Pilot Sample Size", 
+                                        "Log2 Projected Sample Size", "Log2 Pilot Sample Size", 
+                                        "Log2 Power", "Proportion Variability", 
+                                        "Log2 Proportion Variability", "Power")
+
+table_correction_applied <- table_correction_applied %>% 
+  mutate_at(.vars = 2:10, as.numeric) %>% 
+  mutate(`Corrected Sample` = NA)
+
+for (i in 2:nrow(table_correction_applied)){
+  table_correction_applied$`Corrected Sample`[i] <- 
+    sum(table_correction_applied[1, -c(1, ncol(table_correction_applied))] * table_correction_applied[i, -c(1, ncol(table_correction_applied))] )
+}
+
+table_correction_applied$`Data Loss` <- NA
+table_correction_applied$`Data Loss`[2:5] <- unlist(table_correction_applied$`Corrected Sample`[2:5]) * (1/mean(data_loss_c$percent_correct))
+
+table_correction_applied$`Data Loss`[6:9] <- unlist(table_correction_applied$`Corrected Sample`[6:9]) * (1/mean(data_loss_e$percent_correct))
+
+apa_table(table_correction_applied, 
+          caption = "Applied Correction for Each Proposed Sample Size")
+```
+
+The updated proposal steps are in Table \@ref(tab:table-summary).
+The main change occurs in Step 2 with a designated cut-off decile, and Step 5 with a correction score.
+Using the data from the 40% decile in Table \@ref(tab:table-example), we can determine that the stopping rule for concreteness ratings would be `r apa_num(cutoff_c[4])`, and the stopping rule for lexical decision times would be `r apa_num(cutoff_e[4])`.
+For Step 5, we apply our correction formula separately for each one, as they have different variability scores, and these scores are shown in Table \@ref(tab:table-scores-updated).
+Each row was multipled by row one's formula, and then these scores are summed for the final sample size.
+Sample sizes cannot be proportional, so we recommend rounding up to the nearest whole number.
+
+The concreteness ratings data includes `r length(unique(conDF$Word))` concepts that were rated for their concreteness.
+In our fictional study for this example, we selected 100 random words to show participants.
+In the original study, not every participant rated every word, which created uneven sample sizes for each word.
+In our random sample of 100 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size_c` as our pilot sample size for this example.
+All "do not know" ratings were set as missing data.
+The 50% decile for items standard error was `r round(cutoff, digits = 3)` for our cut-off criterion.
+
+For one additional consideration, we calculated the potential amount of data retention given that participants could indicate they did not know a word ($M_{correct}$ = `r round(mean(data_loss$percent_correct), digits = 2)`, *SD* = `r round(sd(data_loss$percent_correct), digits = 2)`).
+In order to account for this facet, the potential sample sizes were multiplied by 1/`r round(mean(data_loss$percent_correct), digits = 2)`,
+
+# Additional Materials
+
+## Package
+
+`simulate_population()`
+
+simulate_population(mu = 4, mu_sigma = .2, sigma = 2, sigma_sigma = .2, number_items = 30, number_scores = 1000, smallest_sigma = .02, min_score = 1, max_score = 7, digits = 0)
+
+## Vignettes
 

### 2023-08-30 simulation_study/simulate_summary_ceiling.Rmd
simulations done!

---
title: "Methods Power Study - Ceiling Simulations"
author: "Erin M. Buchanan"
date: "Last Knitted: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
set.seed(548902) # 233232
library(dplyr)
library(ggplot2)
library(tidyr)
library(broom)
library(papaja)
library(rio)
```

This document includes the informal write up for the method for the manuscript. We exported the simulations to this document in order to run the simulations separately from the manuscript for speed reasons (i.e., it's really slow to run the simulations along with the manuscript markdown). At the end of this document, a `simulated_summary_data.csv` file is created that is imported into the manuscript for plotting and other analyses. Please see the manuscript for formal write up.

# Method

## Data Simulation

*Population*.

- Simulate the data with `rnorm` assuming a normal distribution for 30 scale items
- Population is created with 1000 data points 
- We will run this part 100 times, so create a function that does this for you.

```{r sim_pop, include = T, echo = T}
simulate_population <- function (x = 1){

  # small potential variability overall, sort of 1-7ish scale
  mu.small <- rnorm(30, 6, .25)
  sigma.small.s <- rnorm(30, 2, .2)
  sigma.small.m <- rnorm(30, 2, .4)
  sigma.small.l <- rnorm(30, 2, .8)

  while(sum(sigma.small.s < .02) > 0){
    sigma.small.s <- rnorm(30, 2, .2)
  }
  while(sum(sigma.small.m < .04) > 0){
    sigma.small.m <- rnorm(30, 2, .4)
  }
  while(sum(sigma.small.l < .08) > 0){
    sigma.small.l <- rnorm(30, 2, .8)
  }

  # medium potential variability 0 to 100 scale
  mu.medium <- rnorm(30, 85, 10)
  sigma.medium.s <- rnorm(30, 25, 4)
  sigma.medium.m <- rnorm(30, 25, 8)
  sigma.medium.l <- rnorm(30, 25, 16)

  while(sum(sigma.medium.s < 0.4) > 0){
    sigma.medium.s <- rnorm(30, 25, 4)
  }
  while(sum(sigma.medium.m < 0.8) > 0){
    sigma.medium.m <- rnorm(30, 25, 8)
  }
  while(sum(sigma.medium.l < 1.6) > 0){
    sigma.medium.l <- rnorm(30, 25, 16)
  }

  # large potential variability in the 1000s scale
  mu.large <- rnorm(30, 2500, 150)
  sigma.large.s <- rnorm(30, 400, 50)
  sigma.large.m <- rnorm(30, 400, 100)
  sigma.large.l <- rnorm(30, 400, 200)

  while(sum(sigma.large.s < 5) > 0){
    sigma.large.s <- rnorm(30, 400, 50)
  }
  while(sum(sigma.large.m < 10) > 0){
    sigma.large.m <- rnorm(30, 400, 100)
  }
  while(sum(sigma.large.l < 20) > 0){
    sigma.large.l <- rnorm(30, 400, 200)
  }

  population.small <- data.frame(
    item = rep(1:30, 1000*3),
    variability = rep(c("small", "medium", "large"), each = 1000*30),
    score = c(round(rnorm(1000*30, mean = mu.small, sd = sigma.small.s), digits = 0),
              round(rnorm(1000*30, mean = mu.small, sd = sigma.small.m), digits = 0),
              round(rnorm(1000*30, mean = mu.small, sd = sigma.small.l), digits = 0))
    )

  population.medium <- data.frame(
    item = rep(1:30, 1000*3),
    variability = rep(c("small", "medium", "large"), each = 1000*30),
    score = c(round(rnorm(1000*30, mean = mu.medium, sd = sigma.medium.s), digits = 0),
              round(rnorm(1000*30, mean = mu.medium, sd = sigma.medium.m), digits = 0),
              round(rnorm(1000*30, mean = mu.medium, sd = sigma.medium.l), digits = 0))
    )

  population.large <- data.frame(
    item = rep(1:30, 1000*3),
    variability = rep(c("small", "medium", "large"), each = 1000*30),
    score = c(round(rnorm(1000*30, mean = mu.large, sd = sigma.large.s), digits = 0),
              round(rnorm(1000*30, mean = mu.large, sd = sigma.large.m), digits = 0),
              round(rnorm(1000*30, mean = mu.large, sd = sigma.large.l), digits = 0))
    )

  # truncate scores
  population.small$score <- ifelse(
    population.small$score > 7, 7, ifelse(
      population.small$score < 1, 1, population.small$score
    )
  )

  population.medium$score <- ifelse(
    population.medium$score > 100, 100, ifelse(
      population.medium$score < 0, 0, population.medium$score
    )
  )

  population.large$score <- ifelse(
    population.large$score > 3000, 3000, ifelse(
      population.large$score < 0, 0, population.large$score
    )
  )

  # check means
  tapply(population.small$score, population.small$variability, mean)
  # check var on var
  apply(tapply(population.small$score, list(population.small$item, population.small$variability), sd), 2, sd)

  tapply(population.medium$score, population.medium$variability, mean)
  # check var on var
  apply(tapply(population.medium$score, list(population.medium$item, population.medium$variability), sd), 2, sd)

  tapply(population.large$score, population.large$variability, mean)
  # check var on var
  apply(tapply(population.large$score, list(population.large$item, population.large$variability), sd), 2, sd)

  # return populations 
  return(list(population.small = population.small, 
              population.medium = population.medium, 
              population.large = population.large))
}
```

- Within that function, create the scale size variable: 
  - Small scale data 1-7, mean 6, sd of mean .25
  - Medium scale data 0-100, mean 85, sd of mean 10
  - Large scale data no range, mean 2500, sd of mean 150
  - These choices are meant to mimic popular scale choices that have skewed effects 
- Additionally, create the scale heterogeneity variable:
  - Small scale heterogeneity, sd 2, sd of sd .2 .4 .8
  - Medium scale heterogeneity, sd 25, sd of sd 4, 8, 16
  - Large scale heterogeneity, sd 400, sd of sd 50, 100, 200

*Samples*.

- Create a function that does the samples that a researcher might do
- Includes a start and stop of the number of participants you might consider 
- Includes an increasing value between samples
- We will also run this part 100 times

```{r sim_sample, include = T, echo = T}
simulate_samples <- function(start = 20, stop = 100, increase = 5,
                             population.small, population.medium, population.large){

  # save those samples for small medium large 
  samples.small <- samples.medium <- samples.large <- list()

  # create the list of sizes 
  sizes <- seq(from = start, to = stop, by = increase)

  # loop over sizes and create those samples 
  for (i in 1:length(sizes)){
    samples.small[[i]] <- population.small %>% 
      group_by(item, variability) %>% 
      slice_sample(n = sizes[i])

    samples.medium[[i]] <- population.medium %>%
      group_by(item, variability) %>%
      slice_sample(n = sizes[i])

    samples.large[[i]] <- population.large %>%
      group_by(item, variability) %>%
      slice_sample(n = sizes[i])
  }

  # return your values
  return(list(samples.small = samples.small, 
              samples.medium = samples.medium, 
              samples.large = samples.large,
              sizes = sizes))

}
```

*Cutoff Score Criterions*.

- Calculate the SEs for the items at each decile 
- Save as a function for later

```{r calc_SE, include = T, echo = T}
calculate_deciles <- function(samples.small, samples.medium, samples.large){
  # calculate the SEs and the cutoff scores 
  SES.small <- SES.medium <- SES.large <- list()
  cutoffs.small <- cutoffs.medium <- cutoffs.large <- list()
  sd_items.small <- sd_items.medium <- sd_items.large <- list()

  # loop and calculate for each sample size 
  for (i in 1:length(samples.small)){

  sd_items.small[[i]] <- samples.small[[i]] %>% group_by(item, variability) %>% 
    summarize(sd = sd(score), .groups = "keep") %>% 
    ungroup() %>% group_by(variability) %>% summarize(sd_item = sd(sd))

  sd_items.medium[[i]] <- samples.medium[[i]] %>% group_by(item, variability) %>% 
    summarize(sd = sd(score), .groups = "keep") %>% 
    ungroup() %>% group_by(variability) %>% summarize(sd_item = sd(sd))

  sd_items.large[[i]] <- samples.large[[i]] %>% group_by(item, variability) %>% 
    summarize(sd = sd(score), .groups = "keep") %>% 
    ungroup() %>% group_by(variability) %>% summarize(sd_item = sd(sd))

  SES.small[[i]] <- tapply(samples.small[[i]]$score,
                     list(samples.small[[i]]$item,
                          samples.small[[i]]$variability),
                     function (x){ sd(x)/sqrt(length(x))})

  SES.medium[[i]] <- tapply(samples.medium[[i]]$score,
                   list(samples.medium[[i]]$item,
                        samples.medium[[i]]$variability),
                   function (x){ sd(x)/sqrt(length(x))})

  SES.large[[i]] <- tapply(samples.large[[i]]$score,
                 list(samples.medium[[i]]$item,
                      samples.medium[[i]]$variability),
                 function (x){ sd(x)/sqrt(length(x))})

  cutoffs.small[[i]] <- apply(as.data.frame(SES.small[[i]]), 2, 
                         quantile, 
                         probs = seq(0, .9, by = .1),
                         na.rm = T)

  cutoffs.medium[[i]] <- apply(as.data.frame(SES.medium[[i]]), 2, 
                         quantile, 
                         probs = seq(0, .9, by = .1),
                         na.rm = T)

  cutoffs.large[[i]] <- apply(as.data.frame(SES.large[[i]]), 2, 
                         quantile, 
                         probs = seq(0, .9, by = .1),
                         na.rm = T)

  }

  # return information
  return(list(SES.small = SES.small, SES.medium = SES.medium, SES.large = SES.large, 
              cutoffs.small = cutoffs.small, cutoffs.medium = cutoffs.medium, cutoffs.large = cutoffs.large, 
              sd_items.small = sd_items.small, sd_items.medium = sd_items.medium, sd_items.large = sd_items.large))

}
```

## Researcher Sample Simulation

- Here we are simulating the researcher side of the equation assuming they have a sample pulled from population (which we simulated above)
- We will simulate samples from 20 to 2000 ... mostly only to see the leveling out effect of the simulation procedure, as we likely believe that over 2000 is not reasonable for most researchers. 
- Calculate the SEs here, to see if it matches our cutoff 
- Save this as a function to run within our pipeline (100 populations and 100 samples)

```{r sim_sim, include = T, echo = T}
simulate_researcher <- function(samples.small, samples.medium, samples.large, 
                                start = 20, stop = 2000, 
                                increase = 5){

  # sequence of sample sizes to try
  samplesize_values <- seq(from = start, to = stop, by = increase)

  # place to store everything
  sampled_values.small <- sampled_values.medium <- sampled_values.large <- list()

  # loop over the samples
  for (i in 1:length(samples.small)){

    # create a blank table for us to save the values in 
    sim_table.small <- matrix(NA, 
                        nrow = length(samplesize_values), 
                        ncol = 30*3)

    # make it a data frame
    sim_table.small <- sim_table.medium <- sim_table.large <- as.data.frame(sim_table.small)

    # add a place for sample size values 
    sim_table.small$sample_size <- sim_table.medium$sample_size <- sim_table.large$sample_size <- NA

    # loop over pilot sample sizes
    for (q in 1:length(samplesize_values)){

      # temp dataframe that samples and summarizes
      temp <- samples.small[[i]] %>% 
        group_by(item, variability) %>% 
        slice_sample(n = samplesize_values[q], replace = T) %>% 
        summarize(se = sd(score)/sqrt(length(score)),
                  .groups = "keep")

      sim_table.small[q, 1:90] <- temp$se
      sim_table.small[q, 91] <- samplesize_values[q]

      temp <- samples.medium[[i]] %>% 
        group_by(item, variability) %>% 
        slice_sample(n = samplesize_values[q], replace = T) %>% 
        summarize(se = sd(score)/sqrt(length(score)),
                  .groups = "keep")

      sim_table.medium[q, 1:90] <- temp$se
      sim_table.medium[q, 91] <- samplesize_values[q]

      temp <- samples.large[[i]] %>% 
        group_by(item, variability) %>% 
        slice_sample(n = samplesize_values[q], replace = T) %>% 
        summarize(se = sd(score)/sqrt(length(score)),
                  .groups = "keep")

      sim_table.large[q, 1:90] <- temp$se
      sim_table.large[q, 91] <- samplesize_values[q]

      } # end pilot sample loop

    sampled_values.small[[i]] <- sim_table.small
    sampled_values.medium[[i]] <- sim_table.medium
    sampled_values.large[[i]] <- sim_table.large

  } # end all sample loop

  # return sampled values 
  return(list(sampled_values.small = sampled_values.small, 
              sampled_values.medium = sampled_values.medium, 
              sampled_values.large = sampled_values.large,
              samplesize_values = samplesize_values))

}
```

- Next, calculate the percent of items falling below the decile scores 
- Pick up 80, 85, 90, and 95% of items to mimic power 
- Save a function for our pipeline

```{r calc_percent, include = T, echo = T}
calculate_percent <- function(sampled_values.small, sampled_values.medium, sampled_values.large,
                              cutoffs.small, cutoffs.medium, cutoffs.large, 
                              samplesize_values, sizes){

  # create temporary storage
  summary_list.small <- summary_list.medium <- summary_list.large <- list()

  # loop and calculate
  for (i in 1:length(sampled_values.small)){

    # summary list 1 ----
    summary_list.small[[i]] <- sampled_values.small[[i]] %>% 
      pivot_longer(cols = -c(sample_size)) %>% 
      rename(item = name, se = value) %>% 
      mutate(variability = rep(c("large", "medium", "small"), 
                               30*length(samplesize_values))) %>% 
      mutate(item = rep(rep(1:30, each = 3), length(samplesize_values)))

    # cut offs for 1
    temp.small.s <- summary_list.small[[i]] %>% 
      filter(variability == "small") %>% 
      group_by(sample_size, variability) %>% 
      summarize(Percent_Below0 = sum(se <= cutoffs.small[[i]]["0%", "small"])/30,
             Percent_Below10 = sum(se <= cutoffs.small[[i]]["10%", "small"])/30,
             Percent_Below20 = sum(se <= cutoffs.small[[i]]["20%", "small"])/30,
             Percent_Below30 = sum(se <= cutoffs.small[[i]]["30%", "small"])/30,
             Percent_Below40 = sum(se <= cutoffs.small[[i]]["40%", "small"])/30,
             Percent_Below50 = sum(se <= cutoffs.small[[i]]["50%", "small"])/30, 
             Percent_Below60 = sum(se <= cutoffs.small[[i]]["60%", "small"])/30, 
             Percent_Below70 = sum(se <= cutoffs.small[[i]]["70%", "small"])/30, 
             Percent_Below80 = sum(se <= cutoffs.small[[i]]["80%", "small"])/30, 
             Percent_Below90 = sum(se <= cutoffs.small[[i]]["90%", "small"])/30, 
             .groups = "keep") %>% 
      mutate(original_n = sizes[i], 
             scale_size = "likert")

    temp.small.m <- summary_list.small[[i]] %>% 
      filter(variability == "medium") %>% 
      group_by(sample_size, variability) %>% 
      summarize(Percent_Below0 = sum(se <= cutoffs.small[[i]]["0%", "medium"])/30,
             Percent_Below10 = sum(se <= cutoffs.small[[i]]["10%", "medium"])/30,
             Percent_Below20 = sum(se <= cutoffs.small[[i]]["20%", "medium"])/30,
             Percent_Below30 = sum(se <= cutoffs.small[[i]]["30%", "medium"])/30,
             Percent_Below40 = sum(se <= cutoffs.small[[i]]["40%", "medium"])/30,
             Percent_Below50 = sum(se <= cutoffs.small[[i]]["50%", "medium"])/30, 
             Percent_Below60 = sum(se <= cutoffs.small[[i]]["60%", "medium"])/30, 
             Percent_Below70 = sum(se <= cutoffs.small[[i]]["70%", "medium"])/30, 
             Percent_Below80 = sum(se <= cutoffs.small[[i]]["80%", "medium"])/30, 
             Percent_Below90 = sum(se <= cutoffs.small[[i]]["90%", "medium"])/30, 
             .groups = "keep") %>% 
      mutate(original_n = sizes[i],
             scale_size = "likert")

    temp.small.l <- summary_list.small[[i]] %>% 
      filter(variability == "large") %>% 
      group_by(sample_size, variability) %>% 
      summarize(Percent_Below0 = sum(se <= cutoffs.small[[i]]["0%", "large"])/30,
             Percent_Below10 = sum(se <= cutoffs.small[[i]]["10%", "large"])/30,
             Percent_Below20 = sum(se <= cutoffs.small[[i]]["20%", "large"])/30,
             Percent_Below30 = sum(se <= cutoffs.small[[i]]["30%", "large"])/30,
             Percent_Below40 = sum(se <= cutoffs.small[[i]]["40%", "large"])/30,
             Percent_Below50 = sum(se <= cutoffs.small[[i]]["50%", "large"])/30, 
             Percent_Below60 = sum(se <= cutoffs.small[[i]]["60%", "large"])/30, 
             Percent_Below70 = sum(se <= cutoffs.small[[i]]["70%", "large"])/30, 
             Percent_Below80 = sum(se <= cutoffs.small[[i]]["80%", "large"])/30, 
             Percent_Below90 = sum(se <= cutoffs.small[[i]]["90%", "large"])/30, 
             .groups = "keep") %>% 
      mutate(original_n = sizes[i], 
             scale_size = "likert")

    #rejoin 
    summary_list.small[[i]] <- bind_rows(temp.small.s, temp.small.m, temp.small.l)

    # summary list 2 ----
    summary_list.medium[[i]] <- sampled_values.medium[[i]] %>% 
      pivot_longer(cols = -c(sample_size)) %>% 
      rename(item = name, se = value) %>% 
      mutate(variability = rep(c("large", "medium", "small"), 
                               30*length(samplesize_values))) %>% 
      mutate(item = rep(rep(1:30, each = 3), length(samplesize_values)))

    # cut offs for 2
    temp.medium.s <- summary_list.medium[[i]] %>% 
      filter(variability == "small") %>% 
      group_by(sample_size, variability) %>% 
      summarize(Percent_Below0 = sum(se <= cutoffs.medium[[i]]["0%", "small"])/30,
             Percent_Below10 = sum(se <= cutoffs.medium[[i]]["10%", "small"])/30,
             Percent_Below20 = sum(se <= cutoffs.medium[[i]]["20%", "small"])/30,
             Percent_Below30 = sum(se <= cutoffs.medium[[i]]["30%", "small"])/30,
             Percent_Below40 = sum(se <= cutoffs.medium[[i]]["40%", "small"])/30,
             Percent_Below50 = sum(se <= cutoffs.medium[[i]]["50%", "small"])/30, 
             Percent_Below60 = sum(se <= cutoffs.medium[[i]]["60%", "small"])/30, 
             Percent_Below70 = sum(se <= cutoffs.medium[[i]]["70%", "small"])/30, 
             Percent_Below80 = sum(se <= cutoffs.medium[[i]]["80%", "small"])/30, 
             Percent_Below90 = sum(se <= cutoffs.medium[[i]]["90%", "small"])/30, 
             .groups = "keep") %>% 
      mutate(original_n = sizes[i], 
             scale_size = "percent")

    temp.medium.m <- summary_list.medium[[i]] %>% 
      filter(variability == "medium") %>% 
      group_by(sample_size, variability) %>% 
      summarize(Percent_Below0 = sum(se <= cutoffs.medium[[i]]["0%", "medium"])/30,
             Percent_Below10 = sum(se <= cutoffs.medium[[i]]["10%", "medium"])/30,
             Percent_Below20 = sum(se <= cutoffs.medium[[i]]["20%", "medium"])/30,
             Percent_Below30 = sum(se <= cutoffs.medium[[i]]["30%", "medium"])/30,
             Percent_Below40 = sum(se <= cutoffs.medium[[i]]["40%", "medium"])/30,
             Percent_Below50 = sum(se <= cutoffs.medium[[i]]["50%", "medium"])/30, 
             Percent_Below60 = sum(se <= cutoffs.medium[[i]]["60%", "medium"])/30, 
             Percent_Below70 = sum(se <= cutoffs.medium[[i]]["70%", "medium"])/30, 
             Percent_Below80 = sum(se <= cutoffs.medium[[i]]["80%", "medium"])/30, 
             Percent_Below90 = sum(se <= cutoffs.medium[[i]]["90%", "medium"])/30, 
             .groups = "keep") %>% 
      mutate(original_n = sizes[i],
             scale_size = "percent")

    temp.medium.l <- summary_list.medium[[i]] %>% 
      filter(variability == "large") %>% 
      group_by(sample_size, variability) %>% 
      summarize(Percent_Below0 = sum(se <= cutoffs.medium[[i]]["0%", "large"])/30,
             Percent_Below10 = sum(se <= cutoffs.medium[[i]]["10%", "large"])/30,
             Percent_Below20 = sum(se <= cutoffs.medium[[i]]["20%", "large"])/30,
             Percent_Below30 = sum(se <= cutoffs.medium[[i]]["30%", "large"])/30,
             Percent_Below40 = sum(se <= cutoffs.medium[[i]]["40%", "large"])/30,
             Percent_Below50 = sum(se <= cutoffs.medium[[i]]["50%", "large"])/30, 
             Percent_Below60 = sum(se <= cutoffs.medium[[i]]["60%", "large"])/30, 
             Percent_Below70 = sum(se <= cutoffs.medium[[i]]["70%", "large"])/30, 
             Percent_Below80 = sum(se <= cutoffs.medium[[i]]["80%", "large"])/30, 
             Percent_Below90 = sum(se <= cutoffs.medium[[i]]["90%", "large"])/30, 
             .groups = "keep") %>% 
      mutate(original_n = sizes[i], 
             scale_size = "percent")

    #rejoin 
    summary_list.medium[[i]] <- bind_rows(temp.medium.s, temp.medium.m, temp.medium.l)

    # summary list 3 ----
    summary_list.large[[i]] <- sampled_values.large[[i]] %>% 
      pivot_longer(cols = -c(sample_size)) %>% 
      rename(item = name, se = value) %>% 
      mutate(variability = rep(c("large", "medium", "small"), 
                               30*length(samplesize_values))) %>% 
      mutate(item = rep(rep(1:30, each = 3), length(samplesize_values)))

    # cut offs for 3 
    temp.large.s <- summary_list.large[[i]] %>% 
      filter(variability == "small") %>% 
      group_by(sample_size, variability) %>% 
      summarize(Percent_Below0 = sum(se <= cutoffs.large[[i]]["0%", "small"])/30,
             Percent_Below10 = sum(se <= cutoffs.large[[i]]["10%", "small"])/30,
             Percent_Below20 = sum(se <= cutoffs.large[[i]]["20%", "small"])/30,
             Percent_Below30 = sum(se <= cutoffs.large[[i]]["30%", "small"])/30,
             Percent_Below40 = sum(se <= cutoffs.large[[i]]["40%", "small"])/30,
             Percent_Below50 = sum(se <= cutoffs.large[[i]]["50%", "small"])/30, 
             Percent_Below60 = sum(se <= cutoffs.large[[i]]["60%", "small"])/30, 
             Percent_Below70 = sum(se <= cutoffs.large[[i]]["70%", "small"])/30, 
             Percent_Below80 = sum(se <= cutoffs.large[[i]]["80%", "small"])/30, 
             Percent_Below90 = sum(se <= cutoffs.large[[i]]["90%", "small"])/30, 
             .groups = "keep") %>% 
      mutate(original_n = sizes[i], 
             scale_size = "milliseconds")

    temp.large.m <- summary_list.large[[i]] %>% 
      filter(variability == "medium") %>% 
      group_by(sample_size, variability) %>% 
      summarize(Percent_Below0 = sum(se <= cutoffs.large[[i]]["0%", "medium"])/30,
             Percent_Below10 = sum(se <= cutoffs.large[[i]]["10%", "medium"])/30,
             Percent_Below20 = sum(se <= cutoffs.large[[i]]["20%", "medium"])/30,
             Percent_Below30 = sum(se <= cutoffs.large[[i]]["30%", "medium"])/30,
             Percent_Below40 = sum(se <= cutoffs.large[[i]]["40%", "medium"])/30,
             Percent_Below50 = sum(se <= cutoffs.large[[i]]["50%", "medium"])/30, 
             Percent_Below60 = sum(se <= cutoffs.large[[i]]["60%", "medium"])/30, 
             Percent_Below70 = sum(se <= cutoffs.large[[i]]["70%", "medium"])/30, 
             Percent_Below80 = sum(se <= cutoffs.large[[i]]["80%", "medium"])/30, 
             Percent_Below90 = sum(se <= cutoffs.large[[i]]["90%", "medium"])/30, 
             .groups = "keep") %>% 
      mutate(original_n = sizes[i],
             scale_size = "milliseconds")

    temp.large.l <- summary_list.large[[i]] %>% 
      filter(variability == "large") %>% 
      group_by(sample_size, variability) %>% 
      summarize(Percent_Below0 = sum(se <= cutoffs.large[[i]]["0%", "large"])/30,
             Percent_Below10 = sum(se <= cutoffs.large[[i]]["10%", "large"])/30,
             Percent_Below20 = sum(se <= cutoffs.large[[i]]["20%", "large"])/30,
             Percent_Below30 = sum(se <= cutoffs.large[[i]]["30%", "large"])/30,
             Percent_Below40 = sum(se <= cutoffs.large[[i]]["40%", "large"])/30,
             Percent_Below50 = sum(se <= cutoffs.large[[i]]["50%", "large"])/30, 
             Percent_Below60 = sum(se <= cutoffs.large[[i]]["60%", "large"])/30, 
             Percent_Below70 = sum(se <= cutoffs.large[[i]]["70%", "large"])/30, 
             Percent_Below80 = sum(se <= cutoffs.large[[i]]["80%", "large"])/30, 
             Percent_Below90 = sum(se <= cutoffs.large[[i]]["90%", "large"])/30, 
             .groups = "keep") %>% 
      mutate(original_n = sizes[i], 
             scale_size = "milliseconds")

    #rejoin 
    summary_list.large[[i]] <- bind_rows(temp.large.s, temp.large.m, temp.large.l)

    } # end loop and calculate

  # create end summary 
  summary_DF <- bind_rows(summary_list.small, 
                        summary_list.medium, 
                        summary_list.large)

  # return values 
  return(summary_DF)

}
```

- Last combine everything together grabbing the proposed sample at each decile given the 80, 85, 90, and 95 scores 
- Create a function for our pipeline

```{r calc_percent_combine, include = T, echo = T}
calculate_proposed <- function(summary_DF){

  # 80% summary 
  summary_long_80 <- summary_DF %>% 
    pivot_longer(cols = -c(sample_size, original_n, scale_size, variability)) %>% 
    filter(value >= .80) %>% 
    arrange(sample_size, original_n, scale_size, variability, name) %>% 
    group_by(original_n, name, scale_size, variability) %>% 
    slice_head(n = 1) %>% 
    mutate(power = 80)

  # 85% summary
  summary_long_85 <- summary_DF %>% 
    pivot_longer(cols = -c(sample_size, original_n, scale_size, variability)) %>% 
    filter(value >= .85) %>% 
    arrange(sample_size, original_n, scale_size, variability, name) %>% 
    group_by(original_n, name, scale_size, variability) %>% 
    slice_head(n = 1) %>% 
    mutate(power = 85)

  # 90% summary  
  summary_long_90 <- summary_DF %>% 
    pivot_longer(cols = -c(sample_size, original_n, scale_size, variability)) %>% 
    filter(value >= .90) %>% 
    arrange(sample_size, original_n, scale_size, variability, name) %>% 
    group_by(original_n, name, scale_size, variability) %>% 
    slice_head(n = 1) %>% 
    mutate(power = 90)

  # 95% summary
  summary_long_95 <- summary_DF %>% 
    pivot_longer(cols = -c(sample_size, original_n, scale_size, variability)) %>% 
    filter(value >= .95) %>% 
    arrange(sample_size, original_n, scale_size, variability, name) %>% 
    group_by(original_n, name, scale_size, variability) %>% 
    slice_head(n = 1) %>% 
    mutate(power = 95)

  # combine the summary together
  summary_long <- rbind(summary_long_80, 
                        summary_long_85,
                        summary_long_90,
                        summary_long_95)

  # return values
  return(summary_long)

}
```

## Simulation Pipeline

- Put together all the simulation pieces 
  - Simulate the population 
  - Simulate samples within that population 
  - Calculate the cutoffs and deciles 
  - Simulate like the researcher would 
  - Calculate the percent under the decile
  - Create final summary

- We will simulate multiple times at each simulation point:
  - Samples X 100
  - Simulate Researcher X 100

```{r simulation_pipeline, include = T, echo = T, eval = F}
# a list to save this in 
full_simulation <- list()
populations <- simulate_population(x)

for (p in 1:100){

  # simulate samples from population what pilot data they may have
  samples <- simulate_samples(start = 20, stop = 100, increase = 10, 
                          population.small = populations$population.small,
                          population.medium = populations$population.medium, 
                          population.large = populations$population.large)

  deciles <- calculate_deciles(samples.small = samples$samples.small, 
                           samples.medium = samples$samples.medium,
                           samples.large = samples$samples.large)

  cat(paste(p, Sys.time(), "\n"))

  # reset this with each new sample 
  researcher_summary <- list()

  for (q in 1:100){

    # simulate researcher but work in big steps because this is slow
    researcher_data <- simulate_researcher(samples.small = samples$samples.small, 
                                     samples.medium = samples$samples.medium,
                                     samples.large = samples$samples.large, 
                                     start = 20, stop = 2000, 
                                     increase = 20)

    # calculate percent under each decile
    percent_items <- calculate_percent(
      sampled_values.small = researcher_data$sampled_values.small, 
      sampled_values.medium = researcher_data$sampled_values.medium, 
      sampled_values.large = researcher_data$sampled_values.large, 
      cutoffs.small = deciles$cutoffs.small, 
      cutoffs.medium = deciles$cutoffs.medium, 
      cutoffs.large = deciles$cutoffs.large, 
      samplesize_values = researcher_data$samplesize_values,
      sizes = samples$sizes)

    # calculate final summary data 
    researcher_summary[[q]] <- calculate_proposed(summary_DF = percent_items)

  } # researcher 100 simulations

  # after the researcher summary, summarize and stick in dataframe 
  # so we have a dataframe of each population by sample 
  full_simulation[[p]] <-
   bind_rows(researcher_summary) %>%
   group_by(variability, original_n, scale_size, name, power) %>%
   summarize(sample_size = mean(sample_size),
             value = mean(value), .groups = "keep")

  # clean up the values from the summary while you have the data 
  full_simulation[[p]]$scale_size <- factor(full_simulation[[p]]$scale_size, 
                                levels = c("likert", "percent", "milliseconds"),
                                labels = c("Likert", 
                                           "Percent", 
                                           "Milliseconds"))

  full_simulation[[p]]$variability <- factor(full_simulation[[p]]$variability, 
                                levels = c("small", "medium", "large"),
                                labels = c("Small Heterogeneity", 
                                           "Medium Heterogeneity", 
                                           "Large Heterogeneity"))

  # grab the sd information from the pilot sample
  # note this is not the researcher simulated sample 
  sd_items <- bind_rows(deciles$sd_items.small, deciles$sd_items.medium, deciles$sd_items.large)
  sd_items$scale_size <- rep(c("Likert", 
                           "Percent", 
                           "Milliseconds"), 
                         each = 3*length(samples$sizes))
  sd_items$variability <- factor(sd_items$variability, 
                                levels = c("small", "medium", "large"),
                                labels = c("Small Heterogeneity", 
                                           "Medium Heterogeneity", 
                                           "Large Heterogeneity"))
  sd_items$original_n <- rep(rep(samples$sizes, each = 3), 3)

  # add together that information 
  full_simulation[[p]] <- full_simulation[[p]] %>% 
    full_join(sd_items, 
              by = c("original_n" = "original_n", 
                     "variability" = "variability", 
                     "scale_size" = "scale_size"))

  full_simulation[[p]]$name <- gsub("Percent_Below", "Decile ", full_simulation[[p]]$name)

  # writing out takes time but don't want to lose 100 researcher sims
  # ran this across computers so make clear when run 
  saveRDS(full_simulation, file = paste("simulation_", Sys.time(), ".Rdata", sep = ""))

} # simulate sampling 100 times from the population
```

## Put Together Data

```{r eval = F}
all_sim <- bind_rows(readRDS("simulation_study/simulation_ceiling/simulation_2023-08-30_rstudio_ceiling.Rdata"),
                     bind_rows(readRDS("simulation_study/simulation_ceiling/simulation_2023-08-30_windows_ceiling.Rdata")),
                     bind_rows(readRDS("simulation_study/simulation_ceiling/simulation_2023-08-30_server_ceiling.Rdata"))) %>% unique()

# View(table(all_sim$original_n, all_sim$variability, all_sim$scale_size, all_sim$power, all_sim$name))

summary_long <- all_sim %>% 
  group_by(variability, original_n, scale_size, name, power) %>% 
  summarize(
    n = n(),
    sample_size = mean(sample_size),
    value = mean(value), 
    sd_item = mean(sd_item),
    .groups = "keep")

# 100
```

## Export Final Data

```{r}
export(summary_long, "simulation_study/simulation_ceiling/simulated_summary_data_ceiling.csv", row.names = F)
```

### 2023-08-30 simulation_study/simulate_summary_middle.Rmd
simulations done!

--- 2023-08-30 simulation_study/simulate_summary_ceiling.Rmd
@@
 ---
-title: "Methods Power Study - Ceiling Simulations"
+title: "Methods Power Study - Middle Simulations"
 author: "Erin M. Buchanan"
@@
 ```{r}
-set.seed(548902) # 233232
+set.seed(895893) # 233232
 library(dplyr)
@@
   # small potential variability overall, sort of 1-7ish scale
-  mu.small <- rnorm(30, 6, .25)
+  mu.small <- rnorm(30, 4, .25)
   sigma.small.s <- rnorm(30, 2, .2)
@@
   # medium potential variability 0 to 100 scale
-  mu.medium <- rnorm(30, 85, 10)
+  mu.medium <- rnorm(30, 50, 10)
   sigma.medium.s <- rnorm(30, 25, 4)
@@
   # large potential variability in the 1000s scale
-  mu.large <- rnorm(30, 2500, 150)
+  mu.large <- rnorm(30, 1000, 150)
   sigma.large.s <- rnorm(30, 400, 50)
@@
 - Within that function, create the scale size variable: 
-  - Small scale data 1-7, mean 6, sd of mean .25
-  - Medium scale data 0-100, mean 85, sd of mean 10
-  - Large scale data no range, mean 2500, sd of mean 150
-  - These choices are meant to mimic popular scale choices that have skewed effects 
+  - Small scale data 1-7, mean 4, sd of mean .25
+  - Medium scale data 0-100, mean 50, sd of mean 10
+  - Large scale data no range, mean 1000, sd of mean 150
+  - These choices are meant to mimic popular scale choices 
 - Additionally, create the scale heterogeneity variable:
@@
 ```{r eval = F}
-all_sim <- bind_rows(readRDS("simulation_study/simulation_ceiling/simulation_2023-08-30_rstudio_ceiling.Rdata"),
-                     bind_rows(readRDS("simulation_study/simulation_ceiling/simulation_2023-08-30_windows_ceiling.Rdata")),
-                     bind_rows(readRDS("simulation_study/simulation_ceiling/simulation_2023-08-30_server_ceiling.Rdata"))) %>% unique()
+all_sim <- bind_rows(readRDS("simulation_study/simulation_middle/simulation_2023-08-29_mac_middle.Rdata")) %>% unique()
 
@@
 ```{r}
-export(summary_long, "simulation_study/simulation_ceiling/simulated_summary_data_ceiling.csv", row.names = F)
+export(summary_long, "simulation_study/simulation_middle/simulated_summary_data_middle.csv", row.names = F)
 ```

### 2023-08-30 simulation_study/simulations_combined.Rmd
simulations done!

---
title: "Simulations Combined"
author: "Erin M. Buchanan"
date: "`r Sys.Date()`"
output: html_document
---

## Libraries

```{r}
library(rio)
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
library(papaja)
library(broom)
```

## Data

```{r import-data}
summary_long <- bind_rows(
  import(here("simulation_study/simulation_middle/simulated_summary_data_middle.csv")) %>% mutate(sim = "middle"), 
  import(here("simulation_study/simulation_ceiling/simulated_summary_data_ceiling.csv")) %>% mutate(sim = "ceiling"))
```

## Check for Zero Variance

```{r}
zero <- summary_long %>% 
  mutate(sd_item = round(sd_item, digits = 4)) %>% 
  filter(sd_item < 0.05) %>% 
  group_by(variability, original_n, scale_size, sim) %>% 
  summarize(count = n() / (10*4)) # 10 deciles by 4 powers

zero
```

## Check the Variances

```{r}
rest <- summary_long %>% 
  filter(sd_item > .05) %>% 
  group_by(scale_size, original_n, variability, sim) %>% 
  summarize(variance = mean(sd_item))

rest

ggplot(rest %>% filter(scale_size == "Likert"), 
       aes(variability, variance, color = sim)) + 
  #facet_wrap(~scale_size) + 
  geom_jitter() + 
  geom_boxplot() + 
  theme_linedraw() + 
  xlab("Amount of Simulated Variance") + 
  ylab("Variance Found") + 
  scale_color_discrete(name = "Simulation Type", 
                       labels = c("Ceiling", "Average")) +
  ggtitle("Likert Scale")

ggplot(rest %>% filter(scale_size == "Percent"), 
       aes(variability, variance, color = sim)) + 
  #facet_wrap(~scale_size) + 
  geom_jitter() + 
  geom_boxplot() + 
  theme_linedraw() + 
  xlab("Amount of Simulated Variance") + 
  ylab("Variance Found") + 
  scale_color_discrete(name = "Simulation Type", 
                       labels = c("Ceiling", "Average")) + 
  ggtitle("Percent Scale")

ggplot(rest %>% filter(scale_size == "Milliseconds"), 
       aes(variability, variance, color = sim)) + 
  #facet_wrap(~scale_size) + 
  geom_jitter() + 
  geom_boxplot() + 
  theme_linedraw() + 
  xlab("Amount of Simulated Variance") + 
  ylab("Variance Found" )+ 
  scale_color_discrete(name = "Simulation Type", 
                       labels = c("Ceiling", "Average")) +
  ggtitle("Millisecond Scale")
```

### 2023-09-08 manuscript_draft.Rmd
all the updates

@@
 title       : "Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli"
-shorttitle    : "AIPE SIMULATION"
+shorttitle    : "AIPE BOOTSTRAPPING"
 author: 
@@
 authornote: |
- The authors would like to thank K.D. Valentine for her assistance in formulation of correction scores. 
+
 abstract: |
@@
 draft       : no
-documentclass   : "apa6"
 classoption    : "man"
-output      : papaja::apa6_docx
 bibliography: references.bib
+header-includes:
+  - |
+    \makeatletter
+    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+      {-1em}%
+      {\normalfont\normalsize\bfseries\typesectitle}}
+
+    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+      {-\z@\relax}%
+      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+    \makeatother
+
+csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
+documentclass     : "apa7"
+output            : papaja::apa6_pdf
 ---
 
-```{r setup, include = FALSE}
+```{r setup, include = FALSE, echo = FALSE}
 set.seed(895893)
@@
 One solution to this power analysis problem is the plethora of tools made available for researchers to make power computations accessible to non-statisticians with the caveat that a solid education in power is necessary to use these tools.
-G\*Power is one of the most popular free power software options [@faul2007; @erdfelder1996] that provides a simple point and click graphical user interface for power calculations [however, see @brysbaert].
+G\*Power is one of the most popular free power software options [@faul2007; @erdfelder1996] that provides a simple point and click graphical user interface for power calculations [however, see @brysbaert2019].
 Web-based tools have also sprung up for overall and statistical test specific sample size planning including <https://powerandsamplesize.com> and <https://designingexperiments.com> [@anderson2017].
@@
 
-2)  Calculate the standard error (SE) in the subset of pilot data associated with each item to create a cut-off score that defines when items are considered "accurately measured".
-Next, the simulations below will explore what criterion should be used to determine the cut-off score from the pilot data.
+2)  Calculate the standard error (SE) in the subset of pilot data associated with each item to create a cutoff score that defines when items are considered "accurately measured".
+Next, the simulations below will explore what criterion should be used to determine the cutoff score from the pilot data.
 
@@
 
-5)  Find the minimum sample size so that 80%, 85%, 90%, and 95% of the items meet the cut-off score and can be considered accurately measured.
+5)  Find the minimum sample size so that 80%, 85%, 90%, and 95% of the items meet the cutoff score and can be considered accurately measured.
 We recommend these scores to ensure that most items are accurately measured, in a similar vein to the common power-criterion suggestions.
@@
 
-6)  Report these values, and designate a minimum sample size, the cut-off criterion, and the maximum sample size.
-Each researcher should also report if they plan to use an adaptive design, which would stop data collection after meeting the cut-off criterion for each item.
-
-These steps are summarized in Table \@ref(tab:table-summary) 1.
+6)  Report these values, and designate a minimum sample size, the cutoff criterion, and the maximum sample size.
+Each researcher should also report if they plan to use an adaptive design, which would stop data collection after meeting the cutoff criterion for each item.
+
+These steps are summarized in Table \@ref(tab:table-summary).
 We will first demonstrate the ideas behind the steps using open data [@brysbaert2014; @balota2007].
@@
     "Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.", 
-    "Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cut-off score from Step 2.",
-    "Determine the sample size at which 80%, 85%, 90%, 95% of items are below the cut-off score.",
-    "Report all values. Designate one as the minimum sample size, the cut-off score as the stopping rule for adaptive designs, and the maximum sample size.",
+    "Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cutoff score from Step 2.",
+    "Determine the sample size at which 80%, 85%, 90%, 95% of items are below the cutoff score.",
+    "Report all values. Designate one as the minimum sample size, the cutoff score as the stopping rule for adaptive designs, and the maximum sample size."),
     "Updated Steps" = c("Use representative pilot data.",
-    "Calculate standard error of each of the items in the pilot data. Using the 40%, determine the cut-off and stopping rule for the standard error of the items.",
+    "Calculate standard error of each of the items in the pilot data. Using the 40%, determine the cutoff and stopping rule for the standard error of the items.",
     "Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.", 
-    "Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cut-off score from Step 2.",
-    "Determine the sample size at which 80%, 85%, 90%, 95% of items are below the cut-off score.
+    "Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cutoff score from Step 2.",
+    "Determine the sample size at which 80%, 85%, 90%, 95% of items are below the cutoff score.
 Use the correction formula to adjust your proposed sample size based on pilot data size, power, and percent variability.",
-    "Report all values. Designate one as the minimum sample size, the cut-off score as the stopping rule for adaptive designs, and the maximum sample size.")
+    "Report all values. Designate one as the minimum sample size, the cutoff score as the stopping rule for adaptive designs, and the maximum sample size.")
   ), 
-  caption = "Proposed Procedure for Powering Studies with Multiple Items"
-))
+  caption = "Proposed Procedure for Powering Studies with Multiple Items",
+  col.names = c("Step", "Proposed Steps", "Updated Steps"),
+  align = c("m{1cm}", rep("m{6cm}", 2)),
+  font_size = "footnotesize"
+)
 ```
@@
 In a lexical decision task, participants simply select "word" for real words (e.g., *dog*) and "nonword" for pseudowords (e.g., *wug*).
-The trial level data is available here [<https://elexicon.wustl.edu/>].
+The trial level data is available here: <https://elexicon.wustl.edu/>.
 Critically, in each of these examples, the individual trial level data for each item is available to simulate and calculate standard errors on.
@@
 From the mean and standard deviation for each item, a simulated pilot dataset could be generated for estimating new sample sizes.
-All code to estimate sample sizes is provided on our OSF page.
+All code to estimate sample sizes is provided on our OSF page, and this manuscript was created with a *papaja* [@aust2022] formatted Rmarkdown document.
 
@@
 Further, participants were allowed to indicate they did not know a word, and this data was set to missing data.
-In our sample of 40 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size` as our pilot sample size for this example.
+In our sample of 40 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size_c` as our pilot sample size for this example.
 
@@
 *Step 2*.
-Table \@ref(tab:table-example) demonstrates the cut-off scores for deciles of the standard errors for the items for the concreteness ratings and lexical decision response latencies.
-A researcher could potentially pick any of these cut-offs or other percentage options not shown here.
+Table \@ref(tab:table-example) demonstrates the cutoff scores for deciles of the standard errors for the items for the concreteness ratings and lexical decision response latencies.
+A researcher could potentially pick any of these cutoffs or other percentage options not shown here.
 We will use simulation to determine the suggestion that best captures the balance of adequately powering our sample and feasibility.
@@
 The pilot data was then simulated, with replacement, with samples from 20 to 300 increasing in units of 5, separately for concreteness and lexical decision times (Step 3).
-The standard error of each item was calculated for the bootstrapped samples (Step 4), and the percentage of items below each potential cut-off was gathered (Step 5).
-The smallest sample size with at least 80%, 85%, 90%, and 95% of items below the cut-off are reported in Table \@ref(tab:table-example) (Step 5).
+The standard error of each item was calculated for the bootstrapped samples (Step 4), and the percentage of items below each potential cutoff was gathered (Step 5).
+The smallest sample size with at least 80%, 85%, 90%, and 95% of items below the cutoff are reported in Table \@ref(tab:table-example) (Step 5).
 
@@
   # sample size
-  # decile cut-offs? 
+  # decile cutoffs? 
   # number of participants for each decile? 
@@
              "L Decile SE", "L 80", "L 85", "L 90", "L 95")
+results_table_example$`C Decile SE` <- apa_num(as.numeric(results_table_example$`C Decile SE`), digits = 2)
+results_table_example$`L Decile SE` <- apa_num(as.numeric(results_table_example$`L Decile SE`), digits = 2)
 
@@
 *Step 6*.
-In the last step, the researcher would indicate their smallest sample size, the cut-off standard error criterion if they wanted to adaptively test (e.g., examine the SE after each participant and stop data collection if all items reached criteria), and their maximum sample size.
-As mentioned earlier, the decile for a balanced standard error cut-off is unclear and without guidance, a potential set of researcher degrees of freedom [CITE].
+In the last step, the researcher would indicate their smallest sample size, the cutoff standard error criterion if they wanted to adaptively test (e.g., examine the SE after each participant and stop data collection if all items reached criteria), and their maximum sample size.
+As mentioned earlier, the decile for a balanced standard error cutoff is unclear and without guidance, a potential set of researcher degrees of freedom [CITE].
 Even though both scales appear to converge on similar sample size suggestions for each decile and percent level, the impact of scale size (i.e., 1-5 versus 0-`r max(elp_small$RT, na.rm = T)`) and heterogeneity of item standard errors (concrete $SD_{SD}$ = `r apa_num(sd_items_c)` and lexical $SD_{SD}$ = `r apa_num(sd_items_e)`) is not obvious.
@@
 For each of these scales, we also manipulated item heterogeneity by simulating small differences in item variances to large differences in item variances based on original scale size.
-On each of the simulated datasets, we applied the above proposed method to determine how the procedure would perform and evaluate what criteria should be used for cut-off selection (Step 2).
+On each of the simulated datasets, we applied the above proposed method to determine how the procedure would perform and evaluate what criteria should be used for cutoff selection (Step 2).
 This procedure was performed on distributions in the middle of the scale (i.e., normal) and at the ceiling of the scale (i.e., skewed).
@@
 1)  How do pilot data influence sample size suggestions?
+
     A.
@@
 In theory, the size of the scale used should not impact the power estimates; however, larger scales have a potential for more variability in their item standard deviations (see point C).
+
     B.
@@
 Therefore, we expect skew to impact the estimates in the same way as point C.
+
     C.
@@
 Heterogeneity should decrease power [@alexander1994; @rheinheimer2001], and thus, increased projected sample sizes should be proposed as heterogeneity of item variances increases.
+
 2)  Do the results match what one might expect for traditional power curves?
@@
 For example, if one has a 500-person pilot study, our simulations should suggest a point at which items are likely measured well, which may have happened well before 500.
-3)  What should the suggested cut-off standard error decile be?
+
+3)  What should the suggested cutoff standard error decile be?
 
@@
   Milliseconds = c(0, 3000, 1000, 2500, 150, 400, 50, 100, 200)
-), caption = "Parameter Values for Data Simulation"
+), caption = "Parameter Values for Data Simulation",
+escape = FALSE
 )
@@
 Standard errors were calculated at each decile of the items up to 90% (i.e., 0% smallest SE, 10% ..., 90% largest SE).
-The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cut-off scores, while the higher deciles would represent less strict criteria for cut-off scores.
+The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores.
 
@@
 
-Next, the percentage of items that fall below the cut-off scores and therefore would be considered "well-measured" were calculated for each decile by sample (Step 4).
-From these data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90% and 95% of the items fall below the cut-off criterion (Step 5).
+Next, the percentage of items that fall below the cutoff scores and therefore would be considered "well-measured" were calculated for each decile by sample (Step 4).
+From these data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90% and 95% of the items fall below the cutoff criterion (Step 5).
 These values were chosen as popular measures of "power" in which one could determine the minimum suggested sample size (potentially 80% of the items) and the maximum suggested sample size (selected from a higher percentage, such as 90% or 95%).
@@
 
-Figure \@ref(fig:scale-size-figure) demonstrates the influence of scale size on the results separate by potential cut-off level.
+Figure \@ref(fig:scale-size-figure) demonstrates the influence of scale size on the results separate by potential cutoff level.
 The black dots denote the original sample size for reference.
@@
 This relationship does not appear to be linear with scale size, as percent scales often represent the highest projected sample size.
-Potentially, this finding is due to the larger proportion of possible variance -- the variance of the item standard deviations / total possible variance -- was largest for percent scales in this set of simulations ($p_{Percent}$ = `r apa_num(mean_var["Percent"], gt1 = FALSE)`.
-This finding may be an interaction with heterogeneity, as the Likert scale had the next highest percent variability in item standard errors ($p_{Likert}$ = `r apa_num(mean_var["Likert"], gt1 = FALSE)`, followed by milliseconds ($p_{Milliseconds}$ = `r apa_num(mean_var["Milliseconds"], gt1 = FALSE)`.
+Potentially, this finding is due to the larger proportion of possible variance -- the variance of the item standard deviations / total possible variance -- was largest for percent scales in this set of simulations ($p_{Percent}$ = `r apa_num(mean_var["Percent"], gt1 = FALSE)`).
+This finding may be an interaction with heterogeneity, as the Likert scale had the next highest percent variability in item standard errors ($p_{Likert}$ = `r apa_num(mean_var["Likert"], gt1 = FALSE)`), followed by milliseconds ($p_{Milliseconds}$ = `r apa_num(mean_var["Milliseconds"], gt1 = FALSE)`).
 
@@
 
-```{r scale-hetero-figure, echo = F, fig.cap="Add a good caption here."}
+```{r scale-hetero-figure, echo = F, fig.cap="Simulated pilot sample size compared to projected sample size for differing amounts of heterogeneity on each scale. "}
 ggplot(summary_long %>% 
@@
 
-```{r strongest-predict}
+```{r strongest-predict, include = F, echo = F}
 regression_DF <- summary_long %>% 
@@
   col.names = c("Term", "Estimate", "$SE$", "$t$", "$p$", "$pr^2$"),
-  escape = FALSE
+  caption = "Prediction of Proposed Sample Size from Simulated Variables"
 )
@@
 These simulations from the pilot data would nearly always suggest a larger sample size - mostly in a linear trend increasing with sample sizes.
-This result comes from the nature of the procedure - if we base our estimates on a SE cut-off, we will almost always need a bit more people for items to meet those goals.
+This result comes from the nature of the procedure - if we base our estimates on a SE cutoff, we will almost always need a bit more people for items to meet those goals.
 This result does not achieve our second goal.
@@
 
-```{r corrected-figure, echo = F, fig.cap="A corrected figure update this caption."}
+```{r corrected-figure, echo = F, fig.cap="Corrected projected sample sizes for variability and power levels. "}
 decay <- 1-sqrt((summary_long$original_n-20)/summary_long$original_n)^log2(summary_long$original_n)
@@
 Proposed sample size and original sample size were the largest predictors -- unsurprising given the correction formula employed -- followed by the percent "power" level and proportion of variance.
-This formula approximation captures $R^2$ = `r apa_print(forward)$estimate$modelfit$r2` of the variance in sample size scores and should allow a researcher to estimate based on their own data, `r apa_print(forward)$statistic$modelfit$r2`.
+This formula approximation captures `r apa_print(forward)$estimate$modelfit$r2` of the variance in sample size scores and should allow a researcher to estimate based on their own data, `r apa_print(forward)$statistic$modelfit$r2`.
 We provide convenience functions in our additional materials to assist researchers in estimating the final adjusted sample size.
@@
      col.names = c("Term", "Estimate", "$SE$", "$t$", "$p$", "AIC"),
-     caption = "Parameters for All Decile Cut-off Scores")
-```
-
-## Choosing an Appropriate Cut-off
+     caption = "Parameters for All Decile Cutoff Scores")
+```
+
+## Choosing an Appropriate cutoff
 
@@
 
-Last, we examine the question of an appropriate SE decile.
+Last, we examined the question of an appropriate SE decile.
 First, the 0%, 10%, and 20% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale size, and heterogeneity.
-If we examine the $R^2$ values for each decile of our regression equation separately, we find that the values are all $R^2$ > .99 with very little differences between them.
-Figure \@ref(fig:decile-figure) illustrates the corrected scores for simulations at the the 40% and 50% decile recommended cut-off for item standard errors.
+If we examine the $R^2$ values for each decile of our regression equation separately, we find that the values are all $R^2$ \> .99 with very little differences between them.
+Figure \@ref(fig:decile-figure) illustrates the corrected scores for simulations at the the 40% and 50% decile recommended cutoff for item standard errors.
 For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 40% decile range, especially for scales with larger potential variance.
@@
   geom_point(aes(original_n, original_n), color = "black") + 
-  theme_classic(base_size = 15) + 
+  theme_classic(base_size = 10) + 
   facet_wrap(~ variability*power) + 
@@
   scale_shape_discrete(name = "Scale Size") +
-  scale_color_discrete(name = "Cut-off Decile") + 
+  scale_color_discrete(name = "Cutoff Decile") + 
   theme(legend.position="bottom")
@@
 
-$$\frac{SD_{Item SD}}{\sqrt{\frac{(Maximum - Minimum)^2}{4}}}$$
-where maximum and minimum are the max and min values found in the scale (or the data, if the scale is unbounded).
+$$\frac{SD_{Item SD}}{\sqrt{\frac{(Maximum - Minimum)^2}{4}}}$$ where maximum and minimum are the max and min values found in the scale (or the data, if the scale is unbounded).
 This formula would be applied in Step 5 of the proposed procedure.
@@
      col.names = c("Term", "Estimate", "$SE$", "$t$", "$p$"),
-     caption = "Parameters for 40% Decile Cut-off Scores")
+     caption = "Parameters for 40% Decile Cutoff Scores")
 ```
@@
 
-```{r table-scores-updated, results = 'asis'}
+```{r table-scores-updated, results = 'asis', eval = F}
 c_sample <- c(results_table_example$`C 80`[4], 
@@
 colnames(table_correction_applied) <- c("Formula Estimate", "Intercept", 
-                                        "Projected Sample Size", "Pilot Sample Size", 
-                                        "Log2 Projected Sample Size", "Log2 Pilot Sample Size", 
-                                        "Log2 Power", "Proportion Variability", 
-                                        "Log2 Proportion Variability", "Power")
+                                        "Projected SS", "Pilot SS", 
+                                        "Log2 Projected SS", "Log2 Pilot SS", 
+                                        "Log2 Power", "Prop Var", 
+                                        "Log2 Prop Var", "Power")
 
@@
 apa_table(table_correction_applied, 
-          caption = "Applied Correction for Each Proposed Sample Size")
+          caption = "Applied Correction for Each Proposed Sample Size", 
+          font_size = "footnotesize")
 ```
@@
 The updated proposal steps are in Table \@ref(tab:table-summary).
-The main change occurs in Step 2 with a designated cut-off decile, and Step 5 with a correction score.
+The main change occurs in Step 2 with a designated cutoff decile, and Step 5 with a correction score.
 Using the data from the 40% decile in Table \@ref(tab:table-example), we can determine that the stopping rule for concreteness ratings would be `r apa_num(cutoff_c[4])`, and the stopping rule for lexical decision times would be `r apa_num(cutoff_e[4])`.
 For Step 5, we apply our correction formula separately for each one, as they have different variability scores, and these scores are shown in Table \@ref(tab:table-scores-updated).
-Each row was multipled by row one's formula, and then these scores are summed for the final sample size.
+Each row was multiplied by row one's formula, and then these scores are summed for the final sample size.
 Sample sizes cannot be proportional, so we recommend rounding up to the nearest whole number.
 
-The concreteness ratings data includes `r length(unique(conDF$Word))` concepts that were rated for their concreteness.
-In our fictional study for this example, we selected 100 random words to show participants.
-In the original study, not every participant rated every word, which created uneven sample sizes for each word.
-In our random sample of 100 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size_c` as our pilot sample size for this example.
-All "do not know" ratings were set as missing data.
-The 50% decile for items standard error was `r round(cutoff, digits = 3)` for our cut-off criterion.
-
-For one additional consideration, we calculated the potential amount of data retention given that participants could indicate they did not know a word ($M_{correct}$ = `r round(mean(data_loss$percent_correct), digits = 2)`, *SD* = `r round(sd(data_loss$percent_correct), digits = 2)`).
-In order to account for this facet, the potential sample sizes were multiplied by 1/`r round(mean(data_loss$percent_correct), digits = 2)`,
+For one additional consideration, we calculated the potential amount of data retention given that participants could indicate they did not know a word ($M_{answered}$ = `r round(mean(data_loss_c$percent_correct), digits = 2)`, *SD* = `r round(sd(data_loss_c$percent_correct), digits = 2)`) in the concreteness task or answer a trial incorrectly in the lexical decision task ($M_{correct}$ = `r round(mean(data_loss_e$percent_correct), digits = 2)`, *SD* = `r round(sd(data_loss_e$percent_correct), digits = 2)`).
+In order to account for this facet, the potential sample sizes were multiplied by $\frac{1}{p_{retained}}$ where the denominator is proportion retained for each task.
 
@@
 
-`simulate_population()`
-
-simulate_population(mu = 4, mu_sigma = .2, sigma = 2, sigma_sigma = .2, number_items = 30, number_scores = 1000, smallest_sigma = .02, min_score = 1, max_score = 7, digits = 0)
+We have developed functions to implement the suggested procedure as part of an upcoming package `semanticprimeR`.
+You can install the package from GitHub using: `devtools::install_github("SemanticPriming/semanticprimeR")`.
+We detail the functions below by proposed step in the process.
+
+*Step 1*.
+Ideally, researchers would have pilot data that represented their proposed data collection.
+This data should be formatted in long format wherein each row represents the score from an item by participant, rather than wide format wherein each column represents an item and each row represents a single participant.
+The `tidyr::pivot_longer()` or `reshape::melt()` functions can be used to reformat wide data.
+If no pilot data is available, the `simulate_population()` function can be used with the following arguments (and example numbers, \* indicates optional).
+This function will return a dataframe with the simulated normal values for each item.
+
+```{r sim_pop, include = T, echo = T}
+# devtools::install_github("SemanticPriming/semanticprimeR")
+library(semanticprimeR)
+pops <- simulate_population(mu = 4, # item means
+  mu_sigma = .2, # variability in item means 
+  sigma = 2, # item standard deviations
+  sigma_sigma = .2, # standard deviation of the standard deviations
+  number_items = 30, # number of items
+  number_scores = 20, # number of participants
+  smallest_sigma = .02, #* smallest possible standard deviation
+  min_score = 1, #* minimum score for truncating purposes
+  max_score = 7, #* maximum score for truncating purposes
+  digits = 0) #* number of digits for rounding
+
+head(pops)
+```
+
+*Step 2*.
+In step 2, we can use `calculate_cutoff()` to calculate the standard error of the items, the standard deviation of the standard errors and the corresponding proportion of variance possible, and the 40% decile cutoff score.
+The `pops` dataframe can be used in this function, which has columns named `item` for the item labels (i.e., 1, 2, 3, 4 or characters can be used), and `score` for the dependent variable.
+This function returns a list of values to be used in subsequent steps.
+
+```{r calc_cut, include = T, echo = T}
+cutoff <- calculate_cutoff(population = pops, # pilot data or simulated data
+  grouping_items = "item", # name of the item indicator column
+  score = "score", # name of the dependent variable column
+  minimum = 1, # minimum possible/found score
+  maximum = 7) # maximum possible/found score
+
+cutoff$se_items # all standard errors of items
+cutoff$sd_items # standard deviation of the standard errors
+cutoff$cutoff # 40% decile score
+cutoff$prop_var # proportion of possible variance 
+```
+
+*Step 3*.
+The `bootstrap_samples()` function creates bootstrapped samples from the pilot or simulated population data to create samples to estimate the number of participants needed for item standard error to be below the cutoff calculated in Step 2.
+This function returns a list of samples with sizes that start at the `start` size, increase by `increase`, and end with the `stop` sample size.
+The population or pilot data will be included in `population`, and the item column indicator should be included in `grouping_items`.
+
+```{r boot-sample, include = T, echo = T}
+samples <- bootstrap_samples(start = 20, # starting sample size
+  stop = 100, # stopping sample size
+  increase = 5, # increase bootstrapped samples by this amount
+  population = pops, # population or pilot data
+  replace = TRUE, # bootstrap with replacement? 
+  grouping_items = "item") # item column label
+
+head(samples[[1]])
+```
+
+*Step 4 and 5*.
+The proportion of bootstrapped items across sample sizes below the cutoff score can then be calculated using `calculate_proportion()`.
+This function returns a dataframe of each sample size with the proportion of items below that cutoff to use in the next function.
+The `samples` and `cutoff` were calculated with our previous functions, with the column for item labels and dependent variable to ensure the right calculations.
+
+```{r calc-prop, include = T, echo = T}
+proportion_summary <- calculate_proportion(samples = samples, # samples list
+  cutoff = cutoff$cutoff, # cut off score 
+  grouping_items = "item", # item column name
+  score = "score") # dependent variable column name
+
+head(proportion_summary)
+```
+
+*Step 6*.
+Last, we use the `calculate_correction()` function to correct the sample size scores given the proposed correction formula.
+The `proportion_summary` from above is used in this function, along with required information about the sample size, proportion of variance from our cutoff calculation, and what power levels should be calculated.
+Note that the exact percent of items below a cutoff score will be returned, if the values in `power_levels` are not exactly present.
+The final summary presents the smallest sample size, corrected, for each of the potential power levels.
+
+```{r calc-correct, include = T, echo = T}
+corrected_summary <- calculate_correction(
+  proportion_summary = proportion_summary, # prop from above
+  pilot_sample_size = 20, # number of participants in the pilot data 
+  proportion_variability = cutoff$prop_var, # proportion variance from cutoff scores
+  power_levels = c(80, 85, 90, 95)) # what levels of power to calculate
+
+corrected_summary
+```
 
@@
 While the examples in this manuscript are traditionally cognitive linguistics focused, any research using repeated items can benefit from newer sampling techniques.
-Therefore, we provide XX example vignettes and code examples on our OSF page/GitHub site for this manuscript across a range of examples of data types provided by the authors of this manuscript.
-Examples include psycholinguistics, social psychology, COVID related data, and cognitive psychology.
-The code for this manuscript can be found in *papaja* [@aust2022] formated Rmarkdown document.
-To
+Therefore, we provide 12 example vignettes and varied code examples on our OSF page/GitHub site for this manuscript across a range of data types provided by the authors of this manuscript.
+Examples include psycholinguistics [@dedeyne2008; @heyman2014; @montefinese2022], eye tracking data [@ulloa2014], social psychology [@peterson2022; @grahe2022], COVID related data [@montefinese2021], and cognitive psychology [@röer2013; @10.7554/eLife.71601; @barzykowski2019].
 
@@
 
-In this manuscript, we demonstrated a method using AIPE and simulation/bootstrapping to estimate a minimum and maximum sample size along with a rule for stopping data collection based on narrow confidence intervals on the parameter of interest.
-We believe this procedure is specifically useful for studies with multiple items that intend on using item level focused analyses; however, the utility of measuring each item well can extend to many analysis choices.
-By focusing on gathering quality data, we can suggest that the data is useful, regardless of outcome of any hypothesis test.
-
-One limitation of these methods would be using datasets with very large numbers of items to simulate what might happen within one study.
-For example, the English Lexicon Project includes thousands of items, and by the time we would simulate for all of those, it would likely suggest needing thousands of participants for *most items* to reach criterion.
-Alternatively, as the number of items increases, you also could potentially see very small estimates for sample size due to the correction factor (as with large numbers of items, you could find many items with standard errors below the 50% decile).
-Therefore, it would be beneficial to consider only simulating with what a participant would reasonably complete in a study.
-On the other side, small numbers of repeated items usually result in higher sample sizes proposed from the original pilot data.
+We proposed and demonstrated examples for a method using AIPE, bootstrapping, and simulation to estimate a minimum and maximum sample size along with a rule for stopping data collection based on narrow confidence intervals on the parameter of interest.
+We contend that this procedure is specifically useful for studies with multiple items that intend on using item level focused analyses; furthermore, the utility of measuring each item well can extend to many analysis choices.
+By focusing on collecting quality data, we can suggest that the data is useful, regardless of the outcome of any hypothesis test.
+
+One limitation of these methods would be our decision to use datasets with very large numbers of items to simulate what might happen within one study.
+For example, the English Lexicon Project includes thousands of items, and by the time we would simulate for all of those, it would likely suggest needing thousands of participants for most items to reach the criterion.
+Additionally, as the number of items increases, you may also see very small estimates for sample size due to the correction factor (as with large numbers of items, you could find many items with standard errors below the 40% decile).
+Therefore, it would be beneficial to consider only simulating what a participant would reasonably complete in a study.
+Small numbers of repeated items usually result in larger sample sizes proposed from the original pilot data.
 This result occurs because the smaller number of items means more samples for nearly all to reach the cutoff criteria.
-These results are not too different than what we might expect for a power analysis using a multilevel model - larger number of items tends to decrease necessary sample size, while smaller numbers of items tend to increase sample size.
-
-Second, these methods do not ensure the normal interpretation of power, wherein you know would find a specific effect for a specific test, $\alpha$, and so on.
-As discussed in the introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis, and many of the guesses within a traditional power analysis are just that - best guesses for various parameters.
-These methods could be used together to strengthen our understanding of sample size necessary for both a hypothesis test and well tuned estimation.
+These results are not too different from what we might expect for a power analysis using a multilevel model - larger numbers of items tend to decrease necessary sample size, while smaller numbers of items tend to increase sample size.
+
+Second, these methods do not ensure the normal interpretation of power, where you know that you would find a specific effect for a specific test, $\alpha$, and so on.
+As discussed in the Introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis, many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
+These methods could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
+
+Researchers should consider this hybrid approach for AIPE, bootstrapping, and simulation as a tool for hypothesis testing and parameter estimation.
+The proposed procedure can be beneficial for many different research studies, specifically replication studies, that usually depend on subject sample size but rarely item sample size, in spite of the fact that item sample sizes contribute to language processing [@brysbaert2018; @baayen2008].
+These observed effects can be then replicated and, as a result of several replications, can be applied to meta-analyses.
+As a result, analysts would be able to use the accuracy and high statistical power to calculate the parameters to assess whether the effect is genuine.
+This article helps to achieve this goal by encouraging researchers to conduct studies where the power analysis is not based on the size of the effect but on adequate sampling of the stimuli.
+We argue that this article can be the initial step to apply AIPE in a manner that can allow researchers to use item information to provide a more accurate and statistically reliable measure of the effect we aimed to investigate.
+In conclusion, item power analysis is a tool to avoid the waste of resources while ensuring that adequately measured items can be achieved.
+Well measured data can enable us to counteract the literature that contains false positives, allowing us to achieve replicable, high-quality science to establish answers to scientific questions with precision and accuracy.
 

### 2023-09-08 manuscript_draft.tex
all the updates

--- 2022-06-22 manuscript_draft.tex
@@
 \documentclass[
-  man]{apa6}
+  man]{apa7}
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
 \fi
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Power to the Stimuli: Not the Effect},
-  pdfauthor={Erin M. Buchanan1 \& Other Folks as Per Order on Doc2},
-  pdflang={en-EN},
-  pdfkeywords={power, sampling, accuracy in parameter estimation},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{graphicx}
@@
 {}{\PatchFailed}
-\keywords{power, sampling, accuracy in parameter estimation}
+\keywords{accuracy in parameter estimation, power, sampling, simulation, hypothesis testing}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \usepackage{csquotes}
+\makeatletter
+\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries\typesectitle}}
+
+\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-\z@\relax}%
+  {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+\makeatother
+
 \ifLuaTeX
@@
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  pdftitle={Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli},
+  pdfauthor={Erin M. Buchanan1 \& Other Folks as Per Order on Doc2},
+  pdflang={en-EN},
+  pdfkeywords={accuracy in parameter estimation, power, sampling, simulation, hypothesis testing},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
-\title{Power to the Stimuli: Not the Effect}
+\title{Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli}
 \author{Erin M. Buchanan\textsuperscript{1} \& Other Folks as Per Order on Doc\textsuperscript{2}}
@@
 
-\shorttitle{Title}
+\shorttitle{AIPE BOOTSTRAPPING}
 
 \authornote{
-
-The authors would like to thank K.D. Valentine for her assistance in formulation of correction scores.
 
@@
 \abstract{%
-We will add an abstract that explains that we are going to 1) talk about this cool procedure, 2) show some simulations that the procedure works, and 3) give two examples of the procedure in action.
-You should try it out!
+The planning of the sample size for research studies often focuses on obtaining a significant result given a specified level of power, significance, and an anticipated effect size.
+This planning requires prior knowledge of the study design and a statistical analysis to calculate the proposed sample size.
+However, there may not be one specific testable analysis from which to derive power (Silberzahn et al., 2018) or a hypothesis to test for the project (e.g., creation of a stimuli database).
+Modern power and sample size planning suggestions include accuracy in parameter estimation (AIPE, Kelley, 2007; Maxwell et al., 2008) and simulation of proposed analyses (Chalmers \& Adkins, 2020).
+These toolkits provide flexibility in traditional power analyses that focus on the if-this, then-that approach, yet, both AIPE and simulation require either a specific parameter (e.g., mean, effect size, etc.) or statistical test for planning sample size.
+In this tutorial, we explore how AIPE and simulation approaches can be combined to accommodate studies that may not have a specific hypothesis test or wish to account for the potential of a multiverse of analyses.
+Specifically, the examples focus on studies that adopt multiple items and suggest that sample sizes can be planned to measure those items adequately and precisely, regardless of statistical test.
+We demonstrate that pilot data can be used to determine a sample size that represents well-measured data.
+This tutorial also provides multiple code vignettes that researchers can adapt and apply to their own measures.
 }
@@
 
-This part we will put in a real introduction. Things we need to discuss:
+Statistical power and power analyses are arguably one of the most important components in planning a research study (Cohen, 1990).
+However, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts (Hardwicke et al., 2020, 2022).
+The replication ``crisis'' and credibility revolution have shown that published studies in psychology are underpowered (Open Science Collaboration, 2015; Vazire, 2018).
+Pre-registration of a study involves outlining the study and hypotheses before data collection begins (Chambers et al., 2014; Nosek \& Lakens, 2014; Stewart et al., 2020), and details of a power analyses or other limitations on resources are often used to provide justification for the pre-registered sample quota.
+Given the combined issues of publish-or-perish and that most non-significant results do not result in published manuscripts, one may expect that power analysis would be especially critical for early career researchers (Rosenthal, 1979; Simmons et al., 2011).
+At best, an underpowered study provides limited insight (Halpern, 2002), and it can be difficult to know if a poorly implemented power analysis is better than no power analysis.
 
-\begin{itemize}
-\tightlist
-\item
-  Repeated measures data with items
-\item
-  Accuracy in Parameter Estimation
-\item
-  Adaptive sampling
-\item
-  MLM designs account for item heterogeneity but a lot of people anova
-\end{itemize}
+A recent review of power analyses found - across psychology journal articles - researchers did not provide enough information to understand their power analyses and often chose effect sizes that were unjustified (Beribisky, 2019).
+One solution to this power analysis problem is the plethora of tools made available for researchers to make power computations accessible to non-statisticians with the caveat that a solid education in power is necessary to use these tools.
+G*Power is one of the most popular free power software options (Erdfelder et al., 1996; Faul et al., 2007) that provides a simple point and click graphical user interface for power calculations (however, see Brysbaert, 2019).
+Web-based tools have also sprung up for overall and statistical test specific sample size planning including \url{https://powerandsamplesize.com} and \url{https://designingexperiments.com} (Anderson et al., 2017).
+\emph{R}-coding based packages, such as \emph{pwr} (Champely et al., 2017), \emph{faux} (DeBruine, 2021), and \emph{SimDesign} (Chalmers \& Adkins, 2020), can be used to examine power and sample size planning usually with simulation.
+Researchers must be careful using any toolkit, as errors can occur with the over-reliance on software (e.g., it should not be a substitute for critical thinking, Nuijten et al., 2016).
+When computing sample size estimates, it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result (Batterham \& Atkinson, 2005).
+For example, if any parameter estimated by the researcher was not found in the study (i.e., a smaller effect size than used for the power analysis), then the sample size estimate tied to that specific level of power may be incorrect.
 
-\hypertarget{simulating-sample-size}{%
-\subsection{Simulating Sample Size}\label{simulating-sample-size}}
+Changes in publication practices and research design have also created a new wrinkle in providing a sample size plan for a research study.
+While statistics courses often suggest that a specific research design leads to a specific statistical test, meta-science work has shown that given the same data and hypothesis, researchers can come up with multiple ways to analyze the data Coretta et al.
+(2023).
+Therefore, a single power analysis only supports the specific analysis in which the researcher expects to test.
+Analyses may evolve during the research project or be subject to secondary analysis; thus, power and sample size estimation based on one analysis is potentially less useful than previously imagined.
+Further, research projects often have multiple testable hypotheses, but it is unclear which hypothesis or test should be used to estimate sample size with a power analysis.
+Last, research publications may not even have a specific, testable hypothesis, as some publications are intended to curate a large dataset for future reuse (i.e., stimuli database creation, Buchanan et al., 2019).
 
-Using ideas from accuracy in parameter estimation, we suggest the following procedure to determine a sample size for each item:
+In light of these analytical (or lack thereof) concerns, we propose a new method that combines accuracy in parameter estimation Maxwell et al.
+(2008) and bootstrapped simulation on pilot data (Rousselet et al., 2022).
+This method accounts for a potential lack of hypothesis test (or simply not a good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices.
+Additionally, we consider the nature of cognitive research designs that use multiple items to measure the phenomena of interest.
+For example, semantic priming is measured with multiple paired stimuli (Meyer \& Schvaneveldt, 1971), which traditionally was analyzed by creating person or item-level averages for an ANOVA (Brysbaert \& Stevens, 2018).
+However, the use of multilevel models with random effects for the stimuli used in a study have shown that we should be careful to assume that all items of a research study have the same ``effect'', as there is often variability in their impact on the outcomes of the study.
+
+\hypertarget{accuracy-in-parameter-estimation}{%
+\subsection{Accuracy in Parameter Estimation}\label{accuracy-in-parameter-estimation}}
+
+AIPE shifts the focus away from finding a significant p-value to finding a parameter that is ``accurately measured''.
+For example, researchers may wish to detect a specific correlation in a study, \emph{r} = .35.
+They could then use AIPE to estimate the sample size needed to find a ``sufficiently narrow'' confidence interval around that correlation.
+Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and confidence intervals.
+Therefore, they could decide that their 95\% confidence interval should be approximately between .20 and .50, and sufficiently narrow could be defined as a width of .30 or .15 on each side.
+While confidence intervals are related to null hypothesis significance testing (i.e., 95\% confidence intervals that do not include zero would indicate a significant difference from zero at \(\alpha\) \textless{} .05), AIPE procedures suggest how we can define a sample size with a given width of confidence interval, regardless of whether it includes zero.
+
+\hypertarget{bootstrapping-and-simulation}{%
+\subsection{Bootstrapping and Simulation}\label{bootstrapping-and-simulation}}
+
+Bootstrapping involves using data obtained to simulate similar datasets by drawing from the original data with replacement (Efron, 2000; Rousselet et al., 2022).
+Bootstrapping is a form of data simulation that allows one to calculate parameter estimates, confidence intervals, and more to simulate the potential population distribution, shape, and bias.
+Simulation is often paired with making up data for testing analyses and hypotheses based on proposed effect size or suggested population means.
+Generally, we would suggest starting with pilot data of a smaller sample size to understand the variability in potential items used to represent your phenomenon, especially if they are to be used in a larger study.
+However, given some background knowledge about the potential items, one could simulate example pilot data to use in.
+similar manner in our suggested procedure.
+Pilot or simulated data would be used to estimate the variability within items and select a ``sufficiently narrow'' window for item's confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
+Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein many items meet the required confidence interval.
+
+\hypertarget{sequential-testing}{%
+\subsection{Sequential Testing}\label{sequential-testing}}
+
+Researchers could then use sequential testing to estimate their parameter of interest after each participant's data was collected to determine whether they have achieved their expected width of the confidence interval around that parameter.
+One would set a minimum sample size (i.e., based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the CI is narrow, as defined above).
+Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (i.e., we expect 90\% of items to meet our stopping rule with 100 participants based on simulation).
+By defining each of these components, researchers could ensure a feasible minimum sample size, a way to stop data collection when goals have been met, and a maximum sample size rule to ensure an actual end to data collection.
+Therefore, we proposed that we should be able to leverage the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for repeated measures studies with expected variability in parameter estimates for items.
+
+\hypertarget{proposed-method-to-calculate-sample-size}{%
+\subsection{Proposed Method to Calculate Sample Size}\label{proposed-method-to-calculate-sample-size}}
+
+Using these ideas, we suggest the following procedure to determine a sample size for each item:
 
@@
 \def\labelenumi{\arabic{enumi})}
-\tightlist
 \item
-  Use pilot data that closely resembles your intended data collection, on the same or similar items that will be used in the study.
-In this procedure, we will assume that the pilot data is representative of a larger population of sampled items that you intend to assess.
+  Use pilot data that closely resembles data you intend to collect.
+This dataset should contain items that are identical or similar to those that will be collected in the study.
+In this procedure, it is important to ensure that the data is representative of a larger population of sampled items that you intend to assess.
+Generally, pilot data sample sizes will be small, as the goal would be to determine how many items would be necessary to reach a ``stable'' standard error.
 \item
-  Calculate the standard error of each item from the pilot data to create a cutoff score for when an item is ``accurately measured''.
-The simulations below will explore what criterion to use when determining the cutoff score from the pilot data.
+  Calculate the standard error (SE) in the subset of pilot data associated with each item to create a cutoff score that defines when items are considered ``accurately measured''.
+Next, the simulations below will explore what criterion should be used to determine the cutoff score from the pilot data.
 \item
-  Sample, with replacement, from your pilot data using sample sizes starting at 20 participants and increase in small units (e.g., 20, 25, 30) up to a value that you consider the maximum sample size.
-We will demonstrate example maximum sample sizes based on the data simulation below; however, a practical maximum sample size may be determined by time (e.g, one semester data collection) or researcher resources (e.g., 200 participants worth of funding).
-While 20 participants would likely represent an underpowered study, we simply suggest this starting minimum for simulation purposes.
+  Sample, with replacement, from your pilot data using sample sizes starting at 20 participants and increase in small units up to a value that you consider the maximum sample size.
+We will demonstrate example maximum sample sizes based on the data simulation below; however, a practical maximum sample size may be determined by time (e.g., one semester data collection) or resources (e.g., 200 participants worth of funding).
+Although 20 participants would likely yield imprecise estimates, we suggest this starting minimum for simulation purposes.
 \item
-  For each simulated sample, calculate the standard error for each item, and use these values to ascertain the percentage of items that meet the cutoff score determined in step 2.
+  For each simulation, calculate the standard error for each item, and use these values to determine the percentage of items that meet the cutoff score determined in Step 2.
 \item
-  Find the minimum sample size that meets 80\%, 85\%, 90\%, and 95\% of the items.
-We recommend these scores to ensure that most items are accurately measured, in a similar vein to common power criteria suggestions.
-Each researcher can determine which of these is their minimum or maximum sample size (e.g., individual can choose to use 80\% as a minimum and 90\% as a maximum).
+  Find the minimum sample size so that 80\%, 85\%, 90\%, and 95\% of the items meet the cutoff score and can be considered accurately measured.
+We recommend these scores to ensure that most items are accurately measured, in a similar vein to the common power-criterion suggestions.
+Each researcher can determine which of these is their minimum or maximum sample size (e.g., individuals can choose to use 80\% as a minimum and 90\% as a maximum or use values from Step 3 based on resources).
 \item
@@
 
-\hypertarget{key-issues}{%
-\subsection{Key Issues}\label{key-issues}}
+These steps are summarized in Table \ref{tab:table-summary}.
+We will first demonstrate the ideas behind the steps using open data (Balota et al., 2007; Brysbaert et al., 2014).
+This example will reveal a few areas of needed exploration for the steps.
+Next, we portray simulations for the proposed procedure and find solutions to streamline and improve the sample size estimation procedure.
+Finally, we include additional resources for researchers to use to implement the estimation procedure.
 
-Given the long history of research on power, there are a few key issues that this procedure should address:
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:table-summary}Proposed Procedure for Powering Studies with Multiple Items}
+
+\footnotesize{
+
+\begin{tabular}{m{1cm}m{6cm}m{6cm}}
+\toprule
+Step & Proposed Steps & Updated Steps\\
+\midrule
+1 & Use representative pilot data. & Use representative pilot data.\\
+2 & Calculate standard error of each of the items in the pilot data.
+Determine the appropriate SE for the stopping rule.
+& Calculate standard error of each of the items in the pilot data.
+Using the 40\%, determine the cutoff and stopping rule for the standard error of the items.\\
+3 & Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.
+& Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.\\
+4 & Calculate the standard error of each of the items in the bootstrapped data.
+From these scores, calculate the percent of items below the cutoff score from Step 2.
+& Calculate the standard error of each of the items in the bootstrapped data.
+From these scores, calculate the percent of items below the cutoff score from Step 2.\\
+5 & Determine the sample size at which 80\%, 85\%, 90\%, 95\% of items are below the cutoff score.
+& Determine the sample size at which 80\%, 85\%, 90\%, 95\% of items are below the cutoff score.
+Use the correction formula to adjust your proposed sample size based on pilot data size, power, and percent variability.\\
+6 & Report all values.
+Designate one as the minimum sample size, the cutoff score as the stopping rule for adaptive designs, and the maximum sample size.
+& Report all values.
+Designate one as the minimum sample size, the cutoff score as the stopping rule for adaptive designs, and the maximum sample size.\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\hypertarget{example}{%
+\section{Example}\label{example}}
+
+In this section, we provide two examples of the suggested procedure.
+The first example includes concreteness ratings from Brysbaert et al.
+(2014).
+Instructions given to participants denoted the difference between concrete (i.e., ``refers to something that exists in reality'') and abstract (i.e., ``something you cannot experience directly through your senses or actions'') terms.
+Participants were then asked to rate concreteness of terms using a 1 (\emph{abstract}) to 5 (\emph{concrete}) scale.
+This data represents a small scale dataset that could be used as pilot data for a study using concrete word ratings.
+The data is available at \url{https://osf.io/qpmf4/}.
+
+The second dataset includes a large scale dataset with response latencies, the English Lexicon Project (Balota et al., 2007).
+The English Lexicon Project consists of lexical decision response latencies for English words.
+In a lexical decision task, participants simply select ``word'' for real words (e.g., \emph{dog}) and ``nonword'' for pseudowords (e.g., \emph{wug}).
+The trial level data is available here: \url{https://elexicon.wustl.edu/}.
+Critically, in each of these examples, the individual trial level data for each item is available to simulate and calculate standard errors on.
+Data that has been summarized could potentially be used, as long as the original standard deviations for each item were present.
+From the mean and standard deviation for each item, a simulated pilot dataset could be generated for estimating new sample sizes.
+All code to estimate sample sizes is provided on our OSF page, and this manuscript was created with a \emph{papaja} (Aust et al., 2022) formatted Rmarkdown document.
+
+For this example, our researcher wants to determine the differences in response latencies for abstract and concrete words.
+They will select \emph{n} = 40 words from the rating data that are split evenly into abstract and concrete ends of the rating scale.
+In the experiment, each participant will rate the words for their concreteness, and then complete a lexical decision task with these words as the object of interest.
+Using both datasets, we can determine the sample size necessary to ensure adequately measured ratings and response latencies.
+
+\emph{Step 1}.
+The concreteness ratings data includes 27250 concepts that were rated for their concreteness.
+We randomly selected \emph{n} = 20 abstract words (\(M_{Rating}\) \textless= 2) and \emph{n} = 20 concrete words (\(M_{Rating}\) \textgreater= 4).
+In the original study, not every participant rated every word, which created uneven sample sizes for each word.
+Further, participants were allowed to indicate they did not know a word, and this data was set to missing data.
+In our sample of 40 words, the average pilot sample size was 28.23 (\emph{SD} = 1.35), and we will use 28 as our pilot sample size for this example.
+
+The ELP response latency data includes 27250 word-forms, 219 that are listed as non-words, and 27031 real words.
+We selected the same words as the concreteness subset selected above.
+The average pilot sample size for this random sample was 32.80 (\emph{SD} = 0.56), and \emph{n} = 33 will be our pilot size for this example.
+
+\emph{Step 2}.
+Table \ref{tab:table-example} demonstrates the cutoff scores for deciles of the standard errors for the items for the concreteness ratings and lexical decision response latencies.
+A researcher could potentially pick any of these cutoffs or other percentage options not shown here.
+We will use simulation to determine the suggestion that best captures the balance of adequately powering our sample and feasibility.
+
+\emph{Step 3-5}.
+The pilot data was then simulated, with replacement, with samples from 20 to 300 increasing in units of 5, separately for concreteness and lexical decision times (Step 3).
+The standard error of each item was calculated for the bootstrapped samples (Step 4), and the percentage of items below each potential cutoff was gathered (Step 5).
+The smallest sample size with at least 80\%, 85\%, 90\%, and 95\% of items below the cutoff are reported in Table \ref{tab:table-example} (Step 5).
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:table-example}Sample Size Estimates by Decile for Concreteness Example}
+
+\begin{tabular}{lllllllllll}
+\toprule
+Deciles & \multicolumn{1}{c}{C Decile SE} & \multicolumn{1}{c}{C 80} & \multicolumn{1}{c}{C 85} & \multicolumn{1}{c}{C 90} & \multicolumn{1}{c}{C 95} & \multicolumn{1}{c}{L Decile SE} & \multicolumn{1}{c}{L 80} & \multicolumn{1}{c}{L 85} & \multicolumn{1}{c}{L 90} & \multicolumn{1}{c}{L 95}\\
+\midrule
+Decile 10 & 0.13 & 85 & 85 & 95 & 100 & 39.64 & 115 & 135 & 135 & 175\\
+Decile 20 & 0.14 & 70 & 75 & 80 & 90 & 45.67 & 85 & 90 & 110 & 125\\
+Decile 30 & 0.16 & 50 & 55 & 60 & 65 & 46.76 & 85 & 90 & 105 & 115\\
+Decile 40 & 0.18 & 45 & 50 & 50 & 65 & 51.32 & 75 & 75 & 90 & 105\\
+Decile 50 & 0.19 & 35 & 35 & 45 & 50 & 71.17 & 40 & 45 & 45 & 50\\
+Decile 60 & 0.20 & 30 & 35 & 35 & 45 & 78.95 & 35 & 35 & 45 & 50\\
+Decile 70 & 0.21 & 30 & 30 & 35 & 45 & 87.85 & 25 & 35 & 35 & 35\\
+Decile 80 & 0.23 & 25 & 30 & 30 & 40 & 100.45 & 20 & 25 & 25 & 35\\
+Decile 90 & 0.25 & 20 & 20 & 25 & 40 & 121.11 & 20 & 20 & 20 & 20\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Esimates are based on meeting at least the minimum percent of items (e.g., 80\%) but may be estimated over that amount (e.g., 82.5\%).}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\emph{Step 6}.
+In the last step, the researcher would indicate their smallest sample size, the cutoff standard error criterion if they wanted to adaptively test (e.g., examine the SE after each participant and stop data collection if all items reached criteria), and their maximum sample size.
+As mentioned earlier, the decile for a balanced standard error cutoff is unclear and without guidance, a potential set of researcher degrees of freedom {[}CITE{]}.
+Even though both scales appear to converge on similar sample size suggestions for each decile and percent level, the impact of scale size (i.e., 1-5 versus 0-2852) and heterogeneity of item standard errors (concrete \(SD_{SD}\) = 0.26 and lexical \(SD_{SD}\) = 116.13) is not obvious.
+Last, by selecting the ends of the distribution for our concreteness words, skew of the distribution may additionally impact our estimates.
+Each of these will be explored in our simulation.
+
+\hypertarget{simulation-method}{%
+\section{Simulation Method}\label{simulation-method}}
+
+In order to evaluate our approach, we used data simulation to create representative pilot datasets of several popular cognitive scales (1-7 measurements, 0-100 percentage measurements, and 0-3000 response latency scale data).
+For each of these scales, we also manipulated item heterogeneity by simulating small differences in item variances to large differences in item variances based on original scale size.
+On each of the simulated datasets, we applied the above proposed method to determine how the procedure would perform and evaluate what criteria should be used for cutoff selection (Step 2).
+This procedure was performed on distributions in the middle of the scale (i.e., normal) and at the ceiling of the scale (i.e., skewed).
+With this simulation, we will answer several questions:
 
@@
 \def\labelenumi{\arabic{enumi})}
-\tightlist
 \item
-  We should see differences in projected sample sizes based on the variability in the variance for those items (i.e., heterogeneity should increase projected sample size).
+  How do pilot data influence sample size suggestions?
+
+  A.
+How does scale size impact sample size estimations?
+In theory, the size of the scale used should not impact the power estimates; however, larger scales have a potential for more variability in their item standard deviations (see point C).
+
+  B.
+How does distribution skew impact sample size estimations?
+Skew can potentially decrease heterogeneity (i.e., all items are at ceiling, and therefore, variance between item standard errors is low) or could increase heterogeneity (i.e., some items are skewed, while others are not).
+Therefore, we expect skew to impact the estimates in the same way as point C.
+
+  C.
+How does heterogeneity impact sample size estimations?
+Heterogeneity should decrease power (Alexander \& DeShon, 1994; Rheinheimer \& Penfield, 2001), and thus, increased projected sample sizes should be proposed as heterogeneity of item variances increases.
 \item
-  We should see projected sample sizes that ``level off'' when pilot data increases.
-As with regular power estimates, studies can be ``overpowered'' to detect an effect, and this same idea should be present.
-For example, if one has a 500 person pilot study, our simulations should suggest a point at which items are likely measured well, which may have happened well before 500.
+  Do the results match what one might expect for traditional power curves?
+Power curves are asymptotic, that is, they ``level off'' as sample size increases.
+Therefore, we expect that our procedure should also demonstrate a leveling off effect as pilot data sample size increases.
+For example, if one has a 500-person pilot study, our simulations should suggest a point at which items are likely measured well, which may have happened well before 500.
+\item
+  What should the suggested cutoff standard error decile be?
 \end{enumerate}
-
-\hypertarget{method}{%
-\section{Method}\label{method}}
 
@@
 
+Table \ref{tab:table-sim} presents the variables and information about the simulations as a summary.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:table-sim}Parameter Values for Data Simulation}
+
+\begin{tabular}{llll}
+\toprule
+Information & \multicolumn{1}{c}{Likert} & \multicolumn{1}{c}{Percent} & \multicolumn{1}{c}{Milliseconds}\\
+\midrule
+Minimum & 1.00 & 0.00 & 0.00\\
+Maximum & 7.00 & 100.00 & 3,000.00\\
+$\mu$ & 4.00 & 50.00 & 1,000.00\\
+$Skewed \mu$ & 6.00 & 85.00 & 2,500.00\\
+$\sigma_{\mu}$ & 0.25 & 10.00 & 150.00\\
+$\sigma$ & 2.00 & 25.00 & 400.00\\
+Small $\sigma_{\sigma}$ & 0.20 & 4.00 & 50.00\\
+Medium $\sigma_{\sigma}$ & 0.40 & 8.00 & 100.00\\
+Large $\sigma_{\sigma}$ & 0.80 & 16.00 & 200.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
 \emph{Population}.
-The data was simulated using the \texttt{rnorm} function assuming a normal distribution for 30 scale type items.
-Each population was simulated with 1000 data points.
-No items were rounded for this simulation.
+We simulated data for 30 items using the rnorm function assuming a normal distribution.
+Each item's population data was simulated with 1000 data points.
+Items were rounded to the nearest whole number to mimic scales generally collected by researchers.
+Items were also rounded to their appropriate scale end points (i.e., all items below 0 on a 1-7 scale were replaced with 1, etc.).
 
+\emph{Data Scale}.
 First, the scale of the data was manipulated by creating three sets of scales.
@@
 Last, the final scale included a \(\mu\) = 1000 (\(\sigma\) = 150) simulating a study that may include response latency data in the milliseconds.
-While there are many potential scales, these three represent a large number of potential variables in the social sciences.
-As we are suggesting item variances as a key factor for estimating sample sizes, the scale of the data is influential on the amount of \emph{potential} variance.
-Smaller ranges of data (1-7) cannot necessarily have the same variance as larger ranges (0-100).
+For the skewed distributions, the item means were set to \(\mu\) = 6, 85, and 2500 respectively with the same \(\sigma\) values around the item means.
+Although there are many potential scales, these three represent a large number of potential variables commonly used in the social sciences.
+As we are suggesting item variances is a key factor for estimating sample sizes, the scale of the data is influential on the amount of potential variance.
+Smaller data ranges (1-7) cannot necessarily have the same variance as larger ranges (0-100).
 
-Next, item variance heterogeneity was included by manipulating the potential \(\sigma\) for each individual item.
+\emph{Item Heterogeneity}.
+Next, item heterogeneity was included by manipulating the potential for each individual item.
 For small scales, the \(\sigma\) = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items.
-For the medium scale of data, \(\sigma\) = 25 with a variance of 4, 8, and 16.
-Last, for the large scale of data, \(\sigma\) = 400 with a variance of 50, 100, and 200 for heterogeneity.
+For the medium scale of the data, \(\sigma\) = 25 with a variance of 4, 8, and 16.
+Finally, for the large scale of the data, \(\sigma\) = 400 with a variance of 50, 100, and 200 for heterogeneity.
 
-\emph{Samples}. Each population was then sampled as if a researcher was conducting a pilot study. The sample sizes started at 20 participants per item increasing in units of 5 up to 100 participants.
-
-\emph{Cutoff Score Criterions}.
-The standard errors of each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as standard error functions as the main component in the formula for normal distribution confidence intervals.
-Standard errors were calculated at each decile of the items up to 90\% (i.e., 0\% smallest SE, 10\% \ldots, 90\% largest SE).
-The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criterions for cutoff scores.
+\emph{Pilot Data Samples}.
+Each of the populations shown in Table \ref{tab:table-sim} was then sampled as if a researcher was conducting a pilot study.
+The sample sizes started at 20 participants per item, increasing in units of 10 up to 100 participants.
+Each of these samples would correspond to Step 1 of the proposed method where a researcher would use pilot data to start their estimation.
+Therefore, the simulations included 3 scales X 3 heterogeneity X 2 normal/skewed X 9 pilot sample sizes representing a potential Step 1 of our procedure.
 
@@
 In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items.
-Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 500 to determine what the new sample size suggestion would be.
+Assuming that each pilot sample represents a dataset that a researcher has collected (Step 1), the standard errors for each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as SE functions as the main component of the formula for normal distribution confidence intervals.
+Standard errors were calculated at each decile of the items up to 90\% (i.e., 0\% smallest SE, 10\% \ldots, 90\% largest SE).
+The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores.
+
+We then simulated samples of 20 to 2000 increasing in units of 20 to determine what the new sample size suggestion would be (Step 3).
 We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds.
-The standard error of each item was calculated for each suggested sample size by pilot sample size by population type.
+However, the sample size simulations were estimated over this amount to determine the pattern of suggested sample sizes (i.e., the function between original sample size and projected sample size).
 
-Next, the percent of items that fall below the cutoff scores, and thus, would be considered ``well-measured'' were calculated for each decile by sample.
-From this data, we pinpoint the smallest suggested sample size at which 80\%, 85\%, 90\%, and 95\% of the items fall below the cutoff criterion.
-These values were chosen as popular measures of ``power'' in which one could determine the minimum suggested sample size (potentially 80\% of items) and the maximum suggested sample size (potentially 90\%).
+Next, the percentage of items that fall below the cutoff scores and therefore would be considered ``well-measured'' were calculated for each decile by sample (Step 4).
+From these data, we pinpoint the smallest suggested sample size at which 80\%, 85\%, 90\% and 95\% of the items fall below the cutoff criterion (Step 5).
+These values were chosen as popular measures of ``power'' in which one could determine the minimum suggested sample size (potentially 80\% of the items) and the maximum suggested sample size (selected from a higher percentage, such as 90\% or 95\%).
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+In order to minimize the potential for random quirks to arise, we simulated the sample selection from the population 100 times and the researcher simulation 100 times for each of those selections.
+This resulted in 1,620,000 simulations of all combinations of variables (i.e., scale of the data, heterogeneity, data skew, pilot study size, researcher simulation size).
+The average of these simulations is presented in the results.
 
-\hypertarget{differences-in-item-variance}{%
-\subsection{Differences in Item Variance}\label{differences-in-item-variance}}
+\hypertarget{simulation-results}{%
+\section{Simulation Results}\label{simulation-results}}
 
-We examined if this procedure is sensitive to differences in item heterogeneity, as we should expect to collect larger samples if we wish to have a large number of items reach a threshold of acceptable variance; potentially, assuring we \emph{could} average them if a researcher did not wish to use a more complex analysis such as multilevel modeling.
+\hypertarget{pilot-data-influence-on-sample-size}{%
+\subsection{Pilot Data Influence on Sample Size}\label{pilot-data-influence-on-sample-size}}
 
-Figure \ref{fig:item-figure} illustrates the potential minimum sample size for 80\% of items to achieve a desired cutoff score.
-The black dots denote the original sample size against the suggested sample size.
-By comparing the facets, we can determine that our suggested procedure does capture the differences in heterogeneity.
-As heterogeneity increases in item variances, the proposed sample size also increases, especially at stricter cutoffs.
-Missing cutoff points where sample sizes proposed would be higher than 500.
+For each variable, the plot of the pilot sample size, projected sample size (i.e., what the simulation suggested), and power levels are presented below.
+The large number of variables means we cannot plot them all simultaneously, and therefore, we averaged the results across other variables for each plot.
+The entire datasets can be examined on our OSF page.
+
+\hypertarget{scale-size}{%
+\subsubsection{Scale Size}\label{scale-size}}
 
@@
 \centering
-\includegraphics{manuscript_draft_files/figure-latex/item-figure-1.pdf}
-\caption{\label{fig:item-figure}Add a good caption here.}
+\includegraphics{manuscript_draft_files/figure-latex/scale-size-figure-1.pdf}
+\caption{\label{fig:scale-size-figure}Simulated pilot sample size and final projected sample size to achieve 80\%, 85\%, 90\%, and 95\% of items below threshold.
+These values are averaged over all other variables.
+Black dots represent original sample size for reference.}
 \end{figure}
+
+Figure \ref{fig:scale-size-figure} demonstrates the influence of scale size on the results separate by potential cutoff level.
+The black dots denote the original sample size for reference.
+Larger scales have more potential variability, and therefore, we see that percent and millisecond scales project a larger required sample size.
+This relationship does not appear to be linear with scale size, as percent scales often represent the highest projected sample size.
+Potentially, this finding is due to the larger proportion of possible variance -- the variance of the item standard deviations / total possible variance -- was largest for percent scales in this set of simulations (\(p_{Percent}\) = .13).
+This finding may be an interaction with heterogeneity, as the Likert scale had the next highest percent variability in item standard errors (\(p_{Likert}\) = .10), followed by milliseconds (\(p_{Milliseconds}\) = .06).
+
+\hypertarget{skew}{%
+\subsubsection{Skew}\label{skew}}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_draft_files/figure-latex/scale-skew-figure-1.pdf}
+\caption{\label{fig:scale-skew-figure}Simulated pilot sample size compared to projected sample size for ceiling versus normal distributions on each scale.}
+\end{figure}
+
+Figure \ref{fig:scale-skew-figure} displays that ceiling distributions, averaged over all other variables, show slightly higher estimates than normal distributions.
+This result is consistent across scale type and heterogeneity, as results indicated that they are often the same or slightly higher for ceiling distributions.
+
+\hypertarget{item-heterogeneity}{%
+\subsubsection{Item Heterogeneity}\label{item-heterogeneity}}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_draft_files/figure-latex/scale-hetero-figure-1.pdf}
+\caption{\label{fig:scale-hetero-figure}Simulated pilot sample size compared to projected sample size for differing amounts of heterogeneity on each scale.}
+\end{figure}
+
+Figure \ref{fig:scale-hetero-figure} displays the results for item heterogeneity for different levels of potential power.
+In this figure, we found that our suggested procedure does capture the differences in heterogeneity.
+As heterogeneity increases in item variances, the proposed sample size also increases.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:table-predict-largest}Prediction of Proposed Sample Size from Simulated Variables}
+
+\begin{tabular}{llllll}
+\toprule
+Term & Estimate & $SE$ & $t$ & $p$ & $pr^2$\\
+\midrule
+Intercept & -27.30 & 3.08 & -8.87 & < .001 & .335\\
+Pilot Sample Size & 1.51 & 0.03 & 54.76 & < .001 & .951\\
+Scale: Likert v Percent & 7.00 & 1.80 & 3.89 & < .001 & .088\\
+Scale: Likert v Milllisecond & 25.63 & 1.87 & 13.74 & < .001 & .548\\
+Proportion Variability & 312.44 & 19.86 & 15.73 & < .001 & .613\\
+Data: Ceiling v Normal & -7.16 & 1.41 & -5.08 & < .001 & .142\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+Using a regression model, we predicted proposed sample size using pilot sample size, scale size, proportion variability (i.e., heterogeneity), and data type (normal, ceiling).
+As shown in Table \ref{tab:table-predict-largest}, the largest influence on proposed sample size is the original pilot sample size, followed by proportion of variance/heterogeneity, and then data and scale sizes.
 
@@
 Larger pilot data is more informative, and therefore, we should expect a lower projected sample size.
-As shown in Figure \ref{fig:sensitive-figure} for only the low variability and small scale data, we do not find this effect.
+As shown in each figure presented already, we do not find this effect.
 These simulations from the pilot data would nearly always suggest a larger sample size - mostly in a linear trend increasing with sample sizes.
@@
 This result does not achieve our second goal.
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_draft_files/figure-latex/sensitive-figure-1.pdf}
-\caption{\label{fig:sensitive-figure}Add good description here.}
-\end{figure}
 
@@
 \includegraphics{manuscript_draft_files/figure-latex/corrected-figure-1.pdf}
-\caption{\label{fig:corrected-figure}A corrected figure update this caption.}
+\caption{\label{fig:corrected-figure}Corrected projected sample sizes for variability and power levels.}
 \end{figure}
@@
 
-To explore if we could recover the new projected sample size from data a researcher would have, we used linear models to create a formula for researcher correction.
-First, the corrected projected sample size was predicted by the original projected sample size.
-Next, the standard deviation of the item standard deviations was added to the equation to recreate heterogeneity estimates.
-The scale of the data is embedded into the standard deviation of the items (\emph{r} = 0.81), and therefore, this variable was not included separately.
-Last, we included the pilot sample size.
+To explore if we could recover the new projected sample size from data a researcher would have, we used regression models to create a formula for researcher correction.
+The researcher employing our procedure would have the possible following variables from their simulations on their (one) pilot dataset: 1) proposed sample size, 2) pilot sample size, 3) estimate of heterogeneity for the items, 4) and the estimate percent of items below the threshold.
+Given the non-linear nature of the correction, we added each variable and it's non-linear \texttt{log2} transform to the regression equation, as this function was used to create the correction.
+The intercept only model was used as a starting point (i.e., \texttt{corrected\ sample\ \textasciitilde{}\ 1}), and then all eight variables (each variable and their \texttt{log2} transform) were entered into a forward stepwise regression to capture the corrected scores with the most predictive values.
+Each variable was entered one at a time using the \texttt{step} function from the \emph{stats} library in \emph{R} (R Core Team, 2022).
 
-The first model using pilot sample size to predict new sample size was significant, \emph{F}(1, 5666) = 48,042.64, \emph{p} \textless{} .001, \(R^2 = .89\), 90\% CI \([0.89, 0.90]\), capturing nearly 90\% of the variance, \(b = 0.62\), 95\% CI \([0.62, 0.63]\).
-The second model with item standard deviation was better then the first model \emph{F}(1, 5665) = 55.21, \emph{p} \textless{} .001, \(R^2 = .89\), 90\% CI \([0.89, 0.90]\).
-The item standard deviation predictor was significant, \(b = 0.02\), 95\% CI \([0.01, 0.03]\), \(t(5665) = 4.54\), \(p < .001\).
-The addition of the original pilot sample size was also significant, \emph{F}(1, 5664) = 9,529.83, \emph{p} \textless{} .001, \(R^2 = .96\), 90\% CI \([0.96, 0.96]\).
-
-As shown in the final model Table \ref{tab:table-correction}, the new suggested sample size is proportional to the original suggested sample size (i.e., \emph{b} \textless{} 1), which reduces the sample size suggestion.
-As variability increases, the suggested sample size also increases to capture differences in heterogeneity shown above; however, this predictor is not significant in the final model, and only contributes a small portion of overall variance.
-Last, in order to correct for large pilot data, the original pilot sample size decreases the new suggested sample size.
-This formula approximation captures 96\% of the variance in sample size scores and should allow a researcher to estimate based on their own data.
+As shown in Table \ref{tab:table-correction}, all variables were included in the final equation, each contributing a significant change to the previous model, as defined by \(\Delta\)AIC \textgreater{} 2 points change between each step of the model.
+Proposed sample size and original sample size were the largest predictors -- unsurprising given the correction formula employed -- followed by the percent ``power'' level and proportion of variance.
+This formula approximation captures \(R^2 = .99\), 90\% CI \([0.99, 0.99]\) of the variance in sample size scores and should allow a researcher to estimate based on their own data, \(F(8, 4,527) = 67,497.54\), \(p < .001\).
+We provide convenience functions in our additional materials to assist researchers in estimating the final adjusted sample size.
 
@@
 
-\begin{tabular}{lllll}
+\begin{tabular}{llllll}
 \toprule
-Term & Estimate & $SD$ & $t$ & $p$\\
+Term & Estimate & $SE$ & $t$ & $p$ & AIC\\
 \midrule
-Intercept & 39.269 & 0.437 & 89.843 & < .001\\
-Projected Sample Size & 0.700 & 0.002 & 366.669 & < .001\\
-Item SD & 0.003 & 0.003 & 0.952 & .341\\
-Pilot Sample Size & -0.695 & 0.007 & -97.621 & < .001\\
+Intercept & 111.049 & 78.248 & 1.419 & .156 & 29,996.94\\
+Projected Sample Size & 0.429 & 0.002 & 185.360 & < .001 & 20,327.79\\
+Pilot Sample Size & -0.718 & 0.007 & -103.787 & < .001 & 14,753.61\\
+Log2 Projected Sample Size & 19.522 & 0.215 & 90.693 & < .001 & 8,668.73\\
+Log2 Pilot Sample Size & 4.655 & 0.269 & 17.296 & < .001 & 8,363.69\\
+Log2 Power & -39.367 & 15.640 & -2.517 & .012 & 8,320.82\\
+Proportion Variability & 15.434 & 3.617 & 4.267 & < .001 & 8,297.71\\
+Log2 Proportion Variability & -0.729 & 0.232 & -3.143 & .002 & 8,289.81\\
+Power & 0.606 & 0.259 & 2.343 & .019 & 8,286.31\\
 \bottomrule
@@
 \hypertarget{choosing-an-appropriate-cutoff}{%
-\subsection{Choosing an Appropriate Cutoff}\label{choosing-an-appropriate-cutoff}}
+\subsection{Choosing an Appropriate cutoff}\label{choosing-an-appropriate-cutoff}}
 
-Last, we examine the question of an appropriate SE decile.
-All graphs for power, heterogeneity, scale, and correction are presented online.
-First, the 0\%, 10\%, and 20\% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale, and heterogeneity.
-If we examine the \(R^2\) values for each decile of our regression equation separately, we find that the 50\% (0.96) represents the best match to our corrected sample size suggestions.
-The 50\% decile, in the corrected format, appears to meet all goals: 1) increases with heterogeneity and scale of data, and 2) higher suggested values for small original samples and a leveling effect at larger pilot data.
-Figure \ref{fig:decile-figure} illustrates the corrected scores for simulations at the 50\% decile recommended cutoff for item standard errors.
+Last, we examined the question of an appropriate SE decile.
+First, the 0\%, 10\%, and 20\% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale size, and heterogeneity.
+If we examine the \(R^2\) values for each decile of our regression equation separately, we find that the values are all \(R^2\) \textgreater{} .99 with very little differences between them.
+Figure \ref{fig:decile-figure} illustrates the corrected scores for simulations at the the 40\% and 50\% decile recommended cutoff for item standard errors.
+For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 40\% decile range, especially for scales with larger potential variance.
+Therefore, we would suggest the 40\% decile to overpower each item to determine the minimum and maximum sample size.
 
@@
 \includegraphics{manuscript_draft_files/figure-latex/decile-figure-1.pdf}
-\caption{\label{fig:decile-figure}A picture of the 50\% cutoff.}
+\caption{\label{fig:decile-figure}Comparison of the cutoffs for 40\% and 50\% deciles across heterogeneity (rows), powering of items (columns), and scale size (shape).}
 \end{figure}
 
-The formula for finding the corrected sample size using a 50\% decile is: \(N_{Corrected Projected} = 39.269 + 0.700 \times X_{N_{Projected}} + 0.003 \times X_{SD Items} - 0.694 \times X_{N_{Pilot}}\).
-The suggested sample size will be estimated from the 80\%, 85\%, 90\%, or 95\% selection at the 50\% decile as shown above.
-The item SD can be calculated directly from the data, and the pilot sample size is the sample size of the data from which a researcher is simulating their samples.
-Therefore, we will recommend the 50\% decile of the item standard errors for step 2 of our suggested simulation procedure, and to correct the projected sample sizes found in step 5 using the correction equation above.
+The final formula for 40\% decile correction is provided in Table \ref{tab:table-decile}. Proportion of variance can be calculated with the following:
+
+\[\frac{SD_{Item SD}}{\sqrt{\frac{(Maximum - Minimum)^2}{4}}}\] where maximum and minimum are the max and min values found in the scale (or the data, if the scale is unbounded).
+This formula would be applied in Step 5 of the proposed procedure.
 While the estimated coefficients could change given variations on our simulation parameters, the general size and pattern of coefficients was consistent, and therefore, we believe this correction equation should work for a variety of use cases.
+We will now demonstrate the final procedure on the examples provided earlier.
 
-\hypertarget{examples}{%
-\section{Examples}\label{examples}}
+\begin{table}[tbp]
 
-In this section, we provide two examples of the suggested procedure.
-The first example includes concreteness ratings from Brysbaert et al.~{[}CITE, 2014{]}.
-Instructions given to participants denoted the difference between concrete (i.e., ``refers to something that exists in reality'') and abstract (i.e., ``something you cannot experience directly through your senses or actions'') terms.
-Participants were then asked to rate concreteness of terms using a 1 (abstract) to 5 (concrete) scale.
-This data represents a small scale dataset that could be used as pilot data for a study using concrete word ratings.
-The data is avaliable at OSF LINK {[}CITE{]}.
-The second dataset includes a large scale dataset with response latencies, the English Lexicon Project {[}CITE, Balota et al.{]}.
-The English Lexicon Project consists of lexical decision response latencies for English words.
-In a lexical decision task, participants simply select ``word'' for real words (e.g., \emph{dog}) and ``nonword'' for pseudowords (e.g., \emph{wug}).
-The trial level data is avaliable here {[}\url{https://elexicon.wustl.edu/}, CITE{]}.
-Critically, in each of these examples, the individual trial level data for each item is available to simulate and calculate standard errors on.
-Data that has been summarized could potentially be used, as long as the original standard deviations for each item were present.
-From the mean and standard deviation for each item, a simulated pilot dataset could be generated for estimating new sample sizes.
-All code to estimate sample sizes is provided on our OSF page.
+\begin{center}
+\begin{threeparttable}
 
-\hypertarget{concreteness-ratings}{%
-\subsection{Concreteness Ratings}\label{concreteness-ratings}}
+\caption{\label{tab:table-decile}Parameters for 40\% Decile Cutoff Scores}
 
-The concreteness ratings data includes 63039 concepts that were rated for their concreteness.
-In our fictional study for this example, we selected 100 random words to show participants.
-In the original study, not every particpant rated every word, which created uneven sample sizes for each word.
-In our random sample of 100 words, the average pilot sample size was 27.96 (\emph{SD} = 1.47), and we will use 28 as our pilot sample size for this example.
-All ``do not know'' ratings were set as missing data.
-The 50\% decile for items standard error was 0.25 for our cutoff criterion.
+\begin{tabular}{lllll}
+\toprule
+Term & Estimate & $SE$ & $t$ & $p$\\
+\midrule
+Intercept & 206.589 & 128.861 & 1.603 & .109\\
+Projected Sample Size & 0.368 & 0.005 & 71.269 & < .001\\
+Pilot Sample Size & -0.770 & 0.013 & -59.393 & < .001\\
+Log2 Projected Sample Size & 27.541 & 0.552 & 49.883 & < .001\\
+Log2 Pilot Sample Size & 2.583 & 0.547 & 4.725 & < .001\\
+Log2 Power & -66.151 & 25.760 & -2.568 & .010\\
+Proportion Variability & 16.405 & 6.005 & 2.732 & .006\\
+Log2 Proportion Variability & -1.367 & 0.382 & -3.577 & < .001\\
+Power & 1.088 & 0.426 & 2.552 & .011\\
+\bottomrule
+\end{tabular}
 
-The pilot data was then simulated, with replacement, with samples from 20 to 300 increasing in units of 5.
-On each sample, the percent of items below the cutoff score were calculated.
-After applying our correction equation, we find that a sample size of 44 would allow for at least 80\% of items to meet the cutoff criterion.
-The sample sizes for 85\% (48), 90\% (48), and 95\% (51) are also options for sample size suggestions.
-Finally, we calculated the potential amount of data retention given that participants could indicate they did not know a word (\(M_{correct}\) = 0.82, \emph{SD} = 0.24).
-In order to account for this facet, the potential sample sizes were multiplied by 1/0.82, which results in a suggested sample of 54, 59, and 63.
-Therefore, we could designate our minimum sample per item as 54, stopping rule of 0.25, and maximum sample size of 63.
+\end{threeparttable}
+\end{center}
 
-\hypertarget{response-latencies}{%
-\subsection{Response Latencies}\label{response-latencies}}
+\end{table}
 
-The ELP response latency data includes 80962 word-forms, 40481 that are listed as non-words, and 40481 real words.
-For our example study, we will randomly select 500 real words and 500 non-words to show participants.
-The average pilot sample size for this random sample was 32.68 (\emph{SD} = 0.64), and \emph{n} = 33 will be our pilot size for this example.
-Again, participants are expected to make mistakes, and we calculated percent correct as 0.86, which was roughly even in the two stimulus categories: \(M_{word}\) = 0.84 and \(M_{non-word}\) = 0.87.
-The 50\% decile for items standard error was 61.32 for our cutoff criterion.
-We additionally checked to ensure that the two stimulus types did not have very different cutoff criterions: 50\% decile \(SE_{words}\) = 56.94, 50\% decile \(SE_{nonwords}\) = 65.61.
-In this scenario, we could chose to go with the lower SE to be more conservative (i.e., higher projected sample size).
-Given the values were close for large scale data, we used the 50\% decile of all stimuli taken together.
+\hypertarget{updated-example}{%
+\section{Updated Example}\label{updated-example}}
 
-The pilot response latency data was then simulated in the same way as described above.
-After calculating the percent below our cutoff score, we applied the correction to the projected sample sizes.
-A sample size of 31 would equate to 80\% of the items reaching our cutoff, along with 85\% (34), 90\% (34), and 95\% (38).
-Again, we adjusted for data loss given that participants are expected to incorrectly answer items, resulting in a suggested sample of 36, 40, and 44.
-One other possible consideration for this study is potential fatigue in showing participants 1000 target items.
-Therefore, we could designate in our research design that each participant will only receive 500 of the target items.
-We would need to double our sample sizes to account for splitting of the items across multiple sets of participants.
-Our minimum sample size for the entire study could be 72, stopping rule of 61.32, and maximum sample size of 88.
-This study would benefit from an adaptive design, where smaller sets items are randomly sampled for participants until they reach the minimum sample size or the cutoff criteria.
-At this point, items are probablistically sampled (e.g., higher selection probability for items that have not reached a minimum or stopping rule) until all items have reached criteria.
+The updated proposal steps are in Table \ref{tab:table-summary}.
+The main change occurs in Step 2 with a designated cutoff decile, and Step 5 with a correction score.
+Using the data from the 40\% decile in Table \ref{tab:table-example}, we can determine that the stopping rule for concreteness ratings would be 0.18, and the stopping rule for lexical decision times would be 51.32.
+For Step 5, we apply our correction formula separately for each one, as they have different variability scores, and these scores are shown in Table \ref{tab:table-scores-updated}.
+Each row was multiplied by row one's formula, and then these scores are summed for the final sample size.
+Sample sizes cannot be proportional, so we recommend rounding up to the nearest whole number.
+
+For one additional consideration, we calculated the potential amount of data retention given that participants could indicate they did not know a word (\(M_{answered}\) = 0.90, \emph{SD} = 0.13) in the concreteness task or answer a trial incorrectly in the lexical decision task (\(M_{correct}\) = 0.78, \emph{SD} = 0.22).
+In order to account for this facet, the potential sample sizes were multiplied by \(\frac{1}{p_{retained}}\) where the denominator is proportion retained for each task.
 
 \hypertarget{additional-materials}{%
-\subsection{Additional Materials}\label{additional-materials}}
+\section{Additional Materials}\label{additional-materials}}
+
+\hypertarget{package}{%
+\subsection{Package}\label{package}}
+
+We have developed functions to implement the suggested procedure as part of an upcoming package \texttt{semanticprimeR}.
+You can install the package from GitHub using: \texttt{devtools::install\_github("SemanticPriming/semanticprimeR")}.
+We detail the functions below by proposed step in the process.
+
+\emph{Step 1}.
+Ideally, researchers would have pilot data that represented their proposed data collection.
+This data should be formatted in long format wherein each row represents the score from an item by participant, rather than wide format wherein each column represents an item and each row represents a single participant.
+The \texttt{tidyr::pivot\_longer()} or \texttt{reshape::melt()} functions can be used to reformat wide data.
+If no pilot data is available, the \texttt{simulate\_population()} function can be used with the following arguments (and example numbers, * indicates optional).
+This function will return a dataframe with the simulated normal values for each item.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# devtools::install\_github("SemanticPriming/semanticprimeR")}
+\FunctionTok{library}\NormalTok{(semanticprimeR)}
+\NormalTok{pops }\OtherTok{\textless{}{-}} \FunctionTok{simulate\_population}\NormalTok{(}\AttributeTok{mu =} \DecValTok{4}\NormalTok{, }\CommentTok{\# item means}
+  \AttributeTok{mu\_sigma =}\NormalTok{ .}\DecValTok{2}\NormalTok{, }\CommentTok{\# variability in item means }
+  \AttributeTok{sigma =} \DecValTok{2}\NormalTok{, }\CommentTok{\# item standard deviations}
+  \AttributeTok{sigma\_sigma =}\NormalTok{ .}\DecValTok{2}\NormalTok{, }\CommentTok{\# standard deviation of the standard deviations}
+  \AttributeTok{number\_items =} \DecValTok{30}\NormalTok{, }\CommentTok{\# number of items}
+  \AttributeTok{number\_scores =} \DecValTok{20}\NormalTok{, }\CommentTok{\# number of participants}
+  \AttributeTok{smallest\_sigma =}\NormalTok{ .}\DecValTok{02}\NormalTok{, }\CommentTok{\#* smallest possible standard deviation}
+  \AttributeTok{min\_score =} \DecValTok{1}\NormalTok{, }\CommentTok{\#* minimum score for truncating purposes}
+  \AttributeTok{max\_score =} \DecValTok{7}\NormalTok{, }\CommentTok{\#* maximum score for truncating purposes}
+  \AttributeTok{digits =} \DecValTok{0}\NormalTok{) }\CommentTok{\#* number of digits for rounding}
+
+\FunctionTok{head}\NormalTok{(pops)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##   item score
+## 1    1     3
+## 2    2     2
+## 3    3     7
+## 4    4     2
+## 5    5     5
+## 6    6     3
+\end{verbatim}
+
+\emph{Step 2}.
+In step 2, we can use \texttt{calculate\_cutoff()} to calculate the standard error of the items, the standard deviation of the standard errors and the corresponding proportion of variance possible, and the 40\% decile cutoff score.
+The \texttt{pops} dataframe can be used in this function, which has columns named \texttt{item} for the item labels (i.e., 1, 2, 3, 4 or characters can be used), and \texttt{score} for the dependent variable.
+This function returns a list of values to be used in subsequent steps.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cutoff }\OtherTok{\textless{}{-}} \FunctionTok{calculate\_cutoff}\NormalTok{(}\AttributeTok{population =}\NormalTok{ pops, }\CommentTok{\# pilot data or simulated data}
+  \AttributeTok{grouping\_items =} \StringTok{"item"}\NormalTok{, }\CommentTok{\# name of the item indicator column}
+  \AttributeTok{score =} \StringTok{"score"}\NormalTok{, }\CommentTok{\# name of the dependent variable column}
+  \AttributeTok{minimum =} \DecValTok{1}\NormalTok{, }\CommentTok{\# minimum possible/found score}
+  \AttributeTok{maximum =} \DecValTok{7}\NormalTok{) }\CommentTok{\# maximum possible/found score}
+
+\NormalTok{cutoff}\SpecialCharTok{$}\NormalTok{se\_items }\CommentTok{\# all standard errors of items}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##  [1] 0.2926737 0.4376973 0.3966969 0.4639646 0.4308804 0.4729527 0.3973597
+##  [8] 0.3734618 0.3439324 0.3933660 0.3346247 0.3879772 0.4466248 0.3324550
+## [15] 0.4530598 0.4500000 0.4660867 0.3590924 0.2523573 0.4345294 0.4844965
+## [22] 0.4805425 0.4167544 0.3234274 0.2926737 0.3371709 0.3838859 0.4285840
+## [29] 0.3802700 0.2325488
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cutoff}\SpecialCharTok{$}\NormalTok{sd\_items }\CommentTok{\# standard deviation of the standard errors}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 0.06798869
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cutoff}\SpecialCharTok{$}\NormalTok{cutoff }\CommentTok{\# 40\% decile score}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##       40% 
+## 0.3824396
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cutoff}\SpecialCharTok{$}\NormalTok{prop\_var }\CommentTok{\# proportion of possible variance }
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 0.0226629
+\end{verbatim}
+
+\emph{Step 3}.
+The \texttt{bootstrap\_samples()} function creates bootstrapped samples from the pilot or simulated population data to create samples to estimate the number of participants needed for item standard error to be below the cutoff calculated in Step 2.
+This function returns a list of samples with sizes that start at the \texttt{start} size, increase by \texttt{increase}, and end with the \texttt{stop} sample size.
+The population or pilot data will be included in \texttt{population}, and the item column indicator should be included in \texttt{grouping\_items}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{samples }\OtherTok{\textless{}{-}} \FunctionTok{bootstrap\_samples}\NormalTok{(}\AttributeTok{start =} \DecValTok{20}\NormalTok{, }\CommentTok{\# starting sample size}
+  \AttributeTok{stop =} \DecValTok{100}\NormalTok{, }\CommentTok{\# stopping sample size}
+  \AttributeTok{increase =} \DecValTok{5}\NormalTok{, }\CommentTok{\# increase bootstrapped samples by this amount}
+  \AttributeTok{population =}\NormalTok{ pops, }\CommentTok{\# population or pilot data}
+  \AttributeTok{replace =} \ConstantTok{TRUE}\NormalTok{, }\CommentTok{\# bootstrap with replacement? }
+  \AttributeTok{grouping\_items =} \StringTok{"item"}\NormalTok{) }\CommentTok{\# item column label  }
+
+\FunctionTok{head}\NormalTok{(samples[[}\DecValTok{1}\NormalTok{]])}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 6 x 2
+## # Groups:   item [1]
+##    item score
+##   <int> <dbl>
+## 1     1     1
+## 2     1     4
+## 3     1     1
+## 4     1     1
+## 5     1     2
+## 6     1     5
+\end{verbatim}
+
+\emph{Step 4 and 5}.
+The proportion of bootstrapped items across sample sizes below the cutoff score can then be calculated using \texttt{calculate\_proportion()}.
+This function returns a dataframe of each sample size with the proportion of items below that cutoff to use in the next function.
+The \texttt{samples} and \texttt{cutoff} were calculated with our previous functions, with the column for item labels and dependent variable to ensure the right calculations.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{proportion\_summary }\OtherTok{\textless{}{-}} \FunctionTok{calculate\_proportion}\NormalTok{(}\AttributeTok{samples =}\NormalTok{ samples, }\CommentTok{\# samples list}
+  \AttributeTok{cutoff =}\NormalTok{ cutoff}\SpecialCharTok{$}\NormalTok{cutoff, }\CommentTok{\# cut off score }
+  \AttributeTok{grouping\_items =} \StringTok{"item"}\NormalTok{, }\CommentTok{\# item column name}
+  \AttributeTok{score =} \StringTok{"score"}\NormalTok{) }\CommentTok{\# dependent variable column name }
+
+\FunctionTok{head}\NormalTok{(proportion\_summary)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 6 x 2
+##   percent_below sample_size
+##           <dbl>       <dbl>
+## 1         0.467          20
+## 2         0.533          25
+## 3         0.967          30
+## 4         0.967          35
+## 5         1              40
+## 6         1              45
+\end{verbatim}
+
+\emph{Step 6}.
+Last, we use the \texttt{calculate\_correction()} function to correct the sample size scores given the proposed correction formula.
+The \texttt{proportion\_summary} from above is used in this function, along with required information about the sample size, proportion of variance from our cutoff calculation, and what power levels should be calculated.
+Note that the exact percent of items below a cutoff score will be returned, if the values in \texttt{power\_levels} are not exactly present.
+The final summary presents the smallest sample size, corrected, for each of the potential power levels.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{corrected\_summary }\OtherTok{\textless{}{-}} \FunctionTok{calculate\_correction}\NormalTok{(}
+  \AttributeTok{proportion\_summary =}\NormalTok{ proportion\_summary, }\CommentTok{\# prop from above}
+  \AttributeTok{pilot\_sample\_size =} \DecValTok{20}\NormalTok{, }\CommentTok{\# number of participants in the pilot data }
+  \AttributeTok{proportion\_variability =}\NormalTok{ cutoff}\SpecialCharTok{$}\NormalTok{prop\_var, }\CommentTok{\# proportion variance from cutoff scores}
+  \AttributeTok{power\_levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{80}\NormalTok{, }\DecValTok{85}\NormalTok{, }\DecValTok{90}\NormalTok{, }\DecValTok{95}\NormalTok{)) }\CommentTok{\# what levels of power to calculate }
+
+\NormalTok{corrected\_summary}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 4 x 3
+##   percent_below sample_size corrected_sample_size
+##           <dbl>       <dbl>                 <dbl>
+## 1          96.7          30                  25.3
+## 2          96.7          30                  25.3
+## 3          96.7          30                  25.3
+## 4          96.7          30                  25.3
+\end{verbatim}
+
+\hypertarget{vignettes}{%
+\subsection{Vignettes}\label{vignettes}}
 
 While the examples in this manuscript are traditionally cognitive linguistics focused, any research using repeated items can benefit from newer sampling techniques.
-Therefore, we provide XX example vignettes on our OSF page/GitHub site for this manuscript across a range of examples of data types provided by the authors of this manuscript.
-Examples include \ldots{} ADD HERE AFTER DONE WITH VIGNETTES.
+Therefore, we provide 12 example vignettes and varied code examples on our OSF page/GitHub site for this manuscript across a range of data types provided by the authors of this manuscript.
+Examples include psycholinguistics (De Deyne et al., 2008; Heyman et al., 2014; Montefinese et al., 2022), eye tracking data (Ulloa et al., 2014), social psychology (Grahe et al., 2022; Peterson et al., 2022), COVID related data (Montefinese et al., 2021), and cognitive psychology (Barzykowski et al., 2019; Errington et al., 2021; Röer et al., 2013).
 
@@
 
-To be added once people say this idea doesn't suck.
+We proposed and demonstrated examples for a method using AIPE, bootstrapping, and simulation to estimate a minimum and maximum sample size along with a rule for stopping data collection based on narrow confidence intervals on the parameter of interest.
+We contend that this procedure is specifically useful for studies with multiple items that intend on using item level focused analyses; furthermore, the utility of measuring each item well can extend to many analysis choices.
+By focusing on collecting quality data, we can suggest that the data is useful, regardless of the outcome of any hypothesis test.
+
+One limitation of these methods would be our decision to use datasets with very large numbers of items to simulate what might happen within one study.
+For example, the English Lexicon Project includes thousands of items, and by the time we would simulate for all of those, it would likely suggest needing thousands of participants for most items to reach the criterion.
+Additionally, as the number of items increases, you may also see very small estimates for sample size due to the correction factor (as with large numbers of items, you could find many items with standard errors below the 40\% decile).
+Therefore, it would be beneficial to consider only simulating what a participant would reasonably complete in a study.
+Small numbers of repeated items usually result in larger sample sizes proposed from the original pilot data.
+This result occurs because the smaller number of items means more samples for nearly all to reach the cutoff criteria.
+These results are not too different from what we might expect for a power analysis using a multilevel model - larger numbers of items tend to decrease necessary sample size, while smaller numbers of items tend to increase sample size.
+
+Second, these methods do not ensure the normal interpretation of power, where you know that you would find a specific effect for a specific test, \(\alpha\), and so on.
+As discussed in the Introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis, many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
+These methods could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
+
+Researchers should consider this hybrid approach for AIPE, bootstrapping, and simulation as a tool for hypothesis testing and parameter estimation.
+The proposed procedure can be beneficial for many different research studies, specifically replication studies, that usually depend on subject sample size but rarely item sample size, in spite of the fact that item sample sizes contribute to language processing (Baayen et al., 2008; Brysbaert \& Stevens, 2018).
+These observed effects can be then replicated and, as a result of several replications, can be applied to meta-analyses.
+As a result, analysts would be able to use the accuracy and high statistical power to calculate the parameters to assess whether the effect is genuine.
+This article helps to achieve this goal by encouraging researchers to conduct studies where the power analysis is not based on the size of the effect but on adequate sampling of the stimuli.
+We argue that this article can be the initial step to apply AIPE in a manner that can allow researchers to use item information to provide a more accurate and statistically reliable measure of the effect we aimed to investigate.
+In conclusion, item power analysis is a tool to avoid the waste of resources while ensuring that adequately measured items can be achieved.
+Well measured data can enable us to counteract the literature that contains false positives, allowing us to achieve replicable, high-quality science to establish answers to scientific questions with precision and accuracy.
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-alexander1994}{}}%
+Alexander, R.
+A., \& DeShon, R.
+P.
+(1994).
+Effect of error variance heterogeneity on the power of tests for regression slope differences.
+\emph{Psychological Bulletin}, \emph{115}(2), 308--314.
+\url{https://doi.org/10.1037/0033-2909.115.2.308}
+
+\leavevmode\vadjust pre{\hypertarget{ref-anderson2017}{}}%
+Anderson, S.
+F., Kelley, K., \& Maxwell, S.
+E.
+(2017).
+Sample-Size Planning for More Accurate Statistical Power: A Method Adjusting Sample Effect Sizes for Publication Bias and Uncertainty.
+\emph{Psychological Science}, \emph{28}(11), 1547--1562.
+\url{https://doi.org/10.1177/0956797617723724}
+
+\leavevmode\vadjust pre{\hypertarget{ref-aust2022}{}}%
+Aust, F., Barth, M., Diedenhofen, B., Stahl, C., Casillas, J.
+V., \& Siegel, R.
+(2022).
+\emph{Papaja: Prepare american psychological association journal articles with r markdown}.
+\url{https://CRAN.R-project.org/package=papaja}
+
+\leavevmode\vadjust pre{\hypertarget{ref-baayen2008}{}}%
+Baayen, R.
+H., Davidson, D.
+J., \& Bates, D.
+M.
+(2008).
+Mixed-effects modeling with crossed random effects for subjects and items.
+\emph{Journal of Memory and Language}, \emph{59}(4), 390--412.
+\url{https://doi.org/10.1016/j.jml.2007.12.005}
+
+\leavevmode\vadjust pre{\hypertarget{ref-balota2007}{}}%
+Balota, D.
+A., Yap, M.
+J., Hutchison, K.
+A., Cortese, M.
+J., Kessler, B., Loftis, B., Neely, J.
+H., Nelson, D.
+L., Simpson, G.
+B., \& Treiman, R.
+(2007).
+The English Lexicon Project.
+\emph{Behavior Research Methods}, \emph{39}(3), 445--459.
+\url{https://doi.org/10.3758/BF03193014}
+
+\leavevmode\vadjust pre{\hypertarget{ref-barzykowski2019}{}}%
+Barzykowski, K., Niedźwieńska, A., \& Mazzoni, G.
+(2019).
+How intention to retrieve a memory and expectation that a memory will come to mind influence the retrieval of autobiographical memories.
+\emph{Consciousness and Cognition}, \emph{72}, 31--48.
+\url{https://doi.org/10.1016/j.concog.2019.03.011}
+
+\leavevmode\vadjust pre{\hypertarget{ref-batterham2005}{}}%
+Batterham, A.
+M., \& Atkinson, G.
+(2005).
+How big does my sample need to be?
+A primer on the murky world of sample size estimation.
+\emph{Physical Therapy in Sport}, \emph{6}(3), 153--163.
+\url{https://doi.org/10.1016/j.ptsp.2005.05.004}
+
+\leavevmode\vadjust pre{\hypertarget{ref-beribisky2019}{}}%
+Beribisky, N. (2019). \emph{A Multi-Faceted Mess: A Review of Statistical Power Analysis in Psychology Journal Articles}. \url{https://yorkspace.library.yorku.ca/xmlui/handle/10315/36719}
+
+\leavevmode\vadjust pre{\hypertarget{ref-brysbaert2019}{}}%
+Brysbaert, M.
+(2019).
+How Many Participants Do We Have to Include in Properly Powered Experiments?
+A Tutorial of Power Analysis with Reference Tables.
+\emph{Journal of Cognition}, \emph{2}(1), 16.
+\url{https://doi.org/10.5334/joc.72}
+
+\leavevmode\vadjust pre{\hypertarget{ref-brysbaert2018}{}}%
+Brysbaert, M., \& Stevens, M. (2018). Power Analysis and Effect Size in Mixed Effects Models: A Tutorial. \emph{Journal of Cognition}, \emph{1}(1), 9. \url{https://doi.org/10.5334/joc.10}
+
+\leavevmode\vadjust pre{\hypertarget{ref-brysbaert2014}{}}%
+Brysbaert, M., Warriner, A.
+B., \& Kuperman, V.
+(2014).
+Concreteness ratings for 40 thousand generally known English word lemmas.
+\emph{Behavior Research Methods}, \emph{46}(3), 904--911.
+\url{https://doi.org/10.3758/s13428-013-0403-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-buchanan2019}{}}%
+Buchanan, E.
+M., Valentine, K.
+D., \& Maxwell, N.
+P.
+(2019).
+LAB: Linguistic Annotated Bibliography {\textendash} a searchable portal for normed database information.
+\emph{Behavior Research Methods}, \emph{51}(4), 1878--1888.
+\url{https://doi.org/10.3758/s13428-018-1130-8}
+
+\leavevmode\vadjust pre{\hypertarget{ref-chalmers2020}{}}%
+Chalmers, R.
+P., \& Adkins, M.
+C.
+(2020).
+Writing effective and reliable monte carlo simulations with the SimDesign package.
+\emph{The Quantitative Methods for Psychology}, \emph{16}(4), 248--280.
+\url{https://doi.org/10.20982/tqmp.16.4.p248}
+
+\leavevmode\vadjust pre{\hypertarget{ref-d.chambers2014}{}}%
+Chambers, C.
+D., Feredoes, E., D.
+Muthukumaraswamy, S., J.
+Etchells, P., \& 1 Cardiff University Brain Research Imaging Centre, School of Psychology, Cardiff University; (2014).
+Instead of {``}playing the game{''} it is time to change the rules: Registered Reports at AIMS Neuroscience and beyond.
+\emph{AIMS Neuroscience}, \emph{1}(1), 4--17.
+\url{https://doi.org/10.3934/Neuroscience.2014.1.4}
+
+\leavevmode\vadjust pre{\hypertarget{ref-champely2017}{}}%
+Champely, S., Ekstrom, C., Dalgaard, P., Gill, J., Weibelzahl, S., Anandkumar, A., Ford, C., Volcic, R., \& De Rosario, H. (2017). \emph{Pwr: Basic functions for power analysis}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cohen1990}{}}%
+Cohen, J. (1990). Things I have learned (so far). \emph{American Psychologist}, \emph{45}(12), 1304--1312. \url{https://doi.org/10.1037/0003-066X.45.12.1304}
+
+\leavevmode\vadjust pre{\hypertarget{ref-coretta}{}}%
+Coretta, S., Casillas, J.
+V., Roessig, S., Franke, M., Ahn, B., Al-Hoorie, A.
+H., Al-Tamimi, J., Alotaibi, N.
+E., AlShakhori, M.
+K., Altmiller, R.
+M., Arantes, P., Athanasopoulou, A., Baese-Berk, M.
+M., Bailey, G., Sangma, C.
+B.
+A., Beier, E.
+J., Benavides, G.
+M., Benker, N., BensonMeyer, E.
+P., \ldots{} Roettger, T.
+B.
+(2023).
+Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human-speech analyses.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{6}(3), 25152459231162567.
+\url{https://doi.org/10.1177/25152459231162567}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dedeyne2008}{}}%
+De Deyne, S., Verheyen, S., Ameel, E., Vanpaemel, W., Dry, M.
+J., Voorspoels, W., \& Storms, G.
+(2008).
+Exemplar by feature applicability matrices and other Dutch normative data for semantic concepts.
+\emph{Behavior Research Methods}, \emph{40}(4), 1030--1048.
+\url{https://doi.org/10.3758/brm.40.4.1030}
+
+\leavevmode\vadjust pre{\hypertarget{ref-debruine2021}{}}%
+DeBruine, L. (2021). \emph{Faux: Simulation for factorial designs}. Zenodo. \url{https://doi.org/10.5281/ZENODO.2669586}
+
+\leavevmode\vadjust pre{\hypertarget{ref-efron2000}{}}%
+Efron, B. (2000). The bootstrap and modern statistics. \emph{Journal of the American Statistical Association}, \emph{95}(452), 1293--1296. \url{https://doi.org/10.1080/01621459.2000.10474333}
+
+\leavevmode\vadjust pre{\hypertarget{ref-erdfelder1996}{}}%
+Erdfelder, E., Faul, F., \& Buchner, A.
+(1996).
+GPOWER: A general power analysis program.
+\emph{Behavior Research Methods, Instruments, \& Computers}, \emph{28}(1), 1--11.
+\url{https://doi.org/10.3758/BF03203630}
+
+\leavevmode\vadjust pre{\hypertarget{ref-10.7554ux2feLife.71601}{}}%
+Errington, T.
+M., Mathur, M., Soderberg, C.
+K., Denis, A., Perfito, N., Iorns, E., \& Nosek, B.
+A.
+(2021).
+Investigating the replicability of preclinical cancer biology.
+\emph{eLife}, \emph{10}, e71601.
+\url{https://doi.org/10.7554/eLife.71601}
+
+\leavevmode\vadjust pre{\hypertarget{ref-faul2007}{}}%
+Faul, F., Erdfelder, E., Lang, A.-G., \& Buchner, A.
+(2007).
+G*Power 3: A flexible statistical power analysis program for the social, behavioral, and biomedical sciences.
+\emph{Behavior Research Methods}, \emph{39}(2), 175--191.
+\url{https://doi.org/10.3758/BF03193146}
+
+\leavevmode\vadjust pre{\hypertarget{ref-grahe2022}{}}%
+Grahe, J., Chalk, H., Cramblet Alvarez, L., Faas, C., Hermann, A., McFall, J., \& Molyneux, K. (2022). EAMMi2 public data. \emph{Open Science Framework}. \url{https://doi.org/10.17605/OSF.IO/X7MP2}
+
+\leavevmode\vadjust pre{\hypertarget{ref-halpern2002}{}}%
+Halpern, S. D. (2002). The Continuing Unethical Conduct of Underpowered Clinical Trials. \emph{JAMA}, \emph{288}(3), 358. \url{https://doi.org/10.1001/jama.288.3.358}
+
+\leavevmode\vadjust pre{\hypertarget{ref-doi:10.1177ux2f1745691620979806}{}}%
+Hardwicke, T.
+E., Thibault, R.
+T., Kosie, J.
+E., Wallach, J.
+D., Kidwell, M.
+C., \& Ioannidis, J.
+P.
+A.
+(2022).
+Estimating the prevalence of transparency and reproducibility-related research practices in psychology (2014{\textendash}2017).
+\emph{Perspectives on Psychological Science}, \emph{17}(1), 239--251.
+\url{https://doi.org/10.1177/1745691620979806}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hardwicke2020}{}}%
+Hardwicke, T.
+E., Wallach, J.
+D., Kidwell, M.
+C., Bendixen, T., Crüwell, S., \& Ioannidis, J.
+P.
+A.
+(2020).
+An empirical assessment of transparency and reproducibility-related research practices in the social sciences (2014{\textendash}2017).
+\emph{Royal Society Open Science}, \emph{7}(2), 190806.
+\url{https://doi.org/10.1098/rsos.190806}
+
+\leavevmode\vadjust pre{\hypertarget{ref-heyman2014}{}}%
+Heyman, T., De Deyne, S., Hutchison, K.
+A., \& Storms, G.
+(2014).
+Using the speeded word fragment completion task to examine semantic priming.
+\emph{Behavior Research Methods}, \emph{47}(2), 580--606.
+\url{https://doi.org/10.3758/s13428-014-0496-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kelley2007}{}}%
+Kelley, K.
+(2007).
+Sample size planning for the coefficient of variation from the accuracy in parameter estimation approach.
+\emph{Behavior Research Methods}, \emph{39}(4), 755--766.
+\url{https://doi.org/10.3758/BF03192966}
+
+\leavevmode\vadjust pre{\hypertarget{ref-maxwell2008}{}}%
+Maxwell, S.
+E., Kelley, K., \& Rausch, J.
+R.
+(2008).
+Sample size planning for statistical power and accuracy in parameter estimation.
+\emph{Annual Review of Psychology}, \emph{59}, 537--563.
+\url{https://doi.org/10.1146/annurev.psych.59.103006.093735}
+
+\leavevmode\vadjust pre{\hypertarget{ref-meyer1971}{}}%
+Meyer, D.
+E., \& Schvaneveldt, R.
+W.
+(1971).
+Facilitation in recognizing pairs of words: Evidence of a dependence between retrieval operations.
+\emph{Journal of Experimental Psychology}, \emph{90}(2), 227--234.
+\url{https://doi.org/10.1037/h0031564}
+
+\leavevmode\vadjust pre{\hypertarget{ref-montefinese2021}{}}%
+Montefinese, M., Ambrosini, E., \& Angrilli, A.
+(2021).
+Online search trends and word-related emotional response during COVID-19 lockdown in Italy: a cross-sectional online study.
+\emph{PeerJ}, \emph{9}, e11858.
+\url{https://doi.org/10.7717/peerj.11858}
+
+\leavevmode\vadjust pre{\hypertarget{ref-montefinese2022}{}}%
+Montefinese, M., Vinson, D., Vigliocco, G., \& Ambrosini, E.
+(2022).
+Italian age of acquisition norms for a large set of words (ItAoA).
+\emph{Open Science Framework}.
+\url{https://doi.org/10.17605/OSF.IO/3TRG2}
+
+\leavevmode\vadjust pre{\hypertarget{ref-nosek2014}{}}%
+Nosek, B.
+A., \& Lakens, D.
+(2014).
+Registered Reports: A Method to Increase the Credibility of Published Results.
+\emph{Social Psychology}, \emph{45}(3), 137--141.
+\url{https://doi.org/10.1027/1864-9335/a000192}
+
+\leavevmode\vadjust pre{\hypertarget{ref-nuijten2016}{}}%
+Nuijten, M.
+B., Hartgerink, C.
+H.
+J., Assen, M.
+A.
+L.
+M.
+van, Epskamp, S., \& Wicherts, J.
+M.
+(2016).
+The prevalence of statistical reporting errors in psychology (1985{\textendash}2013).
+\emph{Behavior Research Methods}, \emph{48}(4), 1205--1226.
+\url{https://doi.org/10.3758/s13428-015-0664-2}
+
+\leavevmode\vadjust pre{\hypertarget{ref-opensciencecollaboration2015}{}}%
+Open Science Collaboration. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716--aac4716. \url{https://doi.org/10.1126/science.aac4716}
+
+\leavevmode\vadjust pre{\hypertarget{ref-peterson2022}{}}%
+Peterson, J.
+C., Uddenberg, S., Griffiths, T.
+L., Todorov, A., \& Suchow, J.
+W.
+(2022).
+Deep models of superficial face judgments.
+\emph{Proceedings of the National Academy of Sciences}, \emph{119}(17).
+\url{https://doi.org/10.1073/pnas.2115228119}
+
+\leavevmode\vadjust pre{\hypertarget{ref-R-base}{}}%
+R Core Team. (2022). \emph{R: A language and environment for statistical computing}. \url{https://www.R-project.org/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-rheinheimer2001}{}}%
+Rheinheimer, D.
+C., \& Penfield, D.
+A.
+(2001).
+The effects of type i error rate and power of the ANCOVA f test and selected alternatives under nonnormality and variance heterogeneity.
+\emph{The Journal of Experimental Education}, \emph{69}(4), 373--391.
+\url{https://doi.org/10.1080/00220970109599493}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ruxf6er2013}{}}%
+Röer, J.
+P., Bell, R., \& Buchner, A.
+(2013).
+Is the survival-processing memory advantage due to richness of encoding?
+\emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{39}(4), 1294--1302.
+\url{https://doi.org/10.1037/a0031214}
+
+\leavevmode\vadjust pre{\hypertarget{ref-rosenthal1979}{}}%
+Rosenthal, R. (1979). The file drawer problem and tolerance for null results. \emph{Psychological Bulletin}, \emph{86}(3), 638--641. \url{https://doi.org/10.1037/0033-2909.86.3.638}
+
+\leavevmode\vadjust pre{\hypertarget{ref-rousselet}{}}%
+Rousselet, G., Pernet, D.
+C., \& Wilcox, R.
+R.
+(2022).
+An introduction to the bootstrap: A versatile method to make inferences by using data-driven simulations.
+\emph{Meta-Psychology}.
+\url{https://doi.org/10.31234/osf.io/h8ft7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-silberzahn2018many}{}}%
+Silberzahn, R., Uhlmann, E.
+L., Martin, D.
+P., Anselmi, P., Aust, F., Awtrey, E., Bahník, Š., Bai, F., Bannard, C., Bonnier, E., \& others.
+(2018).
+Many analysts, one data set: Making transparent how variations in analytic choices affect results.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(3), 337356.
+
+\leavevmode\vadjust pre{\hypertarget{ref-simmons2011}{}}%
+Simmons, J.
+P., Nelson, L.
+D., \& Simonsohn, U.
+(2011).
+False-positive psychology: Undisclosed flexibility in data collection and analysis allows presenting anything as significant.
+\emph{Psychological Science}, \emph{22}(11), 1359--1366.
+\url{https://doi.org/10.1177/0956797611417632}
+
+\leavevmode\vadjust pre{\hypertarget{ref-stewart2020}{}}%
+Stewart, S., Rinke, E.
+M., McGarrigle, R., Lynott, D., Lunny, C., Lautarescu, A., Galizzi, M.
+M., Farran, E.
+K., \& Crook, Z.
+(2020).
+\emph{Pre-registration and registered reports: A primer from UKRN}.
+\url{https://doi.org/10.31219/osf.io/8v2n7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ulloa2014}{}}%
+Ulloa, J.
+L., Marchetti, C., Taffou, M., \& George, N.
+(2014).
+Only your eyes tell me what you like: Exploring the liking effect induced by other's gaze.
+\emph{Cognition and Emotion}, \emph{29}(3), 460--470.
+\url{https://doi.org/10.1080/02699931.2014.919899}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vazire2018}{}}%
+Vazire, S.
+(2018).
+Implications of the Credibility Revolution for Productivity, Creativity, and Progress.
+\emph{Perspectives on Psychological Science}, \emph{13}(4), 411--417.
+\url{https://doi.org/10.1177/1745691617751884}
+
 \end{CSLReferences}

### 2023-09-26 manuscript_draft.Rmd
fixing manuscript table issue

@@
 
-```{r table-scores-updated, results = 'asis', eval = F}
+```{r table-scores-updated, results = 'asis', eval = T}
 c_sample <- c(results_table_example$`C 80`[4], 
@@
           caption = "Applied Correction for Each Proposed Sample Size", 
-          font_size = "footnotesize")
+          font_size = "scriptsize", 
+          landscape = TRUE)
 ```

### 2023-09-26 manuscript_draft.tex
fixing manuscript table issue

@@
 
+\begin{lltable}
+
+\scriptsize{
+
+\begin{longtable}{llllllllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:table-scores-updated}Applied Correction for Each Proposed Sample Size}\\
+\toprule
+Formula Estimate & \multicolumn{1}{c}{Intercept} & \multicolumn{1}{c}{Projected SS} & \multicolumn{1}{c}{Pilot SS} & \multicolumn{1}{c}{Log2 Projected SS} & \multicolumn{1}{c}{Log2 Pilot SS} & \multicolumn{1}{c}{Log2 Power} & \multicolumn{1}{c}{Prop Var} & \multicolumn{1}{c}{Log2 Prop Var} & \multicolumn{1}{c}{Power} & \multicolumn{1}{c}{Corrected Sample} & \multicolumn{1}{c}{Data Loss}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:table-scores-updated} continued}}\\
+\toprule
+Formula Estimate & \multicolumn{1}{c}{Intercept} & \multicolumn{1}{c}{Projected SS} & \multicolumn{1}{c}{Pilot SS} & \multicolumn{1}{c}{Log2 Projected SS} & \multicolumn{1}{c}{Log2 Pilot SS} & \multicolumn{1}{c}{Log2 Power} & \multicolumn{1}{c}{Prop Var} & \multicolumn{1}{c}{Log2 Prop Var} & \multicolumn{1}{c}{Power} & \multicolumn{1}{c}{Corrected Sample} & \multicolumn{1}{c}{Data Loss}\\
+\midrule
+\endhead
+Formula & 206.59 & 0.37 & -0.77 & 27.54 & 2.58 & -66.15 & 16.40 & -1.37 & 1.09 & NA & NA\\
+Concrete 80 & 1.00 & 45.00 & 28.00 & 5.49 & 4.81 & 6.32 & 0.13 & -2.94 & 80.00 & 40.25 & 44.84\\
+Concrete 85 & 1.00 & 50.00 & 28.00 & 5.64 & 4.81 & 6.41 & 0.13 & -2.94 & 85.00 & 45.93 & 51.17\\
+Concrete 90 & 1.00 & 50.00 & 28.00 & 5.64 & 4.81 & 6.49 & 0.13 & -2.94 & 90.00 & 45.92 & 51.15\\
+Concrete 95 & 1.00 & 65.00 & 28.00 & 6.02 & 4.81 & 6.57 & 0.13 & -2.94 & 95.00 & 62.14 & 69.22\\
+LDT 80 & 1.00 & 75.00 & 33.00 & 6.23 & 5.04 & 6.32 & 0.08 & -3.61 & 80.00 & 68.47 & 87.62\\
+LDT 85 & 1.00 & 75.00 & 33.00 & 6.23 & 5.04 & 6.41 & 0.08 & -3.61 & 85.00 & 68.13 & 87.18\\
+LDT 90 & 1.00 & 90.00 & 33.00 & 6.49 & 5.04 & 6.49 & 0.08 & -3.61 & 90.00 & 80.88 & 103.50\\
+LDT 95 & 1.00 & 105.00 & 33.00 & 6.71 & 5.04 & 6.57 & 0.08 & -3.61 & 95.00 & 92.80 & 118.76\\
+\bottomrule
+\end{longtable}
+
+}
+
+\end{lltable}
+
 The updated proposal steps are in Table \ref{tab:table-summary}.

### 2023-12-27 manuscript_draft_edits.Rmd
adding this version for edits

--- 2023-09-26 manuscript_draft.Rmd
@@
 
+# New facet label names for power variable
+power.labs <- c("80% of items", "85% of items",
+                "90% of items", "95% of items")
+names(power.labs) <- c("80", "85", "90", "95")
+
+plot_colors <- c("#4FEAAC", "#FF826F", "#4F84FD")
+
 ggplot(summary_long %>% 
@@
   geom_line() + 
+  geom_line(aes(original_n, original_n), color = "black") + 
   theme_linedraw(base_size = 15) + 
-  xlab("Pilot Sample Size") + 
-  ylab("Projected Sample Size") + 
-  scale_color_discrete(name = "Scale Size") +
-  facet_wrap(~power)
+  xlab("\nPilot Sample Size") + 
+  ylab("Projected Sample Size\n") + 
+  scale_color_manual(name = "Scale Size",
+                     values = c(plot_colors)) +
+  facet_wrap(~power,
+             labeller = labeller(power = power.labs))+
+  theme_bw()
 
@@
   geom_line() + 
+  geom_line(aes(original_n, original_n), color = "black") + 
   theme_linedraw(base_size = 15) + 
-  xlab("Pilot Sample Size") + 
-  ylab("Projected Sample Size") + 
-  scale_color_discrete(name = "Scale Skew") +
-  facet_wrap(~power)
+  xlab("\nPilot Sample Size") + 
+  ylab("Projected Sample Size\n") + 
+  scale_color_manual(name = "Scale Skew",
+                     values = c(plot_colors)) +
+  facet_wrap(~power,
+             labeller = labeller(power = power.labs))+
+  theme_bw()
+
 ```
@@
   geom_line() + 
+  geom_line(aes(original_n, original_n), color = "black") +
   theme_linedraw() + 
-  xlab("Pilot Sample Size") + 
-  ylab("Projected Sample Size") + 
-  scale_color_discrete(name = "Item Heterogeneity") +
-  facet_wrap(~power)
+  xlab("\nPilot Sample Size") + 
+  ylab("Projected Sample Size\n") + 
+  scale_color_manual(name = "Item Heterogeneity",
+                     values = c(plot_colors)) +
+  facet_wrap(~power,
+             labeller = labeller(power = power.labs))+
+  theme_bw()
+
 ```
@@
        aes(original_n, new_sample_size_avg, color = variability)) + 
+  geom_line(aes(original_n, original_n), color = "black") +
+  geom_point(aes(original_n, original_n), color = "black") + 
   geom_point() + 
-  geom_point(aes(original_n, original_n), color = "black") + 
   geom_line() + 
-  theme_linedraw() + 
-  xlab("Pilot Sample Size") + 
-  ylab("Projected Sample Size") + 
-  scale_color_discrete(name = "Item Heterogeneity") +
-  facet_wrap(~power)
+  xlab("\nPilot Sample Size") + 
+  ylab("Projected Sample Size\n") + 
+  scale_color_manual(name = "Item Heterogeneity",
+                     values = c(plot_colors)) +
+  facet_wrap(~power,
+             labeller = labeller(power = power.labs))+
+  theme_bw()
 ```
@@
 If we examine the $R^2$ values for each decile of our regression equation separately, we find that the values are all $R^2$ \> .99 with very little differences between them.
-Figure \@ref(fig:decile-figure) illustrates the corrected scores for simulations at the the 40% and 50% decile recommended cutoff for item standard errors.
+Figure \@ref(fig:decile40-figure) and \@ref(fig:decile50-figure) illustrates the corrected scores for simulations at the 40% and 50% decile recommended cutoff for item standard errors.
 For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 40% decile range, especially for scales with larger potential variance.
@@
 
-```{r decile-figure, fig.cap = "Comparison of the cutoffs for 40% and 50% deciles across heterogeneity (rows), powering of items (columns), and scale size (shape)."}
+```{r decile40-figure, fig.cap = "Comparison of the cutoffs for 40% deciles across heterogeneity (columns), powering of items (rows), and scale size (color)."}
 summary_long %>% 
-  filter(name == "Decile 40" | name == "Decile 50") %>% 
+  filter(name == "Decile 40") %>% 
   group_by(variability, original_n, scale_size, name, power) %>% 
   summarize(new_sample = mean(new_sample)) %>% 
-  ggplot(., aes(original_n, new_sample, color = name)) + 
-  geom_point(aes(shape = scale_size)) + 
+  ggplot(., aes(original_n, new_sample, color = scale_size)) + 
   geom_point(aes(original_n, original_n), color = "black") + 
+  geom_line(aes(original_n, original_n), color = "black") + 
+  geom_point() + 
+  geom_line() + 
   theme_classic(base_size = 10) + 
-  facet_wrap(~ variability*power) + 
-  xlab("Pilot Sample Size") + 
-  ylab("Corrected Sample Size") + 
+  facet_grid(power~ variability,
+             labeller = labeller(power = power.labs)) +
+  ggtitle("Results for 40% deciles") +
+  xlab("\nPilot Sample Size") + 
+  ylab("Corrected Sample Size\n") + 
+  scale_color_manual(name = "Scale Size",
+                     values = c(plot_colors)) + 
+  theme_bw() +
+  theme(legend.position="bottom")
+
+```
+
+```{r decile50-figure, fig.cap = "Comparison of the cutoffs for 50% deciles across heterogeneity (columns), powering of items (rows), and scale size (color)."}
+summary_long %>% 
+  filter(name == "Decile 50") %>% 
+  group_by(variability, original_n, scale_size, name, power) %>% 
+  summarize(new_sample = mean(new_sample)) %>% 
+  ggplot(., aes(original_n, new_sample, color = scale_size)) + 
+  geom_point(aes(original_n, original_n), color = "black") + 
+  geom_line(aes(original_n, original_n), color = "black") + 
+  geom_point() + 
+  geom_line() +
+  theme_classic(base_size = 10) + 
+  facet_grid(power~ variability,
+             labeller = labeller(power = power.labs)) + 
+  ggtitle("Results for 40% deciles") +
+  xlab("\nPilot Sample Size") + 
+  ylab("Corrected Sample Size\n") + 
   scale_shape_discrete(name = "Scale Size") +
-  scale_color_discrete(name = "Cutoff Decile") + 
+  scale_color_manual(name = "Cutoff Decile",
+                     values = c(plot_colors)) + 
+  theme_bw() +
   theme(legend.position="bottom")
+
 ```
@@
 ```{r sim_pop, include = T, echo = T}
-# devtools::install_github("SemanticPriming/semanticprimeR")
+devtools::install_github("SemanticPriming/semanticprimeR")
 library(semanticprimeR)

### 2023-12-27 manuscript_draft.Rmd
updating from google docs

--- 2023-09-08 manuscript_draft.Rmd
@@
 
-Statistical power and power analyses are arguably one of the most important components in planning a research study [@cohen1990].
-However, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts [@hardwicke2020; @doi:10.1177/1745691620979806].
-The replication "crisis" and credibility revolution have shown that published studies in psychology are underpowered [@opensciencecollaboration2015; @vazire2018].
-Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020; @d.chambers2014; @nosek2014], and details of a power analyses or other limitations on resources are often used to provide justification for the pre-registered sample quota.
-Given the combined issues of publish-or-perish and that most non-significant results do not result in published manuscripts, one may expect that power analysis would be especially critical for early career researchers [@rosenthal1979; @simmons2011].
+An inevitable decision in almost any empirical research is deciding on the sample size.
+Statistical power and power analyses are arguably some of the most important components in planning a research study and its corresponding sample size [@cohen1990].
+However, if reviews of transparency and openness in research publications are any clue, researchers in the social sciences commonly fail to implement proper power analyses as part of their research workflow [@hardwicke2020; @doi:10.1177/1745691620979806].
+The replication "crisis" and credibility revolution have shown that published studies in psychology are underpowered [@opensciencecollaboration2015; @vazire2018, CITE KORBMACHER].
+Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020; @d.chambers2014; @nosek2014], and details of a power analyses or limitations on resources are often used to provide justification for the pre-registered sample quota [CITE (Pownall et al., 2023; van den Akker et al., 2023a, 2023b)].
+Given the combined issues of publish-or-perish and that most non-significant results do not result in published manuscripts, power analysis may be especially critical for early career researchers to increase the likelihood that they will identify significant effects if they exist [@rosenthal1979; @simmons2011].
+Justified sample sizes may allow for publication of non-significant, yet practically important effects, along with the smallest effect of interest movement [CITE DANIEL/ANVARI], potentially improving the credibility of published work.
 At best, an underpowered study provides limited insight [@halpern2002], and it can be difficult to know if a poorly implemented power analysis is better than no power analysis.
 
-A recent review of power analyses found - across psychology journal articles - researchers did not provide enough information to understand their power analyses and often chose effect sizes that were unjustified [@beribisky2019].
-One solution to this power analysis problem is the plethora of tools made available for researchers to make power computations accessible to non-statisticians with the caveat that a solid education in power is necessary to use these tools.
+A recent review of power analyses found - across behavioral, cognitive, and social science journal articles - researchers did not provide enough information to understand their power analyses and often chose effect sizes that were unjustified [@beribisky2019].
+One solution to this power analysis problem is the plethora of tools made available for researchers to make power computations accessible to non-statisticians; however, a solid education in power is necessary to use these tools properly.
 G\*Power is one of the most popular free power software options [@faul2007; @erdfelder1996] that provides a simple point and click graphical user interface for power calculations [however, see @brysbaert2019].
-Web-based tools have also sprung up for overall and statistical test specific sample size planning including <https://powerandsamplesize.com> and <https://designingexperiments.com> [@anderson2017].
-*R*-coding based packages, such as *pwr* [@champely2017], *faux* [@debruine2021], and *SimDesign* [@chalmers2020], can be used to examine power and sample size planning usually with simulation.
+Web-based tools have also sprung up for overall and statistical test specific sample size planning including <https://powerandsamplesize.com>, <https://jakewestfall.shinyapps.io/pangea/> , <https://pwrss.shinyapps.io/index/>, and <https://designingexperiments.com> [@anderson2017].
+*R*-coding based packages, such as *pwr* [@champely2017], *faux* [@debruine2021], *simr* [CITE GREEN], *mixedpower* [CITE KUMLE], and *SimDesign* [@chalmers2020], can be used to examine power and plan sample sizes, usually with simulation.
 Researchers must be careful using any toolkit, as errors can occur with the over-reliance on software [e.g., it should not be a substitute for critical thinking, @nuijten2016].
+Additionally, many tools assume data normality, place an overemphasis on statistical significance, and may rely on simplified assumptions that do not reflect the actual data.
 When computing sample size estimates, it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result [@batterham2005].
-For example, if any parameter estimated by the researcher was not found in the study (i.e., a smaller effect size than used for the power analysis), then the sample size estimate tied to that specific level of power may be incorrect.
-
-Changes in publication practices and research design have also created a new wrinkle in providing a sample size plan for a research study.
+For example, it is hard to accurately estimate all parameters from a study, and if any were incorrect, then the sample size estimate tied to that specific level of power may be incorrect [CITE ALBERS].
+
+Power analyses are not always the best tool for the job.
+Changes in publication practices and research design have also created new challenges in providing a sample size plan for a research study.
 While statistics courses often suggest that a specific research design leads to a specific statistical test, meta-science work has shown that given the same data and hypothesis, researchers can come up with multiple ways to analyze the data [@silberzahn2018many, @coretta].
-Therefore, a single power analysis only supports the specific analysis in which the researcher expects to test.
+Therefore, a single power analysis only corresponds to the specific analysis that the researcher expects to implement.
 Analyses may evolve during the research project or be subject to secondary analysis; thus, power and sample size estimation based on one analysis is potentially less useful than previously imagined.
 Further, research projects often have multiple testable hypotheses, but it is unclear which hypothesis or test should be used to estimate sample size with a power analysis.
-Last, research publications may not even have a specific, testable hypothesis, as some publications are intended to curate a large dataset for future reuse [i.e., stimuli database creation, @buchanan2019].
-
-In light of these analytical (or lack thereof) concerns, we propose a new method that combines accuracy in parameter estimation [AIPE, @kelley2007, @maxwell2008] and bootstrapped simulation on pilot data [@rousselet].
-This method accounts for a potential lack of hypothesis test (or simply not a good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices.
-Additionally, we consider the nature of cognitive research designs that use multiple items to measure the phenomena of interest.
-For example, semantic priming is measured with multiple paired stimuli [@meyer1971], which traditionally was analyzed by creating person or item-level averages for an ANOVA [@brysbaert2018].
-However, the use of multilevel models with random effects for the stimuli used in a study have shown that we should be careful to assume that all items of a research study have the same "effect", as there is often variability in their impact on the outcomes of the study.
+Last, research investigations may not even have a specific, testable hypothesis, as some projects are intended to curate a large dataset for future reuse [i.e., stimuli database creation, @buchanan2019].
+
+In light of these analytical (or lack thereof) concerns, we propose a new method to determine a sample size in cases where a more traditional power analysis might be less appropriate or even impossible.
+This approach combines accuracy in parameter estimation [AIPE, @kelley2007, @maxwell2008] and bootstrapped simulation on pilot data [@rousselet].
+This method accounts for a potential lack of hypothesis test (or simply no good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices.
+Specifically this manuscript focuses on research designs that use multiple items to measure the phenomena of interest.
+For example, semantic priming is measured with multiple paired stimuli [@meyer1971], which traditionally has been analyzed by creating person or item-level averages to test using an ANOVA [@brysbaert2018].
+However, research implementing multilevel models with random effects for the stimuli used in a study has demonstrated potential variability in their impact on outcomes; thus we should be careful not to assume that all items in a research study have the same "effect".
 
@@
 AIPE shifts the focus away from finding a significant p-value to finding a parameter that is "accurately measured".
-For example, researchers may wish to detect a specific correlation in a study, *r* = .35.
-They could then use AIPE to estimate the sample size needed to find a "sufficiently narrow" confidence interval around that correlation.
+For example, researchers may wish to detect a specific mean in a study, *M* = .35.
+They could then use AIPE to estimate the sample size needed to find a "sufficiently narrow" confidence interval around that mean.
 Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and confidence intervals.
@@
 
-Bootstrapping involves using data obtained to simulate similar datasets by drawing from the original data with replacement [@efron2000; @rousselet].
-Bootstrapping is a form of data simulation that allows one to calculate parameter estimates, confidence intervals, and more to simulate the potential population distribution, shape, and bias.
-Simulation is often paired with making up data for testing analyses and hypotheses based on proposed effect size or suggested population means.
-Generally, we would suggest starting with pilot data of a smaller sample size to understand the variability in potential items used to represent your phenomenon, especially if they are to be used in a larger study.
-However, given some background knowledge about the potential items, one could simulate example pilot data to use in.
-similar manner in our suggested procedure.
-Pilot or simulated data would be used to estimate the variability within items and select a "sufficiently narrow" window for item's confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
-Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein many items meet the required confidence interval.
+One form of data simulation is Bootstrapping, which involves using data obtained to simulate similar datasets by drawing from the original data with replacement [@efron2000; @rousselet].
+Bootstrapping allows one to calculate parameter estimates, confidence intervals, and to simulate the potential population distribution, shape, and bias.
+Simulation is often paired with re-creating a data set with a similar structure for testing analyses and hypotheses based on proposed effect sizes or suggested population means.
+Generally, we would suggest starting with pilot data of a smaller sample size (e.g., 20 to 50) to understand the variability in potential items used to represent your phenomenon, especially if they are to be used in a larger study.
+However, given some background knowledge about the potential items, one could simulate example pilot data to use in a similar manner in our suggested procedure.
+Pilot or simulated data would be used to estimate the variability within items and select a "sufficiently narrow" window for overall item confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
+The advantage to this method over simple power estimation from pilot effect sizes is the multiple simulations to average out potential variability, as well as a shift away from traditional NHST to parameter estimation.
+Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein as many items as required meet the pre-specified confidence interval.
 
@@
 
-Researchers could then use sequential testing to estimate their parameter of interest after each participant's data was collected to determine whether they have achieved their expected width of the confidence interval around that parameter.
-One would set a minimum sample size (i.e., based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the CI is narrow, as defined above).
-Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (i.e., we expect 90% of items to meet our stopping rule with 100 participants based on simulation).
+Researchers could then use sequential testing to estimate their parameter of interest after each participant's data or at regular intervals during data collection to determine whether they have achieved their expected width of the confidence interval around that parameter.
+One would set a minimum sample size (e.g.,, based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the CI is sufficiently narrow, as defined above).
+Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (e.g., , they expect 90% of items to meet their stopping rule with 100 participants based on simulation).
 By defining each of these components, researchers could ensure a feasible minimum sample size, a way to stop data collection when goals have been met, and a maximum sample size rule to ensure an actual end to data collection.
-Therefore, we proposed that we should be able to leverage the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for repeated measures studies with expected variability in parameter estimates for items.
-
-## Proposed Method to Calculate Sample Size
-
-Using these ideas, we suggest the following procedure to determine a sample size for each item:
+The maximum stopping rule could also be defined by resources (e.g., two semesters data collection), but nevertheless should be included.
+Therefore, we propose a method that leverages the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates to measure an overall phenomena.
+
+## Proposed Method for Sample Size Planning
+
+Building on these ideas, we suggest the following procedure to determine a sample size for each item:
+
+### Calculate the Stopping Rule
 
 1) Use pilot data that closely resembles data you intend to collect.
-This dataset should contain items that are identical or similar to those that will be collected in the study.
+This dataset should contain items that are identical or similar to those that will be implemented in the study.
 In this procedure, it is important to ensure that the data is representative of a larger population of sampled items that you intend to assess.
-Generally, pilot data sample sizes will be small, as the goal would be to determine how many items would be necessary to reach a "stable" standard error.
-
-2)  Calculate the standard error (SE) in the subset of pilot data associated with each item to create a cutoff score that defines when items are considered "accurately measured".
-Next, the simulations below will explore what criterion should be used to determine the cutoff score from the pilot data.
-
-3)  Sample, with replacement, from your pilot data using sample sizes starting at 20 participants and increase in small units up to a value that you consider the maximum sample size.
+Generally, pilot data sample sizes will be smaller than the overall intended project (e.g., 20 to 50), as the goal would be to determine how many participants would be necessary to reach a "stable" standard error.
+
+2) For each item in the pilot data, calculate the standard error (SE).
+Select a cutoff SE that defines when items are considered "accurately measured".
+The simulations described in the Data Simulation section will explore what criterion should be used to determine the cutoff SE from the pilot data.
+
+### Bootstrap Samples
+
+3) Sample, with replacement, from your pilot data using sample sizes starting at a value that you consider the minimal sample size per item  and increase in small units up to a value that you consider the maximum sample size.
 We will demonstrate example maximum sample sizes based on the data simulation below; however, a practical maximum sample size may be determined by time (e.g., one semester data collection) or resources (e.g., 200 participants worth of funding).
-Although 20 participants would likely yield imprecise estimates, we suggest this starting minimum for simulation purposes.
-
-4)  For each simulation, calculate the standard error for each item, and use these values to determine the percentage of items that meet the cutoff score determined in Step 2.
+As for the minimal sample size, we suggest using 20 as a reasonable value  for simulation purposes.
+For each sample size simulation, calculate the standard error for each item.
+Use multiple simulations (e.g., n = 500 to 1000) to avoid issues with random sampling variability.
+
+### Determine Minimum, Maximum Sample Size
+
+4) Use the simulated SEs to determine the percentage of items that meet the cutoff score determined in Step 2.
+Each sample size from Step 3 will have multiple bootstrapped simulations, and therefore, create an average percentage score for each sample size for Step 5.
 
@@
 
-6)  Report these values, and designate a minimum sample size, the cutoff criterion, and the maximum sample size.
+### Report Results
+
+6) Report these values, and designate a minimum sample size, the cutoff/stopping rule criterion, and the maximum sample size.
 Each researcher should also report if they plan to use an adaptive design, which would stop data collection after meeting the cutoff criterion for each item.
@@
 This example will reveal a few areas of needed exploration for the steps.
-Next, we portray simulations for the proposed procedure and find solutions to streamline and improve the sample size estimation procedure.
-Finally, we include additional resources for researchers to use to implement the estimation procedure.
+Next, we provide simulations for the proposed procedure to find solutions to overcome these potential limitations and formally define Step 2.
+Finally, we include additional resources for researchers to use to implement the proposed sample size estimation procedure.
 
@@
 
-In this section, we provide two examples of the suggested procedure.
-The first example includes concreteness ratings from @brysbaert2014.
+In this section, we provide an example of the suggested procedure.
+The first dataset includes concreteness ratings from @brysbaert2014.
 Instructions given to participants denoted the difference between concrete (i.e., "refers to something that exists in reality") and abstract (i.e., "something you cannot experience directly through your senses or actions") terms.
@@
 
-The second dataset includes a large scale dataset with response latencies, the English Lexicon Project [@balota2007].
-The English Lexicon Project consists of lexical decision response latencies for English words.
+The second dataset includes a large scale dataset with response latencies, the English Lexicon Project [ELP; @balota2007].
+The ELP consists of lexical decision response latencies for written English words and pseudowords.
 In a lexical decision task, participants simply select "word" for real words (e.g., *dog*) and "nonword" for pseudowords (e.g., *wug*).
 The trial level data is available here: <https://elexicon.wustl.edu/>.
-Critically, in each of these examples, the individual trial level data for each item is available to simulate and calculate standard errors on.
+Critically, in each of these datasets, the individual trial level data for each item is available to simulate and calculate standard errors on.
 Data that has been summarized could potentially be used, as long as the original standard deviations for each item were present.
@@
 
-For this example, our researcher wants to determine the differences in response latencies for abstract and concrete words.
-They will select *n* = 40 words from the rating data that are split evenly into abstract and concrete ends of the rating scale.
-In the experiment, each participant will rate the words for their concreteness, and then complete a lexical decision task with these words as the object of interest.
-Using both datasets, we can determine the sample size necessary to ensure adequately measured ratings and response latencies.
+For this example, imagine a researcher who wants to determine the differences in response latencies for abstract and concrete words.
+They will select *n* = 40 words from the rating data from @brysbaert2014 that are split evenly into abstract and concrete ends of the rating scale.
+In the experiment, each participant will be asked to rate the words for their concreteness, and then complete a lexical decision task with these words as the phenomenon of interest.
+Using both datasets and the procedure outlined above, we can determine the sample size necessary to ensure adequately measured concreteness ratings and response latencies.
 
@@
 In the original study, not every participant rated every word, which created uneven sample sizes for each word.
-Further, participants were allowed to indicate they did not know a word, and this data was set to missing data.
+Further, participants were allowed to indicate they did not know a word, and those responses were set to missing data.
 In our sample of 40 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size_c` as our pilot sample size for this example.
@@
 
-The ELP response latency data includes `r length(unique(elpDF$Stimulus))` word-forms, `r length(unique(elpDF$Stimulus[elpDF$Type == 0]))` that are listed as non-words, and `r length(unique(elpDF$Stimulus[elpDF$Type == 1]))` real words.
-We selected the same words as the concreteness subset selected above.
+We first filtered the ELP data to the same real words as the concreteness subset selected above, and this data includes `r length(unique(elpDF$Stimulus))` real words.
 The average pilot sample size for this random sample was `r format(round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2), nsmall = 2)` (*SD* = `r round(sd(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2)`), and *n* = `r pilot_size_e` will be our pilot size for this example.
@@
 *Step 3-5*.
-The pilot data was then simulated, with replacement, with samples from 20 to 300 increasing in units of 5, separately for concreteness and lexical decision times (Step 3).
-The standard error of each item was calculated for the bootstrapped samples (Step 4), and the percentage of items below each potential cutoff was gathered (Step 5).
-The smallest sample size with at least 80%, 85%, 90%, and 95% of items below the cutoff are reported in Table \@ref(tab:table-example) (Step 5).
+The pilot data was then simulated 500 times, with replacement, with samples of 20 to 300 participants per item increasing in units of 5, for concreteness ratings and lexical decision latencies separately (Step 3).
+The standard error of each item was calculated for the bootstrapped samples separately for concreteness ratings and lexical decision times (Step 4), and the percentage of items below each potential cutoff was gathered for each (Step 5).
+The smallest sample size with at least 80%, 85%, 90%, and 95% of items below the cutoff are reported in Table \@ref(tab:table-example) for each task (Step 5).
 
@@
 In the last step, the researcher would indicate their smallest sample size, the cutoff standard error criterion if they wanted to adaptively test (e.g., examine the SE after each participant and stop data collection if all items reached criteria), and their maximum sample size.
-As mentioned earlier, the decile for a balanced standard error cutoff is unclear and without guidance, a potential set of researcher degrees of freedom [CITE].
-Even though both scales appear to converge on similar sample size suggestions for each decile and percent level, the impact of scale size (i.e., 1-5 versus 0-`r max(elp_small$RT, na.rm = T)`) and heterogeneity of item standard errors (concrete $SD_{SD}$ = `r apa_num(sd_items_c)` and lexical $SD_{SD}$ = `r apa_num(sd_items_e)`) is not obvious.
+As mentioned earlier, the decile for a balanced standard error cutoff is unclear and without guidance, a potential set of researcher degrees of freedom could play a role in the chosen cutoff [CITE SIMONSOHN].
+Even though both measurements (ratings and response latencies) appear to converge on similar sample size suggestions for each decile and percent level, the impact of scale size (i.e., concreteness ratings 1-5 versus response latencies in ms 0-`r max(elp_small$RT, na.rm = T)`) and heterogeneity of item standard errors (concrete $SD_{SD}$ = `r apa_num(sd_items_c)` and lexical $SD_{SD}$ = `r apa_num(sd_items_e)`) is not obvious.
 Last, by selecting the ends of the distribution for our concreteness words, skew of the distribution may additionally impact our estimates.
@@
 
-In order to evaluate our approach, we used data simulation to create representative pilot datasets of several popular cognitive scales (1-7 measurements, 0-100 percentage measurements, and 0-3000 response latency scale data).
+In order to evaluate our approach, we used data simulation to create representative pilot datasets of several popular cognitive scales (1-7 measurements, 0-100 percentage measurements, and 0-3000 response latency type scale data).
 For each of these scales, we also manipulated item heterogeneity by simulating small differences in item variances to large differences in item variances based on original scale size.
-On each of the simulated datasets, we applied the above proposed method to determine how the procedure would perform and evaluate what criteria should be used for cutoff selection (Step 2).
+On each of the simulated datasets, we applied the above proposed method to determine how the procedure would perform and evaluated what criteria should be used for cutoff selection (Step 2).
 This procedure was performed on distributions in the middle of the scale (i.e., normal) and at the ceiling of the scale (i.e., skewed).
@@
 How does distribution skew impact sample size estimations?
-Skew can potentially decrease heterogeneity (i.e., all items are at ceiling, and therefore, variance between item standard errors is low) or could increase heterogeneity (i.e., some items are skewed, while others are not).
+Skew can potentially decrease item variance heterogeneity (i.e., all items are at ceiling, and therefore, variance between item standard errors is low) or could increase heterogeneity (i.e., some items are skewed, while others are not).
 Therefore, we expect skew to impact the estimates in the same way as point C.
@@
 2) Do the results match what one might expect for traditional power curves?
-Power curves are asymptotic, that is, they "level off" as sample size increases.
+Power curves are asymptotic; that is, they "level off" as sample size increases.
 Therefore, we expect that our procedure should also demonstrate a leveling off effect as pilot data sample size increases.
@@
 Items were rounded to the nearest whole number to mimic scales generally collected by researchers.
-Items were also rounded to their appropriate scale end points (i.e., all items below 0 on a 1-7 scale were replaced with 1, etc.).
+Items were also rounded to their appropriate scale endpoints (i.e., all items below 0 on a 1-7 scale were replaced with 1, etc.).
 
@@
 First, the scale of the data was manipulated by creating three sets of scales.
-The first scale was mimicked after small rating scales (i.e., 1-7 type style) using a $\mu$ = 4 with a $\sigma$ = .25 around the mean to create item mean variability.
+The first scale was mimicked after small rating scales (i.e., 1-7 Likert-type style, treated as interval data) using a $\mu$ = 4 with a $\sigma$ = .25 around the mean to create item mean variability.
 The second scale included a larger potential distribution of scores with a $\mu$ = 50 ($\sigma$ = 10) imitating a 0-100 scale.
@@
 Finally, for the large scale of the data, $\sigma$ = 400 with a variance of 50, 100, and 200 for heterogeneity.
+These values were based on the proportion of the overall scale and potential variance.
 
@@
 Each of these samples would correspond to Step 1 of the proposed method where a researcher would use pilot data to start their estimation.
-Therefore, the simulations included 3 scales X 3 heterogeneity X 2 normal/skewed X 9 pilot sample sizes representing a potential Step 1 of our procedure.
+Therefore, the simulations included 3 scales X 3 heterogeneity values X 2 normal/skewed distributions X 9 pilot sample sizes representing a potential Step 1 of our procedure.
 
@@
 Standard errors were calculated at each decile of the items up to 90% (i.e., 0% smallest SE, 10% ..., 90% largest SE).
-The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores.
+The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores(Step 2).
 
@@
 
-Figure \@ref(fig:scale-size-figure) demonstrates the influence of scale size on the results separate by potential cutoff level.
+Figure \@ref(fig:scale-size-figure) demonstrates the influence of scale size on the results separated by potential cutoff level.
 The black dots denote the original sample size for reference.
@@
 To explore if we could recover the new projected sample size from data a researcher would have, we used regression models to create a formula for researcher correction.
-The researcher employing our procedure would have the possible following variables from their simulations on their (one) pilot dataset: 1) proposed sample size, 2) pilot sample size, 3) estimate of heterogeneity for the items, 4) and the estimate percent of items below the threshold.
-Given the non-linear nature of the correction, we added each variable and it's non-linear `log2` transform to the regression equation, as this function was used to create the correction.
+The researcher employing our procedure would have the possible following variables from their simulations on their (one) pilot dataset: 1) proposed sample size, 2) pilot sample size, 3) estimate of heterogeneity for the items, 4) and the estimated percent of items below the threshold.
+Given the non-linear nature of the correction, we added each variable and its non-linear `log2` transform to the regression equation, as this function was used to create the correction.
 The intercept only model was used as a starting point (i.e., `corrected sample ~ 1`), and then all eight variables (each variable and their `log2` transform) were entered into a forward stepwise regression to capture the corrected scores with the most predictive values.
@@
 If we examine the $R^2$ values for each decile of our regression equation separately, we find that the values are all $R^2$ \> .99 with very little differences between them.
-Figure \@ref(fig:decile-figure) illustrates the corrected scores for simulations at the the 40% and 50% decile recommended cutoff for item standard errors.
-For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 40% decile range, especially for scales with larger potential variance.
-Therefore, we would suggest the 40% decile to overpower each item to determine the minimum and maximum sample size.
+Figure \@ref(fig:decile-figure) illustrates the corrected scores for simulations at the 40% and 50% decile recommended cutoff for item standard errors.
+For small heterogeneity, differences in decile are minimal, while large heterogeneity shows more correction at the 40% decile range, especially for scales with larger potential variance.
+Therefore, we would suggest using the 40% decile to overpower each item to determine the minimum and maximum sample size.
 
@@
 While the estimated coefficients could change given variations on our simulation parameters, the general size and pattern of coefficients was consistent, and therefore, we believe this correction equation should work for a variety of use cases.
-We will now demonstrate the final procedure on the examples provided earlier.
+We will now demonstrate the final procedure on the example provided earlier.
 
@@
 For one additional consideration, we calculated the potential amount of data retention given that participants could indicate they did not know a word ($M_{answered}$ = `r round(mean(data_loss_c$percent_correct), digits = 2)`, *SD* = `r round(sd(data_loss_c$percent_correct), digits = 2)`) in the concreteness task or answer a trial incorrectly in the lexical decision task ($M_{correct}$ = `r round(mean(data_loss_e$percent_correct), digits = 2)`, *SD* = `r round(sd(data_loss_e$percent_correct), digits = 2)`).
-In order to account for this facet, the potential sample sizes were multiplied by $\frac{1}{p_{retained}}$ where the denominator is proportion retained for each task.
+In order to account for this data loss, the potential sample sizes were multiplied by $\frac{1}{p_{retained}}$ where the denominator is proportion retained for each task.
 
@@
 You can install the package from GitHub using: `devtools::install_github("SemanticPriming/semanticprimeR")`.
-We detail the functions below by proposed step in the process.
+We detail the functions below with proposed steps in the process.
 
@@
 *Step 3*.
-The `bootstrap_samples()` function creates bootstrapped samples from the pilot or simulated population data to create samples to estimate the number of participants needed for item standard error to be below the cutoff calculated in Step 2.
+The `bootstrap_samples()` function creates bootstrapped samples from the pilot or simulated population data to estimate the number of participants needed for item standard error to be below the cutoff calculated in Step 2.
 This function returns a list of samples with sizes that start at the `start` size, increase by `increase`, and end with the `stop` sample size.
 The population or pilot data will be included in `population`, and the item column indicator should be included in `grouping_items`.
+The `nsim` argument determines the number of bootstrapped simulations to run.
 
@@
 The proportion of bootstrapped items across sample sizes below the cutoff score can then be calculated using `calculate_proportion()`.
-This function returns a dataframe of each sample size with the proportion of items below that cutoff to use in the next function.
-The `samples` and `cutoff` were calculated with our previous functions, with the column for item labels and dependent variable to ensure the right calculations.
+This function returns a dataframe including each sample size with the proportion of items below that cutoff to use in the next function.
+The `samples` and `cutoff` arguments were previously calculated with our functions.
+The column for item labels and dependent variables are included as `grouping_items` and `score` arguments to ensure the right calculations.
 
@@
 The `proportion_summary` from above is used in this function, along with required information about the sample size, proportion of variance from our cutoff calculation, and what power levels should be calculated.
-Note that the exact percent of items below a cutoff score will be returned, if the values in `power_levels` are not exactly present.
+Note that the exact percent of items below a cutoff score will be returned if the values in `power_levels` are not exactly calculated.
 The final summary presents the smallest sample size, corrected, for each of the potential power levels.
@@
 
-While the examples in this manuscript are traditionally cognitive linguistics focused, any research using repeated items can benefit from newer sampling techniques.
+While the example in this manuscript was traditionally cognitive linguistics focused, any research using repeated items as a unit of measure could benefit from the proposed newer sampling techniques.
 Therefore, we provide 12 example vignettes and varied code examples on our OSF page/GitHub site for this manuscript across a range of data types provided by the authors of this manuscript.
-Examples include psycholinguistics [@dedeyne2008; @heyman2014; @montefinese2022], eye tracking data [@ulloa2014], social psychology [@peterson2022; @grahe2022], COVID related data [@montefinese2021], and cognitive psychology [@röer2013; @10.7554/eLife.71601; @barzykowski2019].
+Examples include psycholinguistics [@dedeyne2008; @heyman2014; @montefinese2022], social psychology data [@ulloa2014], social psychology [@peterson2022; @grahe2022], COVID related data [@montefinese2021], and cognitive psychology [@röer2013; @10.7554/eLife.71601; @barzykowski2019].
 
@@
 
-We proposed and demonstrated examples for a method using AIPE, bootstrapping, and simulation to estimate a minimum and maximum sample size along with a rule for stopping data collection based on narrow confidence intervals on the parameter of interest.
+We proposed a method combining AIPE, bootstrapping, and simulation to estimate a minimum and maximum sample size and to define a rule for stopping data collection based on narrow confidence intervals on a parameter of interest.
+In addition, we also demonstrated its practical applications using real-world data.
 We contend that this procedure is specifically useful for studies with multiple items that intend on using item level focused analyses; furthermore, the utility of measuring each item well can extend to many analysis choices.
@@
 One limitation of these methods would be our decision to use datasets with very large numbers of items to simulate what might happen within one study.
-For example, the English Lexicon Project includes thousands of items, and by the time we would simulate for all of those, it would likely suggest needing thousands of participants for most items to reach the criterion.
+For example, the English Lexicon Project includes thousands of items, and if we were to simulate for all of those, our results would likely suggest needing thousands of participants for most items to reach the criterion.
 Additionally, as the number of items increases, you may also see very small estimates for sample size due to the correction factor (as with large numbers of items, you could find many items with standard errors below the 40% decile).
@@
 This result occurs because the smaller number of items means more samples for nearly all to reach the cutoff criteria.
-These results are not too different from what we might expect for a power analysis using a multilevel model - larger numbers of items tend to decrease necessary sample size, while smaller numbers of items tend to increase sample size.
+These results are similar to what we might expect for a power analysis using a multilevel model - larger numbers of items tend to decrease necessary sample size, while smaller numbers of items tend to increase sample size.
 
 Second, these methods do not ensure the normal interpretation of power, where you know that you would find a specific effect for a specific test, $\alpha$, and so on.
-As discussed in the Introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis, many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
-These methods could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
-
-Researchers should consider this hybrid approach for AIPE, bootstrapping, and simulation as a tool for hypothesis testing and parameter estimation.
-The proposed procedure can be beneficial for many different research studies, specifically replication studies, that usually depend on subject sample size but rarely item sample size, in spite of the fact that item sample sizes contribute to language processing [@brysbaert2018; @baayen2008].
-These observed effects can be then replicated and, as a result of several replications, can be applied to meta-analyses.
-As a result, analysts would be able to use the accuracy and high statistical power to calculate the parameters to assess whether the effect is genuine.
+As discussed in the Introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis; many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
+These proposed methods and traditional power analysis could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
+
+Researchers should consider this hybrid approach for AIPE, bootstrapping, and simulation as a powerful tool for hypothesis testing and parameter estimation.
+This procedure holds benefits for various research studies, specifically replication studies, that usually prioritize subject sample size but rarely item sample size, in spite of the fact that item sample sizes can contribute to power in multilevel models [@brysbaert2018].
+Replicated effects, accumulated through multiple studies, contribute to robust meta-analyses, enhancing our understanding of the genuine nature of observed effects.
 This article helps to achieve this goal by encouraging researchers to conduct studies where the power analysis is not based on the size of the effect but on adequate sampling of the stimuli.

### 2023-12-27 manuscript_draft.Rmd
adding @troettge changes back

@@
   geom_line() + 
+  geom_line(aes(original_n, original_n), color = "black") + 
   theme_linedraw(base_size = 15) + 
-  xlab("Pilot Sample Size") + 
-  ylab("Projected Sample Size") + 
-  scale_color_discrete(name = "Scale Size") +
-  facet_wrap(~power)
-
+  xlab("\nPilot Sample Size") + 
+  ylab("Projected Sample Size\n") + 
+  scale_color_manual(name = "Scale Size",
+                     values = c(plot_colors)) +
+  facet_wrap(~power,
+             labeller = labeller(power = power.labs))+
+  theme_bw()
 ```
@@
   geom_line() + 
+  geom_line(aes(original_n, original_n), color = "black") + 
   theme_linedraw(base_size = 15) + 
-  xlab("Pilot Sample Size") + 
-  ylab("Projected Sample Size") + 
-  scale_color_discrete(name = "Scale Skew") +
-  facet_wrap(~power)
+  xlab("\nPilot Sample Size") + 
+  ylab("Projected Sample Size\n") + 
+  scale_color_manual(name = "Scale Skew",
+                     values = c(plot_colors)) +
+  facet_wrap(~power,
+             labeller = labeller(power = power.labs))+
+  theme_bw()
 ```
@@
 If we examine the $R^2$ values for each decile of our regression equation separately, we find that the values are all $R^2$ \> .99 with very little differences between them.
-Figure \@ref(fig:decile-figure) illustrates the corrected scores for simulations at the 40% and 50% decile recommended cutoff for item standard errors.
-For small heterogeneity, differences in decile are minimal, while large heterogeneity shows more correction at the 40% decile range, especially for scales with larger potential variance.
-Therefore, we would suggest using the 40% decile to overpower each item to determine the minimum and maximum sample size.
-
-```{r decile-figure, fig.cap = "Comparison of the cutoffs for 40% and 50% deciles across heterogeneity (rows), powering of items (columns), and scale size (shape)."}
+Figures \@ref(fig:decile40-figure) and \@ref(fig:decile50-figure) illustrate the corrected scores for simulations at the 40% and 50% decile recommended cutoff for item standard errors.
+For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 40% decile range, especially for scales with larger potential variance.
+Therefore, we would suggest the 40% decile to overpower each item to determine the minimum and maximum sample size.
+
+```{r decile40-figure, fig.cap = "Comparison of the cutoffs for 40% deciles across heterogeneity (columns), powering of items (rows), and scale size (color)."}
 summary_long %>% 
-  filter(name == "Decile 40" | name == "Decile 50") %>% 
+  filter(name == "Decile 40") %>% 
   group_by(variability, original_n, scale_size, name, power) %>% 
   summarize(new_sample = mean(new_sample)) %>% 
-  ggplot(., aes(original_n, new_sample, color = name)) + 
-  geom_point(aes(shape = scale_size)) + 
+  ggplot(., aes(original_n, new_sample, color = scale_size)) + 
   geom_point(aes(original_n, original_n), color = "black") + 
+  geom_line(aes(original_n, original_n), color = "black") + 
+  geom_point() + 
+  geom_line() + 
   theme_classic(base_size = 10) + 
-  facet_wrap(~ variability*power) + 
-  xlab("Pilot Sample Size") + 
-  ylab("Corrected Sample Size") + 
+  facet_grid(power~ variability,
+             labeller = labeller(power = power.labs)) +
+  ggtitle("Results for 40% deciles") +
+  xlab("\nPilot Sample Size") + 
+  ylab("Corrected Sample Size\n") + 
+  scale_color_manual(name = "Scale Size",
+                     values = c(plot_colors)) + 
+  theme_bw() +
+  theme(legend.position="bottom")
+
+```
+
+```{r decile50-figure, fig.cap = "Comparison of the cutoffs for 50% deciles across heterogeneity (columns), powering of items (rows), and scale size (color)."}
+summary_long %>% 
+  filter(name == "Decile 50") %>% 
+  group_by(variability, original_n, scale_size, name, power) %>% 
+  summarize(new_sample = mean(new_sample)) %>% 
+  ggplot(., aes(original_n, new_sample, color = scale_size)) + 
+  geom_point(aes(original_n, original_n), color = "black") + 
+  geom_line(aes(original_n, original_n), color = "black") + 
+  geom_point() + 
+  geom_line() +
+  theme_classic(base_size = 10) + 
+  facet_grid(power~ variability,
+             labeller = labeller(power = power.labs)) + 
+  ggtitle("Results for 40% deciles") +
+  xlab("\nPilot Sample Size") + 
+  ylab("Corrected Sample Size\n") + 
   scale_shape_discrete(name = "Scale Size") +
-  scale_color_discrete(name = "Cutoff Decile") + 
+  scale_color_manual(name = "Cutoff Decile",
+                     values = c(plot_colors)) + 
+  theme_bw() +
   theme(legend.position="bottom")
+
 ```

### 2023-12-28 manuscript_draft.Rmd
my updates from feedback

@@
 ```{r trackdown, eval = F}
+# devtools::install_github("ClaudioZandonella/trackdown")
 library(trackdown)
@@
 This example will reveal a few areas of needed exploration for the steps.
-Next, we provide simulations for the proposed procedure to find solutions to overcome these potential limitations and formally define Step 2.
-Finally, we include additional resources for researchers to use to implement the proposed sample size estimation procedure.
+Next, we portray simulations for the proposed procedure and find solutions to streamline and improve the sample size estimation procedure.
+Table \@ref(tab:table-summary) shows the results of the simulations and solutions on the right hand side.
+Finally, we include additional resources for researchers to use to implement the estimation procedure.
 
@@
 elpDF <- elpDF %>% 
-  filter(Stimulus %in% conDF$Word)
+  filter(Stimulus %in% conDF$Word) %>% 
+  filter(Type == 1)
 
@@
 # sequence of sample sizes to try
+nsim <- 5
 samplesize_values <- seq(20, 300, 5)
@@
 sim_table <- matrix(NA, 
-          nrow = length(samplesize_values), 
+          nrow = length(samplesize_values)*nsim, 
           ncol = length(unique(con_use$Word)))
@@
 sim_table$sample_size <- NA
+
+iterate <- 1
+for (p in 1:nsim){
   # loop over sample sizes
@@
    colnames(sim_table)[1:length(unique(con_use$Word))] <- temp$Word
- sim_table[i, 1:length(unique(con_use$Word))] <- temp$se
- sim_table[i, "sample_size"] <- samplesize_values[i]
+   sim_table[iterate, 1:length(unique(con_use$Word))] <- temp$se
+   sim_table[iterate, "sample_size"] <- samplesize_values[i]
+   sim_table[iterate, "nsim"] <- p
+
+   iterate <- iterate + 1
+  }
 }
@@
  sim_table %>% 
- pivot_longer(cols = -c(sample_size)) %>% 
+ pivot_longer(cols = -c(sample_size, nsim)) %>% 
  rename(item = name, se = value) %>% 
- group_by(sample_size) %>% 
+ group_by(sample_size, nsim) %>% 
  summarize(Percent_Below10 = sum(se <= cutoff_c["10%"])/length(unique(con_use$Word)),
@@
           Percent_Below90 = sum(se <= cutoff_c["90%"])/length(unique(con_use$Word))) %>% 
+  ungroup() %>% 
+  group_by(sample_size) %>% 
+  summarize(Percent_Below10 = mean(Percent_Below10),
+            Percent_Below20 = mean(Percent_Below20),
+            Percent_Below30 = mean(Percent_Below30),
+            Percent_Below40 = mean(Percent_Below40),
+            Percent_Below50 = mean(Percent_Below50),
+            Percent_Below60 = mean(Percent_Below60),
+            Percent_Below70 = mean(Percent_Below70),
+            Percent_Below80 = mean(Percent_Below80),
+            Percent_Below90 = mean(Percent_Below90)) %>% 
   pivot_longer(-sample_size) %>% 
@@
 sim_table$sample_size <- NA
+
+iterate <- 1
+
+for (p in 1:nsim){
+
   # loop over sample sizes
@@
    colnames(sim_table)[1:length(unique(elp_use$Stimulus))] <- temp$Stimulus
- sim_table[i, 1:length(unique(elp_use$Stimulus))] <- temp$se
- sim_table[i, "sample_size"] <- samplesize_values[i]
+   sim_table[iterate, 1:length(unique(elp_use$Stimulus))] <- temp$se
+   sim_table[iterate, "sample_size"] <- samplesize_values[i]
+   sim_table[iterate, "nsim"] <- p
+
+   iterate <- iterate + 1
+   }
+
 }
@@
  rename(item = name, se = value) %>% 
- group_by(sample_size) %>% 
+ group_by(sample_size, nsim) %>% 
  summarize(Percent_Below10 = sum(se <= cutoff_e["10%"])/length(unique(elp_use$Stimulus)),
@@
           Percent_Below90 = sum(se <= cutoff_e["90%"])/length(unique(elp_use$Stimulus))) %>% 
+  ungroup() %>% 
+  group_by(sample_size) %>% 
+  summarize(Percent_Below10 = mean(Percent_Below10),
+            Percent_Below20 = mean(Percent_Below20),
+            Percent_Below30 = mean(Percent_Below30),
+            Percent_Below40 = mean(Percent_Below40),
+            Percent_Below50 = mean(Percent_Below50),
+            Percent_Below60 = mean(Percent_Below60),
+            Percent_Below70 = mean(Percent_Below70),
+            Percent_Below80 = mean(Percent_Below80),
+            Percent_Below90 = mean(Percent_Below90)) %>% 
   pivot_longer(-sample_size) %>% 
@@
 results_table_example <- as.data.frame(results_table)
-colnames(results_table_example) <- c("Deciles", "C Decile SE", "C 80", "C 85", "C 90", "C 95",
-             "L Decile SE", "L 80", "L 85", "L 90", "L 95")
+colnames(results_table_example) <- c("Deciles", "C SE", "C 80", "C 85", "C 90", "C 95",
+             "L SE", "L 80", "L 85", "L 90", "L 95")
 results_table_example$`C Decile SE` <- apa_num(as.numeric(results_table_example$`C Decile SE`), digits = 2)
@@
 apa_table(results_table_example, 
-          caption = "Sample Size Estimates by Decile for Concreteness Example",
-          note = "Esimates are based on meeting at least the minimum percent of items (e.g., 80%) but may be estimated over that amount (e.g., 82.5%).")
+          caption = "Sample Size Estimates by Decile for Example Study",
+          note = "C = Concreteness rating, L = Lexical Decision Response Latencies.
+Estimates are based on meeting at least the minimum percent of items (e.g., 80%) but may be estimated over that amount (e.g., 82.5%).
+SE columns represent the standard error value cutoff for each decile, while 80/85/90/95 percent columns represent the sample size needed to have that percent of items below the SE cutoff.
+For example, 100 participants are required to ensure at least 95% of concreteness items SE are below the 10 percent decile SE cutoff, and 175 participants are necessary for the lexical decision SE to be below its 10 percent decile cutoff.")
 ```
@@
 ```{r scale-size-figure, echo = F, fig.cap="Simulated pilot sample size and final projected sample size to achieve 80%, 85%, 90%, and 95% of items below threshold.
-These values are averaged over all other variables.
+These values are averaged over all other variables including decile.
 Black dots represent original sample size for reference.
 "}
-likert_var <- sqrt(((7-1)^2)/4)
+
 percent_var <- sqrt(((100-0)^2)/4)
@@
 
-```{r scale-skew-figure, echo = F, fig.cap="Simulated pilot sample size compared to projected sample size for ceiling versus normal distributions on each scale. " }
+```{r scale-skew-figure, echo = F, fig.cap="Simulated pilot sample size and final projected sample size to achieve 80%, 85%, 90%, and 95% of items below threshold.
+In comparison to Figure 1, this figure shows projected sample size for ceiling versus normal distributions on each scale.
+All other variables are averaged together, and black dots represent original sample size for reference." }
+
 ggplot(summary_long %>% 
@@
 
-```{r scale-hetero-figure, echo = F, fig.cap="Simulated pilot sample size compared to projected sample size for differing amounts of heterogeneity on each scale. "}
+```{r scale-hetero-figure, echo = F, fig.cap="Simulated pilot sample size and final projected sample size to achieve 80%, 85%, 90%, and 95% of items below threshold.
+In comparison to Figure 1 and 2, this figure shows projected sample size or differing amounts of heterogeneity on each scale.
+All other variables are averaged together, and black dots represent original sample size for reference."}
+
 ggplot(summary_long %>% 
@@
 
-```{r corrected-figure, echo = F, fig.cap="Corrected projected sample sizes for variability and power levels. "}
+```{r corrected-figure, echo = F, fig.cap="Corrected projected sample sizes for variability and power levels to achieve 80%, 85%, 90%, and 95% of items below threshold.
+All other variables are averaged together, and black dots represent original sample size for reference."}
+
 decay <- 1-sqrt((summary_long$original_n-20)/summary_long$original_n)^log2(summary_long$original_n)
@@
 
-The updated proposal steps are in Table \@ref(tab:table-summary).
+The updated proposal steps are in Table \@ref(tab:table-summary) on the right hand side.
 The main change occurs in Step 2 with a designated cutoff decile, and Step 5 with a correction score.
@@
   replace = TRUE, # bootstrap with replacement? 
+  nsim = 5, # number of simulations to run
   grouping_items = "item") # item column label

### 2023-12-28 authorship.R

@@ -1,0 +1,1 @@
+library(tenzing)

### 2023-12-28 manuscript_draft.Rmd
so many updates

@@
 author:
-- name     : "Erin M. Buchanan"
-  affiliation  : "1"
-  corresponding : yes  # Define only one corresponding author
-  address    : "326 Market St., Harrisburg, PA, 17101"
-  email     : "ebuchanan@harrisburgu.edu"
-  role:     # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+  - name: Erin M. Buchanan
+    affiliation: '1'
+    role:
       - Conceptualization
-   - Writing - Original Draft Preparation
-   - Writing - Review & Editing
+      - Data curation
+      - Formal analysis
+      - Investigation
+      - Methodology
+      - Project administration
       - Resources
+      - Software
       - Validation
       - Visualization
-   - Project Administration 
-   - Formal Analysis
-- name     : "Other Folks as Per Order on Doc"
-  affiliation  : "2"
-  role:
-   - Writing - Review & Editing
-   - Data Curation
+      - Writing - original draft
+      - Writing - review & editing
+    corresponding: yes
+    email: ebuchanan@harrisburgu.edu
+    address: Enter postal address here
+  - name: Mahmoud M. Elsherif
+    affiliation: '2'
+    role:
+      - Data curation
       - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: mahmoud.medhat.elsherif@gmail.com
+  - name: Jason Geller
+    affiliation: '3'
+    role:
+      - Data curation
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: jg9120@princeton.edu
+  - name: Chris L. Aberson
+    affiliation: '4'
+    role:
+      - Data curation
+      - Writing - original draft
+      - Writing - review & editing
+    email: chris.aberson@gmail.com
+  - name: Necdet Gurkan
+    affiliation: '5'
+    role:
+      - Data curation
+      - Writing - review & editing
+    email: ngurkan@stevens.edu
+  - name: Ettore Ambrosini
+    affiliation: '6,26'
+    role:
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: ettore.ambrosini@unipd.it
+  - name: Tom Heyman
+    affiliation: '7'
+    role:
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: t.d.p.heyman@fsw.leidenuniv.nl
+  - name: Maria Montefinese
+    affiliation: '8'
+    role:
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: maria.montefinese@gmail.com
+  - name: Wolf Vanpaemel
+    affiliation: '9'
+    role:
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: wolf.vanpaemel@kuleuven.be
+  - name: Krystian Barzykowski
+    affiliation: '10'
+    role:
+      - Data curation
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: krystian.barzykowski@uj.edu.pl
+  - name: Carlota Batres
+    affiliation: '11'
+    role:
+      - Resources
+      - Writing - review & editing
+    email: cbatres@fandm.edu
+  - name: Katharina Fellnhofer
+    affiliation: '12'
+    role:
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: kfellnhofer@ethz.ch
+  - name: Guanxiong Huang
+    affiliation: '13'
+    role:
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: g.huang@cityu.edu.hk
+  - name: Joseph McFall
+    affiliation: '14,27'
+    role:
+      - Resources
+      - Writing - review & editing
+    email: jmcfall@ur.rochester.edu
+  - name: Gianni Ribeiro
+    affiliation: '15'
+    role:
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: g.ribeiro@uq.edu.au
+  - name: Jan P. Röer
+    affiliation: '16'
+    role:
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: jan.roeer@uni-wh.de
+  - name: José L. Ulloa
+    affiliation: '17'
+    role:
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: joseluisulloafulgeri@gmail.com
+  - name: Timo B. Roettger
+    affiliation: '18'
+    role:
+      - Visualization
+      - Formal analysis
+      - Writing - original draft
+      - Writing - review & editing
+    email: timo.roettger@iln.uio.no
+  - name: K. D. Valentine
+    affiliation: '19,28'
+    role:
+      - Conceptualization
+      - Writing - original draft
+      - Writing - review & editing
+    email: KVALENTINE2@mgh.harvard.edu
+  - name: Antonino Visalli
+    affiliation: '20'
+    role:
+      - Writing - original draft
+      - Writing - review & editing
+    email: antonino.visalli.av@gmail.com
+  - name: Kathleen Schmidt
+    affiliation: '21'
+    role:
+      - Writing - original draft
+      - Writing - review & editing
+    email: kschmi13@ashland.edu
+  - name: Martin R. Vasilev
+    affiliation: '22'
+    role:
+      - Writing - original draft
+      - Writing - review & editing
+    email: mvasilev@bournemouth.ac.uk
+  - name: Giada Viviani
+    affiliation: '23'
+    role:
+      - Writing - original draft
+      - Writing - review & editing
+    email: giada.viviani@studenti.unipd.it
+  - name: Jacob F. Miranda
+    affiliation: '24'
+    role:
+      - Project administration
+      - Writing - original draft
+      - Writing - review & editing
+    email: jacob.miranda@csueastbay.edu
+  - name: Savannah C. Lewis
+    affiliation: '25'
+    role:
+      - Project administration
+      - Writing - original draft
+      - Writing - review & editing
+    email: Slewis5920@gmail.com
+
 affiliation:
-- id      : "1"
-  institution  : "Harrisburg University of Science and Technology"
-- id      : "2"
-  institution  : "Other Instituions"
+  - id: '1'
+    institution: Harrisburg University of Science and Technology
+  - id: '2'
+    institution: University of Leicester, Department of Vision Sciences
+  - id: '3'
+    institution: Department of Psychology, Princeton University
+  - id: '4'
+    institution: Illumin Analytics
+  - id: '5'
+    institution: Stevens Institute of Technology
+  - id: '6'
+    institution: Department of Neuroscience, University of Padova, Padova, Italy
+  - id: '7'
+    institution: Methodology and Statistics Unit, Institute of Psychology, Leiden
+      University
+  - id: '8'
+    institution: Department of Developmental and Social Psychology, University of
+      Padova, Italy
+  - id: '9'
+    institution: University of Leuven
+  - id: '10'
+    institution: Applied Memory Research Laboratory, Institute of Psychology, Jagiellonian
+      University
+  - id: '11'
+    institution: Franklin and Marshall College
+  - id: '12'
+    institution: ETH Zürich
+  - id: '13'
+    institution: Department of Media and Communication, City University of Hong Kong
+  - id: '14'
+    institution: Department of Psychology, University of Rochester
+  - id: '15'
+    institution: School of Psychology, The University of Queensland
+  - id: '16'
+    institution: Department of Psychology and Psychotherapy, Witten/Herdecke University
+  - id: '17'
+    institution: Programa de Investigación Asociativa (PIA) en Ciencias Cognitivas,
+      Centro de Investigación en Ciencias Cognitivas (CICC), Facultad de Psicología,
+      Universidad de Talca
+  - id: '18'
+    institution: University of Oslo
+  - id: '19'
+    institution: Massachusetts General Hospital
+  - id: '20'
+    institution: IRCCS San Camillo Hospital (Venice Lido, Italy)
+  - id: '21'
+    institution: Ashland University
+  - id: '22'
+    institution: Bournemouth University
+  - id: '23'
+    institution: University of Padova
+  - id: '24'
+    institution: California State University East Bay
+  - id: '25'
+    institution: University of Alabama, Tuscaloosa
+  - id: '26'
+    institution: Padova Neuroscience Center, University of Padova, Padova, Italy
+  - id: '27'
+    institution: "Children's Institute, Inc"
+  - id: '28'
+    institution: Harvard Medical School
+
 authornote: |
+  Authorship order was determined by tier: 1) Lead author, 2) authors who wrote vignettes, 3) authors who contributed datasets, 4) authors who contributed to conceptualization/writing, and 5) project administration team.
+Within these tiers individuals were ordered by number of CRediT contributions and then alphabetically by last name.
+Data curation was defined as writing vignettes, and resources was defined by submitting datasets with their metadata.
+All other CRediT categories are their traditional interpretation.
 
@@
 However, if reviews of transparency and openness in research publications are any clue, researchers in the social sciences commonly fail to implement proper power analyses as part of their research workflow [@hardwicke2020; @doi:10.1177/1745691620979806].
-The replication "crisis" and credibility revolution have shown that published studies in psychology are underpowered [@opensciencecollaboration2015; @vazire2018, CITE KORBMACHER].
-Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020; @d.chambers2014; @nosek2014], and details of a power analyses or limitations on resources are often used to provide justification for the pre-registered sample quota [CITE (Pownall et al., 2023; van den Akker et al., 2023a, 2023b)].
+The replication "crisis" and credibility revolution have shown that published studies in psychology are underpowered [@opensciencecollaboration2015; @vazire2018; @korbmacher2023].
+Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020; @d.chambers2014; @nosek2014], and details of a power analyses or limitations on resources are often used to provide justification for the pre-registered sample quota [@pownall2023; @vandenakker2023; @akker2023].
 Given the combined issues of publish-or-perish and that most non-significant results do not result in published manuscripts, power analysis may be especially critical for early career researchers to increase the likelihood that they will identify significant effects if they exist [@rosenthal1979; @simmons2011].
-Justified sample sizes may allow for publication of non-significant, yet practically important effects, along with the smallest effect of interest movement [CITE DANIEL/ANVARI], potentially improving the credibility of published work.
+Justified sample sizes may allow for publication of non-significant, yet practically important effects, along with the smallest effect of interest movement [@anvari2021], potentially improving the credibility of published work.
 At best, an underpowered study provides limited insight [@halpern2002], and it can be difficult to know if a poorly implemented power analysis is better than no power analysis.
@@
 G\*Power is one of the most popular free power software options [@faul2007; @erdfelder1996] that provides a simple point and click graphical user interface for power calculations [however, see @brysbaert2019].
-Web-based tools have also sprung up for overall and statistical test specific sample size planning including <https://powerandsamplesize.com>, <https://jakewestfall.shinyapps.io/pangea/> , <https://pwrss.shinyapps.io/index/>, and <https://designingexperiments.com> [@anderson2017].
-*R*-coding based packages, such as *pwr* [@champely2017], *faux* [@debruine2021], *simr* [CITE GREEN], *mixedpower* [CITE KUMLE], and *SimDesign* [@chalmers2020], can be used to examine power and plan sample sizes, usually with simulation.
+Web-based tools have also sprung up for overall and statistical test specific sample size planning including <https://powerandsamplesize.com>, <https://jakewestfall.shinyapps.io/pangea/>, <https://pwrss.shinyapps.io/index/>, and <https://designingexperiments.com> [@anderson2017].
+*R*-coding based packages, such as *pwr* [@champely2017], *faux* [@debruine2021], *simr* [@simr:an], *mixedpower* [@kumle2020], and *SimDesign* [@chalmers2020], can be used to examine power and plan sample sizes, usually with simulation.
 Researchers must be careful using any toolkit, as errors can occur with the over-reliance on software [e.g., it should not be a substitute for critical thinking, @nuijten2016].
@@
 When computing sample size estimates, it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result [@batterham2005].
-For example, it is hard to accurately estimate all parameters from a study, and if any were incorrect, then the sample size estimate tied to that specific level of power may be incorrect [CITE ALBERS].
+For example, it is hard to accurately estimate all parameters from a study, and if any were incorrect, then the sample size estimate tied to that specific level of power may be incorrect [@albers2018].
 
@@
 Changes in publication practices and research design have also created new challenges in providing a sample size plan for a research study.
-While statistics courses often suggest that a specific research design leads to a specific statistical test, meta-science work has shown that given the same data and hypothesis, researchers can come up with multiple ways to analyze the data [@silberzahn2018many, @coretta].
+While statistics courses often suggest that a specific research design leads to a specific statistical test, meta-science work has shown that given the same data and hypothesis, researchers can come up with multiple ways to analyze the data [@silberzahn2018many; @coretta].
 Therefore, a single power analysis only corresponds to the specific analysis that the researcher expects to implement.
@@
 In light of these analytical (or lack thereof) concerns, we propose a new method to determine a sample size in cases where a more traditional power analysis might be less appropriate or even impossible.
-This approach combines accuracy in parameter estimation [AIPE, @kelley2007, @maxwell2008] and bootstrapped simulation on pilot data [@rousselet].
+This approach combines accuracy in parameter estimation [AIPE, @kelley2007; @maxwell2008] and bootstrapped simulation on pilot data [@rousselet].
 This method accounts for a potential lack of hypothesis test (or simply no good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices.
@@
 For each sample size simulation, calculate the standard error for each item.
-Use multiple simulations (e.g., n = 500 to 1000) to avoid issues with random sampling variability.
+Use multiple simulations (e.g., *n* = 500 to 1000) to avoid issues with random sampling variability.
 
@@
 
-The second dataset includes a large scale dataset with response latencies, the English Lexicon Project [ELP; @balota2007].
+The second dataset includes a large scale dataset with response latencies, the English Lexicon Project [ELP, @balota2007].
 The ELP consists of lexical decision response latencies for written English words and pseudowords.
@@
 Further, participants were allowed to indicate they did not know a word, and those responses were set to missing data.
-In our sample of 40 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size_c` as our pilot sample size for this example.
+In our sample of 40 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size_c` as our pilot sample size for the concreteness ratings.
 
@@
 We first filtered the ELP data to the same real words as the concreteness subset selected above, and this data includes `r length(unique(elpDF$Stimulus))` real words.
-The average pilot sample size for this random sample was `r format(round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2), nsmall = 2)` (*SD* = `r round(sd(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2)`), and *n* = `r pilot_size_e` will be our pilot size for this example.
+The average pilot sample size for this random sample was `r format(round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2), nsmall = 2)` (*SD* = `r round(sd(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2)`), and *n* = `r pilot_size_e` will be our pilot size for the lexical decision task.
 
@@
 sim_table <- matrix(NA, 
-          nrow = length(samplesize_values), 
+          nrow = length(samplesize_values)*nsim, 
           ncol = length(unique(elp_use$Stimulus)))
@@
  sim_table %>% 
- pivot_longer(cols = -c(sample_size)) %>% 
+ pivot_longer(cols = -c(sample_size, nsim)) %>% 
  rename(item = name, se = value) %>% 
@@
              "L SE", "L 80", "L 85", "L 90", "L 95")
-results_table_example$`C Decile SE` <- apa_num(as.numeric(results_table_example$`C Decile SE`), digits = 2)
-results_table_example$`L Decile SE` <- apa_num(as.numeric(results_table_example$`L Decile SE`), digits = 2)
+results_table_example$`C SE` <- apa_num(as.numeric(results_table_example$`C SE`), digits = 2)
+results_table_example$`L SE` <- apa_num(as.numeric(results_table_example$`L SE`), digits = 2)
 
@@
 In the last step, the researcher would indicate their smallest sample size, the cutoff standard error criterion if they wanted to adaptively test (e.g., examine the SE after each participant and stop data collection if all items reached criteria), and their maximum sample size.
-As mentioned earlier, the decile for a balanced standard error cutoff is unclear and without guidance, a potential set of researcher degrees of freedom could play a role in the chosen cutoff [CITE SIMONSOHN].
+As mentioned earlier, the decile for a balanced standard error cutoff is unclear and without guidance, a potential set of researcher degrees of freedom could play a role in the chosen cutoff [@simmons2011].
 Even though both measurements (ratings and response latencies) appear to converge on similar sample size suggestions for each decile and percent level, the impact of scale size (i.e., concreteness ratings 1-5 versus response latencies in ms 0-`r max(elp_small$RT, na.rm = T)`) and heterogeneity of item standard errors (concrete $SD_{SD}$ = `r apa_num(sd_items_c)` and lexical $SD_{SD}$ = `r apa_num(sd_items_e)`) is not obvious.
@@
 Standard errors were calculated at each decile of the items up to 90% (i.e., 0% smallest SE, 10% ..., 90% largest SE).
-The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores(Step 2).
+The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores (Step 2).
 
@@
 
+likert_var <- sqrt(((7-1)^2)/4)
 percent_var <- sqrt(((100-0)^2)/4)
@@
   geom_line() + 
+  geom_line(aes(original_n, original_n), color = "black") + 
   theme_linedraw() + 
@@
   ylab("Projected Sample Size") + 
-  scale_color_discrete(name = "Item Heterogeneity") +
-  facet_wrap(~power)
+  scale_color_manual(name = "Item Heterogeneity",
+                     values = c(plot_colors)) +
+  facet_wrap(~power,
+             labeller = labeller(power = power.labs))+
+  theme_bw()
 ```
@@
              labeller = labeller(power = power.labs)) + 
-  ggtitle("Results for 40% deciles") +
+  ggtitle("Results for 50% deciles") +
   xlab("\nPilot Sample Size") + 

### 2023-12-28 manuscript_draft.tex
so many updates

--- 2023-09-08 manuscript_draft.tex
@@
   pdftitle={Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli},
-  pdfauthor={Erin M. Buchanan1 \& Other Folks as Per Order on Doc2},
+  pdfauthor={Erin M.
+Buchanan1, Mahmoud M.
+Elsherif2, Jason Geller3, Chris L.
+Aberson4, Necdet Gurkan5, Ettore Ambrosini6,26, Tom Heyman7, Maria Montefinese8, Wolf Vanpaemel9, Krystian Barzykowski10, Carlota Batres11, Katharina Fellnhofer12, Guanxiong Huang13, Joseph McFall14,27, Gianni Ribeiro15, Jan P.
+Röer16, José L.
+Ulloa17, Timo B.
+Roettger18, K.
+D.
+Valentine19,28, Antonino Visalli20, Kathleen Schmidt21, Martin R.
+Vasilev22, Giada Viviani23, Jacob F.
+Miranda24, \& Savannah C.
+Lewis25},
   pdflang={en-EN},
@@
 \title{Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli}
-\author{Erin M. Buchanan\textsuperscript{1} \& Other Folks as Per Order on Doc\textsuperscript{2}}
+\author{Erin M.
+Buchanan\textsuperscript{1}, Mahmoud M.
+Elsherif\textsuperscript{2}, Jason Geller\textsuperscript{3}, Chris L.
+Aberson\textsuperscript{4}, Necdet Gurkan\textsuperscript{5}, Ettore Ambrosini\textsuperscript{6,26}, Tom Heyman\textsuperscript{7}, Maria Montefinese\textsuperscript{8}, Wolf Vanpaemel\textsuperscript{9}, Krystian Barzykowski\textsuperscript{10}, Carlota Batres\textsuperscript{11}, Katharina Fellnhofer\textsuperscript{12}, Guanxiong Huang\textsuperscript{13}, Joseph McFall\textsuperscript{14,27}, Gianni Ribeiro\textsuperscript{15}, Jan P.
+Röer\textsuperscript{16}, José L.
+Ulloa\textsuperscript{17}, Timo B.
+Roettger\textsuperscript{18}, K.
+D.
+Valentine\textsuperscript{19,28}, Antonino Visalli\textsuperscript{20}, Kathleen Schmidt\textsuperscript{21}, Martin R.
+Vasilev\textsuperscript{22}, Giada Viviani\textsuperscript{23}, Jacob F.
+Miranda\textsuperscript{24}, \& Savannah C.
+Lewis\textsuperscript{25}}
 \date{}
@@
 \authornote{
+
+Authorship order was determined by tier: 1) Lead author, 2) authors who wrote vignettes, 3) authors who contributed datasets, 4) authors who contributed to conceptualization/writing, and 5) project administration team.
+Within these tiers individuals were ordered by number of CRediT contributions and then alphabetically by last name.
+Data curation was defined as writing vignettes, and resources was defined by submitting datasets with their metadata.
+All other CRediT categories are their traditional interpretation.
 
@@
 Erin M.
-Buchanan: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing, Resources, Validation, Visualization, Project Administration, Formal Analysis; Other Folks as Per Order on Doc: Writing - Review \& Editing, Data Curation, Resources.
-
-Correspondence concerning this article should be addressed to Erin M.
-Buchanan, 326 Market St., Harrisburg, PA, 17101.
-E-mail: \href{mailto:ebuchanan@harrisburgu.edu}{\nolinkurl{ebuchanan@harrisburgu.edu}}
+Buchanan: Conceptualization, Data curation, Formal analysis, Investigation, Methodology, Project administration, Resources, Software, Validation, Visualization, Writing - original draft, Writing - review \& editing; Mahmoud M.
+Elsherif: Data curation, Resources, Writing - original draft, Writing - review \& editing; Jason Geller: Data curation, Resources, Writing - original draft, Writing - review \& editing; Chris L.
+Aberson: Data curation, Writing - original draft, Writing - review \& editing; Necdet Gurkan: Data curation, Writing - review \& editing; Ettore Ambrosini: Resources, Writing - original draft, Writing - review \& editing; Tom Heyman: Resources, Writing - original draft, Writing - review \& editing; Maria Montefinese: Resources, Writing - original draft, Writing - review \& editing; Wolf Vanpaemel: Resources, Writing - original draft, Writing - review \& editing; Krystian Barzykowski: Data curation, Resources, Writing - original draft, Writing - review \& editing; Carlota Batres: Resources, Writing - review \& editing; Katharina Fellnhofer: Resources, Writing - original draft, Writing - review \& editing; Guanxiong Huang: Resources, Writing - original draft, Writing - review \& editing; Joseph McFall: Resources, Writing - review \& editing; Gianni Ribeiro: Resources, Writing - original draft, Writing - review \& editing; Jan P.
+Röer: Resources, Writing - original draft, Writing - review \& editing; José L.
+Ulloa: Resources, Writing - original draft, Writing - review \& editing; Timo B.
+Roettger: Visualization, Formal analysis, Writing - original draft, Writing - review \& editing; K.
+D.
+Valentine: Conceptualization, Writing - original draft, Writing - review \& editing; Antonino Visalli: Writing - original draft, Writing - review \& editing; Kathleen Schmidt: Writing - original draft, Writing - review \& editing; Martin R.
+Vasilev: Writing - original draft, Writing - review \& editing; Giada Viviani: Writing - original draft, Writing - review \& editing; Jacob F.
+Miranda: Project administration, Writing - original draft, Writing - review \& editing; Savannah C.
+Lewis: Project administration, Writing - original draft, Writing - review \& editing.
+
+Correspondence concerning this article should be addressed to Erin M. Buchanan, Enter postal address here. E-mail: \href{mailto:ebuchanan@harrisburgu.edu}{\nolinkurl{ebuchanan@harrisburgu.edu}}
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Harrisburg University of Science and Technology\\\textsuperscript{2} Other Instituions}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Harrisburg University of Science and Technology\\\textsuperscript{2} University of Leicester, Department of Vision Sciences\\\textsuperscript{3} Department of Psychology, Princeton University\\\textsuperscript{4} Illumin Analytics\\\textsuperscript{5} Stevens Institute of Technology\\\textsuperscript{6} Department of Neuroscience, University of Padova, Padova, Italy\\\textsuperscript{7} Methodology and Statistics Unit, Institute of Psychology, Leiden University\\\textsuperscript{8} Department of Developmental and Social Psychology, University of Padova, Italy\\\textsuperscript{9} University of Leuven\\\textsuperscript{10} Applied Memory Research Laboratory, Institute of Psychology, Jagiellonian University\\\textsuperscript{11} Franklin and Marshall College\\\textsuperscript{12} ETH Zürich\\\textsuperscript{13} Department of Media and Communication, City University of Hong Kong\\\textsuperscript{14} Department of Psychology, University of Rochester\\\textsuperscript{15} School of Psychology, The University of Queensland\\\textsuperscript{16} Department of Psychology and Psychotherapy, Witten/Herdecke University\\\textsuperscript{17} Programa de Investigación Asociativa (PIA) en Ciencias Cognitivas, Centro de Investigación en Ciencias Cognitivas (CICC), Facultad de Psicología, Universidad de Talca\\\textsuperscript{18} University of Oslo\\\textsuperscript{19} Massachusetts General Hospital\\\textsuperscript{20} IRCCS San Camillo Hospital (Venice Lido, Italy)\\\textsuperscript{21} Ashland University\\\textsuperscript{22} Bournemouth University\\\textsuperscript{23} University of Padova\\\textsuperscript{24} California State University East Bay\\\textsuperscript{25} University of Alabama, Tuscaloosa\\\textsuperscript{26} Padova Neuroscience Center, University of Padova, Padova, Italy\\\textsuperscript{27} Children's Institute, Inc\\\textsuperscript{28} Harvard Medical School}
 
@@
 
-Statistical power and power analyses are arguably one of the most important components in planning a research study (Cohen, 1990).
-However, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts (Hardwicke et al., 2020, 2022).
-The replication ``crisis'' and credibility revolution have shown that published studies in psychology are underpowered (Open Science Collaboration, 2015; Vazire, 2018).
-Pre-registration of a study involves outlining the study and hypotheses before data collection begins (Chambers et al., 2014; Nosek \& Lakens, 2014; Stewart et al., 2020), and details of a power analyses or other limitations on resources are often used to provide justification for the pre-registered sample quota.
-Given the combined issues of publish-or-perish and that most non-significant results do not result in published manuscripts, one may expect that power analysis would be especially critical for early career researchers (Rosenthal, 1979; Simmons et al., 2011).
+An inevitable decision in almost any empirical research is deciding on the sample size.
+Statistical power and power analyses are arguably some of the most important components in planning a research study and its corresponding sample size (Cohen, 1990).
+However, if reviews of transparency and openness in research publications are any clue, researchers in the social sciences commonly fail to implement proper power analyses as part of their research workflow (Hardwicke et al., 2020, 2022).
+The replication ``crisis'' and credibility revolution have shown that published studies in psychology are underpowered (Korbmacher et al., 2023; Open Science Collaboration, 2015; Vazire, 2018).
+Pre-registration of a study involves outlining the study and hypotheses before data collection begins (Chambers et al., 2014; Nosek \& Lakens, 2014; Stewart et al., 2020), and details of a power analyses or limitations on resources are often used to provide justification for the pre-registered sample quota (Pownall et al., 2023; van den Akker, Assen, et al., 2023; van den Akker, Bakker, et al., 2023).
+Given the combined issues of publish-or-perish and that most non-significant results do not result in published manuscripts, power analysis may be especially critical for early career researchers to increase the likelihood that they will identify significant effects if they exist (Rosenthal, 1979; Simmons et al., 2011).
+Justified sample sizes may allow for publication of non-significant, yet practically important effects, along with the smallest effect of interest movement (Anvari \& Lakens, 2021), potentially improving the credibility of published work.
 At best, an underpowered study provides limited insight (Halpern, 2002), and it can be difficult to know if a poorly implemented power analysis is better than no power analysis.
 
-A recent review of power analyses found - across psychology journal articles - researchers did not provide enough information to understand their power analyses and often chose effect sizes that were unjustified (Beribisky, 2019).
-One solution to this power analysis problem is the plethora of tools made available for researchers to make power computations accessible to non-statisticians with the caveat that a solid education in power is necessary to use these tools.
+A recent review of power analyses found - across behavioral, cognitive, and social science journal articles - researchers did not provide enough information to understand their power analyses and often chose effect sizes that were unjustified (Beribisky, 2019).
+One solution to this power analysis problem is the plethora of tools made available for researchers to make power computations accessible to non-statisticians; however, a solid education in power is necessary to use these tools properly.
 G*Power is one of the most popular free power software options (Erdfelder et al., 1996; Faul et al., 2007) that provides a simple point and click graphical user interface for power calculations (however, see Brysbaert, 2019).
-Web-based tools have also sprung up for overall and statistical test specific sample size planning including \url{https://powerandsamplesize.com} and \url{https://designingexperiments.com} (Anderson et al., 2017).
-\emph{R}-coding based packages, such as \emph{pwr} (Champely et al., 2017), \emph{faux} (DeBruine, 2021), and \emph{SimDesign} (Chalmers \& Adkins, 2020), can be used to examine power and sample size planning usually with simulation.
+Web-based tools have also sprung up for overall and statistical test specific sample size planning including \url{https://powerandsamplesize.com}, \url{https://jakewestfall.shinyapps.io/pangea/}, \url{https://pwrss.shinyapps.io/index/}, and \url{https://designingexperiments.com} (Anderson et al., 2017).
+\emph{R}-coding based packages, such as \emph{pwr} (Champely et al., 2017), \emph{faux} (DeBruine, 2021), \emph{simr} (Green \& MacLeod, 2016), \emph{mixedpower} (Kumle \& DejanDraschkow, 2020), and \emph{SimDesign} (Chalmers \& Adkins, 2020), can be used to examine power and plan sample sizes, usually with simulation.
 Researchers must be careful using any toolkit, as errors can occur with the over-reliance on software (e.g., it should not be a substitute for critical thinking, Nuijten et al., 2016).
+Additionally, many tools assume data normality, place an overemphasis on statistical significance, and may rely on simplified assumptions that do not reflect the actual data.
 When computing sample size estimates, it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result (Batterham \& Atkinson, 2005).
-For example, if any parameter estimated by the researcher was not found in the study (i.e., a smaller effect size than used for the power analysis), then the sample size estimate tied to that specific level of power may be incorrect.
-
-Changes in publication practices and research design have also created a new wrinkle in providing a sample size plan for a research study.
-While statistics courses often suggest that a specific research design leads to a specific statistical test, meta-science work has shown that given the same data and hypothesis, researchers can come up with multiple ways to analyze the data Coretta et al.
-(2023).
-Therefore, a single power analysis only supports the specific analysis in which the researcher expects to test.
+For example, it is hard to accurately estimate all parameters from a study, and if any were incorrect, then the sample size estimate tied to that specific level of power may be incorrect (Albers \& Lakens, 2018).
+
+Power analyses are not always the best tool for the job.
+Changes in publication practices and research design have also created new challenges in providing a sample size plan for a research study.
+While statistics courses often suggest that a specific research design leads to a specific statistical test, meta-science work has shown that given the same data and hypothesis, researchers can come up with multiple ways to analyze the data (Coretta et al., 2023; Silberzahn et al., 2018).
+Therefore, a single power analysis only corresponds to the specific analysis that the researcher expects to implement.
 Analyses may evolve during the research project or be subject to secondary analysis; thus, power and sample size estimation based on one analysis is potentially less useful than previously imagined.
 Further, research projects often have multiple testable hypotheses, but it is unclear which hypothesis or test should be used to estimate sample size with a power analysis.
-Last, research publications may not even have a specific, testable hypothesis, as some publications are intended to curate a large dataset for future reuse (i.e., stimuli database creation, Buchanan et al., 2019).
-
-In light of these analytical (or lack thereof) concerns, we propose a new method that combines accuracy in parameter estimation Maxwell et al.
-(2008) and bootstrapped simulation on pilot data (Rousselet et al., 2022).
-This method accounts for a potential lack of hypothesis test (or simply not a good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices.
-Additionally, we consider the nature of cognitive research designs that use multiple items to measure the phenomena of interest.
-For example, semantic priming is measured with multiple paired stimuli (Meyer \& Schvaneveldt, 1971), which traditionally was analyzed by creating person or item-level averages for an ANOVA (Brysbaert \& Stevens, 2018).
-However, the use of multilevel models with random effects for the stimuli used in a study have shown that we should be careful to assume that all items of a research study have the same ``effect'', as there is often variability in their impact on the outcomes of the study.
+Last, research investigations may not even have a specific, testable hypothesis, as some projects are intended to curate a large dataset for future reuse (i.e., stimuli database creation, Buchanan et al., 2019).
+
+In light of these analytical (or lack thereof) concerns, we propose a new method to determine a sample size in cases where a more traditional power analysis might be less appropriate or even impossible.
+This approach combines accuracy in parameter estimation (AIPE, Kelley, 2007; Maxwell et al., 2008) and bootstrapped simulation on pilot data (Rousselet et al., 2022).
+This method accounts for a potential lack of hypothesis test (or simply no good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices.
+Specifically this manuscript focuses on research designs that use multiple items to measure the phenomena of interest.
+For example, semantic priming is measured with multiple paired stimuli (Meyer \& Schvaneveldt, 1971), which traditionally has been analyzed by creating person or item-level averages to test using an ANOVA (Brysbaert \& Stevens, 2018).
+However, research implementing multilevel models with random effects for the stimuli used in a study has demonstrated potential variability in their impact on outcomes; thus we should be careful not to assume that all items in a research study have the same ``effect''.
 
@@
 AIPE shifts the focus away from finding a significant p-value to finding a parameter that is ``accurately measured''.
-For example, researchers may wish to detect a specific correlation in a study, \emph{r} = .35.
-They could then use AIPE to estimate the sample size needed to find a ``sufficiently narrow'' confidence interval around that correlation.
+For example, researchers may wish to detect a specific mean in a study, \emph{M} = .35.
+They could then use AIPE to estimate the sample size needed to find a ``sufficiently narrow'' confidence interval around that mean.
 Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and confidence intervals.
@@
 
-Bootstrapping involves using data obtained to simulate similar datasets by drawing from the original data with replacement (Efron, 2000; Rousselet et al., 2022).
-Bootstrapping is a form of data simulation that allows one to calculate parameter estimates, confidence intervals, and more to simulate the potential population distribution, shape, and bias.
-Simulation is often paired with making up data for testing analyses and hypotheses based on proposed effect size or suggested population means.
-Generally, we would suggest starting with pilot data of a smaller sample size to understand the variability in potential items used to represent your phenomenon, especially if they are to be used in a larger study.
-However, given some background knowledge about the potential items, one could simulate example pilot data to use in.
-similar manner in our suggested procedure.
-Pilot or simulated data would be used to estimate the variability within items and select a ``sufficiently narrow'' window for item's confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
-Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein many items meet the required confidence interval.
+One form of data simulation is Bootstrapping, which involves using data obtained to simulate similar datasets by drawing from the original data with replacement (Efron, 2000; Rousselet et al., 2022).
+Bootstrapping allows one to calculate parameter estimates, confidence intervals, and to simulate the potential population distribution, shape, and bias.
+Simulation is often paired with re-creating a data set with a similar structure for testing analyses and hypotheses based on proposed effect sizes or suggested population means.
+Generally, we would suggest starting with pilot data of a smaller sample size (e.g., 20 to 50) to understand the variability in potential items used to represent your phenomenon, especially if they are to be used in a larger study.
+However, given some background knowledge about the potential items, one could simulate example pilot data to use in a similar manner in our suggested procedure.
+Pilot or simulated data would be used to estimate the variability within items and select a ``sufficiently narrow'' window for overall item confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
+The advantage to this method over simple power estimation from pilot effect sizes is the multiple simulations to average out potential variability, as well as a shift away from traditional NHST to parameter estimation.
+Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein as many items as required meet the pre-specified confidence interval.
 
@@
 
-Researchers could then use sequential testing to estimate their parameter of interest after each participant's data was collected to determine whether they have achieved their expected width of the confidence interval around that parameter.
-One would set a minimum sample size (i.e., based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the CI is narrow, as defined above).
-Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (i.e., we expect 90\% of items to meet our stopping rule with 100 participants based on simulation).
+Researchers could then use sequential testing to estimate their parameter of interest after each participant's data or at regular intervals during data collection to determine whether they have achieved their expected width of the confidence interval around that parameter.
+One would set a minimum sample size (e.g.,, based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the CI is sufficiently narrow, as defined above).
+Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (e.g., , they expect 90\% of items to meet their stopping rule with 100 participants based on simulation).
 By defining each of these components, researchers could ensure a feasible minimum sample size, a way to stop data collection when goals have been met, and a maximum sample size rule to ensure an actual end to data collection.
-Therefore, we proposed that we should be able to leverage the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for repeated measures studies with expected variability in parameter estimates for items.
-
-\hypertarget{proposed-method-to-calculate-sample-size}{%
-\subsection{Proposed Method to Calculate Sample Size}\label{proposed-method-to-calculate-sample-size}}
-
-Using these ideas, we suggest the following procedure to determine a sample size for each item:
+The maximum stopping rule could also be defined by resources (e.g., two semesters data collection), but nevertheless should be included.
+Therefore, we propose a method that leverages the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates to measure an overall phenomena.
+
+\hypertarget{proposed-method-for-sample-size-planning}{%
+\subsection{Proposed Method for Sample Size Planning}\label{proposed-method-for-sample-size-planning}}
+
+Building on these ideas, we suggest the following procedure to determine a sample size for each item:
+
+\hypertarget{calculate-the-stopping-rule}{%
+\subsubsection{Calculate the Stopping Rule}\label{calculate-the-stopping-rule}}
 
@@
   Use pilot data that closely resembles data you intend to collect.
-This dataset should contain items that are identical or similar to those that will be collected in the study.
+This dataset should contain items that are identical or similar to those that will be implemented in the study.
 In this procedure, it is important to ensure that the data is representative of a larger population of sampled items that you intend to assess.
-Generally, pilot data sample sizes will be small, as the goal would be to determine how many items would be necessary to reach a ``stable'' standard error.
+Generally, pilot data sample sizes will be smaller than the overall intended project (e.g., 20 to 50), as the goal would be to determine how many participants would be necessary to reach a ``stable'' standard error.
 \item
-  Calculate the standard error (SE) in the subset of pilot data associated with each item to create a cutoff score that defines when items are considered ``accurately measured''.
-Next, the simulations below will explore what criterion should be used to determine the cutoff score from the pilot data.
+  For each item in the pilot data, calculate the standard error (SE).
+Select a cutoff SE that defines when items are considered ``accurately measured''.
+The simulations described in the Data Simulation section will explore what criterion should be used to determine the cutoff SE from the pilot data.
+\end{enumerate}
+
+\hypertarget{bootstrap-samples}{%
+\subsubsection{Bootstrap Samples}\label{bootstrap-samples}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\setcounter{enumi}{2}
+\tightlist
 \item
-  Sample, with replacement, from your pilot data using sample sizes starting at 20 participants and increase in small units up to a value that you consider the maximum sample size.
+  Sample, with replacement, from your pilot data using sample sizes starting at a value that you consider the minimal sample size per item and increase in small units up to a value that you consider the maximum sample size.
 We will demonstrate example maximum sample sizes based on the data simulation below; however, a practical maximum sample size may be determined by time (e.g., one semester data collection) or resources (e.g., 200 participants worth of funding).
-Although 20 participants would likely yield imprecise estimates, we suggest this starting minimum for simulation purposes.
+As for the minimal sample size, we suggest using 20 as a reasonable value for simulation purposes.
+For each sample size simulation, calculate the standard error for each item.
+Use multiple simulations (e.g., n = 500 to 1000) to avoid issues with random sampling variability.
+\end{enumerate}
+
+\hypertarget{determine-minimum-maximum-sample-size}{%
+\subsubsection{Determine Minimum, Maximum Sample Size}\label{determine-minimum-maximum-sample-size}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\setcounter{enumi}{3}
 \item
-  For each simulation, calculate the standard error for each item, and use these values to determine the percentage of items that meet the cutoff score determined in Step 2.
+  Use the simulated SEs to determine the percentage of items that meet the cutoff score determined in Step 2.
+Each sample size from Step 3 will have multiple bootstrapped simulations, and therefore, create an average percentage score for each sample size for Step 5.
 \item
@@
 Each researcher can determine which of these is their minimum or maximum sample size (e.g., individuals can choose to use 80\% as a minimum and 90\% as a maximum or use values from Step 3 based on resources).
+\end{enumerate}
+
+\hypertarget{report-results}{%
+\subsubsection{Report Results}\label{report-results}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\setcounter{enumi}{5}
+\tightlist
 \item
-  Report these values, and designate a minimum sample size, the cutoff criterion, and the maximum sample size.
+  Report these values, and designate a minimum sample size, the cutoff/stopping rule criterion, and the maximum sample size.
 Each researcher should also report if they plan to use an adaptive design, which would stop data collection after meeting the cutoff criterion for each item.
@@
 Next, we portray simulations for the proposed procedure and find solutions to streamline and improve the sample size estimation procedure.
+Table \ref{tab:table-summary} shows the results of the simulations and solutions on the right hand side.
 Finally, we include additional resources for researchers to use to implement the estimation procedure.
@@
 
-In this section, we provide two examples of the suggested procedure.
-The first example includes concreteness ratings from Brysbaert et al.
+In this section, we provide an example of the suggested procedure.
+The first dataset includes concreteness ratings from Brysbaert et al.
 (2014).
@@
 
-The second dataset includes a large scale dataset with response latencies, the English Lexicon Project (Balota et al., 2007).
-The English Lexicon Project consists of lexical decision response latencies for English words.
+The second dataset includes a large scale dataset with response latencies, the English Lexicon Project {[}ELP; Balota et al.
+(2007){]}.
+The ELP consists of lexical decision response latencies for written English words and pseudowords.
 In a lexical decision task, participants simply select ``word'' for real words (e.g., \emph{dog}) and ``nonword'' for pseudowords (e.g., \emph{wug}).
 The trial level data is available here: \url{https://elexicon.wustl.edu/}.
-Critically, in each of these examples, the individual trial level data for each item is available to simulate and calculate standard errors on.
+Critically, in each of these datasets, the individual trial level data for each item is available to simulate and calculate standard errors on.
 Data that has been summarized could potentially be used, as long as the original standard deviations for each item were present.
@@
 
-For this example, our researcher wants to determine the differences in response latencies for abstract and concrete words.
-They will select \emph{n} = 40 words from the rating data that are split evenly into abstract and concrete ends of the rating scale.
-In the experiment, each participant will rate the words for their concreteness, and then complete a lexical decision task with these words as the object of interest.
-Using both datasets, we can determine the sample size necessary to ensure adequately measured ratings and response latencies.
+For this example, imagine a researcher who wants to determine the differences in response latencies for abstract and concrete words.
+They will select \emph{n} = 40 words from the rating data from Brysbaert et al.
+(2014) that are split evenly into abstract and concrete ends of the rating scale.
+In the experiment, each participant will be asked to rate the words for their concreteness, and then complete a lexical decision task with these words as the phenomenon of interest.
+Using both datasets and the procedure outlined above, we can determine the sample size necessary to ensure adequately measured concreteness ratings and response latencies.
 
 \emph{Step 1}.
-The concreteness ratings data includes 27250 concepts that were rated for their concreteness.
+The concreteness ratings data includes 27031 concepts that were rated for their concreteness.
 We randomly selected \emph{n} = 20 abstract words (\(M_{Rating}\) \textless= 2) and \emph{n} = 20 concrete words (\(M_{Rating}\) \textgreater= 4).
 In the original study, not every participant rated every word, which created uneven sample sizes for each word.
-Further, participants were allowed to indicate they did not know a word, and this data was set to missing data.
-In our sample of 40 words, the average pilot sample size was 28.23 (\emph{SD} = 1.35), and we will use 28 as our pilot sample size for this example.
-
-The ELP response latency data includes 27250 word-forms, 219 that are listed as non-words, and 27031 real words.
-We selected the same words as the concreteness subset selected above.
-The average pilot sample size for this random sample was 32.80 (\emph{SD} = 0.56), and \emph{n} = 33 will be our pilot size for this example.
+Further, participants were allowed to indicate they did not know a word, and those responses were set to missing data.
+In our sample of 40 words, the average pilot sample size was 28.52 (\emph{SD} = 1.80), and we will use 29 as our pilot sample size for this example.
+
+We first filtered the ELP data to the same real words as the concreteness subset selected above, and this data includes 27031 real words.
+The average pilot sample size for this random sample was 32.67 (\emph{SD} = 0.57), and \emph{n} = 33 will be our pilot size for this example.
 
@@
 \emph{Step 3-5}.
-The pilot data was then simulated, with replacement, with samples from 20 to 300 increasing in units of 5, separately for concreteness and lexical decision times (Step 3).
-The standard error of each item was calculated for the bootstrapped samples (Step 4), and the percentage of items below each potential cutoff was gathered (Step 5).
-The smallest sample size with at least 80\%, 85\%, 90\%, and 95\% of items below the cutoff are reported in Table \ref{tab:table-example} (Step 5).
+The pilot data was then simulated 500 times, with replacement, with samples of 20 to 300 participants per item increasing in units of 5, for concreteness ratings and lexical decision latencies separately (Step 3).
+The standard error of each item was calculated for the bootstrapped samples separately for concreteness ratings and lexical decision times (Step 4), and the percentage of items below each potential cutoff was gathered for each (Step 5).
+The smallest sample size with at least 80\%, 85\%, 90\%, and 95\% of items below the cutoff are reported in Table \ref{tab:table-example} for each task (Step 5).
 
@@
 
-\caption{\label{tab:table-example}Sample Size Estimates by Decile for Concreteness Example}
+\caption{\label{tab:table-example}Sample Size Estimates by Decile for Example Study}
 
@@
 \toprule
-Deciles & \multicolumn{1}{c}{C Decile SE} & \multicolumn{1}{c}{C 80} & \multicolumn{1}{c}{C 85} & \multicolumn{1}{c}{C 90} & \multicolumn{1}{c}{C 95} & \multicolumn{1}{c}{L Decile SE} & \multicolumn{1}{c}{L 80} & \multicolumn{1}{c}{L 85} & \multicolumn{1}{c}{L 90} & \multicolumn{1}{c}{L 95}\\
+Deciles & \multicolumn{1}{c}{C SE} & \multicolumn{1}{c}{C 80} & \multicolumn{1}{c}{C 85} & \multicolumn{1}{c}{C 90} & \multicolumn{1}{c}{C 95} & \multicolumn{1}{c}{L SE} & \multicolumn{1}{c}{L 80} & \multicolumn{1}{c}{L 85} & \multicolumn{1}{c}{L 90} & \multicolumn{1}{c}{L 95}\\
 \midrule
-Decile 10 & 0.13 & 85 & 85 & 95 & 100 & 39.64 & 115 & 135 & 135 & 175\\
-Decile 20 & 0.14 & 70 & 75 & 80 & 90 & 45.67 & 85 & 90 & 110 & 125\\
-Decile 30 & 0.16 & 50 & 55 & 60 & 65 & 46.76 & 85 & 90 & 105 & 115\\
-Decile 40 & 0.18 & 45 & 50 & 50 & 65 & 51.32 & 75 & 75 & 90 & 105\\
-Decile 50 & 0.19 & 35 & 35 & 45 & 50 & 71.17 & 40 & 45 & 45 & 50\\
-Decile 60 & 0.20 & 30 & 35 & 35 & 45 & 78.95 & 35 & 35 & 45 & 50\\
-Decile 70 & 0.21 & 30 & 30 & 35 & 45 & 87.85 & 25 & 35 & 35 & 35\\
-Decile 80 & 0.23 & 25 & 30 & 30 & 40 & 100.45 & 20 & 25 & 25 & 35\\
-Decile 90 & 0.25 & 20 & 20 & 25 & 40 & 121.11 & 20 & 20 & 20 & 20\\
+Decile 10 & 0.11 & 120 & 125 & 135 & 150 & 33.70 & 165 & 195 & 240 & NA\\
+Decile 20 & 0.14 & 65 & 70 & 75 & 85 & 46.88 & 95 & 100 & 135 & 180\\
+Decile 30 & 0.17 & 50 & 50 & 55 & 70 & 50.45 & 70 & 90 & 115 & 160\\
+Decile 40 & 0.18 & 45 & 45 & 50 & 55 & 56.93 & 55 & 70 & 90 & 130\\
+Decile 50 & 0.19 & 40 & 45 & 45 & 50 & 65.23 & 50 & 55 & 70 & 100\\
+Decile 60 & 0.21 & 35 & 40 & 40 & 45 & 72.51 & 40 & 50 & 50 & 75\\
+Decile 70 & 0.21 & 30 & 35 & 40 & 40 & 81.21 & 35 & 40 & 50 & 65\\
+Decile 80 & 0.23 & 30 & 30 & 35 & 40 & 94.19 & 25 & 30 & 35 & 50\\
+Decile 90 & 0.25 & 25 & 25 & 30 & 30 & 114.51 & 20 & 20 & 25 & 35\\
 \bottomrule
@@
 \begin{tablenotes}[para]
-\normalsize{\textit{Note.} Esimates are based on meeting at least the minimum percent of items (e.g., 80\%) but may be estimated over that amount (e.g., 82.5\%).}
+\normalsize{\textit{Note.} C = Concreteness rating, L = Lexical Decision Response Latencies.
+Estimates are based on meeting at least the minimum percent of items (e.g., 80\%) but may be estimated over that amount (e.g., 82.5\%).
+SE columns represent the standard error value cutoff for each decile, while 80/85/90/95 percent columns represent the sample size needed to have that percent of items below the SE cutoff.
+For example, 100 participants are required to ensure at least 95\% of concreteness items SE are below the 10 percent decile SE cutoff, and 175 participants are necessary for the lexical decision SE to be below its 10 percent decile cutoff.}
 \end{tablenotes}
@@
 In the last step, the researcher would indicate their smallest sample size, the cutoff standard error criterion if they wanted to adaptively test (e.g., examine the SE after each participant and stop data collection if all items reached criteria), and their maximum sample size.
-As mentioned earlier, the decile for a balanced standard error cutoff is unclear and without guidance, a potential set of researcher degrees of freedom {[}CITE{]}.
-Even though both scales appear to converge on similar sample size suggestions for each decile and percent level, the impact of scale size (i.e., 1-5 versus 0-2852) and heterogeneity of item standard errors (concrete \(SD_{SD}\) = 0.26 and lexical \(SD_{SD}\) = 116.13) is not obvious.
+As mentioned earlier, the decile for a balanced standard error cutoff is unclear and without guidance, a potential set of researcher degrees of freedom could play a role in the chosen cutoff (Simmons et al., 2011).
+Even though both measurements (ratings and response latencies) appear to converge on similar sample size suggestions for each decile and percent level, the impact of scale size (i.e., concreteness ratings 1-5 versus response latencies in ms 0-3480) and heterogeneity of item standard errors (concrete \(SD_{SD}\) = 0.28 and lexical \(SD_{SD}\) = 140.83) is not obvious.
 Last, by selecting the ends of the distribution for our concreteness words, skew of the distribution may additionally impact our estimates.
@@
 \def\labelenumi{\arabic{enumi})}
+\tightlist
 \item
   How do pilot data influence sample size suggestions?
+\end{enumerate}
 
@@
 Heterogeneity should decrease power (Alexander \& DeShon, 1994; Rheinheimer \& Penfield, 2001), and thus, increased projected sample sizes should be proposed as heterogeneity of item variances increases.
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\setcounter{enumi}{1}
 \item
   Do the results match what one might expect for traditional power curves?
-Power curves are asymptotic, that is, they ``level off'' as sample size increases.
+Power curves are asymptotic; that is, they ``level off'' as sample size increases.
 Therefore, we expect that our procedure should also demonstrate a leveling off effect as pilot data sample size increases.
@@
 First, the scale of the data was manipulated by creating three sets of scales.
-The first scale was mimicked after small rating scales (i.e., 1-7 type style) using a \(\mu\) = 4 with a \(\sigma\) = .25 around the mean to create item mean variability.
+The first scale was mimicked after small rating scales (i.e., 1-7 Likert-type style, treated as interval data) using a \(\mu\) = 4 with a \(\sigma\) = .25 around the mean to create item mean variability.
 The second scale included a larger potential distribution of scores with a \(\mu\) = 50 (\(\sigma\) = 10) imitating a 0-100 scale.
@@
 Finally, for the large scale of the data, \(\sigma\) = 400 with a variance of 50, 100, and 200 for heterogeneity.
+These values were based on the proportion of the overall scale and potential variance.
 
@@
 Standard errors were calculated at each decile of the items up to 90\% (i.e., 0\% smallest SE, 10\% \ldots, 90\% largest SE).
-The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores.
+The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores(Step 2).
 
@@
 \caption{\label{fig:scale-size-figure}Simulated pilot sample size and final projected sample size to achieve 80\%, 85\%, 90\%, and 95\% of items below threshold.
-These values are averaged over all other variables.
+These values are averaged over all other variables including decile.
 Black dots represent original sample size for reference.}
@@
 
-Figure \ref{fig:scale-size-figure} demonstrates the influence of scale size on the results separate by potential cutoff level.
+Figure \ref{fig:scale-size-figure} demonstrates the influence of scale size on the results separated by potential cutoff level.
 The black dots denote the original sample size for reference.
@@
 \includegraphics{manuscript_draft_files/figure-latex/scale-skew-figure-1.pdf}
-\caption{\label{fig:scale-skew-figure}Simulated pilot sample size compared to projected sample size for ceiling versus normal distributions on each scale.}
+\caption{\label{fig:scale-skew-figure}Simulated pilot sample size and final projected sample size to achieve 80\%, 85\%, 90\%, and 95\% of items below threshold.
+In comparison to Figure 1, this figure shows projected sample size for ceiling versus normal distributions on each scale.
+All other variables are averaged together, and black dots represent original sample size for reference.}
 \end{figure}
@@
 \includegraphics{manuscript_draft_files/figure-latex/scale-hetero-figure-1.pdf}
-\caption{\label{fig:scale-hetero-figure}Simulated pilot sample size compared to projected sample size for differing amounts of heterogeneity on each scale.}
+\caption{\label{fig:scale-hetero-figure}Simulated pilot sample size and final projected sample size to achieve 80\%, 85\%, 90\%, and 95\% of items below threshold.
+In comparison to Figure 1 and 2, this figure shows projected sample size or differing amounts of heterogeneity on each scale.
+All other variables are averaged together, and black dots represent original sample size for reference.}
 \end{figure}
@@
 \includegraphics{manuscript_draft_files/figure-latex/corrected-figure-1.pdf}
-\caption{\label{fig:corrected-figure}Corrected projected sample sizes for variability and power levels.}
+\caption{\label{fig:corrected-figure}Corrected projected sample sizes for variability and power levels to achieve 80\%, 85\%, 90\%, and 95\% of items below threshold.
+All other variables are averaged together, and black dots represent original sample size for reference.}
 \end{figure}
@@
 To explore if we could recover the new projected sample size from data a researcher would have, we used regression models to create a formula for researcher correction.
-The researcher employing our procedure would have the possible following variables from their simulations on their (one) pilot dataset: 1) proposed sample size, 2) pilot sample size, 3) estimate of heterogeneity for the items, 4) and the estimate percent of items below the threshold.
-Given the non-linear nature of the correction, we added each variable and it's non-linear \texttt{log2} transform to the regression equation, as this function was used to create the correction.
+The researcher employing our procedure would have the possible following variables from their simulations on their (one) pilot dataset: 1) proposed sample size, 2) pilot sample size, 3) estimate of heterogeneity for the items, 4) and the estimated percent of items below the threshold.
+Given the non-linear nature of the correction, we added each variable and its non-linear \texttt{log2} transform to the regression equation, as this function was used to create the correction.
 The intercept only model was used as a starting point (i.e., \texttt{corrected\ sample\ \textasciitilde{}\ 1}), and then all eight variables (each variable and their \texttt{log2} transform) were entered into a forward stepwise regression to capture the corrected scores with the most predictive values.
@@
 Proposed sample size and original sample size were the largest predictors -- unsurprising given the correction formula employed -- followed by the percent ``power'' level and proportion of variance.
-This formula approximation captures \(R^2 = .99\), 90\% CI \([0.99, 0.99]\) of the variance in sample size scores and should allow a researcher to estimate based on their own data, \(F(8, 4,527) = 67,497.54\), \(p < .001\).
+This formula approximation captures \(R^2 = .99\), 90\% CI \([0.99, 0.99]\) of the variance in sample size scores and should allow a researcher to estimate based on their own data, \(F(8, 4527) = 67,497.54\), \(p < .001\).
 We provide convenience functions in our additional materials to assist researchers in estimating the final adjusted sample size.
@@
 If we examine the \(R^2\) values for each decile of our regression equation separately, we find that the values are all \(R^2\) \textgreater{} .99 with very little differences between them.
-Figure \ref{fig:decile-figure} illustrates the corrected scores for simulations at the the 40\% and 50\% decile recommended cutoff for item standard errors.
+Figures \ref{fig:decile40-figure} and \ref{fig:decile50-figure} illustrate the corrected scores for simulations at the 40\% and 50\% decile recommended cutoff for item standard errors.
 For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 40\% decile range, especially for scales with larger potential variance.
@@
 \centering
-\includegraphics{manuscript_draft_files/figure-latex/decile-figure-1.pdf}
-\caption{\label{fig:decile-figure}Comparison of the cutoffs for 40\% and 50\% deciles across heterogeneity (rows), powering of items (columns), and scale size (shape).}
+\includegraphics{manuscript_draft_files/figure-latex/decile40-figure-1.pdf}
+\caption{\label{fig:decile40-figure}Comparison of the cutoffs for 40\% deciles across heterogeneity (columns), powering of items (rows), and scale size (color).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_draft_files/figure-latex/decile50-figure-1.pdf}
+\caption{\label{fig:decile50-figure}Comparison of the cutoffs for 50\% deciles across heterogeneity (columns), powering of items (rows), and scale size (color).}
 \end{figure}
@@
 
-The updated proposal steps are in Table \ref{tab:table-summary}.
+The updated proposal steps are in Table \ref{tab:table-summary} on the right hand side.
 The main change occurs in Step 2 with a designated cutoff decile, and Step 5 with a correction score.
-Using the data from the 40\% decile in Table \ref{tab:table-example}, we can determine that the stopping rule for concreteness ratings would be 0.18, and the stopping rule for lexical decision times would be 51.32.
+Using the data from the 40\% decile in Table \ref{tab:table-example}, we can determine that the stopping rule for concreteness ratings would be 0.18, and the stopping rule for lexical decision times would be 56.93.
 For Step 5, we apply our correction formula separately for each one, as they have different variability scores, and these scores are shown in Table \ref{tab:table-scores-updated}.
@@
 
-For one additional consideration, we calculated the potential amount of data retention given that participants could indicate they did not know a word (\(M_{answered}\) = 0.90, \emph{SD} = 0.13) in the concreteness task or answer a trial incorrectly in the lexical decision task (\(M_{correct}\) = 0.78, \emph{SD} = 0.22).
-In order to account for this facet, the potential sample sizes were multiplied by \(\frac{1}{p_{retained}}\) where the denominator is proportion retained for each task.
+For one additional consideration, we calculated the potential amount of data retention given that participants could indicate they did not know a word (\(M_{answered}\) = 0.93, \emph{SD} = 0.11) in the concreteness task or answer a trial incorrectly in the lexical decision task (\(M_{correct}\) = 0.80, \emph{SD} = 0.21).
+In order to account for this data loss, the potential sample sizes were multiplied by \(\frac{1}{p_{retained}}\) where the denominator is proportion retained for each task.
 
@@
 You can install the package from GitHub using: \texttt{devtools::install\_github("SemanticPriming/semanticprimeR")}.
-We detail the functions below by proposed step in the process.
+We detail the functions below with proposed steps in the process.
 
@@
 ##   item score
-## 1    1     3
-## 2    2     2
-## 3    3     7
+## 1    1     7
+## 2    2     3
+## 3    3     5
 ## 4    4     2
-## 5    5     5
-## 6    6     3
+## 5    5     2
+## 6    6     2
 \end{verbatim}
@@
 \begin{verbatim}
-##  [1] 0.2926737 0.4376973 0.3966969 0.4639646 0.4308804 0.4729527 0.3973597
-##  [8] 0.3734618 0.3439324 0.3933660 0.3346247 0.3879772 0.4466248 0.3324550
-## [15] 0.4530598 0.4500000 0.4660867 0.3590924 0.2523573 0.4345294 0.4844965
-## [22] 0.4805425 0.4167544 0.3234274 0.2926737 0.3371709 0.3838859 0.4285840
-## [29] 0.3802700 0.2325488
+##  [1] 0.4110321 0.4247290 0.4817730 0.3966969 0.4369451 0.4142209 0.4129483
+##  [8] 0.4346808 0.4205260 0.4223992 0.4291975 0.3879772 0.4790341 0.3938675
+## [15] 0.4031129 0.3879772 0.4070109 0.4173853 0.4784844 0.3393104 0.4779342
+## [22] 0.3647277 0.4728135 0.3980214 0.3938675 0.4065257 0.3965310 0.4000000
+## [29] 0.4316431 0.4110321
 \end{verbatim}
@@
 \begin{verbatim}
-## [1] 0.06798869
+## [1] 0.03389213
 \end{verbatim}
@@
 ##       40% 
-## 0.3824396
+## 0.4068168
 \end{verbatim}
@@
 \begin{verbatim}
-## [1] 0.0226629
+## [1] 0.01129738
 \end{verbatim}
@@
 \emph{Step 3}.
-The \texttt{bootstrap\_samples()} function creates bootstrapped samples from the pilot or simulated population data to create samples to estimate the number of participants needed for item standard error to be below the cutoff calculated in Step 2.
+The \texttt{bootstrap\_samples()} function creates bootstrapped samples from the pilot or simulated population data to estimate the number of participants needed for item standard error to be below the cutoff calculated in Step 2.
 This function returns a list of samples with sizes that start at the \texttt{start} size, increase by \texttt{increase}, and end with the \texttt{stop} sample size.
 The population or pilot data will be included in \texttt{population}, and the item column indicator should be included in \texttt{grouping\_items}.
+The \texttt{nsim} argument determines the number of bootstrapped simulations to run.
 
@@
   \AttributeTok{replace =} \ConstantTok{TRUE}\NormalTok{, }\CommentTok{\# bootstrap with replacement? }
+  \AttributeTok{nsim =} \DecValTok{5}\NormalTok{, }\CommentTok{\# number of simulations to run}
   \AttributeTok{grouping\_items =} \StringTok{"item"}\NormalTok{) }\CommentTok{\# item column label  }
@@
 ##   <int> <dbl>
-## 1     1     1
-## 2     1     4
-## 3     1     1
-## 4     1     1
-## 5     1     2
-## 6     1     5
+## 1     1     2
+## 2     1     3
+## 3     1     3
+## 4     1     4
+## 5     1     7
+## 6     1     6
 \end{verbatim}
@@
 The proportion of bootstrapped items across sample sizes below the cutoff score can then be calculated using \texttt{calculate\_proportion()}.
-This function returns a dataframe of each sample size with the proportion of items below that cutoff to use in the next function.
-The \texttt{samples} and \texttt{cutoff} were calculated with our previous functions, with the column for item labels and dependent variable to ensure the right calculations.
+This function returns a dataframe including each sample size with the proportion of items below that cutoff to use in the next function.
+The \texttt{samples} and \texttt{cutoff} arguments were previously calculated with our functions.
+The column for item labels and dependent variables are included as \texttt{grouping\_items} and \texttt{score} arguments to ensure the right calculations.
 
@@
 ##           <dbl>       <dbl>
-## 1         0.467          20
-## 2         0.533          25
-## 3         0.967          30
-## 4         0.967          35
+## 1         0.367          20
+## 2         0.867          25
+## 3         1              30
+## 4         1              35
 ## 5         1              40
@@
 The \texttt{proportion\_summary} from above is used in this function, along with required information about the sample size, proportion of variance from our cutoff calculation, and what power levels should be calculated.
-Note that the exact percent of items below a cutoff score will be returned, if the values in \texttt{power\_levels} are not exactly present.
+Note that the exact percent of items below a cutoff score will be returned if the values in \texttt{power\_levels} are not exactly calculated.
 The final summary presents the smallest sample size, corrected, for each of the potential power levels.
@@
 ##           <dbl>       <dbl>                 <dbl>
-## 1          96.7          30                  25.3
-## 2          96.7          30                  25.3
-## 3          96.7          30                  25.3
-## 4          96.7          30                  25.3
+## 1          86.7          25                  16.9
+## 2          86.7          25                  16.9
+## 3         100            30                  26.9
+## 4         100            30                  26.9
 \end{verbatim}
@@
 
-While the examples in this manuscript are traditionally cognitive linguistics focused, any research using repeated items can benefit from newer sampling techniques.
+While the example in this manuscript was traditionally cognitive linguistics focused, any research using repeated items as a unit of measure could benefit from the proposed newer sampling techniques.
 Therefore, we provide 12 example vignettes and varied code examples on our OSF page/GitHub site for this manuscript across a range of data types provided by the authors of this manuscript.
-Examples include psycholinguistics (De Deyne et al., 2008; Heyman et al., 2014; Montefinese et al., 2022), eye tracking data (Ulloa et al., 2014), social psychology (Grahe et al., 2022; Peterson et al., 2022), COVID related data (Montefinese et al., 2021), and cognitive psychology (Barzykowski et al., 2019; Errington et al., 2021; Röer et al., 2013).
+Examples include psycholinguistics (De Deyne et al., 2008; Heyman et al., 2014; Montefinese et al., 2022), social psychology data (Ulloa et al., 2014), social psychology (Grahe et al., 2022; Peterson et al., 2022), COVID related data (Montefinese et al., 2021), and cognitive psychology (Barzykowski et al., 2019; Errington et al., 2021; Röer et al., 2013).
 
@@
 One limitation of these methods would be our decision to use datasets with very large numbers of items to simulate what might happen within one study.
-For example, the English Lexicon Project includes thousands of items, and by the time we would simulate for all of those, it would likely suggest needing thousands of participants for most items to reach the criterion.
+For example, the English Lexicon Project includes thousands of items, and if we were to simulate for all of those, our results would likely suggest needing thousands of participants for most items to reach the criterion.
 Additionally, as the number of items increases, you may also see very small estimates for sample size due to the correction factor (as with large numbers of items, you could find many items with standard errors below the 40\% decile).
@@
 This result occurs because the smaller number of items means more samples for nearly all to reach the cutoff criteria.
-These results are not too different from what we might expect for a power analysis using a multilevel model - larger numbers of items tend to decrease necessary sample size, while smaller numbers of items tend to increase sample size.
+These results are similar to what we might expect for a power analysis using a multilevel model - larger numbers of items tend to decrease necessary sample size, while smaller numbers of items tend to increase sample size.
 
 Second, these methods do not ensure the normal interpretation of power, where you know that you would find a specific effect for a specific test, \(\alpha\), and so on.
-As discussed in the Introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis, many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
-These methods could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
-
-Researchers should consider this hybrid approach for AIPE, bootstrapping, and simulation as a tool for hypothesis testing and parameter estimation.
-The proposed procedure can be beneficial for many different research studies, specifically replication studies, that usually depend on subject sample size but rarely item sample size, in spite of the fact that item sample sizes contribute to language processing (Baayen et al., 2008; Brysbaert \& Stevens, 2018).
-These observed effects can be then replicated and, as a result of several replications, can be applied to meta-analyses.
-As a result, analysts would be able to use the accuracy and high statistical power to calculate the parameters to assess whether the effect is genuine.
+As discussed in the Introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis; many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
+These proposed methods and traditional power analysis could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
+
+Researchers should consider this hybrid approach for AIPE, bootstrapping, and simulation as a powerful tool for hypothesis testing and parameter estimation.
+This procedure holds benefits for various research studies, specifically replication studies, that usually prioritize subject sample size but rarely item sample size, in spite of the fact that item sample sizes can contribute to power in multilevel models (Brysbaert \& Stevens, 2018).
+Replicated effects, accumulated through multiple studies, contribute to robust meta-analyses, enhancing our understanding of the genuine nature of observed effects.
 This article helps to achieve this goal by encouraging researchers to conduct studies where the power analysis is not based on the size of the effect but on adequate sampling of the stimuli.
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-albers2018}{}}%
+Albers, C., \& Lakens, D.
+(2018).
+When power analyses based on pilot data are biased: Inaccurate effect size estimators and follow-up bias.
+\emph{Journal of Experimental Social Psychology}, \emph{74}, 187--195.
+\url{https://doi.org/10.1016/j.jesp.2017.09.004}
+
 \leavevmode\vadjust pre{\hypertarget{ref-alexander1994}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-anvari2021}{}}%
+Anvari, F., \& Lakens, D.
+(2021).
+Using anchor-based methods to determine the smallest effect size of interest.
+\emph{Journal of Experimental Social Psychology}, \emph{96}, 104159.
+\url{https://doi.org/10.1016/j.jesp.2021.104159}
+
 \leavevmode\vadjust pre{\hypertarget{ref-aust2022}{}}%
@@
 \url{https://CRAN.R-project.org/package=papaja}
-
-\leavevmode\vadjust pre{\hypertarget{ref-baayen2008}{}}%
-Baayen, R.
-H., Davidson, D.
-J., \& Bates, D.
-M.
-(2008).
-Mixed-effects modeling with crossed random effects for subjects and items.
-\emph{Journal of Memory and Language}, \emph{59}(4), 390--412.
-\url{https://doi.org/10.1016/j.jml.2007.12.005}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-simr:an}{}}%
+Green, P., \& MacLeod, C.
+J.
+(2016).
+SIMR: An r package for power analysis of generalized linear mixed models by simulation.
+\emph{Methods in Ecology and Evolution}, \emph{7}(4), 493--498.
+https://doi.org/\url{https://doi.org/10.1111/2041-210X.12504}
+
 \leavevmode\vadjust pre{\hypertarget{ref-halpern2002}{}}%
@@
 \url{https://doi.org/10.3758/BF03192966}
+
+\leavevmode\vadjust pre{\hypertarget{ref-korbmacher2023}{}}%
+Korbmacher, M., Azevedo, F., Pennington, C.
+R., Hartmann, H., Pownall, M., Schmidt, K., Elsherif, M., Breznau, N., Robertson, O., Kalandadze, T., Yu, S., Baker, B.
+J., O'Mahony, A., Olsnes, J.
+Ø.-S., Shaw, J.
+J., Gjoneska, B., Yamada, Y., Röer, J.
+P., Murphy, J., \ldots{} Evans, T.
+(2023).
+The replication crisis has led to positive structural, procedural, and community changes.
+\emph{Communications Psychology}, \emph{1}(1), 1--13.
+\url{https://doi.org/10.1038/s44271-023-00003-2}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kumle2020}{}}%
+Kumle, L., \& DejanDraschkow. (2020). \emph{DejanDraschkow/mixedpower: The force awakens}. Zenodo. \url{https://doi.org/10.5281/zenodo.3733023}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-pownall2023}{}}%
+Pownall, M., Pennington, C.
+R., Norris, E., Juanchich, M., Smailes, D., Russell, S., Gooch, D., Evans, T.
+R., Persson, S., Mak, M.
+H.
+C., Tzavella, L., Monk, R., Gough, T., Benwell, C.
+S.
+Y., Elsherif, M., Farran, E., Gallagher-Mitchell, T., Kendrick, L.
+T., Bahnmueller, J., \ldots{} Clark, K.
+(2023).
+Evaluating the Pedagogical Effectiveness of Study Preregistration in the Undergraduate Dissertation.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{6}(4), 25152459231202724.
+\url{https://doi.org/10.1177/25152459231202724}
+
 \leavevmode\vadjust pre{\hypertarget{ref-R-base}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-vandenakker2023}{}}%
+van den Akker, O.
+R., Assen, M.
+A.
+L.
+M.
+van, Bakker, M., Elsherif, M., Wong, T.
+K., \& Wicherts, J.
+M.
+(2023).
+Preregistration in practice: A comparison of preregistered and non-preregistered studies in psychology.
+\emph{Behavior Research Methods}.
+\url{https://doi.org/10.3758/s13428-023-02277-0}
+
+\leavevmode\vadjust pre{\hypertarget{ref-akker2023}{}}%
+van den Akker, O.
+R., Bakker, M., Assen, M.
+A.
+L.
+M.
+van, Pennington, C.
+R., Verweij, L., Elsherif, M., Claesen, A., Gaillard, S.
+D.
+M., Yeung, S.
+K., Frankenberger, J.-L., Krautter, K., Cockcroft, J.
+P., Kreuer, K.
+S., Evans, T.
+R., Heppel, F., Schoch, S.
+F., Korbmacher, M., Yamada, Y., Albayrak-Aydemir, N., \ldots{} Wicherts, J.
+(2023).
+\emph{The effectiveness of preregistration in psychology: Assessing preregistration strictness and preregistration-study consistency}.
+\url{https://doi.org/10.31222/osf.io/h8xjw}
+
 \leavevmode\vadjust pre{\hypertarget{ref-vazire2018}{}}%

### 2023-12-28 authorship.R

@@ -1,1 +1,14 @@
+# devtools::install_github("marton-balazs-kovacs/tenzing")
 library(tenzing)
+library(googlesheets4)
+
+DF <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1KMBSWE_3IPmwN1mgjF6wEWp5rA89iTXqpOQbwdY7Ag0/edit?usp=sharing")
+
+nrow(DF)
+
+DF <- DF %>% 
+  filter(`Authorship Survey` == "yes" | 
+           `Writing - original draft` == TRUE)
+
+nrow(DF)
+cat(print_yaml(DF))

### 2023-12-28 manuscript_draft.Rmd
all the things!

@@
 Black dots represent original sample size for reference.
-"}
+", fig.height=6, fig.width=8}
 
@@
              labeller = labeller(power = power.labs))+
-  theme_bw()
+  theme_bw() +
+  theme(legend.position="bottom")
 ```
@@
 In comparison to Figure 1, this figure shows projected sample size for ceiling versus normal distributions on each scale.
-All other variables are averaged together, and black dots represent original sample size for reference." }
+All other variables are averaged together, and black dots represent original sample size for reference.", fig.height=6, fig.width=8 }
 
@@
 In comparison to Figure 1 and 2, this figure shows projected sample size or differing amounts of heterogeneity on each scale.
-All other variables are averaged together, and black dots represent original sample size for reference."}
+All other variables are averaged together, and black dots represent original sample size for reference.", fig.height=6, fig.width=8}
 
@@
 ```{r corrected-figure, echo = F, fig.cap="Corrected projected sample sizes for variability and power levels to achieve 80%, 85%, 90%, and 95% of items below threshold.
-All other variables are averaged together, and black dots represent original sample size for reference."}
+All other variables are averaged together, and black dots represent original sample size for reference.", fig.height=6, fig.width=8}
 
@@
 
-```{r decile40-figure, fig.cap = "Comparison of the cutoffs for 40% deciles across heterogeneity (columns), powering of items (rows), and scale size (color)."}
+```{r decile40-figure, fig.cap = "Comparison of the cutoffs for 40% deciles across heterogeneity (columns), powering of items (rows), and scale size (color).", fig.height=8, fig.width=8}
 summary_long %>% 
@@
 
-```{r decile50-figure, fig.cap = "Comparison of the cutoffs for 50% deciles across heterogeneity (columns), powering of items (rows), and scale size (color)."}
+```{r decile50-figure, fig.cap = "Comparison of the cutoffs for 50% deciles across heterogeneity (columns), powering of items (rows), and scale size (color).", fig.height=8, fig.width=8}
 summary_long %>% 
@@
           caption = "Applied Correction for Each Proposed Sample Size", 
-          font_size = "footnotesize")
+          font_size = "scriptsize",
+          col.names = c("Formula", "Intercept", 
+                                        "Proj SS", "Pilot SS", 
+                                        "Log Proj SS", "Log Pilot SS", 
+                                        "Log Power", "Prop Var", 
+                                        "Log Prop Var", "Power", 
+                        "Loss", "Cor SS"),
+          landscape = TRUE,
+          note = "SS = Sample Size, Proj = Projected, Prop = Proportion, Var = Variance, Cor = Corrected")
 ```

### 2023-12-28 manuscript_draft.tex
all the things!

@@
 
-A recent review of power analyses found - across behavioral, cognitive, and social science journal articles - researchers did not provide enough information to understand their power analyses and often chose effect sizes that were unjustified (Beribisky, 2019).
+A recent review of power analyses found - across behavioral, cognitive, and social science journal articles - researchers did not provide enough information to understand their power analyses and often chose effect sizes that were unjustified (Beribisky et al., 2019).
 One solution to this power analysis problem is the plethora of tools made available for researchers to make power computations accessible to non-statisticians; however, a solid education in power is necessary to use these tools properly.
@@
 For each sample size simulation, calculate the standard error for each item.
-Use multiple simulations (e.g., n = 500 to 1000) to avoid issues with random sampling variability.
+Use multiple simulations (e.g., \emph{n} = 500 to 1000) to avoid issues with random sampling variability.
 \end{enumerate}
@@
 
-The second dataset includes a large scale dataset with response latencies, the English Lexicon Project {[}ELP; Balota et al.
-(2007){]}.
+The second dataset includes a large scale dataset with response latencies, the English Lexicon Project (ELP, Balota et al., 2007).
 The ELP consists of lexical decision response latencies for written English words and pseudowords.
@@
 Further, participants were allowed to indicate they did not know a word, and those responses were set to missing data.
-In our sample of 40 words, the average pilot sample size was 28.52 (\emph{SD} = 1.80), and we will use 29 as our pilot sample size for this example.
+In our sample of 40 words, the average pilot sample size was 28.52 (\emph{SD} = 1.80), and we will use 29 as our pilot sample size for the concreteness ratings.
 
 We first filtered the ELP data to the same real words as the concreteness subset selected above, and this data includes 27031 real words.
-The average pilot sample size for this random sample was 32.67 (\emph{SD} = 0.57), and \emph{n} = 33 will be our pilot size for this example.
+The average pilot sample size for this random sample was 32.67 (\emph{SD} = 0.57), and \emph{n} = 33 will be our pilot size for the lexical decision task.
 
@@
 Standard errors were calculated at each decile of the items up to 90\% (i.e., 0\% smallest SE, 10\% \ldots, 90\% largest SE).
-The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores(Step 2).
+The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores (Step 2).
 
@@
 
+\begin{lltable}
+
+\begin{TableNotes}[para]
+\normalsize{\textit{Note.} SS = Sample Size, Proj = Projected, Prop = Proportion, Var = Variance, Cor = Corrected}
+\end{TableNotes}
+
+\scriptsize{
+
+\begin{longtable}{llllllllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:table-scores-updated}Applied Correction for Each Proposed Sample Size}\\
+\toprule
+Formula & Intercept & Proj SS & Pilot SS & Log Proj SS & Log Pilot SS & Log Power & Prop Var & Log Prop Var & Power & Loss & Cor SS\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:table-scores-updated} continued}}\\
+\toprule
+Formula & Intercept & Proj SS & Pilot SS & Log Proj SS & Log Pilot SS & Log Power & Prop Var & Log Prop Var & Power & Loss & Cor SS\\
+\midrule
+\endhead
+Formula & 206.59 & 0.37 & -0.77 & 27.54 & 2.58 & -66.15 & 16.40 & -1.37 & 1.09 & NA & NA\\
+Concrete 80 & 1.00 & 45.00 & 29.00 & 5.49 & 4.86 & 6.32 & 0.14 & -2.82 & 80.00 & 39.63 & 42.56\\
+Concrete 85 & 1.00 & 45.00 & 29.00 & 5.49 & 4.86 & 6.41 & 0.14 & -2.82 & 85.00 & 39.29 & 42.19\\
+Concrete 90 & 1.00 & 50.00 & 29.00 & 5.64 & 4.86 & 6.49 & 0.14 & -2.82 & 90.00 & 45.30 & 48.65\\
+Concrete 95 & 1.00 & 55.00 & 29.00 & 5.78 & 4.86 & 6.57 & 0.14 & -2.82 & 95.00 & 51.21 & 54.99\\
+LDT 80 & 1.00 & 55.00 & 33.00 & 5.78 & 5.04 & 6.32 & 0.08 & -3.60 & 80.00 & 48.79 & 61.05\\
+LDT 85 & 1.00 & 70.00 & 33.00 & 6.13 & 5.04 & 6.41 & 0.08 & -3.60 & 85.00 & 63.54 & 79.51\\
+LDT 90 & 1.00 & 90.00 & 33.00 & 6.49 & 5.04 & 6.49 & 0.08 & -3.60 & 90.00 & 80.87 & 101.20\\
+LDT 95 & 1.00 & 130.00 & 33.00 & 7.02 & 5.04 & 6.57 & 0.08 & -3.60 & 95.00 & 110.48 & 138.25\\
+\bottomrule
+\addlinespace
+\insertTableNotes
+\end{longtable}
+
+}
+
+\end{lltable}
+
 The updated proposal steps are in Table \ref{tab:table-summary} on the right hand side.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-beribisky2019}{}}%
-Beribisky, N. (2019). \emph{A Multi-Faceted Mess: A Review of Statistical Power Analysis in Psychology Journal Articles}. \url{https://yorkspace.library.yorku.ca/xmlui/handle/10315/36719}
+Beribisky, N., Alter, U., \& Cribbie, R.
+(2019).
+\emph{A multi-faceted mess: A systematic review of statistical power analysis in psychology journal articles}.
+\url{https://doi.org/10.31234/osf.io/3bdfu}
 
@@
 \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(3), 337356.
+\url{https://doi.org/10.1177/2515245917747646}
 

### 2023-12-29 manuscript_draft.Rmd
finalizing typos and nsim

@@
     email: ebuchanan@harrisburgu.edu
-    address: Enter postal address here
+    address: "326 Market St, Harrisburg, PA, 17101"
   - name: Mahmoud M. Elsherif
@@
     email: ngurkan@stevens.edu
+  - name: Yangyang Yu
+    affiliation: '5'
+    role:
+      - Data curation
+      - Writing - review & editing
+    email: yyu44@stevens.edu
   - name: Ettore Ambrosini
-    affiliation: '6,26'
+    affiliation: '6,27'
     role:
@@
     email: g.huang@cityu.edu.hk
+  - name: Bastian Jaeger
+    affiliation: '14'
+    role:
+      - Resources
+      - Writing - review & editing
+    email: B.Jaeger@tilburguniversity.edu
   - name: Joseph McFall
-    affiliation: '14,27'
+    affiliation: '15,28'
     role:
@@
   - name: Gianni Ribeiro
-    affiliation: '15'
+    affiliation: '16'
     role:
@@
   - name: Jan P. Röer
-    affiliation: '16'
+    affiliation: '17'
     role:
@@
     email: jan.roeer@uni-wh.de
+  - name: Jordan Suchow
+    affiliation: '5'
+    role:
+      - Resources
+      - Writing - review & editing
+    email: jsuchow@stevens.edu
   - name: José L. Ulloa
-    affiliation: '17'
+    affiliation: '18'
     role:
@@
   - name: Timo B. Roettger
-    affiliation: '18'
-    role:
+    affiliation: '19'
+    role:
+      - Formal analysis
       - Visualization
-      - Formal analysis
       - Writing - original draft
@@
   - name: K. D. Valentine
-    affiliation: '19,28'
+    affiliation: '20,29'
     role:
@@
   - name: Antonino Visalli
-    affiliation: '20'
+    affiliation: '21'
     role:
@@
   - name: Kathleen Schmidt
-    affiliation: '21'
+    affiliation: '22'
     role:
@@
   - name: Martin R. Vasilev
-    affiliation: '22'
+    affiliation: '23'
     role:
@@
   - name: Giada Viviani
-    affiliation: '23'
+    affiliation: '24'
     role:
@@
   - name: Jacob F. Miranda
-    affiliation: '24'
+    affiliation: '25'
     role:
@@
   - name: Savannah C. Lewis
-    affiliation: '25'
+    affiliation: '26'
     role:
@@
   - id: '1'
-    institution: Harrisburg University of Science and Technology
+    institution: Analytics, Harrisburg University of Science and Technology
   - id: '2'
-    institution: University of Leicester, Department of Vision Sciences
+    institution: Department of Vision Sciences, University of Leicester
   - id: '3'
@@
   - id: '6'
-    institution: Department of Neuroscience, University of Padova, Padova, Italy
+    institution: Department of Neuroscience, University of Padova
   - id: '7'
@@
     institution: Department of Developmental and Social Psychology, University of
-      Padova, Italy
+      Padova
   - id: '9'
@@
   - id: '14'
+    institution: Tilburg University
+  - id: '15'
     institution: Department of Psychology, University of Rochester
-  - id: '15'
+  - id: '16'
     institution: School of Psychology, The University of Queensland
-  - id: '16'
+  - id: '17'
     institution: Department of Psychology and Psychotherapy, Witten/Herdecke University
-  - id: '17'
+  - id: '18'
     institution: Programa de Investigación Asociativa (PIA) en Ciencias Cognitivas,
@@
       Universidad de Talca
-  - id: '18'
+  - id: '19'
     institution: University of Oslo
-  - id: '19'
+  - id: '20'
     institution: Massachusetts General Hospital
-  - id: '20'
-    institution: IRCCS San Camillo Hospital (Venice Lido, Italy)
   - id: '21'
+    institution: IRCCS San Camillo Hospital
+  - id: '22'
     institution: Ashland University
-  - id: '22'
+  - id: '23'
     institution: Bournemouth University
-  - id: '23'
+  - id: '24'
     institution: University of Padova
-  - id: '24'
+  - id: '25'
     institution: California State University East Bay
-  - id: '25'
-    institution: University of Alabama, Tuscaloosa
   - id: '26'
-    institution: Padova Neuroscience Center, University of Padova, Padova, Italy
+    institution: University of Alabama
   - id: '27'
-    institution: "Children's Institute, Inc"
+    institution: Padova Neuroscience Center, University of Padova
   - id: '28'
+    institution: "Children's Institute Inc."
+  - id: '29'
     institution: Harvard Medical School
@@
 abstract: |
- The planning of the sample size for research studies often focuses on obtaining a significant result given a specified level of power, significance, and an anticipated effect size.
+ The planning of sample size for research studies often focuses on obtaining a significant result given a specified level of power, significance, and an anticipated effect size.
 This planning requires prior knowledge of the study design and a statistical analysis to calculate the proposed sample size.
@@
 In this tutorial, we explore how AIPE and simulation approaches can be combined to accommodate studies that may not have a specific hypothesis test or wish to account for the potential of a multiverse of analyses.
-Specifically, the examples focus on studies that adopt multiple items and suggest that sample sizes can be planned to measure those items adequately and precisely, regardless of statistical test.
-We demonstrate that pilot data can be used to determine a sample size that represents well-measured data.
-This tutorial also provides multiple code vignettes that researchers can adapt and apply to their own measures.
+Specifically, we focus on studies that use multiple items and suggest that sample sizes can be planned to measure those items adequately and precisely, regardless of statistical test.
+This tutorial also provides multiple code vignettes and package functionality that researchers can adapt and apply to their own measures.
 
@@
 Given the combined issues of publish-or-perish and that most non-significant results do not result in published manuscripts, power analysis may be especially critical for early career researchers to increase the likelihood that they will identify significant effects if they exist [@rosenthal1979; @simmons2011].
-Justified sample sizes may allow for publication of non-significant, yet practically important effects, along with the smallest effect of interest movement [@anvari2021], potentially improving the credibility of published work.
-At best, an underpowered study provides limited insight [@halpern2002], and it can be difficult to know if a poorly implemented power analysis is better than no power analysis.
+Justified sample sizes through power analyses may allow for publication of non-significant, yet well measured effects, along with the smallest effect of interest movement [@anvari2021], potentially improving the credibility of published work.
 
@@
 Additionally, many tools assume data normality, place an overemphasis on statistical significance, and may rely on simplified assumptions that do not reflect the actual data.
-When computing sample size estimates, it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result [@batterham2005].
+When computing sample size estimates, it is important to remember that the effects sizes are estimates, not exact calculations guaranteed to produce a specific result [@batterham2005].
 For example, it is hard to accurately estimate all parameters from a study, and if any were incorrect, then the sample size estimate tied to that specific level of power may be incorrect [@albers2018].
 
-Power analyses are not always the best tool for the job.
 Changes in publication practices and research design have also created new challenges in providing a sample size plan for a research study.
@@
 This method accounts for a potential lack of hypothesis test (or simply no good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices.
-Specifically this manuscript focuses on research designs that use multiple items to measure the phenomena of interest.
+Specifically, this manuscript focuses on research designs that use multiple items to measure the phenomena of interest.
 For example, semantic priming is measured with multiple paired stimuli [@meyer1971], which traditionally has been analyzed by creating person or item-level averages to test using an ANOVA [@brysbaert2018].
-However, research implementing multilevel models with random effects for the stimuli used in a study has demonstrated potential variability in their impact on outcomes; thus we should be careful not to assume that all items in a research study have the same "effect".
+However, research implementing multilevel models with random effects for the stimuli has demonstrated potential variability in their impact on outcomes; thus, we should be careful not to assume that all items in a research study have the same "effect".
 
@@
 
-AIPE shifts the focus away from finding a significant p-value to finding a parameter that is "accurately measured".
+AIPE shifts the focus away from finding a significant *p*-value to finding a parameter that is "accurately measured".
 For example, researchers may wish to detect a specific mean in a study, *M* = .35.
 They could then use AIPE to estimate the sample size needed to find a "sufficiently narrow" confidence interval around that mean.
-Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and confidence intervals.
+Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and/or confidence intervals.
 Therefore, they could decide that their 95% confidence interval should be approximately between .20 and .50, and sufficiently narrow could be defined as a width of .30 or .15 on each side.
@@
 
-One form of data simulation is Bootstrapping, which involves using data obtained to simulate similar datasets by drawing from the original data with replacement [@efron2000; @rousselet].
+One form of data simulation is bootstrapping, which involves using data obtained to simulate similar datasets by drawing from the original data with replacement [@efron2000; @rousselet].
 Bootstrapping allows one to calculate parameter estimates, confidence intervals, and to simulate the potential population distribution, shape, and bias.
@@
 However, given some background knowledge about the potential items, one could simulate example pilot data to use in a similar manner in our suggested procedure.
+
 Pilot or simulated data would be used to estimate the variability within items and select a "sufficiently narrow" window for overall item confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
 The advantage to this method over simple power estimation from pilot effect sizes is the multiple simulations to average out potential variability, as well as a shift away from traditional NHST to parameter estimation.
-Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein as many items as required meet the pre-specified confidence interval.
+Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein as many items as required meet the pre-specified well-measured criterion.
 
@@
 Researchers could then use sequential testing to estimate their parameter of interest after each participant's data or at regular intervals during data collection to determine whether they have achieved their expected width of the confidence interval around that parameter.
-One would set a minimum sample size (e.g.,, based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the CI is sufficiently narrow, as defined above).
-Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (e.g., , they expect 90% of items to meet their stopping rule with 100 participants based on simulation).
+One would set a minimum sample size (e.g., based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the confidence interval is sufficiently narrow, as defined above).
+Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (e.g., they expect 90% of items to meet their stopping rule with 100 participants based on simulation).
 By defining each of these components, researchers could ensure a feasible minimum sample size, a way to stop data collection when goals have been met, and a maximum sample size rule to ensure an actual end to data collection.
@@
 In this procedure, it is important to ensure that the data is representative of a larger population of sampled items that you intend to assess.
-Generally, pilot data sample sizes will be smaller than the overall intended project (e.g., 20 to 50), as the goal would be to determine how many participants would be necessary to reach a "stable" standard error.
+Generally, pilot data sample sizes will be smaller than the overall intended project (e.g., 20 to 50), as the goal would be to determine how many participants would be necessary to reach a "stable" standard error for the accurately measured confidence interval rule.
 
@@
 As for the minimal sample size, we suggest using 20 as a reasonable value for simulation purposes.
-For each sample size simulation, calculate the standard error for each item.
+For each sample size simulation, calculate the SE for each item.
 Use multiple simulations (e.g., *n* = 500 to 1000) to avoid issues with random sampling variability.
@@
 
-These steps are summarized in Table \@ref(tab:table-summary).
+These steps are summarized in Table \@ref(tab:table-summary) on the left hand side.
 We will first demonstrate the ideas behind the steps using open data [@brysbaert2014; @balota2007].
@@
 Participants were then asked to rate concreteness of terms using a 1 (*abstract*) to 5 (*concrete*) scale.
-This data represents a small scale dataset that could be used as pilot data for a study using concrete word ratings.
+This data represents a small scale dataset (i.e., the range of the scale of the data is small, 4 points) that could be used as pilot data for a study using concrete word ratings.
 The data is available at <https://osf.io/qpmf4/>.
 
-The second dataset includes a large scale dataset with response latencies, the English Lexicon Project [ELP, @balota2007].
+The second dataset includes a large scale dataset (i.e., wide range of possible data values) with response latencies, the English Lexicon Project [ELP, @balota2007].
 The ELP consists of lexical decision response latencies for written English words and pseudowords.
@@
 Further, participants were allowed to indicate they did not know a word, and those responses were set to missing data.
-In our sample of 40 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size_c` as our pilot sample size for the concreteness ratings.
+In our sample of 40 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size_c` as our pilot sample size for the concreteness ratings (this information will be used in the follow-up to the simulation study).
 
@@
 *Step 2*.
-Table \@ref(tab:table-example) demonstrates the cutoff scores for deciles of the standard errors for the items for the concreteness ratings and lexical decision response latencies.
-A researcher could potentially pick any of these cutoffs or other percentage options not shown here.
+Table \@ref(tab:table-example) demonstrates the cutoff scores for deciles of the SEs for the concreteness ratings and lexical decision response latency items.
+A researcher could potentially pick any of these cutoffs or other percentage options not shown here (e.g., 35%).
 We will use simulation to determine the suggestion that best captures the balance of adequately powering our sample and feasibility.
+This component is explored in the Data Simulation section.
 
 *Step 3-5*.
-The pilot data was then simulated 500 times, with replacement, with samples of 20 to 300 participants per item increasing in units of 5, for concreteness ratings and lexical decision latencies separately (Step 3).
-The standard error of each item was calculated for the bootstrapped samples separately for concreteness ratings and lexical decision times (Step 4), and the percentage of items below each potential cutoff was gathered for each (Step 5).
+The pilot data was then bootstrapped with replacement creating samples of 20 to 300 participants per item increasing in units of 5, for concreteness ratings and lexical decision latencies separately (Step 3).
+Each of these 57 sample sizes was then repeated 500 times.
+The SE of each item was calculated for the bootstrapped samples separately for concreteness ratings and lexical decision times (Step 4), and the average percentage of items for each sample size (averaging across the 500 simulations) below each potential cutoff was gathered for each (Step 5).
 The smallest sample size with at least 80%, 85%, 90%, and 95% of items below the cutoff are reported in Table \@ref(tab:table-example) for each task (Step 5).
@@
 # sequence of sample sizes to try
-nsim <- 5
+nsim <- 500
 samplesize_values <- seq(20, 300, 5)
@@
 *Step 6*.
-In the last step, the researcher would indicate their smallest sample size, the cutoff standard error criterion if they wanted to adaptively test (e.g., examine the SE after each participant and stop data collection if all items reached criteria), and their maximum sample size.
-As mentioned earlier, the decile for a balanced standard error cutoff is unclear and without guidance, a potential set of researcher degrees of freedom could play a role in the chosen cutoff [@simmons2011].
+In the last step, the researcher would indicate their smallest sample size, the cutoff SE criterion if they wanted to adaptively test (e.g., examine the SE after each participant and stop data collection if all items reached criteria), and their maximum sample size.
+As mentioned earlier, the decile for a balanced SE cutoff is unclear and without guidance, a potential set of researcher degrees of freedom could play a role in the chosen cutoff [@simmons2011].
 Even though both measurements (ratings and response latencies) appear to converge on similar sample size suggestions for each decile and percent level, the impact of scale size (i.e., concreteness ratings 1-5 versus response latencies in ms 0-`r max(elp_small$RT, na.rm = T)`) and heterogeneity of item standard errors (concrete $SD_{SD}$ = `r apa_num(sd_items_c)` and lexical $SD_{SD}$ = `r apa_num(sd_items_e)`) is not obvious.
@@
 
-3)  What should the suggested cutoff standard error decile be?
+3)  What should the suggested cutoff standard SE be?
 
@@
 *Population*.
-We simulated data for 30 items using the rnorm function assuming a normal distribution.
-Each item's population data was simulated with 1000 data points.
+We simulated data for 30 items using the `rnorm` function assuming a normal distribution.
+Each items' population data was simulated with 1000 data points.
 Items were rounded to the nearest whole number to mimic scales generally collected by researchers.
@@
 *Data Scale*.
-First, the scale of the data was manipulated by creating three sets of scales.
+The scale of the data was manipulated by creating three sets of scales.
 The first scale was mimicked after small rating scales (i.e., 1-7 Likert-type style, treated as interval data) using a $\mu$ = 4 with a $\sigma$ = .25 around the mean to create item mean variability.
@@
 *Item Heterogeneity*.
-Next, item heterogeneity was included by manipulating the potential for each individual item.
-For small scales, the $\sigma$ = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items.
-For the medium scale of the data, $\sigma$ = 25 with a variance of 4, 8, and 16.
-Finally, for the large scale of the data, $\sigma$ = 400 with a variance of 50, 100, and 200 for heterogeneity.
+Next, item heterogeneity was included by manipulating the potential variance for each individual item.
+For small scales, the variance was set to $\sigma$ = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items.
+For the medium scale of the data, the variance was $\sigma$ = 25 with a variance of 4, 8, and 16.
+Finally, for the large scale of the data, the variance was $\sigma$ = 400 with a variance of 50, 100, and 200 for heterogeneity.
 These values were based on the proportion of the overall scale and potential variance.
@@
 In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items.
-Assuming that each pilot sample represents a dataset that a researcher has collected (Step 1), the standard errors for each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as SE functions as the main component of the formula for normal distribution confidence intervals.
-Standard errors were calculated at each decile of the items up to 90% (i.e., 0% smallest SE, 10% ..., 90% largest SE).
+Assuming that each pilot sample represents a dataset that a researcher has collected (Step 1), the SEs for each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as SE functions as the main component of the formula for normal distribution confidence intervals.
+SEs were calculated at each decile of the items up to 90% (i.e., 0% smallest SE, 10% ..., 90% largest SE).
 The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores (Step 2).
@@
 We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds.
-However, the sample size simulations were estimated over this amount to determine the pattern of suggested sample sizes (i.e., the function between original sample size and projected sample size).
-
-Next, the percentage of items that fall below the cutoff scores and therefore would be considered "well-measured" were calculated for each decile by sample (Step 4).
+However, the sample size simulations were estimated over this amount to determine the pattern of suggested sample sizes (i.e., the function between original pilot sample size and projected sample size).
+
+Next, we calculated the percentage of items that fell below the cutoff score, and therefore, would be considered "well-measured" for each decile by sample (Step 4).
 From these data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90% and 95% of the items fall below the cutoff criterion (Step 5).
@@
 
-Figure \@ref(fig:scale-size-figure) demonstrates the influence of scale size on the results separated by potential cutoff level.
+Figure \@ref(fig:scale-size-figure) demonstrates the influence of scale size on the results separated by potential cutoff decile level.
 The black dots denote the original sample size for reference.
@@
 
-To explore if we could recover the new projected sample size from data a researcher would have, we used regression models to create a formula for researcher correction.
+To explore if we could recover the corrected sample size from data a researcher would have, we used regression models to create a formula for researcher correction.
 The researcher employing our procedure would have the possible following variables from their simulations on their (one) pilot dataset: 1) proposed sample size, 2) pilot sample size, 3) estimate of heterogeneity for the items, 4) and the estimated percent of items below the threshold.
@@
 This formula approximation captures `r apa_print(forward)$estimate$modelfit$r2` of the variance in sample size scores and should allow a researcher to estimate based on their own data, `r apa_print(forward)$statistic$modelfit$r2`.
-We provide convenience functions in our additional materials to assist researchers in estimating the final adjusted sample size.
+We provide convenience functions in our additional materials to assist researchers in estimating the final corrected sample size.
 
@@
 For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 40% decile range, especially for scales with larger potential variance.
-Therefore, we would suggest the 40% decile to overpower each item to determine the minimum and maximum sample size.
+Therefore, we would suggest the 40% decile to overpower each item for Step 2.
 
@@
 The main change occurs in Step 2 with a designated cutoff decile, and Step 5 with a correction score.
-Using the data from the 40% decile in Table \@ref(tab:table-example), we can determine that the stopping rule for concreteness ratings would be `r apa_num(cutoff_c[4])`, and the stopping rule for lexical decision times would be `r apa_num(cutoff_e[4])`.
+Using the data from the 40% decile in Table \@ref(tab:table-example), we can determine that the stopping rule SE for concreteness ratings would be `r apa_num(cutoff_c[4])`, and the stopping rule SE for lexical decision times would be `r apa_num(cutoff_e[4])`.
 For Step 5, we apply our correction formula separately for each one, as they have different variability scores, and these scores are shown in Table \@ref(tab:table-scores-updated).
-Each row was multiplied by row one's formula, and then these scores are summed for the final sample size.
+Each row was multiplied by row one's formula, and then these scores are summed for the final corrected sample size.
 Sample sizes cannot be proportional, so we recommend rounding up to the nearest whole number.
@@
   replace = TRUE, # bootstrap with replacement? 
-  nsim = 5, # number of simulations to run
+  nsim = 500, # number of simulations to run
   grouping_items = "item") # item column label
@@
 
-While the example in this manuscript was traditionally cognitive linguistics focused, any research using repeated items as a unit of measure could benefit from the proposed newer sampling techniques.
+While the example in this manuscript was cognitive linguistics focused, any research using repeated items as a unit of measure could benefit from the proposed newer sampling techniques.
 Therefore, we provide 12 example vignettes and varied code examples on our OSF page/GitHub site for this manuscript across a range of data types provided by the authors of this manuscript.
-Examples include psycholinguistics [@dedeyne2008; @heyman2014; @montefinese2022], social psychology data [@ulloa2014], social psychology [@peterson2022; @grahe2022], COVID related data [@montefinese2021], and cognitive psychology [@röer2013; @10.7554/eLife.71601; @barzykowski2019].
+Examples include psycholinguistics [@dedeyne2008; @heyman2014; @montefinese2022], social psychology data [@ulloa2014; @peterson2022; @grahe2022], COVID related data [@montefinese2021], and cognitive psychology [@röer2013; @10.7554/eLife.71601; @barzykowski2019].
+These can be found on the package tutorial page: https://semanticpriming.github.io/semanticprimeR/.
 
@@
 Second, these methods do not ensure the normal interpretation of power, where you know that you would find a specific effect for a specific test, $\alpha$, and so on.
-As discussed in the Introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis; many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
+As discussed in the introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis; many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
 These proposed methods and traditional power analysis could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
@@
 This procedure holds benefits for various research studies, specifically replication studies, that usually prioritize subject sample size but rarely item sample size, in spite of the fact that item sample sizes can contribute to power in multilevel models [@brysbaert2018].
-Replicated effects, accumulated through multiple studies, contribute to robust meta-analyses, enhancing our understanding of the genuine nature of observed effects.
+Replicated effects, accumulated through multiple studies and accurate measurement, contribute to robust meta-analyses, enhancing our understanding of the genuine nature of observed effects.
 This article helps to achieve this goal by encouraging researchers to conduct studies where the power analysis is not based on the size of the effect but on adequate sampling of the stimuli.

### 2023-12-29 manuscript_draft.tex
finalizing typos and nsim

@@
 Elsherif2, Jason Geller3, Chris L.
-Aberson4, Necdet Gurkan5, Ettore Ambrosini6,26, Tom Heyman7, Maria Montefinese8, Wolf Vanpaemel9, Krystian Barzykowski10, Carlota Batres11, Katharina Fellnhofer12, Guanxiong Huang13, Joseph McFall14,27, Gianni Ribeiro15, Jan P.
-Röer16, José L.
-Ulloa17, Timo B.
-Roettger18, K.
+Aberson4, Necdet Gurkan5, Yangyang Yu5, Ettore Ambrosini6,27, Tom Heyman7, Maria Montefinese8, Wolf Vanpaemel9, Krystian Barzykowski10, Carlota Batres11, Katharina Fellnhofer12, Guanxiong Huang13, Bastian Jaeger14, Joseph McFall15,28, Gianni Ribeiro16, Jan P.
+Röer17, Jordan Suchow5, José L.
+Ulloa18, Timo B.
+Roettger19, K.
 D.
-Valentine19,28, Antonino Visalli20, Kathleen Schmidt21, Martin R.
-Vasilev22, Giada Viviani23, Jacob F.
-Miranda24, \& Savannah C.
-Lewis25},
+Valentine20,29, Antonino Visalli21, Kathleen Schmidt22, Martin R.
+Vasilev23, Giada Viviani24, Jacob F.
+Miranda25, \& Savannah C.
+Lewis26},
   pdflang={en-EN},
@@
 Elsherif\textsuperscript{2}, Jason Geller\textsuperscript{3}, Chris L.
-Aberson\textsuperscript{4}, Necdet Gurkan\textsuperscript{5}, Ettore Ambrosini\textsuperscript{6,26}, Tom Heyman\textsuperscript{7}, Maria Montefinese\textsuperscript{8}, Wolf Vanpaemel\textsuperscript{9}, Krystian Barzykowski\textsuperscript{10}, Carlota Batres\textsuperscript{11}, Katharina Fellnhofer\textsuperscript{12}, Guanxiong Huang\textsuperscript{13}, Joseph McFall\textsuperscript{14,27}, Gianni Ribeiro\textsuperscript{15}, Jan P.
-Röer\textsuperscript{16}, José L.
-Ulloa\textsuperscript{17}, Timo B.
-Roettger\textsuperscript{18}, K.
+Aberson\textsuperscript{4}, Necdet Gurkan\textsuperscript{5}, Yangyang Yu\textsuperscript{5}, Ettore Ambrosini\textsuperscript{6,27}, Tom Heyman\textsuperscript{7}, Maria Montefinese\textsuperscript{8}, Wolf Vanpaemel\textsuperscript{9}, Krystian Barzykowski\textsuperscript{10}, Carlota Batres\textsuperscript{11}, Katharina Fellnhofer\textsuperscript{12}, Guanxiong Huang\textsuperscript{13}, Bastian Jaeger\textsuperscript{14}, Joseph McFall\textsuperscript{15,28}, Gianni Ribeiro\textsuperscript{16}, Jan P.
+Röer\textsuperscript{17}, Jordan Suchow\textsuperscript{5}, José L.
+Ulloa\textsuperscript{18}, Timo B.
+Roettger\textsuperscript{19}, K.
 D.
-Valentine\textsuperscript{19,28}, Antonino Visalli\textsuperscript{20}, Kathleen Schmidt\textsuperscript{21}, Martin R.
-Vasilev\textsuperscript{22}, Giada Viviani\textsuperscript{23}, Jacob F.
-Miranda\textsuperscript{24}, \& Savannah C.
-Lewis\textsuperscript{25}}
+Valentine\textsuperscript{20,29}, Antonino Visalli\textsuperscript{21}, Kathleen Schmidt\textsuperscript{22}, Martin R.
+Vasilev\textsuperscript{23}, Giada Viviani\textsuperscript{24}, Jacob F.
+Miranda\textsuperscript{25}, \& Savannah C.
+Lewis\textsuperscript{26}}
 \date{}
@@
 Elsherif: Data curation, Resources, Writing - original draft, Writing - review \& editing; Jason Geller: Data curation, Resources, Writing - original draft, Writing - review \& editing; Chris L.
-Aberson: Data curation, Writing - original draft, Writing - review \& editing; Necdet Gurkan: Data curation, Writing - review \& editing; Ettore Ambrosini: Resources, Writing - original draft, Writing - review \& editing; Tom Heyman: Resources, Writing - original draft, Writing - review \& editing; Maria Montefinese: Resources, Writing - original draft, Writing - review \& editing; Wolf Vanpaemel: Resources, Writing - original draft, Writing - review \& editing; Krystian Barzykowski: Data curation, Resources, Writing - original draft, Writing - review \& editing; Carlota Batres: Resources, Writing - review \& editing; Katharina Fellnhofer: Resources, Writing - original draft, Writing - review \& editing; Guanxiong Huang: Resources, Writing - original draft, Writing - review \& editing; Joseph McFall: Resources, Writing - review \& editing; Gianni Ribeiro: Resources, Writing - original draft, Writing - review \& editing; Jan P.
-Röer: Resources, Writing - original draft, Writing - review \& editing; José L.
+Aberson: Data curation, Writing - original draft, Writing - review \& editing; Necdet Gurkan: Data curation, Writing - review \& editing; Yangyang Yu: Data curation, Writing - review \& editing; Ettore Ambrosini: Resources, Writing - original draft, Writing - review \& editing; Tom Heyman: Resources, Writing - original draft, Writing - review \& editing; Maria Montefinese: Resources, Writing - original draft, Writing - review \& editing; Wolf Vanpaemel: Resources, Writing - original draft, Writing - review \& editing; Krystian Barzykowski: Data curation, Resources, Writing - original draft, Writing - review \& editing; Carlota Batres: Resources, Writing - review \& editing; Katharina Fellnhofer: Resources, Writing - original draft, Writing - review \& editing; Guanxiong Huang: Resources, Writing - original draft, Writing - review \& editing; Bastian Jaeger: Resources, Writing - review \& editing; Joseph McFall: Resources, Writing - review \& editing; Gianni Ribeiro: Resources, Writing - original draft, Writing - review \& editing; Jan P.
+Röer: Resources, Writing - original draft, Writing - review \& editing; Jordan Suchow: Resources, Writing - review \& editing; José L.
 Ulloa: Resources, Writing - original draft, Writing - review \& editing; Timo B.
-Roettger: Visualization, Formal analysis, Writing - original draft, Writing - review \& editing; K.
+Roettger: Formal analysis, Visualization, Writing - original draft, Writing - review \& editing; K.
 D.
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Harrisburg University of Science and Technology\\\textsuperscript{2} University of Leicester, Department of Vision Sciences\\\textsuperscript{3} Department of Psychology, Princeton University\\\textsuperscript{4} Illumin Analytics\\\textsuperscript{5} Stevens Institute of Technology\\\textsuperscript{6} Department of Neuroscience, University of Padova, Padova, Italy\\\textsuperscript{7} Methodology and Statistics Unit, Institute of Psychology, Leiden University\\\textsuperscript{8} Department of Developmental and Social Psychology, University of Padova, Italy\\\textsuperscript{9} University of Leuven\\\textsuperscript{10} Applied Memory Research Laboratory, Institute of Psychology, Jagiellonian University\\\textsuperscript{11} Franklin and Marshall College\\\textsuperscript{12} ETH Zürich\\\textsuperscript{13} Department of Media and Communication, City University of Hong Kong\\\textsuperscript{14} Department of Psychology, University of Rochester\\\textsuperscript{15} School of Psychology, The University of Queensland\\\textsuperscript{16} Department of Psychology and Psychotherapy, Witten/Herdecke University\\\textsuperscript{17} Programa de Investigación Asociativa (PIA) en Ciencias Cognitivas, Centro de Investigación en Ciencias Cognitivas (CICC), Facultad de Psicología, Universidad de Talca\\\textsuperscript{18} University of Oslo\\\textsuperscript{19} Massachusetts General Hospital\\\textsuperscript{20} IRCCS San Camillo Hospital (Venice Lido, Italy)\\\textsuperscript{21} Ashland University\\\textsuperscript{22} Bournemouth University\\\textsuperscript{23} University of Padova\\\textsuperscript{24} California State University East Bay\\\textsuperscript{25} University of Alabama, Tuscaloosa\\\textsuperscript{26} Padova Neuroscience Center, University of Padova, Padova, Italy\\\textsuperscript{27} Children's Institute, Inc\\\textsuperscript{28} Harvard Medical School}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Analytics, Harrisburg University of Science and Technology\\\textsuperscript{2} Department of Vision Sciences, University of Leicester\\\textsuperscript{3} Department of Psychology, Princeton University\\\textsuperscript{4} Illumin Analytics\\\textsuperscript{5} Stevens Institute of Technology\\\textsuperscript{6} Department of Neuroscience, University of Padova\\\textsuperscript{7} Methodology and Statistics Unit, Institute of Psychology, Leiden University\\\textsuperscript{8} Department of Developmental and Social Psychology, University of Padova\\\textsuperscript{9} University of Leuven\\\textsuperscript{10} Applied Memory Research Laboratory, Institute of Psychology, Jagiellonian University\\\textsuperscript{11} Franklin and Marshall College\\\textsuperscript{12} ETH Zürich\\\textsuperscript{13} Department of Media and Communication, City University of Hong Kong\\\textsuperscript{14} Tilburg University\\\textsuperscript{15} Department of Psychology, University of Rochester\\\textsuperscript{16} School of Psychology, The University of Queensland\\\textsuperscript{17} Department of Psychology and Psychotherapy, Witten/Herdecke University\\\textsuperscript{18} Programa de Investigación Asociativa (PIA) en Ciencias Cognitivas, Centro de Investigación en Ciencias Cognitivas (CICC), Facultad de Psicología, Universidad de Talca\\\textsuperscript{19} University of Oslo\\\textsuperscript{20} Massachusetts General Hospital\\\textsuperscript{21} IRCCS San Camillo Hospital\\\textsuperscript{22} Ashland University\\\textsuperscript{23} Bournemouth University\\\textsuperscript{24} University of Padova\\\textsuperscript{25} California State University East Bay\\\textsuperscript{26} University of Alabama\\\textsuperscript{27} Padova Neuroscience Center, University of Padova\\\textsuperscript{28} Children's Institute Inc.\\\textsuperscript{29} Harvard Medical School}
 

### 2023-12-29 manuscript_draft.Rmd
final final final v 4

@@
     email: ngurkan@stevens.edu
-  - name: Yangyang Yu
-    affiliation: '5'
-    role:
-      - Data curation
-      - Writing - review & editing
-    email: yyu44@stevens.edu
   - name: Ettore Ambrosini
-    affiliation: '6,27'
+    affiliation: '6,26'
     role:
@@
     email: g.huang@cityu.edu.hk
-  - name: Bastian Jaeger
-    affiliation: '14'
-    role:
-      - Resources
-      - Writing - review & editing
-    email: B.Jaeger@tilburguniversity.edu
   - name: Joseph McFall
-    affiliation: '15,28'
+    affiliation: '14,27'
     role:
@@
   - name: Gianni Ribeiro
-    affiliation: '16'
+    affiliation: '15'
     role:
@@
   - name: Jan P. Röer
+    affiliation: '16'
+    role:
+      - Resources
+      - Writing - original draft
+      - Writing - review & editing
+    email: jan.roeer@uni-wh.de
+  - name: José L. Ulloa
     affiliation: '17'
-    role:
-      - Resources
-      - Writing - original draft
-      - Writing - review & editing
-    email: jan.roeer@uni-wh.de
-  - name: Jordan Suchow
-    affiliation: '5'
-    role:
-      - Resources
-      - Writing - review & editing
-    email: jsuchow@stevens.edu
-  - name: José L. Ulloa
-    affiliation: '18'
     role:
@@
   - name: Timo B. Roettger
-    affiliation: '19'
+    affiliation: '18'
     role:
@@
   - name: K. D. Valentine
-    affiliation: '20,29'
+    affiliation: '19,28'
     role:
@@
   - name: Antonino Visalli
-    affiliation: '21'
+    affiliation: '20'
     role:
@@
   - name: Kathleen Schmidt
-    affiliation: '22'
+    affiliation: '21'
     role:
@@
   - name: Martin R. Vasilev
-    affiliation: '23'
+    affiliation: '22'
     role:
@@
   - name: Giada Viviani
-    affiliation: '24'
+    affiliation: '23'
     role:
@@
   - name: Jacob F. Miranda
-    affiliation: '25'
+    affiliation: '24'
     role:
@@
   - name: Savannah C. Lewis
-    affiliation: '26'
+    affiliation: '25'
     role:
@@
   - id: '14'
-    institution: Tilburg University
+    institution: Department of Psychology, University of Rochester
   - id: '15'
-    institution: Department of Psychology, University of Rochester
+    institution: School of Psychology, The University of Queensland
   - id: '16'
-    institution: School of Psychology, The University of Queensland
+    institution: Department of Psychology and Psychotherapy, Witten/Herdecke University
   - id: '17'
-    institution: Department of Psychology and Psychotherapy, Witten/Herdecke University
-  - id: '18'
     institution: Programa de Investigación Asociativa (PIA) en Ciencias Cognitivas,
@@
       Universidad de Talca
+  - id: '18'
+    institution: University of Oslo
   - id: '19'
-    institution: University of Oslo
+    institution: Massachusetts General Hospital
   - id: '20'
-    institution: Massachusetts General Hospital
+    institution: IRCCS San Camillo Hospital
   - id: '21'
-    institution: IRCCS San Camillo Hospital
+    institution: Ashland University
   - id: '22'
-    institution: Ashland University
+    institution: Bournemouth University
   - id: '23'
-    institution: Bournemouth University
+    institution: University of Padova
   - id: '24'
-    institution: University of Padova
+    institution: California State University East Bay
   - id: '25'
-    institution: California State University East Bay
+    institution: University of Alabama
   - id: '26'
-    institution: University of Alabama
+    institution: Padova Neuroscience Center, University of Padova
   - id: '27'
-    institution: Padova Neuroscience Center, University of Padova
+    institution: "Children's Institute Inc."
   - id: '28'
-    institution: "Children's Institute Inc."
-  - id: '29'
     institution: Harvard Medical School
@@
 set.seed(895893)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE)
 library(dplyr)
@@
 nsim <- 500
-samplesize_values <- seq(20, 300, 5)
+samplesize_values <- seq(20, 500, 5)
 # create a blank table for us to save the values in 
@@
 SE columns represent the standard error value cutoff for each decile, while 80/85/90/95 percent columns represent the sample size needed to have that percent of items below the SE cutoff.
-For example, 100 participants are required to ensure at least 95% of concreteness items SE are below the 10 percent decile SE cutoff, and 175 participants are necessary for the lexical decision SE to be below its 10 percent decile cutoff.")
+For example, 150 participants are required to ensure at least 95% of concreteness items SE are below the 10 percent decile SE cutoff, and 345 participants are necessary for the lexical decision SE to be below its 10 percent decile cutoff.")
 ```

### 2023-12-29 manuscript_draft.tex
final final final v 4

--- 2023-12-28 manuscript_draft.tex
@@
 Ulloa: Resources, Writing - original draft, Writing - review \& editing; Timo B.
-Roettger: Visualization, Formal analysis, Writing - original draft, Writing - review \& editing; K.
+Roettger: Formal analysis, Visualization, Writing - original draft, Writing - review \& editing; K.
 D.
@@
 
-Correspondence concerning this article should be addressed to Erin M. Buchanan, Enter postal address here. E-mail: \href{mailto:ebuchanan@harrisburgu.edu}{\nolinkurl{ebuchanan@harrisburgu.edu}}
+Correspondence concerning this article should be addressed to Erin M.
+Buchanan, 326 Market St, Harrisburg, PA, 17101.
+E-mail: \href{mailto:ebuchanan@harrisburgu.edu}{\nolinkurl{ebuchanan@harrisburgu.edu}}
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Harrisburg University of Science and Technology\\\textsuperscript{2} University of Leicester, Department of Vision Sciences\\\textsuperscript{3} Department of Psychology, Princeton University\\\textsuperscript{4} Illumin Analytics\\\textsuperscript{5} Stevens Institute of Technology\\\textsuperscript{6} Department of Neuroscience, University of Padova, Padova, Italy\\\textsuperscript{7} Methodology and Statistics Unit, Institute of Psychology, Leiden University\\\textsuperscript{8} Department of Developmental and Social Psychology, University of Padova, Italy\\\textsuperscript{9} University of Leuven\\\textsuperscript{10} Applied Memory Research Laboratory, Institute of Psychology, Jagiellonian University\\\textsuperscript{11} Franklin and Marshall College\\\textsuperscript{12} ETH Zürich\\\textsuperscript{13} Department of Media and Communication, City University of Hong Kong\\\textsuperscript{14} Department of Psychology, University of Rochester\\\textsuperscript{15} School of Psychology, The University of Queensland\\\textsuperscript{16} Department of Psychology and Psychotherapy, Witten/Herdecke University\\\textsuperscript{17} Programa de Investigación Asociativa (PIA) en Ciencias Cognitivas, Centro de Investigación en Ciencias Cognitivas (CICC), Facultad de Psicología, Universidad de Talca\\\textsuperscript{18} University of Oslo\\\textsuperscript{19} Massachusetts General Hospital\\\textsuperscript{20} IRCCS San Camillo Hospital (Venice Lido, Italy)\\\textsuperscript{21} Ashland University\\\textsuperscript{22} Bournemouth University\\\textsuperscript{23} University of Padova\\\textsuperscript{24} California State University East Bay\\\textsuperscript{25} University of Alabama, Tuscaloosa\\\textsuperscript{26} Padova Neuroscience Center, University of Padova, Padova, Italy\\\textsuperscript{27} Children's Institute, Inc\\\textsuperscript{28} Harvard Medical School}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Analytics, Harrisburg University of Science and Technology\\\textsuperscript{2} Department of Vision Sciences, University of Leicester\\\textsuperscript{3} Department of Psychology, Princeton University\\\textsuperscript{4} Illumin Analytics\\\textsuperscript{5} Stevens Institute of Technology\\\textsuperscript{6} Department of Neuroscience, University of Padova\\\textsuperscript{7} Methodology and Statistics Unit, Institute of Psychology, Leiden University\\\textsuperscript{8} Department of Developmental and Social Psychology, University of Padova\\\textsuperscript{9} University of Leuven\\\textsuperscript{10} Applied Memory Research Laboratory, Institute of Psychology, Jagiellonian University\\\textsuperscript{11} Franklin and Marshall College\\\textsuperscript{12} ETH Zürich\\\textsuperscript{13} Department of Media and Communication, City University of Hong Kong\\\textsuperscript{14} Department of Psychology, University of Rochester\\\textsuperscript{15} School of Psychology, The University of Queensland\\\textsuperscript{16} Department of Psychology and Psychotherapy, Witten/Herdecke University\\\textsuperscript{17} Programa de Investigación Asociativa (PIA) en Ciencias Cognitivas, Centro de Investigación en Ciencias Cognitivas (CICC), Facultad de Psicología, Universidad de Talca\\\textsuperscript{18} University of Oslo\\\textsuperscript{19} Massachusetts General Hospital\\\textsuperscript{20} IRCCS San Camillo Hospital\\\textsuperscript{21} Ashland University\\\textsuperscript{22} Bournemouth University\\\textsuperscript{23} University of Padova\\\textsuperscript{24} California State University East Bay\\\textsuperscript{25} University of Alabama\\\textsuperscript{26} Padova Neuroscience Center, University of Padova\\\textsuperscript{27} Children's Institute Inc.\\\textsuperscript{28} Harvard Medical School}
 
 \abstract{%
-The planning of the sample size for research studies often focuses on obtaining a significant result given a specified level of power, significance, and an anticipated effect size.
+The planning of sample size for research studies often focuses on obtaining a significant result given a specified level of power, significance, and an anticipated effect size.
 This planning requires prior knowledge of the study design and a statistical analysis to calculate the proposed sample size.
@@
 In this tutorial, we explore how AIPE and simulation approaches can be combined to accommodate studies that may not have a specific hypothesis test or wish to account for the potential of a multiverse of analyses.
-Specifically, the examples focus on studies that adopt multiple items and suggest that sample sizes can be planned to measure those items adequately and precisely, regardless of statistical test.
-We demonstrate that pilot data can be used to determine a sample size that represents well-measured data.
-This tutorial also provides multiple code vignettes that researchers can adapt and apply to their own measures.
+Specifically, we focus on studies that use multiple items and suggest that sample sizes can be planned to measure those items adequately and precisely, regardless of statistical test.
+This tutorial also provides multiple code vignettes and package functionality that researchers can adapt and apply to their own measures.
 }
@@
 Given the combined issues of publish-or-perish and that most non-significant results do not result in published manuscripts, power analysis may be especially critical for early career researchers to increase the likelihood that they will identify significant effects if they exist (Rosenthal, 1979; Simmons et al., 2011).
-Justified sample sizes may allow for publication of non-significant, yet practically important effects, along with the smallest effect of interest movement (Anvari \& Lakens, 2021), potentially improving the credibility of published work.
-At best, an underpowered study provides limited insight (Halpern, 2002), and it can be difficult to know if a poorly implemented power analysis is better than no power analysis.
+Justified sample sizes through power analyses may allow for publication of non-significant, yet well measured effects, along with the smallest effect of interest movement (Anvari \& Lakens, 2021), potentially improving the credibility of published work.
 
@@
 Additionally, many tools assume data normality, place an overemphasis on statistical significance, and may rely on simplified assumptions that do not reflect the actual data.
-When computing sample size estimates, it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result (Batterham \& Atkinson, 2005).
+When computing sample size estimates, it is important to remember that the effects sizes are estimates, not exact calculations guaranteed to produce a specific result (Batterham \& Atkinson, 2005).
 For example, it is hard to accurately estimate all parameters from a study, and if any were incorrect, then the sample size estimate tied to that specific level of power may be incorrect (Albers \& Lakens, 2018).
 
-Power analyses are not always the best tool for the job.
 Changes in publication practices and research design have also created new challenges in providing a sample size plan for a research study.
@@
 This method accounts for a potential lack of hypothesis test (or simply no good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices.
-Specifically this manuscript focuses on research designs that use multiple items to measure the phenomena of interest.
+Specifically, this manuscript focuses on research designs that use multiple items to measure the phenomena of interest.
 For example, semantic priming is measured with multiple paired stimuli (Meyer \& Schvaneveldt, 1971), which traditionally has been analyzed by creating person or item-level averages to test using an ANOVA (Brysbaert \& Stevens, 2018).
-However, research implementing multilevel models with random effects for the stimuli used in a study has demonstrated potential variability in their impact on outcomes; thus we should be careful not to assume that all items in a research study have the same ``effect''.
+However, research implementing multilevel models with random effects for the stimuli has demonstrated potential variability in their impact on outcomes; thus, we should be careful not to assume that all items in a research study have the same ``effect''.
 
@@
 
-AIPE shifts the focus away from finding a significant p-value to finding a parameter that is ``accurately measured''.
+AIPE shifts the focus away from finding a significant \emph{p}-value to finding a parameter that is ``accurately measured''.
 For example, researchers may wish to detect a specific mean in a study, \emph{M} = .35.
 They could then use AIPE to estimate the sample size needed to find a ``sufficiently narrow'' confidence interval around that mean.
-Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and confidence intervals.
+Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and/or confidence intervals.
 Therefore, they could decide that their 95\% confidence interval should be approximately between .20 and .50, and sufficiently narrow could be defined as a width of .30 or .15 on each side.
@@
 
-One form of data simulation is Bootstrapping, which involves using data obtained to simulate similar datasets by drawing from the original data with replacement (Efron, 2000; Rousselet et al., 2022).
+One form of data simulation is bootstrapping, which involves using data obtained to simulate similar datasets by drawing from the original data with replacement (Efron, 2000; Rousselet et al., 2022).
 Bootstrapping allows one to calculate parameter estimates, confidence intervals, and to simulate the potential population distribution, shape, and bias.
@@
 However, given some background knowledge about the potential items, one could simulate example pilot data to use in a similar manner in our suggested procedure.
+
 Pilot or simulated data would be used to estimate the variability within items and select a ``sufficiently narrow'' window for overall item confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
 The advantage to this method over simple power estimation from pilot effect sizes is the multiple simulations to average out potential variability, as well as a shift away from traditional NHST to parameter estimation.
-Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein as many items as required meet the pre-specified confidence interval.
+Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein as many items as required meet the pre-specified well-measured criterion.
 
@@
 Researchers could then use sequential testing to estimate their parameter of interest after each participant's data or at regular intervals during data collection to determine whether they have achieved their expected width of the confidence interval around that parameter.
-One would set a minimum sample size (e.g.,, based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the CI is sufficiently narrow, as defined above).
-Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (e.g., , they expect 90\% of items to meet their stopping rule with 100 participants based on simulation).
+One would set a minimum sample size (e.g., based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the confidence interval is sufficiently narrow, as defined above).
+Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (e.g., they expect 90\% of items to meet their stopping rule with 100 participants based on simulation).
 By defining each of these components, researchers could ensure a feasible minimum sample size, a way to stop data collection when goals have been met, and a maximum sample size rule to ensure an actual end to data collection.
@@
 In this procedure, it is important to ensure that the data is representative of a larger population of sampled items that you intend to assess.
-Generally, pilot data sample sizes will be smaller than the overall intended project (e.g., 20 to 50), as the goal would be to determine how many participants would be necessary to reach a ``stable'' standard error.
+Generally, pilot data sample sizes will be smaller than the overall intended project (e.g., 20 to 50), as the goal would be to determine how many participants would be necessary to reach a ``stable'' standard error for the accurately measured confidence interval rule.
 \item
@@
 As for the minimal sample size, we suggest using 20 as a reasonable value for simulation purposes.
-For each sample size simulation, calculate the standard error for each item.
+For each sample size simulation, calculate the SE for each item.
 Use multiple simulations (e.g., \emph{n} = 500 to 1000) to avoid issues with random sampling variability.
@@
 
-These steps are summarized in Table \ref{tab:table-summary}.
+These steps are summarized in Table \ref{tab:table-summary} on the left hand side.
 We will first demonstrate the ideas behind the steps using open data (Balota et al., 2007; Brysbaert et al., 2014).
@@
 Participants were then asked to rate concreteness of terms using a 1 (\emph{abstract}) to 5 (\emph{concrete}) scale.
-This data represents a small scale dataset that could be used as pilot data for a study using concrete word ratings.
+This data represents a small scale dataset (i.e., the range of the scale of the data is small, 4 points) that could be used as pilot data for a study using concrete word ratings.
 The data is available at \url{https://osf.io/qpmf4/}.
 
-The second dataset includes a large scale dataset with response latencies, the English Lexicon Project (ELP, Balota et al., 2007).
+The second dataset includes a large scale dataset (i.e., wide range of possible data values) with response latencies, the English Lexicon Project (ELP, Balota et al., 2007).
 The ELP consists of lexical decision response latencies for written English words and pseudowords.
@@
 Further, participants were allowed to indicate they did not know a word, and those responses were set to missing data.
-In our sample of 40 words, the average pilot sample size was 28.52 (\emph{SD} = 1.80), and we will use 29 as our pilot sample size for the concreteness ratings.
+In our sample of 40 words, the average pilot sample size was 28.52 (\emph{SD} = 1.80), and we will use 29 as our pilot sample size for the concreteness ratings (this information will be used in the follow-up to the simulation study).
 
@@
 \emph{Step 2}.
-Table \ref{tab:table-example} demonstrates the cutoff scores for deciles of the standard errors for the items for the concreteness ratings and lexical decision response latencies.
-A researcher could potentially pick any of these cutoffs or other percentage options not shown here.
+Table \ref{tab:table-example} demonstrates the cutoff scores for deciles of the SEs for the concreteness ratings and lexical decision response latency items.
+A researcher could potentially pick any of these cutoffs or other percentage options not shown here (e.g., 35\%).
 We will use simulation to determine the suggestion that best captures the balance of adequately powering our sample and feasibility.
+This component is explored in the Data Simulation section.
 
 \emph{Step 3-5}.
-The pilot data was then simulated 500 times, with replacement, with samples of 20 to 300 participants per item increasing in units of 5, for concreteness ratings and lexical decision latencies separately (Step 3).
-The standard error of each item was calculated for the bootstrapped samples separately for concreteness ratings and lexical decision times (Step 4), and the percentage of items below each potential cutoff was gathered for each (Step 5).
+The pilot data was then bootstrapped with replacement creating samples of 20 to 300 participants per item increasing in units of 5, for concreteness ratings and lexical decision latencies separately (Step 3).
+Each of these 57 sample sizes was then repeated 500 times.
+The SE of each item was calculated for the bootstrapped samples separately for concreteness ratings and lexical decision times (Step 4), and the average percentage of items for each sample size (averaging across the 500 simulations) below each potential cutoff was gathered for each (Step 5).
 The smallest sample size with at least 80\%, 85\%, 90\%, and 95\% of items below the cutoff are reported in Table \ref{tab:table-example} for each task (Step 5).
@@
 \midrule
-Decile 10 & 0.11 & 120 & 125 & 135 & 150 & 33.70 & 165 & 195 & 240 & NA\\
-Decile 20 & 0.14 & 65 & 70 & 75 & 85 & 46.88 & 95 & 100 & 135 & 180\\
-Decile 30 & 0.17 & 50 & 50 & 55 & 70 & 50.45 & 70 & 90 & 115 & 160\\
-Decile 40 & 0.18 & 45 & 45 & 50 & 55 & 56.93 & 55 & 70 & 90 & 130\\
-Decile 50 & 0.19 & 40 & 45 & 45 & 50 & 65.23 & 50 & 55 & 70 & 100\\
-Decile 60 & 0.21 & 35 & 40 & 40 & 45 & 72.51 & 40 & 50 & 50 & 75\\
-Decile 70 & 0.21 & 30 & 35 & 40 & 40 & 81.21 & 35 & 40 & 50 & 65\\
+Decile 10 & 0.11 & 115 & 125 & 135 & 150 & 33.70 & 170 & 200 & 245 & 345\\
+Decile 20 & 0.14 & 65 & 70 & 75 & 85 & 46.88 & 90 & 105 & 130 & 180\\
+Decile 30 & 0.17 & 50 & 55 & 60 & 65 & 50.45 & 80 & 95 & 115 & 160\\
+Decile 40 & 0.18 & 45 & 45 & 50 & 55 & 56.93 & 60 & 75 & 90 & 125\\
+Decile 50 & 0.19 & 40 & 45 & 45 & 50 & 65.23 & 50 & 60 & 70 & 95\\
+Decile 60 & 0.21 & 35 & 35 & 40 & 45 & 72.51 & 40 & 45 & 60 & 80\\
+Decile 70 & 0.21 & 35 & 35 & 40 & 45 & 81.21 & 30 & 40 & 50 & 65\\
 Decile 80 & 0.23 & 30 & 30 & 35 & 40 & 94.19 & 25 & 30 & 35 & 50\\
-Decile 90 & 0.25 & 25 & 25 & 30 & 30 & 114.51 & 20 & 20 & 25 & 35\\
+Decile 90 & 0.25 & 25 & 30 & 30 & 35 & 114.51 & 20 & 20 & 25 & 35\\
 \bottomrule
@@
 SE columns represent the standard error value cutoff for each decile, while 80/85/90/95 percent columns represent the sample size needed to have that percent of items below the SE cutoff.
-For example, 100 participants are required to ensure at least 95\% of concreteness items SE are below the 10 percent decile SE cutoff, and 175 participants are necessary for the lexical decision SE to be below its 10 percent decile cutoff.}
+For example, 150 participants are required to ensure at least 95\% of concreteness items SE are below the 10 percent decile SE cutoff, and 345 participants are necessary for the lexical decision SE to be below its 10 percent decile cutoff.}
 \end{tablenotes}
@@
 \emph{Step 6}.
-In the last step, the researcher would indicate their smallest sample size, the cutoff standard error criterion if they wanted to adaptively test (e.g., examine the SE after each participant and stop data collection if all items reached criteria), and their maximum sample size.
-As mentioned earlier, the decile for a balanced standard error cutoff is unclear and without guidance, a potential set of researcher degrees of freedom could play a role in the chosen cutoff (Simmons et al., 2011).
+In the last step, the researcher would indicate their smallest sample size, the cutoff SE criterion if they wanted to adaptively test (e.g., examine the SE after each participant and stop data collection if all items reached criteria), and their maximum sample size.
+As mentioned earlier, the decile for a balanced SE cutoff is unclear and without guidance, a potential set of researcher degrees of freedom could play a role in the chosen cutoff (Simmons et al., 2011).
 Even though both measurements (ratings and response latencies) appear to converge on similar sample size suggestions for each decile and percent level, the impact of scale size (i.e., concreteness ratings 1-5 versus response latencies in ms 0-3480) and heterogeneity of item standard errors (concrete \(SD_{SD}\) = 0.28 and lexical \(SD_{SD}\) = 140.83) is not obvious.
@@
 \item
-  What should the suggested cutoff standard error decile be?
+  What should the suggested cutoff standard SE be?
 \end{enumerate}
@@
 \emph{Population}.
-We simulated data for 30 items using the rnorm function assuming a normal distribution.
-Each item's population data was simulated with 1000 data points.
+We simulated data for 30 items using the \texttt{rnorm} function assuming a normal distribution.
+Each items' population data was simulated with 1000 data points.
 Items were rounded to the nearest whole number to mimic scales generally collected by researchers.
@@
 \emph{Data Scale}.
-First, the scale of the data was manipulated by creating three sets of scales.
+The scale of the data was manipulated by creating three sets of scales.
 The first scale was mimicked after small rating scales (i.e., 1-7 Likert-type style, treated as interval data) using a \(\mu\) = 4 with a \(\sigma\) = .25 around the mean to create item mean variability.
@@
 \emph{Item Heterogeneity}.
-Next, item heterogeneity was included by manipulating the potential for each individual item.
-For small scales, the \(\sigma\) = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items.
-For the medium scale of the data, \(\sigma\) = 25 with a variance of 4, 8, and 16.
-Finally, for the large scale of the data, \(\sigma\) = 400 with a variance of 50, 100, and 200 for heterogeneity.
+Next, item heterogeneity was included by manipulating the potential variance for each individual item.
+For small scales, the variance was set to \(\sigma\) = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items.
+For the medium scale of the data, the variance was \(\sigma\) = 25 with a variance of 4, 8, and 16.
+Finally, for the large scale of the data, the variance was \(\sigma\) = 400 with a variance of 50, 100, and 200 for heterogeneity.
 These values were based on the proportion of the overall scale and potential variance.
@@
 In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items.
-Assuming that each pilot sample represents a dataset that a researcher has collected (Step 1), the standard errors for each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as SE functions as the main component of the formula for normal distribution confidence intervals.
-Standard errors were calculated at each decile of the items up to 90\% (i.e., 0\% smallest SE, 10\% \ldots, 90\% largest SE).
+Assuming that each pilot sample represents a dataset that a researcher has collected (Step 1), the SEs for each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as SE functions as the main component of the formula for normal distribution confidence intervals.
+SEs were calculated at each decile of the items up to 90\% (i.e., 0\% smallest SE, 10\% \ldots, 90\% largest SE).
 The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet cutoff scores, while the higher deciles would represent less strict criteria for cutoff scores (Step 2).
@@
 We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds.
-However, the sample size simulations were estimated over this amount to determine the pattern of suggested sample sizes (i.e., the function between original sample size and projected sample size).
-
-Next, the percentage of items that fall below the cutoff scores and therefore would be considered ``well-measured'' were calculated for each decile by sample (Step 4).
+However, the sample size simulations were estimated over this amount to determine the pattern of suggested sample sizes (i.e., the function between original pilot sample size and projected sample size).
+
+Next, we calculated the percentage of items that fell below the cutoff score, and therefore, would be considered ``well-measured'' for each decile by sample (Step 4).
 From these data, we pinpoint the smallest suggested sample size at which 80\%, 85\%, 90\% and 95\% of the items fall below the cutoff criterion (Step 5).
@@
 
-Figure \ref{fig:scale-size-figure} demonstrates the influence of scale size on the results separated by potential cutoff level.
+Figure \ref{fig:scale-size-figure} demonstrates the influence of scale size on the results separated by potential cutoff decile level.
 The black dots denote the original sample size for reference.
@@
 This formula approximation captures \(R^2 = .99\), 90\% CI \([0.99, 0.99]\) of the variance in sample size scores and should allow a researcher to estimate based on their own data, \(F(8, 4527) = 67,497.54\), \(p < .001\).
-We provide convenience functions in our additional materials to assist researchers in estimating the final adjusted sample size.
+We provide convenience functions in our additional materials to assist researchers in estimating the final corrected sample size.
 
@@
 For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 40\% decile range, especially for scales with larger potential variance.
-Therefore, we would suggest the 40\% decile to overpower each item to determine the minimum and maximum sample size.
+Therefore, we would suggest the 40\% decile to overpower each item for Step 2.
 
@@
 Concrete 95 & 1.00 & 55.00 & 29.00 & 5.78 & 4.86 & 6.57 & 0.14 & -2.82 & 95.00 & 51.21 & 54.99\\
-LDT 80 & 1.00 & 55.00 & 33.00 & 5.78 & 5.04 & 6.32 & 0.08 & -3.60 & 80.00 & 48.79 & 61.05\\
-LDT 85 & 1.00 & 70.00 & 33.00 & 6.13 & 5.04 & 6.41 & 0.08 & -3.60 & 85.00 & 63.54 & 79.51\\
+LDT 80 & 1.00 & 60.00 & 33.00 & 5.91 & 5.04 & 6.32 & 0.08 & -3.60 & 80.00 & 54.08 & 67.68\\
+LDT 85 & 1.00 & 75.00 & 33.00 & 6.23 & 5.04 & 6.41 & 0.08 & -3.60 & 85.00 & 68.12 & 85.25\\
 LDT 90 & 1.00 & 90.00 & 33.00 & 6.49 & 5.04 & 6.49 & 0.08 & -3.60 & 90.00 & 80.87 & 101.20\\
-LDT 95 & 1.00 & 130.00 & 33.00 & 7.02 & 5.04 & 6.57 & 0.08 & -3.60 & 95.00 & 110.48 & 138.25\\
+LDT 95 & 1.00 & 125.00 & 33.00 & 6.97 & 5.04 & 6.57 & 0.08 & -3.60 & 95.00 & 107.09 & 134.00\\
 \bottomrule
@@
 The main change occurs in Step 2 with a designated cutoff decile, and Step 5 with a correction score.
-Using the data from the 40\% decile in Table \ref{tab:table-example}, we can determine that the stopping rule for concreteness ratings would be 0.18, and the stopping rule for lexical decision times would be 56.93.
+Using the data from the 40\% decile in Table \ref{tab:table-example}, we can determine that the stopping rule SE for concreteness ratings would be 0.18, and the stopping rule SE for lexical decision times would be 56.93.
 For Step 5, we apply our correction formula separately for each one, as they have different variability scores, and these scores are shown in Table \ref{tab:table-scores-updated}.
-Each row was multiplied by row one's formula, and then these scores are summed for the final sample size.
+Each row was multiplied by row one's formula, and then these scores are summed for the final corrected sample size.
 Sample sizes cannot be proportional, so we recommend rounding up to the nearest whole number.
@@
 ##   item score
-## 1    1     7
-## 2    2     3
-## 3    3     5
-## 4    4     2
-## 5    5     2
-## 6    6     2
+## 1    1     3
+## 2    2     5
+## 3    3     6
+## 4    4     5
+## 5    5     5
+## 6    6     7
 \end{verbatim}
@@
 \begin{verbatim}
-##  [1] 0.4110321 0.4247290 0.4817730 0.3966969 0.4369451 0.4142209 0.4129483
-##  [8] 0.4346808 0.4205260 0.4223992 0.4291975 0.3879772 0.4790341 0.3938675
-## [15] 0.4031129 0.3879772 0.4070109 0.4173853 0.4784844 0.3393104 0.4779342
-## [22] 0.3647277 0.4728135 0.3980214 0.3938675 0.4065257 0.3965310 0.4000000
-## [29] 0.4316431 0.4110321
+##  [1] 0.4285840 0.3618301 0.3561490 0.3211820 0.3938675 0.3661679 0.4679181
+##  [8] 0.2643264 0.3524351 0.2663101 0.4772454 0.4222434 0.4369451 0.4173853
+## [15] 0.3266658 0.3871284 0.3802700 0.3913539 0.4701623 0.3802700 0.4142209
+## [22] 0.3441236 0.3732856 0.4032761 0.4013136 0.3515005 0.3647277 0.3966969
+## [29] 0.3925289 0.3598245
 \end{verbatim}
@@
 \begin{verbatim}
-## [1] 0.03389213
+## [1] 0.05056835
 \end{verbatim}
@@
 ##       40% 
-## 0.4068168
+## 0.3704385
 \end{verbatim}
@@
 \begin{verbatim}
-## [1] 0.01129738
+## [1] 0.01685612
 \end{verbatim}
@@
   \AttributeTok{replace =} \ConstantTok{TRUE}\NormalTok{, }\CommentTok{\# bootstrap with replacement? }
-  \AttributeTok{nsim =} \DecValTok{5}\NormalTok{, }\CommentTok{\# number of simulations to run}
+  \AttributeTok{nsim =} \DecValTok{500}\NormalTok{, }\CommentTok{\# number of simulations to run}
   \AttributeTok{grouping\_items =} \StringTok{"item"}\NormalTok{) }\CommentTok{\# item column label  }
@@
 ##   <int> <dbl>
-## 1     1     2
+## 1     1     4
 ## 2     1     3
-## 3     1     3
-## 4     1     4
-## 5     1     7
-## 6     1     6
+## 3     1     2
+## 4     1     3
+## 5     1     3
+## 6     1     3
 \end{verbatim}
@@
 ##           <dbl>       <dbl>
-## 1         0.367          20
-## 2         0.867          25
-## 3         1              30
-## 4         1              35
+## 1         0.4            20
+## 2         0.8            25
+## 3         0.833          30
+## 4         0.967          35
 ## 5         1              40
@@
 ##           <dbl>       <dbl>                 <dbl>
-## 1          86.7          25                  16.9
-## 2          86.7          25                  16.9
-## 3         100            30                  26.9
-## 4         100            30                  26.9
+## 1          80            25                  16.6
+## 2          96.7          35                  33.7
+## 3          96.7          35                  33.7
+## 4          96.7          35                  33.7
 \end{verbatim}
@@
 
-While the example in this manuscript was traditionally cognitive linguistics focused, any research using repeated items as a unit of measure could benefit from the proposed newer sampling techniques.
+While the example in this manuscript was cognitive linguistics focused, any research using repeated items as a unit of measure could benefit from the proposed newer sampling techniques.
 Therefore, we provide 12 example vignettes and varied code examples on our OSF page/GitHub site for this manuscript across a range of data types provided by the authors of this manuscript.
-Examples include psycholinguistics (De Deyne et al., 2008; Heyman et al., 2014; Montefinese et al., 2022), social psychology data (Ulloa et al., 2014), social psychology (Grahe et al., 2022; Peterson et al., 2022), COVID related data (Montefinese et al., 2021), and cognitive psychology (Barzykowski et al., 2019; Errington et al., 2021; Röer et al., 2013).
+Examples include psycholinguistics (De Deyne et al., 2008; Heyman et al., 2014; Montefinese et al., 2022), social psychology data (Grahe et al., 2022; Peterson et al., 2022; Ulloa et al., 2014), COVID related data (Montefinese et al., 2021), and cognitive psychology (Barzykowski et al., 2019; Errington et al., 2021; Röer et al., 2013).
+These can be found on the package tutorial page: \url{https://semanticpriming.github.io/semanticprimeR/}.
 
@@
 Second, these methods do not ensure the normal interpretation of power, where you know that you would find a specific effect for a specific test, \(\alpha\), and so on.
-As discussed in the Introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis; many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
+As discussed in the introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis; many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
 These proposed methods and traditional power analysis could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
@@
 This procedure holds benefits for various research studies, specifically replication studies, that usually prioritize subject sample size but rarely item sample size, in spite of the fact that item sample sizes can contribute to power in multilevel models (Brysbaert \& Stevens, 2018).
-Replicated effects, accumulated through multiple studies, contribute to robust meta-analyses, enhancing our understanding of the genuine nature of observed effects.
+Replicated effects, accumulated through multiple studies and accurate measurement, contribute to robust meta-analyses, enhancing our understanding of the genuine nature of observed effects.
 This article helps to achieve this goal by encouraging researchers to conduct studies where the power analysis is not based on the size of the effect but on adequate sampling of the stimuli.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-halpern2002}{}}%
-Halpern, S. D. (2002). The Continuing Unethical Conduct of Underpowered Clinical Trials. \emph{JAMA}, \emph{288}(3), 358. \url{https://doi.org/10.1001/jama.288.3.358}
-
 \leavevmode\vadjust pre{\hypertarget{ref-doi:10.1177ux2f1745691620979806}{}}%

### 2023-12-31 manuscript_draft.Rmd
finalization

@@
   - name: Ettore Ambrosini
-    affiliation: '6,26'
+    affiliation: '6'
     role:
@@
   - name: Joseph McFall
-    affiliation: '14,27'
+    affiliation: '14,26'
     role:
@@
   - name: K. D. Valentine
-    affiliation: '19,28'
+    affiliation: '19,27'
     role:
@@
   - id: '26'
-    institution: Padova Neuroscience Center, University of Padova
+    institution: "Children's Institute Inc."
   - id: '27'
-    institution: "Children's Institute Inc."
-  - id: '28'
     institution: Harvard Medical School
@@
 
+# Open Practices
+
+- All data used in this manuscript and vignettes have been cited and can be found on our repository pages. 
+- Manuscript repository with code and data: https://osf.io/swmva/ or https://github.com/SemanticPriming/stimuli-power 
+- Package repository with vignettes and data: https://github.com/SemanticPriming/semanticprimeR
+- We did not pre-register this study, as it was a simulation study. No materials were used.
+
 \newpage

### 2023-12-31 manuscript_draft.tex
finalization

@@
 Elsherif2, Jason Geller3, Chris L.
-Aberson4, Necdet Gurkan5, Ettore Ambrosini6,26, Tom Heyman7, Maria Montefinese8, Wolf Vanpaemel9, Krystian Barzykowski10, Carlota Batres11, Katharina Fellnhofer12, Guanxiong Huang13, Joseph McFall14,27, Gianni Ribeiro15, Jan P.
+Aberson4, Necdet Gurkan5, Ettore Ambrosini6, Tom Heyman7, Maria Montefinese8, Wolf Vanpaemel9, Krystian Barzykowski10, Carlota Batres11, Katharina Fellnhofer12, Guanxiong Huang13, Joseph McFall14,26, Gianni Ribeiro15, Jan P.
 Röer16, José L.
@@
 D.
-Valentine19,28, Antonino Visalli20, Kathleen Schmidt21, Martin R.
+Valentine19,27, Antonino Visalli20, Kathleen Schmidt21, Martin R.
 Vasilev22, Giada Viviani23, Jacob F.
@@
 Elsherif\textsuperscript{2}, Jason Geller\textsuperscript{3}, Chris L.
-Aberson\textsuperscript{4}, Necdet Gurkan\textsuperscript{5}, Ettore Ambrosini\textsuperscript{6,26}, Tom Heyman\textsuperscript{7}, Maria Montefinese\textsuperscript{8}, Wolf Vanpaemel\textsuperscript{9}, Krystian Barzykowski\textsuperscript{10}, Carlota Batres\textsuperscript{11}, Katharina Fellnhofer\textsuperscript{12}, Guanxiong Huang\textsuperscript{13}, Joseph McFall\textsuperscript{14,27}, Gianni Ribeiro\textsuperscript{15}, Jan P.
+Aberson\textsuperscript{4}, Necdet Gurkan\textsuperscript{5}, Ettore Ambrosini\textsuperscript{6}, Tom Heyman\textsuperscript{7}, Maria Montefinese\textsuperscript{8}, Wolf Vanpaemel\textsuperscript{9}, Krystian Barzykowski\textsuperscript{10}, Carlota Batres\textsuperscript{11}, Katharina Fellnhofer\textsuperscript{12}, Guanxiong Huang\textsuperscript{13}, Joseph McFall\textsuperscript{14,26}, Gianni Ribeiro\textsuperscript{15}, Jan P.
 Röer\textsuperscript{16}, José L.
@@
 D.
-Valentine\textsuperscript{19,28}, Antonino Visalli\textsuperscript{20}, Kathleen Schmidt\textsuperscript{21}, Martin R.
+Valentine\textsuperscript{19,27}, Antonino Visalli\textsuperscript{20}, Kathleen Schmidt\textsuperscript{21}, Martin R.
 Vasilev\textsuperscript{22}, Giada Viviani\textsuperscript{23}, Jacob F.
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Analytics, Harrisburg University of Science and Technology\\\textsuperscript{2} Department of Vision Sciences, University of Leicester\\\textsuperscript{3} Department of Psychology, Princeton University\\\textsuperscript{4} Illumin Analytics\\\textsuperscript{5} Stevens Institute of Technology\\\textsuperscript{6} Department of Neuroscience, University of Padova\\\textsuperscript{7} Methodology and Statistics Unit, Institute of Psychology, Leiden University\\\textsuperscript{8} Department of Developmental and Social Psychology, University of Padova\\\textsuperscript{9} University of Leuven\\\textsuperscript{10} Applied Memory Research Laboratory, Institute of Psychology, Jagiellonian University\\\textsuperscript{11} Franklin and Marshall College\\\textsuperscript{12} ETH Zürich\\\textsuperscript{13} Department of Media and Communication, City University of Hong Kong\\\textsuperscript{14} Department of Psychology, University of Rochester\\\textsuperscript{15} School of Psychology, The University of Queensland\\\textsuperscript{16} Department of Psychology and Psychotherapy, Witten/Herdecke University\\\textsuperscript{17} Programa de Investigación Asociativa (PIA) en Ciencias Cognitivas, Centro de Investigación en Ciencias Cognitivas (CICC), Facultad de Psicología, Universidad de Talca\\\textsuperscript{18} University of Oslo\\\textsuperscript{19} Massachusetts General Hospital\\\textsuperscript{20} IRCCS San Camillo Hospital\\\textsuperscript{21} Ashland University\\\textsuperscript{22} Bournemouth University\\\textsuperscript{23} University of Padova\\\textsuperscript{24} California State University East Bay\\\textsuperscript{25} University of Alabama\\\textsuperscript{26} Padova Neuroscience Center, University of Padova\\\textsuperscript{27} Children's Institute Inc.\\\textsuperscript{28} Harvard Medical School}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Analytics, Harrisburg University of Science and Technology\\\textsuperscript{2} Department of Vision Sciences, University of Leicester\\\textsuperscript{3} Department of Psychology, Princeton University\\\textsuperscript{4} Illumin Analytics\\\textsuperscript{5} Stevens Institute of Technology\\\textsuperscript{6} Department of Neuroscience, University of Padova\\\textsuperscript{7} Methodology and Statistics Unit, Institute of Psychology, Leiden University\\\textsuperscript{8} Department of Developmental and Social Psychology, University of Padova\\\textsuperscript{9} University of Leuven\\\textsuperscript{10} Applied Memory Research Laboratory, Institute of Psychology, Jagiellonian University\\\textsuperscript{11} Franklin and Marshall College\\\textsuperscript{12} ETH Zürich\\\textsuperscript{13} Department of Media and Communication, City University of Hong Kong\\\textsuperscript{14} Department of Psychology, University of Rochester\\\textsuperscript{15} School of Psychology, The University of Queensland\\\textsuperscript{16} Department of Psychology and Psychotherapy, Witten/Herdecke University\\\textsuperscript{17} Programa de Investigación Asociativa (PIA) en Ciencias Cognitivas, Centro de Investigación en Ciencias Cognitivas (CICC), Facultad de Psicología, Universidad de Talca\\\textsuperscript{18} University of Oslo\\\textsuperscript{19} Massachusetts General Hospital\\\textsuperscript{20} IRCCS San Camillo Hospital\\\textsuperscript{21} Ashland University\\\textsuperscript{22} Bournemouth University\\\textsuperscript{23} University of Padova\\\textsuperscript{24} California State University East Bay\\\textsuperscript{25} University of Alabama\\\textsuperscript{26} Children's Institute Inc.\\\textsuperscript{27} Harvard Medical School}
 
@@
 
+\hypertarget{open-practices}{%
+\section{Open Practices}\label{open-practices}}
+
+\begin{itemize}
+\tightlist
+\item
+  All data used in this manuscript and vignettes have been cited and can be found on our repository pages.
+\item
+  Manuscript repository with code and data: \url{https://osf.io/swmva/} or \url{https://github.com/SemanticPriming/stimuli-power}
+\item
+  Package repository with vignettes and data: \url{https://github.com/SemanticPriming/semanticprimeR}
+\item
+  We did not pre-register this study, as it was a simulation study. No materials were used.
+\end{itemize}
+
 \newpage

### 2024-05-28 presentations/aps_24/aps_24.qmd
updating with APS presentation

---
title: 'Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning'
author: "Erin M. Buchanan"
institute: "Harrisburg University"
format: 
  revealjs:
    theme: night
editor: source
incremental: true 
scrollable: true 
preview-links: true
code-copy: true 
highlight-style: github 
editor_options: 
  chunk_output_type: inline
---

## Power and Sample Size Planning{.smaller}

```{css}
h1.title {
font-size: 1.5em;
}

.reveal.smaller .slides h2, .reveal .slides section.smaller h2 {
  font-size: 40px;
}

.reveal div.sourceCode pre code {
  min-height: 100%;
  font-size: 25px;
}

.reveal .cell-output-stdout pre code{
  font-size: 20px;
}
```

- Sample Size Planning: New Tools and Innovations
  - Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning, Erin M. Buchanan
  - Power Analyses for Interaction Effects in Observational Studies, David A. Baranger
  - Empowering Sample Size Justification with the Superpower R Package, Aaron Caldwell

## Power and Sample Size Planning{.smaller}

```{r echo = F}
library(knitr)
library(papaja)
library(rio)
library(ggplot2)
library(dplyr)
library(broom)
library(kableExtra)
include_graphics("oprah.png")
```

## A Blender Mix{.smaller}

- Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning
- How we took a bunch of interesting ideas and mixed them together

## Sample Size Planning{.smaller}

- Sample size planning is often thought of as "point and click" 
  - G\*Power: https://www.psychologie.hhu.de/arbeitsgruppen/allgemeine-psychologie-und-arbeitspsychologie/gpower
  - https://jakewestfall.shinyapps.io/pangea/ 
  - https://pwrss.shinyapps.io/index/
  - https://designingexperiments.com/
- Sample size planning is *technically* a closed-form solution for many analysis plans 
- An incredible number of cool *R* packages for sample size planning, such as *pwr*
- So, why do we need new innovations for power?

## The Need{.smaller}

- TOPS movement + pre-registration + grants + registered reports = need for power analyses 
- Power analyses are just our *best guesses* and are likely wrong 
- Many Analyst studies show us that there is no design = analysis answer
- The smallest effect of interest may be unknown
- Some research papers do not have one specific hypothesis (i.e., dataset creations)
- Once you leave the *t*-test behind, power becomes more complicated and often based on simulation

## Our Use Case{.smaller}

- Research studies that use many items to assess the parameter of interest 
- Research studies designed to collect data on many items and share the data 
- We should be careful not to assume all items are equal ... 
- And move away from using item-level averages as parameters of interest

## Combining Toolkits{.smaller}

- Accuracy in Parameter Estimation: finding the sample size that allows for "accurately measured" parameters 
  - Determine a "sufficiently narrow" confidence interval around your parameter
  - Determine the sample size that should provide that CI
- Bootstrapping (sort of) and Simulation
  - Taking pilot data and simulating various sample sizes based on bootstrapping your sample
  - Use this technique to find the sample size for a "sufficiently narrow" CI for items

## Sequential Testing{.smaller}

- Sequential testing: examine the parameter of interest for the intended CI
  - After each participant 
  - At regular intervals during data collection
- Benefits: 
  - Maximizing the usefulness of data collection 
- Cons: 
  - Usually requires code based skill sets

## Proposed Method{.smaller}

```{r table-summary, results = 'asis'}
kable(data.frame(
  "Step Number" = c("1", "2" , "3", "4", "5", "6"),
    "Proposed Steps" = c("Use representative pilot data.",
    "Calculate standard error of each of the items in the pilot data. Using the 40%, determine the cutoff and stopping rule for the standard error of the items.",
    "Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.", 
    "Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cutoff score from Step 2.",
    "Determine the sample size at which 80%, 85%, 90%, 95% of items are below the cutoff score. Use the correction formula to adjust your proposed sample size based on pilot data size, power, and percent variability.",
    "Report all values. Designate one as the minimum sample size, the cutoff score as the stopping rule for adaptive designs, and the maximum sample size.")
  ), 
  caption = "Proposed Procedure for Powering Studies with Multiple Items",
  col.names = c("Step", "Proposed Steps"),
  #align = c("m{1cm}", rep("m{6cm}", 2)),
  #font_size = "footnotesize"
) %>% 
  kable_styling(kable_input = ., 
                font_size = 25)
```

## Package{.smaller}

- Upcoming package `semanticprimeR` as part of a larger project 
- `devtools::install_github("SemanticPriming/semanticprimeR")`
- Functions for each step of the proposed process
- Functionality for when you have example data and when you do not (i.e., simulate example multiple-item data)
- As part of the manuscript and *semanticprimeR* package, we provide 12+ examples online 
- Psycholinguistics, social psychology, COVID related, traditional cognitive psychology

## Example: Step 1 (Pilot Sample){.smaller}

```{r import-data-example, include = F, echo = F}
elpDF <- import("../../data/ELPDecisionData.zip")
conDF <- import("../../data/concreteness_trial.rda")

# grab only the overlap for this fake study
elpDF <- elpDF %>% 
  filter(Stimulus %in% conDF$Word) %>% 
  filter(Type == 1)

conDF <- conDF %>% 
  filter(Word %in% elpDF$Stimulus)

# sample the data of with 20 words in each category
average_item_score <- conDF %>% 
  filter(Rating != "N" & Rating != "n") %>% 
  mutate(Rating = as.numeric(Rating)) %>% 
  group_by(Word) %>% 
  summarize(avgrating = mean(Rating))

con_words <- c(
  average_item_score %>% 
    filter(avgrating <= 2) %>% 
    sample_n(size = 20) %>% 
    pull(Word), 
  average_item_score %>% 
    filter(avgrating >= 4) %>% 
    sample_n(size = 20) %>% 
    pull(Word)
)

elp_small <- elpDF %>% filter(Stimulus %in% con_words)
# elp_small$Stimulus <- droplevels(elp_small$Stimulus) already a character

# figure out pilot size
pilot_size_e <- round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)))

# figure out data loss
elp_small$RT[elp_small$Accuracy == "0"] <- NA
data_loss_e <- elp_small %>% 
 group_by(Stimulus, Type) %>% 
 summarize(percent_correct = sum(!is.na(RT))/n(), .groups = "keep")
elp_use <- elp_small %>% filter(!is.na(RT))

# cutoff score 
SE_e <- tapply(elp_use$RT, elp_use$Stimulus, function (x) { sd(x)/sqrt(length(x))})
cutoff_e <- quantile(SE_e, probs = seq(from = .1, to = .9, by = .1))

elp_use <- elp_use %>% 
  ungroup() %>% 
  as.data.frame()
```

- You want to run a lexical decision project measuring response latencies for concrete and abstract words 
- You can use the English Lexicon Project as pilot data + previous publications of concreteness ratings 
- In these studies, we also have to factor in data loss!
  - Combined data includes `r length(unique(elpDF$Stimulus))` real words filtered down to `r length(unique(elp_small$Stimulus))` selected stimuli
  - Average sample size per word:  `r format(round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2), nsmall = 2)` (*SD* = `r round(sd(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2)`)
  - Pilot sample size: *n* = `r pilot_size_e`

## Example: Step 2 (Calculate Cutoff){.smaller}

```{r echo = T}
library(semanticprimeR)
cutoff <- calculate_cutoff(population = elp_use, # pilot data or simulated data
  grouping_items = "Stimulus", # name of the item indicator column
  score = "RT", # name of the dependent variable column
  minimum = min(elp_use$RT), # minimum possible/found score
  maximum = max(elp_use$RT)) # maximum possible/found score
```

## Example: Step 2 (Calculate Cutoff){.smaller}

```{r echo = T}
cutoff$se_items # all standard errors of items
cutoff$sd_items # standard deviation of the standard errors
cutoff$cutoff # 40% decile score
cutoff$prop_var # proportion of possible variance 
```

## Example: Step 3 (BootSim Samples){.smaller}

```{r boot-sample, include = T, echo = T}
samples <- bootstrap_samples(start = 20, # starting sample size
  stop = 100, # stopping sample size
  increase = 5, # increase bootstrapped samples by this amount
  population = elp_use, # population or pilot data
  replace = TRUE, # bootstrap with replacement? 
  nsim = 500, # number of simulations to run
  grouping_items = "Stimulus") # item column label

head(samples[[1]])
```

```{r echo = F}
for (i in 1:length(samples)){
  samples[[i]] <- samples[[i]] %>% 
    ungroup() %>% 
    as.data.frame()
}
```

## Example: Step 4-5 (Calculate Proportion){.smaller}

```{r echo = F}
calculate_proportion <- function(samples,
                                 cutoff,
                                 grouping_items = NULL,
                                 score){

  if(is.null(samples)){ stop("You must include list of samples to examine.") }
  if(is.null(cutoff)){ stop("You must include the cutoff score for standard error.") }
  if(is.null(grouping_items)){ stop("You must include the grouping variable,
                                    which is normally the item number.") }
  if(is.null(score)){ stop("You must include the score or variable you are
                           estimating sample size for.") }

  summary_list <- list()
  score2 <- sym(score)

  # loop and calculate
  for (i in 1:length(samples)){

    summary_list[[i]] <- samples[[i]] %>%
      group_by(.data[[grouping_items]]) %>%
      summarize(se = sd(.data[[score]])/sqrt(n())) %>%
      ungroup() %>%
      summarize(percent_below = sum(se <= cutoff) / length(se),
                num_items = length(se)) %>%
      mutate(sample_size = nrow(samples[[i]]) / num_items)

  } # end loop and calculate

  # create end summary
  summary_DF <- bind_rows(summary_list) %>%
    select(-num_items) %>%
    group_by(sample_size) %>%
    summarize(percent_below = mean(percent_below))

  # return values
  return(summary_DF)

}

```

```{r calc-prop, include = T, echo = T}
proportion_summary <- calculate_proportion(samples = samples, # samples list
  cutoff = cutoff$cutoff, # cut off score 
  grouping_items = "Stimulus", # item column name
  score = "RT") # dependent variable column name

head(proportion_summary)
```

## Example: Step 6 (Apply Correction){.smaller}

```{r calc-correct, include = T, echo = T}
corrected_summary <- calculate_correction(
  proportion_summary = proportion_summary, # prop from above
  pilot_sample_size = pilot_size_e, # number of participants in the pilot data 
  proportion_variability = cutoff$prop_var, # proportion variance from cutoff scores
  power_levels = c(80, 85, 90, 95)) # what levels of power to calculate

corrected_summary
```

## Last Thoughts{.smaller}

- Use case: multiple items that intend on using item level focused analyses
- Should simulate only what is expected for a participant to do in the study 
  - Large numbers of items may bias estimates 
- Could combine with "traditional" power 
- Provides "well-measured" data --> not a specific decision for a specific sample

## Thanks{.smaller}

- Thanks for listening!
- Reproducible manuscript: https://github.com/SemanticPriming/stimuli-power 
- Package: https://github.com/SemanticPriming/semanticprimeR
- Scan me for a copy of this talk with links:

```{r include = T}
knitr::include_graphics("aps_24_qr.png")
```

## Simulation Method{.smaller}

- To evaluate our approach, we ran a simulation study: 
  - Scale size: popular cognitive scales (1-7 measurements, 0-100 percentage measurements, and 0-3000 response latency type scale data)
  - Item heterogeneity: small, medium, large
  - Skew: normal distributions versus skewed (ceiling) distributions 
  - Pilot sample size: 20 to 100 increasing in units of 10
- 1,620,000 simulations of 3 X 3 X 2 X 9 design

```{r table-sim}
#| fig-cap: " "
#| filters:
#| - parse-latex
#| 
kable(data.frame(
  Information = c("Minimum", "Maximum", "Mu", "Skewed Mu", "Sigma Mu", 
                  "Sigma", "Small Sigma Sigma", 
                  "Medium Sigma Sigma", "Large Sigma Sigma"),
  Likert = c(1, 7, 4, 6, .25, 2, .2, .4, .8),
  Percent = c(0, 100, 50, 85, 10, 25, 4, 8, 16),
  Milliseconds = c(0, 3000, 1000, 2500, 150, 400, 50, 100, 200)
), caption = "Parameter Values for Data Simulation",
escape = FALSE,
#format = "latex"
) %>% 
  kable_styling(kable_input = ., 
                font_size = 25)

```

## Simulation Results: Scale Size{.smaller}

```{r import-data}
summary_long <- bind_rows(
  import("../../simulation_study/simulation_middle/simulated_summary_data_middle.csv") %>% 
    mutate(data_type = "Normal"), 
    import("../../simulation_study/simulation_ceiling/simulated_summary_data_ceiling.csv") %>% 
    mutate(data_type = "Ceiling"))

summary_long$scale_size <- factor(summary_long$scale_size,
                                  levels = c("Likert", "Percent", "Milliseconds"))
summary_long$variability <- factor(summary_long$variability, 
                                   levels = c("Small Heterogeneity", 
                                              "Medium Heterogeneity", 
                                              "Large Heterogeneity"))

### make a decision here ### 
summary_long <- summary_long %>% filter(name != "Decile 0" & name != "Decile 10" & name != "Decile 20")
```

```{r scale-size-figure}

likert_var <- sqrt(((7-1)^2)/4)
percent_var <- sqrt(((100-0)^2)/4)
millisecond_var <- sqrt(((3000-0)^2)/4)

summary_long$percent_var <- ifelse(
  summary_long$scale_size == "Likert", summary_long$sd_item/likert_var, ifelse(
    summary_long$scale_size == "Percent", summary_long$sd_item/percent_var, 
      summary_long$sd_item/millisecond_var
  )
)

mean_var <- tapply(summary_long$percent_var, summary_long$scale_size, mean)

# New facet label names for power variable
power.labs <- c("80% of items", "85% of items",
                "90% of items", "95% of items")
names(power.labs) <- c("80", "85", "90", "95")

plot_colors <- c("#4FEAAC", "#FF826F", "#4F84FD")

ggplot(summary_long %>% 
         group_by(original_n, power, scale_size) %>% 
         summarize(sample_size_avg = mean(sample_size)),
       aes(original_n, sample_size_avg, color = scale_size)) + 
  geom_point() + 
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  geom_line(aes(original_n, original_n), color = "black") + 
  theme_linedraw(base_size = 15) + 
  xlab("\nPilot Sample Size") + 
  ylab("Projected Sample Size\n") + 
  scale_color_manual(name = "Scale Size",
                     values = c(plot_colors)) +
  facet_wrap(~power,
             labeller = labeller(power = power.labs))+
  theme_bw() +
  theme(legend.position="bottom")
```

## Simulation Results: Skew{.smaller}

```{r scale-skew-figure}

ggplot(summary_long %>% 
         group_by(original_n, power, data_type) %>% 
         summarize(sample_size_avg = mean(sample_size)),
       aes(original_n, sample_size_avg, color = data_type)) + 
  geom_point() + 
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  geom_line(aes(original_n, original_n), color = "black") + 
  theme_linedraw(base_size = 15) + 
  xlab("\nPilot Sample Size") + 
  ylab("Projected Sample Size\n") + 
  scale_color_manual(name = "Scale Skew",
                     values = c(plot_colors)) +
  facet_wrap(~power,
             labeller = labeller(power = power.labs))+
  theme_bw() +
  theme(legend.position="bottom")
```

## Simulation Results: Item Heterogeneity{.smaller}

```{r scale-hetero-figure}

ggplot(summary_long %>% 
         group_by(original_n, power, variability) %>% 
         summarize(sample_size_avg = mean(sample_size)),
       aes(original_n, sample_size_avg, color = variability)) + 
  geom_point() + 
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  geom_line(aes(original_n, original_n), color = "black") + 
  theme_linedraw() + 
  xlab("Pilot Sample Size") + 
  ylab("Projected Sample Size") + 
  scale_color_manual(name = "Item Heterogeneity",
                     values = c(plot_colors)) +
  facet_wrap(~power,
             labeller = labeller(power = power.labs))+
  theme_bw() +
  theme(legend.position="bottom")
```

## Dealing with Pilot Sample Size{.smaller}

- At some point, power usually asymptotes with increasing sample size 
- So, we need a correction:

$$ 1 - \sqrt{\frac{N_{Pilot} - min(N_{Simulation})}{N_{Pilot}}}^{log_2(N_{Pilot})}$$

## Dealing with Pilot Sample Size{.smaller}

```{r corrected-figure}

decay <- 1-sqrt((summary_long$original_n-20)/summary_long$original_n)^log2(summary_long$original_n)
summary_long$new_sample <- summary_long$sample_size*decay
ggplot(summary_long %>% 
         group_by(original_n, power, variability) %>% 
         summarize(sample_size_avg = mean(sample_size),
                   new_sample_size_avg = mean(new_sample)),
       aes(original_n, new_sample_size_avg, color = variability)) + 
  geom_line(aes(original_n, original_n), color = "black") +
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_point() + 
  geom_line() + 
  xlab("\nPilot Sample Size") + 
  ylab("Projected Sample Size\n") + 
  scale_color_manual(name = "Item Heterogeneity",
                     values = c(plot_colors)) +
  facet_wrap(~power,
             labeller = labeller(power = power.labs))+
  theme_bw() +
  theme(legend.position="bottom")
```

## Researchers Have *One* Sample{.smaller}

- Long story short: we can provide a function for researchers to use to control pilot sample size
- We also determined *which* level "sufficiently small" was probably best

```{r R2-cutoff, include = F, echo = F}
by_cutoff <- list()
R2 <- list()
for (cutoff in unique(summary_long$name)){
 by_cutoff[[cutoff]] <- lm(new_sample ~ sample_size + percent_var + original_n + power + log2(sample_size) + log2(percent_var) + log2(original_n) + log2(power), data = summary_long %>% filter(name == cutoff))
 R2[cutoff] <- summary(by_cutoff[[cutoff]])$r.squared
}
```

```{r table-decile}
table_correct <- tidy(summary(by_cutoff$`Decile 40`))
table_correct <- table_correct[c(1, 2, 4, 6, 8, 9, 3, 7, 5) , ]
table_correct$term <- c("Intercept", "Projected Sample Size", "Pilot Sample Size", 
                        "Log2 Projected Sample Size", "Log2 Pilot Sample Size", 
                        "Log2 Power", "Proportion Variability", 
                        "Log2 Proportion Variability", "Power")
table_correct$estimate <- printnum(table_correct$estimate, digits = 3)
table_correct$std.error <- printnum(table_correct$std.error, digits = 3)
table_correct$statistic <- printnum(table_correct$statistic, digits = 3)
table_correct$p.value <- printnum(table_correct$p.value, digits = 3,
                 zero = FALSE, gt1 = F)
kable(table_correct, 
     col.names = c("Term", "Estimate", "SE", "t", "p"),
     caption = "Parameters for 40% Decile Cutoff Scores") %>% 
  kable_styling(font_size = 25)
```

### 2024-11-10 manuscript_draft.Rmd
update draft with bimodal

@@
 ---
-title       : "Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli"
+title       : "Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning Accounting for Item Effects"
 shorttitle    : "AIPE BOOTSTRAPPING"
@@
 
-1)  How do pilot data influence sample size suggestions?
+1)  How does pilot data influence sample size suggestions?
 

### 2024-11-10 simulation_study/simulate_summary_bimodal.Rmd
update draft with bimodal

---
title: "Methods Power Study - Bimodal Simulations"
author: "Erin M. Buchanan"
date: "Last Knitted: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r message = F}
set.seed(895893) # 233232
library(dplyr)
library(ggplot2)
library(tidyr)
library(semanticprimeR)
```

This document examines a test of what a researcher might find if they had bimodal distributions on their data. We will simulate a pilot study by generating normal, floor, and ceiling scores. From those populations, we will select a mix of items that are "normal" and a mix of items that are bimodal (by joining together the ceiling and floor effects with a .5 probability). Finally, we will calculate the rest of the steps proposed as normal.

```{r simulate-bimodal}
# simulate population 
normal_pops <- simulate_population(mu = 4, # item means
  mu_sigma = .2, # variability in item means 
  sigma = 2, # item standard deviations
  sigma_sigma = .2, # standard deviation of the standard deviations
  number_items = 30, # number of items
  number_scores = 1000, # number of participants
  smallest_sigma = .02, #* smallest possible standard deviation
  min_score = 1, #* minimum score for truncating purposes
  max_score = 7, #* maximum score for truncating purposes
  digits = 0) #* number of digits for rounding

floor_pops <- simulate_population(mu = 2, # item means
  mu_sigma = .2, # variability in item means 
  sigma = 1, # item standard deviations
  sigma_sigma = .1, # standard deviation of the standard deviations
  number_items = 30, # number of items
  number_scores = 1000, # number of participants
  smallest_sigma = .02, #* smallest possible standard deviation
  min_score = 1, #* minimum score for truncating purposes
  max_score = 7, #* maximum score for truncating purposes
  digits = 0) #* number of digits for rounding

ceiling_pops <- simulate_population(mu = 6, # item means
  mu_sigma = .2, # variability in item means 
  sigma = 1, # item standard deviations
  sigma_sigma = .1, # standard deviation of the standard deviations
  number_items = 30, # number of items
  number_scores = 1000, # number of participants
  smallest_sigma = .02, #* smallest possible standard deviation
  min_score = 1, #* minimum score for truncating purposes
  max_score = 7, #* maximum score for truncating purposes
  digits = 0) #* number of digits for rounding

# simulate sample 
item_number <- 1:30
prop_mix <- 1 # manipulated this number 
pilot_sample_size <- 30
bimodal_items <- sample(item_number, size = round(length(item_number)*prop_mix), replace = FALSE)

researcher_sample <- 
  bind_rows(
    floor_pops %>% 
      filter(item %in% bimodal_items) %>% 
      group_by(item) %>% 
      slice_sample(n = pilot_sample_size / 2), 
    ceiling_pops %>% 
      filter(item %in% bimodal_items) %>% 
      group_by(item) %>% 
      slice_sample(n = pilot_sample_size / 2),
    normal_pops %>% 
      filter(!(item %in% bimodal_items)) %>% 
      group_by(item) %>% 
      slice_sample(n = pilot_sample_size)
  )

# researcher_sample %>% 
#   group_by(item) %>% 
#   summarize(n = n(),
#             mean = mean(score),
#             sd = sd(score), 
#             min = min(score),
#             max = max(score))
# 
# researcher_sample %>% 
#   filter(item %in% bimodal_items) %>% 
#   group_by(item) %>%
#   table()

# calculate cutoff 
cutoff <- calculate_cutoff(population = researcher_sample, # pilot data or simulated data
  grouping_items = "item", # name of the item indicator column
  score = "score", # name of the dependent variable column
  minimum = 1, # minimum possible/found score
  maximum = 7) # maximum possible/found score

# bootstrap samples
samples <- bootstrap_samples(start = 20, # starting sample size
  stop = 100, # stopping sample size
  increase = 5, # increase bootstrapped samples by this amount
  population = researcher_sample, # population or pilot data
  replace = TRUE, # bootstrap with replacement? 
  nsim = 1000, # number of simulations to run
  grouping_items = "item") # item column label

# calculate proportions
proportion_summary <- calculate_proportion(samples = samples, # samples list
  cutoff = cutoff$cutoff, # cut off score 
  grouping_items = "item", # item column name
  score = "score") # dependent variable column name

# calculate correction
corrected_summary <- calculate_correction(
  proportion_summary = proportion_summary, # prop from above
  pilot_sample_size = 30, # number of participants in the pilot data 
  proportion_variability = cutoff$prop_var, # proportion variance from cutoff scores
  power_levels = c(80, 85, 90, 95)) # what levels of power to calculate

rio::export(corrected_summary, paste0("simulation_bimodal/summary_", 
                                 prop_mix, ".csv"), 
                                 row.names = F)
```

```{r}
list_files <- list.files("simulation_bimodal/",
                         pattern = "*.csv",
                         full.names = TRUE)

import_files <- lapply(list_files, rio::import)

full_data <- bind_rows(import_files) %>% 
  mutate(size = rep(c(.1, .2, .3, .4, .5, .6, .7, .8, .9, 0, 1), each = 4))

rio::export(full_data, "simulation_bimodal/summary_bimodal.csv", row.names = F)
```

### 2024-11-11 manuscript_draft.Rmd
updating manuscript

@@
 title       : "Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning Accounting for Item Effects"
-shorttitle    : "AIPE BOOTSTRAPPING"
+shorttitle    : "AIPE MONTE CARLO"
 author:
@@
 The replication "crisis" and credibility revolution have shown that published studies in psychology are underpowered [@opensciencecollaboration2015; @vazire2018; @korbmacher2023].
+Potential reasons for underpowered studies include questionable research practices [@john2012; but see @fiedler2016], weak psychological theories [@proulx2021; @szollosi2021], testing multiple hypotheses [@maxwell2004], and poor intuitions about power [@bakker2016].
+
 Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020; @d.chambers2014; @nosek2014], and details of a power analyses or limitations on resources are often used to provide justification for the pre-registered sample quota [@pownall2023; @vandenakker2023; @akker2023].
@@
 Additionally, many tools assume data normality, place an overemphasis on statistical significance, and may rely on simplified assumptions that do not reflect the actual data.
-When computing sample size estimates, it is important to remember that the effects sizes are estimates, not exact calculations guaranteed to produce a specific result [@batterham2005].
+Further, the social sciences often ignores robust statistical methods as an option for analysis [@field2017; @erceg-hurn2008], and the implementation of these analyses in power software is somewhat sporadic.
+Finally, when computing sample size estimates, it is important to remember that the effects sizes are estimates, not exact calculations guaranteed to produce a specific result [@batterham2005].
 For example, it is hard to accurately estimate all parameters from a study, and if any were incorrect, then the sample size estimate tied to that specific level of power may be incorrect [@albers2018].
@@
 In light of these analytical (or lack thereof) concerns, we propose a new method to determine a sample size in cases where a more traditional power analysis might be less appropriate or even impossible.
-This approach combines accuracy in parameter estimation [AIPE, @kelley2007; @maxwell2008] and bootstrapped simulation on pilot data [@rousselet].
+This approach combines accuracy in parameter estimation [AIPE, @kelley2007; @maxwell2008] and data-driven Monte-Carlo simulation on pilot data [@rousselet].
 This method accounts for a potential lack of hypothesis test (or simply no good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices.
@@
 
-AIPE shifts the focus away from finding a significant *p*-value to finding a parameter that is "accurately measured".
+AIPE shifts the focus away from finding a significant *p*-value to finding a parameter that is accurately measured.
 For example, researchers may wish to detect a specific mean in a study, *M* = .35.
-They could then use AIPE to estimate the sample size needed to find a "sufficiently narrow" confidence interval around that mean.
+They could then use AIPE to estimate the sample size needed to find a sufficiently narrow confidence interval around that mean.
 Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and/or confidence intervals.
@@
 
-## Bootstrapping and Simulation
-
-One form of data simulation is bootstrapping, which involves using data obtained to simulate similar datasets by drawing from the original data with replacement [@efron2000; @rousselet].
-Bootstrapping allows one to calculate parameter estimates, confidence intervals, and to simulate the potential population distribution, shape, and bias.
+## Monte Carlo Simulation
+
+One form of data simulation is data-driven Monte-Carlo simulation, which involves using data obtained to simulate similar datasets by drawing from the original data with replacement [@efron2000; @rousselet].
+This type of simulation allows one to calculate parameter estimates, confidence intervals, and to simulate the potential population distribution, shape, and bias.
 Simulation is often paired with re-creating a data set with a similar structure for testing analyses and hypotheses based on proposed effect sizes or suggested population means.
@@
 
-Pilot or simulated data would be used to estimate the variability within items and select a "sufficiently narrow" window for overall item confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
+Pilot or simulated data would be used to estimate the variability within items and select a sufficiently narrow window for overall item confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
 The advantage to this method over simple power estimation from pilot effect sizes is the multiple simulations to average out potential variability, as well as a shift away from traditional NHST to parameter estimation.
-Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein as many items as required meet the pre-specified well-measured criterion.
+Simulation would then be used to determine how many participants may be necessary to achieve a dataset wherein as many items as required meet the pre-specified well-measured criterion.
 
@@
 The maximum stopping rule could also be defined by resources (e.g., two semesters data collection), but nevertheless should be included.
-Therefore, we propose a method that leverages the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates to measure an overall phenomena.
+Therefore, we propose a method that leverages the ideas behind AIPE, paired with simulation and Monte Carlo simulation, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates to measure an overall phenomena.
 
@@
 2)  For each item in the pilot data, calculate the standard error (SE).
-Select a cutoff SE that defines when items are considered "accurately measured".
+Select a cutoff SE that defines when items are considered accurately measured.
 The simulations described in the Data Simulation section will explore what criterion should be used to determine the cutoff SE from the pilot data.
 
-### Bootstrap Samples
+### Monte Carlo Samples
 
@@
 4)  Use the simulated SEs to determine the percentage of items that meet the cutoff score determined in Step 2.
-Each sample size from Step 3 will have multiple bootstrapped simulations, and therefore, create an average percentage score for each sample size for Step 5.
+Each sample size from Step 3 will have multiple simulations, and therefore, create an average percentage score for each sample size for Step 5.
 
@@
     "Calculate standard error of each of the items in the pilot data. Determine the appropriate SE for the stopping rule.",
-    "Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.", 
-    "Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cutoff score from Step 2.",
+    "Create simulated samples of your pilot data starting with at least 20 participants up to a maximum number of participants.", 
+    "Calculate the standard error of each of the items in the simulated data. From these scores, calculate the percent of items below the cutoff score from Step 2.",
     "Determine the sample size at which 80%, 85%, 90%, 95% of items are below the cutoff score.",
@@
     "Calculate standard error of each of the items in the pilot data. Using the 40%, determine the cutoff and stopping rule for the standard error of the items.",
-    "Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.", 
-    "Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cutoff score from Step 2.",
+    "Create simulated samples of your pilot data starting with at least 20 participants up to a maximum number of participants.", 
+    "Calculate the standard error of each of the items in the simulated data. From these scores, calculate the percent of items below the cutoff score from Step 2.",
     "Determine the sample size at which 80%, 85%, 90%, 95% of items are below the cutoff score.
@@
 *Step 3-5*.
-The pilot data was then bootstrapped with replacement creating samples of 20 to 300 participants per item increasing in units of 5, for concreteness ratings and lexical decision latencies separately (Step 3).
+The pilot data was then simulated with replacement creating samples of 20 to 300 participants per item increasing in units of 5, for concreteness ratings and lexical decision latencies separately (Step 3).
 Each of these 57 sample sizes was then repeated 500 times.
-The SE of each item was calculated for the bootstrapped samples separately for concreteness ratings and lexical decision times (Step 4), and the average percentage of items for each sample size (averaging across the 500 simulations) below each potential cutoff was gathered for each (Step 5).
+The SE of each item was calculated for the simulated samples separately for concreteness ratings and lexical decision times (Step 4), and the average percentage of items for each sample size (averaging across the 500 simulations) below each potential cutoff was gathered for each (Step 5).
 The smallest sample size with at least 80%, 85%, 90%, and 95% of items below the cutoff are reported in Table \@ref(tab:table-example) for each task (Step 5).
@@
 results_table <- matrix(NA, nrow = 9, ncol = 11)
-results_table[ , 1] <- c("Decile 10", "Decile 20", "Decile 30", 
-                         "Decile 40", "Decile 50", "Decile 60", 
-                         "Decile 70", "Decile 80", "Decile 90")
+results_table[ , 1] <- c("Decile 1", "Decile 2", "Decile 3", 
+                         "Decile 4", "Decile 5", "Decile 6", 
+                         "Decile 7", "Decile 8", "Decile 9")
 results_table[1 , 2:11] <- unlist(c(cutoff_c[1], concrete_cuts[1 , ], 
@@
 SE columns represent the standard error value cutoff for each decile, while 80/85/90/95 percent columns represent the sample size needed to have that percent of items below the SE cutoff.
-For example, 150 participants are required to ensure at least 95% of concreteness items SE are below the 10 percent decile SE cutoff, and 345 participants are necessary for the lexical decision SE to be below its 10 percent decile cutoff.")
+For example, 150 participants are required to ensure at least 95% of concreteness items SE are below the 1st decile SE cutoff, and 345 participants are necessary for the lexical decision SE to be below its 1st decile cutoff.")
 ```
@@
 
-Next, we calculated the percentage of items that fell below the cutoff score, and therefore, would be considered "well-measured" for each decile by sample (Step 4).
+Next, we calculated the percentage of items that fell below the cutoff score, and therefore, would be considered "well-measured for each decile by sample (Step 4).
 From these data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90% and 95% of the items fall below the cutoff criterion (Step 5).
-These values were chosen as popular measures of "power" in which one could determine the minimum suggested sample size (potentially 80% of the items) and the maximum suggested sample size (selected from a higher percentage, such as 90% or 95%).
+These values were chosen as popular, yet arbitrary, measures of "power" in which one could determine the minimum suggested sample size (potentially 80% of the items) and the maximum suggested sample size (selected from a higher percentage, such as 90% or 95%).
 
@@
 Last, we examined the question of an appropriate SE decile.
-First, the 0%, 10%, and 20% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale size, and heterogeneity.
+First, the 0, 1st, and 2nd deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale size, and heterogeneity.
 If we examine the $R^2$ values for each decile of our regression equation separately, we find that the values are all $R^2$ \> .99 with very little differences between them.
-Figures \@ref(fig:decile40-figure) and \@ref(fig:decile50-figure) illustrate the corrected scores for simulations at the 40% and 50% decile recommended cutoff for item standard errors.
-For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 40% decile range, especially for scales with larger potential variance.
-Therefore, we would suggest the 40% decile to overpower each item for Step 2.
-
-```{r decile40-figure, fig.cap = "Comparison of the cutoffs for 40% deciles across heterogeneity (columns), powering of items (rows), and scale size (color).", fig.height=8, fig.width=8}
+Figures \@ref(fig:decile40-figure) and \@ref(fig:decile50-figure) illustrate the corrected scores for simulations at the 4th and 5th decile recommended cutoff for item standard errors.
+For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 4th decile range, especially for scales with larger potential variance.
+Therefore, we would suggest the 4th decile to overpower each item for Step 2.
+
+```{r decile40-figure, fig.cap = "Comparison of the cutoffs for 4th deciles across heterogeneity (columns), powering of items (rows), and scale size (color).", fig.height=8, fig.width=8}
 summary_long %>% 
@@
              labeller = labeller(power = power.labs)) +
-  ggtitle("Results for 40% deciles") +
+  ggtitle("Results for 4th deciles") +
   xlab("\nPilot Sample Size") + 
@@
 
-```{r decile50-figure, fig.cap = "Comparison of the cutoffs for 50% deciles across heterogeneity (columns), powering of items (rows), and scale size (color).", fig.height=8, fig.width=8}
+```{r decile50-figure, fig.cap = "Comparison of the cutoffs for 5th decile across heterogeneity (columns), powering of items (rows), and scale size (color).", fig.height=8, fig.width=8}
 summary_long %>% 
@@
              labeller = labeller(power = power.labs)) + 
-  ggtitle("Results for 50% deciles") +
+  ggtitle("Results for 5th decile") +
   xlab("\nPilot Sample Size") + 
@@
 
-The final formula for 40% decile correction is provided in Table \@ref(tab:table-decile). Proportion of variance can be calculated with the following:
+The final formula for 4th decile correction is provided in Table \@ref(tab:table-decile). Proportion of variance can be calculated with the following:
 
@@
      col.names = c("Term", "Estimate", "$SE$", "$t$", "$p$"),
-     caption = "Parameters for 40% Decile Cutoff Scores")
+     caption = "Parameters for 4th decile Cutoff Scores")
 ```
@@
 The main change occurs in Step 2 with a designated cutoff decile, and Step 5 with a correction score.
-Using the data from the 40% decile in Table \@ref(tab:table-example), we can determine that the stopping rule SE for concreteness ratings would be `r apa_num(cutoff_c[4])`, and the stopping rule SE for lexical decision times would be `r apa_num(cutoff_e[4])`.
+Using the data from the 4th decile in Table \@ref(tab:table-example), we can determine that the stopping rule SE for concreteness ratings would be `r apa_num(cutoff_c[4])`, and the stopping rule SE for lexical decision times would be `r apa_num(cutoff_e[4])`.
 For Step 5, we apply our correction formula separately for each one, as they have different variability scores, and these scores are shown in Table \@ref(tab:table-scores-updated).
@@
 *Step 2*.
-In step 2, we can use `calculate_cutoff()` to calculate the standard error of the items, the standard deviation of the standard errors and the corresponding proportion of variance possible, and the 40% decile cutoff score.
+In step 2, we can use `calculate_cutoff()` to calculate the standard error of the items, the standard deviation of the standard errors and the corresponding proportion of variance possible, and the 4th decile cutoff score.
 The `pops` dataframe can be used in this function, which has columns named `item` for the item labels (i.e., 1, 2, 3, 4 or characters can be used), and `score` for the dependent variable.
@@
 cutoff$sd_items # standard deviation of the standard errors
-cutoff$cutoff # 40% decile score
+cutoff$cutoff # 4th decile score
 cutoff$prop_var # proportion of possible variance 
@@
 *Step 3*.
-The `bootstrap_samples()` function creates bootstrapped samples from the pilot or simulated population data to estimate the number of participants needed for item standard error to be below the cutoff calculated in Step 2.
+The `simulate_samples()` function creates simulated samples from the pilot or simulated population data to estimate the number of participants needed for item standard error to be below the cutoff calculated in Step 2.
 This function returns a list of samples with sizes that start at the `start` size, increase by `increase`, and end with the `stop` sample size.
 The population or pilot data will be included in `population`, and the item column indicator should be included in `grouping_items`.
-The `nsim` argument determines the number of bootstrapped simulations to run.
+The `nsim` argument determines the number of simulations to run.
 
 ```{r boot-sample, include = T, echo = T}
-samples <- bootstrap_samples(start = 20, # starting sample size
+samples <- simulate_samples(start = 20, # starting sample size
   stop = 100, # stopping sample size
-  increase = 5, # increase bootstrapped samples by this amount
+  increase = 5, # increase simulated samples by this amount
   population = pops, # population or pilot data
-  replace = TRUE, # bootstrap with replacement? 
+  replace = TRUE, # simulate with replacement? 
   nsim = 500, # number of simulations to run
@@
 *Step 4 and 5*.
-The proportion of bootstrapped items across sample sizes below the cutoff score can then be calculated using `calculate_proportion()`.
+The proportion of simulated items across sample sizes below the cutoff score can then be calculated using `calculate_proportion()`.
 This function returns a dataframe including each sample size with the proportion of items below that cutoff to use in the next function.
@@
 Examples include psycholinguistics [@dedeyne2008; @heyman2014; @montefinese2022], social psychology data [@ulloa2014; @peterson2022; @grahe2022], COVID related data [@montefinese2021], and cognitive psychology [@röer2013; @10.7554/eLife.71601; @barzykowski2019].
-These can be found on the package tutorial page: https://semanticpriming.github.io/semanticprimeR/.
+These can be found on the package tutorial page: <https://semanticpriming.github.io/semanticprimeR/>.
 
@@
 
-We proposed a method combining AIPE, bootstrapping, and simulation to estimate a minimum and maximum sample size and to define a rule for stopping data collection based on narrow confidence intervals on a parameter of interest.
+We proposed a method combining AIPE and Monte Carlo simulation to estimate a minimum and maximum sample size and to define a rule for stopping data collection based on narrow confidence intervals on a parameter of interest.
 In addition, we also demonstrated its practical applications using real-world data.
@@
 For example, the English Lexicon Project includes thousands of items, and if we were to simulate for all of those, our results would likely suggest needing thousands of participants for most items to reach the criterion.
-Additionally, as the number of items increases, you may also see very small estimates for sample size due to the correction factor (as with large numbers of items, you could find many items with standard errors below the 40% decile).
+Additionally, as the number of items increases, you may also see very small estimates for sample size due to the correction factor (as with large numbers of items, you could find many items with standard errors below the 4th decile).
 Therefore, it would be beneficial to consider only simulating what a participant would reasonably complete in a study.
@@
 These proposed methods and traditional power analysis could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
-
-Researchers should consider this hybrid approach for AIPE, bootstrapping, and simulation as a powerful tool for hypothesis testing and parameter estimation.
+We would advise caution when oversampling (i.e., simulation from data with larger sample sizes than the pilot data) for null-hypothesis power estimates, as these may be upwardly biased in small samples creating Type-I error [@burns].
+
+Researchers should consider this hybrid approach for AIPE and simulation as a powerful tool for hypothesis testing and parameter estimation.
 This procedure holds benefits for various research studies, specifically replication studies, that usually prioritize subject sample size but rarely item sample size, in spite of the fact that item sample sizes can contribute to power in multilevel models [@brysbaert2018].
@@
 -   All data used in this manuscript and vignettes have been cited and can be found on our repository pages.
-- Manuscript repository with code and data: https://osf.io/swmva/ or https://github.com/SemanticPriming/stimuli-power 
-- Package repository with vignettes and data: https://github.com/SemanticPriming/semanticprimeR
+-   Manuscript repository with code and data: <https://osf.io/swmva/> or <https://github.com/SemanticPriming/stimuli-power>
+-   Package repository with vignettes and data: <https://github.com/SemanticPriming/semanticprimeR>
 -   We did not pre-register this study, as it was a simulation study. No materials were used.
@@
 
-```{=tex}
 \begingroup
@@
 \setlength{\leftskip}{0.5in}
-```
+
 ::: {#refs custom-style="Bibliography"}
@@
 
-```{=tex}
 \endgroup
-```

### 2024-11-11 manuscript_draft.Rmd
update manuscript

@@
 library(latex2exp)
+library(ggridges)
+library(viridis)
 ```
@@
       gfile = "power2stim_V2_trackdown",
-      gpath = "Studies/PSA007 SPAM-L/Manuscripts/Power",
-      shared_drive = "Psychological Science Accelerator")
+      gpath = "trackdown")
 # to google
@@
       gfile = "power2stim_V2_trackdown",
-      gpath = "Studies/PSA007 SPAM-L/Manuscripts/Power",
-      shared_drive = "Psychological Science Accelerator",
+      gpath = "trackdown",
       rich_text = TRUE,
@@
 While statistics courses often suggest that a specific research design leads to a specific statistical test, meta-science work has shown that given the same data and hypothesis, researchers can come up with multiple ways to analyze the data [@silberzahn2018many; @coretta].
-Therefore, a single power analysis only corresponds to the specific analysis that the researcher expects to implement.
+Therefore, a single power analysis only corresponds to the specific analysis that the researcher expects to implement, and usually, the final expected data ignoring any processing pipeline that a researcher may implement.
 Analyses may evolve during the research project or be subject to secondary analysis; thus, power and sample size estimation based on one analysis is potentially less useful than previously imagined.
@@
 AIPE shifts the focus away from finding a significant *p*-value to finding a parameter that is accurately measured.
+As discussed in Kelley and Maxwell [@kelley2003], one can estimate the sample size necessary to obtain precision in estimation of population parameters.
+Precision is defined as a researcher defined, sufficiently narrow, confidence interval.
 For example, researchers may wish to detect a specific mean in a study, *M* = .35.
 They could then use AIPE to estimate the sample size needed to find a sufficiently narrow confidence interval around that mean.
-Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and/or confidence intervals.
-Therefore, they could decide that their 95% confidence interval should be approximately between .20 and .50, and sufficiently narrow could be defined as a width of .30 or .15 on each side.
-While confidence intervals are related to null hypothesis significance testing (i.e., 95% confidence intervals that do not include zero would indicate a significant difference from zero at $\alpha$ \< .05), AIPE procedures suggest how we can define a sample size with a given width of confidence interval, regardless of whether it includes zero.
+Therefore, they could decide that sufficiently narrow could be defined as a confidence interval width of .30 or .15 on each side of the mean.
+They would then estimate the number of participants needed to find that level of precision.
+Confidence intervals are often misinterpreted [@hoekstra2014; but see @miller2016], and AIPE procedures are not designed to specify sample size for a hypothesis-driven decision (i.e., the confidence interval does not include a specific value of comparison).
+Instead, AIPE focuses on estimating the sample size necessary to ensure precise population parameters.
+Note that any particular confidence interval is not of interest within our procedure, but rather how to define the sufficiently narrow window a researcher should use for sample size estimation when designing studies with multiple items.
 
@@
 Researchers could then use sequential testing to estimate their parameter of interest after each participant's data or at regular intervals during data collection to determine whether they have achieved their expected width of the confidence interval around that parameter.
-One would set a minimum sample size (e.g., based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the confidence interval is sufficiently narrow, as defined above).
-Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (e.g., they expect 90% of items to meet their stopping rule with 100 participants based on simulation).
+One would set a minimum sample size based on our procedure steps below to ensure an appropriate minimum for precise estimates.
+Then a researcher could use the standard error as a stopping rule (i.e., stop data collection when the standard error meets the expected sufficiently narrow width, as defined above).
+A stricter criterion could be defined for a stopping rule as a maximum sample size.
 By defining each of these components, researchers could ensure a feasible minimum sample size, a way to stop data collection when goals have been met, and a maximum sample size rule to ensure an actual end to data collection.
 The maximum stopping rule could also be defined by resources (e.g., two semesters data collection), but nevertheless should be included.
-Therefore, we propose a method that leverages the ideas behind AIPE, paired with simulation and Monte Carlo simulation, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates to measure an overall phenomena.
+The advantage of sequential testing lies within research studies that use a random selection of items across participants (i.e., participants do not see all items).
+Across participants, data can be shifted to items with more uncertainty to increase precision of estimates for those items.
+Sequential testing is not a necessary component to our proposed procedure but we outline how to use the sample size estimates below to set minimum, maximum, and stopping rules.
+If researchers will show all items to participants, they could simply select one of the proposed estimates for their minimum required sample size.
+Therefore, we propose a method that leverages the ideas behind AIPE, paired with simulation and Monte Carlo simulation, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates.
 
@@
 
-For this example, imagine a researcher who wants to determine the differences in response latencies for abstract and concrete words.
+For this example, imagine a researcher who wants to determine the differences in response latencies for abstract and concrete words [see the following for thoughts on analyzing this type of data: @bürkner2019; @liddell2018; @taylor2022].
 They will select *n* = 40 words from the rating data from @brysbaert2014 that are split evenly into abstract and concrete ends of the rating scale.
@@
 
-We first filtered the ELP data to the same real words as the concreteness subset selected above, and this data includes `r length(unique(elpDF$Stimulus))` real words.
+We first selected the same real words in the ELP data as the concreteness subset selected above, and this data includes `r length(unique(elpDF$Stimulus))` real words.
 The average pilot sample size for this random sample was `r format(round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2), nsmall = 2)` (*SD* = `r round(sd(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2)`), and *n* = `r pilot_size_e` will be our pilot size for the lexical decision task.
@@
 On each of the simulated datasets, we applied the above proposed method to determine how the procedure would perform and evaluated what criteria should be used for cutoff selection (Step 2).
-This procedure was performed on distributions in the middle of the scale (i.e., normal) and at the ceiling of the scale (i.e., skewed).
+This procedure was performed on distributions in the middle of the scale (i.e., symmetric) and at the ceiling of the scale (i.e., skewed).
 With this simulation, we will answer several questions:
@@
 Each of these samples would correspond to Step 1 of the proposed method where a researcher would use pilot data to start their estimation.
-Therefore, the simulations included 3 scales X 3 heterogeneity values X 2 normal/skewed distributions X 9 pilot sample sizes representing a potential Step 1 of our procedure.
+Therefore, the simulations included 3 scales X 3 heterogeneity values X 2 symmetric/skewed distributions X 9 pilot sample sizes representing a potential Step 1 of our procedure.
 
@@
   import("simulation_study/simulation_middle/simulated_summary_data_middle.csv") %>% 
-    mutate(data_type = "Normal"), 
+    mutate(data_type = "Symmetric"), 
     import("simulation_study/simulation_ceiling/simulated_summary_data_ceiling.csv") %>% 
@@
 ```{r scale-skew-figure, echo = F, fig.cap="Simulated pilot sample size and final projected sample size to achieve 80%, 85%, 90%, and 95% of items below threshold.
-In comparison to Figure 1, this figure shows projected sample size for ceiling versus normal distributions on each scale.
+In comparison to Figure 1, this figure shows projected sample size for ceiling versus symmetric distributions on each scale.
 All other variables are averaged together, and black dots represent original sample size for reference.", fig.height=6, fig.width=8 }
@@
 
-Figure \@ref(fig:scale-skew-figure) displays that ceiling distributions, averaged over all other variables, show slightly higher estimates than normal distributions.
+Figure \@ref(fig:scale-skew-figure) displays that ceiling distributions, averaged over all other variables, show slightly higher estimates than symmetric distributions.
 This result is consistent across scale type and heterogeneity, as results indicated that they are often the same or slightly higher for ceiling distributions.
@@
 temp_table <- broom::tidy(model.samplesize)
-temp_table$term <- c("Intercept", "Pilot Sample Size", "Scale: Likert v Percent", "Scale: Likert v Milllisecond", "Proportion Variability", "Data: Ceiling v Normal")
+temp_table$term <- c("Intercept", "Pilot Sample Size", "Scale: Likert v Percent", "Scale: Likert v Milllisecond", "Proportion Variability", "Data: Ceiling v Symmetric")
 temp_table$p.value <- apa_p(temp_table$p.value)
@@
 
-Using a regression model, we predicted proposed sample size using pilot sample size, scale size, proportion variability (i.e., heterogeneity), and data type (normal, ceiling).
+Using a regression model, we predicted proposed sample size using pilot sample size, scale size, proportion variability (i.e., heterogeneity), and data type (symmetric, ceiling).
 As shown in Table \@ref(tab:table-predict-largest), the largest influence on proposed sample size is the original pilot sample size, followed by proportion of variance/heterogeneity, and then data and scale sizes.
@@
 Given the non-linear nature of the correction, we added each variable and its non-linear `log2` transform to the regression equation, as this function was used to create the correction.
-The intercept only model was used as a starting point (i.e., `corrected sample ~ 1`), and then all eight variables (each variable and their `log2` transform) were entered into a forward stepwise regression to capture the corrected scores with the most predictive values.
-Each variable was entered one at a time using the `step` function from the *stats* library in *R* [@R-base].
+The intercept only model was used as a starting point (i.e., `corrected sample ~ 1`), and then all eight variables (each variable and their `log2` transform) were entered into the regression equation.
 
@@
 
-model.summary <- tidy(forward)
-model.summary$AIC <- forward$anova$AIC
+model.summary <- tidy(user_model3)
+# model.summary$AIC <- forward$anova$AIC
 ```
@@
 ```{r test-assume, include = F, echo = F}
-resid_model <- scale(residuals(forward))
-fit_model <- scale(fitted.values(forward))
+resid_model <- scale(residuals(user_model3))
+fit_model <- scale(fitted.values(user_model3))
 hist(resid_model)
@@
 
-As shown in Table \@ref(tab:table-correction), all variables were included in the final equation, each contributing a significant change to the previous model, as defined by $\Delta$AIC \> 2 points change between each step of the model.
+As shown in Table \@ref(tab:table-correction), all variables were significant predictors of the new sample size.
 Proposed sample size and original sample size were the largest predictors -- unsurprising given the correction formula employed -- followed by the percent "power" level and proportion of variance.
@@
                  zero = FALSE, gt1 = F)
-table_correct$AIC <- printnum(table_correct$AIC, digits = 2)
 apa_table(table_correct, 
-     col.names = c("Term", "Estimate", "$SE$", "$t$", "$p$", "AIC"),
+     col.names = c("Term", "Estimate", "$SE$", "$t$", "$p$"),
      caption = "Parameters for All Decile Cutoff Scores")
@@
 
-We have developed functions to implement the suggested procedure as part of an upcoming package `semanticprimeR`.
+We have developed functions to implement the suggested procedure as part of a semantic priming focused package `semanticprimeR`.
 You can install the package from GitHub using: `devtools::install_github("SemanticPriming/semanticprimeR")`.
@@
 
+## Special Consideration: Bimodal Data
+
+The above simulations assume symmetric population distributions with small, medium, or large heterogeneity or a skewed distribution with the same heterogeneity.
+However, as shown in @pollock2018, sample rating data that appears to have a mean distribution may represent the average two separate distributions (i.e., a bimodal distribution).
+To examine our procedure with bimodal distributions, we estimated thirty items for three distributions: symmetric, ceiling, and floor distributions for Likert data with small heterogeneity.
+Bimodal data was created by selecting half of the ceiling distribution and floor distribution to combine together.
+The number of bimodal items was varied from 0 percent to 100 percent increasing by 10 percent increments for simulated samples.
+Figure \@\ref(fig:bimodal-figure) shows an example of 10 of the items within a simulation that estimated that half of the Likert items would be bimodal in nature.
+The researcher procedure described above was then carried out using *semanticprimeR*.
+
+```{r import-bimodal, echo = F}
+bi_DF <- import("simulation_study/simulation_bimodal/summary_bimodal.csv") %>% 
+  mutate(new_percent = rep(c(80, 85, 90, 95), times = 11)) %>% 
+  mutate(size = factor(size*100))
+
+# simulate population 
+normal_pops <-semanticprimeR::simulate_population(mu = 4, # item means
+  mu_sigma = .2, # variability in item means 
+  sigma = 2, # item standard deviations
+  sigma_sigma = .2, # standard deviation of the standard deviations
+  number_items = 30, # number of items
+  number_scores = 1000, # number of participants
+  smallest_sigma = .02, #* smallest possible standard deviation
+  min_score = 1, #* minimum score for truncating purposes
+  max_score = 7, #* maximum score for truncating purposes
+  digits = 0) #* number of digits for rounding
+
+floor_pops <- semanticprimeR::simulate_population(mu = 2, # item means
+  mu_sigma = .2, # variability in item means 
+  sigma = 1, # item standard deviations
+  sigma_sigma = .1, # standard deviation of the standard deviations
+  number_items = 30, # number of items
+  number_scores = 1000, # number of participants
+  smallest_sigma = .02, #* smallest possible standard deviation
+  min_score = 1, #* minimum score for truncating purposes
+  max_score = 7, #* maximum score for truncating purposes
+  digits = 0) #* number of digits for rounding
+
+ceiling_pops <- semanticprimeR::simulate_population(mu = 6, # item means
+  mu_sigma = .2, # variability in item means 
+  sigma = 1, # item standard deviations
+  sigma_sigma = .1, # standard deviation of the standard deviations
+  number_items = 30, # number of items
+  number_scores = 1000, # number of participants
+  smallest_sigma = .02, #* smallest possible standard deviation
+  min_score = 1, #* minimum score for truncating purposes
+  max_score = 7, #* maximum score for truncating purposes
+  digits = 0) #* number of digits for rounding
+
+# simulate sample 
+item_number <- 1:30
+prop_mix <- .5 # manipulated this number 
+pilot_sample_size <- 30
+bimodal_items <- sample(item_number, size = round(length(item_number)*prop_mix), replace = FALSE)
+
+researcher_sample <- 
+  bind_rows(
+    floor_pops %>% 
+      filter(item %in% bimodal_items) %>% 
+      group_by(item) %>% 
+      slice_sample(n = pilot_sample_size / 2), 
+    ceiling_pops %>% 
+      filter(item %in% bimodal_items) %>% 
+      group_by(item) %>% 
+      slice_sample(n = pilot_sample_size / 2),
+    normal_pops %>% 
+      filter(!(item %in% bimodal_items)) %>% 
+      group_by(item) %>% 
+      slice_sample(n = pilot_sample_size)
+  )
+```
+
+```{r bimodal-figure, fig.cap = "Simulated example of Likert data with half of the items as bimodal distributions.", fig.height=8, fig.width=8}
+ggplot(researcher_sample %>% 
+         filter(item < 11), aes(x = score, y = factor(item), fill = factor(item))) +
+  geom_density_ridges(alpha = 0.7, scale = 1) +
+  labs(x = "Score", y = "Item") +
+  theme_minimal() +
+  theme(legend.position = "none") + 
+  coord_cartesian(xlim = c(1,7))
+```
+
+Figure \@\ref(fig:bimodal-figure) portrays the results of the suggested procedure.
+The sample with 0 bimodal items demonstrates the same values as above.
+The proposed sample size increases from 10 to 40% because of the heterogeneity in SE across items (i.e., with 10% of items with a larger SE due to their bimodal nature, the required sample size increases).
+From 50% to 100% the sample size decreases because the variance of the SEs across items decreases (i.e., at 100 percent, they are all larger rather than a mix).
+In theory, these results map onto the conceptual framework - if all items are truly bimodal, we have precisely measured the mean of the bimodal distribution, but this result is likely not the intended result.
+If researchers expect these distributions (or have representative pilot data), they could examine the data for bimodal items.
+These items could be estimated separately (i.e., floor effects item 1, ceiling effects item 1) to ensure that both populations of answers are represented in the data.
+
+```{r bimodal-results, fig.cap = "Estimated sample sizes for distributions that vary in the number of bimodal items.
+Note that percent below is treated as a continuous factor", fig.height=8, fig.width=8}
+ggplot(bi_DF, aes(x = percent_below,
+                  y = corrected_sample_size, 
+                  color = factor(size),
+                  shape = factor(size))) +
+  geom_point(size = 3) +
+  geom_line() +
+  theme_classic() +
+  xlab("Percent of Items Below Threshold") + 
+  ylab("Corrected Sample Size") +
+  scale_color_viridis_d(option = "plasma") +  # Use the 'plasma' viridis option
+  scale_shape_manual(values = c(16, 17, 18, 19, 15, 3, 7, 8, 9, 13, 10)) +
+  labs(color = "Size", shape = "Size")
+```
+
 ## Vignettes
@@
 
-Second, these methods do not ensure the normal interpretation of power, where you know that you would find a specific effect for a specific test, $\alpha$, and so on.
+Second, these methods do not ensure the normal interpretation of power, focusing on finding a specific effect for a specific test, $\alpha$, and so on.
 As discussed in the introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis; many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
@@
 Researchers should consider this hybrid approach for AIPE and simulation as a powerful tool for hypothesis testing and parameter estimation.
-This procedure holds benefits for various research studies, specifically replication studies, that usually prioritize subject sample size but rarely item sample size, in spite of the fact that item sample sizes can contribute to power in multilevel models [@brysbaert2018].
+This procedure holds benefits for various research studies, specifically replication studies, that usually prioritize subject sample size but rarely item sample size, in spite of the fact that item sample sizes can contribute to power in multilevel models [@brysbaert2018; however, see @rouder2018 for a discussion of the item-sample size trade off].
 Replicated effects, accumulated through multiple studies and accurate measurement, contribute to robust meta-analyses, enhancing our understanding of the genuine nature of observed effects.
-This article helps to achieve this goal by encouraging researchers to conduct studies where the power analysis is not based on the size of the effect but on adequate sampling of the stimuli.
+This article helps to achieve this goal by encouraging researchers to conduct studies where the power analysis is not based on the size of the effect but on precise measurement of the stimuli.
 We argue that this article can be the initial step to apply AIPE in a manner that can allow researchers to use item information to provide a more accurate and statistically reliable measure of the effect we aimed to investigate.

### 2024-11-11 manuscript_draft.tex
update manuscript

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
   \fi
   % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 \def\languageshorthands#1{}
+\ifLuaTeX
+  \usepackage[english]{selnolig} % disable illegal ligatures
+\fi
 % Manuscript styling
@@
 
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \hypersetup{
-  pdftitle={Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli},
+  pdftitle={Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning Accounting for Item Effects},
   pdfauthor={Erin M.
@@
 
-\title{Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli}
+\title{Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning Accounting for Item Effects}
 \author{Erin M.
@@
 
-\shorttitle{AIPE BOOTSTRAPPING}
+\shorttitle{AIPE MONTE CARLO}
 
@@
 The replication ``crisis'' and credibility revolution have shown that published studies in psychology are underpowered (Korbmacher et al., 2023; Open Science Collaboration, 2015; Vazire, 2018).
+Potential reasons for underpowered studies include questionable research practices (John et al., 2012; but see Fiedler \& Schwarz, 2016), weak psychological theories (Proulx \& Morey, 2021; Szollosi \& Donkin, 2021), testing multiple hypotheses (Maxwell, 2004), and poor intuitions about power (Bakker et al., 2016).
+
 Pre-registration of a study involves outlining the study and hypotheses before data collection begins (Chambers et al., 2014; Nosek \& Lakens, 2014; Stewart et al., 2020), and details of a power analyses or limitations on resources are often used to provide justification for the pre-registered sample quota (Pownall et al., 2023; van den Akker, Assen, et al., 2023; van den Akker, Bakker, et al., 2023).
@@
 Additionally, many tools assume data normality, place an overemphasis on statistical significance, and may rely on simplified assumptions that do not reflect the actual data.
-When computing sample size estimates, it is important to remember that the effects sizes are estimates, not exact calculations guaranteed to produce a specific result (Batterham \& Atkinson, 2005).
+Further, the social sciences often ignores robust statistical methods as an option for analysis (Erceg-Hurn \& Mirosevich, 2008; Field \& Wilcox, 2017), and the implementation of these analyses in power software is somewhat sporadic.
+Finally, when computing sample size estimates, it is important to remember that the effects sizes are estimates, not exact calculations guaranteed to produce a specific result (Batterham \& Atkinson, 2005).
 For example, it is hard to accurately estimate all parameters from a study, and if any were incorrect, then the sample size estimate tied to that specific level of power may be incorrect (Albers \& Lakens, 2018).
@@
 While statistics courses often suggest that a specific research design leads to a specific statistical test, meta-science work has shown that given the same data and hypothesis, researchers can come up with multiple ways to analyze the data (Coretta et al., 2023; Silberzahn et al., 2018).
-Therefore, a single power analysis only corresponds to the specific analysis that the researcher expects to implement.
+Therefore, a single power analysis only corresponds to the specific analysis that the researcher expects to implement, and usually, the final expected data ignoring any processing pipeline that a researcher may implement.
 Analyses may evolve during the research project or be subject to secondary analysis; thus, power and sample size estimation based on one analysis is potentially less useful than previously imagined.
@@
 In light of these analytical (or lack thereof) concerns, we propose a new method to determine a sample size in cases where a more traditional power analysis might be less appropriate or even impossible.
-This approach combines accuracy in parameter estimation (AIPE, Kelley, 2007; Maxwell et al., 2008) and bootstrapped simulation on pilot data (Rousselet et al., 2022).
+This approach combines accuracy in parameter estimation (AIPE, Kelley, 2007; Maxwell et al., 2008) and data-driven Monte-Carlo simulation on pilot data (Rousselet et al., 2022).
 This method accounts for a potential lack of hypothesis test (or simply no good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices.
@@
 
-\hypertarget{accuracy-in-parameter-estimation}{%
-\subsection{Accuracy in Parameter Estimation}\label{accuracy-in-parameter-estimation}}
-
-AIPE shifts the focus away from finding a significant \emph{p}-value to finding a parameter that is ``accurately measured''.
+\subsection{Accuracy in Parameter Estimation}\label{accuracy-in-parameter-estimation}
+
+AIPE shifts the focus away from finding a significant \emph{p}-value to finding a parameter that is accurately measured.
+As discussed in Kelley and Maxwell (Kelley \& Maxwell, 2003), one can estimate the sample size necessary to obtain precision in estimation of population parameters.
+Precision is defined as a researcher defined, sufficiently narrow, confidence interval.
 For example, researchers may wish to detect a specific mean in a study, \emph{M} = .35.
-They could then use AIPE to estimate the sample size needed to find a ``sufficiently narrow'' confidence interval around that mean.
-Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and/or confidence intervals.
-Therefore, they could decide that their 95\% confidence interval should be approximately between .20 and .50, and sufficiently narrow could be defined as a width of .30 or .15 on each side.
-While confidence intervals are related to null hypothesis significance testing (i.e., 95\% confidence intervals that do not include zero would indicate a significant difference from zero at \(\alpha\) \textless{} .05), AIPE procedures suggest how we can define a sample size with a given width of confidence interval, regardless of whether it includes zero.
-
-\hypertarget{bootstrapping-and-simulation}{%
-\subsection{Bootstrapping and Simulation}\label{bootstrapping-and-simulation}}
-
-One form of data simulation is bootstrapping, which involves using data obtained to simulate similar datasets by drawing from the original data with replacement (Efron, 2000; Rousselet et al., 2022).
-Bootstrapping allows one to calculate parameter estimates, confidence intervals, and to simulate the potential population distribution, shape, and bias.
+They could then use AIPE to estimate the sample size needed to find a sufficiently narrow confidence interval around that mean.
+Therefore, they could decide that sufficiently narrow could be defined as a confidence interval width of .30 or .15 on each side of the mean.
+They would then estimate the number of participants needed to find that level of precision.
+Confidence intervals are often misinterpreted (Hoekstra et al., 2014; but see Miller \& Ulrich, 2016), and AIPE procedures are not designed to specify sample size for a hypothesis-driven decision (i.e., the confidence interval does not include a specific value of comparison).
+Instead, AIPE focuses on estimating the sample size necessary to ensure precise population parameters.
+Note that any particular confidence interval is not of interest within our procedure, but rather how to define the sufficiently narrow window a researcher should use for sample size estimation when designing studies with multiple items.
+
+\subsection{Monte Carlo Simulation}\label{monte-carlo-simulation}
+
+One form of data simulation is data-driven Monte-Carlo simulation, which involves using data obtained to simulate similar datasets by drawing from the original data with replacement (Efron, 2000; Rousselet et al., 2022).
+This type of simulation allows one to calculate parameter estimates, confidence intervals, and to simulate the potential population distribution, shape, and bias.
 Simulation is often paired with re-creating a data set with a similar structure for testing analyses and hypotheses based on proposed effect sizes or suggested population means.
@@
 
-Pilot or simulated data would be used to estimate the variability within items and select a ``sufficiently narrow'' window for overall item confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
+Pilot or simulated data would be used to estimate the variability within items and select a sufficiently narrow window for overall item confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
 The advantage to this method over simple power estimation from pilot effect sizes is the multiple simulations to average out potential variability, as well as a shift away from traditional NHST to parameter estimation.
-Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein as many items as required meet the pre-specified well-measured criterion.
-
-\hypertarget{sequential-testing}{%
-\subsection{Sequential Testing}\label{sequential-testing}}
+Simulation would then be used to determine how many participants may be necessary to achieve a dataset wherein as many items as required meet the pre-specified well-measured criterion.
+
+\subsection{Sequential Testing}\label{sequential-testing}
 
 Researchers could then use sequential testing to estimate their parameter of interest after each participant's data or at regular intervals during data collection to determine whether they have achieved their expected width of the confidence interval around that parameter.
-One would set a minimum sample size (e.g., based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the confidence interval is sufficiently narrow, as defined above).
-Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (e.g., they expect 90\% of items to meet their stopping rule with 100 participants based on simulation).
+One would set a minimum sample size based on our procedure steps below to ensure an appropriate minimum for precise estimates.
+Then a researcher could use the standard error as a stopping rule (i.e., stop data collection when the standard error meets the expected sufficiently narrow width, as defined above).
+A stricter criterion could be defined for a stopping rule as a maximum sample size.
 By defining each of these components, researchers could ensure a feasible minimum sample size, a way to stop data collection when goals have been met, and a maximum sample size rule to ensure an actual end to data collection.
 The maximum stopping rule could also be defined by resources (e.g., two semesters data collection), but nevertheless should be included.
-Therefore, we propose a method that leverages the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates to measure an overall phenomena.
-
-\hypertarget{proposed-method-for-sample-size-planning}{%
-\subsection{Proposed Method for Sample Size Planning}\label{proposed-method-for-sample-size-planning}}
+The advantage of sequential testing lies within research studies that use a random selection of items across participants (i.e., participants do not see all items).
+Across participants, data can be shifted to items with more uncertainty to increase precision of estimates for those items.
+Sequential testing is not a necessary component to our proposed procedure but we outline how to use the sample size estimates below to set minimum, maximum, and stopping rules.
+If researchers will show all items to participants, they could simply select one of the proposed estimates for their minimum required sample size.
+Therefore, we propose a method that leverages the ideas behind AIPE, paired with simulation and Monte Carlo simulation, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates.
+
+\subsection{Proposed Method for Sample Size Planning}\label{proposed-method-for-sample-size-planning}
 
@@
 
-\hypertarget{calculate-the-stopping-rule}{%
-\subsubsection{Calculate the Stopping Rule}\label{calculate-the-stopping-rule}}
+\subsubsection{Calculate the Stopping Rule}\label{calculate-the-stopping-rule}
 
@@
   For each item in the pilot data, calculate the standard error (SE).
-Select a cutoff SE that defines when items are considered ``accurately measured''.
+Select a cutoff SE that defines when items are considered accurately measured.
 The simulations described in the Data Simulation section will explore what criterion should be used to determine the cutoff SE from the pilot data.
@@
 
-\hypertarget{bootstrap-samples}{%
-\subsubsection{Bootstrap Samples}\label{bootstrap-samples}}
+\subsubsection{Monte Carlo Samples}\label{monte-carlo-samples}
 
@@
 
-\hypertarget{determine-minimum-maximum-sample-size}{%
-\subsubsection{Determine Minimum, Maximum Sample Size}\label{determine-minimum-maximum-sample-size}}
+\subsubsection{Determine Minimum, Maximum Sample Size}\label{determine-minimum-maximum-sample-size}
 
@@
   Use the simulated SEs to determine the percentage of items that meet the cutoff score determined in Step 2.
-Each sample size from Step 3 will have multiple bootstrapped simulations, and therefore, create an average percentage score for each sample size for Step 5.
+Each sample size from Step 3 will have multiple simulations, and therefore, create an average percentage score for each sample size for Step 5.
 \item
@@
 
-\hypertarget{report-results}{%
-\subsubsection{Report Results}\label{report-results}}
+\subsubsection{Report Results}\label{report-results}
 
@@
 Using the 40\%, determine the cutoff and stopping rule for the standard error of the items.\\
-3 & Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.
-& Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.\\
-4 & Calculate the standard error of each of the items in the bootstrapped data.
+3 & Create simulated samples of your pilot data starting with at least 20 participants up to a maximum number of participants.
+& Create simulated samples of your pilot data starting with at least 20 participants up to a maximum number of participants.\\
+4 & Calculate the standard error of each of the items in the simulated data.
 From these scores, calculate the percent of items below the cutoff score from Step 2.
-& Calculate the standard error of each of the items in the bootstrapped data.
+& Calculate the standard error of each of the items in the simulated data.
 From these scores, calculate the percent of items below the cutoff score from Step 2.\\
@@
 
-\hypertarget{example}{%
-\section{Example}\label{example}}
+\section{Example}\label{example}
 
@@
 
-For this example, imagine a researcher who wants to determine the differences in response latencies for abstract and concrete words.
+For this example, imagine a researcher who wants to determine the differences in response latencies for abstract and concrete words (see the following for thoughts on analyzing this type of data: Bürkner \& Vuorre, 2019; Liddell \& Kruschke, 2018; Taylor et al., 2022).
 They will select \emph{n} = 40 words from the rating data from Brysbaert et al.
@@
 
-We first filtered the ELP data to the same real words as the concreteness subset selected above, and this data includes 27031 real words.
+We first selected the same real words in the ELP data as the concreteness subset selected above, and this data includes 27031 real words.
 The average pilot sample size for this random sample was 32.67 (\emph{SD} = 0.57), and \emph{n} = 33 will be our pilot size for the lexical decision task.
@@
 \emph{Step 3-5}.
-The pilot data was then bootstrapped with replacement creating samples of 20 to 300 participants per item increasing in units of 5, for concreteness ratings and lexical decision latencies separately (Step 3).
+The pilot data was then simulated with replacement creating samples of 20 to 300 participants per item increasing in units of 5, for concreteness ratings and lexical decision latencies separately (Step 3).
 Each of these 57 sample sizes was then repeated 500 times.
-The SE of each item was calculated for the bootstrapped samples separately for concreteness ratings and lexical decision times (Step 4), and the average percentage of items for each sample size (averaging across the 500 simulations) below each potential cutoff was gathered for each (Step 5).
+The SE of each item was calculated for the simulated samples separately for concreteness ratings and lexical decision times (Step 4), and the average percentage of items for each sample size (averaging across the 500 simulations) below each potential cutoff was gathered for each (Step 5).
 The smallest sample size with at least 80\%, 85\%, 90\%, and 95\% of items below the cutoff are reported in Table \ref{tab:table-example} for each task (Step 5).
@@
 \midrule
-Decile 10 & 0.11 & 115 & 125 & 135 & 150 & 33.70 & 170 & 200 & 245 & 345\\
-Decile 20 & 0.14 & 65 & 70 & 75 & 85 & 46.88 & 90 & 105 & 130 & 180\\
-Decile 30 & 0.17 & 50 & 55 & 60 & 65 & 50.45 & 80 & 95 & 115 & 160\\
-Decile 40 & 0.18 & 45 & 45 & 50 & 55 & 56.93 & 60 & 75 & 90 & 125\\
-Decile 50 & 0.19 & 40 & 45 & 45 & 50 & 65.23 & 50 & 60 & 70 & 95\\
-Decile 60 & 0.21 & 35 & 35 & 40 & 45 & 72.51 & 40 & 45 & 60 & 80\\
-Decile 70 & 0.21 & 35 & 35 & 40 & 45 & 81.21 & 30 & 40 & 50 & 65\\
-Decile 80 & 0.23 & 30 & 30 & 35 & 40 & 94.19 & 25 & 30 & 35 & 50\\
-Decile 90 & 0.25 & 25 & 30 & 30 & 35 & 114.51 & 20 & 20 & 25 & 35\\
+Decile 1 & 0.11 & 115 & 125 & 135 & 150 & 33.70 & 170 & 200 & 245 & 345\\
+Decile 2 & 0.14 & 65 & 70 & 75 & 85 & 46.88 & 90 & 105 & 130 & 180\\
+Decile 3 & 0.17 & 50 & 55 & 60 & 65 & 50.45 & 80 & 95 & 115 & 160\\
+Decile 4 & 0.18 & 45 & 45 & 50 & 55 & 56.93 & 60 & 75 & 90 & 125\\
+Decile 5 & 0.19 & 40 & 45 & 45 & 50 & 65.23 & 50 & 60 & 70 & 95\\
+Decile 6 & 0.21 & 35 & 35 & 40 & 45 & 72.51 & 40 & 45 & 60 & 80\\
+Decile 7 & 0.21 & 35 & 35 & 40 & 45 & 81.21 & 30 & 40 & 50 & 65\\
+Decile 8 & 0.23 & 30 & 30 & 35 & 40 & 94.19 & 25 & 30 & 35 & 50\\
+Decile 9 & 0.25 & 25 & 30 & 30 & 35 & 114.51 & 20 & 20 & 25 & 35\\
 \bottomrule
@@
 SE columns represent the standard error value cutoff for each decile, while 80/85/90/95 percent columns represent the sample size needed to have that percent of items below the SE cutoff.
-For example, 150 participants are required to ensure at least 95\% of concreteness items SE are below the 10 percent decile SE cutoff, and 345 participants are necessary for the lexical decision SE to be below its 10 percent decile cutoff.}
+For example, 150 participants are required to ensure at least 95\% of concreteness items SE are below the 1st decile SE cutoff, and 345 participants are necessary for the lexical decision SE to be below its 1st decile cutoff.}
 \end{tablenotes}
@@
 
-\hypertarget{simulation-method}{%
-\section{Simulation Method}\label{simulation-method}}
+\section{Simulation Method}\label{simulation-method}
 
@@
 \item
-  How do pilot data influence sample size suggestions?
+  How does pilot data influence sample size suggestions?
 \end{enumerate}
@@
 
-\hypertarget{data-simulation}{%
-\subsection{Data Simulation}\label{data-simulation}}
+\subsection{Data Simulation}\label{data-simulation}
 
@@
 Each of these samples would correspond to Step 1 of the proposed method where a researcher would use pilot data to start their estimation.
-Therefore, the simulations included 3 scales X 3 heterogeneity values X 2 normal/skewed distributions X 9 pilot sample sizes representing a potential Step 1 of our procedure.
-
-\hypertarget{researcher-sample-simulation}{%
-\subsection{Researcher Sample Simulation}\label{researcher-sample-simulation}}
+Therefore, the simulations included 3 scales X 3 heterogeneity values X 2 symmetric/skewed distributions X 9 pilot sample sizes representing a potential Step 1 of our procedure.
+
+\subsection{Researcher Sample Simulation}\label{researcher-sample-simulation}
 
@@
 
-Next, we calculated the percentage of items that fell below the cutoff score, and therefore, would be considered ``well-measured'' for each decile by sample (Step 4).
+Next, we calculated the percentage of items that fell below the cutoff score, and therefore, would be considered ``well-measured for each decile by sample (Step 4).
 From these data, we pinpoint the smallest suggested sample size at which 80\%, 85\%, 90\% and 95\% of the items fall below the cutoff criterion (Step 5).
-These values were chosen as popular measures of ``power'' in which one could determine the minimum suggested sample size (potentially 80\% of the items) and the maximum suggested sample size (selected from a higher percentage, such as 90\% or 95\%).
+These values were chosen as popular, yet arbitrary, measures of''power'' in which one could determine the minimum suggested sample size (potentially 80\% of the items) and the maximum suggested sample size (selected from a higher percentage, such as 90\% or 95\%).
 
@@
 
-\hypertarget{simulation-results}{%
-\section{Simulation Results}\label{simulation-results}}
-
-\hypertarget{pilot-data-influence-on-sample-size}{%
-\subsection{Pilot Data Influence on Sample Size}\label{pilot-data-influence-on-sample-size}}
+\section{Simulation Results}\label{simulation-results}
+
+\subsection{Pilot Data Influence on Sample Size}\label{pilot-data-influence-on-sample-size}
 
@@
 
-\hypertarget{scale-size}{%
-\subsubsection{Scale Size}\label{scale-size}}
+\subsubsection{Scale Size}\label{scale-size}
 
@@
 \centering
-\includegraphics{manuscript_draft_files/figure-latex/scale-size-figure-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/scale-size-figure-1.pdf}}
 \caption{\label{fig:scale-size-figure}Simulated pilot sample size and final projected sample size to achieve 80\%, 85\%, 90\%, and 95\% of items below threshold.
@@
 
-\hypertarget{skew}{%
-\subsubsection{Skew}\label{skew}}
+\subsubsection{Skew}\label{skew}
 
@@
 \centering
-\includegraphics{manuscript_draft_files/figure-latex/scale-skew-figure-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/scale-skew-figure-1.pdf}}
 \caption{\label{fig:scale-skew-figure}Simulated pilot sample size and final projected sample size to achieve 80\%, 85\%, 90\%, and 95\% of items below threshold.
-In comparison to Figure 1, this figure shows projected sample size for ceiling versus normal distributions on each scale.
+In comparison to Figure 1, this figure shows projected sample size for ceiling versus symmetric distributions on each scale.
 All other variables are averaged together, and black dots represent original sample size for reference.}
@@
 
-Figure \ref{fig:scale-skew-figure} displays that ceiling distributions, averaged over all other variables, show slightly higher estimates than normal distributions.
+Figure \ref{fig:scale-skew-figure} displays that ceiling distributions, averaged over all other variables, show slightly higher estimates than symmetric distributions.
 This result is consistent across scale type and heterogeneity, as results indicated that they are often the same or slightly higher for ceiling distributions.
 
-\hypertarget{item-heterogeneity}{%
-\subsubsection{Item Heterogeneity}\label{item-heterogeneity}}
+\subsubsection{Item Heterogeneity}\label{item-heterogeneity}
 
@@
 \centering
-\includegraphics{manuscript_draft_files/figure-latex/scale-hetero-figure-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/scale-hetero-figure-1.pdf}}
 \caption{\label{fig:scale-hetero-figure}Simulated pilot sample size and final projected sample size to achieve 80\%, 85\%, 90\%, and 95\% of items below threshold.
@@
 Proportion Variability & 312.44 & 19.86 & 15.73 & < .001 & .613\\
-Data: Ceiling v Normal & -7.16 & 1.41 & -5.08 & < .001 & .142\\
+Data: Ceiling v Symmetric & -7.16 & 1.41 & -5.08 & < .001 & .142\\
 \bottomrule
@@
 
-Using a regression model, we predicted proposed sample size using pilot sample size, scale size, proportion variability (i.e., heterogeneity), and data type (normal, ceiling).
+Using a regression model, we predicted proposed sample size using pilot sample size, scale size, proportion variability (i.e., heterogeneity), and data type (symmetric, ceiling).
 As shown in Table \ref{tab:table-predict-largest}, the largest influence on proposed sample size is the original pilot sample size, followed by proportion of variance/heterogeneity, and then data and scale sizes.
 
-\hypertarget{projected-sample-size-sensitivity-to-pilot-sample-size}{%
-\subsection{Projected Sample Size Sensitivity to Pilot Sample Size}\label{projected-sample-size-sensitivity-to-pilot-sample-size}}
+\subsection{Projected Sample Size Sensitivity to Pilot Sample Size}\label{projected-sample-size-sensitivity-to-pilot-sample-size}
 
@@
 \centering
-\includegraphics{manuscript_draft_files/figure-latex/corrected-figure-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/corrected-figure-1.pdf}}
 \caption{\label{fig:corrected-figure}Corrected projected sample sizes for variability and power levels to achieve 80\%, 85\%, 90\%, and 95\% of items below threshold.
@@
 
-\hypertarget{corrections-for-individual-researchers}{%
-\subsection{Corrections for Individual Researchers}\label{corrections-for-individual-researchers}}
+\subsection{Corrections for Individual Researchers}\label{corrections-for-individual-researchers}
 
@@
 Given the non-linear nature of the correction, we added each variable and its non-linear \texttt{log2} transform to the regression equation, as this function was used to create the correction.
-The intercept only model was used as a starting point (i.e., \texttt{corrected\ sample\ \textasciitilde{}\ 1}), and then all eight variables (each variable and their \texttt{log2} transform) were entered into a forward stepwise regression to capture the corrected scores with the most predictive values.
-Each variable was entered one at a time using the \texttt{step} function from the \emph{stats} library in \emph{R} (R Core Team, 2022).
-
-As shown in Table \ref{tab:table-correction}, all variables were included in the final equation, each contributing a significant change to the previous model, as defined by \(\Delta\)AIC \textgreater{} 2 points change between each step of the model.
+The intercept only model was used as a starting point (i.e., \texttt{corrected\ sample\ \textasciitilde{}\ 1}), and then all eight variables (each variable and their \texttt{log2} transform) were entered into the regression equation.
+
+As shown in Table \ref{tab:table-correction}, all variables were significant predictors of the new sample size.
 Proposed sample size and original sample size were the largest predictors -- unsurprising given the correction formula employed -- followed by the percent ``power'' level and proportion of variance.
@@
 
-\begin{tabular}{llllll}
+\begin{tabular}{lllll}
 \toprule
-Term & Estimate & $SE$ & $t$ & $p$ & AIC\\
+Term & Estimate & $SE$ & $t$ & $p$\\
 \midrule
-Intercept & 111.049 & 78.248 & 1.419 & .156 & 29,996.94\\
-Projected Sample Size & 0.429 & 0.002 & 185.360 & < .001 & 20,327.79\\
-Pilot Sample Size & -0.718 & 0.007 & -103.787 & < .001 & 14,753.61\\
-Log2 Projected Sample Size & 19.522 & 0.215 & 90.693 & < .001 & 8,668.73\\
-Log2 Pilot Sample Size & 4.655 & 0.269 & 17.296 & < .001 & 8,363.69\\
-Log2 Power & -39.367 & 15.640 & -2.517 & .012 & 8,320.82\\
-Proportion Variability & 15.434 & 3.617 & 4.267 & < .001 & 8,297.71\\
-Log2 Proportion Variability & -0.729 & 0.232 & -3.143 & .002 & 8,289.81\\
-Power & 0.606 & 0.259 & 2.343 & .019 & 8,286.31\\
+Intercept & 111.049 & 78.248 & 1.419 & .156\\
+Projected Sample Size & 0.429 & 0.002 & 185.360 & < .001\\
+Pilot Sample Size & 15.434 & 3.617 & 4.267 & < .001\\
+Log2 Projected Sample Size & -0.718 & 0.007 & -103.787 & < .001\\
+Log2 Pilot Sample Size & 0.606 & 0.259 & 2.343 & .019\\
+Log2 Power & 19.522 & 0.215 & 90.693 & < .001\\
+Proportion Variability & -0.729 & 0.232 & -3.143 & .002\\
+Log2 Proportion Variability & 4.655 & 0.269 & 17.296 & < .001\\
+Power & -39.367 & 15.640 & -2.517 & .012\\
 \bottomrule
@@
 
-\hypertarget{choosing-an-appropriate-cutoff}{%
-\subsection{Choosing an Appropriate cutoff}\label{choosing-an-appropriate-cutoff}}
+\subsection{Choosing an Appropriate cutoff}\label{choosing-an-appropriate-cutoff}
 
 Last, we examined the question of an appropriate SE decile.
-First, the 0\%, 10\%, and 20\% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale size, and heterogeneity.
+First, the 0, 1st, and 2nd deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale size, and heterogeneity.
 If we examine the \(R^2\) values for each decile of our regression equation separately, we find that the values are all \(R^2\) \textgreater{} .99 with very little differences between them.
-Figures \ref{fig:decile40-figure} and \ref{fig:decile50-figure} illustrate the corrected scores for simulations at the 40\% and 50\% decile recommended cutoff for item standard errors.
-For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 40\% decile range, especially for scales with larger potential variance.
-Therefore, we would suggest the 40\% decile to overpower each item for Step 2.
+Figures \ref{fig:decile40-figure} and \ref{fig:decile50-figure} illustrate the corrected scores for simulations at the 4th and 5th decile recommended cutoff for item standard errors.
+For small heterogeneity, differences in decile are minimal, while larger heterogeneity shows more correction at the 4th decile range, especially for scales with larger potential variance.
+Therefore, we would suggest the 4th decile to overpower each item for Step 2.
 
@@
 \centering
-\includegraphics{manuscript_draft_files/figure-latex/decile40-figure-1.pdf}
-\caption{\label{fig:decile40-figure}Comparison of the cutoffs for 40\% deciles across heterogeneity (columns), powering of items (rows), and scale size (color).}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/decile40-figure-1.pdf}}
+\caption{\label{fig:decile40-figure}Comparison of the cutoffs for 4th deciles across heterogeneity (columns), powering of items (rows), and scale size (color).}
 \end{figure}
@@
 \centering
-\includegraphics{manuscript_draft_files/figure-latex/decile50-figure-1.pdf}
-\caption{\label{fig:decile50-figure}Comparison of the cutoffs for 50\% deciles across heterogeneity (columns), powering of items (rows), and scale size (color).}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/decile50-figure-1.pdf}}
+\caption{\label{fig:decile50-figure}Comparison of the cutoffs for 5th decile across heterogeneity (columns), powering of items (rows), and scale size (color).}
 \end{figure}
 
-The final formula for 40\% decile correction is provided in Table \ref{tab:table-decile}. Proportion of variance can be calculated with the following:
+The final formula for 4th decile correction is provided in Table \ref{tab:table-decile}. Proportion of variance can be calculated with the following:
 
@@
 
-\caption{\label{tab:table-decile}Parameters for 40\% Decile Cutoff Scores}
+\caption{\label{tab:table-decile}Parameters for 4th decile Cutoff Scores}
 
@@
 
-\hypertarget{updated-example}{%
-\section{Updated Example}\label{updated-example}}
+\section{Updated Example}\label{updated-example}
 
@@
 The main change occurs in Step 2 with a designated cutoff decile, and Step 5 with a correction score.
-Using the data from the 40\% decile in Table \ref{tab:table-example}, we can determine that the stopping rule SE for concreteness ratings would be 0.18, and the stopping rule SE for lexical decision times would be 56.93.
+Using the data from the 4th decile in Table \ref{tab:table-example}, we can determine that the stopping rule SE for concreteness ratings would be 0.18, and the stopping rule SE for lexical decision times would be 56.93.
 For Step 5, we apply our correction formula separately for each one, as they have different variability scores, and these scores are shown in Table \ref{tab:table-scores-updated}.
@@
 
-\hypertarget{additional-materials}{%
-\section{Additional Materials}\label{additional-materials}}
-
-\hypertarget{package}{%
-\subsection{Package}\label{package}}
-
-We have developed functions to implement the suggested procedure as part of an upcoming package \texttt{semanticprimeR}.
+\section{Additional Materials}\label{additional-materials}
+
+\subsection{Package}\label{package}
+
+We have developed functions to implement the suggested procedure as part of a semantic priming focused package \texttt{semanticprimeR}.
 You can install the package from GitHub using: \texttt{devtools::install\_github("SemanticPriming/semanticprimeR")}.
@@
 \emph{Step 2}.
-In step 2, we can use \texttt{calculate\_cutoff()} to calculate the standard error of the items, the standard deviation of the standard errors and the corresponding proportion of variance possible, and the 40\% decile cutoff score.
+In step 2, we can use \texttt{calculate\_cutoff()} to calculate the standard error of the items, the standard deviation of the standard errors and the corresponding proportion of variance possible, and the 4th decile cutoff score.
 The \texttt{pops} dataframe can be used in this function, which has columns named \texttt{item} for the item labels (i.e., 1, 2, 3, 4 or characters can be used), and \texttt{score} for the dependent variable.
@@
 \begin{Highlighting}[]
-\NormalTok{cutoff}\SpecialCharTok{$}\NormalTok{cutoff }\CommentTok{\# 40\% decile score}
+\NormalTok{cutoff}\SpecialCharTok{$}\NormalTok{cutoff }\CommentTok{\# 4th decile score}
 \end{Highlighting}
@@
 \emph{Step 3}.
-The \texttt{bootstrap\_samples()} function creates bootstrapped samples from the pilot or simulated population data to estimate the number of participants needed for item standard error to be below the cutoff calculated in Step 2.
+The \texttt{simulate\_samples()} function creates simulated samples from the pilot or simulated population data to estimate the number of participants needed for item standard error to be below the cutoff calculated in Step 2.
 This function returns a list of samples with sizes that start at the \texttt{start} size, increase by \texttt{increase}, and end with the \texttt{stop} sample size.
 The population or pilot data will be included in \texttt{population}, and the item column indicator should be included in \texttt{grouping\_items}.
-The \texttt{nsim} argument determines the number of bootstrapped simulations to run.
+The \texttt{nsim} argument determines the number of simulations to run.
 
@@
 \begin{Highlighting}[]
-\NormalTok{samples }\OtherTok{\textless{}{-}} \FunctionTok{bootstrap\_samples}\NormalTok{(}\AttributeTok{start =} \DecValTok{20}\NormalTok{, }\CommentTok{\# starting sample size}
+\NormalTok{samples }\OtherTok{\textless{}{-}} \FunctionTok{simulate\_samples}\NormalTok{(}\AttributeTok{start =} \DecValTok{20}\NormalTok{, }\CommentTok{\# starting sample size}
   \AttributeTok{stop =} \DecValTok{100}\NormalTok{, }\CommentTok{\# stopping sample size}
-  \AttributeTok{increase =} \DecValTok{5}\NormalTok{, }\CommentTok{\# increase bootstrapped samples by this amount}
+  \AttributeTok{increase =} \DecValTok{5}\NormalTok{, }\CommentTok{\# increase simulated samples by this amount}
   \AttributeTok{population =}\NormalTok{ pops, }\CommentTok{\# population or pilot data}
-  \AttributeTok{replace =} \ConstantTok{TRUE}\NormalTok{, }\CommentTok{\# bootstrap with replacement? }
+  \AttributeTok{replace =} \ConstantTok{TRUE}\NormalTok{, }\CommentTok{\# simulate with replacement? }
   \AttributeTok{nsim =} \DecValTok{500}\NormalTok{, }\CommentTok{\# number of simulations to run}
@@
 ##   <int> <dbl>
-## 1     1     4
-## 2     1     3
-## 3     1     2
-## 4     1     3
-## 5     1     3
-## 6     1     3
+## 1     1     3
+## 2     1     4
+## 3     1     3
+## 4     1     4
+## 5     1     4
+## 6     1     4
 \end{verbatim}
@@
 \emph{Step 4 and 5}.
-The proportion of bootstrapped items across sample sizes below the cutoff score can then be calculated using \texttt{calculate\_proportion()}.
+The proportion of simulated items across sample sizes below the cutoff score can then be calculated using \texttt{calculate\_proportion()}.
 This function returns a dataframe including each sample size with the proportion of items below that cutoff to use in the next function.
@@
 
-\hypertarget{vignettes}{%
-\subsection{Vignettes}\label{vignettes}}
+\subsection{Special Consideration: Bimodal Data}\label{special-consideration-bimodal-data}
+
+The above simulations assume symmetric population distributions with small, medium, or large heterogeneity or a skewed distribution with the same heterogeneity.
+However, as shown in Pollock (2018), sample rating data that appears to have a mean distribution may represent the average two separate distributions (i.e., a bimodal distribution).
+To examine our procedure with bimodal distributions, we estimated thirty items for three distributions: symmetric, ceiling, and floor distributions for Likert data with small heterogeneity.
+Bimodal data was created by selecting half of the ceiling distribution and floor distribution to combine together.
+The number of bimodal items was varied from 0 percent to 100 percent increasing by 10 percent increments for simulated samples.
+Figure @\ref(fig:bimodal-figure) shows an example of 10 of the items within a simulation that estimated that half of the Likert items would be bimodal in nature.
+The researcher procedure described above was then carried out using \emph{semanticprimeR}.
+
+\begin{figure}
+\centering
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/bimodal-figure-1.pdf}}
+\caption{\label{fig:bimodal-figure}Simulated example of Likert data with half of the items as bimodal distributions.}
+\end{figure}
+
+Figure @\ref(fig:bimodal-figure) portrays the results of the suggested procedure.
+The sample with 0 bimodal items demonstrates the same values as above.
+The proposed sample size increases from 10 to 40\% because of the heterogeneity in SE across items (i.e., with 10\% of items with a larger SE due to their bimodal nature, the required sample size increases).
+From 50\% to 100\% the sample size decreases because the variance of the SEs across items decreases (i.e., at 100 percent, they are all larger rather than a mix).
+In theory, these results map onto the conceptual framework - if all items are truly bimodal, we have precisely measured the mean of the bimodal distribution, but this result is likely not the intended result.
+If researchers expect these distributions (or have representative pilot data), they could examine the data for bimodal items.
+These items could be estimated separately (i.e., floor effects item 1, ceiling effects item 1) to ensure that both populations of answers are represented in the data.
+
+\begin{figure}
+\centering
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/bimodal-results-1.pdf}}
+\caption{\label{fig:bimodal-results}Estimated sample sizes for distributions that vary in the number of bimodal items. Note that percent below is treated as a continuous factor}
+\end{figure}
+
+\subsection{Vignettes}\label{vignettes}
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-We proposed a method combining AIPE, bootstrapping, and simulation to estimate a minimum and maximum sample size and to define a rule for stopping data collection based on narrow confidence intervals on a parameter of interest.
+\section{Discussion}\label{discussion}
+
+We proposed a method combining AIPE and Monte Carlo simulation to estimate a minimum and maximum sample size and to define a rule for stopping data collection based on narrow confidence intervals on a parameter of interest.
 In addition, we also demonstrated its practical applications using real-world data.
@@
 For example, the English Lexicon Project includes thousands of items, and if we were to simulate for all of those, our results would likely suggest needing thousands of participants for most items to reach the criterion.
-Additionally, as the number of items increases, you may also see very small estimates for sample size due to the correction factor (as with large numbers of items, you could find many items with standard errors below the 40\% decile).
+Additionally, as the number of items increases, you may also see very small estimates for sample size due to the correction factor (as with large numbers of items, you could find many items with standard errors below the 4th decile).
 Therefore, it would be beneficial to consider only simulating what a participant would reasonably complete in a study.
@@
 
-Second, these methods do not ensure the normal interpretation of power, where you know that you would find a specific effect for a specific test, \(\alpha\), and so on.
+Second, these methods do not ensure the normal interpretation of power, focusing on finding a specific effect for a specific test, \(\alpha\), and so on.
 As discussed in the introduction, there is not necessarily a one-to-one mapping of hypothesis to analysis; many of the estimations within a traditional power analysis are just that - best approximations for various parameters.
 These proposed methods and traditional power analysis could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
-
-Researchers should consider this hybrid approach for AIPE, bootstrapping, and simulation as a powerful tool for hypothesis testing and parameter estimation.
-This procedure holds benefits for various research studies, specifically replication studies, that usually prioritize subject sample size but rarely item sample size, in spite of the fact that item sample sizes can contribute to power in multilevel models (Brysbaert \& Stevens, 2018).
+We would advise caution when oversampling (i.e., simulation from data with larger sample sizes than the pilot data) for null-hypothesis power estimates, as these may be upwardly biased in small samples creating Type-I error (Burns et al., n.d.).
+
+Researchers should consider this hybrid approach for AIPE and simulation as a powerful tool for hypothesis testing and parameter estimation.
+This procedure holds benefits for various research studies, specifically replication studies, that usually prioritize subject sample size but rarely item sample size, in spite of the fact that item sample sizes can contribute to power in multilevel models (Brysbaert \& Stevens, 2018; however, see Rouder \& Haaf, 2018 for a discussion of the item-sample size trade off).
 Replicated effects, accumulated through multiple studies and accurate measurement, contribute to robust meta-analyses, enhancing our understanding of the genuine nature of observed effects.
-This article helps to achieve this goal by encouraging researchers to conduct studies where the power analysis is not based on the size of the effect but on adequate sampling of the stimuli.
+This article helps to achieve this goal by encouraging researchers to conduct studies where the power analysis is not based on the size of the effect but on precise measurement of the stimuli.
 We argue that this article can be the initial step to apply AIPE in a manner that can allow researchers to use item information to provide a more accurate and statistically reliable measure of the effect we aimed to investigate.
@@
 
-\hypertarget{open-practices}{%
-\section{Open Practices}\label{open-practices}}
+\section{Open Practices}\label{open-practices}
 
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+\section{References}\label{references}
 
@@
 
-\hypertarget{refs}{}
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-albers2018}{}}%
+\bibitem[\citeproctext]{ref-albers2018}
 Albers, C., \& Lakens, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-alexander1994}{}}%
+\bibitem[\citeproctext]{ref-alexander1994}
 Alexander, R.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-anderson2017}{}}%
+\bibitem[\citeproctext]{ref-anderson2017}
 Anderson, S.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-anvari2021}{}}%
+\bibitem[\citeproctext]{ref-anvari2021}
 Anvari, F., \& Lakens, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-aust2022}{}}%
+\bibitem[\citeproctext]{ref-aust2022}
 Aust, F., Barth, M., Diedenhofen, B., Stahl, C., Casillas, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-balota2007}{}}%
+\bibitem[\citeproctext]{ref-bakker2016}
+Bakker, M., Hartgerink, C.
+H.
+J., Wicherts, J.
+M., \& Van Der Maas, H.
+L.
+J.
+(2016).
+Researchers{'} Intuitions About Power in Psychological Research.
+\emph{Psychological Science}, \emph{27}(8), 1069--1077.
+\url{https://doi.org/10.1177/0956797616647519}
+
+\bibitem[\citeproctext]{ref-balota2007}
 Balota, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-barzykowski2019}{}}%
+\bibitem[\citeproctext]{ref-barzykowski2019}
 Barzykowski, K., Niedźwieńska, A., \& Mazzoni, G.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-batterham2005}{}}%
+\bibitem[\citeproctext]{ref-batterham2005}
 Batterham, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-beribisky2019}{}}%
+\bibitem[\citeproctext]{ref-beribisky2019}
 Beribisky, N., Alter, U., \& Cribbie, R.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-brysbaert2019}{}}%
+\bibitem[\citeproctext]{ref-brysbaert2019}
 Brysbaert, M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-brysbaert2018}{}}%
+\bibitem[\citeproctext]{ref-brysbaert2018}
 Brysbaert, M., \& Stevens, M. (2018). Power Analysis and Effect Size in Mixed Effects Models: A Tutorial. \emph{Journal of Cognition}, \emph{1}(1), 9. \url{https://doi.org/10.5334/joc.10}
 
-\leavevmode\vadjust pre{\hypertarget{ref-brysbaert2014}{}}%
+\bibitem[\citeproctext]{ref-brysbaert2014}
 Brysbaert, M., Warriner, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-buchanan2019}{}}%
+\bibitem[\citeproctext]{ref-buchanan2019}
 Buchanan, E.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-chalmers2020}{}}%
+\bibitem[\citeproctext]{ref-buxfcrkner2019}
+Bürkner, P.-C., \& Vuorre, M.
+(2019).
+Ordinal Regression Models in Psychology: A Tutorial.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 77--101.
+\url{https://doi.org/10.1177/2515245918823199}
+
+\bibitem[\citeproctext]{ref-burns}
+Burns, C. D. G., Fracasso, A., \& Rousselet, G. A. (n.d.). \emph{Data-driven estimates of the reproducibility of univariate BWAS are biased}. \url{https://doi.org/10.1101/2023.09.21.558661}
+
+\bibitem[\citeproctext]{ref-chalmers2020}
 Chalmers, R.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-d.chambers2014}{}}%
+\bibitem[\citeproctext]{ref-d.chambers2014}
 Chambers, C.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-champely2017}{}}%
+\bibitem[\citeproctext]{ref-champely2017}
 Champely, S., Ekstrom, C., Dalgaard, P., Gill, J., Weibelzahl, S., Anandkumar, A., Ford, C., Volcic, R., \& De Rosario, H. (2017). \emph{Pwr: Basic functions for power analysis}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-cohen1990}{}}%
+\bibitem[\citeproctext]{ref-cohen1990}
 Cohen, J. (1990). Things I have learned (so far). \emph{American Psychologist}, \emph{45}(12), 1304--1312. \url{https://doi.org/10.1037/0003-066X.45.12.1304}
 
-\leavevmode\vadjust pre{\hypertarget{ref-coretta}{}}%
+\bibitem[\citeproctext]{ref-coretta}
 Coretta, S., Casillas, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dedeyne2008}{}}%
+\bibitem[\citeproctext]{ref-dedeyne2008}
 De Deyne, S., Verheyen, S., Ameel, E., Vanpaemel, W., Dry, M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-debruine2021}{}}%
+\bibitem[\citeproctext]{ref-debruine2021}
 DeBruine, L. (2021). \emph{Faux: Simulation for factorial designs}. Zenodo. \url{https://doi.org/10.5281/ZENODO.2669586}
 
-\leavevmode\vadjust pre{\hypertarget{ref-efron2000}{}}%
+\bibitem[\citeproctext]{ref-efron2000}
 Efron, B. (2000). The bootstrap and modern statistics. \emph{Journal of the American Statistical Association}, \emph{95}(452), 1293--1296. \url{https://doi.org/10.1080/01621459.2000.10474333}
 
-\leavevmode\vadjust pre{\hypertarget{ref-erdfelder1996}{}}%
+\bibitem[\citeproctext]{ref-erceg-hurn2008}
+Erceg-Hurn, D.
+M., \& Mirosevich, V.
+M.
+(2008).
+Modern robust statistical methods: An easy way to maximize the accuracy and power of your research.
+\emph{American Psychologist}, \emph{63}(7), 591--601.
+\url{https://doi.org/10.1037/0003-066X.63.7.591}
+
+\bibitem[\citeproctext]{ref-erdfelder1996}
 Erdfelder, E., Faul, F., \& Buchner, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-10.7554ux2feLife.71601}{}}%
+\bibitem[\citeproctext]{ref-10.7554ux2feLife.71601}
 Errington, T.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-faul2007}{}}%
+\bibitem[\citeproctext]{ref-faul2007}
 Faul, F., Erdfelder, E., Lang, A.-G., \& Buchner, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-grahe2022}{}}%
+\bibitem[\citeproctext]{ref-fiedler2016}
+Fiedler, K., \& Schwarz, N. (2016). Questionable Research Practices Revisited. \emph{Social Psychological and Personality Science}, \emph{7}(1), 45--52. \url{https://doi.org/10.1177/1948550615612150}
+
+\bibitem[\citeproctext]{ref-field2017}
+Field, A.
+P., \& Wilcox, R.
+R.
+(2017).
+Robust statistical methods: A primer for clinical psychology and experimental psychopathology researchers.
+\emph{Behaviour Research and Therapy}, \emph{98}, 19--38.
+\url{https://doi.org/10.1016/j.brat.2017.05.013}
+
+\bibitem[\citeproctext]{ref-grahe2022}
 Grahe, J., Chalk, H., Cramblet Alvarez, L., Faas, C., Hermann, A., McFall, J., \& Molyneux, K. (2022). EAMMi2 public data. \emph{Open Science Framework}. \url{https://doi.org/10.17605/OSF.IO/X7MP2}
 
-\leavevmode\vadjust pre{\hypertarget{ref-simr:an}{}}%
+\bibitem[\citeproctext]{ref-simr:an}
 Green, P., \& MacLeod, C.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-doi:10.1177ux2f1745691620979806}{}}%
+\bibitem[\citeproctext]{ref-doi:10.1177ux2f1745691620979806}
 Hardwicke, T.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hardwicke2020}{}}%
+\bibitem[\citeproctext]{ref-hardwicke2020}
 Hardwicke, T.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-heyman2014}{}}%
+\bibitem[\citeproctext]{ref-heyman2014}
 Heyman, T., De Deyne, S., Hutchison, K.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kelley2007}{}}%
+\bibitem[\citeproctext]{ref-hoekstra2014}
+Hoekstra, R., Morey, R.
+D., Rouder, J.
+N., \& Wagenmakers, E.-J.
+(2014).
+Robust misinterpretation of confidence intervals.
+\emph{Psychonomic Bulletin \& Review}, \emph{21}(5), 1157--1164.
+\url{https://doi.org/10.3758/s13423-013-0572-3}
+
+\bibitem[\citeproctext]{ref-john2012}
+John, L.
+K., Loewenstein, G., \& Prelec, D.
+(2012).
+Measuring the Prevalence of Questionable Research Practices With Incentives for Truth Telling.
+\emph{Psychological Science}, \emph{23}(5), 524--532.
+\url{https://doi.org/10.1177/0956797611430953}
+
+\bibitem[\citeproctext]{ref-kelley2007}
 Kelley, K.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-korbmacher2023}{}}%
+\bibitem[\citeproctext]{ref-kelley2003}
+Kelley, K., \& Maxwell, S.
+E.
+(2003).
+Sample Size for Multiple Regression: Obtaining Regression Coefficients That Are Accurate, Not Simply Significant.
+\emph{Psychological Methods}, \emph{8}(3), 305--321.
+\url{https://doi.org/10.1037/1082-989X.8.3.305}
+
+\bibitem[\citeproctext]{ref-korbmacher2023}
 Korbmacher, M., Azevedo, F., Pennington, C.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kumle2020}{}}%
+\bibitem[\citeproctext]{ref-kumle2020}
 Kumle, L., \& DejanDraschkow. (2020). \emph{DejanDraschkow/mixedpower: The force awakens}. Zenodo. \url{https://doi.org/10.5281/zenodo.3733023}
 
-\leavevmode\vadjust pre{\hypertarget{ref-maxwell2008}{}}%
+\bibitem[\citeproctext]{ref-liddell2018}
+Liddell, T.
+M., \& Kruschke, J.
+K.
+(2018).
+Analyzing ordinal data with metric models: What could possibly go wrong?
+\emph{Journal of Experimental Social Psychology}, \emph{79}, 328--348.
+\url{https://doi.org/10.1016/j.jesp.2018.08.009}
+
+\bibitem[\citeproctext]{ref-maxwell2004}
+Maxwell, S.
+E.
+(2004).
+The Persistence of Underpowered Studies in Psychological Research: Causes, Consequences, and Remedies.
+\emph{Psychological Methods}, \emph{9}(2), 147--163.
+\url{https://doi.org/10.1037/1082-989X.9.2.147}
+
+\bibitem[\citeproctext]{ref-maxwell2008}
 Maxwell, S.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-meyer1971}{}}%
+\bibitem[\citeproctext]{ref-meyer1971}
 Meyer, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-montefinese2021}{}}%
+\bibitem[\citeproctext]{ref-miller2016}
+Miller, J., \& Ulrich, R.
+(2016).
+Interpreting confidence intervals: A comment on Hoekstra, Morey, Rouder, and Wagenmakers (2014).
+\emph{Psychonomic Bulletin \& Review}, \emph{23}(1), 124--130.
+\url{https://doi.org/10.3758/s13423-015-0859-7}
+
+\bibitem[\citeproctext]{ref-montefinese2021}
 Montefinese, M., Ambrosini, E., \& Angrilli, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-montefinese2022}{}}%
+\bibitem[\citeproctext]{ref-montefinese2022}
 Montefinese, M., Vinson, D., Vigliocco, G., \& Ambrosini, E.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-nosek2014}{}}%
+\bibitem[\citeproctext]{ref-nosek2014}
 Nosek, B.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-nuijten2016}{}}%
+\bibitem[\citeproctext]{ref-nuijten2016}
 Nuijten, M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-opensciencecollaboration2015}{}}%
+\bibitem[\citeproctext]{ref-opensciencecollaboration2015}
 Open Science Collaboration. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716--aac4716. \url{https://doi.org/10.1126/science.aac4716}
 
-\leavevmode\vadjust pre{\hypertarget{ref-peterson2022}{}}%
+\bibitem[\citeproctext]{ref-peterson2022}
 Peterson, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-pownall2023}{}}%
+\bibitem[\citeproctext]{ref-pollock2018}
+Pollock, L.
+(2018).
+Statistical and methodological problems with concreteness and other semantic variables: A list memory experiment case study.
+\emph{Behavior Research Methods}, \emph{50}(3), 1198--1216.
+\url{https://doi.org/10.3758/s13428-017-0938-y}
+
+\bibitem[\citeproctext]{ref-pownall2023}
 Pownall, M., Pennington, C.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-R-base}{}}%
-R Core Team. (2022). \emph{R: A language and environment for statistical computing}. \url{https://www.R-project.org/}
-
-\leavevmode\vadjust pre{\hypertarget{ref-rheinheimer2001}{}}%
+\bibitem[\citeproctext]{ref-proulx2021}
+Proulx, T., \& Morey, R.
+D.
+(2021).
+Beyond Statistical Ritual: Theory in Psychological Science.
+\emph{Perspectives on Psychological Science}, \emph{16}(4), 671--681.
+\url{https://doi.org/10.1177/17456916211017098}
+
+\bibitem[\citeproctext]{ref-rheinheimer2001}
 Rheinheimer, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ruxf6er2013}{}}%
+\bibitem[\citeproctext]{ref-ruxf6er2013}
 Röer, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-rosenthal1979}{}}%
+\bibitem[\citeproctext]{ref-rosenthal1979}
 Rosenthal, R. (1979). The file drawer problem and tolerance for null results. \emph{Psychological Bulletin}, \emph{86}(3), 638--641. \url{https://doi.org/10.1037/0033-2909.86.3.638}
 
-\leavevmode\vadjust pre{\hypertarget{ref-rousselet}{}}%
+\bibitem[\citeproctext]{ref-rouder2018}
+Rouder, J.
+N., \& Haaf, J.
+M.
+(2018).
+Power, Dominance, and Constraint: A Note on the Appeal of Different Design Traditions.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(1), 19--26.
+\url{https://doi.org/10.1177/2515245917745058}
+
+\bibitem[\citeproctext]{ref-rousselet}
 Rousselet, G., Pernet, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-silberzahn2018many}{}}%
+\bibitem[\citeproctext]{ref-silberzahn2018many}
 Silberzahn, R., Uhlmann, E.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-simmons2011}{}}%
+\bibitem[\citeproctext]{ref-simmons2011}
 Simmons, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-stewart2020}{}}%
+\bibitem[\citeproctext]{ref-stewart2020}
 Stewart, S., Rinke, E.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ulloa2014}{}}%
+\bibitem[\citeproctext]{ref-szollosi2021}
+Szollosi, A., \& Donkin, C.
+(2021).
+Arrested Theory Development: The Misguided Distinction Between Exploratory and Confirmatory Research.
+\emph{Perspectives on Psychological Science}, \emph{16}(4), 717--724.
+\url{https://doi.org/10.1177/1745691620966796}
+
+\bibitem[\citeproctext]{ref-taylor2022}
+Taylor, J.
+E., Rousselet, G.
+A., Scheepers, C., \& Sereno, S.
+C.
+(2022).
+Rating norms should be calculated from cumulative link mixed effects models.
+\emph{Behavior Research Methods}, \emph{55}(5), 2175--2196.
+\url{https://doi.org/10.3758/s13428-022-01814-7}
+
+\bibitem[\citeproctext]{ref-ulloa2014}
 Ulloa, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-vandenakker2023}{}}%
+\bibitem[\citeproctext]{ref-vandenakker2023}
 van den Akker, O.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-akker2023}{}}%
+\bibitem[\citeproctext]{ref-akker2023}
 van den Akker, O.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-vazire2018}{}}%
+\bibitem[\citeproctext]{ref-vazire2018}
 Vazire, S.

### 2024-11-11 simulation_study/simulate_summary_bimodal.Rmd
update manuscript

@@
 # simulate population 
-normal_pops <- simulate_population(mu = 4, # item means
+normal_pops <-semanticprimeR::simulate_population(mu = 4, # item means
   mu_sigma = .2, # variability in item means 
@@
 
-floor_pops <- simulate_population(mu = 2, # item means
+floor_pops <- semanticprimeR::simulate_population(mu = 2, # item means
   mu_sigma = .2, # variability in item means 
@@
 
-ceiling_pops <- simulate_population(mu = 6, # item means
+ceiling_pops <- semanticprimeR::simulate_population(mu = 6, # item means
   mu_sigma = .2, # variability in item means 
@@
 item_number <- 1:30
-prop_mix <- 1 # manipulated this number 
+prop_mix <- .5 # manipulated this number 
 pilot_sample_size <- 30
@@
   )
+
+library(ggridges)
+
+ggplot(researcher_sample %>% 
+         filter(item < 11), aes(x = score, y = factor(item), fill = factor(item))) +
+  geom_density_ridges(alpha = 0.7, scale = 1) +
+  labs(x = "Score", y = "Item") +
+  theme_minimal() +
+  theme(legend.position = "none") + 
+  coord_cartesian(xlim = c(1,7))
 

### 2024-11-11 manuscript_draft.Rmd
updating the manuscript submission

@@
 ```{r bimodal-results, fig.cap = "Estimated sample sizes for distributions that vary in the number of bimodal items.
-Note that percent below is treated as a continuous factor", fig.height=8, fig.width=8}
+Note that percent below is treated as a continuous factor, and therefore, some estimates are the same (i.e., below 80 and 85 percent may be both estimated as 88%).", fig.height=8, fig.width=8}
 ggplot(bi_DF, aes(x = percent_below,

### 2024-11-11 authorship.R

@@ -1,14 +1,0 @@
-# devtools::install_github("marton-balazs-kovacs/tenzing")
-library(tenzing)
-library(googlesheets4)
-
-DF <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1KMBSWE_3IPmwN1mgjF6wEWp5rA89iTXqpOQbwdY7Ag0/edit?usp=sharing")
-
-nrow(DF)
-
-DF <- DF %>% 
-  filter(`Authorship Survey` == "yes" | 
-           `Writing - original draft` == TRUE)
-
-nrow(DF)
-cat(print_yaml(DF))

### 2024-12-01 manuscript_draft.Rmd
updating manuscript draft

--- 2024-11-11 manuscript_draft.Rmd
@@
 This data represents a small scale dataset (i.e., the range of the scale of the data is small, 4 points) that could be used as pilot data for a study using concrete word ratings.
-The data is available at <https://osf.io/qpmf4/>.
+The data is available at <https://osf.io/qpmf4/> [see the following for thoughts on analyzing ordinal rating data: @bürkner2019; @liddell2018; @taylor2022; @kubinec2023].
 
@@
 
-For this example, imagine a researcher who wants to determine the differences in response latencies for abstract and concrete words [see the following for thoughts on analyzing this type of data: @bürkner2019; @liddell2018; @taylor2022].
+For this example, imagine a researcher who wants to determine the differences in response latencies for abstract and concrete words.
 They will select *n* = 40 words from the rating data from @brysbaert2014 that are split evenly into abstract and concrete ends of the rating scale.
@@
 
+```{r elp-estimate, echo = F, include = F}
+elp_small <- elpDF %>% filter(Stimulus %in% con_words)
+# elp_small$Stimulus <- droplevels(elp_small$Stimulus) already a character
+
+# figure out pilot size
+pilot_size_e <- round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)))
+
+# figure out data loss
+elp_small$RT[elp_small$Accuracy == "0"] <- NA
+data_loss_e <- elp_small %>% 
+ group_by(Stimulus, Type) %>% 
+ summarize(percent_correct = sum(!is.na(RT))/n(), .groups = "keep")
+elp_use <- elp_small %>% filter(!is.na(RT))
+
+# cutoff score 
+SE_e <- tapply(elp_use$RT, elp_use$Stimulus, function (x) { sd(x)/sqrt(length(x))})
+cutoff_e <- quantile(SE_e, probs = seq(from = .1, to = .9, by = .1))
+```
+
 *Step 1*.
@@
 In our sample of 40 words, the average pilot sample size was `r round(mean(tapply(con_small$Rating, con_small$Word, length)), digits = 2)` (*SD* = `r round(sd(tapply(con_small$Rating, con_small$Word, length)), digits = 2)`), and we will use `r pilot_size_c` as our pilot sample size for the concreteness ratings (this information will be used in the follow-up to the simulation study).
-
-```{r elp-estimate, echo = F, include = F}
-elp_small <- elpDF %>% filter(Stimulus %in% con_words)
-# elp_small$Stimulus <- droplevels(elp_small$Stimulus) already a character
-
-# figure out pilot size
-pilot_size_e <- round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)))
-
-# figure out data loss
-elp_small$RT[elp_small$Accuracy == "0"] <- NA
-data_loss_e <- elp_small %>% 
- group_by(Stimulus, Type) %>% 
- summarize(percent_correct = sum(!is.na(RT))/n(), .groups = "keep")
-elp_use <- elp_small %>% filter(!is.na(RT))
-
-# cutoff score 
-SE_e <- tapply(elp_use$RT, elp_use$Stimulus, function (x) { sd(x)/sqrt(length(x))})
-cutoff_e <- quantile(SE_e, probs = seq(from = .1, to = .9, by = .1))
-```
-
 We first selected the same real words in the ELP data as the concreteness subset selected above, and this data includes `r length(unique(elpDF$Stimulus))` real words.
@@
 
-## Special Consideration: Bimodal Data
+## Special Considerations
+
+### Bimodal Data
 
@@
 The number of bimodal items was varied from 0 percent to 100 percent increasing by 10 percent increments for simulated samples.
-Figure \@\ref(fig:bimodal-figure) shows an example of 10 of the items within a simulation that estimated that half of the Likert items would be bimodal in nature.
+Figure \@ref(fig:bimodal-figure) shows an example of 10 of the items within a simulation that estimated that half of the Likert items would be bimodal in nature.
 The researcher procedure described above was then carried out using *semanticprimeR*.
 
-```{r import-bimodal, echo = F}
+```{r import-bimodal, echo = F, warning = F, message = F}
 bi_DF <- import("simulation_study/simulation_bimodal/summary_bimodal.csv") %>% 
@@
 
-```{r bimodal-figure, fig.cap = "Simulated example of Likert data with half of the items as bimodal distributions.", fig.height=8, fig.width=8}
+```{r bimodal-figure, fig.cap = "Simulated example of Likert data with half of the items as bimodal distributions.", fig.height=10, fig.width=7, warning = F, include = F, message = F}
 ggplot(researcher_sample %>% 
-         filter(item < 11), aes(x = score, y = factor(item), fill = factor(item))) +
-  geom_density_ridges(alpha = 0.7, scale = 1) +
+         filter(item < 11), aes(x = score)) +
+  geom_histogram(aes(y = ..count..), bins = 7, position = "identity", alpha = 0.7, color = "black") +  # Stacked histograms
   labs(x = "Score", y = "Item") +
   theme_minimal() +
-  theme(legend.position = "none") + 
-  coord_cartesian(xlim = c(1,7))
-```
-
-Figure \@\ref(fig:bimodal-figure) portrays the results of the suggested procedure.
+  facet_wrap(~factor(item), scales = "free_y", nrow = 10) +  # Place each item in a separate row
+  theme(legend.position = "none",
+        axis.text.y = element_blank()) + 
+  coord_cartesian(xlim = c(1, 7))  # Set x-axis limits
+```
+
+Figure \@ref(fig:bimodal-figure) portrays the results of the suggested procedure.
 The sample with 0 bimodal items demonstrates the same values as above.
@@
   ylab("Corrected Sample Size") +
-  scale_color_viridis_d(option = "plasma") +  # Use the 'plasma' viridis option
+  scale_color_manual(values = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", 
+                                "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf", 
+                                "#9edae5")) +  # Custom high-contrast color palette
   scale_shape_manual(values = c(16, 17, 18, 19, 15, 3, 7, 8, 9, 13, 10)) +
@@
 ```
+
+### Combining Tools
+
+Researchers may often be interested in more than just the precision for individual items.
+The estimation of power via traditional power calculations for a statistical test or the reliability of items could also be of interest for an appropriate dataset for analyses.
+We would suggest that researchers combine reliability with precision to collect data that are both reliable and precisely measured.
+For example, one could use our proposed procedure to calculate the sample size for precision.
+Separately, the researcher would determine the level of reliability they would like to find in items based on previous research or practical guidelines.
+The minimum sample size could be set based on estimations for hypothesis testing, the stopping rule based on minimum reliability and desired SE for items, and a maximum sample size based on simulations or practical matters.
+Each data collection represents a unique scenario in which researchers can combine tools based on their needs to collect precise, reliable, and (traditionally defined) adequately powered data.
+While our proposal may bring to mind the problems with researcher degrees of freedom [@simmons2011], transparent practices decisions around sample size planning should be encouraged to limit potential questionable research practices.
 
@@
 
--   All data used in this manuscript and vignettes have been cited and can be found on our repository pages.
+- Data: All data used in this manuscript and vignettes have been cited and can be found on our repository pages (https://github.com/SemanticPriming/stimuli-power).
+- Analysis Code: 
   - Manuscript repository with code and data: <https://osf.io/swmva/> or <https://github.com/SemanticPriming/stimuli-power>
   - Package repository with vignettes and data: <https://github.com/SemanticPriming/semanticprimeR>
--   We did not pre-register this study, as it was a simulation study. No materials were used.
+- Pre-registration: We did not pre-register this study, as it was a simulation study.
+- Materials: No materials were used.
+- Funding: no funding was provided for this study. 
+- Conflicts of interest: the researchers declare no conflicts of interest.
 

### 2024-12-01 manuscript_draft.tex
updating manuscript draft

@@
   {\xapptocmd\section
-    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
+    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi: #1}}
     {}{\InnerPatchFailed}%
@@
 {}{\PatchFailed}
+\makeatother
 \keywords{accuracy in parameter estimation, power, sampling, simulation, hypothesis testing}
@@
 This data represents a small scale dataset (i.e., the range of the scale of the data is small, 4 points) that could be used as pilot data for a study using concrete word ratings.
-The data is available at \url{https://osf.io/qpmf4/}.
+The data is available at \url{https://osf.io/qpmf4/} (see the following for thoughts on analyzing ordinal rating data: Bürkner \& Vuorre, 2019; Kubinec, 2023; Liddell \& Kruschke, 2018; Taylor et al., 2022).
 
@@
 
-For this example, imagine a researcher who wants to determine the differences in response latencies for abstract and concrete words (see the following for thoughts on analyzing this type of data: Bürkner \& Vuorre, 2019; Liddell \& Kruschke, 2018; Taylor et al., 2022).
+For this example, imagine a researcher who wants to determine the differences in response latencies for abstract and concrete words.
 They will select \emph{n} = 40 words from the rating data from Brysbaert et al.
@@
 
-\subsection{Special Consideration: Bimodal Data}\label{special-consideration-bimodal-data}
+\subsection{Special Considerations}\label{special-considerations}
+
+\subsubsection{Bimodal Data}\label{bimodal-data}
 
@@
 
+\begin{verbatim}
+## Warning: The dot-dot notation (`..count..`) was deprecated in ggplot2 3.4.0.
+## i Please use `after_stat(count)` instead.
+## This warning is displayed once every 8 hours.
+## Call `lifecycle::last_lifecycle_warnings()` to see where this warning was
+## generated.
+\end{verbatim}
+
 \begin{figure}
@@
 \pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/bimodal-results-1.pdf}}
-\caption{\label{fig:bimodal-results}Estimated sample sizes for distributions that vary in the number of bimodal items. Note that percent below is treated as a continuous factor}
+\caption{\label{fig:bimodal-results}Estimated sample sizes for distributions that vary in the number of bimodal items.
+Note that percent below is treated as a continuous factor, and therefore, some estimates are the same (i.e., below 80 and 85 percent may be both estimated as 88\%).}
 \end{figure}
+
+\subsubsection{Combine Parameters}\label{combine-parameters}
+
+Researchers may often be interested in more than just the precision for individual items.
+The estimation of power via traditional power calculations for a statistical test or the reliability of items could also be of interest for an appropriate dataset for analyses.
+We would suggest that researchers combine reliability with precision to collect data that are both reliable and precisely measured.
+For example, one could use our proposed procedure to calculate the sample size for precision.
+Separately, the researcher would determine the level of reliability they would like to find in items based on previous research or practical guidelines.
+The minimum sample size could be set based on estimations for hypothesis testing, the stopping rule based on minimum reliability and desired SE for items, and a maximum sample size based on simulations or practical matters.
+Each data collection represents a unique scenario in which researchers can combine tools based on their needs to collect precise, reliable, and (traditionally defined) adequately powered data.
+While our proposal may bring to mind the problems with researcher degrees of freedom (Simmons et al., 2011), transparent practices decisions around sample size planning should be encouraged to limit potential questionable research practices.
 
@@
 \item
-  All data used in this manuscript and vignettes have been cited and can be found on our repository pages.
+  Data: All data used in this manuscript and vignettes have been cited and can be found on our repository pages (\url{https://github.com/SemanticPriming/stimuli-power}).
+\item
+  Scripts:
+
+  \begin{itemize}
+  \tightlist
   \item
@@
     Package repository with vignettes and data: \url{https://github.com/SemanticPriming/semanticprimeR}
+  \end{itemize}
 \item
-  We did not pre-register this study, as it was a simulation study. No materials were used.
+  Pre-registration: We did not pre-register this study, as it was a simulation study.
+\item
+  Materials: No materials were used.
+\item
+  Funding: no funding was provided for this study.
+\item
+  Conflicts of interest: the researchers declare no conflicts of interest.
 \end{itemize}
@@
 
+\bibitem[\citeproctext]{ref-kubinec2023}
+Kubinec, R.
+(2023).
+Ordered Beta Regression: A Parsimonious, Well-Fitting Model for Continuous Data with Lower and Upper Bounds.
+\emph{Political Analysis}, \emph{31}(4), 519--536.
+\url{https://doi.org/10.1017/pan.2022.20}
+
 \bibitem[\citeproctext]{ref-kumle2020}

### 2024-12-01 manuscript_draft.Rmd
fixing figure labels and referencees

@@
 
-1)  How does pilot data influence sample size suggestions?
+1)  How do pilot data influence sample size suggestions?
 
@@
 
-```{r decile50-figure, fig.cap = "Comparison of the cutoffs for 5th decile across heterogeneity (columns), powering of items (rows), and scale size (color).", fig.height=8, fig.width=8}
+```{r decile50-figure, fig.cap = "Comparison of the cutoffs for 5th deciles across heterogeneity (columns), powering of items (rows), and scale size (color).", fig.height=8, fig.width=8}
 summary_long %>% 
@@
              labeller = labeller(power = power.labs)) + 
-  ggtitle("Results for 5th decile") +
+  ggtitle("Results for 5th deciles") +
   xlab("\nPilot Sample Size") + 
@@
 
-```{r bimodal-figure, fig.cap = "Simulated example of Likert data with half of the items as bimodal distributions.", fig.height=10, fig.width=7, warning = F, include = F, message = F}
+```{r bimodal-figure, fig.cap = "Simulated example of Likert data with half of the items as bimodal distributions.", fig.height=10, fig.width=7, warning = F, message = F}
 ggplot(researcher_sample %>% 

### 2024-12-01 manuscript_draft.tex
fixing figure labels and referencees

@@
 In our sample of 40 words, the average pilot sample size was 28.52 (\emph{SD} = 1.80), and we will use 29 as our pilot sample size for the concreteness ratings (this information will be used in the follow-up to the simulation study).
-
 We first selected the same real words in the ELP data as the concreteness subset selected above, and this data includes 27031 real words.
@@
 \item
-  How does pilot data influence sample size suggestions?
+  How do pilot data influence sample size suggestions?
 \end{enumerate}
@@
 \pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/decile50-figure-1.pdf}}
-\caption{\label{fig:decile50-figure}Comparison of the cutoffs for 5th decile across heterogeneity (columns), powering of items (rows), and scale size (color).}
+\caption{\label{fig:decile50-figure}Comparison of the cutoffs for 5th deciles across heterogeneity (columns), powering of items (rows), and scale size (color).}
 \end{figure}
@@
 The number of bimodal items was varied from 0 percent to 100 percent increasing by 10 percent increments for simulated samples.
-Figure @\ref(fig:bimodal-figure) shows an example of 10 of the items within a simulation that estimated that half of the Likert items would be bimodal in nature.
+Figure \ref{fig:bimodal-figure} shows an example of 10 of the items within a simulation that estimated that half of the Likert items would be bimodal in nature.
 The researcher procedure described above was then carried out using \emph{semanticprimeR}.
-
-\begin{verbatim}
-## Warning: The dot-dot notation (`..count..`) was deprecated in ggplot2 3.4.0.
-## i Please use `after_stat(count)` instead.
-## This warning is displayed once every 8 hours.
-## Call `lifecycle::last_lifecycle_warnings()` to see where this warning was
-## generated.
-\end{verbatim}
 
@@
 
-Figure @\ref(fig:bimodal-figure) portrays the results of the suggested procedure.
+Figure \ref{fig:bimodal-figure} portrays the results of the suggested procedure.
 The sample with 0 bimodal items demonstrates the same values as above.
@@
 
-\subsubsection{Combine Parameters}\label{combine-parameters}
+\subsubsection{Combining Tools}\label{combining-tools}
 
@@
 These proposed methods and traditional power analysis could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
-We would advise caution when oversampling (i.e., simulation from data with larger sample sizes than the pilot data) for null-hypothesis power estimates, as these may be upwardly biased in small samples creating Type-I error (Burns et al., n.d.).
+We would advise caution when oversampling (i.e., simulation from data with larger sample sizes than the pilot data) for null-hypothesis power estimates, as these may be upwardly biased in small samples creating Type-I error (Burns et al., 2023).
 
@@
 \item
-  Scripts:
+  Analysis Code:
 
@@
 \bibitem[\citeproctext]{ref-burns}
-Burns, C. D. G., Fracasso, A., \& Rousselet, G. A. (n.d.). \emph{Data-driven estimates of the reproducibility of univariate BWAS are biased}. \url{https://doi.org/10.1101/2023.09.21.558661}
+Burns, C. D. G., Fracasso, A., \& Rousselet, G. A. (2023). \emph{Data-driven estimates of the reproducibility of univariate BWAS are biased}. \url{https://doi.org/10.1101/2023.09.21.558661}
 

### 2024-12-01 manuscript_draft.Rmd
fixing a few more typos

@@
 As discussed in Kelley and Maxwell [@kelley2003], one can estimate the sample size necessary to obtain precision in estimation of population parameters.
-Precision is defined as a researcher defined, sufficiently narrow, confidence interval.
+Precision is defined as a researcher defined, sufficiently narrow, confidence interval in AIPE.
 For example, researchers may wish to detect a specific mean in a study, *M* = .35.
-They could then use AIPE to estimate the sample size needed to find a sufficiently narrow confidence interval around that mean.
-Therefore, they could decide that sufficiently narrow could be defined as a confidence interval width of .30 or .15 on each side of the mean.
+They could then use AIPE to estimate the sample size needed to find a sufficiently narrow window around that mean.
+Therefore, they could decide that sufficiently narrow could be defined as a width of .30 or .15 on each side of the mean.
 They would then estimate the number of participants needed to find that level of precision.
@@
 
-Pilot or simulated data would be used to estimate the variability within items and select a sufficiently narrow window for overall item confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals).
+Pilot or simulated data would be used to estimate the variability within items and select a sufficiently narrow window for overall item SE for AIPE sufficiently narrow windows.
 The advantage to this method over simple power estimation from pilot effect sizes is the multiple simulations to average out potential variability, as well as a shift away from traditional NHST to parameter estimation.
@@
 
-Researchers could then use sequential testing to estimate their parameter of interest after each participant's data or at regular intervals during data collection to determine whether they have achieved their expected width of the confidence interval around that parameter.
 One would set a minimum sample size based on our procedure steps below to ensure an appropriate minimum for precise estimates.
-Then a researcher could use the standard error as a stopping rule (i.e., stop data collection when the standard error meets the expected sufficiently narrow width, as defined above).
+After meeting the minimum sample size, researchers could then use sequential testing to estimate their parameter of interest after each participant's data or at regular intervals during data collection to determine whether they have achieved their expected narrow window around that parameter.
 A stricter criterion could be defined for a stopping rule as a maximum sample size.
@@
 In this procedure, it is important to ensure that the data is representative of a larger population of sampled items that you intend to assess.
-Generally, pilot data sample sizes will be smaller than the overall intended project (e.g., 20 to 50), as the goal would be to determine how many participants would be necessary to reach a "stable" standard error for the accurately measured confidence interval rule.
+Generally, pilot data sample sizes will be smaller than the overall intended project (e.g., 20 to 50), as the goal would be to determine how many participants would be necessary to reach a stable standard error for the accurately measured narrow window rule.
 
@@
     "Updated Steps" = c("Use representative pilot data.",
-    "Calculate standard error of each of the items in the pilot data. Using the 40%, determine the cutoff and stopping rule for the standard error of the items.",
+    "Calculate standard error of each of the items in the pilot data. Using the 4th decile, determine the cutoff and stopping rule for the standard error of the items.",
     "Create simulated samples of your pilot data starting with at least 20 participants up to a maximum number of participants.", 
@@
 Table \@ref(tab:table-example) demonstrates the cutoff scores for deciles of the SEs for the concreteness ratings and lexical decision response latency items.
-A researcher could potentially pick any of these cutoffs or other percentage options not shown here (e.g., 35%).
+A researcher could potentially pick any of these cutoffs or other percentage options not shown here (e.g., 3.5th decile).
 We will use simulation to determine the suggestion that best captures the balance of adequately powering our sample and feasibility.
@@
 In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items.
-Assuming that each pilot sample represents a dataset that a researcher has collected (Step 1), the SEs for each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as SE functions as the main component of the formula for normal distribution confidence intervals.
+Assuming that each pilot sample represents a dataset that a researcher has collected (Step 1), the SEs for each item were calculated to mimic the AIPE procedure of finding a sufficiently narrow window.
 SEs were calculated at each decile of the items up to 90% (i.e., 0% smallest SE, 10% ..., 90% largest SE).
@@
 
-Next, we calculated the percentage of items that fell below the cutoff score, and therefore, would be considered "well-measured for each decile by sample (Step 4).
+Next, we calculated the percentage of items that fell below the cutoff score, and therefore, would be considered well-measured for each decile by sample (Step 4).
 From these data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90% and 95% of the items fall below the cutoff criterion (Step 5).
-These values were chosen as popular, yet arbitrary, measures of "power" in which one could determine the minimum suggested sample size (potentially 80% of the items) and the maximum suggested sample size (selected from a higher percentage, such as 90% or 95%).
+These values were chosen as popular, yet arbitrary, measures of power in which one could determine the minimum suggested sample size (potentially 80% of the items) and the maximum suggested sample size (selected from a higher percentage, such as 90% or 95%).
 
@@
 
-We proposed a method combining AIPE and Monte Carlo simulation to estimate a minimum and maximum sample size and to define a rule for stopping data collection based on narrow confidence intervals on a parameter of interest.
+We proposed a method combining AIPE and Monte Carlo simulation to estimate a minimum and maximum sample size and to define a rule for stopping data collection based on narrow windows on a parameter of interest.
 In addition, we also demonstrated its practical applications using real-world data.

### 2024-12-01 manuscript_draft.tex
fixing a few more typos

@@
 As discussed in Kelley and Maxwell (Kelley \& Maxwell, 2003), one can estimate the sample size necessary to obtain precision in estimation of population parameters.
-Precision is defined as a researcher defined, sufficiently narrow, confidence interval.
+Precision is defined as a researcher defined, sufficiently narrow, confidence interval in AIPE.
 For example, researchers may wish to detect a specific mean in a study, \emph{M} = .35.
-They could then use AIPE to estimate the sample size needed to find a sufficiently narrow confidence interval around that mean.
-Therefore, they could decide that sufficiently narrow could be defined as a confidence interval width of .30 or .15 on each side of the mean.
+They could then use AIPE to estimate the sample size needed to find a sufficiently narrow window around that mean.
+Therefore, they could decide that sufficiently narrow could be defined as a width of .30 or .15 on each side of the mean.
 They would then estimate the number of participants needed to find that level of precision.
@@
 In this procedure, it is important to ensure that the data is representative of a larger population of sampled items that you intend to assess.
-Generally, pilot data sample sizes will be smaller than the overall intended project (e.g., 20 to 50), as the goal would be to determine how many participants would be necessary to reach a ``stable'' standard error for the accurately measured confidence interval rule.
+Generally, pilot data sample sizes will be smaller than the overall intended project (e.g., 20 to 50), as the goal would be to determine how many participants would be necessary to reach a stable standard error for the accurately measured narrow window rule.
 \item
@@
 & Calculate standard error of each of the items in the pilot data.
-Using the 40\%, determine the cutoff and stopping rule for the standard error of the items.\\
+Using the 4th decile, determine the cutoff and stopping rule for the standard error of the items.\\
 3 & Create simulated samples of your pilot data starting with at least 20 participants up to a maximum number of participants.
@@
 Table \ref{tab:table-example} demonstrates the cutoff scores for deciles of the SEs for the concreteness ratings and lexical decision response latency items.
-A researcher could potentially pick any of these cutoffs or other percentage options not shown here (e.g., 35\%).
+A researcher could potentially pick any of these cutoffs or other percentage options not shown here (e.g., 3.5th decile).
 We will use simulation to determine the suggestion that best captures the balance of adequately powering our sample and feasibility.
@@
 In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items.
-Assuming that each pilot sample represents a dataset that a researcher has collected (Step 1), the SEs for each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as SE functions as the main component of the formula for normal distribution confidence intervals.
+Assuming that each pilot sample represents a dataset that a researcher has collected (Step 1), the SEs for each item were calculated to mimic the AIPE procedure of finding a sufficiently narrow window.
 SEs were calculated at each decile of the items up to 90\% (i.e., 0\% smallest SE, 10\% \ldots, 90\% largest SE).
@@
 
-Next, we calculated the percentage of items that fell below the cutoff score, and therefore, would be considered ``well-measured for each decile by sample (Step 4).
+Next, we calculated the percentage of items that fell below the cutoff score, and therefore, would be considered well-measured for each decile by sample (Step 4).
 From these data, we pinpoint the smallest suggested sample size at which 80\%, 85\%, 90\% and 95\% of the items fall below the cutoff criterion (Step 5).
-These values were chosen as popular, yet arbitrary, measures of''power'' in which one could determine the minimum suggested sample size (potentially 80\% of the items) and the maximum suggested sample size (selected from a higher percentage, such as 90\% or 95\%).
+These values were chosen as popular, yet arbitrary, measures of power in which one could determine the minimum suggested sample size (potentially 80\% of the items) and the maximum suggested sample size (selected from a higher percentage, such as 90\% or 95\%).
 

### 2024-12-03 manuscript_draft.tex
updating disclosures

@@
 \item
+  Funding: no funding was provided for this study.
+\item
+  Conflicts of interest: the researchers declare no conflicts of interest.
+\item
+  Ethics approval: not applicable, simulation study.
+\item
+  Consent to participate: not applicable, simulation study.
+\item
+  Consent for publication: all authors approved the manuscript, no participant approval necessary.
+\item
   Data: All data used in this manuscript and vignettes have been cited and can be found on our repository pages (\url{https://github.com/SemanticPriming/stimuli-power}).
+\item
+  Materials: No materials were used.
 \item
@@
 \item
-  Materials: No materials were used.
-\item
-  Funding: no funding was provided for this study.
-\item
-  Conflicts of interest: the researchers declare no conflicts of interest.
+  Author contributions are printed at the top of the manuscript automatically in \emph{papaja}.
 \end{itemize}

### 2025-08-14 simulation_study/simulate_summary_pilots.Rmd
adding new simulation

---
title: "Methods Power Study - Pilot Sample Size Simulation"
author: "Erin M. Buchanan"
date: "Last Knitted: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reviewer-requested simulation using semanticprimeR: Quantifies variability from pilot-based cutoffs and reports: (a) average final N and SE, (b) achieved power and SE

## Libraries

```{r}
# install.packages("remotes") 
# remotes::install_github("SemanticPriming/semanticprimeR")
library(semanticprimeR)
library(dplyr)
library(purrr)
library(tidyr)
library(broom)
library(ggplot2)

set.seed(20250813)
```

## Parameters

```{r}
# ones to change for small, medium, large
scale_min         <- 1
scale_max         <- 7
mu                <- 4
mu_sigma          <- .25
sigma             <- 2
sigma_sigma       <- .8
smallest_sigma    <- .08
digits            <- 0
labels            <- "small_large"

# small - small 1, 7, 4, .25, 2, .2, .02, 0, "small_small"
# small - medium 1, 7, 4, .25, 2, .4, .04, 0, "small_medium"
# small - large 1, 7, 4, .25, 2, .8, .08, 0, "small_large"

# ones to keep consistent 
n_items           <- 30
pop_scores_per_it <- 1000           # large to approximate "truth"
target_q_percent  <- 0.40           # 4th decile (fixed)
percent_levels    <- c(80,85,90,95) # targets in pipeline
n_pilot_reps      <- 100            # 100 researchers all doing the same thing
minN              <- 20
maxN              <- 300
stepN             <- 5
pilot_sizes       <- c(20, 25, 30, 35, 40)
```

## True Population

```{r}
# create a large population of items
pop <- simulate_population(
  mu = mu, 
  mu_sigma = mu_sigma,
  sigma = sigma, 
  sigma_sigma = sigma_sigma,
  number_items = n_items,
  number_scores = pop_scores_per_it,
  smallest_sigma = smallest_sigma,
  min_score = scale_min,
  max_score = scale_max,
  digits = digits
)

# get population SD 
pop_item_sigma <- pop %>%
  dplyr::group_by(item) %>%
  dplyr::summarise(sigma = sd(score), .groups = "drop")
```

## Pilot Helper Function

```{r}
# researcher running one pilot study --> uses variables from above 
run_one_pilot <- function(rep_id, pilot_size) {

  # Draw a pilot sample per item from the population
  pilot <- pop %>%
    group_by(item) %>%
    slice_sample(n = pilot_size, replace = FALSE) %>%
    ungroup()

  # Step 2: get pilot-based cutoff & prop_var
  cutoff <- calculate_cutoff(
    population     = pilot,
    grouping_items = "item",
    score          = "score",
    minimum        = scale_min,
    maximum        = scale_max
  )
  pilot_cut <- cutoff$cutoff
  prop_var  <- cutoff$prop_var

  # Steps 3–5: simulate from pilot; proportion below pilot_cut; correction t0 final N
  samp_list <- simulate_samples(
    start = minN, 
    stop = maxN, 
    increase = stepN,
    population = pilot, 
    replace = TRUE, 
    nsim = 500,
    grouping_items = "item"
  )

  prop_tbl <- calculate_proportion(
    samples = samp_list,
    cutoff = pilot_cut,
    grouping_items = "item",
    score = "score"
  )

  corrected <- calculate_correction(
    proportion_summary   = prop_tbl,
    pilot_sample_size    = pilot_size,
    proportion_variability = prop_var,
    power_levels         = percent_levels
  )

  # Map noisy percent_below to nearest intended target level
  final_Ns <- corrected %>%
    mutate(target_level = percent_levels[
      sapply(percent_below, function(x) which.min(abs(percent_levels - x)))
    ]) %>%
    group_by(target_level) %>%
    summarise(final_N = ceiling(min(corrected_sample_size, na.rm = TRUE)), .groups = "drop")

  # Evaluate precision achieved at final N (vs TRUE population threshold)
  out <- map_dfr(seq_len(nrow(final_Ns)), function(i) {

    # target_pct is just a record of what proportion we were trying for
    target_pct <- final_Ns$target_level[i]

    #	N_final is how many participants/items you’d actually collect in the main study according to the pilot’s advice.
    N_final    <- final_Ns$final_N[i]

    # pretend to actually collect data based on that final sample size 
    dat_final <- pop %>%
      group_by(item) %>% 
      slice_sample(n = N_final, replace = TRUE) %>% 
      ungroup()

    # Get the SE for each item in this simulated final dataset.
    item_SE_final <- dat_final %>%
      group_by(item) %>%
      summarise(SE = sd(score)/sqrt(n()), .groups = "drop")

    # create the SE from the population sigma, gold standard cutoff 
    true_SE_i_at_N      <- pop_item_sigma$sigma / sqrt(N_final)
    pop_threshold_at_N  <- stats::quantile(true_SE_i_at_N, 
                                           probs = target_q_percent, 
                                           names = FALSE)

    # align order just to be safe
    item_ids <- dplyr::arrange(item_SE_final, item)$item
    SE_obs   <- dplyr::arrange(item_SE_final, item)$SE
    SE_true  <- true_SE_i_at_N[order(dplyr::arrange(pop_item_sigma, item)$item)]

    # per-item errors
    se_err        <- SE_obs - SE_true
    se_rel_err    <- se_err / SE_true
    mae           <- mean(abs(se_err))
    rmse          <- sqrt(mean(se_err^2))
    bias          <- mean(se_err)
    mean_rel_err  <- mean(se_rel_err)
    cov_within_5p <- mean(abs(se_rel_err) <= 0.05)  # within ±5% of true SE
    cov_within_10p<- mean(abs(se_rel_err) <= 0.10)  # within ±10%

    # percent-below metrics (as before, now using threshold at this N)
    pct_below_pop   <- mean(SE_obs <= pop_threshold_at_N)
    pct_below_pilot <- mean(SE_obs <= pilot_cut)
    delta_pct_pop   <- pct_below_pop - target_q_percent  # calibration vs truth

    # cutoff error (pilot threshold vs true threshold at this N)
    cutoff_error      <- pilot_cut - pop_threshold_at_N
    cutoff_rel_error  <- cutoff_error / pop_threshold_at_N

    tibble::tibble(
      pilot_rep              = rep_id,
      pilot_size             = pilot_size,
      target_percent         = target_pct,
      final_N                = N_final,
      # thresholded metrics
      pct_items_below_pop    = pct_below_pop,
      pct_items_below_pilot  = pct_below_pilot,
      delta_pct_pop          = delta_pct_pop,
      # continuous accuracy of SEs
      se_bias                = bias,
      se_mae                 = mae,
      se_rmse                = rmse,
      se_mean_rel_err        = mean_rel_err,
      se_cov_within_5pct     = cov_within_5p,
      se_cov_within_10pct    = cov_within_10p,
      # cutoff accuracy
      cutoff_error           = cutoff_error,
      cutoff_rel_error       = cutoff_rel_error
    )
  })

  write.csv(out, paste0("simulation_pilot/pilot_sim", 
                        rep_id, "_", pilot_size, "_", labels, ".csv"), row.names = F)
  out
}
```

## Run Multiple Researcher's Pilot Studies

```{r}
all_results <- map_dfr(pilot_sizes, function(ps) {
  map_dfr(seq_len(n_pilot_reps), function(r) run_one_pilot(r, ps))
})
# beepr::beep()
write.csv(all_results, 
          paste0("simulation_pilot/pilot_simulation_", labels, ".csv"),
          row.names = F)
```

```{r}

# Per target & pilot size
summary_by_size_target <- all_results %>%
  group_by(pilot_size, target_percent) %>%
  summarise(
    avg_final_N           = mean(final_N, na.rm = TRUE),
    se_final_N            = sd(final_N,   na.rm = TRUE)/sqrt(sum(!is.na(final_N))),
    avg_pct_below_pop     = mean(pct_items_below_pop,   na.rm = TRUE),
    se_pct_below_pop      = sd(pct_items_below_pop,     na.rm = TRUE)/sqrt(sum(!is.na(pct_items_below_pop))),
    avg_pct_below_pilot   = mean(pct_items_below_pilot, na.rm = TRUE),
    se_pct_below_pilot    = sd(pct_items_below_pilot,   na.rm = TRUE)/sqrt(sum(!is.na(pct_items_below_pilot))),
    avg_delta_pct_pop     = mean(delta_pct_pop,         na.rm = TRUE),
    avg_se_bias           = mean(se_bias,               na.rm = TRUE),
    avg_se_mae            = mean(se_mae,                na.rm = TRUE),
    avg_se_rmse           = mean(se_rmse,               na.rm = TRUE),
    avg_se_mean_rel_err   = mean(se_mean_rel_err,       na.rm = TRUE),
    avg_cov_within_5pct   = mean(se_cov_within_5pct,    na.rm = TRUE),
    avg_cov_within_10pct  = mean(se_cov_within_10pct,   na.rm = TRUE),
    avg_cutoff_error      = mean(cutoff_error,      na.rm = TRUE),
    avg_cutoff_rel_error  = mean(cutoff_rel_error,  na.rm = TRUE),
    .groups = "drop"
  )

# Collapse to find a pilot-size lower bound (bias & variability vs truth)
# (Here: use mean absolute delta and its sd across targets)
summary_pilot_only <- all_results %>%
  group_by(pilot_size, pilot_rep) %>%
  summarise(abs_delta = mean(abs(delta_pct_pop), na.rm = TRUE), .groups = "drop") %>%
  group_by(pilot_size) %>%
  summarise(
    mean_abs_delta = mean(abs_delta, na.rm = TRUE),
    sd_abs_delta   = sd(abs_delta,   na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
# (A) Mean absolute deviation from population target by pilot size
ggplot(summary_pilot_only, aes(x = pilot_size, y = mean_abs_delta)) +
  geom_line() + geom_point() +
  labs(x = "Pilot size", y = "Mean |Δ % below population|",
       title = "Pilot size vs. deviation from population target")

# (B) For each target, show avg_delta vs pilot size (calibration)
ggplot(summary_by_size_target, aes(x = pilot_size, y = avg_delta_pct_pop, group = factor(target_percent))) +
  geom_line() + geom_point() +
  facet_wrap(~ target_percent, scales = "free_y") +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x = "Pilot size", y = "Avg (final % − population %)",
       title = "Calibration vs. population threshold by target")
```

1. Columns about sample size
	•	target_percent – The intended coverage or accuracy target for the stopping rule (e.g., stop when reaching 80%, 85%, etc.).
	•	avg_final_N – Average number of observations at stopping, across the 10 runs.
	•	se_final_N – Standard error of avg_final_N across runs (so here, how variable final N is between runs).

⸻

2. Proportion below population thresholds
	•	avg_pct_below_pop – Mean proportion of items whose SE (at the end of data collection) is below the population threshold.
	•	se_pct_below_pop – The standard error for that proportion across runs.
	•	avg_pct_below_pilot / se_pct_below_pilot – Same idea, but threshold is based on the pilot’s SE rather than the population.

⸻

3. Difference from population performance
	•	avg_delta_pct_pop – Mean difference between final proportion below threshold and the true (population) proportion below threshold.
	•	Positive = slightly better than the population, negative = slightly worse.

⸻

4. SE bias & error metrics
	•	avg_se_bias – Average signed difference between estimated SE and true SE.
	•	Positive → overestimation; negative → underestimation.
	•	avg_se_mae – Mean absolute error between estimated SE and true SE.
	•	avg_se_rmse – Root mean square error — punishes large deviations more.
	•	avg_se_mean_rel_err – Relative error: (SE_est - SE_true) / SE_true.

⸻

5. Coverage metrics
	•	avg_cov_within_5pct – Proportion of items where the estimated SE is within 5% of the true SE.
	•	avg_cov_within_10pct – Same, but within 10%.

⸻

6. Cutoff error
	•	avg_cutoff_error – Mean absolute difference between the actual cutoff used to stop and the ideal cutoff.
	•	avg_cutoff_rel_error – Same thing but relative to the ideal cutoff.

⸻

💡 Big picture:
In these 10 runs, your bias metrics are very close to zero, suggesting no consistent over/underestimation. But your avg_pct_below_pop values aren’t super high (most ~0.4), meaning fewer than half of items hit the population-based SE threshold before stopping — which matches your earlier suspicion. The coverage columns (avg_cov_within_Xpct) suggest your SE estimates are within 10% of the true SE for about 50–72% of items.

	•	Our diagnostics:
	•	avg_delta_pct_pop (calibration): how close the observed % of items below the population threshold is to the nominal target.
	•	mean_abs_delta (stability): smaller is better; plots show how it shrinks with pilot N.
	•	SE-error metrics (bias/MAE/RMSE, coverage within 5% and 10%) to show absolute accuracy.
	•	We define the minimum viable pilot N as the smallest pilot size where calibration is acceptably close to 0 (e.g., ±0.02) and variability is small (e.g., SD of |Δ| ≤ 0.03).

### 2025-08-22 manuscript_draft.tex
updating manuscript

@@
 AIPE shifts the focus away from finding a significant \emph{p}-value to finding a parameter that is accurately measured.
-As discussed in Kelley and Maxwell (Kelley \& Maxwell, 2003), one can estimate the sample size necessary to obtain precision in estimation of population parameters.
+As discussed in Kelley and Maxwell (2003), one can estimate the sample size necessary to obtain precision in estimation of population parameters.
 Precision is defined as a researcher defined, sufficiently narrow, confidence interval in AIPE.
@@
 They would then estimate the number of participants needed to find that level of precision.
-Confidence intervals are often misinterpreted (Hoekstra et al., 2014; but see Miller \& Ulrich, 2016), and AIPE procedures are not designed to specify sample size for a hypothesis-driven decision (i.e., the confidence interval does not include a specific value of comparison).
+
+Hoekstra et al.
+(2014) argued that confidence intervals are often misinterpreted (see Miller \& Ulrich, 2016 for critique of this claim; see Morey et al., 2016 for original authors' response), and AIPE procedures are not designed to specify sample size for a hypothesis-driven decision (i.e., the confidence interval does not include a specific value of comparison).
 Instead, AIPE focuses on estimating the sample size necessary to ensure precise population parameters.
@@
 Simulation is often paired with re-creating a data set with a similar structure for testing analyses and hypotheses based on proposed effect sizes or suggested population means.
-Generally, we would suggest starting with pilot data of a smaller sample size (e.g., 20 to 50) to understand the variability in potential items used to represent your phenomenon, especially if they are to be used in a larger study.
+Generally, we would suggest starting with pilot data of a smaller sample size (e.g., 20 to 50, see below for tests to determine the appropriate minimum) to understand the variability in potential items used to represent your phenomenon, especially if they are to be used in a larger study.
 However, given some background knowledge about the potential items, one could simulate example pilot data to use in a similar manner in our suggested procedure.
@@
 After meeting the minimum sample size, researchers could then use sequential testing to estimate their parameter of interest after each participant's data or at regular intervals during data collection to determine whether they have achieved their expected narrow window around that parameter.
-A stricter criterion could be defined for a stopping rule as a maximum sample size.
+A stricter criterion could be defined for a stopping rule (i.e., the criterion that specifies when data collection ends, such as reaching a maximum sample size or achieving a desired level of precision).
 By defining each of these components, researchers could ensure a feasible minimum sample size, a way to stop data collection when goals have been met, and a maximum sample size rule to ensure an actual end to data collection.
@@
 If researchers will show all items to participants, they could simply select one of the proposed estimates for their minimum required sample size.
-Therefore, we propose a method that leverages the ideas behind AIPE, paired with simulation and Monte Carlo simulation, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates.
+Therefore, we propose a method that leverages the ideas behind AIPE, paired with Monte Carlo simulation, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates.
+Sequential testing and sequential AIPE methods have a substantial literature of their own (Chow \& Chang, 2006; Kelley, 2007; Kelley et al., 2018; Siegmund, 1985; Wald, 1992), and our approach is complementary rather than a replacement for those frameworks.
 
@@
 
-\subsubsection{Calculate the Stopping Rule}\label{calculate-the-stopping-rule}
+\subsubsection{Define Pilot Data and Cutoff Criterion}\label{define-pilot-data-and-cutoff-criterion}
 
@@
 The simulations described in the Data Simulation section will explore what criterion should be used to determine the cutoff SE from the pilot data.
+Similar concepts appear in classical estimation work where lower bounds of population standard deviations are used as benchmarks (Chattopadhyay \& Banerjee, 2021; Mukhopadhyay, 1980).
 \end{enumerate}
@@
 
+\emph{Assumptions.} Our procedure does not assume normality of the data.
+To illustrate this, we simulated populations with normal, skewed, and bimodal distributions (in special considerations section).
+What the procedure requires is that item-level scores are sampled from a reasonably stable distribution with finite moments (i.e., mean and variance exist).
+The method evaluates variability in standard errors rather than relying on strict distributional forms.
+Thus, while we show results under both normal and non-normal populations, the logic of the procedure is distribution-agnostic provided these basic moment conditions are met.
+
 \subsection{Researcher Sample Simulation}\label{researcher-sample-simulation}
@@
 Last, we examined the question of an appropriate SE decile.
-First, the 0, 1st, and 2nd deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale size, and heterogeneity.
+The minimum and first two deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale size, and heterogeneity.
 If we examine the \(R^2\) values for each decile of our regression equation separately, we find that the values are all \(R^2\) \textgreater{} .99 with very little differences between them.
@@
 
+These corrected values represent the final recommended minimum sample sizes for achieving the specified precision thresholds (e.g., 90\%, 95\%).
+Researchers can use these numbers directly to determine the target N for their main study, and to define a feasible range of sample sizes.
+In practice, the corrected N can serve as the minimum sample size, while researchers may also set a maximum sample size based on available resources, and optionally apply sequential monitoring rules to stop data collection early if the desired precision is reached.
+These corrected estimates translate the pilot-based procedure into concrete targets that researchers can use when planning and monitoring data collection.
+
 \subsection{Special Considerations}\label{special-considerations}
 
+\subsubsection{Pilot Sample Size}\label{pilot-sample-size}
+
+Smaller pilot samples may be expected to show greater variability than larger samples or the full population.
+This variability could influence the sample size recommendations generated by our procedure.
+At the same time, the optimal size of the pilot sample was an open question in our framework.
+To address these questions, we ran simulations designed to identify the smallest pilot sample size that would minimize variability across researchers while also reducing bias when comparing pilot-derived estimates to population values.
+In these simulations, 100 researchers each drew a pilot sample from the same target population and applied our procedure.
+We generated nine populations (the combinations of scale size and heterogeneity described above), each with 30 items (\texttt{simulate\_population}).
+For each population, we varied the pilot sample size across 20, 25, 30, 35, and 40.
+Each simulated researcher then ran 500 replications, drawing between 20 and 300 participants in increments of 5 (\texttt{simulate\_samples}).
+
+We evaluated three statistics from these simulations:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Variability of the recommended sample size.
+For each target level, we calculated the average recommended \emph{N} and the standard deviation of that recommendation.
+Large standard deviations indicate instability, meaning different researchers could receive very different recommendations from similar pilot studies.
+\item
+  Bias in item standard errors.
+For each item, we compared the estimated \emph{SE} from the sample to the true \emph{SE} from the population.
+Smaller bias reflects closer agreement between pilot-based estimates and population values.
+Because larger sample sizes and heterogeneity simulations will naturally show larger SEs in general, we calculated relative error by subtracting estimated \emph{SE} minus the true \emph{SE} divided by the true \emph{SE}.
+\item
+  Cut-off bias.
+The cut-off was fixed at the 4th decile in the population.
+For each pilot, we calculated the difference between the pilot-derived cut-off and the population cut-off.
+Smaller values indicate that pilot-derived thresholds align more closely with the population criterion.
+Again, we used the relative error by dividing by the true \emph{SE} cut off score to be able to compare across simulations.
+\end{enumerate}
+
+Results consistently indicated that smaller pilots produced less stable and a bit more biased recommendations.
+As shown in Figure \ref{fig:fig-pilot-sd-n}, the standard deviation of the recommended sample size was higher at pilot sizes of 20 and 25, particularly in large-scale, high-variance populations.
+This instability means that two researchers running similar small pilots could easily obtain divergent recommended \emph{N}s.
+By contrast, pilot sizes of 30 or larger substantially reduced this variability, and further increases to 35 or 40 provided little additional gain.
+Bias measures told a similar story.
+Figure \ref{fig:fig-pilot-bias} shows that average item \emph{SE} bias was near zero across most conditions, but small pilots (20--25) tended to produce more negative bias in large-scale populations.
+At pilot sizes of 30 and above, \emph{SE} bias converged toward zero across all conditions.
+Likewise, Figure \ref{fig:fig-pilot-cutoff} shows that relative cut-off bias was largest when pilot sizes were small, again especially under high-variance conditions, but shrank considerably by pilot size 30 and remained stable thereafter.
+Accordingly, we propose \textasciitilde30 participants as a practical lower bound for pilot studies using our procedure: large enough to ensure stability, accuracy, and calibration across a wide range of scale and variance conditions, but small enough to remain efficient.
+
+\begin{figure}
+\centering
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/fig-pilot-sd-n-1.pdf}}
+\caption{\label{fig:fig-pilot-sd-n}Average standard deviation of suggested sample sizes for small, medium, and large scale sizes and heterogeneity.}
+\end{figure}
+
+\begin{figure}
+\centering
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/fig-pilot-bias-1.pdf}}
+\caption{\label{fig:fig-pilot-bias}Average relative standard error bias calculated for each simulation of size and heterogeneity combination.
+Relative error subtracts the pilot sample size SE for each item by the true SE for each item and then divides by the true SE for each item to normalize across scales.}
+\end{figure}
+
+\begin{figure}
+\centering
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_draft_files/figure-latex/fig-pilot-cutoff-1.pdf}}
+\caption{\label{fig:fig-pilot-cutoff}Average relative cut-off score bias for scale size and heterogeneity combinations.
+Relative cut-off scores indicate the true population 4th decile SE compared to the pilot sample SE cutoff at the 4th decile, divided by the population SE.}
+\end{figure}
+
 \subsubsection{Bimodal Data}\label{bimodal-data}
@@
 The above simulations assume symmetric population distributions with small, medium, or large heterogeneity or a skewed distribution with the same heterogeneity.
-However, as shown in Pollock (2018), sample rating data that appears to have a mean distribution may represent the average two separate distributions (i.e., a bimodal distribution).
+However, as shown in @pollock2018, rating data that appears normally distributed at the aggregate level may in fact reflect the average of two distinct underlying distributions (i.e., a bimodal distribution).
+In such cases, a pilot sample may misleadingly suggest a single ``middle'' value that does not represent either true subgroup.
 To examine our procedure with bimodal distributions, we estimated thirty items for three distributions: symmetric, ceiling, and floor distributions for Likert data with small heterogeneity.
@@
 These proposed methods and traditional power analysis could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
-We would advise caution when oversampling (i.e., simulation from data with larger sample sizes than the pilot data) for null-hypothesis power estimates, as these may be upwardly biased in small samples creating Type-I error (Burns et al., 2023).
+We would advise caution when resampling pilot data (e.g., oversampling beyond the available pilot sample) for power estimation.
+As Burns et al.
+(2023) note, resampling introduces estimation bias that increases as simulated sample sizes approach the full pilot size.
+This bias can occur not only for hypothesis tests but also when estimating descriptive statistics such as means, and it may distort power estimates, particularly in small samples.
 
@@
 Champely, S., Ekstrom, C., Dalgaard, P., Gill, J., Weibelzahl, S., Anandkumar, A., Ford, C., Volcic, R., \& De Rosario, H. (2017). \emph{Pwr: Basic functions for power analysis}.
+
+\bibitem[\citeproctext]{ref-chattopadhyay2021}
+Chattopadhyay, B., \& Banerjee, S.
+(2021).
+Estimation of location parameter within pre-specified error bound with second-order efficient two-stage procedure.
+\emph{South African Statistical Journal}, \emph{55}(1), 45--54.
+\url{https://doi.org/10.37920/sasj.2021.55.1.4}
+
+\bibitem[\citeproctext]{ref-chow2006}
+Chow, S.-C., \& Chang, M. (2006). \emph{Adaptive Design Methods in Clinical Trials} (0th ed.). Chapman; Hall/CRC. \url{https://doi.org/10.1201/9781584887775}
 
@@
 
+\bibitem[\citeproctext]{ref-kelley2018}
+Kelley, K., Darku, F.
+B., \& Chattopadhyay, B.
+(2018).
+Accuracy in parameter estimation for a general class of effect sizes: A sequential approach.
+\emph{Psychological Methods}, \emph{23}(2), 226--243.
+\url{https://doi.org/10.1037/met0000127}
+
 \bibitem[\citeproctext]{ref-kelley2003}
@@
 \url{https://doi.org/10.17605/OSF.IO/3TRG2}
+
+\bibitem[\citeproctext]{ref-morey2016}
+Morey, R.
+D., Hoekstra, R., Rouder, J.
+N., \& Wagenmakers, E.-J.
+(2016).
+Continued misinterpretation of confidence intervals: response to Miller and Ulrich.
+\emph{Psychonomic Bulletin \& Review}, \emph{23}(1), 131--140.
+\url{https://doi.org/10.3758/s13423-015-0955-8}
+
+\bibitem[\citeproctext]{ref-mukhopadhyay1980}
+Mukhopadhyay, N.
+(1980).
+A consistent and asymptotically efficient two-stage procedure to construct fixed width confidence intervals for the mean.
+\emph{Metrika}, \emph{27}(1), 281--284.
+\url{https://doi.org/10.1007/BF01893607}
 
@@
 
-\bibitem[\citeproctext]{ref-pollock2018}
-Pollock, L.
-(2018).
-Statistical and methodological problems with concreteness and other semantic variables: A list memory experiment case study.
-\emph{Behavior Research Methods}, \emph{50}(3), 1198--1216.
-\url{https://doi.org/10.3758/s13428-017-0938-y}
-
 \bibitem[\citeproctext]{ref-pownall2023}
@@
 \url{https://doi.org/10.31234/osf.io/h8ft7}
+
+\bibitem[\citeproctext]{ref-siegmund1985}
+Siegmund, D. (1985). \emph{Sequential Analysis: Tests and Confidence Intervals}. Springer New York.
 
@@
 
+\bibitem[\citeproctext]{ref-wald1992}
+Wald, A. (1992). \emph{Sequential tests of statistical hypotheses} (S. Kotz \& N. L. Johnson, Eds.; pp. 256--298). Springer New York. \url{http://link.springer.com/10.1007/978-1-4612-0919-5_18}
+
 \end{CSLReferences}

### 2025-08-22 simulation_study/simulate_summary_pilots.Rmd
updating manuscript

@@
 library(ggplot2)
-
+library(stringr)
 set.seed(20250813)
@@
 
-```{r}
+```{r eval = F}
 # ones to change for small, medium, large
-scale_min         <- 1
-scale_max         <- 7
-mu                <- 4
-mu_sigma          <- .25
-sigma             <- 2
-sigma_sigma       <- .8
-smallest_sigma    <- .08
+scale_min         <- 0
+scale_max         <- 3000
+mu                <- 1000
+mu_sigma          <- 150
+sigma             <- 400
+sigma_sigma       <- 200
+smallest_sigma    <- 20
 digits            <- 0
-labels            <- "small_large"
+labels            <- "large_large"
 
-# small - small 1, 7, 4, .25, 2, .2, .02, 0, "small_small"
-# small - medium 1, 7, 4, .25, 2, .4, .04, 0, "small_medium"
-# small - large 1, 7, 4, .25, 2, .8, .08, 0, "small_large"
+# small - small 1, 7, 4, .25, 2, .2, .02, 0, "small_small" done
+# small - medium 1, 7, 4, .25, 2, .4, .04, 0, "small_medium" done
+# small - large 1, 7, 4, .25, 2, .8, .08, 0, "small_large" done
+# medium - small 0, 100, 50, 10, 25, 4, .4, 0, "medium_small" done
+# medium - medium 0, 100, 50, 10, 25, 8, .8, 0, "medium_medium" done
+# medium - large 0, 100, 50, 10, 25, 16, 1.6, 0, "medium_large" done
+# large - small 0, 3000, 1000, 150, 400, 50, 5, 0, "large_small" done
+# large - medium 0, 3000, 1000, 150, 400, 100, 10, "large_medium" done
+# large - large 0, 3000, 1000, 150, 400, 200, 20, "large_large" done
 
@@
 
-```{r}
+```{r eval = F}
 # create a large population of items
@@
 
-```{r}
+```{r eval = F}
 # researcher running one pilot study --> uses variables from above 
@@
 
-```{r}
+```{r eval = F}
 all_results <- map_dfr(pilot_sizes, function(ps) {
@@
 
+## Create Overall File
+
 ```{r}
-
-# Per target & pilot size
-summary_by_size_target <- all_results %>%
-  group_by(pilot_size, target_percent) %>%
-  summarise(
-    avg_final_N           = mean(final_N, na.rm = TRUE),
-    se_final_N            = sd(final_N,   na.rm = TRUE)/sqrt(sum(!is.na(final_N))),
-    avg_pct_below_pop     = mean(pct_items_below_pop,   na.rm = TRUE),
-    se_pct_below_pop      = sd(pct_items_below_pop,     na.rm = TRUE)/sqrt(sum(!is.na(pct_items_below_pop))),
-    avg_pct_below_pilot   = mean(pct_items_below_pilot, na.rm = TRUE),
-    se_pct_below_pilot    = sd(pct_items_below_pilot,   na.rm = TRUE)/sqrt(sum(!is.na(pct_items_below_pilot))),
-    avg_delta_pct_pop     = mean(delta_pct_pop,         na.rm = TRUE),
-    avg_se_bias           = mean(se_bias,               na.rm = TRUE),
-    avg_se_mae            = mean(se_mae,                na.rm = TRUE),
-    avg_se_rmse           = mean(se_rmse,               na.rm = TRUE),
-    avg_se_mean_rel_err   = mean(se_mean_rel_err,       na.rm = TRUE),
-    avg_cov_within_5pct   = mean(se_cov_within_5pct,    na.rm = TRUE),
-    avg_cov_within_10pct  = mean(se_cov_within_10pct,   na.rm = TRUE),
-    avg_cutoff_error      = mean(cutoff_error,      na.rm = TRUE),
-    avg_cutoff_rel_error  = mean(cutoff_rel_error,  na.rm = TRUE),
-    .groups = "drop"
+file.list <- list.files(
+  path = "simulation_pilot",
+  pattern = "pilot_sim.*\\.csv$",
+  full.names = TRUE
 )
 
-# Collapse to find a pilot-size lower bound (bias & variability vs truth)
-# (Here: use mean absolute delta and its sd across targets)
-summary_pilot_only <- all_results %>%
-  group_by(pilot_size, pilot_rep) %>%
-  summarise(abs_delta = mean(abs(delta_pct_pop), na.rm = TRUE), .groups = "drop") %>%
-  group_by(pilot_size) %>%
-  summarise(
-    mean_abs_delta = mean(abs_delta, na.rm = TRUE),
-    sd_abs_delta   = sd(abs_delta,   na.rm = TRUE),
-    .groups = "drop"
+files.df <- lapply(file.list, function(f) {
+
+  # Extract just the filename
+  fname <- basename(f)
+
+  # Remove extension
+  fname_noext <- str_remove(fname, "\\.csv$")
+
+  # Get the part after the last underscore (e.g., "small_large")
+  suffix <- str_extract(fname_noext, "[^_]+_[^_]+$")
+
+  # Split into two parts
+  parts <- str_split(suffix, "_", simplify = TRUE)
+
+  read.csv(f) %>%
+    mutate(
+      source_file = fname,
+      scale_size = parts[1],
+      variance_size   = parts[2]
     )
+
+}) %>%
+  bind_rows()
+
+counts <- files.df %>% 
+  select(scale_size, variance_size, pilot_size, pilot_rep) %>% 
+  unique() %>% 
+  group_by(scale_size, variance_size, pilot_size) %>% 
+  count()
+
+# View(counts)
+write.csv(files.df, "simulation_pilot/pilot_simulation_final.csv", row.names = F)
 ```
-
-```{r}
-# (A) Mean absolute deviation from population target by pilot size
-ggplot(summary_pilot_only, aes(x = pilot_size, y = mean_abs_delta)) +
-  geom_line() + geom_point() +
-  labs(x = "Pilot size", y = "Mean |Δ % below population|",
-       title = "Pilot size vs. deviation from population target")
-
-# (B) For each target, show avg_delta vs pilot size (calibration)
-ggplot(summary_by_size_target, aes(x = pilot_size, y = avg_delta_pct_pop, group = factor(target_percent))) +
-  geom_line() + geom_point() +
-  facet_wrap(~ target_percent, scales = "free_y") +
-  geom_hline(yintercept = 0, linetype = 2) +
-  labs(x = "Pilot size", y = "Avg (final % − population %)",
-       title = "Calibration vs. population threshold by target")
-```
-
-1. Columns about sample size
-	•	target_percent – The intended coverage or accuracy target for the stopping rule (e.g., stop when reaching 80%, 85%, etc.).
-	•	avg_final_N – Average number of observations at stopping, across the 10 runs.
-	•	se_final_N – Standard error of avg_final_N across runs (so here, how variable final N is between runs).
-
-⸻
-
-2. Proportion below population thresholds
-	•	avg_pct_below_pop – Mean proportion of items whose SE (at the end of data collection) is below the population threshold.
-	•	se_pct_below_pop – The standard error for that proportion across runs.
-	•	avg_pct_below_pilot / se_pct_below_pilot – Same idea, but threshold is based on the pilot’s SE rather than the population.
-
-⸻
-
-3. Difference from population performance
-	•	avg_delta_pct_pop – Mean difference between final proportion below threshold and the true (population) proportion below threshold.
-	•	Positive = slightly better than the population, negative = slightly worse.
-
-⸻
-
-4. SE bias & error metrics
-	•	avg_se_bias – Average signed difference between estimated SE and true SE.
-	•	Positive → overestimation; negative → underestimation.
-	•	avg_se_mae – Mean absolute error between estimated SE and true SE.
-	•	avg_se_rmse – Root mean square error — punishes large deviations more.
-	•	avg_se_mean_rel_err – Relative error: (SE_est - SE_true) / SE_true.
-
-⸻
-
-5. Coverage metrics
-	•	avg_cov_within_5pct – Proportion of items where the estimated SE is within 5% of the true SE.
-	•	avg_cov_within_10pct – Same, but within 10%.
-
-⸻
-
-6. Cutoff error
-	•	avg_cutoff_error – Mean absolute difference between the actual cutoff used to stop and the ideal cutoff.
-	•	avg_cutoff_rel_error – Same thing but relative to the ideal cutoff.
-
-⸻
-
-💡 Big picture:
-In these 10 runs, your bias metrics are very close to zero, suggesting no consistent over/underestimation.
-But your avg_pct_below_pop values aren’t super high (most ~0.4), meaning fewer than half of items hit the population-based SE threshold before stopping — which matches your earlier suspicion.
-The coverage columns (avg_cov_within_Xpct) suggest your SE estimates are within 10% of the true SE for about 50–72% of items.
-
-	•	Our diagnostics:
-	•	avg_delta_pct_pop (calibration): how close the observed % of items below the population threshold is to the nominal target.
-	•	mean_abs_delta (stability): smaller is better; plots show how it shrinks with pilot N.
-	•	SE-error metrics (bias/MAE/RMSE, coverage within 5% and 10%) to show absolute accuracy.
-	•	We define the minimum viable pilot N as the smallest pilot size where calibration is acceptably close to 0 (e.g., ±0.02) and variability is small (e.g., SD of |Δ| ≤ 0.03).

### 2025-08-22 manuscript_draft.Rmd
updating disclosures
update manuscript RMD

@@
 # initial set up 
+source("~/GitHub/files/google_creds.R")
 # upload_file(file = "manuscript_draft.Rmd",
-#       gfile = "power2stim_V2_trackdown",
-#       gpath = "Studies/PSA007 SPAM-L/Manuscripts/Power",
-#       shared_drive = "Psychological Science Accelerator")
+#       gfile = "power2stim_V3_trackdown")
 # from google 
 download_file(file = "manuscript_draft.Rmd", 
-      gfile = "power2stim_V2_trackdown",
+      gfile = "power2stim_V3_trackdown",
       gpath = "trackdown")
@@
 update_file(file = "manuscript_draft.Rmd", 
-      gfile = "power2stim_V2_trackdown",
+      gfile = "power2stim_V3_trackdown",
       gpath = "trackdown",
@@
 AIPE shifts the focus away from finding a significant *p*-value to finding a parameter that is accurately measured.
-As discussed in Kelley and Maxwell [@kelley2003], one can estimate the sample size necessary to obtain precision in estimation of population parameters.
+As discussed in Kelley and Maxwell [-@kelley2003], one can estimate the sample size necessary to obtain precision in estimation of population parameters.
 Precision is defined as a researcher defined, sufficiently narrow, confidence interval in AIPE.
@@
 They would then estimate the number of participants needed to find that level of precision.
-Confidence intervals are often misinterpreted [@hoekstra2014; but see @miller2016], and AIPE procedures are not designed to specify sample size for a hypothesis-driven decision (i.e., the confidence interval does not include a specific value of comparison).
+
+@hoekstra2014 argued that confidence intervals are often misinterpreted [see @miller2016 for critique of this claim; see @morey2016 for original authors’ response], and AIPE procedures are not designed to specify sample size for a hypothesis-driven decision (i.e., the confidence interval does not include a specific value of comparison).
 Instead, AIPE focuses on estimating the sample size necessary to ensure precise population parameters.
@@
 If researchers will show all items to participants, they could simply select one of the proposed estimates for their minimum required sample size.
-Therefore, we propose a method that leverages the ideas behind AIPE, paired with simulation and Monte Carlo simulation, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates.
+Therefore, we propose a method that leverages the ideas behind AIPE, paired with Monte Carlo simulation, to estimate the minimum and maximum proposed sample sizes and stopping rules for studies that use multiple items with expected variability in their estimates.
+Sequential testing and sequential AIPE methods have a substantial literature of their own [@wald1992; @kelley2007; @kelley2018; @siegmund1985; @chow2006], and our approach is complementary rather than a replacement for those frameworks.
 
@@
 
-### Calculate the Stopping Rule
+### Define Pilot Data and Cutoff Criterion
 
@@
 The simulations described in the Data Simulation section will explore what criterion should be used to determine the cutoff SE from the pilot data.
+Similar concepts appear in classical estimation work where lower bounds of population standard deviations are used as benchmarks [@mukhopadhyay1980; @chattopadhyay2021].
 
@@
 
+*Assumptions.* Our procedure does not assume normality of the data.
+To illustrate this, we simulated populations with normal, skewed, and bimodal distributions (in special considerations section).
+What the procedure requires is that item-level scores are sampled from a reasonably stable distribution with finite moments (i.e., mean and variance exist).
+The method evaluates variability in standard errors rather than relying on strict distributional forms.
+Thus, while we show results under both normal and non-normal populations, the logic of the procedure is distribution-agnostic provided these basic moment conditions are met.
+
 ## Researcher Sample Simulation
@@
 Last, we examined the question of an appropriate SE decile.
-First, the 0, 1st, and 2nd deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale size, and heterogeneity.
+The minimum and first two deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale size, and heterogeneity.
 If we examine the $R^2$ values for each decile of our regression equation separately, we find that the values are all $R^2$ \> .99 with very little differences between them.
@@
   geom_line() + 
-  theme_classic(base_size = 10) + 
+  theme_linedraw(base_size = 10) + 
   facet_grid(power~ variability,
@@
   geom_line() +
-  theme_classic(base_size = 10) + 
+  theme_linedraw(base_size = 10) + 
   facet_grid(power~ variability,
@@
 
+These corrected values represent the final recommended minimum sample sizes for achieving the specified precision thresholds (e.g., 90%, 95%).
+Researchers can use these numbers directly to determine the target N for their main study, and to define a feasible range of sample sizes.
+In practice, the corrected N can serve as the minimum sample size, while researchers may also set a maximum sample size based on available resources, and optionally apply sequential monitoring rules to stop data collection early if the desired precision is reached.
+These corrected estimates translate the pilot-based procedure into concrete targets that researchers can use when planning and monitoring data collection.
+
 ## Special Considerations
 
+### Pilot Sample Size
+
+```{r pilot-sim-data}
+ps_DF <- import("simulation_study/simulation_pilot/pilot_simulation_final.csv") %>% 
+  select(-source_file) %>% 
+  group_by(pilot_size, target_percent, scale_size, variance_size) %>% 
+summarise(
+    avg_final_N = mean(final_N, na.rm = TRUE),
+    sd_final_N = sd(final_N,   na.rm = TRUE),
+    se_final_N = sd(final_N,   na.rm = TRUE)/sqrt(sum(!is.na(final_N))),
+    avg_pct_below_pop = mean(pct_items_below_pop, na.rm = TRUE),
+    se_pct_below_pop = sd(pct_items_below_pop, na.rm = TRUE)/sqrt(sum(!is.na(pct_items_below_pop))),
+    avg_pct_below_pilot = mean(pct_items_below_pilot, na.rm = TRUE),
+    se_pct_below_pilot = sd(pct_items_below_pilot, na.rm = TRUE)/sqrt(sum(!is.na(pct_items_below_pilot))),
+    avg_delta_pct_pop = mean(delta_pct_pop, na.rm = TRUE),
+    avg_se_bias = mean(se_bias, na.rm = TRUE),
+    avg_se_mae = mean(se_mae, na.rm = TRUE),
+    avg_se_rmse = mean(se_rmse, na.rm = TRUE),
+    avg_se_mean_rel_err = mean(se_mean_rel_err, na.rm = TRUE),
+    avg_cov_within_5pct = mean(se_cov_within_5pct, na.rm = TRUE),
+    avg_cov_within_10pct = mean(se_cov_within_10pct, na.rm = TRUE),
+    avg_cutoff_error = mean(cutoff_error, na.rm = TRUE),
+    avg_cutoff_rel_error = mean(cutoff_rel_error, na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  mutate(scale_size = factor(scale_size, 
+                             levels = c("small", "medium", "large"),
+                             labels = c("Likert", "Percent", "Milliseconds")),
+         variance_size = factor(variance_size, 
+                             levels = c("small", "medium", "large"),
+                             labels = c("Small Heterogeneity", "Medium Heterogeneity", "Large Heterogeneity")))
+```
+
+Smaller pilot samples may be expected to show greater variability than larger samples or the full population.
+This variability could influence the sample size recommendations generated by our procedure.
+At the same time, the optimal size of the pilot sample was an open question in our framework.
+To address these questions, we ran simulations designed to identify the smallest pilot sample size that would minimize variability across researchers while also reducing bias when comparing pilot-derived estimates to population values.
+In these simulations, 100 researchers each drew a pilot sample from the same target population and applied our procedure.
+We generated nine populations (the combinations of scale size and heterogeneity described above), each with 30 items (`simulate_population`).
+For each population, we varied the pilot sample size across 20, 25, 30, 35, and 40.
+Each simulated researcher then ran 500 replications, drawing between 20 and 300 participants in increments of 5 (`simulate_samples`).
+
+We evaluated three statistics from these simulations:
+
+1.
+Variability of the recommended sample size.
+For each target level, we calculated the average recommended *N* and the standard deviation of that recommendation.
+Large standard deviations indicate instability, meaning different researchers could receive very different recommendations from similar pilot studies.
+2.
+Bias in item standard errors.
+For each item, we compared the estimated *SE* from the sample to the true *SE* from the population.
+Smaller bias reflects closer agreement between pilot-based estimates and population values.
+Because larger sample sizes and heterogeneity simulations will naturally show larger SEs in general, we calculated relative error by subtracting estimated *SE* minus the true *SE* divided by the true *SE*.
+3.
+Cut-off bias.
+The cut-off was fixed at the 4th decile in the population.
+For each pilot, we calculated the difference between the pilot-derived cut-off and the population cut-off.
+Smaller values indicate that pilot-derived thresholds align more closely with the population criterion.
+Again, we used the relative error by dividing by the true *SE* cut off score to be able to compare across simulations.
+
+Results consistently indicated that smaller pilots produced less stable and a bit more biased recommendations.
+As shown in Figure \@ref(fig:fig-pilot-sd-n), the standard deviation of the recommended sample size was higher at pilot sizes of 20 and 25, particularly in large-scale, high-variance populations.
+This instability means that two researchers running similar small pilots could easily obtain divergent recommended *N*s.
+By contrast, pilot sizes of 30 or larger substantially reduced this variability, and further increases to 35 or 40 provided little additional gain.
+Bias measures told a similar story.
+Figure \@ref(fig:fig-pilot-bias) shows that average item *SE* bias was near zero across most conditions, but small pilots (20–25) tended to produce more negative bias in large-scale populations.
+At pilot sizes of 30 and above, *SE* bias converged toward zero across all conditions.
+Likewise, Figure \@ref(fig:fig-pilot-cutoff) shows that relative cut-off bias was largest when pilot sizes were small, again especially under high-variance conditions, but shrank considerably by pilot size 30 and remained stable thereafter.
+Accordingly, we propose \~30 participants as a practical lower bound for pilot studies using our procedure: large enough to ensure stability, accuracy, and calibration across a wide range of scale and variance conditions, but small enough to remain efficient.
+
+```{r fig-pilot-sd-n, fig.cap = "Average standard deviation of suggested sample sizes for small, medium, and large scale sizes and heterogeneity.", fig.height=6, fig.width=8, warning = F, message = F}
+ggplot(ps_DF %>% 
+         filter(target_percent == 80), aes(x = pilot_size, y = sd_final_N,
+                  color = variance_size, group = variance_size)) +
+  geom_point() +
+  #geom_line() +
+  # geom_errorbar(aes(ymin = avg_final_N - se_final_N,
+  #                   ymax = avg_final_N + se_final_N), width = 0.2) +
+  facet_wrap(~ scale_size) +
+  labs(x = "Pilot Sample Size", y = "Average SD of N") +
+  theme_linedraw() + 
+  theme_bw() + 
+  theme(legend.position = "bottom") + 
+  scale_color_manual(name = NULL, values = c(plot_colors)) 
+```
+
+```{r fig-pilot-bias, fig.cap = "Average relative standard error bias calculated for each simulation of size and heterogeneity combination.
+Relative error subtracts the pilot sample size SE for each item by the true SE for each item and then divides by the true SE for each item to normalize across scales.
+", fig.height=6, fig.width=8, warning = F, message = F}
+ggplot(ps_DF %>% 
+         filter(target_percent == 80), aes(x = pilot_size, y = avg_se_mean_rel_err,
+                  color = variance_size, group = variance_size)) +
+  geom_point() +
+  #geom_line() +
+  # geom_errorbar(aes(ymin = avg_final_N - se_final_N,
+  #                   ymax = avg_final_N + se_final_N), width = 0.2) +
+  facet_wrap(~ scale_size) +
+  labs(x = "Pilot Sample Size", y = "Average Relative SE Bias") +
+  theme_linedraw() + 
+  theme_bw() + 
+  theme(legend.position = "bottom") + 
+  scale_color_manual(name = NULL, values = c(plot_colors)) 
+```
+
+```{r fig-pilot-cutoff, fig.cap = "Average relative cut-off score bias for scale size and heterogeneity combinations.
+Relative cut-off scores indicate the true population 4th decile SE compared to the pilot sample SE cutoff at the 4th decile, divided by the population SE.
+", fig.height=6, fig.width=8, warning = F, message = F}
+ggplot(ps_DF %>% 
+         filter(target_percent == 80), aes(x = pilot_size, y = avg_cutoff_rel_error,
+                  color = variance_size, group = variance_size)) +
+  geom_point() +
+  #geom_line() +
+  # geom_errorbar(aes(ymin = avg_final_N - se_final_N,
+  #                   ymax = avg_final_N + se_final_N), width = 0.2) +
+  facet_wrap(~ scale_size) +
+  labs(x = "Pilot Sample Size", y = "Average Relative Cut-Off Bias") +
+  theme_linedraw() + 
+  theme_bw() + 
+  theme(legend.position = "bottom") + 
+  scale_color_manual(name = NULL, values = c(plot_colors)) 
+```
+
 ### Bimodal Data
@@
 The above simulations assume symmetric population distributions with small, medium, or large heterogeneity or a skewed distribution with the same heterogeneity.
-However, as shown in @pollock2018, sample rating data that appears to have a mean distribution may represent the average two separate distributions (i.e., a bimodal distribution).
+However, as shown in @pollock2018, rating data that appears normally distributed at the aggregate level may in fact reflect the average of two distinct underlying distributions (i.e., a bimodal distribution).
+In such cases, a pilot sample may misleadingly suggest a single “middle” value that does not represent either true subgroup.
 To examine our procedure with bimodal distributions, we estimated thirty items for three distributions: symmetric, ceiling, and floor distributions for Likert data with small heterogeneity.
@@
 These proposed methods and traditional power analysis could be used together to strengthen our understanding of the sample size necessary for both a hypothesis test and a well-tuned estimation.
-We would advise caution when oversampling (i.e., simulation from data with larger sample sizes than the pilot data) for null-hypothesis power estimates, as these may be upwardly biased in small samples creating Type-I error [@burns].
+We would advise caution when resampling pilot data (e.g., oversampling beyond the available pilot sample) for power estimation.
+As @burns note, resampling introduces estimation bias that increases as simulated sample sizes approach the full pilot size.
+This bias can occur not only for hypothesis tests but also when estimating descriptive statistics such as means, and it may distort power estimates, particularly in small samples.
 
@@
 
-- Data: All data used in this manuscript and vignettes have been cited and can be found on our repository pages (https://github.com/SemanticPriming/stimuli-power).
+-   Funding: no funding was provided for this study.
+-   Conflicts of interest: the researchers declare no conflicts of interest.
+-   Ethics approval: not applicable, simulation study.
+-   Consent to participate: not applicable, simulation study.
+-   Consent for publication: all authors approved the manuscript, no participant approval necessary.
+-   Data: All data used in this manuscript and vignettes have been cited and can be found on our repository pages (<https://github.com/SemanticPriming/stimuli-power>).
+-   Materials: No materials were used.
 -   Analysis Code:
@@
 -   Pre-registration: We did not pre-register this study, as it was a simulation study.
-- Materials: No materials were used.
-- Funding: no funding was provided for this study. 
-- Conflicts of interest: the researchers declare no conflicts of interest.
+-   Author contributions are printed at the top of the manuscript automatically in *papaja*.
 

### 2025-08-22 manuscript_draft.tex
update manuscript RMD

@@
 The above simulations assume symmetric population distributions with small, medium, or large heterogeneity or a skewed distribution with the same heterogeneity.
-However, as shown in @pollock2018, rating data that appears normally distributed at the aggregate level may in fact reflect the average of two distinct underlying distributions (i.e., a bimodal distribution).
+However, as shown in Pollock (2018), rating data that appears normally distributed at the aggregate level may in fact reflect the average of two distinct underlying distributions (i.e., a bimodal distribution).
 In such cases, a pilot sample may misleadingly suggest a single ``middle'' value that does not represent either true subgroup.
@@
 
+\bibitem[\citeproctext]{ref-pollock2018}
+Pollock, L.
+(2018).
+Statistical and methodological problems with concreteness and other semantic variables: A list memory experiment case study.
+\emph{Behavior Research Methods}, \emph{50}(3), 1198--1216.
+\url{https://doi.org/10.3758/s13428-017-0938-y}
+
 \bibitem[\citeproctext]{ref-pownall2023}
